### 2016-09-20 document/XSGRB_sample.tex
Pushing text placeholder

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[XSGRB Sample]{The X-shooter GRB sample. .\thanks{Based on observations made with telescopes at the
	European Southern Observatory at La Silla/Paranal, Chile under
	program  ....}}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[]{
,$^{1}$\thanks{E-mail: jselsing@dark-cosmology.dk}
Everyone in some order,$^{1}$ \\
% List of institutions
$^{1}$Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane
Maries Vej 30, 2100 Copenhagen, Denmark.}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2016}

%% Spectral species

\newcommand{\lya}{Ly$\alpha$} 
\newcommand{\lyb}{Ly$\beta$} 
\newcommand{\hb}{H$\beta$} 
\newcommand{\ha}{H$\alpha$} 
\newcommand{\oi}{[\ion{O}{i}]} 
\newcommand{\sii}{[\ion{S}{ii}]} 
\newcommand{\oii}{[\ion{O}{ii}]$\lambda$3727} 
\newcommand{\oiii}{[\ion{O}{iii}]$\lambda$5007}
\newcommand{\nii}{[\ion{N}{ii}]} 
\newcommand{\feii}{\ion{Fe}{ii}} 
\newcommand{\civ}{\ion{C}{iv}} 
\newcommand{\mgii}{\ion{Mg}{ii}} 
\newcommand{\hi}{\ion{H}{i}}

% TO DOS
\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\changed}[1]{\todo{}{green}{#1}}
\newcommand{\jtodo}[1]{\todo{TODO }{red}{#1}}
\newcommand{\qtodo}[1]{\todo{Question}{red}{#1}}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}

\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

\section{Discussion}

\section{Conclusions}

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{XSGRB_sample} % if your bibtex file is called example.bib

% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
%\begin{thebibliography}{99}
%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
%Others S., 2012, Journal of Interesting Stuff, 17, 198
%\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix

\section{Notes on individual bursts}

\subsection{GRB100205A}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2016-09-20 document/XSGRB_sample.tex
Pushing Johans text

@@
-% mnras_template.tex
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
+\documentclass[iop, twocolappendix, numberedappendix, tighten, appendixfloats]{emulateapj}
+\citestyle{apj}
 
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
+\usepackage[stable]{footmisc}
+\usepackage{graphicx}
+\usepackage{txfonts}
+%\usepackage{hyperref}
+\usepackage{natbib}
+\bibpunct[, ]{(}{)}{;}{a}{}{,}
+%\bibliographystyle{./bibtex/apj}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
-
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-\usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
-
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
-\usepackage[T1]{fontenc}
-\usepackage{ae,aecompl}
-
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[XSGRB Sample]{The X-shooter GRB sample. .\thanks{Based on observations made with telescopes at the
-	European Southern Observatory at La Silla/Paranal, Chile under
-	program  ....}}
-
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[]{
-,$^{1}$\thanks{E-mail: jselsing@dark-cosmology.dk}
-Everyone in some order,$^{1}$ \\
-% List of institutions
-$^{1}$Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane
-Maries Vej 30, 2100 Copenhagen, Denmark.}
-
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2016}
-
-%% Spectral species
-
-\newcommand{\lya}{Ly$\alpha$} 
-\newcommand{\lyb}{Ly$\beta$} 
+\newcommand{\farc}{\hbox{$.\!\!^{\prime\prime}$}} 
+\newcommand{\ergA}{$\rm{erg\,cm^{-2}\,s^{-1}\,\AA^{-1}}$} 
+\newcommand{\erg}{$\rm{erg\,cm^{-2}\,s^{-1}}$} 
+\newcommand{\kms}{$\rm{km\,s^{-1}}\,$}
 \newcommand{\hb}{H$\beta$} 
 \newcommand{\ha}{H$\alpha$}
-\newcommand{\oi}{[\ion{O}{i}]} 
-\newcommand{\sii}{[\ion{S}{ii}]} 
-\newcommand{\oii}{[\ion{O}{ii}]$\lambda$3727} 
-\newcommand{\oiii}{[\ion{O}{iii}]$\lambda$5007}
-\newcommand{\nii}{[\ion{N}{ii}]} 
-\newcommand{\feii}{\ion{Fe}{ii}} 
-\newcommand{\civ}{\ion{C}{iv}} 
-\newcommand{\mgii}{\ion{Mg}{ii}} 
-\newcommand{\hi}{\ion{H}{i}}
+\newcommand{\hg}{H$\gamma$} 
+\newcommand{\hi}{\mbox{H\,{\sc i}}} 
+\newcommand{\oi}{[\ion{O}{2}]} 
+\newcommand{\sii}{[\ion{S}{2}]} 
+\newcommand{\oii}{[\ion{O}{2}]} 
+\newcommand{\oiii}{[\ion{O}{3}]}
+\newcommand{\nii}{[\ion{N}{2}]} 
+\newcommand{\hii}{$\rm{H}_2$}
+\newcommand{\griz}{$g' r' i' z'$}
+\newcommand{\JHK}{$JHK_{\rm{s}}$}
+\newcommand{\gK}{$g' r' i' z' JHK_{\rm{s}}$}
+\newcommand{\Msun}{$M_\odot$}
 
-% TO DOS
-\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
-\newcommand{\changed}[1]{\todo{}{green}{#1}}
-\newcommand{\jtodo}[1]{\todo{TODO }{red}{#1}}
-\newcommand{\qtodo}[1]{\todo{Question}{red}{#1}}
+\begin{document}
 
-% Don't change these lines
-\begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
+	\title{\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra}
+	\altaffiltext{$^{\dag}$}{Based on observations collected at the European Southern Observatory, Paranal, 
+		Chile, Program ID: 084.A-0260, 085.A-009, 086.A-0073, 087.A-0055.}
 
-% Abstract of the paper
+	\author{
+		J.~Selsing\altaffilmark{1}, 
+		D.~Malesani\altaffilmark{1}, 
+		P.~Goldoni\altaffilmark{14}, 
+		J.~P.~U. Fynbo\altaffilmark{1}, 
+		A.~de~Ugarte~Postigo\altaffilmark{11}, 
+		J.~Japelj\altaffilmark{20},
+		P.~D'Avanzo,
+		Z.~Cano,
+		S.~Covino\altaffilmark{10}, 
+		V.~D'Elia\altaffilmark{7, 12}, 
+		H.~Flores,
+		O.~E.~Hartoog\altaffilmark{6},
+		J.~Hjorth\altaffilmark{1}, 
+		P.~Jakobsson\altaffilmark{5}, 
+		T.~Kr\"{u}hler\altaffilmark{1}, 
+		A.~Levan,
+		A.~Melandri,
+		S.~Piranomonte \altaffilmark{7},
+		R.~S\'anchez-Ram\'irez\altaffilmark{11},
+		S.~Schulze\altaffilmark{17, 18}, 
+		N.~R.~Tanvir\altaffilmark{19},
+		C.~Th{\"o}ne,
+		S.~D.~Vergani\altaffilmark{7, 8},
+		P.~M.~Vreeswijk\altaffilmark{3}, 
+		D.~J.~Watson\altaffilmark{1},
+		K.~Wiersema\altaffilmark{19},
+		D.~Xu\altaffilmark{1}
+		L.~Christensen\altaffilmark{1},
+		A.~De~Cia\altaffilmark{3}, 
+		L.~Kaper\altaffilmark{6}, 
+		L.~A.~Antonelli,
+		F.~Fiore,
+		A.~Gomboc,
+		P.~Groot,
+		F.~Hammer,
+		C.~Ledoux\altaffilmark{2}, 
+		E.~Maiorano,
+		B.~Milvang-Jensen\altaffilmark{1}, 
+		E.~Palazzi,
+		E.~Pian,
+		J.~Schaye,
+		G.~Tagliaferri\altaffilmark{7},
+		R.~A.~M.~J.~Wijers\altaffilmark{6}
+	}
+
+	\altaffiltext{1}{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark}
+	\altaffiltext{2}{European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile}
+	\altaffiltext{3}{Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel}
+	\altaffiltext{4}{Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany}
+	\altaffiltext{5}{Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland}
+	\altaffiltext{6}{Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands}
+	\altaffiltext{7}{INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy}
+	\altaffiltext{8}{GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France}
+	\altaffiltext{9}{American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA}
+	\altaffiltext{10}{INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy}
+	\altaffiltext{11}{Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain}
+	\altaffiltext{12}{ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy}
+	\altaffiltext{13}{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic}
+	\altaffiltext{14}{APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/
+		IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et L
+		eonie Duquet, 75205 Paris Cedex 13, France}
+	\altaffiltext{15}{Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachs
+		tra\ss e, 85748 Garching, Germany}
+	\altaffiltext{16}{Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3,
+		20126, Milano, Italy}
+	\altaffiltext{17}{Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astro
+		nom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile}
+	\altaffiltext{18}{Millennium Center for Supernova Science}
+	\altaffiltext{19}{Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK}
+	\altaffiltext{20}{University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI}\
+
 	\begin{abstract}
-
+		Later
 	\end{abstract}
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-keyword1 -- keyword2 -- keyword3
-\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+	\keywords{Gamma-ray burst: individual: GRB~120815A --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
 
@@
 
+	\section{Observations}
+
+	\LongTables
+
+	\begin{deluxetable*}{@{\extracolsep{\fill}}lcccccclc@{}}
+		%\tabletypesize{\scriptsize}
+		%\rotate
+		\tablecaption{The full sample of afterglows or hosts observed in the program.
+			We here list the burst names and details of the spectroscopic observations. The
+			exposure times and slit widths are given in the order UVB/VIS/NIR. The column
+			$\Delta t$ shows the time after trigger when the spectroscopic observation was
+			started. Mag$_\mathrm{acq}$ gives the approximate magnitude (typically in the
+			$R$-band) of the afterglow in the acquisition image.
+			\label{sample}}
+		\tablewidth{0pt}
+		\tablehead{
+			\colhead{GRB} &  \colhead{Exptime} & \colhead{Slit width} & \colhead{Airmass} & \colhead{Seeing} & \colhead{$\Delta t$} & \colhead{Mag$_\mathrm{acq}$} & \colhead{Redshift} & \colhead{Ref}\\
+			&  \colhead{(ks)}   & \colhead{(arcsec)} &   & \colhead{(arcsec)} & \colhead{(hr)}       &  & &  \\
+		}
+		\startdata
+		090313$^1$ & 6.9/6.9/6.9     & 1.0/0.9/0.9 & 1.2--1.4  & 1.0  &    45  &  21.6  & 3.3736 & (1) \\
+		090530$^1$ & 4.8/4.8/4.8     & 1.0/1.2/1.2 & 1.6--2.2  & 1.5  &    20  &  22    & 1.266 & (2) \\
+		090809$^1$ & 7.2/7.2/7.2     & 1.0/0.9/0.9 & 1.2--1.1  & 0.9  &  10.2  &  21    & 2.737  & (2,3) \\
+		090926$^1$ & 7.2/7.2/7.2     & 1.0/0.9/0.9 & 1.4--1.5  & 0.9  &    22  &  17.9  & 2.1062 & (4) \\
+		091018     & 2.4/2.4/2.4     & 1.0/0.9/0.9 & 2.1--1.8  & 0.8  &   3.5  &  19.1  & 0.9710 & (5) \\  
+		091127     & 6.0/6.0/6.0     & 1.0/0.9/0.9 & 1.1--1.2  & 1.0  &   101  &  21.2  & 0.490  & (6) \\
+		100205A     & 10.8/10.8/10.8 & 1.0/0.9/0.9 & 1.9--1.8  & 1.0  &    71  &   --   &  --    & (2) \\
+		100219A     &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.3--1.1  & 0.7  &  12.5  &   23   & 4.667  & (7) \\
+		100316B     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 2.0--2.4  & 0.7  &   0.7  &  18.2  & 1.18   & (2) \\
+		100316D-1$^2$ &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.2--1.5  & 1.0  &  12  &   --   & 0.059  & (8) \\
+		100316D-2   &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.2  & 1.0  &    58  &   --   & 0.059  & (8) \\
+		100316D-3   &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.2  & 0.8  &   192  &   --   & 0.059  & (8) \\
+		100418A-1   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.6--1.3  & 0.7  &   8.4  &  18.1  & 0.6235 & (9) \\
+		100418A-2   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.3  & 0.6  &    34  &   --   & 0.6235 & (9) \\
+		100418A-3   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.4  & 0.7  &    58  &   --   & 0.6235 & (9) \\
+		100424A$^3$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.1--1.2  & 0.8  &   --   &   --   & 2.465  & (2) \\
+		100425A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.5--1.3  & 0.7  &   4.0  &  20.6  & 1.755  & (2,3) \\
+		100621A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.3--1.4  & 1.0  &   7.1  &   --   & 0.542  & (2) \\
+		100625A$^3$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.1--1.0  & 0.8  &    13  &   --   & 0.452  & (2) \\
+		100724A$^4$ &  4.2/4.2/4.2   & 1.0/0.9/0.9 & 1.5--2.3  &      &   0.2  &   --   & 1.288  & (2) \\
+		100728B$^5$ &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.5--1.1  & 0.5  &    22  &   23   & 2.106  & (2) \\
+		100814A-1$^4$ & 0.9/0.9/0.9  & 1.0/0.9/0.9 & 1.9--1.7  & 0.5  &   0.8  &   19   & 1.44   & (2) \\
+		100814A-2   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.5--1.2  & 0.6  &   1.4  &   19   & 1.44   & (2) \\
+		100814A-3   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.0  & 0.6  &   99   &   20   & 1.44   & (2) \\
+		100816A$^6$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.8--1.6  & 0.8  &   3.7  &   --   & 0.806  & (2) \\
+		100901A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.5--1.5  & 1.8  &   66   &   --   & 1.408  & (10) \\
+		101219A     &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.1--1.7  & 2.0  &   3.7  &   --   & 0.718  & (2) \\
+		101219B-1   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.6--2.6  & 1.3  &  11.6  &   20   & 0.5519 & (11) \\
+		101219B-2   &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.2--2.0  & 0.8  &   394  &  22.7  & 0.5519 & (11) \\
+		101219B-3   &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.4--2.1  & 0.9  &   886  &   --   & 0.5519 & (11) \\
+		110128A     &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 2.0--1.6  & 0.9  &   5.5  &  22.5  & 2.339  & (2) \\
+		110407A     &  9.6/9.6/9.6   & 1.0/0.9/0.9 & 1.4--1.3  & 2.0  &  12.4  &   23   &  --    & (2) \\
+		110709B$^1,3$ &  7.2/7.2/7.2 & 1.0/0.9/0.9 & 1.6--1.1  & 1.0  &   --   &   --   &  --    & (2) \\
+		110715A     &  0.6/0.6/0.6   & 1.0/0.9/0.9 & 1.1--1.1  & 1.7  &  12.3  &  18.5  & 0.82  & (2) \\
+		%110721A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.4  & 2.4  &        &        & 0.382  & (2) \\
+		110808A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.2--1.1  & 1.1  &   3.0  &  21.2  & 1.3488 & (2) \\
+		110818A     & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.3  & 1.0  &   6.2  &  22.3  & 3.36   & (2) \\
+		111005A$^3$ & 1.2/1.2/1.2    & 1.0/0.9/0.9 & 1.3--1.3  & 0.7  &   --   &  --    & 0.013? & (2) \\
+		111008A-1   & 8.8/8.8/8.4    & 1.0/0.9/0.9 & 1.1--1.0  & 1.2  &   8.5  &  21?   & 4.9898 & (12) \\
+		111008A-2   & 8.0/8.0/7.2    & 1.0/0.9/0.9 & 1.3--1.0  & 1.0  &  20.1  &  22?   & 4.9898 & (12) \\
+		111107A     & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.8--1.5  & 0.7  &   5.3  &  21.5  & 2.893  & (2) \\
+		111117A$^6$ & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.5--1.4  & 0.6  &    38  &  --    & 1.3?   & (2) \\
+		111123A-1   & 6.2/6.6/6.6    & 1.0/0.9/0.9 & 1.6--1.1  & 1.0  &  12.2  &  $>$24 & 3.1516 & (2) \\
+		111123A-2$^3$ & 2.4/2.4/2.4  & 1.0/0.9/0.9 & 1.0--1.0  & 0.5  &   --   &  --    & 3.1516 & (2) \\
+		111129A     & 3.6/3.6/3.6    & 1.0/0.9/0.9 & 1.6--2.1  & 1.7  &        &        &  --    & (2) \\
+		111209A-1   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.1--1.2  & 0.8  &  17.7  &  20.1  & 0.677  & (13) \\
+		111209A-2   & 9.6/9.6/9.6    & 1.0/0.9/0.9 & 1.2--2.0  & 0.8  &  497   &  23    & 0.677  & (13) \\
+		111211A$^1$ & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.4--1.6  & 0.6  &   31   &  19.5  & 0.478  & (2) \\
+		111228A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.4--1.4  & 0.9  &  15.9  &  20.1  & 0.716  & (2) \\
+		120118B$^3$ & 3.6/3.6/3.6    & 1.0/0.9/0.9 & 1.1--1.0  & 1.0  &   --   &  --    & 2.943  & (2) \\
+		120119A-1   & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.1--1.1  & 0.6  &   1.4  &   17   & 1.728  & (2) \\
+		120119A-2   & 1.2/1.2/1.2    & 1.0/0.9/0.9 & 1.8--1.9  & 0.6  &   4.5  &   20   & 1.728  & (2) \\
+		120119A-3$^3$ & 4.8/4.8/4.8  & 1.0/0.9/0.6JH & 1.0--1.1  & 1.1 &   --   &   --   & 1.728  & (2) \\
+		120211      &                &             &           &      &        &         & 2.346 & (2) \\
+		120224A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.7--2.1  & 1.4  &  19.8  &   22.3 &  --    & (2) \\
+		120311A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.6--1.4  & 0.6  &   3.7  &   21.6 &  --    & (2) \\
+		120327A-1   & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.6--1.4  & 0.5  &   2.1  &   18.8 & 2.815  & (14) \\
+		120327A-2   & 4.2/4.2/4.2    & 1.0/0.9/0.9 & 1.0--1.1  & 1.0  &    29  &   22.5 & 2.815  & (14) \\
+		120404A     & 9.6/9.6/9.6    & 1.0/0.9/0.9JH & 1.7--1.3 & 1.3 &  15.7  &   21.3 & 2.876  & (2) \\
+		120422A-1   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.3  & 0.6  &  17.2  &   22.0 & 0.283  & (15) \\
+		120422A-2   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.4  & 0.9  &  113   &   --   & 0.283  & (15) \\
+		120422A-3   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.4--1.7  & 1.0  &  210   &   --   & 0.283  & (15) \\
+		120422A-4   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.3--1.4 & 0.6  & 449   &   --   & 0.283  & (15) \\
+		120422A-5   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.3--1.6 & 0.8  & 593   &   --   & 0.283  & (15) \\
+		120422A-6   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.7--2.4 & 2.5  & 882   &   --   & 0.283  & (15) \\
+		120422A-7   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.5--1.9 & 1.3  & 906   &   --   & 0.283  & (15) \\
+		120712A     & 4.8/4.8/4.8    & 1.0/0.9/0.9   & 1.5--2.5 & 1.3  & 10.4  &   21.5 & 4.175  & (2) \\
+		120714B     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.5--1.2 & 1.2  &  7.8  &   22.1 & 0.398  & (2)\\
+		120716A$^1$ & 3.6/3.6/3.6    & 1.0/0.9/0.9JH & 1.8--2.6 & 1.0  &  62   &   20.9 & 2.486  & (2) \\
+		120722A$^2$ & 4.8/4.8/4.8    & 1.0/0.9/0.9   & 1.3--1.3 & 1.1  & 10.3  &   23.6 & 0.959  & (2) \\
+		120805A$^2$ & 3.6/3.6/3.6    & 1.0/0.9/0.9JH & 1.3--1.7 & 0.9  & 218   &   --   & 2.8?   & (2) \\
+		120815A     & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.3--1.4 & 0.6  &  1.7  &   20   & 2.358  & (16) \\
+		120909A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.6--1.6 & 1.4  &  1.7  &   21   & 3.929  & (2) \\
+		120923A     & 9.6/9.6/9.6    & 1.0/0.9/0.9JH & 1.2--1.4 & 1.0  & 18.5  &   --   & $\gtrsim8$ & (2) \\
+		121024A     & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.2--1.1 & 0.6  &  1.8  &   20   & 2.300  & (17) \\
+		121027A     & 8.4/8.4/8.4    & 1.0/0.9/0.9   & 1.3--1.3 & 0.9  &  69   &  21.1  & 1.773  & (2) \\
+		121201A     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.1--1.1 & 0.9  & 12.9  &   23   & 3.385  & (2) \\
+		121229A     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.4--1.2 & 1.4  &  2.0  &  21.5  & 2.707  & (2) \\
+		130131B$^3$ & 7.2/7.2/7.2    & 1.0/0.9/0.9JH & 1.3--1.6 & 0.8  &  --   &   --   & 2.539  & (2) \\
+		130408A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.0--1.0 & 1.0  &  1.9  &   20   & 3.758  & (2) \\
+		130418A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.4--1.3 & 1.3  &  4.6  &   18.5 & 1.218  & (2) \\
+		130427A     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 1.8--1.8 & 0.8  & 16.5  &   19   & 0.340  & (18) \\
+		130427B     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 1.2--1.0 & 0.8  & 20.3  &   22.7 & 2.78   &  (2) \\
+		130603B$^6$ & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.4--1.4 & 1.1  &  8.2  &   21.5 & 0.356  & (19) \\
+		130606A     & 4.2/4.2/4.2    & 1.0/0.9/0.9JH & 1.7--1.9 & 1.1  &  7.1  &   19   & 5.91   & (20) \\
+		130612A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.3--1.3 & 1.4  &  1.1  &   21.5 & 2.006  & (2) \\
+		130615A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 2.1--2.2 & 1.0  &  0.8  &   21   & $~3$?  & (2) \\
+		130701A     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 2.0--2.0 & 1.6  &  5.5  &   19.9 & 1.155  & (2) \\
+		\enddata
+		\tablenotetext{1}{Not part of the statistical sample}
+		\tablenotetext{2}{Spectrum dominated by light from the host galaxy}
+		\tablenotetext{3}{Spectrum of the host galaxy taken long after the burst}
+		\tablenotetext{4}{RRM observation}
+		\tablenotetext{5}{ADC malfunction during observation}
+		\tablenotetext{6}{Short burst}
+		\tablerefs{
+			(1) \citet{Antonio090313}; (2) This work ; (3) Skuladottir (2010);
+			(4) \citet{Valerio090926A}; (5) \citet{Klaas091018}; (6) \citet{Vergani091127,Cobb091127}; 
+			(7) \citet{christina100219A}; (8) \citet{Buffano} ; (9) de Ugarte Postigo et al. ;
+			(10) \citet{Haartog100901A}; (11) \citet{Sparre101219B}; (12) \citet{Sparre111008A};
+			(13) \citet{Levanultralong}; (14) D'Elia et al., submitted; (15) \citet{SchulzeSNGRB};
+			(16) \citet{Kruehler120815A}; (17} \citet{friis121024A}
+	\end{deluxetable*}
+	%-------------------------END TABLE---------------------------------------
+
+	\section{Results}
+
+	\subsection{Spectral resolution}
+	The afterglow spectra described in this paper are obtained in Target-of-Opportunity (override) mode.
+	In most cases there is therefore little possibility to tweak slit widths to the seeing at the time of observations
+	(i.e. to optimise spectral resolution and signal to noise), and almost all our data is therefore taken with a fixed
+	set of slit widths and binning, described above. In a fair number of cases, the seeing full width at half maximum (FWHM) 
+	is considerably smaller than the silt width, and the delivered spectral resolution will then be determined by the seeing rather 
+	than slit width, as afterglows are point sources (this is evidently not the case for extended sources, e.g for host galaxies).
+	The delivered resolution for slit width dominated spectra post-reduction and extraction can easily be determined 
+	from the bright sky emission lines. For afterglow spectra with very high signal to noise, the delivered spectral resolution 
+	can at times be determined from the science data themselves. However, in the presence of multiple velocity components in 
+	absorption, other forms of line broadening, and a lack of lines at some redshifts, this is difficult to do at poorer signal to noise ratios
+	(the majority of spectra in our sample). A broad starting value for the expect resolution will help fitting of these spectra, and can be important
+	in upper limit determination, and for this reason we construct a aim to construct a crude relation between the seeing and the delivered
+	resolution at our slit width, binning, and reduction pipeline settings.  
+	To this end we use observations of telluric standard stars that are taken with identical instrument settings as our afterglow spectra, usually just after 
+	the science data, as part of the ESO X-shooter calibration plan. These spectra have been reduced together with the afterglow spectra, using identical
+	pipeline settings with the same version of the pipeline.
+	First we fit a Gaussian function in the spatial direction of the trace of the standard star at 792 nm (i.e. in the VIS arm). 
+	After this, we fit a series of  20 telluric absorption lines in the telluric standard star spectra with Gaussians, taking care to select transitions 
+	that are not almost-resolved multiples, should be intrinsically unresolved, and are in areas with well defined continuum flux.  We pick 34 
+	telluric standard stars spanning a range of DIMM seeing values, with the majority between $0.5-1.5 arcsec$. The resulting distribution of 
+	spectral FWHM (km/s) as a function of spatial FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with
+	0.15 arc sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$. We use this linear relation as a 
+	way to estimate the spectral resolution for medium to poor signal to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR arm, we measured
+	a series of lines in NIR arm spectra of a subset of 19 sources used for the VIS arm above, and find that the resulting distribution is consistent with 
+	a simple scaling of the VIS arm relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit filling, sources as given on the ESO instrument website.
+	The UVB arm contains no suitable absorption lines to use, and we therefore use a scaled value as in the NIR arm.
+	While this simple method is not terribly accurate (for one, the spatial profile of the trace is not a perfect Gaussian), but it gives a sufficiently accurate estimate for the analysis of these poor signal to noise science spectra.
+
+%	\begin{figure}
+%		%\vspace{11pc}   
+%		%\includegraphics[width=10cm]{060218/6273fig1.ps}
+%		\centerline{\includegraphics[width=8cm]{Resolution/resolution_paper.eps}}
+%		\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
+The lower horizontal axis is in units pixels, the top axis in arc seconds.
+The red datapoints show a subsample of NIR spectra.
+%			The grey line shows a linear fit to the VIS datapoints. }
+%		\label{fig:res}
+%	\end{figure}
+
+	\subsection{Correction for offsets in the wavelength calibration}
+
+	X-shooter, being installed at the VLT Cassegrain focusi is  prone to
+	flexures during operations. The flexures modify the projection of the slit
+	on the detector with respect to the one obtained in daytime calibration. 
+	This require a modification of the wavelength solution in order to
+	process correctly the night-time data. Part of this correction
+	is performed by the pipeline using the frames taken
+	during X-shooter Active Flexure Compensation procedure
+	\footnote{X-shooter User Manual available at https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}
+	We corrected the remaining part using as a reference the sky
+	emission lines present in the observed data.
+
+	For every afterglow observation we reduced one frame individually in STARE mode without
+	sky subtraction obtaining $\sim$ 100 sky spectra. The sky line list compiled at ESO for
+	the E-ELT
+	study\footnote{http://www.eso.org/sci/facilities/eelt/science/drm/tech\_data/data/optical\_ir\_sky\_lines.dat}
+	from the work of (Hanuschik, 2003, A\&A, 407, 1157) and (Rousselot et al. (2000, A\&A, 354, 1134), was used as a reference.
+	From this list, we selected a subset of bright and isolated lines.
+In the case of the OH doublets, unresolved at X-shooter resolution, we took as line position the average between the blue and red components.
+To find the offsets of the spectra, we fitted gaussians near the
+	expected positions under IDL using the MPFIT software (Markwardt, 2009, Astronomical Society of the Pacific Conference Series, Vol.
+411, ADASS XVII, ed.
+D.A.
+Bohlender, D.
+Durand, \& P.
+Dowler, 251) and we compared the result to the tabulated values.
+	The resulting offsets, which were smaller than 0.1 \AA~in the UVB and VIS data
+	and smaller than 0.5 \AA~in the NIR spectra, were applied to the corresponding spectra.
+
+	\subsection{Redshifts}
+%	\subsection{Notes on Individual objects}
+
 	\section{Discussion}
 
-\section{Conclusions}
+	\begin{acknowledgements}
+		JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
+		Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
+		acknowledges support by the European Commission under the Marie Curie
+		Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
+		European Commission under the Marie Curie Career Integration Grant programme
+		(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
+		{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
+		acknowledge expert support from the ESO staff at the Paranal and La Silla
+		observatories in obtaining these target of opportunity data.
 
-\section*{Acknowledgements}
+	\end{acknowledgements}
 
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
+	\def\aj{AJ}
+	\def\araa{ARA\&A}
+	\def\apj{ApJ}
+	\def\apjl{ApJL}
+	\def\apjs{ApJS}
+	\def\apss{Ap\&SS}
+	\def\aap{A\&A}
+	\def\aapr{A\&A~Rev.}
+	\def\aaps{A\&AS}
+	\def\mnras{MNRAS}
+	\def\nat{Nature}
+	\def\pasp{PASP}
+	\def\aplett{Astrophys.~Lett.}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\bibliographystyle{apj}
+	\bibliography{XSGRB_sample}
+	%\bibliography{thebib}
 
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-% The best way to enter references is to use BibTeX:
-
-\bibliographystyle{mnras}
-\bibliography{XSGRB_sample} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
 	\newpage
 	\appendix
+	\section{Notes on Individual objects}
 
-\section{Notes on individual bursts}
+	\subsection{GRB090313}
 
-\subsection{GRB100205A}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
 \end{document}
-
-% End of mnras_template.tex

### 2016-09-21 document/XSGRB_sample.tex

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \usepackage{txfonts}
@@
 \newcommand{\kms}{$\rm{km\,s^{-1}}\,$}
-\newcommand{\hb}{H$\beta$} 
-\newcommand{\ha}{H$\alpha$}
-\newcommand{\hg}{H$\gamma$} 
-\newcommand{\hi}{\mbox{H\,{\sc i}}} 
-\newcommand{\oi}{[\ion{O}{2}]} 
-\newcommand{\sii}{[\ion{S}{2}]} 
-\newcommand{\oii}{[\ion{O}{2}]} 
-\newcommand{\oiii}{[\ion{O}{3}]}
-\newcommand{\nii}{[\ion{N}{2}]} 
-\newcommand{\hii}{$\rm{H}_2$}
+
 \newcommand{\griz}{$g' r' i' z'$}
@@
 \newcommand{\Msun}{$M_\odot$}
-
+\newcommand{\lya}{Ly$\alpha$} 
+\newcommand{\lyb}{Ly$\beta$} 
+\newcommand{\hb}{H$\beta$} 
+\newcommand{\ha}{H$\alpha$} 
+\newcommand{\oi}{[\ion{O}{1}]} 
+\newcommand{\sii}{[\ion{S}{2}]} 
+\newcommand{\oii}{[\ion{O}{2}]$\lambda$3727} 
+\newcommand{\oiii}{[\ion{O}{3}]$\lambda$5007}
+\newcommand{\nii}{[\ion{N}{2}]} 
+\newcommand{\feii}{\ion{Fe}{2}} 
+\newcommand{\civ}{\ion{C}{4}} 
+\newcommand{\cii}{\ion{C}{2}}
+\newcommand{\mgii}{\ion{Mg}{2}} 
+\newcommand{\hi}{\ion{H}{1}} 
+\newcommand{\ali}{\ion{Al}{1}} 
+\newcommand{\SIii}{\ion{Si}{2}} 
 \begin{document}
@@
 		P.~Goldoni\altaffilmark{14}, 
+		T.~Kr\"{u}hler\altaffilmark{1}, 
 		J.~P.~U. Fynbo\altaffilmark{1}, 
@@
 		P.~Jakobsson\altaffilmark{5}, 
-		T.~Kr\"{u}hler\altaffilmark{1}, 
 		A.~Levan,
@@
 		\tablerefs{
-			(1) \citet{Antonio090313}; (2) This work ; (3) Skuladottir (2010);
-			(4) \citet{Valerio090926A}; (5) \citet{Klaas091018}; (6) \citet{Vergani091127,Cobb091127}; 
-			(7) \citet{christina100219A}; (8) \citet{Buffano} ; (9) de Ugarte Postigo et al. ;
-			(10) \citet{Haartog100901A}; (11) \citet{Sparre101219B}; (12) \citet{Sparre111008A};
-			(13) \citet{Levanultralong}; (14) D'Elia et al., submitted; (15) \citet{SchulzeSNGRB};
-			(16) \citet{Kruehler120815A}; (17} \citet{friis121024A}
+			(1) \citet{DeUgartePostigo2010}; (2) This work ; (3) Skuladottir (2010);
+			(4) \citet{DElia2010}; (5) \citet{Wiersema2012}; (6) \citet{Vergani2011, Cobb2010}; 
+			(7) \citet{Thone2013}; (8) \citet{Bufano2011} ; (9) \citet{DeUgartePostigo2011} ;
+			(10) \citet{Hartoog2013}; (11) \citet{Sparre2011}; (12) \citet{Sparre2014};
+			(13) \citet{Levan2014}; (14) \citet{DElia2014}; (15) \citet{Schulze2014};
+			(16) \citet{Kruhler2013}; (17} \citet{Friis2015}
 	\end{deluxetable*}
@@
 	\subsection{Spectral resolution}
-	The afterglow spectra described in this paper are obtained in Target-of-Opportunity (override) mode.
-	In most cases there is therefore little possibility to tweak slit widths to the seeing at the time of observations
-	(i.e. to optimise spectral resolution and signal to noise), and almost all our data is therefore taken with a fixed
-	set of slit widths and binning, described above. In a fair number of cases, the seeing full width at half maximum (FWHM) 
-	is considerably smaller than the silt width, and the delivered spectral resolution will then be determined by the seeing rather 
-	than slit width, as afterglows are point sources (this is evidently not the case for extended sources, e.g for host galaxies).
-	The delivered resolution for slit width dominated spectra post-reduction and extraction can easily be determined 
-	from the bright sky emission lines. For afterglow spectra with very high signal to noise, the delivered spectral resolution 
-	can at times be determined from the science data themselves. However, in the presence of multiple velocity components in 
-	absorption, other forms of line broadening, and a lack of lines at some redshifts, this is difficult to do at poorer signal to noise ratios
-	(the majority of spectra in our sample). A broad starting value for the expect resolution will help fitting of these spectra, and can be important
-	in upper limit determination, and for this reason we construct a aim to construct a crude relation between the seeing and the delivered
-	resolution at our slit width, binning, and reduction pipeline settings.  
-	To this end we use observations of telluric standard stars that are taken with identical instrument settings as our afterglow spectra, usually just after 
-	the science data, as part of the ESO X-shooter calibration plan. These spectra have been reduced together with the afterglow spectra, using identical
-	pipeline settings with the same version of the pipeline.
-	First we fit a Gaussian function in the spatial direction of the trace of the standard star at 792 nm (i.e. in the VIS arm). 
-	After this, we fit a series of  20 telluric absorption lines in the telluric standard star spectra with Gaussians, taking care to select transitions 
-	that are not almost-resolved multiples, should be intrinsically unresolved, and are in areas with well defined continuum flux.  We pick 34 
-	telluric standard stars spanning a range of DIMM seeing values, with the majority between $0.5-1.5 arcsec$. The resulting distribution of 
-	spectral FWHM (km/s) as a function of spatial FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with
-	0.15 arc sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$. We use this linear relation as a 
-	way to estimate the spectral resolution for medium to poor signal to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR arm, we measured
-	a series of lines in NIR arm spectra of a subset of 19 sources used for the VIS arm above, and find that the resulting distribution is consistent with 
-	a simple scaling of the VIS arm relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit filling, sources as given on the ESO instrument website.
-	The UVB arm contains no suitable absorption lines to use, and we therefore use a scaled value as in the NIR arm.
-	While this simple method is not terribly accurate (for one, the spatial profile of the trace is not a perfect Gaussian), but it gives a sufficiently accurate estimate for the analysis of these poor signal to noise science spectra.
+	The afterglow spectra described in this paper are obtained in
+	Target-of-Opportunity (override) mode.
+	In most cases there is therefore little possibility to tweak slit widths to the
+	seeing at the time of observations (i.e. to optimise spectral resolution and
+	signal to noise), and almost all our data is therefore taken with a fixed set
+	of slit widths and binning, described above.
+	In a fair number of cases, the seeing full width at half maximum (FWHM) is
+	considerably smaller than the silt width, and the delivered spectral resolution
+	will then be determined by the seeing rather than slit width, as afterglows are
+	point sources (this is evidently not the case for extended sources, e.g for
+	host galaxies).
+	The delivered resolution for slit width dominated spectra post-reduction and
+	extraction can easily be determined from the bright sky emission lines.
+	For afterglow spectra with very high signal to noise, the delivered spectral
+	resolution can at times be determined from the science data themselves.
+	However, in the presence of multiple velocity components in absorption, other
+	forms of line broadening, and a lack of lines at some redshifts, this is
+	difficult to do at poorer signal to noise ratios (the majority of spectra in
+	our sample).
+	A broad starting value for the expect resolution will help fitting of these
+	spectra, and can be important in upper limit determination, and for this reason
+	we construct a aim to construct a crude relation between the seeing and the
+	delivered resolution at our slit width, binning, and reduction pipeline
+	settings.
+	To this end we use observations of telluric standard stars that are taken with
+	identical instrument settings as our afterglow spectra, usually just after the
+	science data, as part of the ESO X-shooter calibration plan.
+	These spectra have been reduced together with the afterglow spectra, using
+	identical pipeline settings with the same version of the pipeline.
+	First we fit a Gaussian function in the spatial direction of the trace of the
+	standard star at 792 nm (i.e. in the VIS arm).
+	After this, we fit a series of  20 telluric absorption lines in the telluric
+	standard star spectra with Gaussians, taking care to select transitions that
+	are not almost-resolved multiples, should be intrinsically unresolved, and are
+	in areas with well defined continuum flux.
+	We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
+	the majority between $0.5-1.5 arcsec$.
+	The resulting distribution of spectral FWHM (km/s) as a function of spatial
+	FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
+	$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
+	1.3$ km/s, $b=1.4 \pm 0.2$.
+	We use this linear relation as a way to estimate the spectral resolution for
+	medium to poor signal to noise afterglow spectra in the VIS arm.
+	To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
+	spectra of a subset of 19 sources used for the VIS arm above, and find that the
+	resulting distribution is consistent with a simple scaling of the VIS arm
+	relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
+	slit filling, sources as given on the ESO instrument website.
+	The UVB arm contains no suitable absorption lines to use, and we therefore use
+	a scaled value as in the NIR arm.
+	While this simple method is not terribly accurate (for one, the spatial profile
+	of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
+	estimate for the analysis of these poor signal to noise science spectra.
 
@@
 
+	\subsection{GRB120327A (z = 2.813)}
+	The data presented here also formed the basis of GCN \#
+	13134\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13134.gcn3}} and is published
+	in \citet{DElia2014}.
+	Consisting of two observations, 2.13 hrs and 29.98 hrs after the burst,
+	continuum is visible in all arms for both.
+	Absorption features from Ly-limit, \lya, \cii/\cii$^*$, \SIii/\SIii$^*$, \ali,
+	\feii ~and \mgii ~are detected at a consistent redshift, $z = 2.813$.
+
 \end{document}

### 2016-09-21 document/XSGRB_sample.tex

@@
 	in \citet{DElia2014}.
-	Consisting of two observations, 2.13 hrs and 29.98 hrs after the burst,
-	continuum is visible in all arms for both.
-	Absorption features from Ly-limit, \lya, \cii/\cii$^*$, \SIii/\SIii$^*$, \ali,
+	The observation consists of two visits, 2.13 hrs and 29.98 hrs after the burst,
+	with an afteglow continuum visible in all arms for both visits.
+	We detect absorption features from Ly-limit, \lya, \cii/\cii$^*$, \SIii/\SIii$^*$, \ali,
 	\feii ~and \mgii ~are detected at a consistent redshift, $z = 2.813$.
 
+        \subsection{GRB130408A (z = 3.758)}
+	The data presented here also formed the basis of GCN \#
+	14365\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14365.gcn3}}. The spectrum 
+	has not otherwise been published previously.
+	The observations consists of two 600sec spectra taken 1.9hr after the burst.
+	We detect absorption features from a wide range of ions. We also detect intervening absorption at $z=1.255$
+	and $z=3.248$.
+
 \end{document}

### 2016-09-21 document/XSGRB_sample.tex
Update XSGRB_sample.tex

@@
 
+	\subsection{GRB130606A (z = 5.913)}
+	The data presented here also formed the basis of GCN \#
+	14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is published
+	in \citet{Hartoog2015}.
+	The observations consists of three 2x600sec visits starting 7.1 hr after the burst at fairly
+	high airmass.
+	We detect absorption features from a wide range of ions at z=5.913 
+	as well as intervening absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
+
 \end{document}

### 2016-09-23 document/XSGRB_sample.tex
Update XSGRB_sample.tex

@@
 
+	\subsection{GRB160203A (z = 3.517)}
+	The data presented here also formed the basis of GCN \#
+	18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
+	elsewhere.
+	The observation was carried out in RRM starting 18 min af the GRB trigger.
+	We detect absorption features from a wide range of ions at z=3.517 
+	as well as intervening absorption at $z=2.203$.
+
 \end{document}

### 2016-09-24 document/XSGRB_sample.tex
Update XSGRB_sample.tex

@@
 
+	\subsection{GRB151021A (z = 2.330)}
+	The data presented here also formed the basis of GCN \#
+	18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
+	elsewhere.
+	The observation was carried out in RRM starting 44 min af the GRB trigger.
+	We detect absorption features from a wide range of ions at z=2.330 
+	as well as intervening absorption at $z=1.49$.
+
 	\subsection{GRB160203A (z = 3.517)}

### 2016-09-24 document/XSGRB_sample.tex
Moving table to separate file. Update for GRB100724A

--- 2016-09-21 document/XSGRB_sample.tex
@@
 
-	\LongTables
-
-	\begin{deluxetable*}{@{\extracolsep{\fill}}lcccccclc@{}}
-		%\tabletypesize{\scriptsize}
-		%\rotate
-		\tablecaption{The full sample of afterglows or hosts observed in the program.
-			We here list the burst names and details of the spectroscopic observations. The
-			exposure times and slit widths are given in the order UVB/VIS/NIR. The column
-			$\Delta t$ shows the time after trigger when the spectroscopic observation was
-			started. Mag$_\mathrm{acq}$ gives the approximate magnitude (typically in the
-			$R$-band) of the afterglow in the acquisition image.
-			\label{sample}}
-		\tablewidth{0pt}
-		\tablehead{
-			\colhead{GRB} &  \colhead{Exptime} & \colhead{Slit width} & \colhead{Airmass} & \colhead{Seeing} & \colhead{$\Delta t$} & \colhead{Mag$_\mathrm{acq}$} & \colhead{Redshift} & \colhead{Ref}\\
-			&  \colhead{(ks)}   & \colhead{(arcsec)} &   & \colhead{(arcsec)} & \colhead{(hr)}       &  & &  \\
-		}
-		\startdata
-		090313$^1$ & 6.9/6.9/6.9     & 1.0/0.9/0.9 & 1.2--1.4  & 1.0  &    45  &  21.6  & 3.3736 & (1) \\
-		090530$^1$ & 4.8/4.8/4.8     & 1.0/1.2/1.2 & 1.6--2.2  & 1.5  &    20  &  22    & 1.266 & (2) \\
-		090809$^1$ & 7.2/7.2/7.2     & 1.0/0.9/0.9 & 1.2--1.1  & 0.9  &  10.2  &  21    & 2.737  & (2,3) \\
-		090926$^1$ & 7.2/7.2/7.2     & 1.0/0.9/0.9 & 1.4--1.5  & 0.9  &    22  &  17.9  & 2.1062 & (4) \\
-		091018     & 2.4/2.4/2.4     & 1.0/0.9/0.9 & 2.1--1.8  & 0.8  &   3.5  &  19.1  & 0.9710 & (5) \\  
-		091127     & 6.0/6.0/6.0     & 1.0/0.9/0.9 & 1.1--1.2  & 1.0  &   101  &  21.2  & 0.490  & (6) \\
-		100205A     & 10.8/10.8/10.8 & 1.0/0.9/0.9 & 1.9--1.8  & 1.0  &    71  &   --   &  --    & (2) \\
-		100219A     &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.3--1.1  & 0.7  &  12.5  &   23   & 4.667  & (7) \\
-		100316B     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 2.0--2.4  & 0.7  &   0.7  &  18.2  & 1.18   & (2) \\
-		100316D-1$^2$ &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.2--1.5  & 1.0  &  12  &   --   & 0.059  & (8) \\
-		100316D-2   &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.2  & 1.0  &    58  &   --   & 0.059  & (8) \\
-		100316D-3   &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.2  & 0.8  &   192  &   --   & 0.059  & (8) \\
-		100418A-1   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.6--1.3  & 0.7  &   8.4  &  18.1  & 0.6235 & (9) \\
-		100418A-2   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.3  & 0.6  &    34  &   --   & 0.6235 & (9) \\
-		100418A-3   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.4  & 0.7  &    58  &   --   & 0.6235 & (9) \\
-		100424A$^3$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.1--1.2  & 0.8  &   --   &   --   & 2.465  & (2) \\
-		100425A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.5--1.3  & 0.7  &   4.0  &  20.6  & 1.755  & (2,3) \\
-		100621A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.3--1.4  & 1.0  &   7.1  &   --   & 0.542  & (2) \\
-		100625A$^3$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.1--1.0  & 0.8  &    13  &   --   & 0.452  & (2) \\
-		100724A$^4$ &  4.2/4.2/4.2   & 1.0/0.9/0.9 & 1.5--2.3  &      &   0.2  &   --   & 1.288  & (2) \\
-		100728B$^5$ &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.5--1.1  & 0.5  &    22  &   23   & 2.106  & (2) \\
-		100814A-1$^4$ & 0.9/0.9/0.9  & 1.0/0.9/0.9 & 1.9--1.7  & 0.5  &   0.8  &   19   & 1.44   & (2) \\
-		100814A-2   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.5--1.2  & 0.6  &   1.4  &   19   & 1.44   & (2) \\
-		100814A-3   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.2--1.0  & 0.6  &   99   &   20   & 1.44   & (2) \\
-		100816A$^6$ &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.8--1.6  & 0.8  &   3.7  &   --   & 0.806  & (2) \\
-		100901A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.5--1.5  & 1.8  &   66   &   --   & 1.408  & (10) \\
-		101219A     &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.1--1.7  & 2.0  &   3.7  &   --   & 0.718  & (2) \\
-		101219B-1   &  4.8/4.8/4.8   & 1.0/0.9/0.9 & 1.6--2.6  & 1.3  &  11.6  &   20   & 0.5519 & (11) \\
-		101219B-2   &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.2--2.0  & 0.8  &   394  &  22.7  & 0.5519 & (11) \\
-		101219B-3   &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 1.4--2.1  & 0.9  &   886  &   --   & 0.5519 & (11) \\
-		110128A     &  7.2/7.2/7.2   & 1.0/0.9/0.9 & 2.0--1.6  & 0.9  &   5.5  &  22.5  & 2.339  & (2) \\
-		110407A     &  9.6/9.6/9.6   & 1.0/0.9/0.9 & 1.4--1.3  & 2.0  &  12.4  &   23   &  --    & (2) \\
-		110709B$^1,3$ &  7.2/7.2/7.2 & 1.0/0.9/0.9 & 1.6--1.1  & 1.0  &   --   &   --   &  --    & (2) \\
-		110715A     &  0.6/0.6/0.6   & 1.0/0.9/0.9 & 1.1--1.1  & 1.7  &  12.3  &  18.5  & 0.82  & (2) \\
-		%110721A     &  2.4/2.4/2.4   & 1.0/0.9/0.9 & 1.2--1.4  & 2.4  &        &        & 0.382  & (2) \\
-		110808A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.2--1.1  & 1.1  &   3.0  &  21.2  & 1.3488 & (2) \\
-		110818A     & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.3  & 1.0  &   6.2  &  22.3  & 3.36   & (2) \\
-		111005A$^3$ & 1.2/1.2/1.2    & 1.0/0.9/0.9 & 1.3--1.3  & 0.7  &   --   &  --    & 0.013? & (2) \\
-		111008A-1   & 8.8/8.8/8.4    & 1.0/0.9/0.9 & 1.1--1.0  & 1.2  &   8.5  &  21?   & 4.9898 & (12) \\
-		111008A-2   & 8.0/8.0/7.2    & 1.0/0.9/0.9 & 1.3--1.0  & 1.0  &  20.1  &  22?   & 4.9898 & (12) \\
-		111107A     & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.8--1.5  & 0.7  &   5.3  &  21.5  & 2.893  & (2) \\
-		111117A$^6$ & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.5--1.4  & 0.6  &    38  &  --    & 1.3?   & (2) \\
-		111123A-1   & 6.2/6.6/6.6    & 1.0/0.9/0.9 & 1.6--1.1  & 1.0  &  12.2  &  $>$24 & 3.1516 & (2) \\
-		111123A-2$^3$ & 2.4/2.4/2.4  & 1.0/0.9/0.9 & 1.0--1.0  & 0.5  &   --   &  --    & 3.1516 & (2) \\
-		111129A     & 3.6/3.6/3.6    & 1.0/0.9/0.9 & 1.6--2.1  & 1.7  &        &        &  --    & (2) \\
-		111209A-1   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.1--1.2  & 0.8  &  17.7  &  20.1  & 0.677  & (13) \\
-		111209A-2   & 9.6/9.6/9.6    & 1.0/0.9/0.9 & 1.2--2.0  & 0.8  &  497   &  23    & 0.677  & (13) \\
-		111211A$^1$ & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.4--1.6  & 0.6  &   31   &  19.5  & 0.478  & (2) \\
-		111228A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.4--1.4  & 0.9  &  15.9  &  20.1  & 0.716  & (2) \\
-		120118B$^3$ & 3.6/3.6/3.6    & 1.0/0.9/0.9 & 1.1--1.0  & 1.0  &   --   &  --    & 2.943  & (2) \\
-		120119A-1   & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.1--1.1  & 0.6  &   1.4  &   17   & 1.728  & (2) \\
-		120119A-2   & 1.2/1.2/1.2    & 1.0/0.9/0.9 & 1.8--1.9  & 0.6  &   4.5  &   20   & 1.728  & (2) \\
-		120119A-3$^3$ & 4.8/4.8/4.8  & 1.0/0.9/0.6JH & 1.0--1.1  & 1.1 &   --   &   --   & 1.728  & (2) \\
-		120211      &                &             &           &      &        &         & 2.346 & (2) \\
-		120224A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.7--2.1  & 1.4  &  19.8  &   22.3 &  --    & (2) \\
-		120311A     & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.6--1.4  & 0.6  &   3.7  &   21.6 &  --    & (2) \\
-		120327A-1   & 2.4/2.4/2.4    & 1.0/0.9/0.9 & 1.6--1.4  & 0.5  &   2.1  &   18.8 & 2.815  & (14) \\
-		120327A-2   & 4.2/4.2/4.2    & 1.0/0.9/0.9 & 1.0--1.1  & 1.0  &    29  &   22.5 & 2.815  & (14) \\
-		120404A     & 9.6/9.6/9.6    & 1.0/0.9/0.9JH & 1.7--1.3 & 1.3 &  15.7  &   21.3 & 2.876  & (2) \\
-		120422A-1   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.3  & 0.6  &  17.2  &   22.0 & 0.283  & (15) \\
-		120422A-2   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.3--1.4  & 0.9  &  113   &   --   & 0.283  & (15) \\
-		120422A-3   & 4.8/4.8/4.8    & 1.0/0.9/0.9 & 1.4--1.7  & 1.0  &  210   &   --   & 0.283  & (15) \\
-		120422A-4   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.3--1.4 & 0.6  & 449   &   --   & 0.283  & (15) \\
-		120422A-5   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.3--1.6 & 0.8  & 593   &   --   & 0.283  & (15) \\
-		120422A-6   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.7--2.4 & 2.5  & 882   &   --   & 0.283  & (15) \\
-		120422A-7   & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.5--1.9 & 1.3  & 906   &   --   & 0.283  & (15) \\
-		120712A     & 4.8/4.8/4.8    & 1.0/0.9/0.9   & 1.5--2.5 & 1.3  & 10.4  &   21.5 & 4.175  & (2) \\
-		120714B     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.5--1.2 & 1.2  &  7.8  &   22.1 & 0.398  & (2)\\
-		120716A$^1$ & 3.6/3.6/3.6    & 1.0/0.9/0.9JH & 1.8--2.6 & 1.0  &  62   &   20.9 & 2.486  & (2) \\
-		120722A$^2$ & 4.8/4.8/4.8    & 1.0/0.9/0.9   & 1.3--1.3 & 1.1  & 10.3  &   23.6 & 0.959  & (2) \\
-		120805A$^2$ & 3.6/3.6/3.6    & 1.0/0.9/0.9JH & 1.3--1.7 & 0.9  & 218   &   --   & 2.8?   & (2) \\
-		120815A     & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.3--1.4 & 0.6  &  1.7  &   20   & 2.358  & (16) \\
-		120909A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.6--1.6 & 1.4  &  1.7  &   21   & 3.929  & (2) \\
-		120923A     & 9.6/9.6/9.6    & 1.0/0.9/0.9JH & 1.2--1.4 & 1.0  & 18.5  &   --   & $\gtrsim8$ & (2) \\
-		121024A     & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.2--1.1 & 0.6  &  1.8  &   20   & 2.300  & (17) \\
-		121027A     & 8.4/8.4/8.4    & 1.0/0.9/0.9   & 1.3--1.3 & 0.9  &  69   &  21.1  & 1.773  & (2) \\
-		121201A     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.1--1.1 & 0.9  & 12.9  &   23   & 3.385  & (2) \\
-		121229A     & 4.8/4.8/4.8    & 1.0/0.9/0.9JH & 1.4--1.2 & 1.4  &  2.0  &  21.5  & 2.707  & (2) \\
-		130131B$^3$ & 7.2/7.2/7.2    & 1.0/0.9/0.9JH & 1.3--1.6 & 0.8  &  --   &   --   & 2.539  & (2) \\
-		130408A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.0--1.0 & 1.0  &  1.9  &   20   & 3.758  & (2) \\
-		130418A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.4--1.3 & 1.3  &  4.6  &   18.5 & 1.218  & (2) \\
-		130427A     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 1.8--1.8 & 0.8  & 16.5  &   19   & 0.340  & (18) \\
-		130427B     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 1.2--1.0 & 0.8  & 20.3  &   22.7 & 2.78   &  (2) \\
-		130603B$^6$ & 2.4/2.4/2.4    & 1.0/0.9/0.9   & 1.4--1.4 & 1.1  &  8.2  &   21.5 & 0.356  & (19) \\
-		130606A     & 4.2/4.2/4.2    & 1.0/0.9/0.9JH & 1.7--1.9 & 1.1  &  7.1  &   19   & 5.91   & (20) \\
-		130612A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 1.3--1.3 & 1.4  &  1.1  &   21.5 & 2.006  & (2) \\
-		130615A     & 1.2/1.2/1.2    & 1.0/0.9/0.9   & 2.1--2.2 & 1.0  &  0.8  &   21   & $~3$?  & (2) \\
-		130701A     & 1.2/1.2/1.2    & 1.0/0.9/0.9JH & 2.0--2.0 & 1.6  &  5.5  &   19.9 & 1.155  & (2) \\
-		\enddata
-		\tablenotetext{1}{Not part of the statistical sample}
-		\tablenotetext{2}{Spectrum dominated by light from the host galaxy}
-		\tablenotetext{3}{Spectrum of the host galaxy taken long after the burst}
-		\tablenotetext{4}{RRM observation}
-		\tablenotetext{5}{ADC malfunction during observation}
-		\tablenotetext{6}{Short burst}
-		\tablerefs{
-			(1) \citet{DeUgartePostigo2010}; (2) This work ; (3) Skuladottir (2010);
-			(4) \citet{DElia2010}; (5) \citet{Wiersema2012}; (6) \citet{Vergani2011, Cobb2010}; 
-			(7) \citet{Thone2013}; (8) \citet{Bufano2011} ; (9) \citet{DeUgartePostigo2011} ;
-			(10) \citet{Hartoog2013}; (11) \citet{Sparre2011}; (12) \citet{Sparre2014};
-			(13) \citet{Levan2014}; (14) \citet{DElia2014}; (15) \citet{Schulze2014};
-			(16) \citet{Kruhler2013}; (17} \citet{Friis2015}
-	\end{deluxetable*}
-	%-------------------------END TABLE---------------------------------------
+	\input{tables/burst_overview.tex}
 
@@
 
-	\subsection{GRB130606A (z = 5.913)}
-	The data presented here also formed the basis of GCN \#
-	14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is published
-	in \citet{Hartoog2015}.
-	The observations consists of three 2x600sec visits starting 7.1 hr after the burst at fairly
-	high airmass.
-	We detect absorption features from a wide range of ions at z=5.913 
-	as well as intervening absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
-
 \end{document}

### 2016-09-24 document/XSGRB_sample.tex

@@
 
+	\subsection{GRB130606A (z = 5.913)}
+	The data presented here also formed the basis of GCN \#
+	14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is published
+	in \citet{Hartoog2015}.
+	The observations consists of three 2x600sec visits starting 7.1 hr after the burst at fairly
+	high airmass.
+	We detect absorption features from a wide range of ions at z=5.913 
+	as well as intervening absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
+
+	\subsection{GRB151021A (z = 2.330)}
+	The data presented here also formed the basis of GCN \#
+	18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
+	elsewhere.
+	The observation was carried out in RRM starting 44 min af the GRB trigger.
+	We detect absorption features from a wide range of ions at z=2.330 
+	as well as intervening absorption at $z=1.49$.
+
+	\subsection{GRB160203A (z = 3.517)}
+	The data presented here also formed the basis of GCN \#
+	18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
+	elsewhere.
+	The observation was carried out in RRM starting 18 min af the GRB trigger.
+	We detect absorption features from a wide range of ions at z=3.517 
+	as well as intervening absorption at $z=2.203$.
+
 \end{document}

### 2016-09-24 document/XSGRB_sample.tex
GRB100724A

@@
 	\input{tables/burst_overview.tex}
+
+	\subsection{RRM observations} \label{RRM}
+	The rapid-response mode is
 
@@
 
+	\subsection{GRB100724A (z = 1.288)}
+	The data presented here also formed the basis of GCN \#
+	10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. 
+	The spectrum has not otherwise been published previously. The observations were
+	carried out in RRM starting 11 min after the GRB trigger. 
+	See section \ref{RRM}, for a description of the RRM scheme.
+	Absorption lines from several ionic species are detected in the afterglow continuum at a common redshift
+	of $z = 1.288$.
+
 	\subsection{GRB120327A (z = 2.813)}
@@
 	elsewhere.
-	The observation was carried out in RRM starting 18 min af the GRB trigger.
+	The observation was carried out in RRM starting 18 min after the GRB trigger.
 	We detect absorption features from a wide range of ions at z=3.517 

### 2016-09-24 document/XSGRB_sample.tex
Small edit

@@
 
-	\subsection{GRB100724A (z = 1.288)}
+	\subsection{GRB100724A* (z = 1.288)}
 	The data presented here also formed the basis of GCN \#
@@
 	Absorption lines from several ionic species are detected in the afterglow continuum at a common redshift
-	of $z = 1.288$.
+	of $z = 1.288$. This burst is not a part of the statistical sample.
 

### 2016-09-24 document/XSGRB_sample.tex
GRB100205A, GRB100219A and GRB100316B

@@
 \newcommand{\cii}{\ion{C}{2}}
+\newcommand{\mgi}{\ion{Mg}{1}} 
 \newcommand{\mgii}{\ion{Mg}{2}} 
@@
 \newcommand{\ali}{\ion{Al}{1}} 
+\newcommand{\alii}{\ion{Al}{2}} 
+\newcommand{\aliii}{\ion{Al}{3}} 
 \newcommand{\SIii}{\ion{Si}{2}} 
@@
 
+	\subsection{GRB100205A}
+	Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected in 10.8 ks. GRB likely located at high redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}.
+
+	\subsection{GRB100219A (z = 4.667)}
+	The data presented here also formed the basis of GCN \#
+	10441\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10441.gcn3}} and is published
+	in \citet{Thone2013}.
+	Observations started 12.5 hours after the \textit{Swift} trigger and has a
+	total exposure time of 4.8 ks.
+	Absorption features, including those of \lya, from a multitude of ions are
+	detected against the afterglow continuum at $z = 4.667$. Additionally,
+	absorption from an intervening system is found at $z = 2.181$.
+
+	\subsection{GRB100316B (z = 1.180)}
+	The data presented here also formed the basis of GCN \#
+	10495\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10495.gcn3}}.
+	Observations started 44 minutes after the \textit{Swift} trigger and has a
+	total exposure time of 2.4 ks.
+	Absorption features from \feii, \alii, \aliii,	\mgii~and \mgi~are well detected against the afterglow continuum at $z = 1.180$.
+Additionally, strong absorption lines from \feii~and \mgii~from an intervening system are found at $z = 1.063$.
+
 	\subsection{GRB100724A* (z = 1.288)}
@@
 	See section \ref{RRM}, for a description of the RRM scheme.
-	Absorption lines from several ionic species are detected in the afterglow continuum at a common redshift
-	of $z = 1.288$. This burst is not a part of the statistical sample.
+	Absorption lines from several ionic species are detected in the afterglow
+	continuum at a common redshift of $z = 1.288$.
+	This burst has been flux calibrated using the master response function and is
+	not a part of the statistical sample.
 
@@
 	high airmass.
-	We detect absorption features from a wide range of ions at z=5.913 
+	We detect absorption features from a wide range of ions at $z=5.913$
 	as well as intervening absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
@@
 	The observation was carried out in RRM starting 44 min af the GRB trigger.
-	We detect absorption features from a wide range of ions at z=2.330 
+	We detect absorption features from a wide range of ions at $z=2.330$ 
 	as well as intervening absorption at $z=1.49$.
@@
 	The observation was carried out in RRM starting 18 min after the GRB trigger.
-	We detect absorption features from a wide range of ions at z=3.517 
+	We detect absorption features from a wide range of ions at $z=3.517$ 
 	as well as intervening absorption at $z=2.203$.

### 2016-09-24 document/XSGRB_sample.tex

@@
 \newcommand{\ha}{H$\alpha$}
+
+\newcommand{\hi}{\ion{H}{1}} 
+\newcommand{\hei}{[\ion{He}{1}]} 
 \newcommand{\oi}{[\ion{O}{1}]} 
 \newcommand{\sii}{[\ion{S}{2}]} 
+\newcommand{\siii}{[\ion{S}{3}]} 
 \newcommand{\oii}{[\ion{O}{2}]$\lambda$3727} 
@@
 \newcommand{\nii}{[\ion{N}{2}]} 
+\newcommand{\neiii}{[\ion{Ne}{3}]} 
 \newcommand{\feii}{\ion{Fe}{2}} 
@@
 \newcommand{\mgii}{\ion{Mg}{2}} 
-\newcommand{\hi}{\ion{H}{1}} 
 \newcommand{\ali}{\ion{Al}{1}} 
@@
 \newcommand{\SIii}{\ion{Si}{2}} 
+\newcommand{\znii}{\ion{Zn}{2}} 
+\newcommand{\crii}{\ion{Cr}{2}} 
+\newcommand{\mnii}{\ion{Mn}{2}} 
+\newcommand{\tiii}{\ion{Ti}{2}} 
+\newcommand{\caii}{\ion{Ca}{2}}
+
 \begin{document}
@@
 
+	\subsection{GRB100316D (z = 0.059)}
+	The data presented here also formed the basis of GCN \#
+	10512\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10512.gcn3}}, GCN \#
+	10513\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10513.gcn3}}, GCN \#
+	10543\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10543.gcn3}} and is published in \citet{Bufano2012} and \citet{Starling2011}.
+This GRB has an associated SN and has therefore undergone intense follow-up.
+	The data presented here consists only of a subset of the entire VLT/X-shooter campaign. The transition from
+
+	\subsection{GRB100418A (z=0.624)}
+	The data presented here also formed the basis of GCN \#
+	10620\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10620.gcn3}} and GCN \#
+	10631\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10631.gcn3}} and is published
+	in \citet{DeUgartePostigo2011}.
+	The burst have been followed up in three epochs of observations, 0.4, 1.4, and
+	2.4 days after the burst, each lasting 4.8 ks.
+	The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous detection of emission features belonging to nebular lines, including \hi, \oii, \oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all at a consistent redshift.
+	Temporal evolution of the fine-structure lines belonging to \feii$^*$ is found between the epochs.
+
 	\subsection{GRB100724A* (z = 1.288)}

### 2016-09-24 document/XSGRB_sample.tex
GRB100424A, GRB100425A, GRB100615A and GRB100621A
Notes on individuals bursts

@@
 	\subsection{GRB100205A}
-	Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected in 10.8 ks. GRB likely located at high redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}.
+	Observed 3 days after the \textit{Swift} trigger.
+No afterglow or host detected in 10.8 ks.
+GRB likely located at high redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}.
+The spectrum has not otherwise been published previously.
 
@@
 	10495\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10495.gcn3}}.
+	The spectrum has not otherwise been published previously.
 	Observations started 44 minutes after the \textit{Swift} trigger and has a
@@
 
+	\subsection{GRB100424A (z=2.465)}
+	The data presented here also formed the basis of GCN \#
+	14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}.
+	The spectrum has not otherwise been published previously.
+	Observations carried out, long after the burst has faded.  Emission lines from
+	the host are detected at $z=2.465$.
+
+	\subsection{GRB100425A (z=1.1755)}
+	The data presented here also formed the basis of GCN \#
+	10684\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10684.gcn3}} and is used in
+	\citet{Skuladottir2010}, but not published elsewhere.
+	Observations started 4 hours after the \textit{Swift} trigger, totaling 2.4 ks. 
+	Absorption features from \mgii~and \feii~in the afterglow continuum are
+	detected at $z=1.1755$.
+
+	\subsection{GRB100615A (z=1.398)}
+	The data presented here also formed the basis of GCN \#
+	14264\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14264.gcn3}}, 
+	but not published elsewhere.
+	Host observation of a dark burst\citep{DElia2011} taken long after the afterglow has faded.
+Emission lines from the host belonging to \oii, \neiii, \oiii~and \ha~are detected at a common redshift of $z=1.398$.
+
+	\subsection{GRB100621A (z=0.542)}
+	The data presented here also formed the basis of GCN \#
+	10876\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10876.gcn3}}, 
+	but not published elsewhere.
+	Beginning 7.1 hours after the GRB, 2.4 ks observations reveal emission lines from \oii, \hb~and \oiii~at a common redshift of $z=0.542$.
+
 	\subsection{GRB100724A* (z = 1.288)}
@@
 	continuum at a common redshift of $z = 1.288$.
-	This burst has been flux calibrated using the master response function and is
-	not a part of the statistical sample.
+	This is not a part of the statistical sample.
 

### 2016-09-25 document/XSGRB_sample.tex

@@
 	but not published elsewhere.
-	Beginning 7.1 hours after the GRB, 2.4 ks observations reveal emission lines from \oii, \hb~and \oiii~at a common redshift of $z=0.542$.
+	Beginning 7.1 hours after the GRB, 2.4 ks observations reveal emission lines from \oii, \hb~and \oiii~at a common redshift of $z=0.542$ and a very weak afterglow continuum.
+
+	\subsection{GRB100625A (z=0.452)}
+	The data presented here is of the candidate host galaxy, taken long after the burst has faded and have not previously been published.
+4.8 ks of exposure reveals a weak continuum present in all arms, but an absence of emission lines.
+This could indicate that the host primarily contains a older stellar population.
+The redshift, $z=0.452$, is taken from \citet{Fong2013}.
 

### 2016-09-25 document/XSGRB_sample.tex

@@
 	\subsection{GRB100205A}
-	Observed 3 days after the \textit{Swift} trigger.
-No afterglow or host detected in 10.8 ks.
-GRB likely located at high redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}.
-The spectrum has not otherwise been published previously.
+	Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected
+	in 10.8 ks. GRB likely located at high
+	redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}. The spectrum
+	has not otherwise been published previously.
 
@@
 	10441\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10441.gcn3}} and is published
-	in \citet{Thone2013}.
-	Observations started 12.5 hours after the \textit{Swift} trigger and has a
-	total exposure time of 4.8 ks.
-	Absorption features, including those of \lya, from a multitude of ions are
-	detected against the afterglow continuum at $z = 4.667$. Additionally,
-	absorption from an intervening system is found at $z = 2.181$.
+	in \citet{Thone2013}. Observations started 12.5 hours after the \textit{Swift}
+	trigger and has a total exposure time of 4.8 ks. Absorption features, including
+	those of \lya, from a multitude of ions are detected against the afterglow
+	continuum at $z = 4.667$. Additionally, absorption from an intervening system
+	is found at $z = 2.181$.
 
@@
 	The data presented here also formed the basis of GCN \#
-	10495\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10495.gcn3}}.
-	The spectrum has not otherwise been published previously.
-	Observations started 44 minutes after the \textit{Swift} trigger and has a
-	total exposure time of 2.4 ks.
-	Absorption features from \feii, \alii, \aliii,	\mgii~and \mgi~are well detected against the afterglow continuum at $z = 1.180$.
-Additionally, strong absorption lines from \feii~and \mgii~from an intervening system are found at $z = 1.063$.
+	10495\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10495.gcn3}}. The spectrum
+	has not otherwise been published previously. Observations started 44 minutes
+	after the \textit{Swift} trigger and has a total exposure time of 2.4 ks.
+	Absorption features from \feii, \alii, \aliii,	\mgii~and \mgi~are well detected
+	against the afterglow continuum at $z = 1.180$. Additionally, strong absorption
+	lines from \feii~and \mgii~from an intervening system are found at $z = 1.063$.
 
@@
 	10513\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10513.gcn3}}, GCN \#
-	10543\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10543.gcn3}} and is published in \citet{Bufano2012} and \citet{Starling2011}.
-This GRB has an associated SN and has therefore undergone intense follow-up.
-	The data presented here consists only of a subset of the entire VLT/X-shooter campaign. The transition from
+	10543\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10543.gcn3}} and is published
+	in \citet{Bufano2012} and \citet{Starling2011}. This GRB has an associated SN
+	and has therefore undergone intense follow-up. The data presented here consists
+	only of a subset of the entire VLT/X-shooter campaign. The transition from
 
@@
 	10631\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10631.gcn3}} and is published
-	in \citet{DeUgartePostigo2011}.
-	The burst have been followed up in three epochs of observations, 0.4, 1.4, and
-	2.4 days after the burst, each lasting 4.8 ks.
-	The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous detection of emission features belonging to nebular lines, including \hi, \oii, \oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all at a consistent redshift.
-	Temporal evolution of the fine-structure lines belonging to \feii$^*$ is found between the epochs.
+	in \citet{DeUgartePostigo2011}. The burst have been followed up in three epochs
+	of observations, 0.4, 1.4, and 2.4 days after the burst, each lasting 4.8 ks.
+	The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous
+	detection of emission features belonging to nebular lines, including \hi, \oii,
+	\oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due
+	to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii,
+	all at a consistent redshift. Temporal evolution of the fine-structure lines
+	belonging to \feii$^*$ is found between the epochs.
 
@@
 	The data presented here also formed the basis of GCN \#
-	14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}.
-	The spectrum has not otherwise been published previously.
-	Observations carried out, long after the burst has faded.  Emission lines from
-	the host are detected at $z=2.465$.
+	14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}. The spectrum
+	has not otherwise been published previously. Observations carried out, long
+	after the burst has faded.  Emission lines from the host are detected at
+	$z=2.465$.
 
@@
 	10684\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10684.gcn3}} and is used in
-	\citet{Skuladottir2010}, but not published elsewhere.
-	Observations started 4 hours after the \textit{Swift} trigger, totaling 2.4 ks. 
-	Absorption features from \mgii~and \feii~in the afterglow continuum are
-	detected at $z=1.1755$.
+	\citet{Skuladottir2010}, but not published elsewhere. Observations started 4
+	hours after the \textit{Swift} trigger, totaling 2.4 ks. Absorption features
+	from \mgii~and \feii~in the afterglow continuum are detected at $z=1.1755$.
 
@@
 	The data presented here also formed the basis of GCN \#
-	14264\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14264.gcn3}}, 
-	but not published elsewhere.
-	Host observation of a dark burst\citep{DElia2011} taken long after the afterglow has faded.
-Emission lines from the host belonging to \oii, \neiii, \oiii~and \ha~are detected at a common redshift of $z=1.398$.
+	14264\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14264.gcn3}}, but not
+	published elsewhere. Host observation of a dark burst\citep{DElia2011} taken
+	long after the afterglow has faded. Emission lines from the host belonging to
+	\oii, \neiii, \oiii~and \ha~are detected at a common redshift of $z=1.398$.
 
@@
 	The data presented here also formed the basis of GCN \#
-	10876\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10876.gcn3}}, 
-	but not published elsewhere.
-	Beginning 7.1 hours after the GRB, 2.4 ks observations reveal emission lines from \oii, \hb~and \oiii~at a common redshift of $z=0.542$ and a very weak afterglow continuum.
+	10876\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10876.gcn3}}, but not
+	published elsewhere. Beginning 7.1 hours after the GRB, 2.4 ks observations
+	reveal emission lines from \oii, \hb~and \oiii~at a common redshift of
+	$z=0.542$ and a very weak afterglow continuum.
 
 	\subsection{GRB100625A (z=0.452)}
-	The data presented here is of the candidate host galaxy, taken long after the burst has faded and have not previously been published.
-4.8 ks of exposure reveals a weak continuum present in all arms, but an absence of emission lines.
-This could indicate that the host primarily contains a older stellar population.
-The redshift, $z=0.452$, is taken from \citet{Fong2013}.
+	The data presented here is of the candidate host galaxy, taken long after the
+	burst has faded and have not previously been published. 4.8 ks of exposure
+	reveals a weak continuum present in all arms, but an absence of emission lines.
+	This could indicate that the host primarily contains a older stellar
+	population. The redshift, $z=0.452$, is taken from \citet{Fong2013}.
 
@@
 	The data presented here also formed the basis of GCN \#
-	10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. 
-	The spectrum has not otherwise been published previously. The observations were
-	carried out in RRM starting 11 min after the GRB trigger. 
-	See section \ref{RRM}, for a description of the RRM scheme.
-	Absorption lines from several ionic species are detected in the afterglow
-	continuum at a common redshift of $z = 1.288$.
-	This is not a part of the statistical sample.
+	10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. The spectrum
+	has not otherwise been published previously. The observations were carried out
+	in RRM starting 11 min after the GRB trigger. See section \ref{RRM}, for a
+	description of the RRM scheme. Absorption lines from several ionic species are
+	detected in the afterglow continuum at a common redshift of $z = 1.288$. This
+	is not a part of the statistical sample.
 
@@
 	13134\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13134.gcn3}} and is published
-	in \citet{DElia2014}.
-	The observation consists of two visits, 2.13 hrs and 29.98 hrs after the burst,
-	with an afteglow continuum visible in all arms for both visits.
-	We detect absorption features from Ly-limit, \lya, \cii/\cii$^*$, \SIii/\SIii$^*$, \ali,
-	\feii ~and \mgii ~are detected at a consistent redshift, $z = 2.813$.
+	in \citet{DElia2014}. The observation consists of two visits, 2.13 hrs and
+	29.98 hrs after the burst, with an afteglow continuum visible in all arms for
+	both visits. We detect absorption features from Ly-limit, \lya, \cii/\cii$^*$,
+	\SIii/\SIii$^*$, \ali, \feii ~and \mgii ~are detected at a consistent redshift,
+	$z = 2.813$.
 
@@
 	14365\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14365.gcn3}}. The spectrum
-	has not otherwise been published previously.
-	The observations consists of two 600sec spectra taken 1.9hr after the burst.
-	We detect absorption features from a wide range of ions. We also detect intervening absorption at $z=1.255$
-	and $z=3.248$.
+	has not otherwise been published previously. The observations consists of two
+	600sec spectra taken 1.9hr after the burst. We detect absorption features from
+	a wide range of ions. We also detect intervening absorption at $z=1.255$ and
+	$z=3.248$.
 
@@
 	14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is published
-	in \citet{Hartoog2015}.
-	The observations consists of three 2x600sec visits starting 7.1 hr after the burst at fairly
-	high airmass.
-	We detect absorption features from a wide range of ions at $z=5.913$
-	as well as intervening absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
+	in \citet{Hartoog2015}. The observations consists of three 2x600sec visits
+	starting 7.1 hr after the burst at fairly high airmass. We detect absorption
+	features from a wide range of ions at $z=5.913$ as well as intervening
+	absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
 
@@
 	The data presented here also formed the basis of GCN \#
-	18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
-	elsewhere.
-	The observation was carried out in RRM starting 44 min af the GRB trigger.
-	We detect absorption features from a wide range of ions at $z=2.330$ 
-	as well as intervening absorption at $z=1.49$.
+	18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
+	published elsewhere. The observation was carried out in RRM starting 44 min af
+	the GRB trigger. We detect absorption features from a wide range of ions at
+	$z=2.330$ as well as intervening absorption at $z=1.49$.
 
@@
 	The data presented here also formed the basis of GCN \#
-	18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not published
-	elsewhere.
-	The observation was carried out in RRM starting 18 min after the GRB trigger.
-	We detect absorption features from a wide range of ions at $z=3.517$ 
-	as well as intervening absorption at $z=2.203$.
+	18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
+	published elsewhere. The observation was carried out in RRM starting 18 min
+	after the GRB trigger. We detect absorption features from a wide range of ions
+	at $z=3.517$ as well as intervening absorption at $z=2.203$.
 

### 2016-09-25 document/XSGRB_sample.tex

@@
 
+	\subsection{GRB100728B (z=2.106)}
+	The data presented here also formed the basis of GCN \#
+	11317\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11317.gcn3}}. The spectrum
+	has not otherwise been published previously. Starting 22 hours after the burst
+	trigger, 7.2 ks of observations reveals a faint afterglow continuum with \lya-
+	and \mgii-absorption at $z=2.106$. Due to a malfunctioning ADC, the sensitivity
+	of X-shooter is depressed with respect to normal operations, resulting in a
+	poorer throughout. Additionally, the position of the trace on the slit moves
+	due to atmospheric differential refraction.
+
 	\subsection{GRB120327A (z = 2.813)}
@@
 	features from a wide range of ions at $z=5.913$ as well as intervening
-	absorption at $z=2.313, 2.521, 3.415, 4.6448, 4.6468, 4 and 4.6495$.
+	absorption at $z=2.3103, 2.5207, 3.4515, 4.4660, 4.5309, 4.5427, 4.6497 $ and $
+	4.7244$.
 

### 2016-09-26 document/XSGRB_sample.tex
text and table update for GRB100814A

@@
 
+	\subsection{GRB100814A (z=1.439)}
+    The spectra presented here has not been published previously. The
+    observations consists of three visits, the first beginning only 0.9 hours
+    after the \textit{Swift} trigger, the other two visits were 2.13 and 98.40
+    hours after the trigger, respectively. A bright afterglow continuum is
+    present in all visits, allowing identification of absorption features
+    belonging to a wide range of ions at $z=1.439$. A complex velocity structure
+    in the absorption features belonging to \mgii, shows several components,
+    separated by as much as 5000km/s, pointing to a likely merger scenario in
+    the host.
+
 	\subsection{GRB120327A (z = 2.813)}
@@
 	has not otherwise been published previously. The observations consists of two
-	600sec spectra taken 1.9hr after the burst. We detect absorption features from
+	600sec spectra taken 1.9hrs after the burst. We detect absorption features from
 	a wide range of ions. We also detect intervening absorption at $z=1.255$ and
@@
 	in \citet{Hartoog2015}. The observations consists of three 2x600sec visits
-	starting 7.1 hr after the burst at fairly high airmass. We detect absorption
+	starting 7.1 hrs after the burst at fairly high airmass. We detect absorption
 	features from a wide range of ions at $z=5.913$ as well as intervening
@@
 	18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
-	published elsewhere. The observation was carried out in RRM starting 44 min af
-	the GRB trigger. We detect absorption features from a wide range of ions at
-	$z=2.330$ as well as intervening absorption at $z=1.49$.
+	published elsewhere. The observation was carried out in RRM starting 44 minutes
+	after the GRB trigger. We detect absorption features from a wide range of ions
+	at $z=2.330$ as well as intervening absorption at $z=1.49$.
 
@@
 	18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
-	published elsewhere. The observation was carried out in RRM starting 18 min
+	published elsewhere. The observation was carried out in RRM starting 18 minutes
 	after the GRB trigger. We detect absorption features from a wide range of ions

### 2016-09-26 document/XSGRB_sample.tex
GRB100816A

@@
     in the absorption features belonging to \mgii, shows several components,
-    separated by as much as 5000km/s, pointing to a likely merger scenario in
+    separated by as much as 500km/s, pointing to a likely merger scenario in
     the host.
+
+	\subsection{GRB100816A (z=0.805)}
+	The data presented here also formed the basis of GCN \#
+	11123\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11123.gcn3}}. The spectrum
+	has not otherwise been published previously. This short GRB was observed 28.4
+	hours after the GRB trigger. 4 x 1200 s of exposure reveals two distinct sets
+	of emission lines, spatially offset $\lesssim 1 \arcsec $, very close in
+	redshift space, $z=0.8034$ and $z=0.8049$, indicating either an interacting
+	host or some complex velocity structure of the host. Faint underlying continua
+	are present under both sets of lines.
 

### 2016-09-27 document/XSGRB_sample.tex
GRB100901A

@@
 
+	\subsection{GRB100901A(z=1.408)}
+	The data presented here has been published in \citet{Hartoog2013}. Because of
+	the unusual lingering brightness of this GRB, 2.4s of observations taken 65.98
+	hours after the GRB trigger still reveals an afterglow continuum visible across
+	the entire spectral coverage of X-shooter. Absorption lines from a wide range
+	ion put the redshift at $z=1.408$, with intervening absorption systems at $z =
+	1.3147$ and $z = 1.3179$.
+
 	\subsection{GRB120327A (z = 2.813)}

### 2016-09-29 document/XSGRB_sample.tex
Adding GRB100316D

@@
 	\begin{abstract}
-		Later
+	The {\it Swift} satellite allows us to use gamma-ray bursts (GRBs) to peer into 
+	the hearts of star forming galaxies through cosmic time. Our open
+	collaboration, representing most of the active ESO member researchers in this
+	field, seeks to build a public legacy sample of GRB X-shooter spectroscopy
+	while {\it Swift} continues to fly. We propose to continue our programme to target
+	all suitably observable GRB afterglows (up to 15 bursts per semester), with the
+	primary goal of producing a well-defined, homogeneous, statistically useful
+	sample. To date, our spectroscopy covers a redshift range from 0.059 to about
+	8, with more than 20 robust metallicity measurements from absorption lines (over 
+	the redshift range 1.7--5.9) and 4 secure detections of H$_2$ or CH molecular
+	absorption. Such information is extremely difficult to obtain by other means.
+	In terms of studying the spread and redshift evolution in gas-phase properties, the
+	sample is still limited by low-number statistics. 
 	\end{abstract}
@@
 
+	Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+	towards about 10$^5$ QSOs have 5 systems with 
+	$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$
+	been identified (Noterdaeme et al.\ 2012, A\&A, 547, L1). Long GRB afterglow
+	spectra, by contrast, reveal such systems in the majority of cases (e.g., Jakobsson et
+	al.\ 2006, A\&A, 460, L13; Fynbo et al.\ 2009, ApJS, 185, 526).
+	Whereas DLAs towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due
+	to the atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing
+	redshifts (e.g., Rafelski et al.\ 2014, ApJL, 782, L29), GRBs allow us to see 
+	into the hearts of star-forming galaxies over
+	the full history of cosmic star formation from $z \approx 0$ to $z > 8$ (e.g.,
+	Tanvir et al.\ 2009, Nature, 461, 1254; Salvaterra et al.\ 2009, Nature, 461,
+	1258; Jakobsson et al.\ 2012, ApJ, 752, 62). With afterglow spectroscopy
+	(throughout the electromagnetic spectrum from X-rays to the sub-mm) we can hence
+	characterize the properties of star-forming galaxies over cosmic history in
+	terms of  redshifts, metallicities, molecular contents, ISM temperatures,
+	UV-flux densities, etc.. This is, however, only possible as long as there are
+	satellites in orbit that rapidly and accurately locate GRBs. The currently
+	operating \textit{Swift} satellite, launched in 2004 and still fully-functioning,
+	allows for very efficient follow-up observations of GRBs due to its
+	unprecedented rate, speed and precision of localisations.
+
+	%With GTO time on X-shooter we
+	%have started building a sample of afterglows with X-shooter spectroscopy.
+	%X-shooter is ideal for this with its high efficiency, wide wavelength coverage
+	%and spectral resolution sufficient for inferring the fundamental physical
+	%parameters just listed.
+
+	\smallskip
+	\smallskip
+
+	{\bf The case for a large sample of GRBs with X-shooter spectroscopy.}
+
+	There are
+	several unanswered but fundamental questions that must be addressed in order
+	to exploit the full potential of GRBs as cosmological probes.
+	%Among them: what is the true redshift distribution and corresponding luminosity
+	%function of long-duration GRBs?
+	More than 50\% of the \textit{Swift} bursts with measured redshift are at $z >
+	2$, and  5--10\% are expected to be above $z=5$ (Salvaterra et al. 2008, MNRAS, 380, L45; Salvaterra et al.\ 2012, ApJ, 749,
+	68; Jakobsson et al.\ 2012, ApJ, 752, 62; Perley et al.\ 2016, ApJ, 817, 7). 
+	A high redshift completeness is
+	crucial for our understanding of the link between the number density of GRBs
+	per unit redshift and the global star-formation history of the Universe, as
+	measured by other means (UV, FIR, sub-mm, see Robertson \& Ellis 2011, ApJ, 744,
+	95). The detection of GRBs at $z > 6$ shows that GRBs have become competitive as
+	a tool to identifying galaxies at the highest redshifts and unsurpassed in
+	providing detailed abundance information via absorption line spectroscopy
+	(Tanvir et al.\ 2012, ApJ, 754, 46, McGuire et al.\ 2016, ApJ submitted, 
+	arxiv:1512.07808).
+
+	\smallskip
+
+	%The sample of short GRBs with redshifts remains too small and too
+	%incomplete to really
+	%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
+	%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
+	%enlarge as far as possible the number of known redshifts for this class of
+	%sources.
+
+	From March 2005 to March 2016 there have been about 350 {\it Swift} bursts 
+	complying with
+	our sample selection criteria (see ``Immediate objective''), and about half of them have
+	measured redshifts. Among the latter subset, the team proposing these observations has
+	measured about two thirds of the redshifts, mainly with FORS1/2 and X-shooter (see
+	Fynbo et al.\ 2009, ApJS, 185, 526;  Jakobsson et al.\ 2012, ApJ, 752, 62;
+	Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our current aim is to build a sample superior to
+	our previous low-resolution survey (Fynbo et al.\ 2009, ApJS, 185, 526), both in
+	terms of quantity and  quality of the spectra. Our program started as 
+	guaranteed time observations during
+	periods 84-91, and we have continued in open time since then.
+
+	%In P84 we started the X-shooter legacy survey in GTO time. Our aim is to  build
+	%a sample of GRBs superior to our previous  low-resolution  survey, Fynbo et al.\
+	%2009, ApJS, 185, 526) both in terms of quantity and  quality of the spectra. The
+	%last period of GTO time on the survey was P90 (plus 6 hr of leftover GTO time
+	%in  P91) and for P92 a proposal like this was  accepted for continuation of the
+	%program in open time (concerning P93 see box 8).
+
+	\smallskip
+
+	X-shooter is in many ways
+	the ideal GRB follow-up instrument and indeed GRB follow-up was one of the
+	primary science cases behind the instrument design and implementation. Our
+	program secures general purpose GRB afterglow spectroscopic follow-up 
+	that adds strong legacy value to the \textit{Swift} GRB sample.
+	Due to the wide wavelength range of X-shooter  
+	with the same observation cover molecular H$_2$ absorption near the
+	atmospheric cut-off and all the strong emission lines from the host in the
+	NIR arm (e.g., Friis et al., 2015, MNRAS, 451, 167). In general, the wide wavelength coverage ensures that we always have
+	features on which to base a redshift measurement as long as the afterglow is
+	brighter than about 23 mag in either the $R$- or $z$-band. 
+	Frequently, emission lines are also detected from the underlying host, which
+	also provide further information such as SFR and metallicity (the top right 
+	panel in Fig.~1 shows an example). Only for 
+	7 out of more than 70 secured spectra could we not measure a redshift.   
+	With the X-shooter survey we
+	provide {\bf metallicity measurements} for about 30\% (Voigt-profile fits)
+	of the $z>1.7$ events.
+	So far we have
+	measured metallicities for more than 20 GRB afterglows with X-shooter.
+	With the wide wavelength coverage of X-shooter we can
+	study important chemical species as Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$
+	elements over a much wider redshift range than what is possible with other
+	instruments.
+	As an example, we have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A
+	at $z=4.669$ (Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for 
+	GRB\,111008A at $z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$
+	for the $z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139).  
+	Reconciling the
+	abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs
+	in particular, and old stars in the Local Group is an important long-term
+	goal (see also Sparre et al.\ 2014, ApJ, 785, 150).
+	Metallicities are also measured from host emission lines (Kr{\"u}hler et al.\ 2015,
+	A\&A, 581, A125).
+	GRB spectroscopy also allows us to determine the dust content of their environments,
+	both through analysis of the depletion pattern and through measurement of the
+	associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). 
+	This allows us to quantify the dust-to-metals ratio and its
+	evolution with redshift (e.g., De Cia et al.\ 2013, A\&A, 560, 88; Zafar \&
+	Watson 2013, A\&A, 560, 26).
+	%Moreover, GRB
+	%afterglows allow investigation of random foreground absorbers, complementing
+	%QSO sightline studies (e.g., Schulze et
+	%al.\ 2012, A\&A, 546, 20).
+
+	%In Fynbo et al. (2009, ApJS, 185,
+	%526), we compare the host-galaxy HI
+	%column density and metallicity (most of which has been measured by our group)
+	%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
+	%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
+	%column density and metallicity for GRB and QSO sightlines shows that the total
+	%cross-section for environments like those of GRBs must be less than a few
+	%percent of the total cross-section for QSO-DLAs, and suggests that GRB
+	%sightlines probe regions with high gas density.
+
+	%The issue of production and {\bf release of Lyman continuum radiation} is
+	%currently not fully understood, but the evidence suggests that star-forming
+	%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
+	%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
+	%X-shooter spectroscopy will allow major progress due to the excellent blue
+	%sensitivity.
+
+	%GRB afterglows can probe dust in high-redshift star-forming galaxies
+	%including exploring the {\bf 2175~\AA{} extinction bump}
+	%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
+	%732, 143).
+	%Due to the broad wavelength  coverage we can do
+	%this more accurately and over a broader range of redshifts with X-shooter.
+
+	\smallskip
+
+	We will also determine the frequency and properties of {\bf molecular
+		absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
+	the process of star formation, but prior to our program H$_2$ had been detected
+	just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
+	in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
+	found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
+	2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). 
+	We are currently analysing more of our spectra for less obvious molecular
+	absorption and we expect to find more (a dedicated sample paper is 
+	addressing this issue).
+
+	\vspace{0.2cm}
+	A natural question to ask is: {\bf how long should this work continue?} Our
+	view is that we need to keep observing the afterglows as long as we have
+	\textit{Swift} in operation. Also note that the program is still producing many 
+	papers and provides data for many theses (Box 9 and 10). \textit{Swift} is currently 
+	funded until 2018, but is likely to get more extensions given its overwhelming 
+	success. As
+	mentioned above GRBs allow us to probe star-forming galaxies that are almost
+	impossible to study in other ways both in terms of redshifts, galaxy luminosity
+	function, and regions within galaxies. After 7 periods we have 
+	secured seven spectra of $z>4$ GRBs, of which three were of sufficient quality to
+	allow abundance measurements (Th\"one et al.\ 2013, Sparre et al.\ 2014, 
+	Hartoog et al.\ 2015). GRBs offer the only way to derive chemical abundances 
+	for the gas phase of central, actively star-forming regions of high-$z$ galaxies. 
+	The program also maintains a very high discovery
+	potential where we occasionally find something completely unexpected that 
+	provides interesting clues to puzzles in other fields, e.g. extinction of 
+	type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of these spectra are like
+	precious jewels – it is a type of observation that can never be repeated and a
+	class of sightlines that can only be studied while we have operating GRB
+	satellites.
+
+	It is also worth adding that we have build up a rather unique team spread over
+	Europe from Granada to Reykjavik, which by now has reached a point where the 
+	distribution of night shifts, the scientific exploitation of the data is 
+	efficient and where we are open to all new members who wish to participate.
+	As mentioned all data are public immediately.
+
+	{\bf For all of these reasons, we need to keep
+		building up the sample of GRB afterglow spectra now as we may have to wait many
+		years before a mission like \textit{Swift} becomes available again.}
+
+	A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark bursts}'', e.g.,
+	Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012, MNRAS, 421, 1265). 
+	Some of these are at the highest
+	redshifts ($z > 6$) and their observer-frame optical emission is absorbed by
+	the IGM. The majority, however, suffer from large dust obscuration (e.g., Perley
+	et al. 2009, AJ, 138, 1690; Greiner et al.\ 2011, A\&A, 526, 30).
+	Identifying such GRBs is important for
+	constraining the fraction of obscured star formation. In both cases,
+	NIR emission is
+	expected. X-shooter
+	can adequately study these objects, provided that a NIR counterpart is timely
+	identified, for which we have the dedicated HAWKI run D.
+
+	The detection of {\bf absorption line variability} can reveal the burst
+	influence on the surrounding medium and in turn the absorber distance from the
+	burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
+	2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\
+	2012, A\&A, 545, 64; Hartoog et al.\ 2013).
+	{\bf Short GRBs} originate in a substantially different
+	environment compared to long GRBs. Short GRBs may be related to the merging of
+	compact binaries and the coalescence time can be long enough to allow the
+	progenitor system to move far away from the star formation site (Belczynski et
+	al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum with a sufficient 
+	signal-to-noise ratio
+	of a short GRB afterglow has been secured.
+	A knowledge of the redshift distribution
+	of short bursts is of key importance for the next generation of gravitational
+	wave experiments, as they are the likely EM counterparts to their
+	primary targets.
+
+	\cleardoublepage
+
+	\section{Sample selection criteria}
+
+	Being of transient nature, it is difficult to impose strong sample selection
+	criteria on GRBs, without hampering the follow-up effort. Many natural follow-up
+	restrictions exists from already, being it weather conditions, pointing
+	restrictions of the telescope or poorly localized bursts as reported by the
+	\textit{Swift}-telescope. To maximize the return of the follow-up campaign
+	we have chosen a few selection criteria that attempts to provide an unbiased
+	selection of bursts, while allowing for a high success-rate
+
+	\footnote{Note that in the P84 proposal the criteria have
+	been stated a bit differently, the visibility constraint being replaced by a
+	declination + Sun angle constraint. The above criteria are however those
+	defining the sample.}
+
+	\begin{enumerate}
+		\item GRB triggered onboard by Swift.
+		\item Galactic $A_V \lesssim 0.5$ mag.
+		\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
+		\item The target must be visible for at least 60 min at least 30° above horizon, with the Sun below -12°.
+		\item No bright closeby stars.
+
+	\end{enumerate}
+
+	\cleardoublepage
+
 	\section{Observations}
-
-	\input{tables/burst_overview.tex}
 
@@
 	The rapid-response mode is
+
+	\section{Reduction scheme}
+	\input{tables/burst_overview.tex}
 
@@
 	10543\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10543.gcn3}} and is published
-	in \citet{Bufano2012} and \citet{Starling2011}. This GRB has an associated SN
-	and has therefore undergone intense follow-up. The data presented here consists
-	only of a subset of the entire VLT/X-shooter campaign. The transition from
+	in \citet{Bufano2012} and \citet{Starling2011}. This GRB is very close by and
+	has an associated SN and has therefore undergone intense follow-up. The data
+	presented here consists of a subset of the entire VLT/X-shooter campaign,
+	covering the four first observing days while the afterglow still contributes
+	significantly to the total emission. The first observations started 10 hours
+	after the burst, before the SN was discovered, and targeted the star-forming
+	'A'-region\citep{Starling2011}, not the GRB. A very rich spectrum containing a
+	multitude of emission lines puts the host at $z = 0.059$. For three consequtive
+	nights, 58, 79 and 101 hours after the \textit{Swift} trigger, the afterglow
+	was observed as it transitioned into the spectrum of a high-velocity Ic-BL SN.
+	The observations taken 79 and 101 hours after the burst are taken under
+	programme 084.D-0265(A) (PI: Benetti), but with a similar setup to first two
+	observations.
 

### 2016-09-29 document/XSGRB_sample.tex
Adding GRB101219A and GRB101219B

@@
 	in \citet{Bufano2012} and \citet{Starling2011}. This GRB is very close by and
-	has an associated SN and has therefore undergone intense follow-up. The data
+	has an associated SN, SN2010bh, and has therefore undergone intense follow-up. The data
 	presented here consists of a subset of the entire VLT/X-shooter campaign,
@@
 	The observations taken 79 and 101 hours after the burst are taken under
-	programme 084.D-0265(A) (PI: Benetti), but with a similar setup to first two
-	observations.
+	programme 084.D-0265(A) (PI: Benetti), but with an identical setup to the first
+	two observations.
 
@@
 
-	\subsection{GRB100901A(z=1.408)}
+	\subsection{GRB100901A (z=1.408)}
 	The data presented here has been published in \citet{Hartoog2013}. Because of
@@
 	1.3147$ and $z = 1.3179$.
+
+	\subsection{GRB101219A (z=0.718)}
+	This data has not been published before. Starting 3.7 hours after the GRB
+	trigger, 7.2 ks of exposure time reveals a very faint continuum in the visual
+	and near-infrared, only visible when heavily binning the images. No redshift
+	estimate is available from these observations.  Late-time Gemini-North
+	observations reveal emission lines from the host at
+	$z=0.718$\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11518.gcn3}}.
+
+	\subsection{GRB101219B (z=0.552)}
+	The data presented here also formed the basis of GCN \#
+	11579\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11579.gcn3}} and is published
+	in \citet{Sparre2011}.	The first observation, taken 11.6 hours after the burst
+	trigger and lasting 4.8ks, reveals absorption from \mgii~and \mgi~in the host
+	located at $z = 0.552$ on a featureless continuum visible across the entire
+	coverage of X-shooter.  Subsequent observations taken 16 and 37 days after the
+	trigger shows the fading spectral signature of a SN, SN2010ma.
 

### 2016-09-30 document/XSGRB_sample.tex

@@
 \newcommand{\ha}{H$\alpha$} 
-
+\newcommand{\hg}{H$\gamma$} 
 \newcommand{\hi}{\ion{H}{1}} 
@@
 
+	\subsection{GRB160804A (z = 0.736)}
+	The data presented here also formed the basis of GCN \#
+	19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}} and is not
+	published elsewhere. Observations started 22.37 hours after the BAT trigger and
+	lasted for 2.4ks. The afterglow continuum is detected across the entire
+	spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii and
+	\alii~are found at $z = 0.736$. At the same redshift, emission lines from \oii,
+	\oiii, \ha, \hb, \hg, \nii, \sii, \siii are found. A second epoch, lasting
+	3.6ks, is obtained after the afterglow has faded, confirming the emission line
+	detections.
+
 \end{document}

### 2016-10-07 document/XSGRB_sample.tex
Update XSGRB_sample.tex
GRB120118B, GRB120119A, GRB120211A

@@
 	\section{Reduction scheme}
-	\input{tables/burst_overview.tex}
+
+	Additionally, the wavelength solution for all bursts have been re-calibrated.
+Using a synthetic sky spectrum \citep{Noll2012, Jones2013}, a first refinement of the wavelength solution have been obtained by cross-correlating with the observed sky.
+
+	Both a multiplicative and an additive offset has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
 
 	\section{Results}
-
+	\input{tables/burst_overview.tex}	
 	\subsection{Spectral resolution}
@@
 
+	\subsection{GRB120118B (z =	2.943)}
+	The data presented here also formed the basis of GCN \#
+	14225\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14225.gcn3}}, but is not
+	published otherwise. This late-time observation of the host of GRB120118B
+	consists of 3.6 ks exposures and contains emission lines belonging to \oii and
+	\oiii at $z =	2.943$, suggested to be redshift of the host.
+
+	\subsection{GRB120119A (z =	1.728)}
+	The data presented here has not been published before. Three epochs of
+	observations have been obtained, the first two immediately after the burst, and
+	the last one long after the afterglow had faded. Starting 1.4 hours after the
+	\textit{Swift} trigger, the first epoch contains bright afterglow continuum.
+	Rich in absorption features belonging to a multitude of ions, $z =	1.728$ is
+	estimated for the host with intervening systems at $z =	1.476$, $z = 1.214$, $z
+	= 0.662$ and $z = 0.632$. The second epoch, obtained 4.5 hours after the burst
+	contains the fading afterglow. A third epoch is obtained $>1$ year after the
+	GRB in which emission lines from \hb~and \ha~are found at the redshift of the
+	host, confirming the association of the absorption line system and the host.
+
+	\subsection{GRB120211A (z =	2.346)}
+	The data presented here has been published in \citet{Kruhler2015}. Two
+	observations of the host of GRB120211A has been obtained, starting 2013.02.17,
+	$> 1 year$ after the burst has faded. A redshift for this object has been
+	reported by \citet{Kruhler2015} and the features seen by those authors are
+	reproduced in these reductions, confirming $z =	2.346$.
+
 	\subsection{GRB120327A (z = 2.813)}

### 2016-10-14 document/XSGRB_sample.tex
Notes on individual bursts

@@
 
+	In cases where multiple traces are visible in the slit, additional components
+	for the profile are used in the optimal extraction. The additional components
+	share the PSF parameters and in cases where the additional component is an
+	extended object, the fits have been inspected to ensure that the additional
+	component does now skew the fit towards a broader PSF. The additional
+	components are not used for the weights.
+
 	Additionally, the wavelength solution for all bursts have been re-calibrated.
@@
 
-%	\begin{figure}
-%		%\vspace{11pc}   
-%		%\includegraphics[width=10cm]{060218/6273fig1.ps}
-%		\centerline{\includegraphics[width=8cm]{Resolution/resolution_paper.eps}}
-%		\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
+	\begin{figure}
+		\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+		\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
 The lower horizontal axis is in units pixels, the top axis in arc seconds.
 The red datapoints show a subsample of NIR spectra.
-%			The grey line shows a linear fit to the VIS datapoints. }
-%		\label{fig:res}
-%	\end{figure}
+			The grey line shows a linear fit to the VIS datapoints. }
+		\label{fig:res}
+	\end{figure}
 
@@
 
+	\subsection{GRB120224A (z = 1.10 NEW!!!)}
+	The data presented here has formed the basis of GCN \#
+	12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, and has also
+	been published in \citet{Kruhler2015}. Starting 19.8 hours after the GRB
+	trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting at
+	$\sim$ 7000 \AA~and extending all the way through 25000 \AA. We detect a $\sim
+	2 \sigma$ emission line which, if interpreted as \ha, gives $z = 1.10$,
+	supporting the redshift reported by \citet{Kruhler2015}.
+
+	\subsection{GRB120311 (z = 0.350 NEW!!!)}
+	The data presented here has formed the basis of GCN \#
+	12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, but is not
+	published otherwise. Starting just before twilight, 3.65 hours after the burst,
+	a faint afterglow continuum is detected at all wavelengths. Due to the
+	faintness of the afterglow, no absorption features are discernible superposed
+	on the continuum. Displaced from the afterglow continuum by 1\farc4, emission
+	lines belonging to \hb, \oiii~and \ha~are detected at $z = 0.350$. The line
+	belonging to \ha~shows some extended emission toward the afterglow continuum.
+	The angular distance between the two sources correspond to a projected distance
+	in the host plane of 6 kpc, posing a potential problem for the host redshift,
+	unless the GRB ocurred in a merging system. The extended emission in \ha,
+	supports this interpretation. This burst is not apart of the statistical sample.
+
 	\subsection{GRB120327A (z = 2.813)}
@@
 	$z = 2.813$.
+
+	\subsection{GRB120404A (z = 2.876)}
+	The data presented here has formed the basis of GCN \#
+	13227\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13227.gcn3}}, but is not published otherwise.
+9.6 ks integration, starting 15.7 hours after the \textit{Swift}-trigger reveals a low-intensity afterglow continuum on which absorption from \lya is detected in two distinct regions at redshifts $z=2.876$ and $z=.255$.
+These absorption systems are confirmed by ionic absorption features at both of these redshifts.
+
+	\subsection{GRB120422A (z = 0.283)}
+	The data presented here also formed the basis of GCN \#
+	13257\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13257.gcn3}} and is published in \citet{Schulze2014}.
+A GRB-SN, this burst has been followed up multiple times.
+The data presented here only contain the first epoch in which the afterglow is still visible and before the rise of SN2012bz.
+Starting 16.5 hours after the burst, 4.8 ks integration time captures both the host and the burst in emission.
+A blue afterglow continuum is detected at all wavelengths covered by X-shooter, on which \mgii absorption at $z = 0.283$ is found.
+Offset by 1\farc75, the host is clearly detected at a consistent redshift with a rich emission line spectrum, the lines extending towards to burst.
+
+	\subsection{GRB120712A (z = 4.175)}
+	The data presented here also formed the basis of GCN \#
+	13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
+	published elsewhere.
+4.8 ks integration time, starting 10.5 hours after the BAT trigger, shows a bright afterglow continuum starting at $\sim$ 4720 \AA, signifying the onset of the Lyman alpha forest, for a GRB located at $z = 4.175$.
+Absorption features from \lya, \feii, \mgii~and \SIii~are readily detected at a consistent redshift.
 

### 2016-10-15 document/XSGRB_sample.tex
More notes on individual bursts

@@
 \newcommand{\neiii}{[\ion{Ne}{3}]} 
+\newcommand{\NIii}{[\ion{Ni}{2}]} 
 \newcommand{\feii}{\ion{Fe}{2}} 
@@
 \newcommand{\SIii}{\ion{Si}{2}} 
+\newcommand{\SIiv}{\ion{Si}{4}} 
 \newcommand{\znii}{\ion{Zn}{2}} 
@@
 \newcommand{\caii}{\ion{Ca}{2}}
+
+\usepackage{todonotes}
 
@@
 
+	A significant fraction of the bursts presented here \todo[inline]{Insert exact number} have already had their hosts investigated in \citet{Kruhler2015}, for which extractions of the hosts exist.
+The focus of the data presented here are on the afterglows themselves and in the absence of a clear afterglow, the host.
+We will not, however, investigate the hosts.
+
 	\cleardoublepage
@@
 	\appendix
+
 	\section{Notes on Individual objects}
@@
 	The data presented here has formed the basis of GCN \#
-	13227\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13227.gcn3}}, but is not published otherwise.
-9.6 ks integration, starting 15.7 hours after the \textit{Swift}-trigger reveals a low-intensity afterglow continuum on which absorption from \lya is detected in two distinct regions at redshifts $z=2.876$ and $z=.255$.
-These absorption systems are confirmed by ionic absorption features at both of these redshifts.
+	13227\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13227.gcn3}}, but is not
+	published otherwise. 9.6 ks integration, starting 15.7 hours after the
+	\textit{Swift}-trigger reveals a low-intensity afterglow continuum on which
+	absorption from \lya is detected in two distinct regions at redshifts $z=2.876$
+	and $z=.255$. These absorption systems are confirmed by ionic absorption
+	features at both of these redshifts.
 
@@
 	The data presented here also formed the basis of GCN \#
-	13257\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13257.gcn3}} and is published in \citet{Schulze2014}.
-A GRB-SN, this burst has been followed up multiple times.
-The data presented here only contain the first epoch in which the afterglow is still visible and before the rise of SN2012bz.
-Starting 16.5 hours after the burst, 4.8 ks integration time captures both the host and the burst in emission.
-A blue afterglow continuum is detected at all wavelengths covered by X-shooter, on which \mgii absorption at $z = 0.283$ is found.
-Offset by 1\farc75, the host is clearly detected at a consistent redshift with a rich emission line spectrum, the lines extending towards to burst.
+	13257\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13257.gcn3}} and is published
+	in \citet{Schulze2014}. A GRB-SN, this burst has been followed up multiple
+	times. The data presented here only contain the first epoch in which the
+	afterglow is still visible and before the rise of SN2012bz. Starting 16.5 hours
+	after the burst, 4.8 ks integration time captures both the host and the burst
+	in emission. A blue afterglow continuum is detected at all wavelengths covered
+	by X-shooter, on which \mgii absorption at $z = 0.283$ is found. Offset by
+	1\farc75, the host is clearly detected at a consistent redshift with a rich
+	emission line spectrum, the lines extending towards to burst.
 
@@
 	13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
-	published elsewhere.
-4.8 ks integration time, starting 10.5 hours after the BAT trigger, shows a bright afterglow continuum starting at $\sim$ 4720 \AA, signifying the onset of the Lyman alpha forest, for a GRB located at $z = 4.175$.
-Absorption features from \lya, \feii, \mgii~and \SIii~are readily detected at a consistent redshift.
+	published elsewhere. 4.8 ks integration time, starting 10.5 hours after the BAT
+	trigger, shows a bright afterglow continuum starting at $\sim$ 4720 \AA,
+	signifying the onset of the Lyman alpha forest, for a GRB located at $z =
+	4.175$. Absorption features from \lya, \feii, \mgii~and \SIii~are readily
+	detected at a consistent redshift.
+
+	\subsection{GRB120714B (z = 0.398)}
+	The data presented here also formed the basis of GCN \#
+	13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
+	published elsewhere. Observations of this burst started 7.8 hours after the GRB
+	trigger, lasting 4.8 ks. A continuum is visible across the entire spectral
+	coverage of X-shooter, with both emission lines from  \oii, \hb, \oiii~and \ha,
+	as well as absorption from \mgii~detected at $z = 0.398$, securely setting it
+	as the redshift of the GRB.
+
+	\subsection{GRB120716A (z = 2.486)}
+	The data presented here also formed the basis of GCN \#
+	13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
+	published elsewhere. Despite observations starting 62 hours after the
+	\textit{Swift} trigger and lasting 3.6 ks, a bright afterglow is clearly seen,
+	along with a plethora of absorption features. Absorption of \lya-photons in the
+	host leaves a broad trough, from which the Lyman alpha forest is visible
+	bluewards, all the way down to the Lyman limit. Metal absorption lines from
+	\cii, \SIii, \oi, \feii, \civ, \SIiv, including fine structure transitions
+	identified as \cii$^*$, \SIii$^*$, \feii$^*$ and metastable \NIii~lines are all
+	detected at $z = 2.486$
+
+	\subsection{GRB120722A (z = 0.959)}
+	The data presented here also formed the basis of GCN \#
+	13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
+	published elsewhere. On 4.8 ks integration time, starting 10 hours after the
+	burst trigger, the simultaneous detection of absorption features belonging to
+	\mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
+	\hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the
+	redshift of the GRB.
 
@@
 	lasted for 2.4ks. The afterglow continuum is detected across the entire
-	spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii and
+	spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii~and
 	\alii~are found at $z = 0.736$. At the same redshift, emission lines from \oii,
-	\oiii, \ha, \hb, \hg, \nii, \sii, \siii are found. A second epoch, lasting
+	\oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch, lasting
 	3.6ks, is obtained after the afterglow has faded, confirming the emission line

### 2016-11-01 document/XSGRB_sample.tex
GRB120805A GRB120815A GRB120909A GRB121024A GRB121027A GRB161007A GRB161014A

@@
 \newcommand{\lyb}{Ly$\beta$} 
+\newcommand{\lyg}{Ly$\gamma$} 
 \newcommand{\hb}{H$\beta$} 
@@
 \newcommand{\siii}{[\ion{S}{3}]} 
-\newcommand{\oii}{[\ion{O}{2}]$\lambda$3727} 
-\newcommand{\oiii}{[\ion{O}{3}]$\lambda$5007}
+\newcommand{\oii}{[\ion{O}{2}]} 
+\newcommand{\oiii}{[\ion{O}{3}]}
 \newcommand{\nii}{[\ion{N}{2}]} 
+\newcommand{\nv}{[\ion{N}{5}]} 
 \newcommand{\neiii}{[\ion{Ne}{3}]} 
@@
 
+	\subsection{GRB120805A (z $\sim$ 3.9 NEW!!!)}
+	A separate reduction of this burst has been published in \citet{Kruhler2015},
+	but not otherwise. Starting 9 days after the burst trigger, this is host
+	observation and does not contain any afterglow continuum. In 3.6 ks integration
+	time, we detect a faint continuum visible from 4500 \AA~and all the way through
+	21000 \AA, in contrast to what is found previously. The continuum from 4500 -
+	6000 \AA is detected at very low significance. If the drop at 4500 \AA~is the
+	Lyman limit, this fits with Lyman alpha at $\sim$ 6000 \AA, giving $z \sim
+	3.9$. The absence of nebular lines if due to \oii~falling in a telluric
+	absorption band and the rest being shifted out of the wavelength coverage.
+
+	\subsection{GRB120815A* (z = 2.358)} 
+	Not a part of the statistical sample, thisburst also formed the basis of  GCN
+	\# 13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
+	published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
+	trigger and consist of 2.4 ks integration. A bright afterglow continuum is
+	detected across the entire spectral coverage of X-shooter, with a multitude of
+	absorption lines superposed. Absorption features from the host at $z = 2.358$
+	include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
+	\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
+	fine-structure lines from \NIii and \feii are exited local to the GRB.
+	Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
+
 	\subsection{GRB120722A (z = 0.959)}
@@
 	redshift of the GRB.
+
+	\subsection{GRB120909A (z = 3.929)}
+	The data presented here has formed the basis of GCN \#
+	13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
+	published otherwise. A very rapid follow-up, starting only 1.7 hours after the
+	BAT trigger, this 1.2 ks observation captures a very bright afterglow
+	continuum, starting at 4500 \AA, signifying the onset of the Lyman limit for a 
+	system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
+	prominent absorption features in the form of \lya, \lyb, and \lyg, visible in
+	the Lyman alpha forest. Metal absorption lines arising from \feii, \NIii,
+	\SIii, \sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along
+	with the corresponding fine-structure lines from (\feii$^*$, \SIii$^*$,
+	\oi$^*$, \oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
+
+	\subsection{GRB121024A (z = 2.300)}
+	The data presented here also formed the basis of GCN \#
+	13890\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13890.gcn3}} and is published
+	in \citet{Friis2015}. Also rapid, starting 1.8 hours after the \textit{Swift}
+	trigger, a bright afterglow continuum is visible across all arms. A broad
+	absorption feature from Lyman alpha, along with narrow lines from \civ, \SIii,
+	\SIiv, \feii, \sii, and \alii, as well as fine-structure lines associated with
+	\SIii$^*$ are all detected at $z = 2.300$, securely setting it as the redshift
+	of the GRB.
+
+	\subsection{GRB121027A (z = 1.773)}		
+	The data presented here has formed the basis of GCN \#
+	13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
+	published otherwise. Starting 69.6 hours after the GRB trigger, that we detect
+	the afterglow continuum a so high significance in all arms with 8.4 ks
+	integration, testifies to the brightness of this burst. The concurrent
+	identification of emission lines from \oiii~and absorption from \civ, \alii,
+	\aliii, \mgi, \mgii, and \feii, tightly constrains the redshift of the burst to
+	be $(z = 1.773$
 
@@
 	The data presented here also formed the basis of GCN \#
-	19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}} and is not
+	19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, but is not
 	published elsewhere. Observations started 22.37 hours after the BAT trigger and
@@
 
+	\subsection{GRB161007A (z =4.6 ??? NEW!!!)}
+	This data has not been published elsewhere. Observations for GRB161007A started
+	323 hours after the burst trigger and contains the potential host. 4 x 600 seconds
+	of observations reveals a faint continuum rising abruptly above the noise at
+	$\sim$ 6850 \AA~and continuing through 21000 \AA. A very low significance
+	continuum is detected at shorter wavelengths, down to $\sim$ 6000 \AA. If the
+	host is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha
+	break and the absence of nebular emission lines is due to \oii being shifted
+	out of the wavelength coverage. Alternatively, an early-type host at $z = 0.71$
+	could exhibit the 4000 \AA break at 6000 \AA, but due to the preference of
+	long-duration GRBs for star-forming galaxies, this is the least likely
+	explanation, why we believe the high-z solution.
+
+	\subsection{GRB161014A (z =2.823)}
+	The data presented here also formed the basis of GCN \#
+	20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}}, but is not
+	published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8 ks of
+	integration time captures the afterglow continuum across all three
+	spectroscopic arms. A broad absorption trough due to Lyman alpha is visible,
+	along with metal absorption features from \mgii, \SIii, \cii, \civ, \alii,
+	\aliii, and	\feii, all at $z =2.823$. Similar to GRB140506 \citep{Fynbo2014}, a
+	break in the continuum shape is detected bluewards of 6000\AA, possible
+	signifying some anomalous form of extinction.
+
 \end{document}

### 2016-11-09 document/XSGRB_sample.tex
Skeleton

@@
 		D.~Malesani\altaffilmark{1}, 
+		T.~Kr\"{u}hler\altaffilmark{1}, 
 		P.~Goldoni\altaffilmark{14}, 
-		T.~Kr\"{u}hler\altaffilmark{1}, 
 		J.~P.~U. Fynbo\altaffilmark{1}, 
@@
 
+	We simultaneously want to minimize any biases against astrophysical conditions while at the same time maximizing likelihood of observations.
+By restricting the selection criteria to conditions local to Galaxy.
+
 	\cleardoublepage
@@
 
-	\section{Results}
-	\input{tables/burst_overview.tex}	
 	\subsection{Spectral resolution}
@@
 
+	\section{Results}
+	\input{tables/burst_overview.tex}
+
+	\subsection{Complete sample}
+
+	\subsection{Follow-up timing and afterglow brightness}
+	This sections should contain a plot og the follow-up delay and the acquisition image brightness at the time of observations.
+
+	\subsection{Properties of rejected triggers}
+	How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
+
+	\subsection{On the nature of triggered GRBs without detected optical afterglow.}
+	This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?
+
 	\subsection{Redshifts}
-%	\subsection{Notes on Individual objects}
+
+	\subsubsection{Afterglow-based redshifts}
+
+	\subsubsection{Emissionline-based redshifts}	
+	Redshift completeness
+
+	\subsection{Hydrogen column densities}
+
+	\subsubsection{Comparison to X-ray determined columns}
 
 	\section{Discussion}
+
+	\subsection{Comparison to previous samples.}	
+	This should also include a discussion of dark bursts. 
+	Cuchiara. BAT6 (Salvaterra), TOUGH
+
+	\subsection{Comparison to previous samples.}
+
+	\subsection{X-ray column densities vs. optically derived ones}
+
+	\subsection{Why is the HI column truncated?}
+
+	What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
+
+	\section{Conclusion}
 

### 2016-11-09 py/Buchner_NH.py

@@ -1,0 +1,27 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+
+def main():
+    grbcat = fits.open("../data/grbcat-annotated.fits")
+    print(grbcat[1].header)
+    names = grbcat[1].data.field("Name")
+    t90 = grbcat[1].data.field("duration")
+    nH_gal = grbcat[1].data.field("nhgal")
+    z = grbcat[1].data.field("z")
+    nH = grbcat[1].data.field("NH_sphere_mean")
+    nH_l, hH_h = grbcat[1].data.field("NH_sphere_hi"), grbcat[1].data.field("NH_sphere_lo")
+    print(names, nH)
+    # print(hH_h)
+    # pl.errorbar(z, nH, yerr=[nH - nH_l, hH_h - nH], fmt=".k", capsize=0, elinewidth=0.5, ms=3)
+    # pl.show()
+
+
+
+if __name__ == '__main__':
+    main()

### 2016-11-14 py/Buchner_NH.py
Adding technical description

@@ -1,27 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-from astropy.io import fits
-import matplotlib.pyplot as pl
-import seaborn; seaborn.set_style('ticks')
-
-def main():
-    grbcat = fits.open("../data/grbcat-annotated.fits")
-    print(grbcat[1].header)
-    names = grbcat[1].data.field("Name")
-    t90 = grbcat[1].data.field("duration")
-    nH_gal = grbcat[1].data.field("nhgal")
-    z = grbcat[1].data.field("z")
-    nH = grbcat[1].data.field("NH_sphere_mean")
-    nH_l, hH_h = grbcat[1].data.field("NH_sphere_hi"), grbcat[1].data.field("NH_sphere_lo")
-    print(names, nH)
-    # print(hH_h)
-    # pl.errorbar(z, nH, yerr=[nH - nH_l, hH_h - nH], fmt=".k", capsize=0, elinewidth=0.5, ms=3)
-    # pl.show()
-
-
-
-if __name__ == '__main__':
-    main()

### 2016-11-14 document/XSGRB_sample.tex
resolving conflict in pull #124

@@
 
-\usepackage{todonotes}
+\usepackage{xcolor}
+\newcommand\todo[1]{\textcolor{red}{(#1)}}
 
@@
 
-	\cleardoublepage
-
-	\section{Sample selection criteria}
+\subsection{Sample selection criteria and observations}
+
+\subsection{Sample selection criteria} \label{samplecrit}
 
@@
 
-	A significant fraction of the bursts presented here \todo[inline]{Insert exact number} have already had their hosts investigated in \citet{Kruhler2015}, for which extractions of the hosts exist.
-The focus of the data presented here are on the afterglows themselves and in the absence of a clear afterglow, the host.
+A significant fraction of the bursts presented here \todo{Insert exact
+number} have already had their hosts investigated in \citet{Kruhler2015}, for
+which extractions of the hosts exist. The focus of the data presented here are
+on the afterglows themselves and in the absence of a clear afterglow, the host.
 We will not, however, investigate the hosts.
 
-	We simultaneously want to minimize any biases against astrophysical conditions while at the same time maximizing likelihood of observations.
-By restricting the selection criteria to conditions local to Galaxy.
-
-	\cleardoublepage
-
-	\section{Observations}
-
-	\subsection{RRM observations} \label{RRM}
+We simultaneously want to minimize any biases against astrophysical conditions
+while at the same time maximizing likelihood of observations. By restricting
+the selection criteria to conditions local to Galaxy.
+
+\subsection{Follow-up procedure}
+
+\subsubsection{RRM observations} \label{RRM}
 The rapid-response mode is
 
-	\section{Reduction scheme}
+\subsection{Observations}
+All the observations in this sample has been carried out using the
+cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
+two of the four Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal)
+during the duration of this follow-up campaign, which covers the entire lifetime
+of X-shooter. Observations have been carried out from the ESO period P84
+through P98 under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
+095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
+089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.
+
+The first GRB followed up is GRB090313, observed 15th of March 2009, which was
+while X-shooter was mounted on UT3 during the commissioning of the instrument.
+The bursts observed during the commisioning period is not a part of our
+statistical sample, but are nonetheless published as part of the bursts observed
+by the X-shooter GRB team. The first burst observed after science verification
+was completed when X-shooter was moved to UT2 is GRB091018 which thereby
+constitute the first burst entering our statistical sample. For all bursts that
+fulfill our sample selection criteria, described in Sect. \ref{samplecrit},
+spectroscopic follow-up has been attempted with X-shooter. Various conditions
+can affect our ability to follow up a given burst, and a discussion of this
+effect is included in Sect. \ref{badbursts}.
+
+X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800 \AA~in
+a single exposure, by splitting the light in three separate spectroscopic arms
+using dichroics.
+This way each arm work as a separate instrument, each functioning as its own
+echelle spectrograph.
+The ultraviolet blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm
+covering 5500 - 10200 \AA, and the near-infrared()NIR) arm covering 10200 -
+24800 \AA~ with the possibility of applying a k-band
+blocking filter cutting the coverage of the NIR arm at 10200 - 21000 \AA.
+
+\todo{ABBA}
+
+For the large majority of the bursts we have observed with a slit width of
+1\farc0, 0\farc9, and 0\farc9 for the UVB, VIS, and NIR-arm respectively, which
+puts a lower limit on the delivered resolution of the spectra based on the
+tabulated values of the delivered resolutions, which is 4350, 7450, and 5300 for
+the UVB, VIS and NIR-arm respectively\footnote{\url{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}}.
+The slit width also sets the width of the atmospheric sky lines and determines
+the amount of light lost due to the wavelength-dependent seeing PSF extending
+outside the coverage of the slit, where the width of sky-lines is always set by
+the slit width whereas both the delivered resolution and the slitloss changes
+for the better as the seeing PSF drops below the slit width.
+For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9
+observed with a 0\farc9 slit only 76.1 percent of the light will enter the slit,
+meaning that for almost all observations a slitloss correction is required. We
+describe how slitlosses were corrected for in Sect. \ref{postproc}.
+For accurate measurements of velocity widths, a precise instrumental resolution
+is required and this becomes better when the delivered seeing is better than the
+slit width. We discuss this in Sect. \ref{resolution}.
+
+\todo{ADC discussion}
+
+\section{Data processing}
+
+In this section we describe how the final data products are produced.
+
+Before any reductions are done, the raw object images are run through the
+cosmic-ray removal algorithm \citep{VanDokkum2001} implementation,
+\textit{Astro-SCRAPPY} \footnote{\url{https://github.com/astropy/astroscrappy}},
+where a wide clipping radius have been used around detected cosmics to ensure that edge residuals are
+robustly rejected.
+
+The basis for the reductions is the VLT/X-shooter pipeline, version
+\texttt{2.7.1} or newer \citep{Goldoni2006, Modigliani2010}. The pipeline is
+managed with the Reflex interface \citep{Freudling2013} and is used for
+subtraction of bias level, flat-fielding, tracing of the echelle orders,
+wavelength calibrations with the use of arc-line lamps, flux calibration using
+spectrophotometric standards \citep{Vernet2010, Hamuy1994}, mirror flexure
+compensation, initial sky-subtraction and lastly the rectification and merging
+of the orders. For the initial sky-subtraction, the background has been
+estimated in regions adjacent to the object trace clear of contaminating
+sources. Because of the broken ADC, for some objects there is a lot of curvature
+in the object trace. This means that for some bursts, the initial sky-estimate
+has been made from a limited number of pixels in the spatial direction. By doing
+an initial subtracting the sky on the un-rectified image we ensure that bulk of
+the sky background is not redistributed by the rectification process.
+
+The image is rectified onto an equidistant grid with a dispersion sampling of
+0.2 \AA/pixel and a 0.16 "/pixel spatial sampling for the UVB and VIS arm and
+0.6 \AA/pixel with a 0.21 "/pixel in the NIR arm.  Because the tabulated
+resolution is a lower limit, by choosing a sampling of 0.2 \AA/pixel, we ensure
+that the bluest part of neither of the arms have a sampling lower than the
+Nyquist sampling rate of 2 pixels per resolution FWHM.
+
+\subsection{Post-processing} \label{postproc}
+
+For a typical observation, each of the exposures in the nodding sequence has
+been reduced as single observation and then subsequently combined to form a
+single image. Because this strategy is employed, we can reject outliers in the
+stack and weight by the average inverse variance of the background. When
+weighting images where the noise in each pixel is dominated by Poisson noise it
+important to estimate the background variance in a large enough region, so that
+to remove the correlation between the signal and the weights. To this end, the
+weight map is generated by a running median window over the variance map, where
+the trace as been masked and width of the window is chosen to be wide enough for
+median variance to be generated on the basis of several hundred pixels. This
+weighting scheme automatically also optimally combines images of different
+exposure times or images where the background is varying, which is often the
+case when a burst has been observed close to twilight
+
+Because the background is very bright and there is a high abundance of broad
+sky-lines in the NIR arm, when there are no contaminating sources in the slit,
+the sky has been put back on the images and they have been reduced in pairs of
+two, subtracting the two from each other, keeping the WCS static. This amounts
+to the regular nodding reduction, only we can reject outliers and weight by the
+averaged inverse variance map.
+
+\subsection{Correction for offsets in the wavelength calibration}    \label{wavecal}
+
+Additionally, the wavelength solution for all bursts have been re-calibrated.
+Using a synthetic sky spectrum \citep{Noll2012, Jones2013}, a first refinement of the wavelength solution have been obtained by cross-correlating with the observed sky.
+Both a multiplicative and an additive offset has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
+
+X-shooter, being installed at the VLT Cassegrain focus is  prone to
+flexures during operations. The flexures modify the projection of the slit
+on the detector with respect to the one obtained in daytime calibration. 
+This require a modification of the wavelength solution in order to
+process correctly the night-time data. Part of this correction
+is performed by the pipeline using the frames taken
+during X-shooter Active Flexure Compensation procedure
+\footnote{X-shooter User Manual available at https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}
+We corrected the remaining part using as a reference the sky
+emission lines present in the observed data.
+
+For every afterglow observation we reduced one frame individually in STARE mode without
+sky subtraction obtaining $\sim$ 100 sky spectra. The sky line list compiled at ESO for
+the E-ELT
+study\footnote{http://www.eso.org/sci/facilities/eelt/science/drm/tech\_data/data/optical\_ir\_sky\_lines.dat}
+from the work of (Hanuschik, 2003, A\&A, 407, 1157) and (Rousselot et al. (2000, A\&A, 354, 1134), was used as a reference.
+From this list, we selected a subset of bright and isolated lines.
+In the case of the OH doublets, unresolved at X-shooter resolution, we took as line position the average between the blue and red components.
+To find the offsets of the spectra, we fitted gaussians near the
+expected positions under IDL using the MPFIT software (Markwardt, 2009, Astronomical Society of the Pacific Conference Series, Vol.
+411, ADASS XVII, ed.
+D.A.
+Bohlender, D.
+Durand, \& P.
+Dowler, 251) and we compared the result to the tabulated values.
+The resulting offsets, which were smaller than 0.1 \AA~in the UVB and VIS data
+and smaller than 0.5 \AA~in the NIR spectra, were applied to the corresponding spectra.
+
+\subsection{Spectral extraction}    \label{extract}
 
@@
 
-	Additionally, the wavelength solution for all bursts have been re-calibrated.
-Using a synthetic sky spectrum \citep{Noll2012, Jones2013}, a first refinement of the wavelength solution have been obtained by cross-correlating with the observed sky.
-
-	Both a multiplicative and an additive offset has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
-
-	\subsection{Spectral resolution}
+\subsection{Spectral resolution} \label{resolution}
 The afterglow spectra described in this paper are obtained in
@@
 
-	\subsection{Correction for offsets in the wavelength calibration}
-
-	X-shooter, being installed at the VLT Cassegrain focusi is  prone to
-	flexures during operations. The flexures modify the projection of the slit
-	on the detector with respect to the one obtained in daytime calibration. 
-	This require a modification of the wavelength solution in order to
-	process correctly the night-time data. Part of this correction
-	is performed by the pipeline using the frames taken
-	during X-shooter Active Flexure Compensation procedure
-	\footnote{X-shooter User Manual available at https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}
-	We corrected the remaining part using as a reference the sky
-	emission lines present in the observed data.
-
-	For every afterglow observation we reduced one frame individually in STARE mode without
-	sky subtraction obtaining $\sim$ 100 sky spectra. The sky line list compiled at ESO for
-	the E-ELT
-	study\footnote{http://www.eso.org/sci/facilities/eelt/science/drm/tech\_data/data/optical\_ir\_sky\_lines.dat}
-	from the work of (Hanuschik, 2003, A\&A, 407, 1157) and (Rousselot et al. (2000, A\&A, 354, 1134), was used as a reference.
-	From this list, we selected a subset of bright and isolated lines.
-In the case of the OH doublets, unresolved at X-shooter resolution, we took as line position the average between the blue and red components.
-To find the offsets of the spectra, we fitted gaussians near the
-	expected positions under IDL using the MPFIT software (Markwardt, 2009, Astronomical Society of the Pacific Conference Series, Vol.
-411, ADASS XVII, ed.
-D.A.
-Bohlender, D.
-Durand, \& P.
-Dowler, 251) and we compared the result to the tabulated values.
-	The resulting offsets, which were smaller than 0.1 \AA~in the UVB and VIS data
-	and smaller than 0.5 \AA~in the NIR spectra, were applied to the corresponding spectra.
-
 \section{Results}
@@
 
-	\subsection{Properties of rejected triggers}
+\subsection{Properties of rejected triggers} \label{badbursts}
 How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
@@
 \subsection{Hydrogen column densities}
-
-	\subsubsection{Comparison to X-ray determined columns}
 
@@
 
+\begin{figure}
+	\centerline{\includegraphics[width=10cm]{figures/OAXR_NH_sphere.pdf}}
+	\caption{Optically derived HI column densities vs.
+X-ray derived ones taken from \citet{Buchner2016}.
+The colors of the points reflect the redshift of the burst.
+Circles mark the column densities published in \citet{Fynbo2009} and squares mark the ones derived as part of the sample presented here.
+Overplot in blue is the 1:1 correspondence, expected if the conditions local to the GRBs are similar to the local ISM abundances.
+If the metallicity is lower at the GRB site, the point move upwards in the plot.
+}
+	\label{fig:OAXR}
+\end{figure}
+
 \subsection{Why is the HI column truncated?}
@@
 What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
+Number of bursts with detected H2 and their HI column density.
+Rahmati Schaye, Krongold Prochaska.
 

### 2016-11-14 document/XSGRB_sample.tex
Reduction stuff

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Sample selection criteria and observations}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Sample selection criteria} \label{samplecrit}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Being of transient nature, it is difficult to impose strong sample selection
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Follow-up procedure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{RRM observations} \label{RRM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The rapid-response mode is
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Observations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 All the observations in this sample has been carried out using the
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data processing}
-
-In this section we describe how the final data products are produced.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we describe how the final data products are produced and
+subsequently post-processed. All post-processing scripts developed for this
+dataset are made publicly available at
+\url{https://github.com/jselsing/XSGRB_reduction_scripts}, along with
+instructions of use.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Post-processing} \label{postproc}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+By STARE reducing all observations we additionally get a spectrum of the sky,
+which we can use to calibrate the wavelength solution in Sect. \ref{wavecal}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Correction for offsets in the wavelength calibration}    \label{wavecal}
-
-Additionally, the wavelength solution for all bursts have been re-calibrated.
-Using a synthetic sky spectrum \citep{Noll2012, Jones2013}, a first refinement of the wavelength solution have been obtained by cross-correlating with the observed sky.
-Both a multiplicative and an additive offset has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
-
-X-shooter, being installed at the VLT Cassegrain focus is  prone to
-flexures during operations. The flexures modify the projection of the slit
-on the detector with respect to the one obtained in daytime calibration. 
-This require a modification of the wavelength solution in order to
-process correctly the night-time data. Part of this correction
-is performed by the pipeline using the frames taken
-during X-shooter Active Flexure Compensation procedure
-\footnote{X-shooter User Manual available at https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}
-We corrected the remaining part using as a reference the sky
-emission lines present in the observed data.
-
-For every afterglow observation we reduced one frame individually in STARE mode without
-sky subtraction obtaining $\sim$ 100 sky spectra. The sky line list compiled at ESO for
-the E-ELT
-study\footnote{http://www.eso.org/sci/facilities/eelt/science/drm/tech\_data/data/optical\_ir\_sky\_lines.dat}
-from the work of (Hanuschik, 2003, A\&A, 407, 1157) and (Rousselot et al. (2000, A\&A, 354, 1134), was used as a reference.
-From this list, we selected a subset of bright and isolated lines.
-In the case of the OH doublets, unresolved at X-shooter resolution, we took as line position the average between the blue and red components.
-To find the offsets of the spectra, we fitted gaussians near the
-expected positions under IDL using the MPFIT software (Markwardt, 2009, Astronomical Society of the Pacific Conference Series, Vol.
-411, ADASS XVII, ed.
-D.A.
-Bohlender, D.
-Durand, \& P.
-Dowler, 251) and we compared the result to the tabulated values.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+X-shooter, being installed at the VLT Cassegrain focus is  prone to flexures
+during operations. The flexures modify the projection of the slit on the
+detector with respect to the one obtained in daytime calibration. This requires a
+modification of the wavelength solution in order to process correctly the
+night-time data. Part of this correction is performed by the pipeline using the
+frames taken during X-shooter Active Flexure Compensation procedure
+\footnote{X-shooter User Manual available at
+	https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}.
+
+The remaining offset is corrected by cross-correlation with a synthetic sky
+spectrum \citep{Noll2012, Jones2013}. To get the correct seeing PSF with which
+to convolve the synthetic sky an initial refinement of the wavelength solution
+have been obtained by cross-correlating the observed sky with an unconvolved
+synthetic sky. This preliminary wavelength calibration is applied to the
+observed sky. The synthetic spectrum is then convolved with an increasing seeing
+PSF and the width that minimizes $\chi^2$ with the updated observed sky is
+chosen to be the effective sky-PSF. Using the synthetic sky with the matched
+resolution, a final wavelength calibration can then be calculated by
+cross-correlating the observed sky with the correctly broadened sky spectrum, as
+a function of a velocity offset. Both a multiplicative and an additive offset to
+the wavelength calibration has been tested, but in terms of $\chi^2$, the model
+with only a multiplicative offset is preferred.
+
 The resulting offsets, which were smaller than 0.1 \AA~in the UVB and VIS data
-and smaller than 0.5 \AA~in the NIR spectra, were applied to the corresponding spectra.
-
+and smaller than 0.5 \AA~in the NIR spectra, but changing over short period of
+time were applied to the corresponding spectra.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spectral extraction}    \label{extract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spectral resolution} \label{resolution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 The afterglow spectra described in this paper are obtained in
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{tables/burst_overview.tex}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Complete sample}	
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Follow-up timing and afterglow brightness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 This sections should contain a plot og the follow-up delay and the acquisition image brightness at the time of observations.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Properties of rejected triggers} \label{badbursts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{On the nature of triggered GRBs without detected optical afterglow.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Redshifts}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Afterglow-based redshifts}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Emissionline-based redshifts}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Redshift completeness
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Hydrogen column densities}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Comparison to previous samples.}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 This should also include a discussion of dark bursts. 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
 \subsection{Comparison to previous samples.}		
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{X-ray column densities vs. optically derived ones}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Why is the HI column truncated?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-12-02 document/XSGRB_sample.tex

@@
-\documentclass[iop, twocolappendix, numberedappendix, tighten, appendixfloats]{emulateapj}
-\citestyle{apj}
-
-\usepackage[stable]{footmisc}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{txfonts}
-%\usepackage{hyperref}
+%                                                                 aa.dem
+% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
+% demonstration file
+%                                                       (c) EDP Sciences
+%-----------------------------------------------------------------------
+%
+%\documentclass[referee]{aa} % for a referee version
+%\documentclass[onecolumn]{aa} % for a paper on 1 column  
+%\documentclass[longauth]{aa} % for the long lists of affiliations 
+%\documentclass[rnote]{aa} % for the research notes
+%\documentclass[letter]{aa} % for the letters 
+%\documentclass[bibyear]{aa} % if the references are not structured 
+% according to the author-year natbib style
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
+%\documentclass[referee]{aa}
+
+\usepackage{graphicx,url,twoopt,natbib}
+\usepackage[varg]{txfonts}           %% A&A font choice
+
+\usepackage{pdfcomment}              %% for popup acronym meanings
+\usepackage{acronym}                 %% for popup acronym meanings
+
+\usepackage{url}
+\usepackage{color,hyperref}
+\definecolor{linkcolor}{rgb}{0,0,0.7}
+\hypersetup{colorlinks=true,
+	linkcolor=linkcolor, 
+	citecolor=linkcolor,
+	filecolor=linkcolor, 
+	urlcolor=linkcolor}
+
 \usepackage{natbib}
-\bibpunct[, ]{(}{)}{;}{a}{}{,}
-%\bibliographystyle{./bibtex/apj}
+
+\bibliographystyle{aa}
+\bibpunct{(}{)}{;}{a}{}{,}    %% natbib cite format used by A&A and ApJ
 
@@
 \title{\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra}
-\altaffiltext{$^{\dag}$}{Based on observations collected at the European Southern Observatory, Paranal, 
-	Chile, Program ID: 084.A-0260, 085.A-009, 086.A-0073, 087.A-0055.}
+\thanks{Based on observations collected at the European Southern Observatory, Paranal, 
+	Chile, Program ID: 098.A-0055, 097.A-0036, 096.A-0079, 095.B-0811(B), 
+	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
+	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}
+
+\titlerunning{XSGRB sample}
 
 \author{
-	J.~Selsing\altaffilmark{1}, 
-	D.~Malesani\altaffilmark{1}, 
-	T.~Kr\"{u}hler\altaffilmark{1}, 
-	P.~Goldoni\altaffilmark{14}, 
-	J.~P.~U. Fynbo\altaffilmark{1}, 
-	A.~de~Ugarte~Postigo\altaffilmark{11}, 
-	J.~Japelj\altaffilmark{20},
+	J.~Selsing\inst{1}, 
+	D.~Malesani\inst{1}, 
+	T.~Kr\"{u}hler\inst{1}, 
+	P.~Goldoni\inst{14}, 
+	J.~P.~U. Fynbo\inst{1}, 
+	A.~de~Ugarte~Postigo\inst{11}, 
+	J.~Japelj\inst{20},
 	P.~D'Avanzo,
 	Z.~Cano,
-	S.~Covino\altaffilmark{10}, 
-	V.~D'Elia\altaffilmark{7, 12}, 
+	S.~Covino\inst{10}, 
+	V.~D'Elia\inst{7, 12}, 
 	H.~Flores,
-	O.~E.~Hartoog\altaffilmark{6},
-	J.~Hjorth\altaffilmark{1}, 
-	P.~Jakobsson\altaffilmark{5}, 
+	O.~E.~Hartoog\inst{6},
+	J.~Hjorth\inst{1}, 
+	P.~Jakobsson\inst{5}, 
 	A.~Levan,
 	A.~Melandri,
-	S.~Piranomonte \altaffilmark{7},
-	R.~S\'anchez-Ram\'irez\altaffilmark{11},
-	S.~Schulze\altaffilmark{17, 18}, 
-	N.~R.~Tanvir\altaffilmark{19},
+	S.~Piranomonte \inst{7},
+	R.~S\'anchez-Ram\'irez\inst{11},
+	S.~Schulze\inst{17, 18}, 
+	N.~R.~Tanvir\inst{19},
 	C.~Th{\"o}ne,
-	S.~D.~Vergani\altaffilmark{7, 8},
-	P.~M.~Vreeswijk\altaffilmark{3}, 
-	D.~J.~Watson\altaffilmark{1},
-	K.~Wiersema\altaffilmark{19},
-	D.~Xu\altaffilmark{1}
-	L.~Christensen\altaffilmark{1},
-	A.~De~Cia\altaffilmark{3}, 
-	L.~Kaper\altaffilmark{6}, 
+	S.~D.~Vergani\inst{7, 8},
+	P.~M.~Vreeswijk\inst{3}, 
+	D.~J.~Watson\inst{1},
+	K.~Wiersema\inst{19},
+	D.~Xu\inst{1}
+	L.~Christensen\inst{1},
+	A.~De~Cia\inst{3}, 
+	L.~Kaper\inst{6}, 
 	L.~A.~Antonelli,
@@
 	F.~Hammer,
-	C.~Ledoux\altaffilmark{2}, 
+	C.~Ledoux\inst{2}, 
 	E.~Maiorano,
-	B.~Milvang-Jensen\altaffilmark{1}, 
+	B.~Milvang-Jensen\inst{1}, 
 	E.~Palazzi,
@@
 	J.~Schaye,
-	G.~Tagliaferri\altaffilmark{7},
-	R.~A.~M.~J.~Wijers\altaffilmark{6}
+	G.~Tagliaferri\inst{7},
+	R.~A.~M.~J.~Wijers\inst{6}
 }
 
-\altaffiltext{1}{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark}
-\altaffiltext{2}{European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile}
-\altaffiltext{3}{Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel}
-\altaffiltext{4}{Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany}
-\altaffiltext{5}{Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland}
-\altaffiltext{6}{Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands}
-\altaffiltext{7}{INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy}
-\altaffiltext{8}{GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France}
-\altaffiltext{9}{American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA}
-\altaffiltext{10}{INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy}
-\altaffiltext{11}{Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain}
-\altaffiltext{12}{ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy}
-\altaffiltext{13}{Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic}
-\altaffiltext{14}{APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/
+\authorrunning{Selsing et al.}
+
+\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
+\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
+\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
+\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
+\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
+\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
+\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
+\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
+\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
+\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
+\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
+\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
+\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/
 	IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et L
-	eonie Duquet, 75205 Paris Cedex 13, France}
-\altaffiltext{15}{Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachs
-	tra\ss e, 85748 Garching, Germany}
-\altaffiltext{16}{Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3,
-	20126, Milano, Italy}
-\altaffiltext{17}{Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astro
-	nom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile}
-\altaffiltext{18}{Millennium Center for Supernova Science}
-\altaffiltext{19}{Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK}
-\altaffiltext{20}{University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI}\
-
-\begin{abstract}
+	eonie Duquet, 75205 Paris Cedex 13, France
+\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachs
+	tra\ss e, 85748 Garching, Germany
+\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3,
+	20126, Milano, Italy
+\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astro
+	nom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
+\and Millennium Center for Supernova Science
+\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
+\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI}
+
+\date{Received/ accepted}
+
+\abstract{
 The {\it Swift} satellite allows us to use gamma-ray bursts (GRBs) to peer into 
@@
 sample is still limited by low-number statistics. 
-\end{abstract}
+}
 
 \keywords{Gamma-ray burst: individual: GRB~120815A --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
+
+\maketitle
 
@@
 	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 min at least 30° above horizon, with the Sun below -12°.
+	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
 	\item No bright closeby stars.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Observations}
+\subsection{Observations} \label{obs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The remaining offset is corrected by cross-correlation with a synthetic sky
-spectrum \citep{Noll2012, Jones2013}. To get the correct seeing PSF with which
+spectrum \citep{Noll2012, Jones2013} after the continuum, estimated as the mode
+of all flux values, has been subtrated. To get the correct seeing PSF with which
 to convolve the synthetic sky an initial refinement of the wavelength solution
@@
 
+Using the convolved synthetic sky, the $\gtrsim 3 \sigma$ sky brightness pixels
+have been added to the bad pixel map to avoid the cores of the brightest
+sky-lines.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To extract the afterglow spectrum from the rectified 2D-image, several
+techniques have been employed based on the brightness of the afterglow and the
+complexity of the objects entering the slit. Due to the malfunctioning ADC, see
+Sect. \ref{obs}, the spectral trace moves across the slit as a function of
+wavelength for a large fraction of the bursts observed meaning that using a
+single aperture for the spectral extraction is inadequate due to the large
+amount of background that would then enter the slit. To optimally select the
+extraction regions we therefore need to model the trace position.
+
+To get the shape and the position of the spectral PSF as a function of location
+on the image, we need to chose a model which can represent how the light falls
+on the slit. We know from \citet{Trujillo2001} that the Moffat function
+\citep{Moffat1969} adequately describes an imaging PSF due to atmospheric
+turbulence, but due to the aberrations in the dispersion elements and the
+rectification process, the PSF we are trying to model different from this
+profile. To allow for flexibility in the model, we have chosen the Voigt
+function as a model for the spectral PSF and we describe how this is evaluated
+in App. \ref{voigt}. Since additionally, the host galaxy could also have a
+contribution the image profile, this choice allows for the required freedom if
+additional flux is in the wings of the profile.
+
+To guide the guess position of the trace on the slit as a function of
+wavelength, we have used the analytic prescription for the trace position
+described in \citet{Filippenko1982}, where the header keywords of the
+observations have been queried for the ambient conditions which controls the
+degree to which the trace moves.
+
+Based on the signal-to-noise of the afterglow continuum, the 2D-image has been
+binned down in the spectral direction to a number of elements that allows for an
+accurate tracing of the PSF, typically 200 bins for moderate signal-to-noise.
+For each of the bins and using the guess position, the spectral PSF has been fit
+using the unweighted chi-squared minimization algorithm implemented in
+\texttt{scipy.optimize.curve\_fit} \citep{scipy}. Since we know that the trace
+varies slowly as a function of wavelength, we have then fitted a low-order
+polynomial to the fit parameters as a function of wavelength, which allows us to
+evaluate the spectral PSF at all wavelengths and this way accurately model the
+entire spectral PSF.
+
+Equipped with a model for how the light falls on the entire image, we can then
+employ the optimal extraction algorithm \cite{Horne1986}, which weights the
+extraction aperture by the spectral profile, or alternatively sum all pixels
+within 1 FWHM of the modeled profile. Where possible, we have used the optimal
+extraction. In cases where the trace is very weak, even in the binned images, an
+aperture has been selected manually which covers emission lines, if present, and
+when nothing is immediately visible, the entire nodding window. The error- and
+bad pixel maps are in all cases propagated throughout the extraction.
 
@@
 for the profile are used in the optimal extraction. The additional components
-share the PSF parameters and in cases where the additional component is an
+do not share the PSF parameters and in cases where the additional component is an
 extended object, the fits have been inspected to ensure that the additional
-component does now skew the fit towards a broader PSF. The additional
+component does now skew the fit towards a different PSF. The additional
 components are not used for the weights.
+
+The spectra are corrected for galactic extinction using the E(B-V) value from
+the dust maps of \citet{Schlegel1998} with the update in
+\citet{Schlafly2011}\footnote{Queried from
+	http://irsa.ipac.caltech.edu/applications/DUST/index.html using
+	\citet{astroquery}}, and the extinction curve by \cite{Cardelli1989} with a
+total to selective extinction $R_V = 3.1$. The wavelengths of the extracted
+1D-spectra are moved to vacuum, moved to the barycentric frame, and the
+wavelength recalibration described in Sect. \ref{wavecal} is applied. Pixels
+with pixel-to-pixel variation large than $50 \sigma$ are added to the bad pixel
+map.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Telluric correction} \label{tell_corr}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For all earth-based telescopes the light of interest has to pass through Earths
+atmosphere, where the atmospheric content and conditions make an imprint on the
+received spectrum. These telluric features can be calculated in a multitude of
+ways and we employ a prioritized list of methods here, depending on the
+availability of the different method. Since the observation are often taken at
+odd times under varying conditions, this prioritized list ensures that we are
+always doing the optimal correction.
+
+The highest priority method is using the GRB afterglow continuum itself, where
+the atmospheric conditions have directly been imprinted on the spectrum. The
+telluric features can directly be fit with an atmospheric model
+\citep{Smette2015, Kausch2015}, which can then be to correct for the absorption.
+The requirement here is that the afterglow continuum spectrum has a
+signal-to-noise higher than a value of 10.
+
+If the afterglow is not sufficiently bright, telluric standard stars observed
+close in the time to the GRB can be used as a proxy for the atmospheric
+condition during the GRB observation. Here we employ the telluric correction
+method that has been developed in \citet{Selsing2015}, where a library of
+synthetic templates is fit to the observed telluric standard.
+
+In the last case, where the object is neither bright enough, or there for some
+reason have not been observed a telluric standard, we rely on the synthetic sky
+model by \citep{Noll2012, Jones2013} for which we generate a synthetic
+transmission spectrum, which we then use, where the ambient parameters for the
+observations have been used.
 
@@
 
-\input{tables/burst_overview.tex}
+%\input{tables/burst_overview.tex}
 
@@
 
-\def\aj{AJ}
-\def\araa{ARA\&A}
-\def\apj{ApJ}
-\def\apjl{ApJL}
-\def\apjs{ApJS}
-\def\apss{Ap\&SS}
-\def\aap{A\&A}
-\def\aapr{A\&A~Rev.}
-\def\aaps{A\&AS}
-\def\mnras{MNRAS}
-\def\nat{Nature}
-\def\pasp{PASP}
-\def\aplett{Astrophys.~Lett.}
-
-\bibliographystyle{apj}
 \bibliography{XSGRB_sample}
-%\bibliography{thebib}
 
@@
 \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{The complex error function and the Voigt profile} \label{voigt}
+When modeling the spectral PSF, we need to evaluate the Voigt-profile.
+The Voigt profile, which is the convolution of the Gaussian and Lorentzian profiles, can, centered at zero, be written as \citep{pagnini2010}
+\begin{equation} 
+\begin{split}
+V(\lambda,\sigma, \gamma)  
+& = G(\lambda, \sigma)  \otimes L(\lambda, \gamma) \\
+& = \int_{-\infty}^{\infty} G(\xi, \sigma) L(\lambda - \xi, \gamma) d\xi \\
+& = \int_{-\infty}^{\infty} \frac{1}{\sqrt{2 \pi} \sigma} e^{- \left( \frac{\xi}{\sqrt{2} \sigma}  \right)^2 } \frac{1}{\gamma \pi} \frac{\gamma^2}{(\lambda - \xi)^2 + \gamma^2} d\xi \\
+& = \frac{\gamma}{\sqrt{2} \sigma} \frac{1}{ \pi^{3/2}}   \int_{-\infty}^{\infty} \frac{e^{- \left( \frac{\xi}{\sqrt{2} \sigma}  \right)^2 }}{(\lambda - \xi)^2 + \gamma^2}.
+\end{split}
+\end{equation}
+We can by making the following substitution, $\xi = \sqrt{2} \sigma$ and $d\xi = \sqrt{2} \sigma dt$, write it as
+\begin{equation} 
+\begin{split}
+V(\lambda,\sigma, \gamma)  
+& =  \frac{\sqrt{2} \sigma}{ \sqrt{{\pi}}} \frac{\frac{\gamma}{\sqrt{2} \sigma}}{\pi}  \int_{-\infty}^{\infty} \frac{e^{- t^2 }}{(\lambda - \sqrt{2} \sigma t)^2 + \gamma^2} dt \\
+& = \frac{1}{\sqrt{2 \pi} \sigma}  \frac{\frac{\gamma}{\sqrt{2} \sigma}}{\pi}  \int_{-\infty}^{\infty} \frac{e^{- t^2 }}{\left(\frac{\lambda}{\sqrt{2} \sigma} -  t\right)^2 + \left(\frac{\gamma}{\sqrt{2} \sigma}\right)^2} dt.
+\end{split}
+\end{equation}
+This form of the convolution is closely related to the complex probability function \citep{letchworth2007, abrarov2015a},
+\begin{equation} 
+\begin{split}
+W(z)  
+& = \frac{i}{\pi} \int_{-\infty}^{\infty} \frac{e^{-t^2}}{z - t}  
+\end{split}
+\end{equation}
+for any complex argument, $z = x + iy$. The complex probability function can be expressed as a sum of a real an imaginary part \citep{benner1995, abrarov2015b},
+\begin{equation} 
+\begin{split}
+W(x, y)  
+& = K(x, y) + i L(x, y) \\
+& = \frac{y}{\pi}  \int_{-\infty}^{\infty} \frac{e^{- t^2 }}{(x -  t)^2 +y^2} dt  + \frac{i}{\pi}  \int_{-\infty}^{\infty} \frac{(x - t)e^{- t^2 }}{(x -  t)^2 +y^2} dt,
+\end{split}
+\end{equation}
+where is the real part, $\mathtt{Re}[W(x, y)] =  \sqrt{2 \pi} \sigma V(\lambda,\sigma, \gamma)$ if $x = \frac{\lambda}{\sqrt{2} \sigma}$ and $y = \frac{\gamma}{\sqrt{2} \sigma}$, which can be obtained by using the complex argument, $z = \frac{\lambda + i\gamma}{\sqrt{2} \sigma}$, in the complex probability function.
+If $\mathtt{Im}[z] \geq 0$, which is always guaranteed for the width of a spectral profile, the complex probability function equals the complex error function.
+The complex error function has numerous, fast, numerical approximations where in this work we use the \texttt{scipy.special.wofz} \citep{scipy} implementation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-12-07 document/XSGRB_sample.tex
Change to A&A and HI

@@
 
-\bibliographystyle{aa}
+\bibliographystyle{mnras}
 \bibpunct{(}{)}{;}{a}{}{,}    %% natbib cite format used by A&A and ApJ
 
-\newcommand{\farc}{\hbox{$.\!\!^{\prime\prime}$}} 
-\newcommand{\ergA}{$\rm{erg\,cm^{-2}\,s^{-1}\,\AA^{-1}}$} 
-\newcommand{\erg}{$\rm{erg\,cm^{-2}\,s^{-1}}$} 
-\newcommand{\kms}{$\rm{km\,s^{-1}}\,$}
-
-\newcommand{\griz}{$g' r' i' z'$}
-\newcommand{\JHK}{$JHK_{\rm{s}}$}
-\newcommand{\gK}{$g' r' i' z' JHK_{\rm{s}}$}
-\newcommand{\Msun}{$M_\odot$}
-\newcommand{\lya}{Ly$\alpha$} 
-\newcommand{\lyb}{Ly$\beta$} 
-\newcommand{\lyg}{Ly$\gamma$} 
-\newcommand{\hb}{H$\beta$} 
-\newcommand{\ha}{H$\alpha$} 
-\newcommand{\hg}{H$\gamma$} 
-\newcommand{\hi}{\ion{H}{1}} 
-\newcommand{\hei}{[\ion{He}{1}]} 
-\newcommand{\oi}{[\ion{O}{1}]} 
-\newcommand{\sii}{[\ion{S}{2}]} 
-\newcommand{\siii}{[\ion{S}{3}]} 
-\newcommand{\oii}{[\ion{O}{2}]} 
-\newcommand{\oiii}{[\ion{O}{3}]}
-\newcommand{\nii}{[\ion{N}{2}]} 
-\newcommand{\nv}{[\ion{N}{5}]} 
-\newcommand{\neiii}{[\ion{Ne}{3}]} 
-\newcommand{\NIii}{[\ion{Ni}{2}]} 
-\newcommand{\feii}{\ion{Fe}{2}} 
-\newcommand{\civ}{\ion{C}{4}} 
-\newcommand{\cii}{\ion{C}{2}}
-\newcommand{\mgi}{\ion{Mg}{1}} 
-\newcommand{\mgii}{\ion{Mg}{2}} 
-\newcommand{\ali}{\ion{Al}{1}} 
-\newcommand{\alii}{\ion{Al}{2}} 
-\newcommand{\aliii}{\ion{Al}{3}} 
-\newcommand{\SIii}{\ion{Si}{2}} 
-\newcommand{\SIiv}{\ion{Si}{4}} 
-\newcommand{\znii}{\ion{Zn}{2}} 
-\newcommand{\crii}{\ion{Cr}{2}} 
-\newcommand{\mnii}{\ion{Mn}{2}} 
-\newcommand{\tiii}{\ion{Ti}{2}} 
-\newcommand{\caii}{\ion{Ca}{2}}
+% Manually specified definitions
+\input{definitions.tex}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample selection criteria and observations}
+\section{Sample selection criteria and observations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 blocking filter cutting the coverage of the NIR arm at 10200 - 21000 \AA.
+\\
 
@@
 slit width. We discuss this in Sect. \ref{resolution}.
-
+\\
 \todo{ADC discussion}
+
+We list the overview of all the observations in Tab. \ref{tab:sample_overview}.
+\input{tables/burst_overview.tex}
 
@@
 
-For all earth-based telescopes the light of interest has to pass through Earths
+For all earth-based telescopes, the light of interest has to pass through Earths
 atmosphere, where the atmospheric content and conditions make an imprint on the
@@
 telluric features can directly be fit with an atmospheric model
-\citep{Smette2015, Kausch2015}, which can then be to correct for the absorption.
+\citep{Smette2015, Kausch2015}, which can then be used to correct for the absorption. The accuracy of the correction depends on the 
 The requirement here is that the afterglow continuum spectrum has a
@@
 
-%\input{tables/burst_overview.tex}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Because the locations of GRBs are associated with the most intensely starforming
+regions \citep{Hogg1999, Bloom2002, Fruchter2006}, the GRB afterglow light has
+to propagate through the large amounts of hydrogen fueling the star formation.
+Because a significant fraction of the hydrogen has not been ionized yet, the
+optical depth at the wavelength of \lya~is very high, saturating at the line
+center. This causes a strong absorption system to appear in the afterglow
+continuum. For bursts with $z \gtrsim 1.7$, the position of \lya~moves into the
+spectroscopic coverage of X-shooter, meaning that we are able to detect this
+absorption trough due to \lya.
+
+Due of the stochastic nature of the \lya-forest, the blue wing of the Lyman
+alpha absorption line is randomly superposed with forest systems, making it
+notoriously difficult to model. Similarly so, the red wing has got the ISM
+signatures imprinted on it, especially strong absorption due to \SIii~which can
+exhibit significant velocity structure. Along with instrumental effects, the
+generative model for the data that we would use in a likelihood-based analysis
+would be almost boundless in complexity, making formal fitting of the column
+densities unfeasible.
+
+Using an analytic approximation to an absorption line profile and it's
+dependence on the column density \citep{TepperGarcia2006}, we overplot a
+synthetic absorption line with a specified column density on our observed
+spectrum. By tuning the value of the hydrogen column density until the synthetic
+absorption line matches the spectrum, we can thereby infer the actual column
+density of the GRB sight line in a manual way. In a similar fashion, the errors
+on the hydrogen column can be inferred. We show the results of this procedure
+for all bursts where possible in Fig. \ref{fig:HI1} and the inferred hydrogen
+column densities in Tab. \ref{tab:HI}
+
+\input{tables/HI_columns.tex}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
+	\caption{ Measurements of the hydrogen column-densities for all bursts with a clear Lyman alpha absorption system.
+In solid black is shown the spectrum with black dotted giving the corresponding 1-$\sigma$ error.
+Black dashed shows zero flux density.
+The solid red line is the absorption of column density equal to the value presented in Tab.
+\ref{HI_tab} with the 1-$\sigma$ interval shown with dashed lines.}
+	\label{fig:HI1}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
+	\caption{Same as Fig. \ref{fig:HI1}}
+	\label{fig:HI2}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
+	\caption{Same as Fig. \ref{fig:HI1}}
+	\label{fig:HI3}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
+	\caption{Same as Fig. \ref{fig:HI1}}
+	\label{fig:HI4}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
+	\caption{Same as Fig. \ref{fig:HI1}}
+	\label{fig:HI5}
+\end{figure*}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{X-ray column densities vs. optically derived ones}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-	\centerline{\includegraphics[width=10cm]{figures/OAXR_NH_sphere.pdf}}
-	\caption{Optically derived HI column densities vs.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{X-ray column densities vs. optically derived ones}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
+%
+%\begin{figure}
+%	\centerline{\includegraphics[width=10cm]{figures/OAXR_NH_sphere.pdf}}
+%	\caption{Optically derived HI column densities vs.
 X-ray derived ones taken from \citet{Buchner2016}.
@@
 }
-	\label{fig:OAXR}
-\end{figure}
+%	\label{fig:OAXR}
+%\end{figure}
 
@@
 10684\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10684.gcn3}} and is used in
-\citet{Skuladottir2010}, but not published elsewhere. Observations started 4
+Skuladottir (2010), but not published elsewhere. Observations started 4
 hours after the \textit{Swift} trigger, totaling 2.4 ks. Absorption features

### 2016-12-14 document/XSGRB_sample.tex
Additional data and text. Plot for follow-up timing

@@
 \usepackage{natbib}
-
-\bibliographystyle{mnras}
+\usepackage{caption}
+
 \bibpunct{(}{)}{;}{a}{}{,}    %% natbib cite format used by A&A and ApJ
@@
 
-\title{\vspace{-0.5cm}The X-shooter GTO sample of GRB afterglow and host Galaxy spectra}
+\title{\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample}
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	J.~Selsing\inst{1}, 
+	J.~P.~U. Fynbo\inst{1}, 
 	D.~Malesani\inst{1}, 
@@
 	P.~Goldoni\inst{14}, 
-	J.~P.~U. Fynbo\inst{1}, 
 	A.~de~Ugarte~Postigo\inst{11}, 
@@
 towards about 10$^5$ QSOs have 5 systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$
-been identified (Noterdaeme et al.\ 2012, A\&A, 547, L1). Long GRB afterglow
-spectra, by contrast, reveal such systems in the majority of cases (e.g., Jakobsson et
-al.\ 2006, A\&A, 460, L13; Fynbo et al.\ 2009, ApJS, 185, 526).
-Whereas DLAs towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due
-to the atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing
-redshifts (e.g., Rafelski et al.\ 2014, ApJL, 782, L29), GRBs allow us to see 
-into the hearts of star-forming galaxies over
-the full history of cosmic star formation from $z \approx 0$ to $z > 8$ (e.g.,
-Tanvir et al.\ 2009, Nature, 461, 1254; Salvaterra et al.\ 2009, Nature, 461,
-1258; Jakobsson et al.\ 2012, ApJ, 752, 62). With afterglow spectroscopy
-(throughout the electromagnetic spectrum from X-rays to the sub-mm) we can hence
-characterize the properties of star-forming galaxies over cosmic history in
-terms of  redshifts, metallicities, molecular contents, ISM temperatures,
-UV-flux densities, etc.. This is, however, only possible as long as there are
-satellites in orbit that rapidly and accurately locate GRBs. The currently
-operating \textit{Swift} satellite, launched in 2004 and still fully-functioning,
-allows for very efficient follow-up observations of GRBs due to its
-unprecedented rate, speed and precision of localisations.
-
-%With GTO time on X-shooter we
-%have started building a sample of afterglows with X-shooter spectroscopy.
-%X-shooter is ideal for this with its high efficiency, wide wavelength coverage
-%and spectral resolution sufficient for inferring the fundamental physical
-%parameters just listed.
-
-\smallskip
-\smallskip
-
-{\bf The case for a large sample of GRBs with X-shooter spectroscopy.}
-
-There are
-several unanswered but fundamental questions that must be addressed in order
-to exploit the full potential of GRBs as cosmological probes.
-%Among them: what is the true redshift distribution and corresponding luminosity
-%function of long-duration GRBs?
-More than 50\% of the \textit{Swift} bursts with measured redshift are at $z >
-2$, and  5--10\% are expected to be above $z=5$ (Salvaterra et al. 2008, MNRAS, 380, L45; Salvaterra et al.\ 2012, ApJ, 749,
-68; Jakobsson et al.\ 2012, ApJ, 752, 62; Perley et al.\ 2016, ApJ, 817, 7). 
-A high redshift completeness is
-crucial for our understanding of the link between the number density of GRBs
-per unit redshift and the global star-formation history of the Universe, as
-measured by other means (UV, FIR, sub-mm, see Robertson \& Ellis 2011, ApJ, 744,
-95). The detection of GRBs at $z > 6$ shows that GRBs have become competitive as
-a tool to identifying galaxies at the highest redshifts and unsurpassed in
-providing detailed abundance information via absorption line spectroscopy
-(Tanvir et al.\ 2012, ApJ, 754, 46, McGuire et al.\ 2016, ApJ submitted, 
-arxiv:1512.07808).
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
+\citep{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep{Jakobsson2006, Fynbo2009}. Whereas DLAs
+towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
+atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
+\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
+star-forming galaxies over the full history of cosmic star formation from $z
+\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
+	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we can hence characterize the properties of
+star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
+molecular contents, ISM temperatures, UV-flux densities, etc.. This is, however,
+only possible as long as there are satellites in orbit that rapidly and
+accurately locate GRBs. The currently operating \textit{Swift} satellite,
+launched in 2004 and still fully-functioning, allows for very efficient
+follow-up observations of GRBs due to its unprecedented rate, speed and
+precision of localisations.
+
+There are several unanswered but fundamental questions that must be addressed in
+order to exploit the full potential of GRBs as cosmological probes. More than
+50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and 
+5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
+	Salvaterra2012,Jakobsson2012, Perley2016}. A high redshift completeness is
+crucial for our understanding of the link between the number density of GRBs per
+unit redshift and the global star-formation history of the Universe, as measured
+by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
+GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
+galaxies at the highest redshifts and unsurpassed in providing detailed
+abundance information via absorption line spectroscopy \citep{Tanvir2012,
+	McGuire2016}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Complete sample}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In this section we describe the efficiency of the follow-up effort and the characteristics of the observed bursts. We also assess the degree to which the sample obtained is complete.
 
@@
 
-This sections should contain a plot og the follow-up delay and the acquisition image brightness at the time of observations.
+A rapid follow-up is essential for the accurate designation of GRB the origin. The redshift determination of burst where the host is not easily
+
+\begin{figure}
+	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}}
+	\caption{Follow-up timing and afterglow magnitude at the start of observation. The points have been colored based on the redshift of the corresponding burst. }
+	\label{fig:timing}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample completeness}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Based on the sample criteria specified in Sect.
+\ref{samplecrit}, a total of 152 \todo{Get number} bursts has been triggered by \textit{Swift} in the period since the comissioning of VLT/X-shooter in \todo{Get date}.
+This constitutes the \textit{statistical sample}.
 
@@
 Due of the stochastic nature of the \lya-forest, the blue wing of the Lyman
-alpha absorption line is randomly superposed with forest systems, making it
-notoriously difficult to model. Similarly so, the red wing has got the ISM
-signatures imprinted on it, especially strong absorption due to \SIii~which can
-exhibit significant velocity structure. Along with instrumental effects, the
+alpha absorption line is randomly superposed with forest systems, along with
+strong absorption from \mnii~and \SIiii, making it notoriously difficult to
+model. Additionally, Similarly so, the red wing has got the ISM signatures
+imprinted on it, especially strong absorption due to \SIii, \sii and \nv~which
+can exhibit significant velocity structure. Along with instrumental effects, the
 generative model for the data that we would use in a likelihood-based analysis
 would be almost boundless in complexity, making formal fitting of the column
-densities unfeasible.
+densities outside the scope of this paper.
 
@@
 	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Same as Fig. \ref{fig:HI1}}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
 	\label{fig:HI2}
@@
 	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Same as Fig. \ref{fig:HI1}}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
 	\label{fig:HI3}
@@
 	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Same as Fig. \ref{fig:HI1}}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
 	\label{fig:HI4}
@@
 	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Same as Fig. \ref{fig:HI1}}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
 	\label{fig:HI5}
@@
 
+\bibliographystyle{mnras}
 \bibliography{XSGRB_sample}
@@
 
-\subsection{GRB090313}
-
-\subsection{GRB100205A}
+\subsection{GRB090313 (z = 3.373)}
+
+\subsection{GRB090530 (z = 1.266)}
+
+\subsection{GRB090809 (z = 2.737)}
+
+\subsection{GRB090926 (z = 2.106)}
+
+\subsection{GRB091018 (z = 0.971)}
+
+\subsection{GRB091127 (z = 0.49)}
+
+\subsection{GRB100205A  (z = na)}
 Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected
@@
 
+\subsection{GRB110128A (z=2.339)}
+
+\subsection{GRB110407A (z=na)}
+
+\subsection{GRB110709B (z=na)}
+
+\subsection{GRB110715A (z=0.82)}
+
+\subsection{GRB110721A (z=0.382)}
+
+\subsection{GRB110808A (z=1.348)}
+
+\subsection{GRB110818A (z=3.36)}
+
+\subsection{GRB111005A (z=0.013)}
+
+\subsection{GRB111008A  (z=4.989)}
+
+\subsection{GRB111107A (z=2.893)}
+
+\subsection{GRB111117A (z=1.3?)}
+
+\subsection{GRB111123A  (z=3.151)}
+
+\subsection{GRB111129A (z=na)}
+
+\subsection{GRB111209A (z=0.677)}
+
+\subsection{GRB111211A (z=0.478)}
+
+\subsection{GRB111228A (z=0.716)}
+
 \subsection{GRB120118B (z = 2.943)}
@@
 
+\subsection{GRB120722A (z = 0.959)}
+The data presented here also formed the basis of GCN \#
+13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
+published elsewhere. On 4.8 ks integration time, starting 10 hours after the
+burst trigger, the simultaneous detection of absorption features belonging to
+\mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
+\hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the
+redshift of the GRB.
+
 \subsection{GRB120805A (z $\sim$ 3.9 NEW!!!)}
@@
 Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
-
-\subsection{GRB120722A (z = 0.959)}
-The data presented here also formed the basis of GCN \#
-13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
-published elsewhere. On 4.8 ks integration time, starting 10 hours after the
-burst trigger, the simultaneous detection of absorption features belonging to
-\mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
-\hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the
-redshift of the GRB.
 
@@
 
+\subsection{GRB120923A (z = $\gtrsim8$?)}
+
 \subsection{GRB121024A (z = 2.300)}
@@
 
+\subsection{GRB121201A (z = 3.385)}
+
+\subsection{GRB121229A (z = 2.707)}
+
+\subsection{GRB130131B (z = 2.539)}
+
 \subsection{GRB130408A (z = 3.758)}
@@
 $z=3.248$.
+
+\subsection{GRB130418A (z = 1.218)}
+
+\subsection{GRB130427A (z = 0.34)}
+
+\subsection{GRB130427B (z = 2.78)}
+
+\subsection{GRB130603B (z = 0.356)}
 
@@
 
+\subsection{GRB130612A (z = 2.006)}
+
+\subsection{GRB130615A (z = $~3$?)}
+
+\subsection{GRB130701A (z = 1.155)}
+
+\subsection{GRB131011A (z = 1.874)}
+
+\subsection{GRB131030A (z = 1.296)}
+
+\subsection{GRB131103A (z = 0.599)}
+
+\subsection{GRB131105A (z = 1.686)}
+
+\subsection{GRB131117A (z = 4.042)}
+
+\subsection{GRB131231A (z =0.642)}
+
+\subsection{GRB140114A (z = $\sim2.8$?)}
+
+\subsection{GRB140213A (z = 1.208)}
+
+\subsection{GRB140301A (z = 1.416)}
+
+\subsection{GRB140311A (z = 4.95)}
+
+\subsection{GRB140430A (z = 1.601)}
+
+\subsection{GRB140506A  (z = 0.889)}
+
+\subsection{GRB140515A (z = $\sim6.32$?)}
+
+\subsection{GRB140614A (z = 4.233)}
+
+\subsection{GRB140622A (z = 0.959)}
+
+\subsection{GRB141028A (z = 2.332)}
+
+\subsection{GRB141031A  (z = na)}
+
+\subsection{GRB141109A (z = 2.993)}
+
+\subsection{GRB150206A (z = 2.087)}
+
+\subsection{GRB150301B (z = 1.517)}
+
+\subsection{GRB150403A (z = 2.06)}
+
+\subsection{GRB150423A (z = 1.394)}
+
+\subsection{GRB150428A (z = na)}
+
+\subsection{GRB150514A (z = 0.807)}
+
+\subsection{GRB150518A (z = 0.256)}
+
+\subsection{GRB150616A (z = 1.188)}
+
+\subsection{GRB150727A (z = 0.313)}
+
+\subsection{GRB150821A (z = 0.755)}
+
+\subsection{GRB150831A (z = na)}
+
+\subsection{GRB150910A (z = 1.359)}
+
+\subsection{GRB150915A (z = 1.968)}
+
 \subsection{GRB151021A (z = 2.330)}
@@
 
+\subsection{GRB151027B (z = 4.063)}
+
+\subsection{GRB151029A (z = 1.423)}
+
+\subsection{GRB151031A (z = 1.167)}
+
+\subsection{GRB160117B (z = 0.87)}
+
 \subsection{GRB160203A (z = 3.517)}
@@
 at $z=3.517$ as well as intervening absorption at $z=2.203$.
+
+\subsection{GRB160228A (z = 1.64)}
+
+\subsection{GRB160303A (z = na)}
+
+\subsection{GRB160314A (z = 0.726)}
+
+\subsection{GRB160410A (z = 1.717)}
+
+\subsection{GRB160425A (z = 0.555)}
+
+\subsection{GRB160625B (z = 1.406)}
 
@@
 detections.
+
+\subsection{GRB161001A (z = 0.891?)}
 
@@
 
+\subsection{GRB161023A (z = 2.710)}
+
+\subsection{GRB161117A (z = 1.549)}
+
 \end{document}

### 2016-12-28 document/XSGRB_sample.tex
Writing completeness section and adding figures

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Being of transient nature, it is difficult to impose strong sample selection
@@
 been stated a bit differently, the visibility constraint being replaced by a
-declination + Sun angle constraint. The above criteria are however those
+declination + Sun angle constraint. The below criteria are however those
 defining the sample.}
@@
 
-A significant fraction of the bursts presented here \todo{Insert exact
+A fraction of the bursts presented here \todo{Insert exact
 number} have already had their hosts investigated in \citet{Kruhler2015}, for
@@
 
-All the observations in this sample has been carried out using the
-cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
-two of the four Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal)
-during the duration of this follow-up campaign, which covers the entire lifetime
-of X-shooter. Observations have been carried out from the ESO period P84
-through P98 under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
-095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.
+Observations in this sample has been carried out using the cross-dispersed
+echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on two of the four
+Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the duration
+of this follow-up campaign, which covers the entire lifetime of X-shooter.
+Observations have been carried out from the ESO period P84 through P98 under the
+following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.A-0045,
+094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088, 089.A-0067,
+088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. A few additional
+bursts, not a part of the \textit{statistical sample}, have been included from
+092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI: Greiner), and Levan under
+095.B-0811(B) (PI: Levan). This sample represents \textit{all} GRBs that have
+ever been followed up by X-shooter.
 
@@
 while X-shooter was mounted on UT3 during the commissioning of the instrument.
-The bursts observed during the commisioning period is not a part of our
+The bursts observed during the commissioning period is not a part of our
 statistical sample, but are nonetheless published as part of the bursts observed
@@
 a single exposure, by splitting the light in three separate spectroscopic arms
-using dichroics.
-This way each arm work as a separate instrument, each functioning as its own
-echelle spectrograph.
-The ultraviolet blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm
-covering 5500 - 10200 \AA, and the near-infrared()NIR) arm covering 10200 -
-24800 \AA~ with the possibility of applying a k-band
-blocking filter cutting the coverage of the NIR arm at 10200 - 21000 \AA.
+using dichroics. This way each arm work as a separate instrument, each
+functioning as its own echelle spectrograph. The ultraviolet blue(UVB) arm
+covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 - 10200 \AA, and the
+near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the possibility of
+applying a k-band blocking filter cutting the coverage of the NIR arm at 10200 -
+21000 \AA.
 \\
@@
 
-In this section we describe the efficiency of the follow-up effort and the characteristics of the observed bursts. We also assess the degree to which the sample obtained is complete.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up timing and afterglow brightness}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A rapid follow-up is essential for the accurate designation of GRB the origin. The redshift determination of burst where the host is not easily
+In this section we describe the efficiency of the follow-up effort and the
+characteristics of the observed bursts. We also assess the degree to which the
+sample obtained is complete. An important note is that here we present
+\textit{all} GRBs that have ever been observed with X-shooter, while only a
+subset of these constitute our \textit{statistical} sample. The statistical
+sample is based on the selection criteria described in Sect. \ref{samplecrit}
+and should represent an unbiased sample of the underlying, parent GRB
+population. Single bursts outside the sample criteria have been followed up on
+the basis of the curiosity of their light curve or their brightness, but is not
+discussed as part of the investigation of the GRB population. A prime example of
+a burst outside the statistical sample is GRB161023A with it's rich
+spectrum containing 11 intervening absorption systems and pending further analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up timing and afterglow brightness} \label{timing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Redshift determination of bursts where the host is too faint for a spectroscopic
+redshift relies on the absorption spectrum imprinted on the GRB afterglow
+continuum, which again depends on the afterglow brightness at time of follow-up.
+Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better references} a rapid
+follow-up is essential for the accurate designation of the GRB host \todo{Get
+reference}. In Fig. \ref{fig:timing} we plot the follow-up delay from the BAT
+trigger to the start of the spectroscopic observation. Visible as the
+shortest delays are the bursts observed in RRM-mode in which where the fastest follow-up
+between BAT trigger and start of spectroscopic observations is for the short, $z = 1.717$, GRB160410A which was
+followed up only 8.4 minutes after the BAT trigger.
 
@@
 	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}}
-	\caption{Follow-up timing and afterglow magnitude at the start of observation. The points have been colored based on the redshift of the corresponding burst. }
+	\caption{Follow-up timing and afterglow magnitude at the start of observation.
+	The points have been colored based on the redshift of the corresponding burst.
+	Points with black faces does not have a redshift determination and downarrows
+	indicate a non-detection in the acquisition image. In blue is shown the
+	fractional redshift completess as a function of delay-time.}
 	\label{fig:timing}
@@
 
+To illustrate the effect of redshift completeness loss due to late follow-up, we
+overplot the redshift completeness fraction over the delay time. Visible is the
+increasing fraction of GRBs without a good redshift determination for increasing
+delays. That the redshift completeness of the bursts we have followed up are $>$
+90 percent \todo{Get number}, shows the ability of VLT/X-shooter to successfully
+get redshifts. Not shown in the figure is an additional 12 \todo{Get number}
+bursts that we have redshift determinations from late-time host observations
+with delay-times longer than $\sim$10 days.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Based on the sample criteria specified in Sect.
-\ref{samplecrit}, a total of 152 \todo{Get number} bursts has been triggered by \textit{Swift} in the period since the comissioning of VLT/X-shooter in \todo{Get date}.
-This constitutes the \textit{statistical sample}.
+Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
+\todo{Get number} bursts has been triggered by \textit{Swift} in the period
+since the commissioning of VLT/X-shooter in \todo{Insert date}. This sample
+constitutes the \textit{statistical sample} on which we will try to address the
+GRB population properties. Of this sample, 86 \todo{Get number} have been
+spectroscopically followed up with X-shooter. In order to assess whether our
+sample criteria and the subset of bursts followed up represent the underlying
+parent population of GRB, we compare some intrinsic GRB properties of our sample
+to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
+that \textit{Swift} randomly samples the underlying GRB population, if the
+sample statistics of our subsample are similar to the \textit{Swift}-sample, we
+can assume that the population properties should be conserved. There is some
+evidence that the burst duration does not scale with redshift as expected
+through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
+that we could be missing an increasing fraction of the bursts at higher
+redshifts, but \citet{Lien2016} argues that we should only be missing the bright
+"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
+the sample presented here is complete. We show the comparison between the burst
+properties in Fig. \ref{fig:swift_complete}.
+
+\begin{figure*}
+	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
+	\caption{Comparison between bursts properties of all bursts observed with
+	\textit{Swift}, the properties of the ones that fullfill the sample criteria
+	specified in Sect. \ref{samplecrit}, and the subset that has been observed as
+	part of the statistical sample. The left plot shows the fluence in 15-150 keV
+	as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
+	after the bursts as measured by XRT. Based on the XRT spectrum,
+	\citep{Evans2009}, a hydrogen column density has been inferred which we show in
+	the right most panel.}
+	\label{fig:swift_complete}
+\end{figure*}
+
+Using the observational characteristics of the 1244 \todo{Get number} bursts
+observed by \textit{Swift} to date \cite{Gehrels2009}, and the X-ray derived
+hydrogen column densities \citep{Evans2009}, we can quantify the degree to which
+our sample is biased against the overall \textit{Swift} sample. The values are
+queried from then Three main samples are of interest to compare to assess the
+completenessof the follow-up campaign; the full \textit{Swift} sample consisting
+of all the burst observed by \textit{Swift} (1), the burst that fulfill the
+selection criteria imposed in Sect. \ref{samplecrit} (2), and the bursts
+actually followed up with X-shooter (3).
+
+\input{tables/sample_properties.tex}
+
+Using a
+2-sided Kolmogorov-Smirnov (KS), we can assess the degree to which the null that
+the two distribution are drawn from the same distribution can be rejected.
+
+\begin{figure}
+	\centerline{\includegraphics[width=9cm]{figures/p_values.pdf}}
+	\caption{Relational graph shown the .}
+	\label{fig:p_values}
+\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{On the nature of triggered GRBs without detected optical afterglow.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?
+\subsection{The nature of triggered GRBs without detected optical afterglow.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This should include observations where the acquisition image has been taken, but
+observations were discontinued due to non-detection. What are the X-ray
+properties of these bursts?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{On the redshift distribution of GRBs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Because the redshift completeness of our statistical sample is  54 percent
+\todo{Get number}, making inference of the underlying redshift distribution of
+GRBs is difficult. Because a large fraction of the bursts for which we were
+unable to secure a redshift are missing due to terrestrial reasons, see Sect.
+\ref{badbursts}, imposing additional unbiasing cuts on the sample can increase
+the redshift completeness at the cost of sample size. As long as the cuts we
+impose are not selective in terms of the burst properties, the sample properties
+stay the same. The first cut we try is telescope availability such as bad
+weather, visitor rejected the TOO, or unavailability of the instrument. Because
+telescope availability is independent on the burst properties, by cutting away
+these from our sample, the underlying redshift distribution should be conserved.
+By trimming away the 30 bursts \todo{Get number} that can be rejected due to
+telescope availability, the redshift completeness increases to 63 percent
+\todo{Get number}. As is shown in Sect. \ref{timing}, the redshift completeness
+decreases with increasing follow-up delay, we can additionally increase the
+redshift completeness by imposing cuts on the follow-up delay. Our ability to
+rapidly follow up a burst is also not dependent on the burst properties and
+therefore we can additionally cut delays longer than 24 hours (all burst in the
+sample are in principle observable within 24 hours).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The biased introduced in the redshift distribution by the increased ability to
+secure redshift of optically brighter bursts has been investigated by
+\citet{Turpin2016} which find that we systematically find redshift for longer
+GRBs. Additionally only the brightest GRBs are seen above redshift $z \gtrsim
+1$. \todo{Additional text}
 
@@
 model. Additionally, Similarly so, the red wing has got the ISM signatures
-imprinted on it, especially strong absorption due to \SIii, \sii and \nv~which
+imprinted on it, especially strong absorption due to \SIii, \sii~and \nv~which
 can exhibit significant velocity structure. Along with instrumental effects, the
@@
 to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii,
-all at a consistent redshift. Temporal evolution of the fine-structure lines
+all at a consistent redshift. Temporal evolution of the fine structure lines
 belonging to \feii$^*$ is found between the epochs.
@@
 \civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
-fine-structure lines from \NIii and \feii are exited local to the GRB.
+fine structure lines from \NIii and \feii are exited local to the GRB.
 Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
@@
 \SIii, \sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along
-with the corresponding fine-structure lines from (\feii$^*$, \SIii$^*$,
+with the corresponding fine structure lines from (\feii$^*$, \SIii$^*$,
 \oi$^*$, \oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
@@
 absorption feature from Lyman alpha, along with narrow lines from \civ, \SIii,
-\SIiv, \feii, \sii, and \alii, as well as fine-structure lines associated with
+\SIiv, \feii, \sii, and \alii, as well as fine structure lines associated with
 \SIii$^*$ are all detected at $z = 2.300$, securely setting it as the redshift

### 2016-12-28 document/tables/sample_properties.tex
Writing completeness section and adding figures

\begin{table*}[!ht]

	\centering
	\begin{tabular}{cccc}
		\hline
		\hline\noalign{\smallskip}
		{} & {Full \textit{Swift} sample} & {Statistical sample} &  {Followed up bursts} \\
		\hline\noalign{\smallskip}
		N$_{BAT}$ & 1071 & 156 & 85\\
		$\log(15-150$keV fluence)  & $-5.9_{-0.7}^{+0.6}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$   \\
		N$_{XRT}$ & 1004 & 154 & 84\\
		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.6}^{+0.8}$  \\
		N$_{HI}$ & 1034 & 155 & 85\\
		$\log(N_H)$ & $21.4_{-0.5}^{+0.5}$ &  $-21.3_{-0.6}^{+0.6}$ &  $21.4_{-0.7}^{+0.7}$  \\

		\hline\noalign{\smallskip}

	\end{tabular}
	\caption{Population properties for the \textit{Swift sample} and the subset of bursts fulfulling the sample criteria. The population characteristics for the three samples exhibit a large degree of similarity, signifying the ability of our selection criteria to effectively capture the underlying population.  \label{tab:sample_properties}}
\end{table*}

### 2016-12-28 py/Sample_completeness.py

@@ -1,0 +1,170 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib
+from matplotlib.ticker import FormatStrFormatter
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [9, 4.5]
+   }
+matplotlib.rcParams.update(params)
+
+def filt_nan(input_array):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except:
+            pass
+    return np.array(holder)
+
+
+def main():
+    """
+    Small script to produce sample completeness plots.
+    """
+
+
+    # Read in burst list
+    # burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
+    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
+
+    name, z, observed, OA, BAT, XRT, HI = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values, burst_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values.astype("float"), burst_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values.astype("float"), burst_table["XRT Column Density (NH) [10^21 cm^-2]"].values.astype("float")
+
+    BAT_o = BAT[observed == "Yes"]
[…]

### 2016-12-28 py/follow_up_timing.py

@@ -4,6 +4,7 @@
 from __future__ import division, print_function
 
 from astropy.io import fits
+import pandas as pd
 import matplotlib.pyplot as pl
 import seaborn; seaborn.set_style('ticks')
 import numpy as np
@@ -15,7 +16,7 @@ params = {
    'legend.fontsize': 10,
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
-   'text.usetex': True,
+   'text.usetex': False,
    'figure.figsize': [4.5, 4.5]
    }
 matplotlib.rcParams.update(params)
@@ -29,62 +30,81 @@ def main():
 
 
     # Read in burst list
-    burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
+    # burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
+    burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
 
-    name, z, delay, mag = burst_table[:, 0], burst_table[:, 1].astype(float), burst_table[:, 2].astype(float), burst_table[:, 3].astype(float)
+    name, z, delay, mag = burst_table["GRB"].values, burst_table["z"].values, burst_table["Follow-up delay"].values, burst_table["Acquisition mag"].values
+
+    idx_limits = ~(mag == ">24")
+    magnondet = 24 * (~idx_limits).astype("int")
+
+    delay_sort = np.argsort(delay)
+    sorted_delay = delay[delay_sort]
+    sorted_z = z[delay_sort]
+    fractional_completeness = (1 - np.cumsum(np.isnan(sorted_z).astype("int"))/len(sorted_z))*100
+    print(len(sorted_z), np.sum(np.isnan(sorted_z).astype("int")))
+    print(fractional_completeness[-1])
+    exit()
+    idx_hasz = ~np.isnan(z)
 
     # Plot
-    fig, ax = pl.subplots()
+    fig, ax1 = pl.subplots()
 
     cmap = pl.get_cmap("plasma")
+    # With redshift and acq mag
[…]

### 2016-12-31 document/XSGRB_sample.tex
Adding skymap and text

@@
 
-\title{\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample}
+\title{XSGRB - The X-shooter GRB afterglow sample}
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 
-\titlerunning{XSGRB sample}
+\titlerunning{The X-shooter sample of GRBs}
 
@@
 	J.~Selsing\inst{1}, 
-	J.~P.~U. Fynbo\inst{1}, 
 	D.~Malesani\inst{1}, 
 	T.~Kr\"{u}hler\inst{1}, 
+	J.~P.~U. Fynbo\inst{1}, 
 	P.~Goldoni\inst{14}, 
@@
 From March 2005 to March 2016 there have been about 350 {\it Swift} bursts
-complying with
-our sample selection criteria (see ``Immediate objective''), and about half of them have
-measured redshifts. Among the latter subset, the team proposing these observations has
-measured about two thirds of the redshifts, mainly with FORS1/2 and X-shooter (see
-Fynbo et al.\ 2009, ApJS, 185, 526;  Jakobsson et al.\ 2012, ApJ, 752, 62;
-Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our current aim is to build a sample superior to
-our previous low-resolution survey (Fynbo et al.\ 2009, ApJS, 185, 526), both in
-terms of quantity and  quality of the spectra. Our program started as 
-guaranteed time observations during
-periods 84-91, and we have continued in open time since then.
+complying with our sample selection criteria (see ``Immediate objective''), and
+about half of them have measured redshifts. Among the latter subset, the team
+proposing these observations has measured about two thirds of the redshifts,
+mainly with FORS1/2 and X-shooter (see Fynbo et al.\ 2009, ApJS, 185, 526; 
+Jakobsson et al.\ 2012, ApJ, 752, 62; Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our
+current aim is to build a sample superior to our previous low-resolution survey
+(Fynbo et al.\ 2009, ApJS, 185, 526), both in terms of quantity and  quality of
+the spectra. Our program started as guaranteed time observations during periods
+84-91, and we have continued in open time since then.
 
 %In P84 we started the X-shooter legacy survey in GTO time. Our aim is to  build
-%a sample of GRBs superior to our previous  low-resolution  survey, Fynbo et al.\
-%2009, ApJS, 185, 526) both in terms of quantity and  quality of the spectra. The
+%a sample of GRBs superior to our previous  low-resolution  survey, Fynbo et
+% al.\
+%2009, ApJS, 185, 526) both in terms of quantity and  quality of the spectra.
+% The
 %last period of GTO time on the survey was P90 (plus 6 hr of leftover GTO time
@@
 
-X-shooter is in many ways
-the ideal GRB follow-up instrument and indeed GRB follow-up was one of the
-primary science cases behind the instrument design and implementation. Our
-program secures general purpose GRB afterglow spectroscopic follow-up 
-that adds strong legacy value to the \textit{Swift} GRB sample.
-Due to the wide wavelength range of X-shooter  
-with the same observation cover molecular H$_2$ absorption near the
-atmospheric cut-off and all the strong emission lines from the host in the
-NIR arm (e.g., Friis et al., 2015, MNRAS, 451, 167). In general, the wide wavelength coverage ensures that we always have
-features on which to base a redshift measurement as long as the afterglow is
-brighter than about 23 mag in either the $R$- or $z$-band. 
-Frequently, emission lines are also detected from the underlying host, which
-also provide further information such as SFR and metallicity (the top right 
-panel in Fig.~1 shows an example). Only for 
-7 out of more than 70 secured spectra could we not measure a redshift.   
-With the X-shooter survey we
-provide {\bf metallicity measurements} for about 30\% (Voigt-profile fits)
-of the $z>1.7$ events.
-So far we have
-measured metallicities for more than 20 GRB afterglows with X-shooter.
-With the wide wavelength coverage of X-shooter we can
-study important chemical species as Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$
-elements over a much wider redshift range than what is possible with other
-instruments.
-As an example, we have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A
-at $z=4.669$ (Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for 
-GRB\,111008A at $z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$
-for the $z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139).  
-Reconciling the
-abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs
-in particular, and old stars in the Local Group is an important long-term
-goal (see also Sparre et al.\ 2014, ApJ, 785, 150).
-Metallicities are also measured from host emission lines (Kr{\"u}hler et al.\ 2015,
-A\&A, 581, A125).
-GRB spectroscopy also allows us to determine the dust content of their environments,
+X-shooter is in many ways the ideal GRB follow-up instrument and indeed GRB
+follow-up was one of the primary science cases behind the instrument design and
+implementation. Our program secures general purpose GRB afterglow spectroscopic
+follow-up that adds strong legacy value to the \textit{Swift} GRB sample. Due to
+the wide wavelength range of X-shooter with the same observation cover molecular
+H$_2$ absorption near the atmospheric cut-off and all the strong emission lines
+from the host in the NIR arm (e.g., Friis et al., 2015, MNRAS, 451, 167). In
+general, the wide wavelength coverage ensures that we always have features on
+which to base a redshift measurement as long as the afterglow is brighter than
+about 23 mag in either the $R$- or $z$-band. Frequently, emission lines are also
+detected from the underlying host, which also provide further information such
+as SFR and metallicity (the top right panel in Fig.~1 shows an example). Only
+for 7 out of more than 70 secured spectra could we not measure a redshift. With
+the X-shooter survey we provide {\bf metallicity measurements} for about 30\%
+(Voigt-profile fits) of the $z>1.7$ events. So far we have measured
+metallicities for more than 20 GRB afterglows with X-shooter. With the wide
+wavelength coverage of X-shooter we can study important chemical species as
+Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$ elements over a much wider
+redshift range than what is possible with other instruments. As an example, we
+have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A at $z=4.669$
+(Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for GRB\,111008A at
+$z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$ for the
+$z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139). Reconciling the
+abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs in
+particular, and old stars in the Local Group is an important long-term goal (see
+also Sparre et al.\ 2014, ApJ, 785, 150). Metallicities are also measured from
+host emission lines (Kr{\"u}hler et al.\ 2015, A\&A, 581, A125). GRB
+spectroscopy also allows us to determine the dust content of their environments,
 both through analysis of the depletion pattern and through measurement of the
-associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). 
-This allows us to quantify the dust-to-metals ratio and its
-evolution with redshift (e.g., De Cia et al.\ 2013, A\&A, 560, 88; Zafar \&
-Watson 2013, A\&A, 560, 26).
-%Moreover, GRB
+associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). This allows us to
+quantify the dust-to-metals ratio and its evolution with redshift (e.g., De Cia
+et al.\ 2013, A\&A, 560, 88; Zafar \& Watson 2013, A\&A, 560, 26). %Moreover,
+% GRB
 %afterglows allow investigation of random foreground absorbers, complementing
@@
 found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
-2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). 
-We are currently analysing more of our spectra for less obvious molecular
-absorption and we expect to find more (a dedicated sample paper is 
-addressing this issue).
-
-\vspace{0.2cm}
-A natural question to ask is: {\bf how long should this work continue?} Our
-view is that we need to keep observing the afterglows as long as we have
-\textit{Swift} in operation. Also note that the program is still producing many 
-papers and provides data for many theses (Box 9 and 10). \textit{Swift} is currently 
-funded until 2018, but is likely to get more extensions given its overwhelming 
-success. As
-mentioned above GRBs allow us to probe star-forming galaxies that are almost
-impossible to study in other ways both in terms of redshifts, galaxy luminosity
-function, and regions within galaxies. After 7 periods we have 
-secured seven spectra of $z>4$ GRBs, of which three were of sufficient quality to
-allow abundance measurements (Th\"one et al.\ 2013, Sparre et al.\ 2014, 
-Hartoog et al.\ 2015). GRBs offer the only way to derive chemical abundances 
-for the gas phase of central, actively star-forming regions of high-$z$ galaxies. 
-The program also maintains a very high discovery
-potential where we occasionally find something completely unexpected that 
-provides interesting clues to puzzles in other fields, e.g. extinction of 
-type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of these spectra are like
-precious jewels – it is a type of observation that can never be repeated and a
-class of sightlines that can only be studied while we have operating GRB
-satellites.
+2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). We are currently
+analysing more of our spectra for less obvious molecular absorption and we
+expect to find more (a dedicated sample paper is addressing this issue).
+
+\vspace{0.2cm} A natural question to ask is: {\bf how long should this work
+	continue?} Our view is that we need to keep observing the afterglows as long as
+we have \textit{Swift} in operation. Also note that the program is still
+producing many papers and provides data for many theses (Box 9 and 10).
+\textit{Swift} is currently funded until 2018, but is likely to get more
+extensions given its overwhelming success. As mentioned above GRBs allow us to
+probe star-forming galaxies that are almost impossible to study in other ways
+both in terms of redshifts, galaxy luminosity function, and regions within
+galaxies. After 7 periods we have secured seven spectra of $z>4$ GRBs, of which
+three were of sufficient quality to allow abundance measurements (Th\"one et
+al.\ 2013, Sparre et al.\ 2014, Hartoog et al.\ 2015). GRBs offer the only way
+to derive chemical abundances for the gas phase of central, actively
+star-forming regions of high-$z$ galaxies. The program also maintains a very
+high discovery potential where we occasionally find something completely
+unexpected that provides interesting clues to puzzles in other fields, e.g.
+extinction of type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of
+these spectra are like precious jewels – it is a type of observation that can
+never be repeated and a class of sightlines that can only be studied while we
+have operating GRB satellites.
 
@@
 distribution of night shifts, the scientific exploitation of the data is
-efficient and where we are open to all new members who wish to participate.
-As mentioned all data are public immediately.
-
-{\bf For all of these reasons, we need to keep
-	building up the sample of GRB afterglow spectra now as we may have to wait many
-	years before a mission like \textit{Swift} becomes available again.}
-
-A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark bursts}'', e.g.,
-Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012, MNRAS, 421, 1265). 
-Some of these are at the highest
-redshifts ($z > 6$) and their observer-frame optical emission is absorbed by
-the IGM. The majority, however, suffer from large dust obscuration (e.g., Perley
-et al. 2009, AJ, 138, 1690; Greiner et al.\ 2011, A\&A, 526, 30).
-Identifying such GRBs is important for
-constraining the fraction of obscured star formation. In both cases,
-NIR emission is
-expected. X-shooter
-can adequately study these objects, provided that a NIR counterpart is timely
-identified, for which we have the dedicated HAWKI run D.
+efficient and where we are open to all new members who wish to participate. As
+mentioned all data are public immediately.
+
+{\bf For all of these reasons, we need to keep building up the sample of GRB
+	afterglow spectra now as we may have to wait many years before a mission like
+	\textit{Swift} becomes available again.}
+
+A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark
+	bursts}'', e.g., Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012,
+MNRAS, 421, 1265). Some of these are at the highest redshifts ($z > 6$) and
+their observer-frame optical emission is absorbed by the IGM. The majority,
+however, suffer from large dust obscuration (e.g., Perley et al. 2009, AJ, 138,
+1690; Greiner et al.\ 2011, A\&A, 526, 30). Identifying such GRBs is important
+for constraining the fraction of obscured star formation. In both cases, NIR
+emission is expected. X-shooter can adequately study these objects, provided
+that a NIR counterpart is timely identified, for which we have the dedicated
+HAWKI run D.
 
@@
 burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
-2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\
-2012, A\&A, 545, 64; Hartoog et al.\ 2013).
-{\bf Short GRBs} originate in a substantially different
-environment compared to long GRBs. Short GRBs may be related to the merging of
-compact binaries and the coalescence time can be long enough to allow the
-progenitor system to move far away from the star formation site (Belczynski et
-al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum with a sufficient 
-signal-to-noise ratio
-of a short GRB afterglow has been secured.
-A knowledge of the redshift distribution
-of short bursts is of key importance for the next generation of gravitational
-wave experiments, as they are the likely EM counterparts to their
-primary targets.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria and observations}
+2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\ 2012,
+A\&A, 545, 64; Hartoog et al.\ 2013). {\bf Short GRBs} originate in a
+substantially different environment compared to long GRBs. Short GRBs may be
+related to the merging of compact binaries and the coalescence time can be long
+enough to allow the progenitor system to move far away from the star formation
+site (Belczynski et al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum
+with a sufficient signal-to-noise ratio of a short GRB afterglow has been
+secured. A knowledge of the redshift distribution of short bursts is of key
+importance for the next generation of gravitational wave experiments, as they
+are the likely EM counterparts to their primary targets.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sample selection criteria}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Observations} \label{obs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Observations} \label{obs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[!ht]
+	\centerline{\includegraphics{figures/skymap.pdf}}
+	\caption{Mollweide projection in galactic coordinates of the full sky showing
+		the positions on the sky of the bursts presented in this work. The empty
+		circles are the positions of all the \textit{Swift} bursts detected until the
+		end of 2016. In blue is shown the position of the bursts fulfilling the sample
+		criteria specified in Sect. \ref{samplecrit}. The red circles are added around
+		GRBs which have been followed up by X-shooter since the commissioning of the
+		instrument. Blue points with red circles represent GRBs that enter our sample
+		and have X-shooter spectroscopy. The different samples are compared in Sect.
+		\ref{results}. The background shows the dust maps presented in
+		\citet{Schlegel1998} where the sample criteria cut with $A_V \lesssim 0.5$ mag
+		is removed.}
+	\label{fig:skymap}
+\end{figure*}
 
@@
 We list the overview of all the observations in Tab. \ref{tab:sample_overview}.
-\input{tables/burst_overview.tex}
+\input{tables/burst_overview.tex} and plot the positions off all the bursts on
+the night sky in Fig. \ref{fig:skymap}. The central zone of avoidance due to
+high galactic extinction cutoff is visible with an apparent isotropic
+distribution of the bursts away from the galactic center except in the upper
+right corner, where Earth casts a shadow as seen from the telescope.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section{Results} \label{results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample completeness}	
+\subsection{Sample completeness} \label{completeness}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 hydrogen column densities \citep{Evans2009}, we can quantify the degree to which
-our sample is biased against the overall \textit{Swift} sample. The values are
-queried from then Three main samples are of interest to compare to assess the
-completenessof the follow-up campaign; the full \textit{Swift} sample consisting
-of all the burst observed by \textit{Swift} (1), the burst that fulfill the
-selection criteria imposed in Sect. \ref{samplecrit} (2), and the bursts
-actually followed up with X-shooter (3).
+our sample is biased against the overall \textit{Swift} sample in terms of
+intrinsic GRB properties. The values are queried from the online
+\textit{Swift} database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
+Three main samples are of interest to compare to assess the completeness of the
+follow-up campaign; the full \textit{Swift} sample consisting of all the bursts
+observed by \textit{Swift} (1), the burst that fulfill the selection criteria
+imposed in Sect. \ref{samplecrit} (2), and the bursts actually followed up with
+X-shooter (3).
 
@@
 
-Using a
-2-sided Kolmogorov-Smirnov (KS), we can assess the degree to which the null that
-the two distribution are drawn from the same distribution can be rejected.
+For each of the samples we calculate the median and 14th and 86th percentile
+which can be used as point estimates for the population distribution. These are
+shown in \ref{tab:sample_properties}. It can be seen from the values that the
+three samples have very similar distributions in terms of the point estimates
+chosen, suggesting that our selection criteria are unbiasing against the
+\textit{Swift}-sample and that additionally, the follow-up effort is unbiasing
+towards the intrinsic GRB properties. This is despite spectroscopic follow-up
+only being carried out in cases where either a detectable optical afterglow or a
+clear counterpart are seen, which naively should be biased against dark bursts
+occurring in more obscured galaxies, which is shown to exhibit different
+galactic properties \citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b,
+	Perley2015b}.
+
+Additionally, using a 2-sided Kolmogorov-Smirnov test (KS-test), we can assess
+the degree to which the null that the two distribution are drawn from the same
+distribution can be rejected. We show a graphic representation of the test
+statistics in Fig. \ref{fig:p_values}. We have run the test on whether is can be
+rejected that the sample criteria is drawn from the same distribution as the
+\textit{Swift} sample, if the observed sample is dissimilar from the statistical
+sample, and again whether the observed sample differs from the full
+\textit{Swift} sample. A high p-value indicates little evidence against the null
+meaning for the parameters investigated here, only the X-ray derived column
+densities exhibits the highest degree of dissimilarity, but not to a significant
+level. A large part of the tension between the two samples in $N_H$ is driven by
+a small number of low column densities($\log(N_H) \lesssim 19 \mathrm{cm}^2$)
+bursts.
 
@@
 
+We therefore conclude that the sample presented here, to a high degree
+represents the intrinsic properties of the GRBs in the full \textit{Swift} sample.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
+GRB follow-up requires the immediate availability of the telescope in order to
+successfully carry out spectroscopic observations. Out of the 158 \todo{Get
+exact number} bursts in the statistical sample, 32($\sim$ 20 percent) \todo{Get
+exact number} have not been observed due to conditions, not relating the GRB or
+afterglow properties. The reasons includes unavailability of the telescope due
+to technical maintenance such af mirror re-coating, a visitor rejecting the TOO
+trigger, or bad weather at the observing site. Because this cut is irrespective
+of the GRB properties, it is not expected that this should change sample
+properties. Addressing whether this cut introduces a bias in the followed sample
+is not trivial, but in terms of intrinsic GRB properties, the sample
+distributions does not change by much, as shown in Sect. \ref{completeness}.
+\todo{Should this be quantified somehow?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A subset of all GRBs exhibit no detectable or very faint optical afterglow \citep{Groot1998,
+	Djorgovski2001, Fynbo2001}, which had been parametrized in term of "darkness"
+\citep{Jakobsson2004, VanderHorst2009}. The X-ray properties of this subset has
+previously been investigated \citep{DePasquale2003, Melandri2012}, finding a
+slightly higher X-ray luminosity compared to the optically brights bursts
+combined with similar prompt characteristics. This indicates along with
+investigations of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
+extinction of the optical afterglow is primarily driven by the presence of dust,
+local to the burst \todo{Local dust reference}.
 
@@
 	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
-	\caption{ Measurements of the hydrogen column-densities for all bursts with a clear Lyman alpha absorption system.
-In solid black is shown the spectrum with black dotted giving the corresponding 1-$\sigma$ error.
-Black dashed shows zero flux density.
-The solid red line is the absorption of column density equal to the value presented in Tab.
-\ref{HI_tab} with the 1-$\sigma$ interval shown with dashed lines.}
+	\caption{Measurements of the hydrogen column-densities for all bursts with a
+	clear Lyman alpha absorption system. In solid black is shown the spectrum with
+	black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
+	flux density. The solid red line is the absorption of column density equal to
+	the value presented in Tab. \ref{HI_tab} with the 1-$\sigma$ interval shown
+	with dashed lines.}
 	\label{fig:HI1}
@@
 Cuchiara. BAT6 (Salvaterra), TOUGH
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
-\subsection{Comparison to previous samples.}		
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-12-31 py/NH_distribution_plot.py

@@ -1,0 +1,64 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import healpy as hp
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import corner
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+def main():
+    """
+    # Script to produce NH distribution plot.
+    """
+
+    # Read DLA NH distribution from Noterdaeme et al. 2012b
+    Noter_NH, Noter_z = np.zeros(12081), np.zeros(12081)
+    count = 0
+    with open("../data/J_A+A_547_L1/table3.dat") as fp:
+        for line in fp:
+            Noter_NH[count] = float(line[82:87])
+            Noter_z[count] = float(line[51:56])
+            count += 1
+
+    # Read in GRB NH
+    burst_table = pd.read_csv("../data/Burst list - HI columns.csv")
+    name, z, NH, NHe = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values
+
+    fig, ax = pl.subplots()
+    # gs = pl.GridSpec(6, 6)
+
+    # ax = fig.add_subplot(gs[1:, :-1])
+    # ax_marg_x = fig.add_subplot(gs[0, :-1], sharex=ax)
+    # ax_marg_y = fig.add_subplot(gs[1:, -1], sharey=ax)
+    g = sns.JointGrid(x=Noter_z, y=Noter_NH, ax=ax)
+    d = sns.JointGrid(x=z, y=NH, ax=ax)
+    # corner.hist2d(Noter_z, Noter_NH, bins=100, ax=ax, smooth=2, color="#C44E52")
+    # ax.scatter(Noter_z, Noter_NH)
+    ax.scatter(z, NH, color="#4C72B0")
+    # sns.distplot(NH, hist=True, kde=False, ax=ax_marg_y, hist_kws={"bins": 20, "color": "#4C72B0", "gridsize": 200, "alpha": 0.2},
[…]

### 2016-12-31 py/Sample_completeness.py

@@ -133,7 +133,7 @@ def main():
     print(len(HI_o[~np.isnan(HI_o)]))
     l, m, h = np.percentile(np.log10(1e21*HI_o), [16, 50, 84])
     print(m, m - l, h - m)
-    exit()
+    # exit()
 
     for ax in (ax1, ax2, ax3):
         ax.spines['top'].set_visible(False)

### 2016-12-31 py/skymap.py

@@ -1,0 +1,125 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import healpy as hp
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib
+from matplotlib.ticker import FormatStrFormatter
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [9, 4.5]
+   }
+matplotlib.rcParams.update(params)
+from astropy.coordinates import SkyCoord
+from astropy.coordinates import Galactic, FK5
+from astropy.coordinates import Angle, Latitude, Longitude
+import astropy.units as u
+
+def filt_nan(input_array):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except:
+            pass
+    return np.array(holder)
+
+
+def main():
+    """
+    # Script to produce sky plots.
+    """
+    ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
+    A_Vmap = ebv_map*3.1
+    A_Vmap[A_Vmap >= 0.5] = 0
+    ax = hp.mollview(A_Vmap, max=0.5, cbar=False, title="")
+    hp.graticule()
+
+    # Overplot swift positions
+    # Read in master sample
+    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
+    name, z, observed, OA, = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values
+    # Read in observed bursts
+    burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
+    name_o = np.array([ii[3:] for ii in burst_table["GRB"].values])
+    # Read in Swift table
[…]

### 2016-12-31 py/database/build_composite.py
adding database scripts

@@ -1,0 +1,286 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import sys
+sys.path.append("/Users/jselsing/github/XSGRB_reduction_scrips/py/")
+import stitch_arms, util
+
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import glob
+import itertools
+from scipy.signal import medfilt
+from scipy import interpolate
+from NonnegMFPy import nmf
+
+def mask_intervening_systems(wl, bpmap, z, mask_width = 5.):
+    """Small script to update bad-pixel map to mask out lines in intervening systems.
+
+    Args:
+        wl (np.array): Wavelength array corresponding to bad pixel map
+        bpmap (np.array): Bap pixel map of input spectrum
+        z (float): Redshift of intervening system
+
+    Returns:
+        np.array: Updated bad pixel map, where pixels in the vicinity of lines in intervening systems has been replaced by the value 1216
+    """
+    # Read in the GRB linelist from Christensen et al. 2011
+    linelist = np.genfromtxt('static/grblines_strong.dat', dtype=None)
+    # Pick out all wavlengths and move to the redshift of the intervening system
+    waves = [kk*(1. + z) for ii, (kk, jj) in enumerate(linelist)]
+
+    # List comprehension for each line where pixels within 5 AA on each side of the line are replaced by ones.
+    # 50 km/s mask intervening systems are
+    c = 299792.458
+    mask_list = [((wl < kk + mask_width*kk / (c))  & (wl > kk - mask_width*kk / (c))).astype("int") for kk in waves]
+
+    # Sum all the individual line masks
+    mask = np.sum(np.array(mask_list), axis = 0)
+
+    # Replace all masked values by 1216
+    bpmap[mask != 0] = 1216
+
+    # Return updated bap pixel map
+    return bpmap
+
+# Execution of scipt
+if __name__ == '__main__':
+
[…]

### 2016-12-31 py/database/fits_to_ext.py

@@ -1,0 +1,58 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import numpy as np
+import glob
+from astropy.io import fits
+import matplotlib.pyplot as pl
+
+
+def main():
+    """
+    Small script to convert BinTableHDU to extensions to use with IRAF
+    """
+
+    files = glob.glob("../specdb/*")
+
+    # Loop through files
+    for ii in files:
+        outname = ii[:3] + "fits" + ii[9:]
+        bintab = fits.open(ii)
+        header = bintab[0].header
+
+        wl, flux, error, bpmap, cont, tell = bintab[1].data.field("WAVE"), bintab[1].data.field("FLUX"), bintab[1].data.field("ERR"), bintab[1].data.field("QUAL"), bintab[1].data.field("CONTINUUM"), bintab[1].data.field("TELL_CORR")
+
+        minlam, maxlam, dlam = min(wl), max(wl), np.median(np.diff(wl))
+        header["CUNIT1"] = "angstrom"
+        header["CRVAL1"] = minlam
+        header["CDELT"] = dlam
+        header["CD1_1"] = dlam
+
+        f = fits.PrimaryHDU(flux, header= header)
+        e = fits.ImageHDU(error, name ="ERR", header= header)
+        bp = fits.ImageHDU(bpmap, name ="QUAL", header= header)
+        cont = fits.ImageHDU(cont, name ="CONTINUUM", header= header)
+        tell = fits.ImageHDU(tell, name ="TELL_CORR", header= header)
+        hdulist = fits.HDUList([f, e, bp, cont, tell])
+
+        for kk in xrange(1, 5):
+            hdulist[kk].header["XTENSION"]
+            hdulist[kk].header["PCOUNT"]
+            hdulist[kk].header["GCOUNT"]
+            hdulist[kk].header["PCOUNT"]
+
+        hdulist.writeto(outname, clobber=True)
+        # exit()
+
+
+
+
+    pass
+
+
+
+
+if __name__ == '__main__':
+    main()

### 2016-12-31 py/database/get_fits.py

@@ -1,0 +1,57 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import os
+import glob
+from shutil import copyfile
+import time
+
+def main():
+    """
+    Small script to crawl through ERDA can collect burst extractions
+    """
+    dirs = [ii for ii in glob.glob("/Volumes/io.erda.dk/XSGRB/*") if "GRB" in ii.split("/")[-1]]
+    burst_names = [ii.split("/")[-1] for ii in dirs]
+    data_dir = "../specdb_input/"
+
+    # Loop through data directories and copy files to local directories
+    for ii, kk in zip(burst_names, dirs):
+        # if not "GRB160203A" in kk:
+        #     continue
+        # Create directory to contain data
+        if not os.path.exists(data_dir+ii):
+            os.makedirs(data_dir+ii)
+        # Files to be moved to local dir
+        print(kk)
+        filelist = [ll for ll in glob.glob(kk+"/reduced_data/*/*/*/*") if ("too" in ll.lower() or "rrm" in ll.lower()) and "IDP" in ll]
+        OBs = list(set([ll.split("/")[-4] for ll in filelist]))
+        lst = [ll for ll in glob.glob(kk+"/*") if "combine" in ll]
+        if len(lst) != 0:
+            OBs.append("com")
+
+        arms = list(set([ll.split("/")[-3] for ll in filelist]))
+
+        copy_filelist, filenames, count = [0]*len(OBs)*len(arms), [0]*len(OBs)*len(arms), 0
+
+        for dd in OBs:
+            for ss in arms:
+                if dd == "com":
+                    copy_filelist[count] = [ll for ll in glob.glob(kk+"/reduced_data/OB1/"+ss+"/*/*") if ("too" in ll.lower() or "rrm" in ll.lower()) and "IDP" in ll][0]
+                else:
+                    copy_filelist[count] = [ll for ll in glob.glob(kk+"/reduced_data/"+dd+"/"+ss+"/*/*") if ("too" in ll.lower() or "rrm" in ll.lower()) and "IDP" in ll][0]
+                filenames[count] = dd+ss+"_IDP.fits"
+                count += 1
+
+        # print(copy_filelist, filenames)
+        # Move files
+        for tt, pp in zip(copy_filelist, filenames):
+            if not os.path.exists(data_dir+ii+"/"+pp):
[…]

### 2016-12-31 py/database/get_fits_t.py

@@ -1,0 +1,55 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import os
+import glob
+from shutil import copyfile
+import time
+
+def main():
+    """
+    Small script to crawl through ERDA can collect burst extractions
+    """
+    dirs = [ii for ii in glob.glob("/Volumes/io.erda.dk/XSGRB/*") if "GRB" in ii.split("/")[-1]]
+    burst_names = [ii.split("/")[-1] for ii in dirs]
+    data_dir = "../specdb_input/"
+
+    # Loop through data directories and copy files to local directories
+    for ii, kk in zip(burst_names, dirs):
+
+        # if not ii == "GRB150616A_II":
+        #     continue
+
+        # Create directory to contain data
+        if not os.path.exists(data_dir+ii):
+            os.makedirs(data_dir+ii)
+        # Files to be moved to local dir
+        print(kk)
+
+        filelist = [ll for ll in glob.glob(kk+"/*/*") if ii[:-3] in ll.split("/")[-1] and ll[-4:] == "fits"]
+
+        OBs = ["OB1"]
+
+        arms = list(set([ll.split("/")[-1].split(".")[0].upper()[-3:] for ll in filelist]))
+
+        copy_filelist, filenames, count = [0]*len(OBs)*len(arms), [0]*len(OBs)*len(arms), 0
+
+        for dd in OBs:
+            for ss in arms:
+                copy_filelist[count] = [ll for ll in glob.glob(kk+"/*/*") if ss.lower() in ll and ll[-4:] == "fits"][0]
+                filenames[count] = dd+ss+"_IDP.fits"
+                count += 1
+
+        # print(copy_filelist, filenames)
+        # Move files
+        for tt, pp in zip(copy_filelist, filenames):
+            if not os.path.exists(data_dir+ii+"/"+pp):
+                print("Moving file: "+tt)
+                print("To file: "+data_dir+ii+"/"+pp)
+                copyfile(tt, data_dir+ii+"/"+pp)
+
+
+if __name__ == '__main__':
+    main()

### 2016-12-31 py/database/make_fits.py

@@ -1,0 +1,136 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+
+import glob
+import numpy as np
+import matplotlib.pyplot as pl
+import os
+
+def load_array(data_array):
+
+    # Check is reposnse has been written based on value in sixt column
+    thomas = False
+
+    if np.shape(data_array)[1] == 3:
+        thomas = True
+
+
+    if thomas:
+        return data_array[:, 0], data_array[:, 1], data_array[:, 2], np.zeros_like(data_array[:, 0])
+
+    has_resp = False
+    if np.nanmean(data_array[:, 6]) < 1e-10:
+        has_resp = True
+
+
+
+    if has_resp:
+        try:
+            tellcorr = data_array[:, 8]
+            tellcorr[np.isnan(tellcorr)] = 1
+            slitcorr = np.median(data_array[:, 7])
+            return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4], tellcorr
+        except:
+            slitcorr = np.median(data_array[:, 7])
+            return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4]
+    elif not has_resp:
+        try:
+            tellcorr = data_array[:, 7]
+            tellcorr[np.isnan(tellcorr)] = 1
+            slitcorr = np.median(data_array[:, 6])
+            return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4], tellcorr
+        except:
+            slitcorr = np.median(data_array[:, 6])
+            return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4]
+
+
+
+def main():
+    """
+    Script to combine extractions, telluric correction and continua in a single fits-file for ingestion into specdb
+    """
+
+    extract_dir = "../data/"
+    dirs = [ii for ii in glob.glob(extract_dir+"*") if "GRB" in ii.split("/")[-1]]
+    burst_names = [ii.split("/")[-1] for ii in dirs]
+
+    for kk, ll in zip(burst_names, dirs):
+        # if not "GRB160117B" in ll:
+        #     continue
+        print(kk)
+        fitsfiles = glob.glob(ll+"/*")
[…]

### 2016-12-31 py/database/normalise_spectra.py

@@ -1,0 +1,201 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import sys
+
+import stitch_arms, util
+
+from glob import glob
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy import optimize
+import os
+import xsh_norm.interactive
+from scipy.signal import medfilt
+
+from numpy.polynomial import chebyshev
+from scipy.interpolate import interp1d
+from scipy.interpolate import splrep, splev
+
+
+def main():
+    """
+    Script to loop though bursts and normalise arms
+    """
+    data_dir = "../data/"
+    bursts = glob(data_dir+"*")
+
+    # Loop through burst names
+    for ii, kk in enumerate(bursts):
+        # List of extractions in burst dir
+        extractions = [ll for ll in glob(kk+"/*") if ".dat" in ll or ".spec" in ll]
+        # print(extractions)
+
+        # Loop through extractions and normalize
+        for ll in extractions:
+
+            if "GRB" in ll:
+                print("Normalizing file:")
+                print(ll)
+                burst_name = ll.split("/")[2]
+                if "spec" in ll:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-5]
+                else:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-4]
+                print(kk+"/"+outname+"_norm.npy")
+
+                if not os.path.exists(kk+"/"+outname+"_norm.npy"):
+                    wl, flux, error, bpmap = stitch_arms.load_array(np.genfromtxt(ll))
+                    mask = ~bpmap.astype("bool")
+                    normalise = xsh_norm.interactive.xsh_norm(wl[mask], flux[mask], error[mask], bpmap, wl, flux, error, kk+"/"+outname)
+                    # change = raw_input('Change default maskparameters(y/n)? ')
+                    # if change == 'y':
+
+                    if burst_name == "GRB090809A":
+                        normalise.leg_order = 4
+                        normalise.endpoint_order = 4
+                    elif burst_name == "GRB091018A":
+                        normalise.leg_order = 4
+                        normalise.lover_mask = -5e-17
+                        normalise.exclude_width = 2
+                        normalise.sigma_mask = 10
+                    elif burst_name == "GRB100205A":
+                        normalise.lover_mask = -1e-16
+                    elif burst_name == "GRB100219A":
+                        normalise.leg_order = 1
[…]

### 2016-12-31 py/database/normalise_spectra_10.py

@@ -1,0 +1,83 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import sys
+
+import stitch_arms, util
+
+from glob import glob
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy import optimize
+import os
+import xsh_norm.automatic
+from scipy.signal import medfilt
+
+from numpy.polynomial import chebyshev
+from scipy.interpolate import interp1d
+from scipy.interpolate import splrep, splev
+
+
+def main():
+    """
+    Script to loop though bursts and normalise arms
+    """
+    data_dir = "../data/"
+    bursts = glob(data_dir+"*")
+
+    # Loop through burst names
+    for ii, kk in enumerate(bursts):
+        # List of extractions in burst dir
+        extractions = [ll for ll in glob(kk+"/*") if ".dat" in ll or ".spec" in ll]
+        # print(extractions)
+
+        # Loop through extractions and normalize
+        for ll in extractions:
+
+            if "GRB10" in ll:
+                print("Normalizing file:")
+                print(ll)
+                burst_name = ll.split("/")[2]
+                if "spec" in ll:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-5]
+                else:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-4]
+                print(kk+"/"+outname+"_norm.npy")
+
+                if not os.path.exists(kk+"/"+outname+"_norm.npy"):
+                    wl, flux, error, bpmap = stitch_arms.load_array(np.genfromtxt(ll))
+                    mask = ~bpmap.astype("bool")
+                    normalise = xsh_norm.automatic.xsh_norm(wl[mask], flux[mask], error[mask], bpmap, wl, flux, error, kk+"/"+outname)
+
+                    # if burst_name == "GRB090809A":
+                    #     normalise.leg_order = 4
+                    #     normalise.endpoint_order = 4
+                    # elif burst_name == "GRB091018A":
+                    #     normalise.leg_order = 4
+                    #     normalise.lover_mask = -5e-17
+                    #     normalise.exclude_width = 2
+                    #     normalise.sigma_mask = 10
+
+                    # normalise.leg_order = 5 #float(raw_input('Filtering Chebyshev Order (default = 3) = '))
+                    # normalise.endpoint_order = 5#float(raw_input('Order of polynomial used for placing endpoints (default = 3) = '))
[…]

### 2016-12-31 py/database/normalise_spectra_11.py

@@ -1,0 +1,83 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import sys
+
+import stitch_arms, util
+
+from glob import glob
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy import optimize
+import os
+import xsh_norm.automatic
+from scipy.signal import medfilt
+
+from numpy.polynomial import chebyshev
+from scipy.interpolate import interp1d
+from scipy.interpolate import splrep, splev
+
+
+def main():
+    """
+    Script to loop though bursts and normalise arms
+    """
+    data_dir = "../data/"
+    bursts = glob(data_dir+"*")
+
+    # Loop through burst names
+    for ii, kk in enumerate(bursts):
+        # List of extractions in burst dir
+        extractions = [ll for ll in glob(kk+"/*") if ".dat" in ll or ".spec" in ll]
+        # print(extractions)
+
+        # Loop through extractions and normalize
+        for ll in extractions:
+
+            if "GRB11" in ll:
+                print("Normalizing file:")
+                print(ll)
+                burst_name = ll.split("/")[2]
+                if "spec" in ll:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-5]
+                else:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-4]
+                print(kk+"/"+outname+"_norm.npy")
+
+                if not os.path.exists(kk+"/"+outname+"_norm.npy"):
+                    wl, flux, error, bpmap = stitch_arms.load_array(np.genfromtxt(ll))
+                    mask = ~bpmap.astype("bool")
+                    normalise = xsh_norm.automatic.xsh_norm(wl[mask], flux[mask], error[mask], bpmap, wl, flux, error, kk+"/"+outname)
+
+                    # if burst_name == "GRB090809A":
+                    #     normalise.leg_order = 4
+                    #     normalise.endpoint_order = 4
+                    # elif burst_name == "GRB091018A":
+                    #     normalise.leg_order = 4
+                    #     normalise.lover_mask = -5e-17
+                    #     normalise.exclude_width = 2
+                    #     normalise.sigma_mask = 10
+
+                    # normalise.leg_order = 5 #float(raw_input('Filtering Chebyshev Order (default = 3) = '))
+                    # normalise.endpoint_order = 5#float(raw_input('Order of polynomial used for placing endpoints (default = 3) = '))
[…]

### 2016-12-31 py/database/normalise_spectra_16.py

@@ -1,0 +1,83 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import sys
+
+import stitch_arms, util
+
+from glob import glob
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy import optimize
+import os
+import xsh_norm.automatic
+from scipy.signal import medfilt
+
+from numpy.polynomial import chebyshev
+from scipy.interpolate import interp1d
+from scipy.interpolate import splrep, splev
+
+
+def main():
+    """
+    Script to loop though bursts and normalise arms
+    """
+    data_dir = "../data/"
+    bursts = glob(data_dir+"*")
+
+    # Loop through burst names
+    for ii, kk in enumerate(bursts):
+        # List of extractions in burst dir
+        extractions = [ll for ll in glob(kk+"/*") if ".dat" in ll or ".spec" in ll]
+        # print(extractions)
+
+        # Loop through extractions and normalize
+        for ll in extractions:
+
+            if "GRB16" in ll:
+                print("Normalizing file:")
+                print(ll)
+                burst_name = ll.split("/")[2]
+                if "spec" in ll:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-5]
+                else:
+                    outname = (ll.split("/")[-1]).replace("skysub", "")[:-4]
+                print(kk+"/"+outname+"_norm.npy")
+
+                if not os.path.exists(kk+"/"+outname+"_norm.npy"):
+                    wl, flux, error, bpmap = stitch_arms.load_array(np.genfromtxt(ll))
+                    mask = ~bpmap.astype("bool")
+                    normalise = xsh_norm.automatic.xsh_norm(wl[mask], flux[mask], error[mask], bpmap, wl, flux, error, kk+"/"+outname)
+
+                    # if burst_name == "GRB090809A":
+                    #     normalise.leg_order = 4
+                    #     normalise.endpoint_order = 4
+                    # elif burst_name == "GRB091018A":
+                    #     normalise.leg_order = 4
+                    #     normalise.lover_mask = -5e-17
+                    #     normalise.exclude_width = 2
+                    #     normalise.sigma_mask = 10
+
+                    # normalise.leg_order = 5 #float(raw_input('Filtering Chebyshev Order (default = 3) = '))
+                    # normalise.endpoint_order = 5#float(raw_input('Order of polynomial used for placing endpoints (default = 3) = '))
[…]

### 2016-12-31 py/database/res_calc.py

@@ -1,0 +1,149 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import glob
+from astropy.io import fits
+from scipy.special import wofz, erf
+from scipy.optimize import curve_fit
+import seaborn; seaborn.set_style('ticks')
+import numpy as np
+import matplotlib.pyplot as pl
+from scipy import interpolate
+
+
+def voigt_base(x, amp=1, cen=0, sigma=1, gamma=0):
+    """1 dimensional voigt function.
+    see http://en.wikipedia.org/wiki/Voigt_profile
+    """
+    z = (x-cen + 1j*gamma)/ (sigma*np.sqrt(2.0))
+    return amp * wofz(z).real / (sigma*np.sqrt(2*np.pi))
+
+
+def multi_voigt(x, *params):
+    # Multiple voigt-profiles for telluric resolution estimate
+    sigma = params[0]
+    gamma = params[1]
+    c = params[2]
+    a = params[3]
+
+
+
+    multivoigt = 0
+    for ii in range(4, int(4 + len(params[4:])/2)):
+        if sigma < 0 or gamma < 0 or params[ii] > 0:
+            a = 1e10
+        multivoigt += voigt_base(x, params[ii], params[int(len(params[4:])/2) + ii], sigma, gamma)
+    return multivoigt + c + a * x
+
+# ------------------------------------------------------------------------------
+# Copied these two function from ppxf_util.py by Michelle Cappellari
+def vac_to_air(lam_vac):
+    """
+    Convert vacuum to air wavelengths using
+    equation (1) of Ciddor 1996, Applied Optics 35, 1566
+        http://dx.doi.org/10.1364/AO.35.001566
+
+    :param lam_vac - Wavelength in Angstroms
+    :return: lam_air - Wavelength in Angstroms
+
+    """
+    sigma2 = (1e4/lam_vac)**2
+    fact = 1 + 5.792105e-2/(238.0185 - sigma2) + 1.67917e-3/(57.362 - sigma2)
+
+    return lam_vac/fact
+
+# ------------------------------------------------------------------------------
+
+def air_to_vac(lam_air):
+    """
+    Convert air to vacuum wavelengths using
+    equation (1) of Ciddor 1996, Applied Optics 35, 1566
+        http://dx.doi.org/10.1364/AO.35.001566
[…]

### 2016-12-31 py/database/rescale_spec.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import numpy as np
+import glob
+
+def main():
+    files = [ii for ii in glob.glob("../data/*/*") if "spec" in ii]
+
+    for ii in files:
+        file = np.genfromtxt(ii)
+        data = zip(file[:, 0], file[:, 1]*1e-17, file[:, 2]*1e-17)
+        np.savetxt(ii, data)
+        # print(file[:, 0])
+    # print(files)
+    # pass
+
+if __name__ == '__main__':
+    main()

### 2016-12-31 py/database/timediff.py

@@ -1,0 +1,22 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import datetime
+
+
+def timediff(t1, t2):
+
+    time1 = datetime.datetime(int(t1[0:4]), int(t1[5:7]), int(t1[8:10]), hour=int(round(float(t1[11:13]))), minute=int(round(float(t1[14:16]))), second=int(round(float(t1[17:19]))))
+    time2 = datetime.datetime(int(t2[0:4]), int(t2[5:7]), int(t2[8:10]), hour=int(round(float(t2[11:13]))), minute=int(round(float(t2[14:16]))), second=int(round(float(t2[17:19]))))
+
+    dt = time2 - time1
+    return dt.total_seconds()/(3600)
+
+if __name__ == '__main__':
+
+    t_trigger = "2016:12:19T18:48:39"
+    t_obs = "2016-12-21T06:33:30.354"
+
+    print(timediff(t_trigger, t_obs))

### 2016-12-31 py/database/wave_recalib.py

@@ -1,0 +1,166 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import glob
+import itertools
+from scipy.signal import medfilt
+from scipy import interpolate
+from astropy.convolution import Gaussian1DKernel, Gaussian2DKernel, convolve
+
+
+def mask_intervening_systems(wl, bpmap, z, mask_width = 5.):
+    """Small script to update bad-pixel map to mask out lines in intervening systems.
+
+    Args:
+        wl (np.array): Wavelength array corresponding to bad pixel map
+        bpmap (np.array): Bap pixel map of input spectrum
+        z (float): Redshift of intervening system
+
+    Returns:
+        np.array: Updated bad pixel map, where pixels in the vicinity of lines in intervening systems has been replaced by the value 1216
+    """
+    # Read in the GRB linelist from Christensen et al. 2011
+    linelist = np.genfromtxt('static/grblines_strong.dat', dtype=None)
+    # Pick out all wavlengths and move to the redshift of the intervening system
+    waves = [kk*(1. + z) for ii, (kk, jj) in enumerate(linelist)]
+
+    # List comprehension for each line where pixels within 5 AA on each side of the line are replaced by ones.
+    # 50 km/s mask intervening systems are
+    c = 299792.458
+    mask_list = [((wl < kk + mask_width*kk / (c))  & (wl > kk - mask_width*kk / (c))).astype("int") for kk in waves]
+
+    # Sum all the individual line masks
+    mask = np.sum(np.array(mask_list), axis = 0)
+
+    # Replace all masked values by 1216
+    bpmap[mask != 0] = 1216
+
+    # Return updated bap pixel map
+    return bpmap
+
+
+
+def find_nearest(array, value):
+        idx = (np.abs(array-value)).argmin()
+        return idx
+
+
+def main():
+    """
+    Script to recalibrate wavelengths by cross-correlation with preliminary composite
+    """
+
+    # Get stitched spectra
[…]

### 2016-12-31 py/database/xsh_norm/automatic.py

@@ -1,0 +1,260 @@
+# -*- coding: utf-8 -*-
+
+
+# from matplotlib import rc_file
+# rc_file('/Users/jselsing/Pythonlibs/plotting/matplotlibstyle.rc')
+import matplotlib.pyplot as plt
+
+from xsh_norm import methods_auto
+from pylab import pause
+from matplotlib.lines import Line2D
+import numpy as np
+import seaborn as sns; sns.set_style('ticks')
+cmap = [sns.xkcd_rgb["denim blue"], sns.xkcd_rgb["medium green"], sns.xkcd_rgb["pale red"]]
+
+
+__all__ = ["xsh_norm"]
+
+
+def binning1d(array,bin):
+    """
+    Used to bin low S/N 1D  response data from xshooter.
+    Calculates the biweighted mean (a la done in Kriek'10). 
+    Returns binned 1dimage
+    """
+#    ;--------------
+    s=len((array))
+    outsize=s/bin
+    res = np.zeros((outsize))
+    for i in np.arange(0,s-(bin+1),bin):
+             res[((i+bin)/bin-1)] = np.sum(array[i:i+bin])/bin
+    return res
+
+
+
+def f8(seq):
+    seen = set()
+    return np.array([i for i, x in enumerate(seq) if x not in seen and not seen.add(x)])
+
+
+class xsh_norm(object):
+    """
+    A normalisation suite.
+
+    Key-bindings
+      'left-mouse-button'
+              place spline point
+
+      'enter'
+              1. sort the continuum-point-array according to the x-values
+              2. fit a spline and evaluate it in the wavelength points
+              3. plot the continuum
+      'n'
+
+              Apply normalisation
+      'w'
+
+              Write to file
+
+      'y'
+
+              Insert evenly spaced points - every 100'th pixel
+
+      'u'
+              Insert evenly spaced points - every 50'th pixel
+
+      'd'
+
+              Delete point under mouse
+
+      't'
+
+              Filter points by low-order chebyshev fitting and clipping values of high sigma iteratively until continuum is found
+
+      'm'
+
+              Mask values (spectrum - continuum fit) > 2 sigma
+
+      'i'
+              Run through the process y-t-enter-mask until the median value of the realisations converge
+    """
+
+
+    def __init__(self, w, f, e, bpmap, w_ori, f_ori, e_ori, filename):
+        self.showverts = True
+        self.epsilon = 10  # max pixel distance to count as a vertex hit
+        self.pointx = []
+        self.pointy = []
[…]

### 2017-01-03 document/XSGRB_sample.tex
NH distribution

@@
 $\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
+\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
 systems in the majority of cases \citep{Jakobsson2006, Fynbo2009}. Whereas DLAs
@@
 
+Because we only follow up targets with either a clear afterglow or counterpart, correctly associating a galaxy with a burst when there is no optical afterglow is more difficult.
+
 This should include observations where the acquisition image has been taken, but
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/NH_dist.pdf}
+	\caption{Distributions of DLA hydrogen column densities for DLAs found in
+	quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
+	is the similar distribution, only for DLAs in GRB sightlines. Values are taken
+	from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
+	with the new values presented in this sample. The marginal distributions for
+	the two distribution are also shown, where the different environments probed
+	are clearly visible in the hydrogen column densities, as previously also noted
+	in \citet{Fynbo2009}}
+	\label{fig:NH_dist}
+\end{figure}
+
 Because the locations of GRBs are associated with the most intensely starforming
@@
 \end{figure*}
+
+In a similar manner to how the X-ray derived column densities was used to assess
+the sample completeness in Sect. \ref{completeness}, we addresses the same
+question in terms of the optically derived hydrogen column densities. Using the
+compilation by \citet{Tanvir2017} we have all 81 published HI values previously,
+we can compare to the 39 new HI columns added by this sample. We note that this
+sample increase the number of optically derived hydrogen column densities with
+$\sim$ 50 percent. We compare the median, the 14th, and 86th percentiles of the
+two distributions where the sample presented here has $21.8_{0.8}^{0.3}$ and the
+rest of the litterature values has $21.5_{1.5}^{0.4}$. We see that the two
+distribution has a large degree of overlap due to the large width of the
+distributions, but find a slightly higher median value for the new sample
+presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
+relatively large evidence against the null that the two samples are drawn from
+the same underlying distribution. Because the bursts that have measurements of
+the hydrogen column density are selected sole based on our ability to infer a
+column, it is difficult to make any strong conclusions about the completeness in
+terms of gas content. We show the comparison between the two samples in Fig.
+\ref{fig:NH_dist}, where we additionally show the column-distribution for 12081
+quasar DLAs published in \citet{Noterdaeme2012b}. The ability of the GRBs to
+select the highest N$_H$ is immediately visible as is also noted previously
+\citep{Prochaska2007, Fynbo2009}. The reason for this is the selection of sight
+lines as random position of the QSO DLAs whereas the GRB forms in dense regions
+in the host. The N$_H$ distribution has also be used the infer the position of
+the gas in host \cite{Buchner2016} where a galaxy-wide gas reservoir is
+preferred.
 

### 2017-01-03 py/NH_distribution_plot.py

@@ -10,7 +10,7 @@ import healpy as hp
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
 from scipy import stats
-import corner
+
 
 params = {
    'axes.labelsize': 10,
@@ -38,27 +38,53 @@ def main():
 
     # Read in GRB NH
     burst_table = pd.read_csv("../data/Burst list - HI columns.csv")
-    name, z, NH, NHe = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values
+    name, z, NH, NHe, pub = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values, burst_table["Sample"].values
+    XSGRB = (pub == "Selsing2017")
+    print(np.sum((~XSGRB).astype("int")))
 
-    fig, ax = pl.subplots()
-    # gs = pl.GridSpec(6, 6)
 
-    # ax = fig.add_subplot(gs[1:, :-1])
-    # ax_marg_x = fig.add_subplot(gs[0, :-1], sharex=ax)
-    # ax_marg_y = fig.add_subplot(gs[1:, -1], sharey=ax)
-    g = sns.JointGrid(x=Noter_z, y=Noter_NH, ax=ax)
-    d = sns.JointGrid(x=z, y=NH, ax=ax)
-    # corner.hist2d(Noter_z, Noter_NH, bins=100, ax=ax, smooth=2, color="#C44E52")
-    # ax.scatter(Noter_z, Noter_NH)
-    ax.scatter(z, NH, color="#4C72B0")
-    # sns.distplot(NH, hist=True, kde=False, ax=ax_marg_y, hist_kws={"bins": 20, "color": "#4C72B0", "gridsize": 200, "alpha": 0.2},
-                 # vertical=True, axlabel=False)
+    g = sns.jointplot(x=Noter_z, y=Noter_NH, kind="hex", stat_func=None, xlim = (1, 7), ylim = (20, 23), marginal_kws = {"kde": False, "norm_hist": True}, label="Noterdaeme et al. 2012b", space=0)
+
+    # Plot old distribution
+    NH_old = NH[~XSGRB]
+    z_old = z[~XSGRB]
+    g.x = z_old[NH_old > 20]
+    g.y = NH_old[NH_old > 20]
[…]

### 2017-01-03 py/Sample_completeness.py

@@ -69,9 +69,9 @@ def main():
 
 
     # BAT fluence
-    sns.distplot(np.log10(1e-7*BAT), ax=ax1, kde=True)
-    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=True)
-    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=True)
+    sns.distplot(np.log10(1e-7*BAT), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
     print(len(BAT[~np.isnan(BAT)]))
     l, m, h = np.percentile(np.log10(1e-7*BAT), [16, 50, 84])
     print(m, m - l, h - m)
@@ -90,9 +90,9 @@ def main():
 
 
     # XRT flux
-    sns.distplot(np.log10(1e-11*XRT), ax=ax2, kde=True)
-    sns.distplot(np.log10(1e-11*XRT_s), ax=ax2, kde=True)
-    sns.distplot(np.log10(1e-11*XRT_o), ax=ax2, kde=True)
+    sns.distplot(np.log10(1e-11*XRT), ax=ax2, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-11*XRT_s), ax=ax2, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-11*XRT_o), ax=ax2, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
 
 
 
@@ -115,9 +115,9 @@ def main():
 
 
     # HI column
-    sns.distplot(np.log10(1e21*HI), ax=ax3, kde=True)
[…]

### 2017-01-03 document/XSGRB_sample.tex
Selection Criteria section

@@
 \usepackage{color,hyperref}
-\definecolor{linkcolor}{rgb}{0,0,0.7}
+\definecolor{linkcolor}{rgb}{0,0.3,0.7}
 \hypersetup{colorlinks=true,
@@
 \usepackage{xcolor}
-\newcommand\todo[1]{\textcolor{red}{(#1)}}
+\newcommand\todo[1]{\textbf{(#1)}}
 
@@
 \citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep{Jakobsson2006, Fynbo2009}. Whereas DLAs
+systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
 towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
@@
 5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
-	Salvaterra2012,Jakobsson2012, Perley2016}. A high redshift completeness is
+	Salvaterra2012,Jakobsson2012, Perley2016a}. A high redshift completeness is
 crucial for our understanding of the link between the number density of GRBs per
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria}
+\section{Sample selection criteria and follow-up}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 criteria on GRBs, without hampering the follow-up effort. Many natural follow-up
-restrictions exists from already, being it weather conditions, pointing
-restrictions of the telescope or poorly localized bursts as reported by the
-\textit{Swift}-telescope. To maximize the return of the follow-up campaign
-we have chosen a few selection criteria that attempts to provide an unbiased
-selection of bursts, while allowing for a high success-rate
-
+restrictions exists already, being it weather conditions, pointing restrictions
+of the telescope or poorly localized bursts as reported by the alerting
+telescope. To maximize the return of the follow-up campaign we have chosen a few
+selection criteria that attempts to provide an unbiased selection of bursts,
+while allowing for a high follow-up success rate. The importance of defining
+unbiasing selection criteria has been highlighted previously
+\citep{Salvaterra2012, Hjorth2012, Vergani2015, Perley2016a}, when trying to
+address the specific underlying distribution functions such as the redshift
+distribution, host metallicity distribution, or brightness distribution. When
+investigating a specific distribution function, a high degree of
+\textit{completeness} is desired meaning that in terms of the parameter of
+interest, the largest possible fraction of targets in the sample has
+determinations of that parameter \citep{Perley2016b}.
+
+We simultaneously want to minimize any biases against astrophysical conditions
+while at the same time maximizing likelihood of observations allowing us to
+obtain a higher degree of observing completeness. By restricting the selection
+criteria to conditions local to Galaxy and therefore irrespective of intrinsic
+GRB properties, the hope is that the sample collect represents the underlying
+distribution of GRBs in a fair way while at the same time, additional unbiasing
+cuts can be imposed that can improve the follow-up success. The selection
+criteria used here is based on previous successes \cite{Jakobsson2006b,
+	Fynbo2009, Hjorth2012}, which has shown to represent the underlying distribution
+to a high degree. We test the degree of completeness in Sect.
+\ref{completeness}.
+
+The selection criteria are defined as follows:
+
+\begin{enumerate}
+	\item GRB triggered onboard by Swift.
+	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
+	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
 	\footnote{Note that in the P84 proposal the criteria have
@@
 	defining the sample.}
-
-\begin{enumerate}
-	\item GRB triggered onboard by Swift.
+	\item No bright closeby stars.
 	\item Galactic $A_V \lesssim 0.5$ mag.
-	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
-	\item No bright closeby stars.
-
 \end{enumerate}
 
-A fraction of the bursts presented here \todo{Insert exact
-number} have already had their hosts investigated in \citet{Kruhler2015}, for
-which extractions of the hosts exist. The focus of the data presented here are
-on the afterglows themselves and in the absence of a clear afterglow, the host.
-We will not, however, investigate the hosts.
-
-We simultaneously want to minimize any biases against astrophysical conditions
-while at the same time maximizing likelihood of observations. By restricting
-the selection criteria to conditions local to Galaxy.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up procedure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Because our ability to observe GRB afterglows is strongly correlated with the
+rapidness and precision of the target positions delivered by the triggering
+facilities, by selecting only bursts that have been triggered on board the
+\textit{Swift} space telescope \cite{Gehrels2004}, by the Burst Alert Telescope
+(BAT) \citep{Barthelmy2005}, we start out with a homogeneously selected sample
+where burst characteristics are delivered immediately, allowing for an informed
+follow-up strategy. Despite the complexity of the triggering mechanism on board
+\textit{Swift} \citep{Band2006}, attempts at inferring properties of the
+underlying cosmic GRB population based on the detection thresholds and
+triggering algorithms have been made \citep{Lien2014, Graff2016} and thereby by
+restricting the follow-up effort to bursts detected by \textit{Swift}, the
+limitations of the parent sample is well studied.
+
+Because the localization accuracy of BAT is 1 - 4 arcminutes
+\citep{Barthelmy2005}, we additionally require an X-ray position from the X-ray
+Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
+\citep{Barthelmy2000} with 10 hours and to account for observing constraints on
+\textit{Swift} that XRT began observations within 10 minutes. The additional
+rapidness requirement of the XRT follow-up means that the number of detected
+X-ray afterglows for BAT triggers is 100 percent \citep{Burrows2007}. The
+\textit{Swift} constraints on the GRB follow-up means that the sample presented
+here inherits the \textit{Swift} sample properties, except the additional XRT
+constraint. Because the XRT completeness is very high\citep{Burrows2007}, this
+should change the parent sample significantly. This issue is addressed again in
+Sect. \ref{completeness}.
+
+To allow for observability we additionally require that the GRB is visible
+(duh!) from the telescope site at Cerro Paranal, Antofagasta, Región de
+Antofagasta, Chile, for a least 1 hour allowing time for the spectroscopic
+observations to completed. Since the GRB population is isotropically distributed
+seen from Earth and that GRB properties does not depend on position on the sky
+\citep{Briggs1996}, this cut does not influence our ability to sample the
+underlying GRB population. The same arguments applies to the requirement that
+there are no nearby bright stars, which is additionally supplemented by the
+requirement what the galactic dust content does not exert an extinction higher
+than $A_V \lesssim 0.5$ mag based on the extinction maps by
+\citet{Schlegel1998}. Because of  isotropy, these additional cuts should not
+influence the optical properties of the bursts themselves, only our ability to
+successfully secure observations that allows us to investigate the spectroscopic
+properties of GRBs.
+
+We address the succesfullness of these criteria and the actual follow-up effort
+in representing the full \textit{Swift} sample ion Sect. \ref{completeness}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up procedure and practicalities}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\todo{Will be written by Daniele?}
 
@@
 distribution of the bursts away from the galactic center except in the upper
-right corner, where Earth casts a shadow as seen from the telescope.
+right corner, where Earth casts a shadow on the sky as seen from the telescope.
+
+A fraction of the bursts presented here \todo{Insert exact
+number} have already had their hosts investigated in \citet{Kruhler2015}, for
+which extractions of the hosts exist. The focus of the data presented here are
+on the afterglows themselves and in the absence of a clear afterglow, the host.
+We will not, however, investigate the hosts.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
+The lower horizontal axis is in units pixels, the top axis in arc seconds.
+The red datapoints show a subsample of NIR spectra.
+		The grey line shows a linear fit to the VIS datapoints. }
+	\label{fig:res}
+\end{figure}
 
@@
 
-\begin{figure}
-	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
-The lower horizontal axis is in units pixels, the top axis in arc seconds.
-The red datapoints show a subsample of NIR spectra.
-		The grey line shows a linear fit to the VIS datapoints. }
-	\label{fig:res}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}
-	\centerline{\includegraphics[width=9cm]{figures/p_values.pdf}}
-	\caption{Relational graph shown the .}
+	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
+	\caption{Relational graph shown the respective p-values that the different
+	sample are drawn from the same underlying distribution. The arrows represent
+	the comparison with each of the p-values are listed next to. Only in the HI
+	column density distribution is there mild evidence against the null, but the
+	discrepancy is mainly driven by a relatively larger fractional contribution
+	from low-column hosts.}
 	\label{fig:p_values}
@@
 represents the intrinsic properties of the GRBs in the full \textit{Swift} sample.
+\todo{This doesn't quite feel exhaustive? Is there a better way of addressing the issue of completeness}
 

### 2017-01-03 py/Sample_completeness.py

@@ -51,17 +51,23 @@ def main():
     XRT_o = XRT[observed == "Yes"]
     HI_o = HI[observed == "Yes"]
 
+    BAT_c = BAT[observed == "No"]
+    XRT_c = XRT[observed == "No"]
+    HI_c = HI[observed == "No"]
+
     swift_table = pd.read_table("../data/grb_table_1482495106.txt", delimiter="\t", dtype=None)
     name_s, BAT_s, XRT_s, HI_s = swift_table["GRB"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values, swift_table["XRT Column Density (NH) [10^21 cm^-2]"].values
-    print(len(name_s))
-    BAT_s = filt_nan(BAT_s)
-    XRT_s = filt_nan(XRT_s)
-    HI_s = filt_nan(HI_s)
+    # Exclude sample bursts
+    idx = [ii for ii, kk in enumerate(name_s) if kk not in name]
+
+    BAT_s = filt_nan(BAT_s[idx])
+    XRT_s = filt_nan(XRT_s[idx])
+    HI_s = filt_nan(HI_s[idx])
 
 
-    BAT, BAT_o, BAT_s = BAT[(BAT > 0)], BAT_o[(BAT_o > 0)], BAT_s[(BAT_s > 0)]
-    XRT, XRT_o, XRT_s = XRT[(XRT > 0)], XRT_o[(XRT_o > 0)], XRT_s[(XRT_s > 0)]
-    HI, HI_o, HI_s = HI[(HI > 0)], HI_o[(HI_o > 0)], HI_s[(HI_s > 0)]
+    BAT_c, BAT_o, BAT_s = BAT_c[(BAT_c > 0)], BAT_o[(BAT_o > 0)], BAT_s[(BAT_s > 0)]
+    XRT_c, XRT_o, XRT_s = XRT_c[(XRT_c > 0)], XRT_o[(XRT_o > 0)], XRT_s[(XRT_s > 0)]
+    HI_c, HI_o, HI_s = HI_c[(HI_c > 0)], HI_o[(HI_o > 0)], HI_s[(HI_s > 0)]
 
     # Plot
     fig, (ax1, ax2, ax3) = pl.subplots(ncols=3)
@@ -69,11 +75,11 @@ def main():
 
 
     # BAT fluence
[…]

### 2017-01-06 document/XSGRB_sample.tex
BetaOX

@@
 
-\title{XSGRB - The X-shooter GRB afterglow sample}
+\title{XSGRB - The X-shooter GRB afterglow sample
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}
+	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}}
 
@@
 proposing these observations has measured about two thirds of the redshifts,
-mainly with FORS1/2 and X-shooter (see Fynbo et al.\ 2009, ApJS, 185, 526; 
+mainly with FORS1/2 and X-shooter \citep[see][]{Fynbo2009, Jakobsson2012, Kruhler2012}(see Fynbo et al.\ 2009, ApJS, 185, 526; 
 Jakobsson et al.\ 2012, ApJ, 752, 62; Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our
@@
 
-\begin{figure*}[!ht]
+\begin{figure*}
 	\centerline{\includegraphics{figures/skymap.pdf}}
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
@@
 our sample is biased against the overall \textit{Swift} sample in terms of
-intrinsic GRB properties. The values are queried from the online
-\textit{Swift} database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
+intrinsic GRB properties. The values are queried from the online \textit{Swift}
+database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
 Three main samples are of interest to compare to assess the completeness of the
 follow-up campaign; the full \textit{Swift} sample consisting of all the bursts
-observed by \textit{Swift} (1), the burst that fulfill the selection criteria
-imposed in Sect. \ref{samplecrit} (2), and the bursts actually followed up with
-X-shooter (3).
+observed by \textit{Swift}, excluding the bursts entering our sample (1), the
+burst that fulfill the selection criteria imposed in Sect. \ref{samplecrit}, but
+has not been followed up (2), and the bursts actually followed up with X-shooter
+(3).
 
@@
 is not trivial, but in terms of intrinsic GRB properties, the sample
-distributions does not change by much, as shown in Sect. \ref{completeness}.
+distributions does not change significantly, as shown in Sect. \ref{completeness}.
 \todo{Should this be quantified somehow?}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The nature of triggered GRBs without detected optical afterglow.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A subset of all GRBs exhibit no detectable or very faint optical afterglow \citep{Groot1998,
-	Djorgovski2001, Fynbo2001}, which had been parametrized in term of "darkness"
-\citep{Jakobsson2004, VanderHorst2009}. The X-ray properties of this subset has
-previously been investigated \citep{DePasquale2003, Melandri2012}, finding a
-slightly higher X-ray luminosity compared to the optically brights bursts
-combined with similar prompt characteristics. This indicates along with
-investigations of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
+\subsection{Sample darkness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A subset of all GRBs exhibit no detectable or very faint optical afterglow
+\citep{Groot1998, Djorgovski2001, Fynbo2001}, which had been parametrized in
+term of "darkness" \citep{Jakobsson2004, VanderHorst2009}. The X-ray properties
+of this subset has previously been investigated \citep{DePasquale2003,
+	Melandri2012}, finding a slightly higher X-ray luminosity compared to the
+optically brights bursts combined with similar prompt characteristics pointing
+towards a denser circumburst material. This indicates along with investigations
+of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
 extinction of the optical afterglow is primarily driven by the presence of dust,
-local to the burst \todo{Local dust reference}.
-
-Because we only follow up targets with either a clear afterglow or counterpart, correctly associating a galaxy with a burst when there is no optical afterglow is more difficult.
-
-This should include observations where the acquisition image has been taken, but
-observations were discontinued due to non-detection. What are the X-ray
-properties of these bursts?
+local to the burst \todo{Local dust reference}. Because the dust content of a
+galaxy is correlated with the gas amount \citep{Bohlin1978, Guver2009}, we
+expect the X-ray derived column densities for the bursts without an optical
+afterglow to be higher than the ones without.
+
+For all bursts with follow-up within 100 hours we can calculate the
+"darkness"-parameter, $\beta_{OX}$ \citep{Milvang-Jensen2012}. This required the
+simultaneous measurement of the X-ray flux density and the optical flux density
+which is in practice never possible. We use the measured acquisition camera
+magnitude reported in Tab. \ref{tab:sample_overview} to get the optical flux
+density at the beginning of the spectroscopic integration. Because we know the
+delay between the follow-up and the \textit{Swift} trigger, we can use the
+measured XRT burst lightcurve \citep{Evans2007,
+	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
+to infer the corresponding X-ray flux density. This is done by either linearly
+interpolating between temporally neighboring XRT measurements of by
+extrapolating the last few X-ray data points to the spectroscopic delay. When
+the afterglow is not detected in the acquisition camera, an upper limit of $> 24$
+mags is used, which propagates into an upper limit on $\beta_{OX}$.
+
+In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
+the one presented in \citet{Fynbo2009}. We have replaced limits with values,
+meaning that we artificially bias the distribution towards higher
+$\beta_{OX}$-values because darker bursts naturally fainter in the optical
+making the fraction of non-detections higher for lower $\beta_{OX}$. We
+marginally confirm the result by \citet{Fynbo2009} that dark bursts $\beta_{OX}
+< 0.5$) have slightly higher X-ray derived column densities, by calculating the
+median, 14th, and 86th percentile of the distributions arising by separating the
+sample at $\beta_{OX} = 0.5$. For bursts with $\beta_{OX} \geq 0.5$ we find
+$\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
+$21.8_{-0.9}^{+0.5}$, meaning that the median value for dark bursts is higher,
+but with a great deal of overlap. A 2-sided KS test fails to reject the null
+that they are drawn from the same distribution with $p = 0.43$, meaning there is
+no clear evidence for a discrepancy.
+
+\begin{figure}
+	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
+	\caption{Optical extinction relative to X-ray against X-ray derived hydrogen
+	column densities. In red is shown the sample presented in \citet{Fynbo2009}
+	where is lines indicate the kernel density estimate of the distribution where
+	limits have been replaced with values. Darker colors represented denser points.
+	The correcspoing marginal distributions are shown. In blue is the points for
+	the bursts presented here along with the marginal histograms. Limits on
+	$\beta_{OX}$ is shown by downwards facing arrows.}
+	\label{fig:betaOX}
+\end{figure}
+
+Because we only follow up targets with either a clear afterglow or counterpart,
+correctly associating a galaxy with a burst when there is no detected optical afterglow
+is more difficult, meaning that we should not have attempted follow-up for a
+larger fraction of bursts with no optical afterglow. The X-ray properties for
+the followed sample does not differ signicantly from the statistical sample, as
+shown in Sect. \ref{completeness}. Using the table maintained by
+Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How to reference
+this}, we can see how the pressence of an optical afterglow affects the
+follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do not
+have an detected optical afterglow, but this number also includes bursts where
+no optical telescopes were available, so is a lower limit. For the bursts
+entering our statistical sample, this number is 27 percent
+
+Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/NH_dist.pdf}
-	\caption{Distributions of DLA hydrogen column densities for DLAs found in
-	quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
-	is the similar distribution, only for DLAs in GRB sightlines. Values are taken
-	from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
-	with the new values presented in this sample. The marginal distributions for
-	the two distribution are also shown, where the different environments probed
-	are clearly visible in the hydrogen column densities, as previously also noted
-	in \citet{Fynbo2009}}
-	\label{fig:NH_dist}
-\end{figure}
-
 Because the locations of GRBs are associated with the most intensely starforming
@@
 \input{tables/HI_columns.tex}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Measurements of the hydrogen column-densities for all bursts with a
-	clear Lyman alpha absorption system. In solid black is shown the spectrum with
-	black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
-	flux density. The solid red line is the absorption of column density equal to
-	the value presented in Tab. \ref{HI_tab} with the 1-$\sigma$ interval shown
-	with dashed lines.}
-	\label{fig:HI1}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI2}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI3}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI4}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI5}
-\end{figure*}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/NH_dist.pdf}
+	\caption{Distributions of DLA hydrogen column densities for DLAs found in
+		quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
+		is the similar distribution, only for DLAs in GRB sightlines. Values are taken
+		from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
+		with the new values presented in this sample. The marginal distributions for
+		the two distribution are also shown, where the different environments probed
+		are clearly visible in the hydrogen column densities, as previously also noted
+		in \citet{Fynbo2009}}
+	\label{fig:NH_dist}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+This work made use of data supplied by the UK Swift Science Data Centre at the University of Leicester.
 \end{acknowledgements}
@@
 
-\newpage
+\clearpage
+
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
+	\caption{Measurements of the hydrogen column-densities for all bursts with a
+		clear Lyman alpha absorption system. In solid black is shown the spectrum with
+		black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
+		flux density. The solid red line is the absorption of column density equal to
+		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
+		with dashed lines.}
+	\label{fig:HI1}
+\end{figure*}
+\clearpage
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI2}
+\end{figure*}
+\clearpage
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI3}
+\end{figure*}
+\clearpage
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI4}
+\end{figure*}
+\clearpage
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI5}
+\end{figure*}
+\clearpage
+
 \appendix
@@
 \subsection{GRB090313 (z = 3.373)}
+The first GRB ever observed with X-shooter, during the commissioning of the
+instrument, this data formed the basis of GCN
+\#9015\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9015.gcn3}} and is published
+in \citet{DeUgartePostigo2010}. Due to the lingering brightness of GRB090313,
+6.9 ks spectroscopic integration starting 45 hours after the BAT trigger reveals
+a wealth of absorption features superposed on the afterglow continuum at a
+common redshift of $z = 3.373$. Two intervening systems at $z = 1.959$ and $z =
+1.800$ is identified based on strong \mgii-absorption. Because this burst is
+observed before the instrument is science-verified, it does not enter into the
+statistical sample.
 
 \subsection{GRB090530 (z = 1.266)}
+Observed during paranalization of the instrument, this data forms the basis of
+GCN \#15571\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15571.gcn3}}, but is not
+published elsewhere. Observations began 20.6 hours after the BAT trigger and 4.8
+ks spectroscopic integration in all three arms reveals the absorption signature
+for a host at $z = 1.266$ from the detection of \mgii, \mgi, \SIii, \feii,
+\aliii. Because this burst is observed before the instrument is
+science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB090809 (z = 2.737)}
+Observed during the first science verification period and was used for GCN
+\#9761\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9761.gcn3}} and is
+additionally used as the basis for the master thesis by \'Asa Sk\'ulad\'ottir
+(2010). 7.2 ks integration starting 10.2 hours after the GRB trigger notice
+yields the clear afterglow continuum in all arms from with Lyman Limit located
+in the beginning of the UVB coverage. The simultaneous detection of absorption
+lines identified as \lya, \SIii, \oi, \SIi$^*$, \SIiv, \civ, \feii, \alii,
+\aliii and \mgii at $z = 2.737$ sets it as the redshift of the GRB. Because this burst is observed before the instrument is
+science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB090926 (z = 2.106)}
+Starting during the second science verification period, this dataset forms the
+basis of GCN \#9942\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9942.gcn3}} and
+is additionally published in \citet{DElia2010}. Spectroscopic integration
+started 22 hours after the BAT trigger and from the acquistion camera the
+optical afterglow has R = 17.9 mag (vega) at the beginning of the observations
+which causes a very continuum to be seen in all arms. An absorption trough due
+to \lya is clearly visible along with numerous metal resonance lines \civ,
+\SIii, \SIii$^*$ \feii, \mgii, all at $z = 2.106$, marks this as the redshift of
+the GRB. Because this burst is observed before the instrument is
+science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB091018 (z = 0.971)}
-
-\subsection{GRB091127 (z = 0.49)}
+The first burst observed during normal operation after science verification was
+completed and there is the first burst that enter the statistical sample. This
+data is the basis for GCN \#
+10042\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10042.gcn3}} and is published
+in \citet{Wiersema2012}. With a bright afterglow and a rapid follow-up, this
+spectrum is of pristine quality. The afterglow continuum is bright throughout
+all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
+\mnii, \mgii, \mgi, and \caii al located at $z = 0.971$, setting is as the
+redshift of the host.
+
+\subsection{GRB091127 (z = 0.490)}
+Obtained 4 days after the burst trigger, this data forms the basis for GCN \#
+10233\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10233.gcn3}} and is published
+in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the S/N of
+the afterglow continuum is poor especially in the UVB arm, why no clear
+absorption lines are detected against the afterglow continuum, although see
+\citet{Vergani2011} which report a tentative detection of \mgii. Emission lines
+from the underlying host is clearly visible with lines from \oii, \hb, \oiii,
+and \ha~all at $z = 0.490$. This bursts is additionally associated with
+SN2009nz.
 
@@
 trigger and has a total exposure time of 4.8 ks. Absorption features, including
-those of \lya, from a multitude of ions are detected against the afterglow
+those of \lya~and from a multitude of ions are detected against the afterglow
 continuum at $z = 4.667$. Additionally, absorption from an intervening system

### 2017-01-06 py/DarknessHI.py

@@ -1,0 +1,149 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+
+def iqr(a):
+    """Calculate the IQR for an array of numbers."""
+    a = np.asarray(a)
+    q1 = stats.scoreatpercentile(a, 25)
+    q3 = stats.scoreatpercentile(a, 75)
+    return q3 - q1
+
+
+def _freedman_diaconis_bins(a):
+    """Calculate number of hist bins using Freedman-Diaconis rule."""
+    # From http://stats.stackexchange.com/questions/798/
+    a = np.asarray(a)
+    h = 2 * iqr(a) / (len(a) ** (1 / 3))
+    # fall back to sqrt(a) bins if iqr is 0
+    if h == 0:
+        return int(np.sqrt(a.size))
+    else:
+        return int(np.ceil((a.max() - a.min()) / h))
+
+
+def filt_nan(input_array):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except:
+            holder.append(1e99)
+    return np.array(holder)
+
+
+def main():
+    """
+    # Script to produce NH distribution plots for dark bursts.
+    """
+    # Read in burst list
+    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
[…]

### 2017-01-06 py/betaOX.py

@@ -1,0 +1,165 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+
+def iqr(a):
+    """Calculate the IQR for an array of numbers."""
+    a = np.asarray(a)
+    q1 = stats.scoreatpercentile(a, 25)
+    q3 = stats.scoreatpercentile(a, 75)
+    return q3 - q1
+
+
+def _freedman_diaconis_bins(a):
+    """Calculate number of hist bins using Freedman-Diaconis rule."""
+    # From http://stats.stackexchange.com/questions/798/
+    a = np.asarray(a)
+    h = 2 * iqr(a) / (len(a) ** (1 / 3))
+    # fall back to sqrt(a) bins if iqr is 0
+    if h == 0:
+        return int(np.sqrt(a.size))
+    else:
+        return int(np.ceil((a.max() - a.min()) / h))
+
+
+def filt_nan(input_array, fill_value=np.nan):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except ValueError:
+            try:
+                holder.append(float(ii.split("=")[1]))
+            except IndexError:
+                try:
+                    holder.append(float(ii))
+                except:
+                    holder.append(fill_value)
+
+    return np.array(holder)
+
+
+def main():
+    """
+    # Script to produce beta OX.
+    """
+    # Read in burst list
[…]

### 2017-01-09 document/XSGRB_sample.tex
Beginning introduction

@@
 \abstract{
-The {\it Swift} satellite allows us to use gamma-ray bursts (GRBs) to peer into 
-the hearts of star forming galaxies through cosmic time. Our open
-collaboration, representing most of the active ESO member researchers in this
-field, seeks to build a public legacy sample of GRB X-shooter spectroscopy
-while {\it Swift} continues to fly. We propose to continue our programme to target
-all suitably observable GRB afterglows (up to 15 bursts per semester), with the
-primary goal of producing a well-defined, homogeneous, statistically useful
-sample. To date, our spectroscopy covers a redshift range from 0.059 to about
-8, with more than 20 robust metallicity measurements from absorption lines (over 
-the redshift range 1.7--5.9) and 4 secure detections of H$_2$ or CH molecular
-absorption. Such information is extremely difficult to obtain by other means.
-In terms of studying the spread and redshift evolution in gas-phase properties, the
-sample is still limited by low-number statistics. 
+In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
+afterglows that have \textit{ever} been observed at the VLT/X-shooter. This
+totals 137 spectroscopic observation of 118 individual bursts, of which
+redshifts have been secured for 91 \% covering a redshift range from 0.059 to
+$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
+has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
+of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
+We here provide a public release of all the spectra, including continuum
+estimates and telluric corrections. We provide an assessment of the degree of
+completeness with respect to the parent GRB population, in terms of the X-ray
+properties of the bursts in the sample and find that the sample presented here
+can not be significantly distinguished from the \textit{Swift} sample.
+For all bursts (39) for which it is possible, we provide a measurement of the
+optically derived hydrogen column density, increasing the total number of
+published HI column densities by $\sim$ 50 \% and characterize the distribution
+compared with literature values.
+We constrain the fraction of dark burst to be 27\% - 42\% and confirm previous
+results that higher optical darkness has increased X-ray excess
+absorption.
+%This dataset provides a unique possibility to study the ISM across cosmic time.
 }
@@
 
-Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ QSOs have 5 systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
-towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
-atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
-\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
-star-forming galaxies over the full history of cosmic star formation from $z
-\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
-	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we can hence characterize the properties of
-star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities, etc.. This is, however,
-only possible as long as there are satellites in orbit that rapidly and
-accurately locate GRBs. The currently operating \textit{Swift} satellite,
-launched in 2004 and still fully-functioning, allows for very efficient
-follow-up observations of GRBs due to its unprecedented rate, speed and
-precision of localisations.
-
-There are several unanswered but fundamental questions that must be addressed in
-order to exploit the full potential of GRBs as cosmological probes. More than
-50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and
-5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
-	Salvaterra2012,Jakobsson2012, Perley2016a}. A high redshift completeness is
-crucial for our understanding of the link between the number density of GRBs per
-unit redshift and the global star-formation history of the Universe, as measured
-by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
-GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
-galaxies at the highest redshifts and unsurpassed in providing detailed
-abundance information via absorption line spectroscopy \citep{Tanvir2012,
-	McGuire2016}.
-
-\smallskip
-
-%The sample of short GRBs with redshifts remains too small and too
-%incomplete to really
-%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
-%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
-%enlarge as far as possible the number of known redshifts for this class of
-%sources.
-
-From March 2005 to March 2016 there have been about 350 {\it Swift} bursts
-complying with our sample selection criteria (see ``Immediate objective''), and
-about half of them have measured redshifts. Among the latter subset, the team
-proposing these observations has measured about two thirds of the redshifts,
-mainly with FORS1/2 and X-shooter \citep[see][]{Fynbo2009, Jakobsson2012, Kruhler2012}(see Fynbo et al.\ 2009, ApJS, 185, 526; 
-Jakobsson et al.\ 2012, ApJ, 752, 62; Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our
-current aim is to build a sample superior to our previous low-resolution survey
-(Fynbo et al.\ 2009, ApJS, 185, 526), both in terms of quantity and  quality of
-the spectra. Our program started as guaranteed time observations during periods
-84-91, and we have continued in open time since then.
-
-%In P84 we started the X-shooter legacy survey in GTO time. Our aim is to  build
-%a sample of GRBs superior to our previous  low-resolution  survey, Fynbo et
-% al.\
-%2009, ApJS, 185, 526) both in terms of quantity and  quality of the spectra.
-% The
-%last period of GTO time on the survey was P90 (plus 6 hr of leftover GTO time
-%in  P91) and for P92 a proposal like this was  accepted for continuation of the
-%program in open time (concerning P93 see box 8).
-
-\smallskip
-
-X-shooter is in many ways the ideal GRB follow-up instrument and indeed GRB
-follow-up was one of the primary science cases behind the instrument design and
-implementation. Our program secures general purpose GRB afterglow spectroscopic
-follow-up that adds strong legacy value to the \textit{Swift} GRB sample. Due to
-the wide wavelength range of X-shooter with the same observation cover molecular
-H$_2$ absorption near the atmospheric cut-off and all the strong emission lines
-from the host in the NIR arm (e.g., Friis et al., 2015, MNRAS, 451, 167). In
-general, the wide wavelength coverage ensures that we always have features on
-which to base a redshift measurement as long as the afterglow is brighter than
-about 23 mag in either the $R$- or $z$-band. Frequently, emission lines are also
-detected from the underlying host, which also provide further information such
-as SFR and metallicity (the top right panel in Fig.~1 shows an example). Only
-for 7 out of more than 70 secured spectra could we not measure a redshift. With
-the X-shooter survey we provide {\bf metallicity measurements} for about 30\%
-(Voigt-profile fits) of the $z>1.7$ events. So far we have measured
-metallicities for more than 20 GRB afterglows with X-shooter. With the wide
-wavelength coverage of X-shooter we can study important chemical species as
-Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$ elements over a much wider
-redshift range than what is possible with other instruments. As an example, we
-have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A at $z=4.669$
-(Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for GRB\,111008A at
-$z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$ for the
-$z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139). Reconciling the
-abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs in
-particular, and old stars in the Local Group is an important long-term goal (see
-also Sparre et al.\ 2014, ApJ, 785, 150). Metallicities are also measured from
-host emission lines (Kr{\"u}hler et al.\ 2015, A\&A, 581, A125). GRB
-spectroscopy also allows us to determine the dust content of their environments,
-both through analysis of the depletion pattern and through measurement of the
-associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). This allows us to
-quantify the dust-to-metals ratio and its evolution with redshift (e.g., De Cia
-et al.\ 2013, A\&A, 560, 88; Zafar \& Watson 2013, A\&A, 560, 26). %Moreover,
-% GRB
-%afterglows allow investigation of random foreground absorbers, complementing
-%QSO sightline studies (e.g., Schulze et
-%al.\ 2012, A\&A, 546, 20).
-
-%In Fynbo et al. (2009, ApJS, 185,
-%526), we compare the host-galaxy HI
-%column density and metallicity (most of which has been measured by our group)
-%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
-%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
-%column density and metallicity for GRB and QSO sightlines shows that the total
-%cross-section for environments like those of GRBs must be less than a few
-%percent of the total cross-section for QSO-DLAs, and suggests that GRB
-%sightlines probe regions with high gas density.
-
-%The issue of production and {\bf release of Lyman continuum radiation} is
-%currently not fully understood, but the evidence suggests that star-forming
-%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
-%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
-%X-shooter spectroscopy will allow major progress due to the excellent blue
-%sensitivity.
-
-%GRB afterglows can probe dust in high-redshift star-forming galaxies
-%including exploring the {\bf 2175~\AA{} extinction bump}
-%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
-%732, 143).
-%Due to the broad wavelength  coverage we can do
-%this more accurately and over a broader range of redshifts with X-shooter.
-
-\smallskip
-
-We will also determine the frequency and properties of {\bf molecular
-	absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
-the process of star formation, but prior to our program H$_2$ had been detected
-just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
-in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
-found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
-2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). We are currently
-analysing more of our spectra for less obvious molecular absorption and we
-expect to find more (a dedicated sample paper is addressing this issue).
-
-\vspace{0.2cm} A natural question to ask is: {\bf how long should this work
-	continue?} Our view is that we need to keep observing the afterglows as long as
-we have \textit{Swift} in operation. Also note that the program is still
-producing many papers and provides data for many theses (Box 9 and 10).
-\textit{Swift} is currently funded until 2018, but is likely to get more
-extensions given its overwhelming success. As mentioned above GRBs allow us to
-probe star-forming galaxies that are almost impossible to study in other ways
-both in terms of redshifts, galaxy luminosity function, and regions within
-galaxies. After 7 periods we have secured seven spectra of $z>4$ GRBs, of which
-three were of sufficient quality to allow abundance measurements (Th\"one et
-al.\ 2013, Sparre et al.\ 2014, Hartoog et al.\ 2015). GRBs offer the only way
-to derive chemical abundances for the gas phase of central, actively
-star-forming regions of high-$z$ galaxies. The program also maintains a very
-high discovery potential where we occasionally find something completely
-unexpected that provides interesting clues to puzzles in other fields, e.g.
-extinction of type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of
-these spectra are like precious jewels – it is a type of observation that can
-never be repeated and a class of sightlines that can only be studied while we
-have operating GRB satellites.
-
-It is also worth adding that we have build up a rather unique team spread over
-Europe from Granada to Reykjavik, which by now has reached a point where the
-distribution of night shifts, the scientific exploitation of the data is
-efficient and where we are open to all new members who wish to participate. As
-mentioned all data are public immediately.
-
-{\bf For all of these reasons, we need to keep building up the sample of GRB
-	afterglow spectra now as we may have to wait many years before a mission like
-	\textit{Swift} becomes available again.}
-
-A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark
-	bursts}'', e.g., Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012,
-MNRAS, 421, 1265). Some of these are at the highest redshifts ($z > 6$) and
-their observer-frame optical emission is absorbed by the IGM. The majority,
-however, suffer from large dust obscuration (e.g., Perley et al. 2009, AJ, 138,
-1690; Greiner et al.\ 2011, A\&A, 526, 30). Identifying such GRBs is important
-for constraining the fraction of obscured star formation. In both cases, NIR
-emission is expected. X-shooter can adequately study these objects, provided
-that a NIR counterpart is timely identified, for which we have the dedicated
-HAWKI run D.
-
-The detection of {\bf absorption line variability} can reveal the burst
-influence on the surrounding medium and in turn the absorber distance from the
-burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
-2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\ 2012,
-A\&A, 545, 64; Hartoog et al.\ 2013). {\bf Short GRBs} originate in a
-substantially different environment compared to long GRBs. Short GRBs may be
-related to the merging of compact binaries and the coalescence time can be long
-enough to allow the progenitor system to move far away from the star formation
-site (Belczynski et al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum
-with a sufficient signal-to-noise ratio of a short GRB afterglow has been
-secured. A knowledge of the redshift distribution of short bursts is of key
-importance for the next generation of gravitational wave experiments, as they
-are the likely EM counterparts to their primary targets.
+For a brief moment, the energy output of a $\gamma$-ray burst (GRB) can rival
+that of the stellar light integrated over the entirety of the observable
+universe \citep{Kumar2015}. Bright across the entire electromagnetic spectrum,
+the following afterglow makes possible the exploration of the intervening space,
+otherwise dark. Since the discovery of the first $\gamma$-ray counterparts
+nearly two decades ago\citep{Costa1997a, VanParadijs1997, Djorgovski1997,
+	Frail1997}, significant advances in the understanding of the underlying
+phenomenology and potential usage has been made.
+
+Categorized on the basis of their duration and spectral hardness, two main classes of GRBs exist, long and short \citep{Kouveliotou1993}.
+Now thoroughly established, the connection between long GRBs and supernovae (SNe) \citep{Galama1998, Hjorth2003a, Stanek2003, Woosley2006c} and the association of the location of long GRBs with the most star-forming regions in their host galaxies \citep{Fruchter2006, Svensson2010a}, make a stellar origin a likely explanation for the progenitor system.
+The seeming metal aversion \citep{Graham2013, Kruhler2015, Perley2016b} in selection of host at low redshift along with theoretical single-progenitor models preferring more pristine starting conditions \citep{Woosley2006d}, paints a picture where the core of a rapidly spinning, massive star undergoes gravitational collapse, although a certain degree of binarity \citep{Fryer2005, Podsiadlowski2010} in the progenitor population is required to explain the host metallicity distribution to super solar values \citep{Niino2011}.
+The driving engine in short GRBs have been established as the merger of two compact objects \citep{Metzger2010} through the detection of an associated kilonova \citep{Tanvir2013, Berger2013a}.
+
+ see \citet{Levan2016} for recent review.
+
+%Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+%towards about 10$^5$ QSOs have 5 systems with
+%$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
+%\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
+%systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
+%towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
+%atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
+%\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
+%star-forming galaxies over the full history of cosmic star formation from $z
+%\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
+%	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
+%spectrum from X-rays to the sub-mm) we can hence characterize the properties of
+%star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
+%molecular contents, ISM temperatures, UV-flux densities, etc.. 
+%
+%
+%This is, however,
+%only possible as long as there are satellites in orbit that rapidly and
+%accurately locate GRBs. The currently operating \textit{Swift} satellite,
+%launched in 2004 and still fully-functioning, allows for very efficient
+%follow-up observations of GRBs due to its unprecedented rate, speed and
+%precision of localisations.
+%
+%There are several unanswered but fundamental questions that must be addressed in
+%order to exploit the full potential of GRBs as cosmological probes. More than
+%50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and
+%5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
+%	Salvaterra2012,Jakobsson2012, Perley2016a}. A high redshift completeness is
+%crucial for our understanding of the link between the number density of GRBs per
+%unit redshift and the global star-formation history of the Universe, as measured
+%by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
+%GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
+%galaxies at the highest redshifts and unsurpassed in providing detailed
+%abundance information via absorption line spectroscopy \citep{Tanvir2012,
+%	McGuire2016}.
+%
+%\smallskip
+%
+%%The sample of short GRBs with redshifts remains too small and too
+%%incomplete to really
+%%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
+%%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
+%%enlarge as far as possible the number of known redshifts for this class of
+%%sources.
+%
+%From March 2005 to March 2016 there have been about 350 {\it Swift} bursts
+%complying with our sample selection criteria, and
+%about half of them have measured redshifts. Among the latter subset, the team
+%proposing these observations has measured about two thirds of the redshifts,
+%mainly with FORS1/2 and X-shooter \citep[see][]{Fynbo2009, Jakobsson2012, Kruhler2012}. Our
+%current aim is to build a sample superior to our previous low-resolution survey
+%\citep{Fynbo2009}, both in terms of quantity and  quality of
+%the spectra. 
+%
+%Due to
+%the wide wavelength range of X-shooter with the same observation cover molecular
+%H$_2$ absorption near the atmospheric cut-off and all the strong emission lines
+%from the host in the NIR arm \citep[e.g.,][]{Friis2015}. In
+%general, the wide wavelength coverage ensures that we always have features on
+%which to base a redshift measurement as long as the afterglow is brighter than
+%about 23 mag in either the $R$- or $z$-band. Frequently, emission lines are also
+%detected from the underlying host, which also provide further information such
+%as SFR and metallicity (the top right panel in Fig.~1 shows an example). Only
+%for 7 out of more than 70 secured spectra could we not measure a redshift. With
+%the X-shooter survey we provide {\bf metallicity measurements} for about 30\%
+%(Voigt-profile fits) of the $z>1.7$ events. So far we have measured
+%metallicities for more than 20 GRB afterglows with X-shooter. With the wide
+%wavelength coverage of X-shooter we can study important chemical species as
+%Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$ elements over a much wider
+%redshift range than what is possible with other instruments. As an example, we
+%have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A at $z=4.669$
+%(Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for GRB\,111008A at
+%$z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$ for the
+%$z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139). Reconciling the
+%abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs in
+%particular, and old stars in the Local Group is an important long-term goal (see
+%also Sparre et al.\ 2014, ApJ, 785, 150). Metallicities are also measured from
+%host emission lines (Kr{\"u}hler et al.\ 2015, A\&A, 581, A125). GRB
+%spectroscopy also allows us to determine the dust content of their environments,
+%both through analysis of the depletion pattern and through measurement of the
+%associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). This allows us to
+%quantify the dust-to-metals ratio and its evolution with redshift (e.g., De Cia
+%et al.\ 2013, A\&A, 560, 88; Zafar \& Watson 2013, A\&A, 560, 26). %Moreover,
+%% GRB
+%%afterglows allow investigation of random foreground absorbers, complementing
+%%QSO sightline studies (e.g., Schulze et
+%%al.\ 2012, A\&A, 546, 20).
+%
+%%In Fynbo et al. (2009, ApJS, 185,
+%%526), we compare the host-galaxy HI
+%%column density and metallicity (most of which has been measured by our group)
+%%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
+%%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
+%%column density and metallicity for GRB and QSO sightlines shows that the total
+%%cross-section for environments like those of GRBs must be less than a few
+%%percent of the total cross-section for QSO-DLAs, and suggests that GRB
+%%sightlines probe regions with high gas density.
+%
+%%The issue of production and {\bf release of Lyman continuum radiation} is
+%%currently not fully understood, but the evidence suggests that star-forming
+%%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
+%%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
+%%X-shooter spectroscopy will allow major progress due to the excellent blue
+%%sensitivity.
+%
+%%GRB afterglows can probe dust in high-redshift star-forming galaxies
+%%including exploring the {\bf 2175~\AA{} extinction bump}
+%%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
+%%732, 143).
+%%Due to the broad wavelength  coverage we can do
+%%this more accurately and over a broader range of redshifts with X-shooter.
+%
+%\smallskip
+%
+%We will also determine the frequency and properties of {\bf molecular
+%	absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
+%the process of star formation, but prior to our program H$_2$ had been detected
+%just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
+%in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
+%found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
+%2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). We are currently
+%analysing more of our spectra for less obvious molecular absorption and we
+%expect to find more (a dedicated sample paper is addressing this issue).
+%
+%\vspace{0.2cm} A natural question to ask is: {\bf how long should this work
+%	continue?} Our view is that we need to keep observing the afterglows as long as
+%we have \textit{Swift} in operation. Also note that the program is still
+%producing many papers and provides data for many theses (Box 9 and 10).
+%\textit{Swift} is currently funded until 2018, but is likely to get more
+%extensions given its overwhelming success. As mentioned above GRBs allow us to
+%probe star-forming galaxies that are almost impossible to study in other ways
+%both in terms of redshifts, galaxy luminosity function, and regions within
+%galaxies. After 7 periods we have secured seven spectra of $z>4$ GRBs, of which
+%three were of sufficient quality to allow abundance measurements (Th\"one et
+%al.\ 2013, Sparre et al.\ 2014, Hartoog et al.\ 2015). GRBs offer the only way
+%to derive chemical abundances for the gas phase of central, actively
+%star-forming regions of high-$z$ galaxies. The program also maintains a very
+%high discovery potential where we occasionally find something completely
+%unexpected that provides interesting clues to puzzles in other fields, e.g.
+%extinction of type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of
+%these spectra are like precious jewels – it is a type of observation that can
+%never be repeated and a class of sightlines that can only be studied while we
+%have operating GRB satellites.
+%
+%It is also worth adding that we have build up a rather unique team spread over
+%Europe from Granada to Reykjavik, which by now has reached a point where the
+%distribution of night shifts, the scientific exploitation of the data is
+%efficient and where we are open to all new members who wish to participate. As
+%mentioned all data are public immediately.
+%
+%{\bf For all of these reasons, we need to keep building up the sample of GRB
+%	afterglow spectra now as we may have to wait many years before a mission like
+%	\textit{Swift} becomes available again.}
+%
+%A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark
+%	bursts}'', e.g., Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012,
+%MNRAS, 421, 1265). Some of these are at the highest redshifts ($z > 6$) and
+%their observer-frame optical emission is absorbed by the IGM. The majority,
+%however, suffer from large dust obscuration (e.g., Perley et al. 2009, AJ, 138,
+%1690; Greiner et al.\ 2011, A\&A, 526, 30). Identifying such GRBs is important
+%for constraining the fraction of obscured star formation. In both cases, NIR
+%emission is expected. X-shooter can adequately study these objects, provided
+%that a NIR counterpart is timely identified, for which we have the dedicated
+%HAWKI run D.
+%
+%The detection of {\bf absorption line variability} can reveal the burst
+%influence on the surrounding medium and in turn the absorber distance from the
+%burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
+%2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\ 2012,
+%A\&A, 545, 64; Hartoog et al.\ 2013). {\bf Short GRBs} originate in a
+%substantially different environment compared to long GRBs. Short GRBs may be
+%related to the merging of compact binaries and the coalescence time can be long
+%enough to allow the progenitor system to move far away from the star formation
+%site (Belczynski et al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum
+%with a sufficient signal-to-noise ratio of a short GRB afterglow has been
+%secured. A knowledge of the redshift distribution of short bursts is of key
+%importance for the next generation of gravitational wave experiments, as they
+%are the likely EM counterparts to their primary targets.
 
@@
 Because we only follow up targets with either a clear afterglow or counterpart,
-correctly associating a galaxy with a burst when there is no detected optical afterglow
-is more difficult, meaning that we should not have attempted follow-up for a
-larger fraction of bursts with no optical afterglow. The X-ray properties for
-the followed sample does not differ signicantly from the statistical sample, as
-shown in Sect. \ref{completeness}. Using the table maintained by
-Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How to reference
-this}, we can see how the pressence of an optical afterglow affects the
-follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do not
-have an detected optical afterglow, but this number also includes bursts where
-no optical telescopes were available, so is a lower limit. For the bursts
-entering our statistical sample, this number is 27 percent
-
-Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts.
+correctly associating a galaxy with a burst when there is no detected optical
+afterglow is more difficult, see \citet{Levesque2010} and \citet{Perley2017} for
+a recent example. Because the presence of optical transient is highly dependent
+on the "darkness", we expect that the $\beta_{OX}$ sample is biased against dark
+bursts. The larger relative fraction of upper limits for increased darkness is
+visible both in the sample presented here and the one in \citet{Fynbo2009}. The
+X-ray properties for the followed sample does not differ signicantly from the
+statistical sample, as shown in Sect. \ref{completeness}. Using the table
+maintained by Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How
+to reference this}, we can see how the presence of an optical afterglow affects
+the follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do
+not have an detected optical afterglow, but this number also includes bursts
+where no optical telescopes were available, so is a lower limit. For the bursts
+entering our statistical sample, this number is $\sim$ 27 percent, close to the
+maximum fraction of dark bursts \citep{Melandri2012}. Of the subsample (3) that
+have actually been attempted, this number is $\sim$ 18 percent, confirming that
+we are indeed biased against bursts with an optical afterglow in the
+spectroscopic sample.
+
+Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts. \todo{Daniele, help.}
 

### 2017-01-09 py/DarknessHI.py

@@ -94,7 +94,7 @@ def main():
     XRT_not = np.log10(1e-11*filt_nan(XRT_s[idx_not]))
     HI_not = np.log10(1e21*filt_nan(HI_s[idx_not]))
 
-    idx = [ii for ii, kk in enumerate(name_g) if kk in name]
+    idx = [ii for ii, kk in enumerate(name_g) if kk in name_o]
     name_g_ot = name_g[idx][(OT[idx] == "y") & (inst[idx] == "Swift")]
     name_g_not = name_g[idx][~(OT[idx] == "y") & (inst[idx] == "Swift")]
 

### 2017-01-09 py/NH_distribution_plot.py

@@ -41,6 +41,8 @@ def main():
     name, z, NH, NHe, pub = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values, burst_table["Sample"].values
     XSGRB = (pub == "Selsing2017")
     print(np.sum((~XSGRB).astype("int")))
+    print(len(name))
+    exit()
 
 
     g = sns.jointplot(x=Noter_z, y=Noter_NH, kind="hex", stat_func=None, xlim = (1, 7), ylim = (20, 23), marginal_kws = {"kde": False, "norm_hist": True}, label="Noterdaeme et al. 2012b", space=0)

### 2017-01-09 py/betaOX.py

@@ -150,7 +150,7 @@ def main():
     #Fraction of dark bursts
     print(len(HI[betaOX[idx_2] < 0.5])/(len(HI[betaOX[idx_2] < 0.5]) + len(HI[betaOX[idx_2] >= 0.5])))
     print(stats.ks_2samp(HI[betaOX[idx_2] < 0.5], HI[betaOX[idx_2] >= 0.5]))
-    # exit()
+    exit()
     # print(len(f_HI))
     # l, m, h = np.percentile(f_HI, [16, 50, 84])
     # print(m, m - l, h - m)

### 2017-01-10 document/XSGRB_sample.tex
Spec plot and intro

@@
 Now thoroughly established, the connection between long GRBs and supernovae (SNe) \citep{Galama1998, Hjorth2003a, Stanek2003, Woosley2006c} and the association of the location of long GRBs with the most star-forming regions in their host galaxies \citep{Fruchter2006, Svensson2010a}, make a stellar origin a likely explanation for the progenitor system.
-The seeming metal aversion \citep{Graham2013, Kruhler2015, Perley2016b} in selection of host at low redshift along with theoretical single-progenitor models preferring more pristine starting conditions \citep{Woosley2006d}, paints a picture where the core of a rapidly spinning, massive star undergoes gravitational collapse, although a certain degree of binarity \citep{Fryer2005, Podsiadlowski2010} in the progenitor population is required to explain the host metallicity distribution to super solar values \citep{Niino2011}.
-The driving engine in short GRBs have been established as the merger of two compact objects \citep{Metzger2010} through the detection of an associated kilonova \citep{Tanvir2013, Berger2013a}.
-
- see \citet{Levan2016} for recent review.
+The seeming metal aversion \citep{Graham2013, Kruhler2015, Perley2016b} in selection of host at low redshift along with theoretical single-progenitor models preferring more pristine starting conditions \citep{Woosley2006d}, paints a picture where the core of a rapidly spinning, massive star undergoes gravitational collapse, although a certain degree of binarity \citep{Fryer2005, Podsiadlowski2010} in the progenitor population is required to explain the host metallicity distribution to super solar values \citep{Niino2011}, see \citet{Levan2016} for recent review.
+The driving engine in short GRBs has been established as the merger of two compact objects \citep{Metzger2010} through the detection of an associated kilonova \citep{Tanvir2013a, Berger2013a}.
+With the advent of the first gravitational wave (GW) detection \cite{Abbott2016a}, the electromagnetic counterpart in the form of a short
+
+GRBs as cosmological probes
+
+Do GRBs trace star formation
+
+Afterglow spectroscopy
+
+Complete samples
+
+Content of paper
 
@@
 
+\begin{figure*}
+	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
+	\caption{Telluric corrected, normalized spectrum of GRB121024A at $z = 2.300$,
+	meant to illustrate the typical data quality. The acquisition magnitude is r =
+	20, meaning it is in the brighter end of the sample presented here, but not the
+	brightest. The spectrum is rich in absorption, including molecular absorption
+	of $H_2$ with the absorption due to \lya is visible at $\sim$ 4000 $\AA$.
+	Additionally, three intervening systems are seen in this sightline  This
+	spectrum has been analyzed in detail in \citet{Friis2015}. }
+	\label{fig:spectrum}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Science data products} \label{products}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results} \label{results}

### 2017-01-10 py/spec_plot.py

@@ -1,0 +1,63 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats, signal
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [8.27, 11.7]
+   }
+mpl.rcParams.update(params)
+
+def main():
+    """
+    # Script to produce plot of GRB121024A
+    """
+    # Load data from file
+    data = np.genfromtxt("../data/GRB121024A_OB1_stitched_spectrum.dat")
+    wl = data[:, 0]
+    flux = data[:, 1]
+    error = data[:, 2]
+    n_elem = len(wl)
+
+    # Make plot
+    n_axes = 7
+    fig, (ax1, ax2, ax3, ax4, ax5, ax6, ax7) = pl.subplots(n_axes, 1)
+
+    for ii, ax in enumerate([ax1, ax2, ax3, ax4, ax5, ax6, ax7]):
+
+        cut = slice(int(np.round(n_elem*(ii/n_axes))), int(np.round(n_elem*(ii/n_axes + 1/n_axes))))
+        ax.plot(wl[cut], flux[cut], color="black", lw=0.5, linestyle="steps-mid", rasterized=True, zorder=2)
+        ax.plot(wl[cut][::10], signal.medfilt(error[cut][::10], 11), color=sns.color_palette()[0], alpha=0.9, zorder=1)
+        ax.axhline(1, linestyle="dashed", color=sns.color_palette()[2], alpha=0.9, zorder=10)
+        ax.axhline(0, linestyle="dotted", color="black", alpha=0.7)
+        ax.set_xlim(min(wl[cut]), max(wl[cut]))
+        ax.set_ylim(-0.2, 2)
+        ax.tick_params(axis='x', direction='in')
+        # ax.spines['top'].set_visible(False)
+        ax.yaxis.set_ticks_position('left')
+        ax.xaxis.set_ticks_position('bottom')
+    # Save figure for tex
[…]

### 2017-01-15 document/XSGRB_sample.tex
Notes on bursts

@@
 
-\author{
-	J.~Selsing\inst{1}, 
-	D.~Malesani\inst{1}, 
-	T.~Kr\"{u}hler\inst{1}, 
-	J.~P.~U. Fynbo\inst{1}, 
-	P.~Goldoni\inst{14}, 
-	A.~de~Ugarte~Postigo\inst{11}, 
-	J.~Japelj\inst{20},
-	P.~D'Avanzo,
-	Z.~Cano,
-	S.~Covino\inst{10}, 
-	V.~D'Elia\inst{7, 12}, 
-	H.~Flores,
-	O.~E.~Hartoog\inst{6},
-	J.~Hjorth\inst{1}, 
-	P.~Jakobsson\inst{5}, 
-	A.~Levan,
-	A.~Melandri,
-	S.~Piranomonte \inst{7},
-	R.~S\'anchez-Ram\'irez\inst{11},
-	S.~Schulze\inst{17, 18}, 
-	N.~R.~Tanvir\inst{19},
-	C.~Th{\"o}ne,
-	S.~D.~Vergani\inst{7, 8},
-	P.~M.~Vreeswijk\inst{3}, 
-	D.~J.~Watson\inst{1},
-	K.~Wiersema\inst{19},
-	D.~Xu\inst{1}
-	L.~Christensen\inst{1},
-	A.~De~Cia\inst{3}, 
-	L.~Kaper\inst{6}, 
-	L.~A.~Antonelli,
-	F.~Fiore,
-	A.~Gomboc,
-	P.~Groot,
-	F.~Hammer,
-	C.~Ledoux\inst{2}, 
-	E.~Maiorano,
-	B.~Milvang-Jensen\inst{1}, 
-	E.~Palazzi,
-	E.~Pian,
-	J.~Schaye,
-	G.~Tagliaferri\inst{7},
-	R.~A.~M.~J.~Wijers\inst{6}
-}
+\input{authorlist.tex}
+
+\date{Received/ accepted}
 
 \authorrunning{Selsing et al.}
-
-\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
-\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
-\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
-\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
-\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
-\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
-\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
-\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
-\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
-\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
-\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
-\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
-\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
-\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/
-	IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et L
-	eonie Duquet, 75205 Paris Cedex 13, France
-\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachs
-	tra\ss e, 85748 Garching, Germany
-\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3,
-	20126, Milano, Italy
-\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astro
-	nom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
-\and Millennium Center for Supernova Science
-\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
-\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI}
-
-\date{Received/ accepted}
 
@@
 In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
-afterglows that have \textit{ever} been observed at the VLT/X-shooter. This
+afterglows that have \textit{ever} been observed at the VLT/X-shooter up to the end of 2016. This
 totals 137 spectroscopic observation of 118 individual bursts, of which
@@
 compared with literature values.
-We constrain the fraction of dark burst to be 27\% - 42\% and confirm previous
-results that higher optical darkness has increased X-ray excess
-absorption.
+We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and tentatively confirm previous
+results that larger optical darkness leads to increased X-ray absorption.
 %This dataset provides a unique possibility to study the ISM across cosmic time.
@@
 nearly two decades ago\citep{Costa1997a, VanParadijs1997, Djorgovski1997,
-	Frail1997}, significant advances in the understanding of the underlying
+	Frail1997, Wijers1998}, significant advances in the understanding of the underlying
 phenomenology and potential usage has been made.
@@
 
+Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+towards about 10$^5$ QSOs have 5 systems with
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
+\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
+towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
+atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
+\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
+star-forming galaxies over the full history of cosmic star formation from $z
+\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
+	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we can hence characterize the properties of
+star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
+molecular contents, ISM temperatures, UV-flux densities.
+
 Content of paper
 
-%Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-%towards about 10$^5$ QSOs have 5 systems with
-%$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-%\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-%systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
-%towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
-%atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
-%\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
-%star-forming galaxies over the full history of cosmic star formation from $z
-%\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
-%	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
-%spectrum from X-rays to the sub-mm) we can hence characterize the properties of
-%star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
-%molecular contents, ISM temperatures, UV-flux densities, etc.. 
-%
-%
-%This is, however,
-%only possible as long as there are satellites in orbit that rapidly and
-%accurately locate GRBs. The currently operating \textit{Swift} satellite,
-%launched in 2004 and still fully-functioning, allows for very efficient
-%follow-up observations of GRBs due to its unprecedented rate, speed and
-%precision of localisations.
-%
 %There are several unanswered but fundamental questions that must be addressed in
@@
 In this section we describe how the final data products are produced and
-subsequently post-processed. All post-processing scripts developed for this
-dataset are made publicly available at
-\url{https://github.com/jselsing/XSGRB_reduction_scripts}, along with
+subsequently post-processed. $\sim$ 30 percent of the data released here have
+been reduced in a manner similar to what is described in \citet{Kruhler2015}.
+The two reduction schemes have been tested and are in good agreement. All
+post-processing scripts developed for this dataset are made publicly available
+at \url{https://github.com/jselsing/XSGRB_reduction_scripts}, along with
 instructions of use.
@@
 	brightest. The spectrum is rich in absorption, including molecular absorption
-	of $H_2$ with the absorption due to \lya is visible at $\sim$ 4000 $\AA$.
+	of $H_2$. The absorption trough visible at $\sim$ 4000 $\AA$ is due to \lya in
+	the host. We have overplot the most prominent lines seen in GRB afterglows from
+	\citet{Christensen2011a}. The continuum estimate is shown in dashed red.
 	Additionally, three intervening systems are seen in this sightline  This
-	spectrum has been analyzed in detail in \citet{Friis2015}. }
+	spectrum has been analyzed in detail in \citet{Friis2015}.}
 	\label{fig:spectrum}
@@
 \todo{Get number} bursts has been triggered by \textit{Swift} in the period
-since the commissioning of VLT/X-shooter in \todo{Insert date}. This sample
+since the commissioning of VLT/X-shooter in November, 2009. This sample
 constitutes the \textit{statistical sample} on which we will try to address the
@@
 term of "darkness" \citep{Jakobsson2004, VanderHorst2009}. The X-ray properties
-of this subset has previously been investigated \citep{DePasquale2003,
+of this subset has previously been investigated \citep{DePasquale2003, Fynbo2009,
 	Melandri2012}, finding a slightly higher X-ray luminosity compared to the
@@
 For all bursts with follow-up within 100 hours we can calculate the
-"darkness"-parameter, $\beta_{OX}$ \citep{Milvang-Jensen2012}. This required the
+"darkness"-parameter, $\beta_{OX}$ \citep{Jakobsson2004, Milvang-Jensen2012}. This required the
 simultaneous measurement of the X-ray flux density and the optical flux density
@@
 
-Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts. \todo{Daniele, help.}
+Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts.
+\todo{Hmm, how do I finish this?}
 
@@
 
-Because the redshift completeness of our statistical sample is  54 percent
+Because the redshift completeness of our statistical sample is just 54 percent
 \todo{Get number}, making inference of the underlying redshift distribution of
-GRBs is difficult. Because a large fraction of the bursts for which we were
+GRBs on this ample is difficult. Because a large fraction of the bursts for which we were
 unable to secure a redshift are missing due to terrestrial reasons, see Sect.
@@
 telescope availability, the redshift completeness increases to 63 percent
-\todo{Get number}. As is shown in Sect. \ref{timing}, the redshift completeness
-decreases with increasing follow-up delay, we can additionally increase the
-redshift completeness by imposing cuts on the follow-up delay. Our ability to
-rapidly follow up a burst is also not dependent on the burst properties and
-therefore we can additionally cut delays longer than 24 hours (all burst in the
-sample are in principle observable within 24 hours).
+\todo{Get number}.
+
+%As is shown in Sect. \ref{timing}, the redshift completeness
+%decreases with increasing follow-up delay, we can additionally increase the
+%redshift completeness by imposing cuts on the follow-up delay. Our ability to
+%rapidly follow up a burst is also not dependent on the burst properties and
+%therefore we can additionally cut delays longer than 24 hours (all burst in the
+%sample are in principle observable within 24 hours).
 
@@
 
-\subsection{GRB110709B (z=na)}
+\subsection{GRB110709B (z=2.1092 (NEW!!!!))}
+This is a late-time observation (> 1 year) and has previously been used in
+\citet{Perley2016a}. In this reduction of the 7.2 ks spectroscopic integration,
+the tentative detection of \oiii~reported in \citet{Perley2016a} is confirmed
+along with low-significance detection of \ha~at the end of the spectral
+coverage, both at a consistent redshift, $z=2.1092$, securing it as the redshift
+of the GRB.
 
@@
 
-\subsection{GRB111117A (z=1.3?)}
+\subsection{GRB111117A (z=2.211)}
 

### 2017-01-15 document/authorlist.tex
Notes on bursts

\author{J.~Selsing\inst{1}
\and XSGRB collaboration
%\and D.~Malesani\inst{1}
%\and T.~Kr\"{u}hler\inst{1}
%\and J.~P.~U. Fynbo\inst{1}
%\and P.~Goldoni\inst{14}
%\and A.~de~Ugarte~Postigo\inst{11}
%\and J.~Japelj\inst{20}
%\and P.~D'Avanzo
%\and Z.~Cano
%\and S.~Covino\inst{10}
%\and V.~D'Elia\inst{712}
%\and H.~Flores
%\and O.~E.~Hartoog\inst{6}
%\and J.~Hjorth\inst{1}
%\and P.~Jakobsson\inst{5}
%\and A.~Levan
%\and A.~Melandri
%\and S.~Piranomonte \inst{7}
%\and R.~S\'anchez-Ram\'irez\inst{11}
%\and S.~Schulze\inst{1718}
%\and N.~R.~Tanvir\inst{19}
%\and C.~Th{\"o}ne
%\and S.~D.~Vergani\inst{78}
%\and P.~M.~Vreeswijk\inst{3}
%\and D.~J.~Watson\inst{1}
%\and K.~Wiersema\inst{19}
%\and D.~Xu\inst{1}
%\and L.~Christensen\inst{1}
%\and A.~De~Cia\inst{3}
%\and L.~Kaper\inst{6}
%\and L.~A.~Antonelli
%\and F.~Fiore
%\and A.~Gomboc
%\and P.~Groot
%\and F.~Hammer
%\and C.~Ledoux\inst{2}
%\and E.~Maiorano
%\and B.~Milvang-Jensen\inst{1}
%\and E.~Palazzi
%\and E.~Pian
%\and J.~Schaye
%\and G.~Tagliaferri\inst{7},
%\and R.~A.~M.~J.~Wijers\inst{6}
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
%\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
%\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
%\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
%\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
%\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
%\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
%\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
%\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
%\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
%\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
%\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
%\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et Leonie Duquet, 75205 Paris Cedex 13, France
%\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3, 20126, Milano, Italy
%\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astronom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
%\and Millennium Center for Supernova Science
%\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
%\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI
}

### 2017-01-15 py/Sample_completeness.py

@@ -9,8 +9,9 @@ import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
 from scipy import stats
-import matplotlib
+import matplotlib as mpl
 from matplotlib.ticker import FormatStrFormatter
+
 params = {
    'axes.labelsize': 10,
    'font.size': 10,
@@ -20,7 +21,7 @@ params = {
    'text.usetex': False,
    'figure.figsize': [9, 4.5]
    }
-matplotlib.rcParams.update(params)
+mpl.rcParams.update(params)
 
 def filt_nan(input_array):
     """

### 2017-01-15 py/follow_up_timing.py

@@ -6,20 +6,20 @@ from __future__ import division, print_function
 from astropy.io import fits
 import pandas as pd
 import matplotlib.pyplot as pl
-import seaborn; seaborn.set_style('ticks')
+import seaborn as sns; sns.set_style('ticks')
 import numpy as np
-import matplotlib
+import matplotlib as mpl
 from matplotlib.ticker import FormatStrFormatter
 params = {
-   'axes.labelsize': 10,
-   'font.size': 10,
-   'legend.fontsize': 10,
-   'xtick.labelsize': 10,
-   'ytick.labelsize': 10,
+   'axes.labelsize': 12,
+   'font.size': 12,
+   'legend.fontsize': 12,
+   'xtick.labelsize': 12,
+   'ytick.labelsize': 12,
    'text.usetex': False,
-   'figure.figsize': [4.5, 4.5]
+   'figure.figsize': [6, 6]
    }
-matplotlib.rcParams.update(params)
+mpl.rcParams.update(params)
 
 
 
@@ -34,6 +34,12 @@ def main():
     burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
 
     name, z, delay, mag = burst_table["GRB"].values, burst_table["z"].values, burst_table["Follow-up delay"].values, burst_table["Acquisition mag"].values
+    for ii, ll in enumerate(mag):
+        # print(ll)
+        if "R" in str(ll):
+            print(ll)
+            mag[ii] = float(ll.split("=")[1]) + 0.21
+
 
     idx_limits = ~(mag == ">24")
     magnondet = 24 * (~idx_limits).astype("int")
@@ -42,15 +48,19 @@ def main():
     sorted_delay = delay[delay_sort]
     sorted_z = z[delay_sort]
     fractional_completeness = (1 - np.cumsum(np.isnan(sorted_z).astype("int"))/len(sorted_z))*100
-    print(len(sorted_z), np.sum(np.isnan(sorted_z).astype("int")))
-    print(fractional_completeness[-1])
-    exit()
+    # print(len(sorted_z), np.sum(np.isnan(sorted_z).astype("int")))
+    # print(fractional_completeness[-1])
+    # exit()
     idx_hasz = ~np.isnan(z)
 
     # Plot
     fig, ax1 = pl.subplots()
 
-    cmap = pl.get_cmap("plasma")
[…]

### 2017-01-15 py/spec_plot.py

@@ -8,7 +8,7 @@ import pandas as pd
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
-from scipy import stats, signal
+from scipy import stats, signal, interpolate
 import matplotlib as mpl
 
 params = {
@@ -33,6 +33,20 @@ def main():
     error = data[:, 2]
     n_elem = len(wl)
 
+    # Get linelist
+    import lineid_plot
+    fit_line_positions = np.genfromtxt("../data/grblines_Christensen2011.dat", dtype=None)
+    # fit_line_positions = np.genfromtxt("../data/grblines.dat", dtype=None)
+    linelist = []
+    linenames = []
+    for n in fit_line_positions:
+        linelist.append(float(n[0]))
+        linenames.append(str(n[1]))
+    linelist = np.array(linelist)
+    linenames = np.array(linenames)
+    # Redshifts of systems
+    z = [2.300]
+    ls = ["dashed"]
     # Make plot
     n_axes = 7
     fig, (ax1, ax2, ax3, ax4, ax5, ax6, ax7) = pl.subplots(n_axes, 1)
@@ -41,6 +55,15 @@ def main():
 
         cut = slice(int(np.round(n_elem*(ii/n_axes))), int(np.round(n_elem*(ii/n_axes + 1/n_axes))))
         ax.plot(wl[cut], flux[cut], color="black", lw=0.5, linestyle="steps-mid", rasterized=True, zorder=2)
+        interp = interpolate.interp1d(wl[cut], signal.medfilt(flux[cut], 11))
+        for oo, pp in enumerate(z):
+            idx = [ii for ii, kk in enumerate(linelist*(1 + pp)) if (kk > wl[cut][0] and kk < wl[cut][-1])]
+            # for aa, ss in zip(linelist[idx]*(1 + pp), linenames[idx]):
+                # ax.axvline(aa, ymin=interp(aa)/2, linestyle=ls[oo], color="black", alpha=0.7, lw=0.5)
+                # ax.text(aa, 1.5, ss)
+            lineid_plot.plot_line_ids(wl[cut], flux[cut], linelist[idx]*(1 + pp), linenames[idx], arrow_tip=1.2, box_loc=1.6, ax = ax)
+
+
         ax.plot(wl[cut][::10], signal.medfilt(error[cut][::10], 11), color=sns.color_palette()[0], alpha=0.9, zorder=1)
[…]

### 2017-09-18 document/XSGRB_sample.tex
Trimming sections

@@
 
-A fraction of the bursts presented here \todo{Insert exact
-number} have already had their hosts investigated in \citet{Kruhler2015}, for
-which extractions of the hosts exist. The focus of the data presented here are
-on the afterglows themselves and in the absence of a clear afterglow, the host.
-We will not, however, investigate the hosts.
+%A fraction of the bursts presented here \todo{Insert exact
+%number} have already had their hosts investigated in \citet{Kruhler2015}, for
+%which extractions of the hosts exist. The focus of the data presented here are
+%on the afterglows themselves and in the absence of a clear afterglow, the host.
+%We will not, however, investigate the hosts.
 
@@
 atmosphere, where the atmospheric content and conditions make an imprint on the
-received spectrum. These telluric features can be calculated in a multitude of
-ways and we employ a prioritized list of methods here, depending on the
+received spectrum. These telluric features can be corrected for in a multitude
+of ways. We employ a prioritized list of methods here, depending on the
 availability of the different method. Since the observation are often taken at
@@
 telluric features can directly be fit with an atmospheric model
-\citep{Smette2015, Kausch2015}, which can then be used to correct for the absorption. The accuracy of the correction depends on the 
-The requirement here is that the afterglow continuum spectrum has a
-signal-to-noise higher than a value of 10.
+\citep{Smette2015, Kausch2015}, which can then be used to correct for the
+absorption. The accuracy of the correction depends on the S/N for the target
+spectrum, where we have chosen the requirement that the afterglow continuum
+spectrum has a median signal-to-noise higher than a value of 10.
 
@@
 
-%As is shown in Sect. \ref{timing}, the redshift completeness
-%decreases with increasing follow-up delay, we can additionally increase the
-%redshift completeness by imposing cuts on the follow-up delay. Our ability to
-%rapidly follow up a burst is also not dependent on the burst properties and
-%therefore we can additionally cut delays longer than 24 hours (all burst in the
-%sample are in principle observable within 24 hours).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Redshifts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+As is shown in Sect. \ref{timing}, the redshift completeness
+decreases with increasing follow-up delay, we can additionally increase the
+redshift completeness by imposing cuts on the follow-up delay. Our ability to
+rapidly follow up a burst is also not dependent on the burst properties and
+therefore we can additionally cut delays longer than 24 hours (all burst in the
+sample are in principle observable within 24 hours). \todo{Unfinished section}
 
@@
 GRBs. Additionally only the brightest GRBs are seen above redshift $z \gtrsim
-1$. \todo{Additional text}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Afterglow-based redshifts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Emissionline-based redshifts}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Redshift completeness
+1$. \todo{Unfinished section}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{X-ray column densities vs. optically derived ones}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{X-ray column densities vs. optically derived ones (Remove section)}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
 %
@@
 %\end{figure}	
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Why is the HI column truncated?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
+%	
+%	
+%	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Why is the HI column truncated? (Remove section)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
 Number of bursts with detected H2 and their HI column density.

### 2017-09-18 py/betaOX.py

@@ -86,7 +86,9 @@ def main():
 
     betaOX = betaOX[~mask][::-1]
     name = name[~mask][::-1]
-
+    for ii, kk in enumerate(betaOX):
+        print(name[ii], kk)
+    exit()
     fynbo_table = np.genfromtxt("../data/Fynbo2009betaOX.dat", dtype=None)
     f_name = fynbo_table[:, 0]
     fynbo_table[:, 1] = [ii.replace('<', '') for ii in fynbo_table[:, 1]]

### 2017-10-24 document/XSGRB_sample.tex
Adding ontro

--- 2017-01-10 document/XSGRB_sample.tex
@@
 \usepackage{color,hyperref}
-\definecolor{linkcolor}{rgb}{0,0.3,0.7}
+\definecolor{linkcolor}{rgb}{0,0,0.7}
 \hypersetup{colorlinks=true,
@@
 \usepackage{xcolor}
-\newcommand\todo[1]{\textbf{(#1)}}
+\newcommand\todo[1]{\textcolor{red}{(#1)}}
 
@@
 
-\title{XSGRB - The X-shooter GRB afterglow sample
+\title{\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample}
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}}
-
-\titlerunning{The X-shooter sample of GRBs}
+	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}
+
+\titlerunning{XSGRB sample}
 
@@
 	J.~Selsing\inst{1}, 
+	T.~Kr\"{u}hler\inst{1}, 
+	P.~Goldoni\inst{14}, 
 	D.~Malesani\inst{1}, 
-	T.~Kr\"{u}hler\inst{1}, 
 	J.~P.~U. Fynbo\inst{1}, 
-	P.~Goldoni\inst{14}, 
 	A.~de~Ugarte~Postigo\inst{11}, 
@@
 \abstract{
-In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
-afterglows that have \textit{ever} been observed at the VLT/X-shooter. This
-totals 137 spectroscopic observation of 118 individual bursts, of which
-redshifts have been secured for 91 \% covering a redshift range from 0.059 to
-$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
-has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
-of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
-We here provide a public release of all the spectra, including continuum
-estimates and telluric corrections. We provide an assessment of the degree of
-completeness with respect to the parent GRB population, in terms of the X-ray
-properties of the bursts in the sample and find that the sample presented here
-can not be significantly distinguished from the \textit{Swift} sample.
-For all bursts (39) for which it is possible, we provide a measurement of the
-optically derived hydrogen column density, increasing the total number of
-published HI column densities by $\sim$ 50 \% and characterize the distribution
-compared with literature values.
-We constrain the fraction of dark burst to be 27\% - 42\% and confirm previous
-results that higher optical darkness has increased X-ray excess
-absorption.
-%This dataset provides a unique possibility to study the ISM across cosmic time.
+The {\it Swift} satellite allows us to use gamma-ray bursts (GRBs) to peer into 
+the hearts of star forming galaxies through cosmic time. Our open
+collaboration, representing most of the active ESO member researchers in this
+field, seeks to build a public legacy sample of GRB X-shooter spectroscopy
+while {\it Swift} continues to fly. We propose to continue our programme to target
+all suitably observable GRB afterglows (up to 15 bursts per semester), with the
+primary goal of producing a well-defined, homogeneous, statistically useful
+sample. To date, our spectroscopy covers a redshift range from 0.059 to about
+8, with more than 20 robust metallicity measurements from absorption lines (over 
+the redshift range 1.7--5.9) and 4 secure detections of H$_2$ or CH molecular
+absorption. Such information is extremely difficult to obtain by other means.
+In terms of studying the spread and redshift evolution in gas-phase properties, the
+sample is still limited by low-number statistics. 
 }
@@
 
-For a brief moment, the energy output of a $\gamma$-ray burst (GRB) can rival
-that of the stellar light integrated over the entirety of the observable
-universe \citep{Kumar2015}. Bright across the entire electromagnetic spectrum,
-the following afterglow makes possible the exploration of the intervening space,
-otherwise dark. Since the discovery of the first $\gamma$-ray counterparts
-nearly two decades ago\citep{Costa1997a, VanParadijs1997, Djorgovski1997,
-	Frail1997}, significant advances in the understanding of the underlying
-phenomenology and potential usage has been made.
-
-Categorized on the basis of their duration and spectral hardness, two main classes of GRBs exist, long and short \citep{Kouveliotou1993}.
-Now thoroughly established, the connection between long GRBs and supernovae (SNe) \citep{Galama1998, Hjorth2003a, Stanek2003, Woosley2006c} and the association of the location of long GRBs with the most star-forming regions in their host galaxies \citep{Fruchter2006, Svensson2010a}, make a stellar origin a likely explanation for the progenitor system.
-The seeming metal aversion \citep{Graham2013, Kruhler2015, Perley2016b} in selection of host at low redshift along with theoretical single-progenitor models preferring more pristine starting conditions \citep{Woosley2006d}, paints a picture where the core of a rapidly spinning, massive star undergoes gravitational collapse, although a certain degree of binarity \citep{Fryer2005, Podsiadlowski2010} in the progenitor population is required to explain the host metallicity distribution to super solar values \citep{Niino2011}, see \citet{Levan2016} for recent review.
-The driving engine in short GRBs has been established as the merger of two compact objects \citep{Metzger2010} through the detection of an associated kilonova \citep{Tanvir2013a, Berger2013a}.
-With the advent of the first gravitational wave (GW) detection \cite{Abbott2016a}, the electromagnetic counterpart in the form of a short
-
-GRBs as cosmological probes
-
-Do GRBs trace star formation
-
-Afterglow spectroscopy
-
-Complete samples
-
-Content of paper
-
-%Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-%towards about 10$^5$ QSOs have 5 systems with
-%$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-%\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-%systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
-%towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
-%atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
-%\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
-%star-forming galaxies over the full history of cosmic star formation from $z
-%\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
-%	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
-%spectrum from X-rays to the sub-mm) we can hence characterize the properties of
-%star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
-%molecular contents, ISM temperatures, UV-flux densities, etc.. 
-%
-%
-%This is, however,
-%only possible as long as there are satellites in orbit that rapidly and
-%accurately locate GRBs. The currently operating \textit{Swift} satellite,
-%launched in 2004 and still fully-functioning, allows for very efficient
-%follow-up observations of GRBs due to its unprecedented rate, speed and
-%precision of localisations.
-%
-%There are several unanswered but fundamental questions that must be addressed in
-%order to exploit the full potential of GRBs as cosmological probes. More than
-%50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and
-%5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
-%	Salvaterra2012,Jakobsson2012, Perley2016a}. A high redshift completeness is
-%crucial for our understanding of the link between the number density of GRBs per
-%unit redshift and the global star-formation history of the Universe, as measured
-%by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
-%GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
-%galaxies at the highest redshifts and unsurpassed in providing detailed
-%abundance information via absorption line spectroscopy \citep{Tanvir2012,
-%	McGuire2016}.
-%
-%\smallskip
-%
-%%The sample of short GRBs with redshifts remains too small and too
-%%incomplete to really
-%%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
-%%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
-%%enlarge as far as possible the number of known redshifts for this class of
-%%sources.
-%
-%From March 2005 to March 2016 there have been about 350 {\it Swift} bursts
-%complying with our sample selection criteria, and
-%about half of them have measured redshifts. Among the latter subset, the team
-%proposing these observations has measured about two thirds of the redshifts,
-%mainly with FORS1/2 and X-shooter \citep[see][]{Fynbo2009, Jakobsson2012, Kruhler2012}. Our
-%current aim is to build a sample superior to our previous low-resolution survey
-%\citep{Fynbo2009}, both in terms of quantity and  quality of
-%the spectra. 
-%
-%Due to
-%the wide wavelength range of X-shooter with the same observation cover molecular
-%H$_2$ absorption near the atmospheric cut-off and all the strong emission lines
-%from the host in the NIR arm \citep[e.g.,][]{Friis2015}. In
-%general, the wide wavelength coverage ensures that we always have features on
-%which to base a redshift measurement as long as the afterglow is brighter than
-%about 23 mag in either the $R$- or $z$-band. Frequently, emission lines are also
-%detected from the underlying host, which also provide further information such
-%as SFR and metallicity (the top right panel in Fig.~1 shows an example). Only
-%for 7 out of more than 70 secured spectra could we not measure a redshift. With
-%the X-shooter survey we provide {\bf metallicity measurements} for about 30\%
-%(Voigt-profile fits) of the $z>1.7$ events. So far we have measured
-%metallicities for more than 20 GRB afterglows with X-shooter. With the wide
-%wavelength coverage of X-shooter we can study important chemical species as
-%Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$ elements over a much wider
-%redshift range than what is possible with other instruments. As an example, we
-%have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A at $z=4.669$
-%(Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for GRB\,111008A at
-%$z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$ for the
-%$z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139). Reconciling the
-%abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs in
-%particular, and old stars in the Local Group is an important long-term goal (see
-%also Sparre et al.\ 2014, ApJ, 785, 150). Metallicities are also measured from
-%host emission lines (Kr{\"u}hler et al.\ 2015, A\&A, 581, A125). GRB
-%spectroscopy also allows us to determine the dust content of their environments,
-%both through analysis of the depletion pattern and through measurement of the
-%associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). This allows us to
-%quantify the dust-to-metals ratio and its evolution with redshift (e.g., De Cia
-%et al.\ 2013, A\&A, 560, 88; Zafar \& Watson 2013, A\&A, 560, 26). %Moreover,
-%% GRB
-%%afterglows allow investigation of random foreground absorbers, complementing
-%%QSO sightline studies (e.g., Schulze et
-%%al.\ 2012, A\&A, 546, 20).
-%
-%%In Fynbo et al. (2009, ApJS, 185,
-%%526), we compare the host-galaxy HI
-%%column density and metallicity (most of which has been measured by our group)
-%%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
-%%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
-%%column density and metallicity for GRB and QSO sightlines shows that the total
-%%cross-section for environments like those of GRBs must be less than a few
-%%percent of the total cross-section for QSO-DLAs, and suggests that GRB
-%%sightlines probe regions with high gas density.
-%
-%%The issue of production and {\bf release of Lyman continuum radiation} is
-%%currently not fully understood, but the evidence suggests that star-forming
-%%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
-%%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
-%%X-shooter spectroscopy will allow major progress due to the excellent blue
-%%sensitivity.
-%
-%%GRB afterglows can probe dust in high-redshift star-forming galaxies
-%%including exploring the {\bf 2175~\AA{} extinction bump}
-%%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
-%%732, 143).
-%%Due to the broad wavelength  coverage we can do
-%%this more accurately and over a broader range of redshifts with X-shooter.
-%
-%\smallskip
-%
-%We will also determine the frequency and properties of {\bf molecular
-%	absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
-%the process of star formation, but prior to our program H$_2$ had been detected
-%just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
-%in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
-%found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
-%2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). We are currently
-%analysing more of our spectra for less obvious molecular absorption and we
-%expect to find more (a dedicated sample paper is addressing this issue).
-%
-%\vspace{0.2cm} A natural question to ask is: {\bf how long should this work
-%	continue?} Our view is that we need to keep observing the afterglows as long as
-%we have \textit{Swift} in operation. Also note that the program is still
-%producing many papers and provides data for many theses (Box 9 and 10).
-%\textit{Swift} is currently funded until 2018, but is likely to get more
-%extensions given its overwhelming success. As mentioned above GRBs allow us to
-%probe star-forming galaxies that are almost impossible to study in other ways
-%both in terms of redshifts, galaxy luminosity function, and regions within
-%galaxies. After 7 periods we have secured seven spectra of $z>4$ GRBs, of which
-%three were of sufficient quality to allow abundance measurements (Th\"one et
-%al.\ 2013, Sparre et al.\ 2014, Hartoog et al.\ 2015). GRBs offer the only way
-%to derive chemical abundances for the gas phase of central, actively
-%star-forming regions of high-$z$ galaxies. The program also maintains a very
-%high discovery potential where we occasionally find something completely
-%unexpected that provides interesting clues to puzzles in other fields, e.g.
-%extinction of type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of
-%these spectra are like precious jewels – it is a type of observation that can
-%never be repeated and a class of sightlines that can only be studied while we
-%have operating GRB satellites.
-%
-%It is also worth adding that we have build up a rather unique team spread over
-%Europe from Granada to Reykjavik, which by now has reached a point where the
-%distribution of night shifts, the scientific exploitation of the data is
-%efficient and where we are open to all new members who wish to participate. As
-%mentioned all data are public immediately.
-%
-%{\bf For all of these reasons, we need to keep building up the sample of GRB
-%	afterglow spectra now as we may have to wait many years before a mission like
-%	\textit{Swift} becomes available again.}
-%
-%A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark
-%	bursts}'', e.g., Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012,
-%MNRAS, 421, 1265). Some of these are at the highest redshifts ($z > 6$) and
-%their observer-frame optical emission is absorbed by the IGM. The majority,
-%however, suffer from large dust obscuration (e.g., Perley et al. 2009, AJ, 138,
-%1690; Greiner et al.\ 2011, A\&A, 526, 30). Identifying such GRBs is important
-%for constraining the fraction of obscured star formation. In both cases, NIR
-%emission is expected. X-shooter can adequately study these objects, provided
-%that a NIR counterpart is timely identified, for which we have the dedicated
-%HAWKI run D.
-%
-%The detection of {\bf absorption line variability} can reveal the burst
-%influence on the surrounding medium and in turn the absorber distance from the
-%burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
-%2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\ 2012,
-%A\&A, 545, 64; Hartoog et al.\ 2013). {\bf Short GRBs} originate in a
-%substantially different environment compared to long GRBs. Short GRBs may be
-%related to the merging of compact binaries and the coalescence time can be long
-%enough to allow the progenitor system to move far away from the star formation
-%site (Belczynski et al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum
-%with a sufficient signal-to-noise ratio of a short GRB afterglow has been
-%secured. A knowledge of the redshift distribution of short bursts is of key
-%importance for the next generation of gravitational wave experiments, as they
-%are the likely EM counterparts to their primary targets.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria and follow-up}
+GRBs are interesting as they provide the opportunity to derive new insights
+into a very wide range of topics in astrophysics. In that sense GRB
+research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
+which was launched in 2004, has made
+it possible to harvest the great potential in using GRBs as probes of
+galaxies in the intergalactic medium, which was already hinted at by
+results from earlier missions.
+
+A crucial aspect of the great success of the {\it Swift} mission has been
+the extensive ground based follow-up observations of the afterglows and of
+the host galaxies of the GRBs. This work has involved a large community
+of researchers both inside and outside the official {\it Swift} team. This fruitful
+collaboration has been greatly facilitated by the open data access policy
+of the {\it Swift} mission.
+
+In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
+was secured using low-resolution spectrographs (typically
+R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
+very bright afterglows high-resolution (typically R$>$20000) spectra were
+secured.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
+instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
+echelle spectrograph
+covering the full spectra range from the UV atmospheric cutoff to the near-infrared
+(near-IR)
+K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
+transient source follow-up in mind as the fading luminosities of such sources 
+makes in urgent to secure as extensive coverage as possible in the shortest possible 
+time. At the same time the resolution was set to be in the range 4000--9000 in order
+to be able to get a large useful spectra range between the many sky-background
+emission lines in the red and near-IR.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sample selection criteria and observations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Being of transient nature, it is difficult to impose strong sample selection
 criteria on GRBs, without hampering the follow-up effort. Many natural follow-up
-restrictions exists already, being it weather conditions, pointing restrictions
-of the telescope or poorly localized bursts as reported by the alerting
-telescope. To maximize the return of the follow-up campaign we have chosen a few
-selection criteria that attempts to provide an unbiased selection of bursts,
-while allowing for a high follow-up success rate. The importance of defining
-unbiasing selection criteria has been highlighted previously
-\citep{Salvaterra2012, Hjorth2012, Vergani2015, Perley2016a}, when trying to
-address the specific underlying distribution functions such as the redshift
-distribution, host metallicity distribution, or brightness distribution. When
-investigating a specific distribution function, a high degree of
-\textit{completeness} is desired meaning that in terms of the parameter of
-interest, the largest possible fraction of targets in the sample has
-determinations of that parameter \citep{Perley2016b}.
-
-We simultaneously want to minimize any biases against astrophysical conditions
-while at the same time maximizing likelihood of observations allowing us to
-obtain a higher degree of observing completeness. By restricting the selection
-criteria to conditions local to Galaxy and therefore irrespective of intrinsic
-GRB properties, the hope is that the sample collect represents the underlying
-distribution of GRBs in a fair way while at the same time, additional unbiasing
-cuts can be imposed that can improve the follow-up success. The selection
-criteria used here is based on previous successes \cite{Jakobsson2006b,
-	Fynbo2009, Hjorth2012}, which has shown to represent the underlying distribution
-to a high degree. We test the degree of completeness in Sect.
-\ref{completeness}.
-
-The selection criteria are defined as follows:
+restrictions exists from already, being it weather conditions, pointing
+restrictions of the telescope or poorly localized bursts as reported by the
+\textit{Swift}-telescope. To maximize the return of the follow-up campaign
+we have chosen a few selection criteria that attempts to provide an unbiased
+selection of bursts, while allowing for a high success-rate
+
+\footnote{Note that in the P84 proposal the criteria have
+been stated a bit differently, the visibility constraint being replaced by a
+declination + Sun angle constraint. The above criteria are however those
+defining the sample.}
 
@@
 	\item GRB triggered onboard by Swift.
+	\item Galactic $A_V \lesssim 0.5$ mag.
 	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
 	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
-	\footnote{Note that in the P84 proposal the criteria have
-	been stated a bit differently, the visibility constraint being replaced by a
-	declination + Sun angle constraint. The below criteria are however those
-	defining the sample.}
 	\item No bright closeby stars.
-	\item Galactic $A_V \lesssim 0.5$ mag.
+
 \end{enumerate}
 
-Because our ability to observe GRB afterglows is strongly correlated with the
-rapidness and precision of the target positions delivered by the triggering
-facilities, by selecting only bursts that have been triggered on board the
-\textit{Swift} space telescope \cite{Gehrels2004}, by the Burst Alert Telescope
-(BAT) \citep{Barthelmy2005}, we start out with a homogeneously selected sample
-where burst characteristics are delivered immediately, allowing for an informed
-follow-up strategy. Despite the complexity of the triggering mechanism on board
-\textit{Swift} \citep{Band2006}, attempts at inferring properties of the
-underlying cosmic GRB population based on the detection thresholds and
-triggering algorithms have been made \citep{Lien2014, Graff2016} and thereby by
-restricting the follow-up effort to bursts detected by \textit{Swift}, the
-limitations of the parent sample is well studied.
-
-Because the localization accuracy of BAT is 1 - 4 arcminutes
-\citep{Barthelmy2005}, we additionally require an X-ray position from the X-ray
-Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
-\citep{Barthelmy2000} with 10 hours and to account for observing constraints on
-\textit{Swift} that XRT began observations within 10 minutes. The additional
-rapidness requirement of the XRT follow-up means that the number of detected
-X-ray afterglows for BAT triggers is 100 percent \citep{Burrows2007}. The
-\textit{Swift} constraints on the GRB follow-up means that the sample presented
-here inherits the \textit{Swift} sample properties, except the additional XRT
-constraint. Because the XRT completeness is very high\citep{Burrows2007}, this
-should change the parent sample significantly. This issue is addressed again in
-Sect. \ref{completeness}.
-
-To allow for observability we additionally require that the GRB is visible
-(duh!) from the telescope site at Cerro Paranal, Antofagasta, Región de
-Antofagasta, Chile, for a least 1 hour allowing time for the spectroscopic
-observations to completed. Since the GRB population is isotropically distributed
-seen from Earth and that GRB properties does not depend on position on the sky
-\citep{Briggs1996}, this cut does not influence our ability to sample the
-underlying GRB population. The same arguments applies to the requirement that
-there are no nearby bright stars, which is additionally supplemented by the
-requirement what the galactic dust content does not exert an extinction higher
-than $A_V \lesssim 0.5$ mag based on the extinction maps by
-\citet{Schlegel1998}. Because of  isotropy, these additional cuts should not
-influence the optical properties of the bursts themselves, only our ability to
-successfully secure observations that allows us to investigate the spectroscopic
-properties of GRBs.
-
-We address the succesfullness of these criteria and the actual follow-up effort
-in representing the full \textit{Swift} sample ion Sect. \ref{completeness}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up procedure and practicalities}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\todo{Will be written by Daniele?}
+A significant fraction of the bursts presented here \todo{Insert exact
+number} have already had their hosts investigated in \citet{Kruhler2015}, for
+which extractions of the hosts exist. The focus of the data presented here are
+on the afterglows themselves and in the absence of a clear afterglow, the host.
+We will not, however, investigate the hosts.
+
+We simultaneously want to minimize any biases against astrophysical conditions
+while at the same time maximizing likelihood of observations. By restricting
+the selection criteria to conditions local to Galaxy.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up procedure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Observations} \label{obs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-	\centerline{\includegraphics{figures/skymap.pdf}}
-	\caption{Mollweide projection in galactic coordinates of the full sky showing
-		the positions on the sky of the bursts presented in this work. The empty
-		circles are the positions of all the \textit{Swift} bursts detected until the
-		end of 2016. In blue is shown the position of the bursts fulfilling the sample
-		criteria specified in Sect. \ref{samplecrit}. The red circles are added around
-		GRBs which have been followed up by X-shooter since the commissioning of the
-		instrument. Blue points with red circles represent GRBs that enter our sample
-		and have X-shooter spectroscopy. The different samples are compared in Sect.
-		\ref{results}. The background shows the dust maps presented in
-		\citet{Schlegel1998} where the sample criteria cut with $A_V \lesssim 0.5$ mag
-		is removed.}
-	\label{fig:skymap}
-\end{figure*}
-
-Observations in this sample has been carried out using the cross-dispersed
-echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on two of the four
-Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the duration
-of this follow-up campaign, which covers the entire lifetime of X-shooter.
-Observations have been carried out from the ESO period P84 through P98 under the
-following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.A-0045,
-094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088, 089.A-0067,
-088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. A few additional
-bursts, not a part of the \textit{statistical sample}, have been included from
-092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI: Greiner), and Levan under
-095.B-0811(B) (PI: Levan). This sample represents \textit{all} GRBs that have
-ever been followed up by X-shooter.
+\subsection{Observations} \label{obs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+All the observations in this sample has been carried out using the
+cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
+two of the four Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal)
+during the duration of this follow-up campaign, which covers the entire lifetime
+of X-shooter. Observations have been carried out from the ESO period P84
+through P98 under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
+095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
+089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.
 
@@
 while X-shooter was mounted on UT3 during the commissioning of the instrument.
-The bursts observed during the commissioning period is not a part of our
+The bursts observed during the commisioning period is not a part of our
 statistical sample, but are nonetheless published as part of the bursts observed
@@
 a single exposure, by splitting the light in three separate spectroscopic arms
-using dichroics. This way each arm work as a separate instrument, each
-functioning as its own echelle spectrograph. The ultraviolet blue(UVB) arm
-covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 - 10200 \AA, and the
-near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the possibility of
-applying a k-band blocking filter cutting the coverage of the NIR arm at 10200 -
-21000 \AA.
+using dichroics.
+This way each arm work as a separate instrument, each functioning as its own
+echelle spectrograph.
+The ultraviolet blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm
+covering 5500 - 10200 \AA, and the near-infrared()NIR) arm covering 10200 -
+24800 \AA~ with the possibility of applying a k-band
+blocking filter cutting the coverage of the NIR arm at 10200 - 21000 \AA.
 \\
@@
 We list the overview of all the observations in Tab. \ref{tab:sample_overview}.
-\input{tables/burst_overview.tex} and plot the positions off all the bursts on
-the night sky in Fig. \ref{fig:skymap}. The central zone of avoidance due to
-high galactic extinction cutoff is visible with an apparent isotropic
-distribution of the bursts away from the galactic center except in the upper
-right corner, where Earth casts a shadow on the sky as seen from the telescope.
-
-A fraction of the bursts presented here \todo{Insert exact
-number} have already had their hosts investigated in \citet{Kruhler2015}, for
-which extractions of the hosts exist. The focus of the data presented here are
-on the afterglows themselves and in the absence of a clear afterglow, the host.
-We will not, however, investigate the hosts.
+\input{tables/burst_overview.tex}
 
@@
 
-\begin{figure*}
-	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
-	\caption{Telluric corrected, normalized spectrum of GRB121024A at $z = 2.300$,
-	meant to illustrate the typical data quality. The acquisition magnitude is r =
-	20, meaning it is in the brighter end of the sample presented here, but not the
-	brightest. The spectrum is rich in absorption, including molecular absorption
-	of $H_2$ with the absorption due to \lya is visible at $\sim$ 4000 $\AA$.
-	Additionally, three intervening systems are seen in this sightline  This
-	spectrum has been analyzed in detail in \citet{Friis2015}. }
-	\label{fig:spectrum}
-\end{figure*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!ht]
-	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
-The lower horizontal axis is in units pixels, the top axis in arc seconds.
-The red datapoints show a subsample of NIR spectra.
-		The grey line shows a linear fit to the VIS datapoints. }
-	\label{fig:res}
-\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Science data products} \label{products}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results} \label{results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this section we describe the efficiency of the follow-up effort and the
-characteristics of the observed bursts. We also assess the degree to which the
-sample obtained is complete. An important note is that here we present
-\textit{all} GRBs that have ever been observed with X-shooter, while only a
-subset of these constitute our \textit{statistical} sample. The statistical
-sample is based on the selection criteria described in Sect. \ref{samplecrit}
-and should represent an unbiased sample of the underlying, parent GRB
-population. Single bursts outside the sample criteria have been followed up on
-the basis of the curiosity of their light curve or their brightness, but is not
-discussed as part of the investigation of the GRB population. A prime example of
-a burst outside the statistical sample is GRB161023A with it's rich
-spectrum containing 11 intervening absorption systems and pending further analysis.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up timing and afterglow brightness} \label{timing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Redshift determination of bursts where the host is too faint for a spectroscopic
-redshift relies on the absorption spectrum imprinted on the GRB afterglow
-continuum, which again depends on the afterglow brightness at time of follow-up.
-Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better references} a rapid
-follow-up is essential for the accurate designation of the GRB host \todo{Get
-reference}. In Fig. \ref{fig:timing} we plot the follow-up delay from the BAT
-trigger to the start of the spectroscopic observation. Visible as the
-shortest delays are the bursts observed in RRM-mode in which where the fastest follow-up
-between BAT trigger and start of spectroscopic observations is for the short, $z = 1.717$, GRB160410A which was
-followed up only 8.4 minutes after the BAT trigger.
+\begin{figure}
+	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
+The lower horizontal axis is in units pixels, the top axis in arc seconds.
+The red datapoints show a subsample of NIR spectra.
+		The grey line shows a linear fit to the VIS datapoints. }
+	\label{fig:res}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we describe the efficiency of the follow-up effort and the characteristics of the observed bursts. We also assess the degree to which the sample obtained is complete.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up timing and afterglow brightness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A rapid follow-up is essential for the accurate designation of GRB the origin. The redshift determination of burst where the host is not easily
 
@@
 	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}}
-	\caption{Follow-up timing and afterglow magnitude at the start of observation.
-	The points have been colored based on the redshift of the corresponding burst.
-	Points with black faces does not have a redshift determination and downarrows
-	indicate a non-detection in the acquisition image. In blue is shown the
-	fractional redshift completess as a function of delay-time.}
+	\caption{Follow-up timing and afterglow magnitude at the start of observation. The points have been colored based on the redshift of the corresponding burst. }
 	\label{fig:timing}
@@
 
-To illustrate the effect of redshift completeness loss due to late follow-up, we
-overplot the redshift completeness fraction over the delay time. Visible is the
-increasing fraction of GRBs without a good redshift determination for increasing
-delays. That the redshift completeness of the bursts we have followed up are $>$
-90 percent \todo{Get number}, shows the ability of VLT/X-shooter to successfully
-get redshifts. Not shown in the figure is an additional 12 \todo{Get number}
-bursts that we have redshift determinations from late-time host observations
-with delay-times longer than $\sim$10 days.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample completeness} \label{completeness}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
-\todo{Get number} bursts has been triggered by \textit{Swift} in the period
-since the commissioning of VLT/X-shooter in \todo{Insert date}. This sample
-constitutes the \textit{statistical sample} on which we will try to address the
-GRB population properties. Of this sample, 86 \todo{Get number} have been
-spectroscopically followed up with X-shooter. In order to assess whether our
-sample criteria and the subset of bursts followed up represent the underlying
-parent population of GRB, we compare some intrinsic GRB properties of our sample
-to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
-that \textit{Swift} randomly samples the underlying GRB population, if the
-sample statistics of our subsample are similar to the \textit{Swift}-sample, we
-can assume that the population properties should be conserved. There is some
-evidence that the burst duration does not scale with redshift as expected
-through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
-that we could be missing an increasing fraction of the bursts at higher
-redshifts, but \citet{Lien2016} argues that we should only be missing the bright
-"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
-the sample presented here is complete. We show the comparison between the burst
-properties in Fig. \ref{fig:swift_complete}.
-
-\begin{figure*}
-	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
-	\caption{Comparison between bursts properties of all bursts observed with
-	\textit{Swift}, the properties of the ones that fullfill the sample criteria
-	specified in Sect. \ref{samplecrit}, and the subset that has been observed as
-	part of the statistical sample. The left plot shows the fluence in 15-150 keV
-	as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
-	after the bursts as measured by XRT. Based on the XRT spectrum,
-	\citep{Evans2009}, a hydrogen column density has been inferred which we show in
-	the right most panel.}
-	\label{fig:swift_complete}
-\end{figure*}
-
-Using the observational characteristics of the 1244 \todo{Get number} bursts
-observed by \textit{Swift} to date \cite{Gehrels2009}, and the X-ray derived
-hydrogen column densities \citep{Evans2009}, we can quantify the degree to which
-our sample is biased against the overall \textit{Swift} sample in terms of
-intrinsic GRB properties. The values are queried from the online \textit{Swift}
-database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
-Three main samples are of interest to compare to assess the completeness of the
-follow-up campaign; the full \textit{Swift} sample consisting of all the bursts
-observed by \textit{Swift}, excluding the bursts entering our sample (1), the
-burst that fulfill the selection criteria imposed in Sect. \ref{samplecrit}, but
-has not been followed up (2), and the bursts actually followed up with X-shooter
-(3).
-
-\input{tables/sample_properties.tex}
-
-For each of the samples we calculate the median and 14th and 86th percentile
-which can be used as point estimates for the population distribution. These are
-shown in \ref{tab:sample_properties}. It can be seen from the values that the
-three samples have very similar distributions in terms of the point estimates
-chosen, suggesting that our selection criteria are unbiasing against the
-\textit{Swift}-sample and that additionally, the follow-up effort is unbiasing
-towards the intrinsic GRB properties. This is despite spectroscopic follow-up
-only being carried out in cases where either a detectable optical afterglow or a
-clear counterpart are seen, which naively should be biased against dark bursts
-occurring in more obscured galaxies, which is shown to exhibit different
-galactic properties \citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b,
-	Perley2015b}.
-
-Additionally, using a 2-sided Kolmogorov-Smirnov test (KS-test), we can assess
-the degree to which the null that the two distribution are drawn from the same
-distribution can be rejected. We show a graphic representation of the test
-statistics in Fig. \ref{fig:p_values}. We have run the test on whether is can be
-rejected that the sample criteria is drawn from the same distribution as the
-\textit{Swift} sample, if the observed sample is dissimilar from the statistical
-sample, and again whether the observed sample differs from the full
-\textit{Swift} sample. A high p-value indicates little evidence against the null
-meaning for the parameters investigated here, only the X-ray derived column
-densities exhibits the highest degree of dissimilarity, but not to a significant
-level. A large part of the tension between the two samples in $N_H$ is driven by
-a small number of low column densities($\log(N_H) \lesssim 19 \mathrm{cm}^2$)
-bursts.
-
-\begin{figure}
-	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
-	\caption{Relational graph shown the respective p-values that the different
-	sample are drawn from the same underlying distribution. The arrows represent
-	the comparison with each of the p-values are listed next to. Only in the HI
-	column density distribution is there mild evidence against the null, but the
-	discrepancy is mainly driven by a relatively larger fractional contribution
-	from low-column hosts.}
-	\label{fig:p_values}
-\end{figure}
-
-We therefore conclude that the sample presented here, to a high degree
-represents the intrinsic properties of the GRBs in the full \textit{Swift} sample.
-\todo{This doesn't quite feel exhaustive? Is there a better way of addressing the issue of completeness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample completeness}	
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Based on the sample criteria specified in Sect.
+\ref{samplecrit}, a total of 152 \todo{Get number} bursts has been triggered by \textit{Swift} in the period since the comissioning of VLT/X-shooter in \todo{Get date}.
+This constitutes the \textit{statistical sample}.
 
@@
 
-GRB follow-up requires the immediate availability of the telescope in order to
-successfully carry out spectroscopic observations. Out of the 158 \todo{Get
-exact number} bursts in the statistical sample, 32($\sim$ 20 percent) \todo{Get
-exact number} have not been observed due to conditions, not relating the GRB or
-afterglow properties. The reasons includes unavailability of the telescope due
-to technical maintenance such af mirror re-coating, a visitor rejecting the TOO
-trigger, or bad weather at the observing site. Because this cut is irrespective
-of the GRB properties, it is not expected that this should change sample
-properties. Addressing whether this cut introduces a bias in the followed sample
-is not trivial, but in terms of intrinsic GRB properties, the sample
-distributions does not change significantly, as shown in Sect. \ref{completeness}.
-\todo{Should this be quantified somehow?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample darkness}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A subset of all GRBs exhibit no detectable or very faint optical afterglow
-\citep{Groot1998, Djorgovski2001, Fynbo2001}, which had been parametrized in
-term of "darkness" \citep{Jakobsson2004, VanderHorst2009}. The X-ray properties
-of this subset has previously been investigated \citep{DePasquale2003,
-	Melandri2012}, finding a slightly higher X-ray luminosity compared to the
-optically brights bursts combined with similar prompt characteristics pointing
-towards a denser circumburst material. This indicates along with investigations
-of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
-extinction of the optical afterglow is primarily driven by the presence of dust,
-local to the burst \todo{Local dust reference}. Because the dust content of a
-galaxy is correlated with the gas amount \citep{Bohlin1978, Guver2009}, we
-expect the X-ray derived column densities for the bursts without an optical
-afterglow to be higher than the ones without.
-
-For all bursts with follow-up within 100 hours we can calculate the
-"darkness"-parameter, $\beta_{OX}$ \citep{Milvang-Jensen2012}. This required the
-simultaneous measurement of the X-ray flux density and the optical flux density
-which is in practice never possible. We use the measured acquisition camera
-magnitude reported in Tab. \ref{tab:sample_overview} to get the optical flux
-density at the beginning of the spectroscopic integration. Because we know the
-delay between the follow-up and the \textit{Swift} trigger, we can use the
-measured XRT burst lightcurve \citep{Evans2007,
-	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
-to infer the corresponding X-ray flux density. This is done by either linearly
-interpolating between temporally neighboring XRT measurements of by
-extrapolating the last few X-ray data points to the spectroscopic delay. When
-the afterglow is not detected in the acquisition camera, an upper limit of $> 24$
-mags is used, which propagates into an upper limit on $\beta_{OX}$.
-
-In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
-the one presented in \citet{Fynbo2009}. We have replaced limits with values,
-meaning that we artificially bias the distribution towards higher
-$\beta_{OX}$-values because darker bursts naturally fainter in the optical
-making the fraction of non-detections higher for lower $\beta_{OX}$. We
-marginally confirm the result by \citet{Fynbo2009} that dark bursts $\beta_{OX}
-< 0.5$) have slightly higher X-ray derived column densities, by calculating the
-median, 14th, and 86th percentile of the distributions arising by separating the
-sample at $\beta_{OX} = 0.5$. For bursts with $\beta_{OX} \geq 0.5$ we find
-$\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
-$21.8_{-0.9}^{+0.5}$, meaning that the median value for dark bursts is higher,
-but with a great deal of overlap. A 2-sided KS test fails to reject the null
-that they are drawn from the same distribution with $p = 0.43$, meaning there is
-no clear evidence for a discrepancy.
-
-\begin{figure}
-	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
-	\caption{Optical extinction relative to X-ray against X-ray derived hydrogen
-	column densities. In red is shown the sample presented in \citet{Fynbo2009}
-	where is lines indicate the kernel density estimate of the distribution where
-	limits have been replaced with values. Darker colors represented denser points.
-	The correcspoing marginal distributions are shown. In blue is the points for
-	the bursts presented here along with the marginal histograms. Limits on
-	$\beta_{OX}$ is shown by downwards facing arrows.}
-	\label{fig:betaOX}
-\end{figure}
-
-Because we only follow up targets with either a clear afterglow or counterpart,
-correctly associating a galaxy with a burst when there is no detected optical
-afterglow is more difficult, see \citet{Levesque2010} and \citet{Perley2017} for
-a recent example. Because the presence of optical transient is highly dependent
-on the "darkness", we expect that the $\beta_{OX}$ sample is biased against dark
-bursts. The larger relative fraction of upper limits for increased darkness is
-visible both in the sample presented here and the one in \citet{Fynbo2009}. The
-X-ray properties for the followed sample does not differ signicantly from the
-statistical sample, as shown in Sect. \ref{completeness}. Using the table
-maintained by Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How
-to reference this}, we can see how the presence of an optical afterglow affects
-the follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do
-not have an detected optical afterglow, but this number also includes bursts
-where no optical telescopes were available, so is a lower limit. For the bursts
-entering our statistical sample, this number is $\sim$ 27 percent, close to the
-maximum fraction of dark bursts \citep{Melandri2012}. Of the subsample (3) that
-have actually been attempted, this number is $\sim$ 18 percent, confirming that
-we are indeed biased against bursts with an optical afterglow in the
-spectroscopic sample.
-
-Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts. \todo{Daniele, help.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{On the redshift distribution of GRBs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Because the redshift completeness of our statistical sample is  54 percent
-\todo{Get number}, making inference of the underlying redshift distribution of
-GRBs is difficult. Because a large fraction of the bursts for which we were
-unable to secure a redshift are missing due to terrestrial reasons, see Sect.
-\ref{badbursts}, imposing additional unbiasing cuts on the sample can increase
-the redshift completeness at the cost of sample size. As long as the cuts we
-impose are not selective in terms of the burst properties, the sample properties
-stay the same. The first cut we try is telescope availability such as bad
-weather, visitor rejected the TOO, or unavailability of the instrument. Because
-telescope availability is independent on the burst properties, by cutting away
-these from our sample, the underlying redshift distribution should be conserved.
-By trimming away the 30 bursts \todo{Get number} that can be rejected due to
-telescope availability, the redshift completeness increases to 63 percent
-\todo{Get number}. As is shown in Sect. \ref{timing}, the redshift completeness
-decreases with increasing follow-up delay, we can additionally increase the
-redshift completeness by imposing cuts on the follow-up delay. Our ability to
-rapidly follow up a burst is also not dependent on the burst properties and
-therefore we can additionally cut delays longer than 24 hours (all burst in the
-sample are in principle observable within 24 hours).
+How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{On the nature of triggered GRBs without detected optical afterglow.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The biased introduced in the redshift distribution by the increased ability to
-secure redshift of optically brighter bursts has been investigated by
-\citet{Turpin2016} which find that we systematically find redshift for longer
-GRBs. Additionally only the brightest GRBs are seen above redshift $z \gtrsim
-1$. \todo{Additional text}
 
@@
 model. Additionally, Similarly so, the red wing has got the ISM signatures
-imprinted on it, especially strong absorption due to \SIii, \sii~and \nv~which
+imprinted on it, especially strong absorption due to \SIii, \sii and \nv~which
 can exhibit significant velocity structure. Along with instrumental effects, the
@@
 
-In a similar manner to how the X-ray derived column densities was used to assess
-the sample completeness in Sect. \ref{completeness}, we addresses the same
-question in terms of the optically derived hydrogen column densities. Using the
-compilation by \citet{Tanvir2017} we have all 81 published HI values previously,
-we can compare to the 39 new HI columns added by this sample. We note that this
-sample increase the number of optically derived hydrogen column densities with
-$\sim$ 50 percent. We compare the median, the 14th, and 86th percentiles of the
-two distributions where the sample presented here has $21.8_{0.8}^{0.3}$ and the
-rest of the litterature values has $21.5_{1.5}^{0.4}$. We see that the two
-distribution has a large degree of overlap due to the large width of the
-distributions, but find a slightly higher median value for the new sample
-presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
-relatively large evidence against the null that the two samples are drawn from
-the same underlying distribution. Because the bursts that have measurements of
-the hydrogen column density are selected sole based on our ability to infer a
-column, it is difficult to make any strong conclusions about the completeness in
-terms of gas content. We show the comparison between the two samples in Fig.
-\ref{fig:NH_dist}, where we additionally show the column-distribution for 12081
-quasar DLAs published in \citet{Noterdaeme2012b}. The ability of the GRBs to
-select the highest N$_H$ is immediately visible as is also noted previously
-\citep{Prochaska2007, Fynbo2009}. The reason for this is the selection of sight
-lines as random position of the QSO DLAs whereas the GRB forms in dense regions
-in the host. The N$_H$ distribution has also be used the infer the position of
-the gas in host \cite{Buchner2016} where a galaxy-wide gas reservoir is
-preferred.
-
-\begin{figure}
+\begin{figure*}
 	\centering
-	\includegraphics[width=9cm]{figures/NH_dist.pdf}
-	\caption{Distributions of DLA hydrogen column densities for DLAs found in
-		quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
-		is the similar distribution, only for DLAs in GRB sightlines. Values are taken
-		from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
-		with the new values presented in this sample. The marginal distributions for
-		the two distribution are also shown, where the different environments probed
-		are clearly visible in the hydrogen column densities, as previously also noted
-		in \citet{Fynbo2009}}
-	\label{fig:NH_dist}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison to previous samples.}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This should also include a discussion of dark bursts. 
-Cuchiara. BAT6 (Salvaterra), TOUGH
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{X-ray column densities vs. optically derived ones}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
-%
-%\begin{figure}
-%	\centerline{\includegraphics[width=10cm]{figures/OAXR_NH_sphere.pdf}}
-%	\caption{Optically derived HI column densities vs.
-X-ray derived ones taken from \citet{Buchner2016}.
-The colors of the points reflect the redshift of the burst.
-Circles mark the column densities published in \citet{Fynbo2009} and squares mark the ones derived as part of the sample presented here.
-Overplot in blue is the 1:1 correspondence, expected if the conditions local to the GRBs are similar to the local ISM abundances.
-If the metallicity is lower at the GRB site, the point move upwards in the plot.
-}
-%	\label{fig:OAXR}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Why is the HI column truncated?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
-Number of bursts with detected H2 and their HI column density.
-Rahmati Schaye, Krongold Prochaska.
+	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
+	\caption{ Measurements of the hydrogen column-densities for all bursts with a clear Lyman alpha absorption system.
+In solid black is shown the spectrum with black dotted giving the corresponding 1-$\sigma$ error.
+Black dashed shows zero flux density.
+The solid red line is the absorption of column density equal to the value presented in Tab.
+\ref{HI_tab} with the 1-$\sigma$ interval shown with dashed lines.}
+	\label{fig:HI1}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI2}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI3}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI4}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI5}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper we have presented the results
+of a dedicated effort over the years 2009 -- 2016 to use the X-shooter 
+spectrograph on the ESO-VLT to secure spectroscopic observations of 
+afterglows and host galaxies of GRBs detected by {\it Swift}. This 
+work was initiated by a consortium using Guaranteed Time on X-shooter, 
+but over the years the project continuued in open time and the team was
+opened to include any researchers interested in contribution to this 
+effort.
+
+The sample we present here include spectroscopic observations of 86 
+systems fulfilling our sample criteria of which 72 are afterglow spectra 
+and 14 are observations of the underlying host galaxies. 
+UPDATE THIS.
 
@@
 
-This work made use of data supplied by the UK Swift Science Data Centre at the University of Leicester.
 \end{acknowledgements}
@@
 
-\clearpage
-
-\begin{figure*}[!h]
-	\centering
-	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Measurements of the hydrogen column-densities for all bursts with a
-		clear Lyman alpha absorption system. In solid black is shown the spectrum with
-		black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
-		flux density. The solid red line is the absorption of column density equal to
-		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
-		with dashed lines.}
-	\label{fig:HI1}
-\end{figure*}
-\clearpage
-\begin{figure*}[!h]
-	\centering
-	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI2}
-\end{figure*}
-\clearpage
-\begin{figure*}[!h]
-	\centering
-	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI3}
-\end{figure*}
-\clearpage
-\begin{figure*}[!h]
-	\centering
-	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI4}
-\end{figure*}
-\clearpage
-\begin{figure*}[!h]
-	\centering
-	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI5}
-\end{figure*}
-\clearpage
-
+\newpage
 \appendix
@@
 \subsection{GRB090313 (z = 3.373)}
-The first GRB ever observed with X-shooter, during the commissioning of the
-instrument, this data formed the basis of GCN
-\#9015\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9015.gcn3}} and is published
-in \citet{DeUgartePostigo2010}. Due to the lingering brightness of GRB090313,
-6.9 ks spectroscopic integration starting 45 hours after the BAT trigger reveals
-a wealth of absorption features superposed on the afterglow continuum at a
-common redshift of $z = 3.373$. Two intervening systems at $z = 1.959$ and $z =
-1.800$ is identified based on strong \mgii-absorption. Because this burst is
-observed before the instrument is science-verified, it does not enter into the
-statistical sample.
 
 \subsection{GRB090530 (z = 1.266)}
-Observed during paranalization of the instrument, this data forms the basis of
-GCN \#15571\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15571.gcn3}}, but is not
-published elsewhere. Observations began 20.6 hours after the BAT trigger and 4.8
-ks spectroscopic integration in all three arms reveals the absorption signature
-for a host at $z = 1.266$ from the detection of \mgii, \mgi, \SIii, \feii,
-\aliii. Because this burst is observed before the instrument is
-science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB090809 (z = 2.737)}
-Observed during the first science verification period and was used for GCN
-\#9761\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9761.gcn3}} and is
-additionally used as the basis for the master thesis by \'Asa Sk\'ulad\'ottir
-(2010). 7.2 ks integration starting 10.2 hours after the GRB trigger notice
-yields the clear afterglow continuum in all arms from with Lyman Limit located
-in the beginning of the UVB coverage. The simultaneous detection of absorption
-lines identified as \lya, \SIii, \oi, \SIi$^*$, \SIiv, \civ, \feii, \alii,
-\aliii and \mgii at $z = 2.737$ sets it as the redshift of the GRB. Because this burst is observed before the instrument is
-science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB090926 (z = 2.106)}
-Starting during the second science verification period, this dataset forms the
-basis of GCN \#9942\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9942.gcn3}} and
-is additionally published in \citet{DElia2010}. Spectroscopic integration
-started 22 hours after the BAT trigger and from the acquistion camera the
-optical afterglow has R = 17.9 mag (vega) at the beginning of the observations
-which causes a very continuum to be seen in all arms. An absorption trough due
-to \lya is clearly visible along with numerous metal resonance lines \civ,
-\SIii, \SIii$^*$ \feii, \mgii, all at $z = 2.106$, marks this as the redshift of
-the GRB. Because this burst is observed before the instrument is
-science-verified, it does not enter into the statistical sample.
 
 \subsection{GRB091018 (z = 0.971)}
-The first burst observed during normal operation after science verification was
-completed and there is the first burst that enter the statistical sample. This
-data is the basis for GCN \#
-10042\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10042.gcn3}} and is published
-in \citet{Wiersema2012}. With a bright afterglow and a rapid follow-up, this
-spectrum is of pristine quality. The afterglow continuum is bright throughout
-all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
-\mnii, \mgii, \mgi, and \caii al located at $z = 0.971$, setting is as the
-redshift of the host.
-
-\subsection{GRB091127 (z = 0.490)}
-Obtained 4 days after the burst trigger, this data forms the basis for GCN \#
-10233\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10233.gcn3}} and is published
-in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the S/N of
-the afterglow continuum is poor especially in the UVB arm, why no clear
-absorption lines are detected against the afterglow continuum, although see
-\citet{Vergani2011} which report a tentative detection of \mgii. Emission lines
-from the underlying host is clearly visible with lines from \oii, \hb, \oiii,
-and \ha~all at $z = 0.490$. This bursts is additionally associated with
-SN2009nz.
+
+\subsection{GRB091127 (z = 0.49)}
 
@@
 trigger and has a total exposure time of 4.8 ks. Absorption features, including
-those of \lya~and from a multitude of ions are detected against the afterglow
+those of \lya, from a multitude of ions are detected against the afterglow
 continuum at $z = 4.667$. Additionally, absorption from an intervening system
@@
 to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii,
-all at a consistent redshift. Temporal evolution of the fine structure lines
+all at a consistent redshift. Temporal evolution of the fine-structure lines
 belonging to \feii$^*$ is found between the epochs.
@@
 \civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
-fine structure lines from \NIii and \feii are exited local to the GRB.
+fine-structure lines from \NIii and \feii are exited local to the GRB.
 Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
@@
 \SIii, \sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along
-with the corresponding fine structure lines from (\feii$^*$, \SIii$^*$,
+with the corresponding fine-structure lines from (\feii$^*$, \SIii$^*$,
 \oi$^*$, \oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
@@
 absorption feature from Lyman alpha, along with narrow lines from \civ, \SIii,
-\SIiv, \feii, \sii, and \alii, as well as fine structure lines associated with
+\SIiv, \feii, \sii, and \alii, as well as fine-structure lines associated with
 \SIii$^*$ are all detected at $z = 2.300$, securely setting it as the redshift

### 2017-10-24 document/XSGRB_sample.tex
Adding ABBA and ADC

--- 2017-09-18 document/XSGRB_sample.tex
@@
 applying a k-band blocking filter cutting the coverage of the NIR arm at 10200 -
-21000 \AA.
-\\
-
-\todo{ABBA}
+21000 \AA. For all observations, a nod-and-shuffle observing scheme has been
+employed, with a nodding throw of 5 \arcsec. The individual observations has
+been carried out in a standard ABBA pattern, which allows flexibility in the sky
+subtraction method for the different arms.
 
@@
 slit width. We discuss this in Sect. \ref{resolution}.
-\\
-\todo{ADC discussion}
+
+Due to mechanical issues, the atmospheric dispersion corrector (ADC) were
+disabled from 1st of August 2012. No bursts were affected by the failing ADC
+prior to disablement. To avoid chromatic slit losses due to atmospheric
+dispersion, all subsequent observations have been carried out at parallactic
+angle. A consequence of this is that for all observations following 1st of
+August 2012, the source moves across the spatial direction of the slit as a
+function of wavelength. This effect has been modeled in the extraction
+procedure, as described in \ref{extract}.
 
@@
 
-%A fraction of the bursts presented here \todo{Insert exact
-%number} have already had their hosts investigated in \citet{Kruhler2015}, for
-%which extractions of the hosts exist. The focus of the data presented here are
-%on the afterglows themselves and in the absence of a clear afterglow, the host.
-%We will not, however, investigate the hosts.
+A subset of the bursts presented here  have already had their hosts investigated
+in \citet{Kruhler2015}, for which extractions of the hosts exist. The focus of
+the data presented here are on the afterglows themselves and in the absence of a
+clear afterglow, the host. We will not, however, investigate the hosts.
 

### 2017-10-24 document/XSGRB_sample.tex
Merge branch 'master' into writing_branch
# Conflicts:
#	document/XSGRB_sample.tex

@@
 
-For a brief moment, the energy output of a $\gamma$-ray burst (GRB) can rival
-that of the stellar light integrated over the entirety of the observable
-universe \citep{Kumar2015}. Bright across the entire electromagnetic spectrum,
-the following afterglow makes possible the exploration of the intervening space,
-otherwise dark. Since the discovery of the first $\gamma$-ray counterparts
-nearly two decades ago\citep{Costa1997a, VanParadijs1997, Djorgovski1997,
-	Frail1997, Wijers1998}, significant advances in the understanding of the underlying
-phenomenology and potential usage has been made.
-
-Categorized on the basis of their duration and spectral hardness, two main classes of GRBs exist, long and short \citep{Kouveliotou1993}.
-Now thoroughly established, the connection between long GRBs and supernovae (SNe) \citep{Galama1998, Hjorth2003a, Stanek2003, Woosley2006c} and the association of the location of long GRBs with the most star-forming regions in their host galaxies \citep{Fruchter2006, Svensson2010a}, make a stellar origin a likely explanation for the progenitor system.
-The seeming metal aversion \citep{Graham2013, Kruhler2015, Perley2016b} in selection of host at low redshift along with theoretical single-progenitor models preferring more pristine starting conditions \citep{Woosley2006d}, paints a picture where the core of a rapidly spinning, massive star undergoes gravitational collapse, although a certain degree of binarity \citep{Fryer2005, Podsiadlowski2010} in the progenitor population is required to explain the host metallicity distribution to super solar values \citep{Niino2011}, see \citet{Levan2016} for recent review.
-The driving engine in short GRBs has been established as the merger of two compact objects \citep{Metzger2010} through the detection of an associated kilonova \citep{Tanvir2013a, Berger2013a}.
-With the advent of the first gravitational wave (GW) detection \cite{Abbott2016a}, the electromagnetic counterpart in the form of a short
-
-GRBs as cosmological probes
-
-Do GRBs trace star formation
-
-Afterglow spectroscopy
-
-Complete samples
-
-Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ QSOs have 5 systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep{Jakobsson2006a, Fynbo2009}. Whereas DLAs
-towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
-atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
-\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
-star-forming galaxies over the full history of cosmic star formation from $z
-\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
-	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we can hence characterize the properties of
-star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities.
-
-Content of paper
-
-%There are several unanswered but fundamental questions that must be addressed in
-%order to exploit the full potential of GRBs as cosmological probes. More than
-%50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and
-%5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
-%	Salvaterra2012,Jakobsson2012, Perley2016a}. A high redshift completeness is
-%crucial for our understanding of the link between the number density of GRBs per
-%unit redshift and the global star-formation history of the Universe, as measured
-%by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
-%GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
-%galaxies at the highest redshifts and unsurpassed in providing detailed
-%abundance information via absorption line spectroscopy \citep{Tanvir2012,
-%	McGuire2016}.
-%
-%\smallskip
-%
-%%The sample of short GRBs with redshifts remains too small and too
-%%incomplete to really
-%%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
-%%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
-%%enlarge as far as possible the number of known redshifts for this class of
-%%sources.
-%
-%From March 2005 to March 2016 there have been about 350 {\it Swift} bursts
-%complying with our sample selection criteria, and
-%about half of them have measured redshifts. Among the latter subset, the team
-%proposing these observations has measured about two thirds of the redshifts,
-%mainly with FORS1/2 and X-shooter \citep[see][]{Fynbo2009, Jakobsson2012, Kruhler2012}. Our
-%current aim is to build a sample superior to our previous low-resolution survey
-%\citep{Fynbo2009}, both in terms of quantity and  quality of
-%the spectra. 
-%
-%Due to
-%the wide wavelength range of X-shooter with the same observation cover molecular
-%H$_2$ absorption near the atmospheric cut-off and all the strong emission lines
-%from the host in the NIR arm \citep[e.g.,][]{Friis2015}. In
-%general, the wide wavelength coverage ensures that we always have features on
-%which to base a redshift measurement as long as the afterglow is brighter than
-%about 23 mag in either the $R$- or $z$-band. Frequently, emission lines are also
-%detected from the underlying host, which also provide further information such
-%as SFR and metallicity (the top right panel in Fig.~1 shows an example). Only
-%for 7 out of more than 70 secured spectra could we not measure a redshift. With
-%the X-shooter survey we provide {\bf metallicity measurements} for about 30\%
-%(Voigt-profile fits) of the $z>1.7$ events. So far we have measured
-%metallicities for more than 20 GRB afterglows with X-shooter. With the wide
-%wavelength coverage of X-shooter we can study important chemical species as
-%Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$ elements over a much wider
-%redshift range than what is possible with other instruments. As an example, we
-%have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A at $z=4.669$
-%(Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for GRB\,111008A at
-%$z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$ for the
-%$z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139). Reconciling the
-%abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs in
-%particular, and old stars in the Local Group is an important long-term goal (see
-%also Sparre et al.\ 2014, ApJ, 785, 150). Metallicities are also measured from
-%host emission lines (Kr{\"u}hler et al.\ 2015, A\&A, 581, A125). GRB
-%spectroscopy also allows us to determine the dust content of their environments,
-%both through analysis of the depletion pattern and through measurement of the
-%associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). This allows us to
-%quantify the dust-to-metals ratio and its evolution with redshift (e.g., De Cia
-%et al.\ 2013, A\&A, 560, 88; Zafar \& Watson 2013, A\&A, 560, 26). %Moreover,
-%% GRB
-%%afterglows allow investigation of random foreground absorbers, complementing
-%%QSO sightline studies (e.g., Schulze et
-%%al.\ 2012, A\&A, 546, 20).
-%
-%%In Fynbo et al. (2009, ApJS, 185,
-%%526), we compare the host-galaxy HI
-%%column density and metallicity (most of which has been measured by our group)
-%%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
-%%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
-%%column density and metallicity for GRB and QSO sightlines shows that the total
-%%cross-section for environments like those of GRBs must be less than a few
-%%percent of the total cross-section for QSO-DLAs, and suggests that GRB
-%%sightlines probe regions with high gas density.
-%
-%%The issue of production and {\bf release of Lyman continuum radiation} is
-%%currently not fully understood, but the evidence suggests that star-forming
-%%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
-%%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
-%%X-shooter spectroscopy will allow major progress due to the excellent blue
-%%sensitivity.
-%
-%%GRB afterglows can probe dust in high-redshift star-forming galaxies
-%%including exploring the {\bf 2175~\AA{} extinction bump}
-%%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
-%%732, 143).
-%%Due to the broad wavelength  coverage we can do
-%%this more accurately and over a broader range of redshifts with X-shooter.
-%
-%\smallskip
-%
-%We will also determine the frequency and properties of {\bf molecular
-%	absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
-%the process of star formation, but prior to our program H$_2$ had been detected
-%just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
-%in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
-%found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
-%2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). We are currently
-%analysing more of our spectra for less obvious molecular absorption and we
-%expect to find more (a dedicated sample paper is addressing this issue).
-%
-%\vspace{0.2cm} A natural question to ask is: {\bf how long should this work
-%	continue?} Our view is that we need to keep observing the afterglows as long as
-%we have \textit{Swift} in operation. Also note that the program is still
-%producing many papers and provides data for many theses (Box 9 and 10).
-%\textit{Swift} is currently funded until 2018, but is likely to get more
-%extensions given its overwhelming success. As mentioned above GRBs allow us to
-%probe star-forming galaxies that are almost impossible to study in other ways
-%both in terms of redshifts, galaxy luminosity function, and regions within
-%galaxies. After 7 periods we have secured seven spectra of $z>4$ GRBs, of which
-%three were of sufficient quality to allow abundance measurements (Th\"one et
-%al.\ 2013, Sparre et al.\ 2014, Hartoog et al.\ 2015). GRBs offer the only way
-%to derive chemical abundances for the gas phase of central, actively
-%star-forming regions of high-$z$ galaxies. The program also maintains a very
-%high discovery potential where we occasionally find something completely
-%unexpected that provides interesting clues to puzzles in other fields, e.g.
-%extinction of type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of
-%these spectra are like precious jewels – it is a type of observation that can
-%never be repeated and a class of sightlines that can only be studied while we
-%have operating GRB satellites.
-%
-%It is also worth adding that we have build up a rather unique team spread over
-%Europe from Granada to Reykjavik, which by now has reached a point where the
-%distribution of night shifts, the scientific exploitation of the data is
-%efficient and where we are open to all new members who wish to participate. As
-%mentioned all data are public immediately.
-%
-%{\bf For all of these reasons, we need to keep building up the sample of GRB
-%	afterglow spectra now as we may have to wait many years before a mission like
-%	\textit{Swift} becomes available again.}
-%
-%A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark
-%	bursts}'', e.g., Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012,
-%MNRAS, 421, 1265). Some of these are at the highest redshifts ($z > 6$) and
-%their observer-frame optical emission is absorbed by the IGM. The majority,
-%however, suffer from large dust obscuration (e.g., Perley et al. 2009, AJ, 138,
-%1690; Greiner et al.\ 2011, A\&A, 526, 30). Identifying such GRBs is important
-%for constraining the fraction of obscured star formation. In both cases, NIR
-%emission is expected. X-shooter can adequately study these objects, provided
-%that a NIR counterpart is timely identified, for which we have the dedicated
-%HAWKI run D.
-%
-%The detection of {\bf absorption line variability} can reveal the burst
-%influence on the surrounding medium and in turn the absorber distance from the
-%burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
-%2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\ 2012,
-%A\&A, 545, 64; Hartoog et al.\ 2013). {\bf Short GRBs} originate in a
-%substantially different environment compared to long GRBs. Short GRBs may be
-%related to the merging of compact binaries and the coalescence time can be long
-%enough to allow the progenitor system to move far away from the star formation
-%site (Belczynski et al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum
-%with a sufficient signal-to-noise ratio of a short GRB afterglow has been
-%secured. A knowledge of the redshift distribution of short bursts is of key
-%importance for the next generation of gravitational wave experiments, as they
-%are the likely EM counterparts to their primary targets.
+GRBs are interesting as they provide the opportunity to derive new insights
+into a very wide range of topics in astrophysics. In that sense GRB
+research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
+which was launched in 2004, has made
+it possible to harvest the great potential in using GRBs as probes of
+galaxies in the intergalactic medium, which was already hinted at by
+results from earlier missions.
+
+A crucial aspect of the great success of the {\it Swift} mission has been
+the extensive ground based follow-up observations of the afterglows and of
+the host galaxies of the GRBs. This work has involved a large community
+of researchers both inside and outside the official {\it Swift} team. This fruitful
+collaboration has been greatly facilitated by the open data access policy
+of the {\it Swift} mission.
+
+In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
+was secured using low-resolution spectrographs (typically
+R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
+very bright afterglows high-resolution (typically R$>$20000) spectra were
+secured.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
+instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
+echelle spectrograph
+covering the full spectra range from the UV atmospheric cutoff to the near-infrared
+(near-IR)
+K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
+transient source follow-up in mind as the fading luminosities of such sources 
+makes in urgent to secure as extensive coverage as possible in the shortest possible 
+time. At the same time the resolution was set to be in the range 4000--9000 in order
+to be able to get a large useful spectra range between the many sky-background
+emission lines in the red and near-IR.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison to previous samples.}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This should also include a discussion of dark bursts. 
-Cuchiara. BAT6 (Salvaterra), TOUGH
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{X-ray column densities vs. optically derived ones (Remove section)}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
-%
-%\begin{figure}
-%	\centerline{\includegraphics[width=10cm]{figures/OAXR_NH_sphere.pdf}}
-%	\caption{Optically derived HI column densities vs.
-X-ray derived ones taken from \citet{Buchner2016}.
-The colors of the points reflect the redshift of the burst.
-Circles mark the column densities published in \citet{Fynbo2009} and squares mark the ones derived as part of the sample presented here.
-Overplot in blue is the 1:1 correspondence, expected if the conditions local to the GRBs are similar to the local ISM abundances.
-If the metallicity is lower at the GRB site, the point move upwards in the plot.
-}
-%	\label{fig:OAXR}
-%\end{figure}	
-%	
-%	
-%	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Why is the HI column truncated? (Remove section)}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%What is the expected transmission for a high-column HI, where the hydrogen primarily will be in the form of H2?
-Number of bursts with detected H2 and their HI column density.
-Rahmati Schaye, Krongold Prochaska.
+\section{Discussion and conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper we have presented the results
+of a dedicated effort over the years 2009 -- 2016 to use the X-shooter 
+spectrograph on the ESO-VLT to secure spectroscopic observations of 
+afterglows and host galaxies of GRBs detected by {\it Swift}. This 
+work was initiated by a consortium using Guaranteed Time on X-shooter, 
+but over the years the project continuued in open time and the team was
+opened to include any researchers interested in contribution to this 
+effort.
+
+The sample we present here include spectroscopic observations of 86 
+systems fulfilling our sample criteria of which 72 are afterglow spectra 
+and 14 are observations of the underlying host galaxies. 
+UPDATE THIS.
 

### 2017-10-24 document/intro.tex
Merge branch 'master' into writing_branch
# Conflicts:
#	document/XSGRB_sample.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ QSOs have 5 systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
\citep{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
systems in the majority of cases \citep{Jakobsson2006, Fynbo2009}. Whereas DLAs
towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
star-forming galaxies over the full history of cosmic star formation from $z
\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
	Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we can hence characterize the properties of
star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, etc.. This is, however,
only possible as long as there are satellites in orbit that rapidly and
accurately locate GRBs. The currently operating \textit{Swift} satellite,
launched in 2004 and still fully-functioning, allows for very efficient
follow-up observations of GRBs due to its unprecedented rate, speed and
precision of localisations.

There are several unanswered but fundamental questions that must be addressed in
order to exploit the full potential of GRBs as cosmological probes. More than
50\% of the \textit{Swift} bursts with measured redshift are at $z > 2$, and 
5--10\% are expected to be above $z=5$ \citep{Salvaterra2007,
	Salvaterra2012,Jakobsson2012, Perley2016}. A high redshift completeness is
crucial for our understanding of the link between the number density of GRBs per
unit redshift and the global star-formation history of the Universe, as measured
by other means \citep[UV, FIR, sub-mm, see][]{Robertson2012}. The detection of
GRBs at $z > 6$ shows that GRBs have become competitive as a tool to identifying
galaxies at the highest redshifts and unsurpassed in providing detailed
abundance information via absorption line spectroscopy \citep{Tanvir2012,
	McGuire2016}.

\smallskip

%The sample of short GRBs with redshifts remains too small and too
%incomplete to really
%constrain their redshift distribution (Guetta \& Piran 2006, A\&A, 453, 823;
%Salvaterra et al.\ 2008, MNRAS, 388, 6) making it of extreme importance to
%enlarge as far as possible the number of known redshifts for this class of
%sources.

From March 2005 to March 2016 there have been about 350 {\it Swift} bursts 
complying with
our sample selection criteria (see ``Immediate objective''), and about half of them have
measured redshifts. Among the latter subset, the team proposing these observations has
measured about two thirds of the redshifts, mainly with FORS1/2 and X-shooter (see
Fynbo et al.\ 2009, ApJS, 185, 526;  Jakobsson et al.\ 2012, ApJ, 752, 62;
Kr\"uhler et al.\ 2012, ApJ, 758, 46). Our current aim is to build a sample superior to
our previous low-resolution survey (Fynbo et al.\ 2009, ApJS, 185, 526), both in
terms of quantity and  quality of the spectra. Our program started as 
guaranteed time observations during
periods 84-91, and we have continued in open time since then.

%In P84 we started the X-shooter legacy survey in GTO time. Our aim is to  build
%a sample of GRBs superior to our previous  low-resolution  survey, Fynbo et al.\
%2009, ApJS, 185, 526) both in terms of quantity and  quality of the spectra. The
%last period of GTO time on the survey was P90 (plus 6 hr of leftover GTO time
%in  P91) and for P92 a proposal like this was  accepted for continuation of the
%program in open time (concerning P93 see box 8).

\smallskip

X-shooter is in many ways
the ideal GRB follow-up instrument and indeed GRB follow-up was one of the
primary science cases behind the instrument design and implementation. Our
program secures general purpose GRB afterglow spectroscopic follow-up 
that adds strong legacy value to the \textit{Swift} GRB sample.
Due to the wide wavelength range of X-shooter  
with the same observation cover molecular H$_2$ absorption near the
atmospheric cut-off and all the strong emission lines from the host in the
NIR arm (e.g., Friis et al., 2015, MNRAS, 451, 167). In general, the wide wavelength coverage ensures that we always have
features on which to base a redshift measurement as long as the afterglow is
brighter than about 23 mag in either the $R$- or $z$-band. 
Frequently, emission lines are also detected from the underlying host, which
also provide further information such as SFR and metallicity (the top right 
panel in Fig.~1 shows an example). Only for 
7 out of more than 70 secured spectra could we not measure a redshift.   
With the X-shooter survey we
provide {\bf metallicity measurements} for about 30\% (Voigt-profile fits)
of the $z>1.7$ events.
So far we have
measured metallicities for more than 20 GRB afterglows with X-shooter.
With the wide wavelength coverage of X-shooter we can
study important chemical species as Zn\,\textsc{ii}, Cr\,\textsc{ii} and $\alpha$
elements over a much wider redshift range than what is possible with other
instruments.
As an example, we have measured a metallicity of $0.1 Z_\odot$ for GRB\,100219A
at $z=4.669$ (Th{\"o}ne et al.\ 2013, MNRAS, 428, 3590), $0.02Z_\odot$ for 
GRB\,111008A at $z=4.991$ (Sparre et al.\ 2014, ApJ, 785, 150) and $0.05Z_\odot$
for the $z=5.9125$ GRB\,130606A (Hartoog et al.\ 2015, A\&A, 580, 139).  
Reconciling the
abundance patterns of GRB absorbers, other types of absorbers, QSO DLAs
in particular, and old stars in the Local Group is an important long-term
goal (see also Sparre et al.\ 2014, ApJ, 785, 150).
Metallicities are also measured from host emission lines (Kr{\"u}hler et al.\ 2015,
A\&A, 581, A125).
GRB spectroscopy also allows us to determine the dust content of their environments,
both through analysis of the depletion pattern and through measurement of the
associated extinction (Japelj et al.\ 2015, A\&A, 451, 2050). 
This allows us to quantify the dust-to-metals ratio and its
evolution with redshift (e.g., De Cia et al.\ 2013, A\&A, 560, 88; Zafar \&
Watson 2013, A\&A, 560, 26).
%Moreover, GRB
%afterglows allow investigation of random foreground absorbers, complementing
%QSO sightline studies (e.g., Schulze et
%al.\ 2012, A\&A, 546, 20).

%In Fynbo et al. (2009, ApJS, 185,
%526), we compare the host-galaxy HI
%column density and metallicity (most of which has been measured by our group)
%with a sample of damped \lya\ systems (DLAs) along QSO sightlines. We show that
%most GRB-DLAs have a much larger column than QSO-DLAs.  The difference in
%column density and metallicity for GRB and QSO sightlines shows that the total
%cross-section for environments like those of GRBs must be less than a few
%percent of the total cross-section for QSO-DLAs, and suggests that GRB
%sightlines probe regions with high gas density.

%The issue of production and {\bf release of Lyman continuum radiation} is
%currently not fully understood, but the evidence suggests that star-forming
%galaxies are dominating, in particular at the highest redshifts (e.g., Bianchi
%et al.\ 2001, A\&A, 376, 1; Chen et al.\ 2007, ApJ, 667, L25). Also here
%X-shooter spectroscopy will allow major progress due to the excellent blue
%sensitivity.

%GRB afterglows can probe dust in high-redshift star-forming galaxies
%including exploring the {\bf 2175~\AA{} extinction bump}
%(El\'iasd\'ottir et al.\ 2009, ApJ, 697, 1725; Zafar et al.\ 2011, A\&A,
%732, 143).
%Due to the broad wavelength  coverage we can do
%this more accurately and over a broader range of redshifts with X-shooter.

\smallskip

We will also determine the frequency and properties of {\bf molecular
	absorption} towards GRB absorbers. Molecular gas is a key element to catalyze
the process of star formation, but prior to our program H$_2$ had been detected
just in two cases (tentatively in Fynbo et al.\ 2006, A\&A, 451, L47; securely
in Prochaska et al.\ 2009, ApJ, 691, L27). With our X-shooter program we have
found three more systems (Kr\"uhler et al.\ 2013, A\&A, 557, 18; D'Elia et al.\
2014, A\&A, 564, 38; Friis et al.\, 2015, MNRAS, 451, 167). 
We are currently analysing more of our spectra for less obvious molecular
absorption and we expect to find more (a dedicated sample paper is 
addressing this issue).

\vspace{0.2cm}
A natural question to ask is: {\bf how long should this work continue?} Our
view is that we need to keep observing the afterglows as long as we have
\textit{Swift} in operation. Also note that the program is still producing many 
papers and provides data for many theses (Box 9 and 10). \textit{Swift} is currently 
funded until 2018, but is likely to get more extensions given its overwhelming 
success. As
mentioned above GRBs allow us to probe star-forming galaxies that are almost
impossible to study in other ways both in terms of redshifts, galaxy luminosity
function, and regions within galaxies. After 7 periods we have 
secured seven spectra of $z>4$ GRBs, of which three were of sufficient quality to
allow abundance measurements (Th\"one et al.\ 2013, Sparre et al.\ 2014, 
Hartoog et al.\ 2015). GRBs offer the only way to derive chemical abundances 
for the gas phase of central, actively star-forming regions of high-$z$ galaxies. 
The program also maintains a very high discovery
potential where we occasionally find something completely unexpected that 
provides interesting clues to puzzles in other fields, e.g. extinction of 
type Ia supernovae (Fynbo et al.\ 2014, A\&A, 572, 12). Each of these spectra are like
precious jewels – it is a type of observation that can never be repeated and a
class of sightlines that can only be studied while we have operating GRB
satellites.

It is also worth adding that we have build up a rather unique team spread over
Europe from Granada to Reykjavik, which by now has reached a point where the 
distribution of night shifts, the scientific exploitation of the data is 
efficient and where we are open to all new members who wish to participate.
As mentioned all data are public immediately.

{\bf For all of these reasons, we need to keep
	building up the sample of GRB afterglow spectra now as we may have to wait many
	years before a mission like \textit{Swift} becomes available again.}

A significant proportion of GRBs lack a bright optical afterglow (``{\bf dark bursts}'', e.g.,
Jakobsson et al.\ 2004, ApJ, 617, L21; Melandri et al.\ 2012, MNRAS, 421, 1265). 
Some of these are at the highest
redshifts ($z > 6$) and their observer-frame optical emission is absorbed by
the IGM. The majority, however, suffer from large dust obscuration (e.g., Perley
et al. 2009, AJ, 138, 1690; Greiner et al.\ 2011, A\&A, 526, 30).
Identifying such GRBs is important for
constraining the fraction of obscured star formation. In both cases,
NIR emission is
expected. X-shooter
can adequately study these objects, provided that a NIR counterpart is timely
identified, for which we have the dedicated HAWKI run D.

The detection of {\bf absorption line variability} can reveal the burst
influence on the surrounding medium and in turn the absorber distance from the
burst and its metallicity (Vreeswijk et al.\ 2007, A\&A 468, 83; D'Elia et al.\
2009, ApJ, 694, 332; Th\"one et al.\ 2011, MNRAS, 414, 479; De Cia et al.\
2012, A\&A, 545, 64; Hartoog et al.\ 2013).
{\bf Short GRBs} originate in a substantially different
environment compared to long GRBs. Short GRBs may be related to the merging of
compact binaries and the coalescence time can be long enough to allow the
progenitor system to move far away from the star formation site (Belczynski et
al.\ 2002, ApJ, 571, 147). Up to now, however, no spectrum with a sufficient 
signal-to-noise ratio
of a short GRB afterglow has been secured.
A knowledge of the redshift distribution
of short bursts is of key importance for the next generation of gravitational
wave experiments, as they are the likely EM counterparts to their
primary targets.

### 2017-10-24 document/XSGRB_sample.tex
Intro work
More work on introduction and conclusions

--- 2017-10-24 document/XSGRB_sample.tex
@@
 
-GRBs are interesting as they provide the opportunity to derive new insights
-into a very wide range of topics in astrophysics. In that sense GRB
-research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
-which was launched in 2004, has made
-it possible to harvest the great potential in using GRBs as probes of
-galaxies in the intergalactic medium, which was already hinted at by
-results from earlier missions.
+Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
+derive new insights into a very wide range of topics in astrophysics. Examples
+range from small scale phenomena relating to the properties of highly
+relativistic jets, supernova explosions, the interstellar medium, dust extinction
+curves, starburst galaxies, chemical and molecular abundances
+the ionization state of the intergalactic medium, intervening
+absoption systems to standard candles in cosmology.  In that sense GRB research
+is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, which
+was launched in 2004, has made it possible to harvest the great potential in
+using GRBs as probes of galaxies in the intergalactic medium, which was already
+hinted at by results from earlier missions \citep[e.g.][]{Ricker04}.
 
@@
 was secured using low-resolution spectrographs (typically
-R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
+R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution 
+spectroscopy is powerful for a number of reasons, first of all is the ability 
+to secure information for even very faint targets. This allows the 
+measurements of a number of important parameters such as redshifts and spectral
+slopes. For a handful of
 very bright afterglows high-resolution (typically R$>$20000) spectra were
-secured.
+secured and for these events much more information about conditions inside the
+host galaxies were extracted 
+\citep[e.g.,]{}{Fiore, Thoene07, Prochaska07, Vreejswijk07}.
 
@@
 
+Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+towards about 10$^5$ QSOs have a handful such systems with
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
+\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
+reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
+Whereas DLAs
+towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
+atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
+\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
+star-forming galaxies over the full history of cosmic star formation from $z
+\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
+    Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we can hence characterize the properties of
+star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
+molecular contents, ISM temperatures, UV-flux densities, etc.
+
+Bias still, more to do, ELTs, SVOM
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-10-25 document/XSGRB_sample.tex
More on intro
I have written more on the introduction.

@@
 derive new insights into a very wide range of topics in astrophysics. Examples
-range from small scale phenomena relating to the properties of highly
-relativistic jets, supernova explosions, the interstellar medium, dust extinction
-curves, starburst galaxies, chemical and molecular abundances
-the ionization state of the intergalactic medium, intervening
+range from small scale phenomena relating to the magnetars, properties of highly
+relativistic jets, hyper/supernova explosions, the interstellar medium, dust extinction
+curves, starburst galaxies, chemical and molecular abundances, escape of ionizing 
+radiation, the ionization state of the intergalactic medium, intervening
 absoption systems to standard candles in cosmology.  In that sense GRB research
-is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, which
-was launched in 2004, has made it possible to harvest the great potential in
-using GRBs as probes of galaxies in the intergalactic medium, which was already
-hinted at by results from earlier missions \citep[e.g.][]{Ricker04}.
-
-A crucial aspect of the great success of the {\it Swift} mission has been
-the extensive ground based follow-up observations of the afterglows and of
-the host galaxies of the GRBs. This work has involved a large community
-of researchers both inside and outside the official {\it Swift} team. This fruitful
-collaboration has been greatly facilitated by the open data access policy
-of the {\it Swift} mission.
+is very multidisciplinary.
+
+The {\it Swift} satellite \citep{Gehrels09}, which was launched in 2004, has
+made it possible to harvest much of the great potential in using GRBs as probes
+of galaxies in the intergalactic medium, which was already hinted at by results
+from earlier missions \citep[e.g.][]{Ricker04}.  A crucial aspect of the great
+success of the {\it Swift} mission has been the extensive ground based
+follow-up observations of the afterglows and of the host galaxies of the GRBs.
+This work has involved a large community of researchers both inside and outside
+the official {\it Swift} team. This fruitful collaboration has been greatly
+facilitated by the open data access policy of the {\it Swift} mission.
 
@@
 spectroscopy is powerful for a number of reasons, first of all is the ability
-to secure information for even very faint targets. This allows the 
-measurements of a number of important parameters such as redshifts and spectral
-slopes. For a handful of
-very bright afterglows high-resolution (typically R$>$20000) spectra were
-secured and for these events much more information about conditions inside the
-host galaxies were extracted 
-\citep[e.g.,]{}{Fiore, Thoene07, Prochaska07, Vreejswijk07}.
-
-The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
-instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
-echelle spectrograph
-covering the full spectra range from the UV atmospheric cutoff to the near-infrared
-(near-IR)
-K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
-transient source follow-up in mind as the fading luminosities of such sources 
-makes in urgent to secure as extensive coverage as possible in the shortest possible 
-time. At the same time the resolution was set to be in the range 4000--9000 in order
-to be able to get a large useful spectra range between the many sky-background
-emission lines in the red and near-IR.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria and observations}
+to secure information for even very faint targets. This allows the measurements
+of a number of important parameters such as redshifts and spectral slopes. For
+a handful of very bright afterglows high-resolution (typically R$>$20000)
+spectra were secured and for these events much more information about
+conditions inside the host galaxies were extracted \citep[e.g.,]{}{Fiore,
+Thoene07, Prochaska07, Vreejswijk07, CastroTirado2010}.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
+generation instruments at the ESO Very Large Telescope (VLT). It is a
+high-throughput echelle spectrograph covering the full spectra range from the
+UV atmospheric cutoff to the near-infrared (near-IR) K-band with three spectral
+arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
+in mind as the fading luminosities of such sources makes in urgent to secure as
+extensive coverage as possible in the shortest possible time. At the same time
+the resolution was set to be in the range 4000--9000 in order to be able to get
+a large useful spectra range between the many sky-background emission lines in
+the red and near-IR spectral ranges. In this paper we present the results
+of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
+spectrograph to secure spectroscopic observations of
+afterglows and host galaxies of GRBs detected by {\it Swift}. 
+All data resulting from the survey are publicly availble in reduced form.
+
+The paper is
+organized in the following way: In Sect.~\ref{sample} we describe the sample
+including the sample selection criteria and the observational strategy.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sample selection criteria and observations}\label{sample}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Bias still, more to do, ELTs, SVOM
+Bias still, more to do, ELTs, SVOM unharvested potential Kruehler? 2012, GRB08607
 
@@
 \begin{acknowledgements}
+We should add an ode to Neil Gehrels here.
 JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The

### 2017-10-25 document/XSGRB_sample.tex
further work on intro and conclusions

@@
 
-The paper is
-organized in the following way: In Sect.~\ref{sample} we describe the sample
-including the sample selection criteria and the observational strategy.
+The paper is organized in the following way: In Sect.~\ref{sample} we describe
+the sample including the sample selection criteria and the observational
+strategy.  In Sec.t~\ref{results} we describe the results of the survey, i.e.
+the efficiency of the follow-up effort and the characteristics of the observed
+bursts. We also here assess the completeness of the finally realized sample.
+Finally, we offer our conclusions in Sect.~\ref{conclusions}.
 
@@
 
-X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800 \AA~in
-a single exposure, by splitting the light in three separate spectroscopic arms
-using dichroics.
-This way each arm work as a separate instrument, each functioning as its own
-echelle spectrograph.
-The ultraviolet blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm
-covering 5500 - 10200 \AA, and the near-infrared()NIR) arm covering 10200 -
-24800 \AA~ with the possibility of applying a k-band
-blocking filter cutting the coverage of the NIR arm at 10200 - 21000 \AA.
+X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800
+\AA~in a single exposure, by splitting the light in three separate
+spectroscopic arms using dichroics.  This way each arm work as a separate
+instrument, each functioning as its own echelle spectrograph.  The ultraviolet
+blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 -
+10200 \AA, and the near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the
+possibility of applying a k-band blocking filter cutting the coverage of the
+NIR arm at 10200 - 21000 \AA.
 \\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this section we describe the efficiency of the follow-up effort and the characteristics of the observed bursts. We also assess the degree to which the sample obtained is complete.
+\section{Results}\label{results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we describe the efficiency of the follow-up effort and the
+characteristics of the observed bursts. We also assess the degree to which the
+sample obtained is complete.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion and conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this paper we have presented the results
-of a dedicated effort over the years 2009 -- 2016 to use the X-shooter 
-spectrograph on the ESO-VLT to secure spectroscopic observations of 
-afterglows and host galaxies of GRBs detected by {\it Swift}. This 
-work was initiated by a consortium using Guaranteed Time on X-shooter, 
-but over the years the project continuued in open time and the team was
-opened to include any researchers interested in contribution to this 
+\section{Discussion and conclusions}\label{conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper we have presented the results of a dedicated effort over the
+years 2009 -- 2016 to use the X-shooter spectrograph on the ESO-VLT to secure
+spectroscopic observations of afterglows and host galaxies of GRBs detected by
+{\it Swift}. This work was initiated by a consortium using Guaranteed Time on
+X-shooter, but over the years the project continuued in open time and the team
+was opened to include any researchers interested in contribution to this
 effort.
@@
 
+Our sample serves the purpose to further characterize the environments 
+of GRBs that was also much advanced by the previous surveys based primarily on 
+lower-resolution spectographs. The most striking result found is that GRBs
+typically are located behind very large column densities of neutral hydrogen.
 Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ QSOs have a handful such systems with
+towards about 10$^5$ quasars have a handful such systems with
 $\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
@@
 systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
-Whereas DLAs
-towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
-atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
-\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
-star-forming galaxies over the full history of cosmic star formation from $z
-\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
-    Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we can hence characterize the properties of
+With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we hence characterize the properties of
 star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
@@
 
-Bias still, more to do, ELTs, SVOM unharvested potential Kruehler? 2012, GRB08607
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+It seems important to end this paper by stressing that the potential of 
+using GRB sightlines as probes is far from fully harvested. The sample of
+sightlines probed by our spectra is not representative for all GRB 
+sightlines as we have shown (CONFIRM) and consistent with earlier findings
+\citep[e.g.,][]{Fynbo2009}. \cite{Kruhler2013} convincingly argue, that 
+very rich sightlines like that probed by the remarkable GRB\,080607 
+\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
+sampled by our spectra. However, with current instrumentation these sightlines are
+out of reach except under very fortunate circumstances as in the case of 
+GRB\,080607 when the afterglow could be observed only a few minutes after the 
+burst with a 10-m class telescope. Observations of such sightlines with 
+X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
+is likely to be very rewarding.
 
 \begin{acknowledgements}
-We should add an ode to Neil Gehrels here.
+*** We should add an ode to Neil Gehrels here ***.
 JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The

### 2017-10-25 document/XSGRB_sample.tex
More
More

@@
 spectra were secured and for these events much more information about
-conditions inside the host galaxies were extracted \citep[e.g.,]{}{Fiore,
-Thoene07, Prochaska07, Vreejswijk07, CastroTirado2010}.
+conditions inside the host galaxies were extracted 
+\citep[e.g.,][]{Fiore,Thoene07,Prochaska07,Vreejswijk07,CastroTirado2010}.
 

### 2017-10-25 document/XSGRB_sample.tex
Final today
Conclusions

@@
 
+All spectra have been made publicly available in the reduced form used
+in this paper.
+
+A number of independent papers have been published or submitted for publication
+focusing on specific 
+issues such as escape of ionizing radiation \citep{Tanvir17}, extinction
+curves \citep{Jure16,Zafar17}, emission lines from the underlying host 
+galaxies \citep{Kruhler2015}, more here.
+
 It seems important to end this paper by stressing that the potential of 

### 2017-10-26 document/XSGRB_sample.tex
Update 26/10
More intro and conclusions
AA changed to nm

--- 2017-10-25 document/XSGRB_sample.tex
@@
 In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
-was secured using low-resolution spectrographs (typically
-R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution
+was secured using low-resolution spectrographs 
+\citep[typically
+R=$\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Low-resolution
 spectroscopy is powerful for a number of reasons, first of all is the ability
 to secure information for even very faint targets. This allows the measurements
-of a number of important parameters such as redshifts and spectral slopes. For
+of a number of important parameters such as redshifts, spectral slopes,
+and extinction. For
 a handful of very bright afterglows high-resolution (typically R$>$20000)
@@
 generation instruments at the ESO Very Large Telescope (VLT). It is a
-high-throughput echelle spectrograph covering the full spectra range from the
-UV atmospheric cutoff to the near-infrared (near-IR) K-band with three spectral
+high-throughput echelle spectrograph covering the full spectral range from the
+UV atmospheric cutoff around 300 nm to the near-infrared (near-IR) K-band 
+ending around 2500 nm with three spectral
 arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
@@
 the resolution was set to be in the range 4000--9000 in order to be able to get
-a large useful spectra range between the many sky-background emission lines in
+a large useful spectral range between the many sky-background emission lines in
 the red and near-IR spectral ranges. In this paper we present the results
@@
 
-X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800
-\AA~in a single exposure, by splitting the light in three separate
+X-shooter can cover the spectral wavelength region from 300 nm to 2480
+nm in a single exposure, by splitting the light in three separate
 spectroscopic arms using dichroics.  This way each arm work as a separate
 instrument, each functioning as its own echelle spectrograph.  The ultraviolet
-blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 -
-10200 \AA, and the near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the
+blue(UVB) arm covering 3000 - 550 nm, the visual(VIS) arm covering 550 -
+1020 nm, and the near-infrared()NIR) arm covering 10200 - 2480 nm with the
 possibility of applying a k-band blocking filter cutting the coverage of the
-NIR arm at 10200 - 21000 \AA.
+NIR arm at 1020 - 2100 nm.
 \\
@@
 The image is rectified onto an equidistant grid with a dispersion sampling of
-0.2 \AA/pixel and a 0.16 "/pixel spatial sampling for the UVB and VIS arm and
-0.6 \AA/pixel with a 0.21 "/pixel in the NIR arm.  Because the tabulated
-resolution is a lower limit, by choosing a sampling of 0.2 \AA/pixel, we ensure
+0.02 nm/pixel and a 0.16 "/pixel spatial sampling for the UVB and VIS arm and
+0.06 nm/pixel with a 0.21 "/pixel in the NIR arm.  Because the tabulated
+resolution is a lower limit, by choosing a sampling of 0.02 nm/pixel, we ensure
 that the bluest part of neither of the arms have a sampling lower than the
@@
 
-The resulting offsets, which were smaller than 0.1 \AA~in the UVB and VIS data
-and smaller than 0.5 \AA~in the NIR spectra, but changing over short period of
+The resulting offsets, which were smaller than 0.01 nm in the UVB and VIS data
+and smaller than 0.05 nm in the NIR spectra, but changing over short period of
 time were applied to the corresponding spectra.
@@
 UPDATE THIS.
-
-Our sample serves the purpose to further characterize the environments 
-of GRBs that was also much advanced by the previous surveys based primarily on 
+All spectra have been made publicly available in the reduced form used
+in this paper.
+
+Our sample serves the purpose to further characterize the environments of GRBs
+that was also much advanced by the previous surveys based primarily on
 lower-resolution spectographs. The most striking result found is that GRBs
@@
 towards about 10$^5$ quasars have a handful such systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
-reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
-With afterglow spectroscopy (throughout the electromagnetic
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
+in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006,
+Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
 spectrum from X-rays to the sub-mm) we hence characterize the properties of
 star-forming galaxies over cosmic history in terms of redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities, etc.
+molecular contents, ISM temperatures, UV-flux densities, extinction curves,
+etc.  A number of independent papers have been published or submitted for
+publication focusing on many of these specific issues of our sample such as
+extinction curves [][see also
+\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
+underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
+\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
+radiation \citep{Tanvir17}.
 
@@
 sightlines probed by our spectra is not representative for all GRB 
-sightlines as we have shown (CONFIRM) and consistent with earlier findings
-\citep[e.g.,][]{Fynbo2009}. \cite{Kruhler2013} convincingly argue, that 
+sightlines as we have shown and consistent with earlier findings
+from samples based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009}
+and from studies of complete samples of GRB host galaxies 
+\citep{Hjorth2012,Covinto2013,Perley2016}. 
+\cite{Kruhler2013} convincingly argue, that 
 very rich sightlines like that probed by the remarkable GRB\,080607 
@@
 trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting at
-$\sim$ 7000 \AA~and extending all the way through 25000 \AA. We detect a $\sim
+$\sim$ 700 nm~and extending all the way through 2500 nm. We detect a $\sim
 2 \sigma$ emission line which, if interpreted as \ha, gives $z = 1.10$,
@@
 published elsewhere. 4.8 ks integration time, starting 10.5 hours after the BAT
-trigger, shows a bright afterglow continuum starting at $\sim$ 4720 \AA,
+trigger, shows a bright afterglow continuum starting at $\sim$ 472 nm,
 signifying the onset of the Lyman alpha forest, for a GRB located at $z =
@@
 observation and does not contain any afterglow continuum. In 3.6 ks integration
-time, we detect a faint continuum visible from 4500 \AA~and all the way through
-21000 \AA, in contrast to what is found previously. The continuum from 4500 -
-6000 \AA is detected at very low significance. If the drop at 4500 \AA~is the
-Lyman limit, this fits with Lyman alpha at $\sim$ 6000 \AA, giving $z \sim
+time, we detect a faint continuum visible from 450 nm and all the way through
+2100 nm, in contrast to what is found previously. The continuum from 4500 -
+600 nm is detected at very low significance. If the drop at 450 nm is the
+Lyman limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim
 3.9$. The absence of nebular lines if due to \oii~falling in a telluric
@@
 BAT trigger, this 1.2 ks observation captures a very bright afterglow
-continuum, starting at 4500 \AA, signifying the onset of the Lyman limit for a 
+continuum, starting at 450 nm, signifying the onset of the Lyman limit for a 
 system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
@@
 of observations reveals a faint continuum rising abruptly above the noise at
-$\sim$ 6850 \AA~and continuing through 21000 \AA. A very low significance
-continuum is detected at shorter wavelengths, down to $\sim$ 6000 \AA. If the
+$\sim$ 685 nm and continuing through 2100 nm. A very low significance
+continuum is detected at shorter wavelengths, down to $\sim$ 600 nm. If the
 host is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha
@@
 out of the wavelength coverage. Alternatively, an early-type host at $z = 0.71$
-could exhibit the 4000 \AA break at 6000 \AA, but due to the preference of
+could exhibit the 400 nm break at 600 nm, but due to the preference of
 long-duration GRBs for star-forming galaxies, this is the least likely
@@
 \aliii, and	\feii, all at $z =2.823$. Similar to GRB140506 \citep{Fynbo2014}, a
-break in the continuum shape is detected bluewards of 6000\AA, possible
+break in the continuum shape is detected bluewards of 600 nm, possible
 signifying some anomalous form of extinction.

### 2017-10-26 document/XSGRB_sample.tex
Adding continuum estimate

--- 2017-10-25 document/XSGRB_sample.tex
@@
 
-\title{\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample}
+\title{XSGRB - The X-shooter GRB afterglow sample
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}
-
-\titlerunning{XSGRB sample}
-
-\author{
-	J.~Selsing\inst{1}, 
-	T.~Kr\"{u}hler\inst{1}, 
-	P.~Goldoni\inst{14}, 
-	D.~Malesani\inst{1}, 
-	J.~P.~U. Fynbo\inst{1}, 
-	A.~de~Ugarte~Postigo\inst{11}, 
-	J.~Japelj\inst{20},
-	P.~D'Avanzo,
-	Z.~Cano,
-	S.~Covino\inst{10}, 
-	V.~D'Elia\inst{7, 12}, 
-	H.~Flores,
-	O.~E.~Hartoog\inst{6},
-	J.~Hjorth\inst{1}, 
-	P.~Jakobsson\inst{5}, 
-	A.~Levan,
-	A.~Melandri,
-	S.~Piranomonte \inst{7},
-	R.~S\'anchez-Ram\'irez\inst{11},
-	S.~Schulze\inst{17, 18}, 
-	N.~R.~Tanvir\inst{19},
-	C.~Th{\"o}ne,
-	S.~D.~Vergani\inst{7, 8},
-	P.~M.~Vreeswijk\inst{3}, 
-	D.~J.~Watson\inst{1},
-	K.~Wiersema\inst{19},
-	D.~Xu\inst{1}
-	L.~Christensen\inst{1},
-	A.~De~Cia\inst{3}, 
-	L.~Kaper\inst{6}, 
-	L.~A.~Antonelli,
-	F.~Fiore,
-	A.~Gomboc,
-	P.~Groot,
-	F.~Hammer,
-	C.~Ledoux\inst{2}, 
-	E.~Maiorano,
-	B.~Milvang-Jensen\inst{1}, 
-	E.~Palazzi,
-	E.~Pian,
-	J.~Schaye,
-	G.~Tagliaferri\inst{7},
-	R.~A.~M.~J.~Wijers\inst{6}
+	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}}
+
+\titlerunning{The X-shooter sample of GRBs}
+
+\input{authorlist.tex}
+
+\date{Received/ accepted}
+
+\authorrunning{Selsing et al.}
+
+\abstract{
+In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
+afterglows that have \textit{ever} been observed at the VLT/X-shooter up to the end of 2016. This
+totals 137 spectroscopic observation of 118 individual bursts, of which
+redshifts have been secured for 91 \% covering a redshift range from 0.059 to
+$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
+has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
+of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
+We here provide a public release of all the spectra, including continuum
+estimates and telluric corrections. We provide an assessment of the degree of
+completeness with respect to the parent GRB population, in terms of the X-ray
+properties of the bursts in the sample and find that the sample presented here
+can not be significantly distinguished from the \textit{Swift} sample.
+For all bursts (39) for which it is possible, we provide a measurement of the
+optically derived hydrogen column density, increasing the total number of
+published HI column densities by $\sim$ 50 \% and characterize the distribution
+compared with literature values.
+We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and tentatively confirm previous
+results that larger optical darkness leads to increased X-ray absorption.
+%This dataset provides a unique possibility to study the ISM across cosmic time.
 }
 
-\authorrunning{Selsing et al.}
-
-\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
-\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
-\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
-\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
-\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
-\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
-\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
-\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
-\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
-\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
-\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
-\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
-\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
-\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/
-	IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et L
-	eonie Duquet, 75205 Paris Cedex 13, France
-\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachs
-	tra\ss e, 85748 Garching, Germany
-\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3,
-	20126, Milano, Italy
-\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astro
-	nom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
-\and Millennium Center for Supernova Science
-\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
-\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI}
-
-\date{Received/ accepted}
-
-\abstract{
-The {\it Swift} satellite allows us to use gamma-ray bursts (GRBs) to peer into 
-the hearts of star forming galaxies through cosmic time. Our open
-collaboration, representing most of the active ESO member researchers in this
-field, seeks to build a public legacy sample of GRB X-shooter spectroscopy
-while {\it Swift} continues to fly. We propose to continue our programme to target
-all suitably observable GRB afterglows (up to 15 bursts per semester), with the
-primary goal of producing a well-defined, homogeneous, statistically useful
-sample. To date, our spectroscopy covers a redshift range from 0.059 to about
-8, with more than 20 robust metallicity measurements from absorption lines (over 
-the redshift range 1.7--5.9) and 4 secure detections of H$_2$ or CH molecular
-absorption. Such information is extremely difficult to obtain by other means.
-In terms of studying the spread and redshift evolution in gas-phase properties, the
-sample is still limited by low-number statistics. 
-}
-
 \keywords{Gamma-ray burst: individual: GRB~120815A --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
@@
 
-Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
-derive new insights into a very wide range of topics in astrophysics. Examples
-range from small scale phenomena relating to the magnetars, properties of highly
-relativistic jets, hyper/supernova explosions, the interstellar medium, dust extinction
-curves, starburst galaxies, chemical and molecular abundances, escape of ionizing 
-radiation, the ionization state of the intergalactic medium, intervening
-absoption systems to standard candles in cosmology.  In that sense GRB research
-is very multidisciplinary.
-
-The {\it Swift} satellite \citep{Gehrels09}, which was launched in 2004, has
-made it possible to harvest much of the great potential in using GRBs as probes
-of galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.][]{Ricker04}.  A crucial aspect of the great
-success of the {\it Swift} mission has been the extensive ground based
-follow-up observations of the afterglows and of the host galaxies of the GRBs.
-This work has involved a large community of researchers both inside and outside
-the official {\it Swift} team. This fruitful collaboration has been greatly
-facilitated by the open data access policy of the {\it Swift} mission.
+GRBs are interesting as they provide the opportunity to derive new insights
+into a very wide range of topics in astrophysics. In that sense GRB
+research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
+which was launched in 2004, has made
+it possible to harvest the great potential in using GRBs as probes of
+galaxies in the intergalactic medium, which was already hinted at by
+results from earlier missions.
+
+A crucial aspect of the great success of the {\it Swift} mission has been
+the extensive ground based follow-up observations of the afterglows and of
+the host galaxies of the GRBs. This work has involved a large community
+of researchers both inside and outside the official {\it Swift} team. This fruitful
+collaboration has been greatly facilitated by the open data access policy
+of the {\it Swift} mission.
 
@@
 was secured using low-resolution spectrographs (typically
-R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution
-spectroscopy is powerful for a number of reasons, first of all is the ability
-to secure information for even very faint targets. This allows the measurements
-of a number of important parameters such as redshifts and spectral slopes. For
-a handful of very bright afterglows high-resolution (typically R$>$20000)
-spectra were secured and for these events much more information about
-conditions inside the host galaxies were extracted 
-\citep[e.g.,][]{Fiore,Thoene07,Prochaska07,Vreejswijk07,CastroTirado2010}.
-
-The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
-generation instruments at the ESO Very Large Telescope (VLT). It is a
-high-throughput echelle spectrograph covering the full spectra range from the
-UV atmospheric cutoff to the near-infrared (near-IR) K-band with three spectral
-arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
-in mind as the fading luminosities of such sources makes in urgent to secure as
-extensive coverage as possible in the shortest possible time. At the same time
-the resolution was set to be in the range 4000--9000 in order to be able to get
-a large useful spectra range between the many sky-background emission lines in
-the red and near-IR spectral ranges. In this paper we present the results
-of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
-spectrograph to secure spectroscopic observations of
-afterglows and host galaxies of GRBs detected by {\it Swift}. 
-All data resulting from the survey are publicly availble in reduced form.
-
-The paper is organized in the following way: In Sect.~\ref{sample} we describe
-the sample including the sample selection criteria and the observational
-strategy.  In Sec.t~\ref{results} we describe the results of the survey, i.e.
-the efficiency of the follow-up effort and the characteristics of the observed
-bursts. We also here assess the completeness of the finally realized sample.
-Finally, we offer our conclusions in Sect.~\ref{conclusions}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria and observations}\label{sample}
+R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
+very bright afterglows high-resolution (typically R$>$20000) spectra were
+secured.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
+instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
+echelle spectrograph
+covering the full spectra range from the UV atmospheric cutoff to the near-infrared
+(near-IR)
+K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
+transient source follow-up in mind as the fading luminosities of such sources 
+makes in urgent to secure as extensive coverage as possible in the shortest possible 
+time. At the same time the resolution was set to be in the range 4000--9000 in order
+to be able to get a large useful spectra range between the many sky-background
+emission lines in the red and near-IR.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sample selection criteria and follow-up}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Being of transient nature, it is difficult to impose strong sample selection
 criteria on GRBs, without hampering the follow-up effort. Many natural follow-up
-restrictions exists from already, being it weather conditions, pointing
-restrictions of the telescope or poorly localized bursts as reported by the
-\textit{Swift}-telescope. To maximize the return of the follow-up campaign
-we have chosen a few selection criteria that attempts to provide an unbiased
-selection of bursts, while allowing for a high success-rate
-
+restrictions exists already, being it weather conditions, pointing restrictions
+of the telescope or poorly localized bursts as reported by the alerting
+telescope. To maximize the return of the follow-up campaign we have chosen a few
+selection criteria that attempts to provide an unbiased selection of bursts,
+while allowing for a high follow-up success rate. The importance of defining
+unbiasing selection criteria has been highlighted previously
+\citep{Salvaterra2012, Hjorth2012, Vergani2015, Perley2016a}, when trying to
+address the specific underlying distribution functions such as the redshift
+distribution, host metallicity distribution, or brightness distribution. When
+investigating a specific distribution function, a high degree of
+\textit{completeness} is desired meaning that in terms of the parameter of
+interest, the largest possible fraction of targets in the sample has
+determinations of that parameter \citep{Perley2016b}.
+
+We simultaneously want to minimize any biases against astrophysical conditions
+while at the same time maximizing likelihood of observations allowing us to
+obtain a higher degree of observing completeness. By restricting the selection
+criteria to conditions local to Galaxy and therefore irrespective of intrinsic
+GRB properties, the hope is that the sample collect represents the underlying
+distribution of GRBs in a fair way while at the same time, additional unbiasing
+cuts can be imposed that can improve the follow-up success. The selection
+criteria used here is based on previous successes \cite{Jakobsson2006b,
+	Fynbo2009, Hjorth2012}, which has shown to represent the underlying distribution
+to a high degree. We test the degree of completeness in Sect.
+\ref{completeness}.
+
+The selection criteria are defined as follows:
+
+\begin{enumerate}
+	\item GRB triggered onboard by Swift.
+	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
+	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
 	\footnote{Note that in the P84 proposal the criteria have
 	been stated a bit differently, the visibility constraint being replaced by a
-declination + Sun angle constraint. The above criteria are however those
+	declination + Sun angle constraint. The below criteria are however those
 	defining the sample.}
-
-\begin{enumerate}
-	\item GRB triggered onboard by Swift.
+	\item No bright closeby stars.
 	\item Galactic $A_V \lesssim 0.5$ mag.
-	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
-	\item No bright closeby stars.
-
 \end{enumerate}
 
-A significant fraction of the bursts presented here \todo{Insert exact
-number} have already had their hosts investigated in \citet{Kruhler2015}, for
-which extractions of the hosts exist. The focus of the data presented here are
-on the afterglows themselves and in the absence of a clear afterglow, the host.
-We will not, however, investigate the hosts.
-
-We simultaneously want to minimize any biases against astrophysical conditions
-while at the same time maximizing likelihood of observations. By restricting
-the selection criteria to conditions local to Galaxy.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up procedure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Because our ability to observe GRB afterglows is strongly correlated with the
+rapidness and precision of the target positions delivered by the triggering
+facilities, by selecting only bursts that have been triggered on board the
+\textit{Swift} space telescope \cite{Gehrels2004}, by the Burst Alert Telescope
+(BAT) \citep{Barthelmy2005}, we start out with a homogeneously selected sample
+where burst characteristics are delivered immediately, allowing for an informed
+follow-up strategy. Despite the complexity of the triggering mechanism on board
+\textit{Swift} \citep{Band2006}, attempts at inferring properties of the
+underlying cosmic GRB population based on the detection thresholds and
+triggering algorithms have been made \citep{Lien2014, Graff2016} and thereby by
+restricting the follow-up effort to bursts detected by \textit{Swift}, the
+limitations of the parent sample is well studied.
+
+Because the localization accuracy of BAT is 1 - 4 arcminutes
+\citep{Barthelmy2005}, we additionally require an X-ray position from the X-ray
+Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
+\citep{Barthelmy2000} with 10 hours and to account for observing constraints on
+\textit{Swift} that XRT began observations within 10 minutes. The additional
+rapidness requirement of the XRT follow-up means that the number of detected
+X-ray afterglows for BAT triggers is 100 percent \citep{Burrows2007}. The
+\textit{Swift} constraints on the GRB follow-up means that the sample presented
+here inherits the \textit{Swift} sample properties, except the additional XRT
+constraint. Because the XRT completeness is very high\citep{Burrows2007}, this
+should change the parent sample significantly. This issue is addressed again in
+Sect. \ref{completeness}.
+
+To allow for observability we additionally require that the GRB is visible
+(duh!) from the telescope site at Cerro Paranal, Antofagasta, Región de
+Antofagasta, Chile, for a least 1 hour allowing time for the spectroscopic
+observations to completed. Since the GRB population is isotropically distributed
+seen from Earth and that GRB properties does not depend on position on the sky
+\citep{Briggs1996}, this cut does not influence our ability to sample the
+underlying GRB population. The same arguments applies to the requirement that
+there are no nearby bright stars, which is additionally supplemented by the
+requirement what the galactic dust content does not exert an extinction higher
+than $A_V \lesssim 0.5$ mag based on the extinction maps by
+\citet{Schlegel1998}. Because of  isotropy, these additional cuts should not
+influence the optical properties of the bursts themselves, only our ability to
+successfully secure observations that allows us to investigate the spectroscopic
+properties of GRBs.
+
+We address the succesfullness of these criteria and the actual follow-up effort
+in representing the full \textit{Swift} sample ion Sect. \ref{completeness}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up procedure and practicalities}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\todo{Will be written by Daniele?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Observations} \label{obs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All the observations in this sample has been carried out using the
-cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
-two of the four Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal)
-during the duration of this follow-up campaign, which covers the entire lifetime
-of X-shooter. Observations have been carried out from the ESO period P84
-through P98 under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
-095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.
+\section{Observations} \label{obs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}
+	\centerline{\includegraphics{figures/skymap.pdf}}
+	\caption{Mollweide projection in galactic coordinates of the full sky showing
+		the positions on the sky of the bursts presented in this work. The empty
+		circles are the positions of all the \textit{Swift} bursts detected until the
+		end of 2016. In blue is shown the position of the bursts fulfilling the sample
+		criteria specified in Sect. \ref{samplecrit}. The red circles are added around
+		GRBs which have been followed up by X-shooter since the commissioning of the
+		instrument. Blue points with red circles represent GRBs that enter our sample
+		and have X-shooter spectroscopy. The different samples are compared in Sect.
+		\ref{results}. The background shows the dust maps presented in
+		\citet{Schlegel1998} where the sample criteria cut with $A_V \lesssim 0.5$ mag
+		is removed.}
+	\label{fig:skymap}
+\end{figure*}
+
+Observations in this sample has been carried out using the cross-dispersed
+echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on two of the four
+Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the duration
+of this follow-up campaign, which covers the entire lifetime of X-shooter.
+Observations have been carried out from the ESO period P84 through P98 under the
+following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.A-0045,
+094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088, 089.A-0067,
+088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. A few additional
+bursts, not a part of the \textit{statistical sample}, have been included from
+092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI: Greiner), and Levan under
+095.B-0811(B) (PI: Levan). This sample represents \textit{all} GRBs that have
+ever been followed up by X-shooter.
 
@@
 
-X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800
-\AA~in a single exposure, by splitting the light in three separate
-spectroscopic arms using dichroics.  This way each arm work as a separate
-instrument, each functioning as its own echelle spectrograph.  The ultraviolet
-blue(UVB) arm covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 -
-10200 \AA, and the near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the
-possibility of applying a k-band blocking filter cutting the coverage of the
-NIR arm at 10200 - 21000 \AA.
-\\
-
-\todo{ABBA}
+X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800 \AA~in
+a single exposure, by splitting the light in three separate spectroscopic arms
+using dichroics. This way each arm work as a separate instrument, each
+functioning as its own echelle spectrograph. The ultraviolet blue(UVB) arm
+covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 - 10200 \AA, and the
+near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the possibility of
+applying a k-band blocking filter cutting the coverage of the NIR arm at 10200 -
+21000 \AA. For all observations, a nod-and-shuffle observing scheme has been
+employed, with a nodding throw of 5 \arcsec. The individual observations has
+been carried out in a standard ABBA pattern, which allows flexibility in the sky
+subtraction method for the different arms.
 
@@
 slit width. We discuss this in Sect. \ref{resolution}.
-\\
-\todo{ADC discussion}
+
+Due to mechanical issues, the atmospheric dispersion corrector (ADC) were
+disabled from 1st of August 2012. Only GRB~100728B was affected by the failing
+ADC prior to disablement, resulting in a lower-than-nominal throughput. To avoid
+chromatic slit losses due to atmospheric dispersion, all subsequent observations
+have been carried out at parallactic angle. A consequence of this is that for
+all observations following 1st of August 2012, the source moves across the
+spatial direction of the slit as a function of wavelength. This effect has been
+modeled in the extraction procedure, as described in \ref{extract}.
 
 We list the overview of all the observations in Tab. \ref{tab:sample_overview}.
-\input{tables/burst_overview.tex}
+\input{tables/burst_overview.tex} and plot the positions off all the bursts on
+the night sky in Fig. \ref{fig:skymap}. The central zone of avoidance due to
+high galactic extinction cutoff is visible with an apparent isotropic
+distribution of the bursts away from the galactic center except in the upper
+right corner, where Earth casts a shadow on the sky as seen from the telescope.
+
+A subset of the bursts presented here  have already had their hosts investigated
+in \citet{Kruhler2015}, for which extractions of the hosts exist. The focus of
+the data presented here are on the afterglows themselves and in the absence of a
+clear afterglow, the host. We will not, however, investigate the hosts.
 
@@
 
+\begin{figure*}
+	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
+	\caption{Telluric corrected, normalized spectrum of GRB121024A at $z = 2.300$,
+	meant to illustrate the typical data quality. The acquisition magnitude is r =
+	20, meaning it is in the brighter end of the sample presented here, but not the
+	brightest. The spectrum is rich in absorption, including molecular absorption
+	of $H_2$. The absorption trough visible at $\sim$ 4000 $\AA$ is due to \lya in
+	the host. We have overplot the most prominent lines seen in GRB afterglows from
+	\citet{Christensen2011a}. The continuum estimate is shown in dashed red.
+	Additionally, three intervening systems are seen in this sightline  This
+	spectrum has been analyzed in detail in \citet{Friis2015}.}
+	\label{fig:spectrum}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Continuum estimate} \label{continuum}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We additionally provide an estimate of the continuum for all the spectra
+presented here. We have developed an algorithm that attempts to automatically
+find the continuum placement along with the error on the continuum estimate. For
+each arm,
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!ht]
+	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
+The lower horizontal axis is in units pixels, the top axis in arc seconds.
+The red datapoints show a subsample of NIR spectra.
+		The grey line shows a linear fit to the VIS datapoints. }
+	\label{fig:res}
+\end{figure}
 
@@
 
-\begin{figure}
-	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
-The lower horizontal axis is in units pixels, the top axis in arc seconds.
-The red datapoints show a subsample of NIR spectra.
-		The grey line shows a linear fit to the VIS datapoints. }
-	\label{fig:res}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}\label{results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Science data products} \label{products}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+All the spectra are made available at
+\url{http://www.dark-cosmology.dk/~jselsing/XSGRB/} and additionally at , including the ones outside
+the statistical sample. This release includes both prompt afterglow observations
+as well as late time observations. An overview of all the spectra provided here
+are given in Table. \ref{sample_overview}. For each burst, each individual
+observation is provided in a separate reduction and in cases where observations
+have been repeated, a combined spectrum is also provided.  No attempt has been
+made to combine the arms, so for each observation, three spectra exists.
+
+We have
+adopted the following naming convention GRB*OB*.fits
+
+In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results} \label{results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 characteristics of the observed bursts. We also assess the degree to which the
-sample obtained is complete.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up timing and afterglow brightness}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A rapid follow-up is essential for the accurate designation of GRB the origin. The redshift determination of burst where the host is not easily
+sample obtained is complete. An important note is that here we present
+\textit{all} GRBs that have ever been observed with X-shooter, while only a
+subset of these constitute our \textit{statistical} sample. The statistical
+sample is based on the selection criteria described in Sect. \ref{samplecrit}
+and should represent an unbiased sample of the underlying, parent GRB
+population. Single bursts outside the sample criteria have been followed up on
+the basis of the curiosity of their light curve or their brightness, but is not
+discussed as part of the investigation of the GRB population. A prime example of
+a burst outside the statistical sample is GRB161023A with it's rich
+spectrum containing 11 intervening absorption systems and pending further analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Follow-up timing and afterglow brightness} \label{timing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Redshift determination of bursts where the host is too faint for a spectroscopic
+redshift relies on the absorption spectrum imprinted on the GRB afterglow
+continuum, which again depends on the afterglow brightness at time of follow-up.
+Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better references} a rapid
+follow-up is essential for the accurate designation of the GRB host \todo{Get
+reference}. In Fig. \ref{fig:timing} we plot the follow-up delay from the BAT
+trigger to the start of the spectroscopic observation. Visible as the
+shortest delays are the bursts observed in RRM-mode in which where the fastest follow-up
+between BAT trigger and start of spectroscopic observations is for the short, $z = 1.717$, GRB160410A which was
+followed up only 8.4 minutes after the BAT trigger.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample completeness}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Based on the sample criteria specified in Sect.
-\ref{samplecrit}, a total of 152 \todo{Get number} bursts has been triggered by \textit{Swift} in the period since the comissioning of VLT/X-shooter in \todo{Get date}.
-This constitutes the \textit{statistical sample}.
+To illustrate the effect of redshift completeness loss due to late follow-up, we
+overplot the redshift completeness fraction over the delay time. Visible is the
+increasing fraction of GRBs without a good redshift determination for increasing
+delays. That the redshift completeness of the bursts we have followed up are $>$
+90 percent \todo{Get number}, shows the ability of VLT/X-shooter to successfully
+get redshifts. Not shown in the figure is an additional 12 \todo{Get number}
+bursts that we have redshift determinations from late-time host observations
+with delay-times longer than $\sim$10 days.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample completeness} \label{completeness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
+\todo{Get number} bursts has been triggered by \textit{Swift} in the period
+since the commissioning of VLT/X-shooter in November, 2009. This sample
+constitutes the \textit{statistical sample} on which we will try to address the
+GRB population properties. Of this sample, 86 \todo{Get number} have been
+spectroscopically followed up with X-shooter. In order to assess whether our
+sample criteria and the subset of bursts followed up represent the underlying
+parent population of GRB, we compare some intrinsic GRB properties of our sample
+to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
+that \textit{Swift} randomly samples the underlying GRB population, if the
+sample statistics of our subsample are similar to the \textit{Swift}-sample, we
+can assume that the population properties should be conserved. There is some
+evidence that the burst duration does not scale with redshift as expected
+through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
+that we could be missing an increasing fraction of the bursts at higher
+redshifts, but \citet{Lien2016} argues that we should only be missing the bright
+"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
+the sample presented here is complete. We show the comparison between the burst
+properties in Fig. \ref{fig:swift_complete}.
+
+\begin{figure*}
+	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
+	\caption{Comparison between bursts properties of all bursts observed with
+	\textit{Swift}, the properties of the ones that fullfill the sample criteria
+	specified in Sect. \ref{samplecrit}, and the subset that has been observed as
+	part of the statistical sample. The left plot shows the fluence in 15-150 keV
+	as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
+	after the bursts as measured by XRT. Based on the XRT spectrum,
+	\citep{Evans2009}, a hydrogen column density has been inferred which we show in
+	the right most panel.}
+	\label{fig:swift_complete}
+\end{figure*}
+
+Using the observational characteristics of the 1244 \todo{Get number} bursts
+observed by \textit{Swift} to date \cite{Gehrels2009}, and the X-ray derived
+hydrogen column densities \citep{Evans2009}, we can quantify the degree to which
+our sample is biased against the overall \textit{Swift} sample in terms of
+intrinsic GRB properties. The values are queried from the online \textit{Swift}
+database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
+Three main samples are of interest to compare to assess the completeness of the
+follow-up campaign; the full \textit{Swift} sample consisting of all the bursts
+observed by \textit{Swift}, excluding the bursts entering our sample (1), the
+burst that fulfill the selection criteria imposed in Sect. \ref{samplecrit}, but
+has not been followed up (2), and the bursts actually followed up with X-shooter
+(3).
+
+\input{tables/sample_properties.tex}
+
+For each of the samples we calculate the median and 14th and 86th percentile
+which can be used as point estimates for the population distribution. These are
+shown in \ref{tab:sample_properties}. It can be seen from the values that the
+three samples have very similar distributions in terms of the point estimates
+chosen, suggesting that our selection criteria are unbiasing against the
+\textit{Swift}-sample and that additionally, the follow-up effort is unbiasing
+towards the intrinsic GRB properties. This is despite spectroscopic follow-up
+only being carried out in cases where either a detectable optical afterglow or a
+clear counterpart are seen, which naively should be biased against dark bursts
+occurring in more obscured galaxies, which is shown to exhibit different
+galactic properties \citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b,
+	Perley2015b}.
+
+Additionally, using a 2-sided Kolmogorov-Smirnov test (KS-test), we can assess
+the degree to which the null that the two distribution are drawn from the same
+distribution can be rejected. We show a graphic representation of the test
+statistics in Fig. \ref{fig:p_values}. We have run the test on whether is can be
+rejected that the sample criteria is drawn from the same distribution as the
+\textit{Swift} sample, if the observed sample is dissimilar from the statistical
+sample, and again whether the observed sample differs from the full
+\textit{Swift} sample. A high p-value indicates little evidence against the null
+meaning for the parameters investigated here, only the X-ray derived column
+densities exhibits the highest degree of dissimilarity, but not to a significant
+level. A large part of the tension between the two samples in $N_H$ is driven by
+a small number of low column densities($\log(N_H) \lesssim 19 \mathrm{cm}^2$)
+bursts.
+
+\begin{figure}
+	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
+	\caption{Relational graph shown the respective p-values that the different
+	sample are drawn from the same underlying distribution. The arrows represent
+	the comparison with each of the p-values are listed next to. Only in the HI
+	column density distribution is there mild evidence against the null, but the
+	discrepancy is mainly driven by a relatively larger fractional contribution
+	from low-column hosts.}
+	\label{fig:p_values}
+\end{figure}
+
+We therefore conclude that the sample presented here, to a high degree
+represents the intrinsic properties of the GRBs in the full \textit{Swift} sample.
+\todo{This doesn't quite feel exhaustive? Is there a better way of addressing the issue of completeness}
 
@@
 
-How many bursts have been excluded due to particularities in the follow-up attempt. Visitor, Weather, technical issues.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{On the nature of triggered GRBs without detected optical afterglow.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This should include observations where the acquisition image has been taken, but observations were discontinued due to non-detection. What are the X-ray properties of these bursts?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Redshifts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Afterglow-based redshifts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Emissionline-based redshifts}	
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Redshift completeness
+GRB follow-up requires the immediate availability of the telescope in order to
+successfully carry out spectroscopic observations. Out of the 158 \todo{Get
+exact number} bursts in the statistical sample, 32($\sim$ 20 percent) \todo{Get
+exact number} have not been observed due to conditions, not relating the GRB or
+afterglow properties. The reasons includes unavailability of the telescope due
+to technical maintenance such af mirror re-coating, a visitor rejecting the TOO
+trigger, or bad weather at the observing site. Because this cut is irrespective
+of the GRB properties, it is not expected that this should change sample
+properties. Addressing whether this cut introduces a bias in the followed sample
+is not trivial, but in terms of intrinsic GRB properties, the sample
+distributions does not change significantly, as shown in Sect. \ref{completeness}.
+\todo{Should this be quantified somehow?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample darkness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A subset of all GRBs exhibit no detectable or very faint optical afterglow
+\citep{Groot1998, Djorgovski2001, Fynbo2001}, which had been parametrized in
+term of "darkness" \citep{Jakobsson2004, VanderHorst2009}. The X-ray properties
+of this subset has previously been investigated \citep{DePasquale2003, Fynbo2009,
+	Melandri2012}, finding a slightly higher X-ray luminosity compared to the
+optically brights bursts combined with similar prompt characteristics pointing
+towards a denser circumburst material. This indicates along with investigations
+of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
+extinction of the optical afterglow is primarily driven by the presence of dust,
+local to the burst \todo{Local dust reference}. Because the dust content of a
+galaxy is correlated with the gas amount \citep{Bohlin1978, Guver2009}, we
+expect the X-ray derived column densities for the bursts without an optical
+afterglow to be higher than the ones without.
+
+For all bursts with follow-up within 100 hours we can calculate the
+"darkness"-parameter, $\beta_{OX}$ \citep{Jakobsson2004, Milvang-Jensen2012}. This required the
+simultaneous measurement of the X-ray flux density and the optical flux density
+which is in practice never possible. We use the measured acquisition camera
+magnitude reported in Tab. \ref{tab:sample_overview} to get the optical flux
+density at the beginning of the spectroscopic integration. Because we know the
+delay between the follow-up and the \textit{Swift} trigger, we can use the
+measured XRT burst lightcurve \citep{Evans2007,
+	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
+to infer the corresponding X-ray flux density. This is done by either linearly
+interpolating between temporally neighboring XRT measurements of by
+extrapolating the last few X-ray data points to the spectroscopic delay. When
+the afterglow is not detected in the acquisition camera, an upper limit of $> 24$
+mags is used, which propagates into an upper limit on $\beta_{OX}$.
+
+In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
+the one presented in \citet{Fynbo2009}. We have replaced limits with values,
+meaning that we artificially bias the distribution towards higher
+$\beta_{OX}$-values because darker bursts naturally fainter in the optical
+making the fraction of non-detections higher for lower $\beta_{OX}$. We
+marginally confirm the result by \citet{Fynbo2009} that dark bursts $\beta_{OX}
+< 0.5$) have slightly higher X-ray derived column densities, by calculating the
+median, 14th, and 86th percentile of the distributions arising by separating the
+sample at $\beta_{OX} = 0.5$. For bursts with $\beta_{OX} \geq 0.5$ we find
+$\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
+$21.8_{-0.9}^{+0.5}$, meaning that the median value for dark bursts is higher,
+but with a great deal of overlap. A 2-sided KS test fails to reject the null
+that they are drawn from the same distribution with $p = 0.43$, meaning there is
+no clear evidence for a discrepancy.
+
+\begin{figure}
+	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
+	\caption{Optical extinction relative to X-ray against X-ray derived hydrogen
+	column densities. In red is shown the sample presented in \citet{Fynbo2009}
+	where is lines indicate the kernel density estimate of the distribution where
+	limits have been replaced with values. Darker colors represented denser points.
+	The correcspoing marginal distributions are shown. In blue is the points for
+	the bursts presented here along with the marginal histograms. Limits on
+	$\beta_{OX}$ is shown by downwards facing arrows.}
+	\label{fig:betaOX}
+\end{figure}
+
+Because we only follow up targets with either a clear afterglow or counterpart,
+correctly associating a galaxy with a burst when there is no detected optical
+afterglow is more difficult, see \citet{Levesque2010} and \citet{Perley2017} for
+a recent example. Because the presence of optical transient is highly dependent
+on the "darkness", we expect that the $\beta_{OX}$ sample is biased against dark
+bursts. The larger relative fraction of upper limits for increased darkness is
+visible both in the sample presented here and the one in \citet{Fynbo2009}. The
+X-ray properties for the followed sample does not differ signicantly from the
+statistical sample, as shown in Sect. \ref{completeness}. Using the table
+maintained by Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How
+to reference this}, we can see how the presence of an optical afterglow affects
+the follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do
+not have an detected optical afterglow, but this number also includes bursts
+where no optical telescopes were available, so is a lower limit. For the bursts
+entering our statistical sample, this number is $\sim$ 27 percent, close to the
+maximum fraction of dark bursts \citep{Melandri2012}. Of the subsample (3) that
+have actually been attempted, this number is $\sim$ 18 percent, confirming that
+we are indeed biased against bursts with an optical afterglow in the
+spectroscopic sample.
+
+Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts.
+\todo{Hmm, how do I finish this?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{On the redshift distribution of GRBs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Because the redshift completeness of our statistical sample is just 54 percent
+\todo{Get number}, making inference of the underlying redshift distribution of
+GRBs on this ample is difficult. Because a large fraction of the bursts for which we were
+unable to secure a redshift are missing due to terrestrial reasons, see Sect.
+\ref{badbursts}, imposing additional unbiasing cuts on the sample can increase
+the redshift completeness at the cost of sample size. As long as the cuts we
+impose are not selective in terms of the burst properties, the sample properties
+stay the same. The first cut we try is telescope availability such as bad
+weather, visitor rejected the TOO, or unavailability of the instrument. Because
+telescope availability is independent on the burst properties, by cutting away
+these from our sample, the underlying redshift distribution should be conserved.
+By trimming away the 30 bursts \todo{Get number} that can be rejected due to
+telescope availability, the redshift completeness increases to 63 percent
+\todo{Get number}.
+
+As is shown in Sect. \ref{timing}, the redshift completeness
+decreases with increasing follow-up delay, we can additionally increase the
+redshift completeness by imposing cuts on the follow-up delay. Our ability to
+rapidly follow up a burst is also not dependent on the burst properties and
+therefore we can additionally cut delays longer than 24 hours (all burst in the
+sample are in principle observable within 24 hours). \todo{Unfinished section}
+
+The biased introduced in the redshift distribution by the increased ability to
+secure redshift of optically brighter bursts has been investigated by
+\citet{Turpin2016} which find that we systematically find redshift for longer
+GRBs. Additionally only the brightest GRBs are seen above redshift $z \gtrsim
+1$. \todo{Unfinished section}
 
@@
 
-\begin{figure*}
+In a similar manner to how the X-ray derived column densities was used to assess
+the sample completeness in Sect. \ref{completeness}, we addresses the same
+question in terms of the optically derived hydrogen column densities. Using the
+compilation by \citet{Tanvir2017} we have all 81 published HI values previously,
+we can compare to the 39 new HI columns added by this sample. We note that this
+sample increase the number of optically derived hydrogen column densities with
+$\sim$ 50 percent. We compare the median, the 14th, and 86th percentiles of the
+two distributions where the sample presented here has $21.8_{0.8}^{0.3}$ and the
+rest of the litterature values has $21.5_{1.5}^{0.4}$. We see that the two
+distribution has a large degree of overlap due to the large width of the
+distributions, but find a slightly higher median value for the new sample
+presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
+relatively large evidence against the null that the two samples are drawn from
+the same underlying distribution. Because the bursts that have measurements of
+the hydrogen column density are selected sole based on our ability to infer a
+column, it is difficult to make any strong conclusions about the completeness in
+terms of gas content. We show the comparison between the two samples in Fig.
+\ref{fig:NH_dist}, where we additionally show the column-distribution for 12081
+quasar DLAs published in \citet{Noterdaeme2012b}. The ability of the GRBs to
+select the highest N$_H$ is immediately visible as is also noted previously
+\citep{Prochaska2007, Fynbo2009}. The reason for this is the selection of sight
+lines as random position of the QSO DLAs whereas the GRB forms in dense regions
+in the host. The N$_H$ distribution has also be used the infer the position of
+the gas in host \cite{Buchner2016} where a galaxy-wide gas reservoir is
+preferred.
+
+\begin{figure}
 	\centering
-	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
-	\caption{ Measurements of the hydrogen column-densities for all bursts with a clear Lyman alpha absorption system.
-In solid black is shown the spectrum with black dotted giving the corresponding 1-$\sigma$ error.
-Black dashed shows zero flux density.
-The solid red line is the absorption of column density equal to the value presented in Tab.
-\ref{HI_tab} with the 1-$\sigma$ interval shown with dashed lines.}
-	\label{fig:HI1}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI2}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI3}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI4}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
-	\caption*{Fig. \ref{fig:HI1}. continued.}
-	\label{fig:HI5}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion and conclusions}\label{conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this paper we have presented the results of a dedicated effort over the
-years 2009 -- 2016 to use the X-shooter spectrograph on the ESO-VLT to secure
-spectroscopic observations of afterglows and host galaxies of GRBs detected by
-{\it Swift}. This work was initiated by a consortium using Guaranteed Time on
-X-shooter, but over the years the project continuued in open time and the team
-was opened to include any researchers interested in contribution to this
+	\includegraphics[width=9cm]{figures/NH_dist.pdf}
+	\caption{Distributions of DLA hydrogen column densities for DLAs found in
+		quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
+		is the similar distribution, only for DLAs in GRB sightlines. Values are taken
+		from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
+		with the new values presented in this sample. The marginal distributions for
+		the two distribution are also shown, where the different environments probed
+		are clearly visible in the hydrogen column densities, as previously also noted
+		in \citet{Fynbo2009}}
+	\label{fig:NH_dist}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper we have presented the results
+of a dedicated effort over the years 2009 -- 2016 to use the X-shooter 
+spectrograph on the ESO-VLT to secure spectroscopic observations of 
+afterglows and host galaxies of GRBs detected by {\it Swift}. This 
+work was initiated by a consortium using Guaranteed Time on X-shooter, 
+but over the years the project continuued in open time and the team was
+opened to include any researchers interested in contribution to this 
 effort.
@@
 
-Our sample serves the purpose to further characterize the environments 
-of GRBs that was also much advanced by the previous surveys based primarily on 
-lower-resolution spectographs. The most striking result found is that GRBs
-typically are located behind very large column densities of neutral hydrogen.
-Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ quasars have a handful such systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
-reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
-With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we hence characterize the properties of
-star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities, etc.
-
-It seems important to end this paper by stressing that the potential of 
-using GRB sightlines as probes is far from fully harvested. The sample of
-sightlines probed by our spectra is not representative for all GRB 
-sightlines as we have shown (CONFIRM) and consistent with earlier findings
-\citep[e.g.,][]{Fynbo2009}. \cite{Kruhler2013} convincingly argue, that 
-very rich sightlines like that probed by the remarkable GRB\,080607 
-\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
-sampled by our spectra. However, with current instrumentation these sightlines are
-out of reach except under very fortunate circumstances as in the case of 
-GRB\,080607 when the afterglow could be observed only a few minutes after the 
-burst with a 10-m class telescope. Observations of such sightlines with 
-X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
-is likely to be very rewarding.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{acknowledgements}
-*** We should add an ode to Neil Gehrels here ***.
 JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The

### 2017-10-26 document/XSGRB_sample.tex

@@
 
-GRBs are interesting as they provide the opportunity to derive new insights
-into a very wide range of topics in astrophysics. In that sense GRB
-research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
-which was launched in 2004, has made
-it possible to harvest the great potential in using GRBs as probes of
-galaxies in the intergalactic medium, which was already hinted at by
-results from earlier missions.
-
-A crucial aspect of the great success of the {\it Swift} mission has been
-the extensive ground based follow-up observations of the afterglows and of
-the host galaxies of the GRBs. This work has involved a large community
-of researchers both inside and outside the official {\it Swift} team. This fruitful
-collaboration has been greatly facilitated by the open data access policy
-of the {\it Swift} mission.
+Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
+derive new insights into a very wide range of topics in astrophysics. Examples
+range from small scale phenomena relating to the magnetars, properties of highly
+relativistic jets, hyper/supernova explosions, the interstellar medium, dust extinction
+curves, starburst galaxies, chemical and molecular abundances, escape of ionizing 
+radiation, the ionization state of the intergalactic medium, intervening
+absoption systems to standard candles in cosmology.  In that sense GRB research
+is very multidisciplinary.
+
+The {\it Swift} satellite \citep{Gehrels09}, which was launched in 2004, has
+made it possible to harvest much of the great potential in using GRBs as probes
+of galaxies in the intergalactic medium, which was already hinted at by results
+from earlier missions \citep[e.g.][]{Ricker04}.  A crucial aspect of the great
+success of the {\it Swift} mission has been the extensive ground based
+follow-up observations of the afterglows and of the host galaxies of the GRBs.
+This work has involved a large community of researchers both inside and outside
+the official {\it Swift} team. This fruitful collaboration has been greatly
+facilitated by the open data access policy of the {\it Swift} mission.
 
 In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
-was secured using low-resolution spectrographs (typically
-R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
-very bright afterglows high-resolution (typically R$>$20000) spectra were
-secured.
-
-The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
-instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
-echelle spectrograph
-covering the full spectra range from the UV atmospheric cutoff to the near-infrared
-(near-IR)
-K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
-transient source follow-up in mind as the fading luminosities of such sources 
-makes in urgent to secure as extensive coverage as possible in the shortest possible 
-time. At the same time the resolution was set to be in the range 4000--9000 in order
-to be able to get a large useful spectra range between the many sky-background
-emission lines in the red and near-IR.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sample selection criteria and follow-up}
+was secured using low-resolution spectrographs 
+\citep[typically
+R=$\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Low-resolution
+spectroscopy is powerful for a number of reasons, first of all is the ability
+to secure information for even very faint targets. This allows the measurements
+of a number of important parameters such as redshifts, spectral slopes,
+and extinction. For
+a handful of very bright afterglows high-resolution (typically R$>$20000)
+spectra were secured and for these events much more information about
+conditions inside the host galaxies were extracted 
+\citep[e.g.,][]{Fiore,Thoene07,Prochaska07,Vreejswijk07,CastroTirado2010}.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
+generation instruments at the ESO Very Large Telescope (VLT). It is a
+high-throughput echelle spectrograph covering the full spectral range from the
+UV atmospheric cutoff around 300 nm to the near-infrared (near-IR) K-band 
+ending around 2500 nm with three spectral
+arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
+in mind as the fading luminosities of such sources makes in urgent to secure as
+extensive coverage as possible in the shortest possible time. At the same time
+the resolution was set to be in the range 4000--9000 in order to be able to get
+a large useful spectral range between the many sky-background emission lines in
+the red and near-IR spectral ranges. In this paper we present the results
+of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
+spectrograph to secure spectroscopic observations of
+afterglows and host galaxies of GRBs detected by {\it Swift}. 
+All data resulting from the survey are publicly availble in reduced form.
+
+The paper is organized in the following way: In Sect.~\ref{sample} we describe
+the sample including the sample selection criteria and the observational
+strategy.  In Sec.t~\ref{results} we describe the results of the survey, i.e.
+the efficiency of the follow-up effort and the characteristics of the observed
+bursts. We also here assess the completeness of the finally realized sample.
+Finally, we offer our conclusions in Sect.~\ref{conclusions}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sample selection criteria and observations}\label{sample}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-X-shooter can cover the spectral wavelength region from 3000 \AA~to 24800 \AA~in
-a single exposure, by splitting the light in three separate spectroscopic arms
+X-shooter can cover the spectral wavelength region from 300 nm to 2480 nm in a
+single exposure, by splitting the light in three separate spectroscopic arms
 using dichroics.  This way each arm work as a separate instrument, each
 functioning as its own echelle spectrograph.  The ultraviolet blue(UVB) arm
-covering 3000 - 5500 \AA, the visual(VIS) arm covering 5500 - 10200 \AA, and the
-near-infrared()NIR) arm covering 10200 - 24800 \AA~ with the possibility of
-applying a k-band blocking filter cutting the coverage of the NIR arm at 10200 -
-21000 \AA. For all observations, a nod-and-shuffle observing scheme has been
+covering 3000 - 550 nm, the visual(VIS) arm covering 550 - 1020 nm, and the
+near-infrared()NIR) arm covering 10200 - 2480 nm with the possibility of
+applying a k-band blocking filter cutting the coverage of the NIR arm at 1020 -
+2100 nm. For all observations, a nod-and-shuffle observing scheme has been
 employed, with a nodding throw of 5 \arcsec. The individual observations has
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and conclusions}\label{conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper we have presented the results of a dedicated effort over the
+years 2009 -- 2016 to use the X-shooter spectrograph on the ESO-VLT to secure
+spectroscopic observations of afterglows and host galaxies of GRBs detected by
+{\it Swift}. This work was initiated by a consortium using Guaranteed Time on
+X-shooter, but over the years the project continuued in open time and the team
+was opened to include any researchers interested in contribution to this
+effort.
+
+The sample we present here include spectroscopic observations of 86 
+systems fulfilling our sample criteria of which 72 are afterglow spectra 
+and 14 are observations of the underlying host galaxies. 
+UPDATE THIS.
+All spectra have been made publicly available in the reduced form used
+in this paper.
+
+Our sample serves the purpose to further characterize the environments of GRBs
+that was also much advanced by the previous surveys based primarily on
+lower-resolution spectographs. The most striking result found is that GRBs
+typically are located behind very large column densities of neutral hydrogen.
+Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+towards about 10$^5$ quasars have a handful such systems with
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
+in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006,
+Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we hence characterize the properties of
+star-forming galaxies over cosmic history in terms of redshifts, metallicities,
+molecular contents, ISM temperatures, UV-flux densities, extinction curves,
+etc.  A number of independent papers have been published or submitted for
+publication focusing on many of these specific issues of our sample such as
+extinction curves [][see also
+\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
+underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
+\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
+radiation \citep{Tanvir17}.
+
+It seems important to end this paper by stressing that the potential of 
+using GRB sightlines as probes is far from fully harvested. The sample of
+sightlines probed by our spectra is not representative for all GRB 
+sightlines as we have shown and consistent with earlier findings
+from samples based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009}
+and from studies of complete samples of GRB host galaxies 
+\citep{Hjorth2012,Covinto2013,Perley2016}. 
+\cite{Kruhler2013} convincingly argue, that 
+very rich sightlines like that probed by the remarkable GRB\,080607 
+\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
+sampled by our spectra. However, with current instrumentation these sightlines are
+out of reach except under very fortunate circumstances as in the case of 
+GRB\,080607 when the afterglow could be observed only a few minutes after the 
+burst with a 10-m class telescope. Observations of such sightlines with 
+X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
+is likely to be very rewarding.
+
+\begin{acknowledgements}
+*** We should add an ode to Neil Gehrels here ***.
+JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
+Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
+acknowledges support by the European Commission under the Marie Curie
+Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
+European Commission under the Marie Curie Career Integration Grant programme
+(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
+{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
+acknowledge expert support from the ESO staff at the Paranal and La Silla
+observatories in obtaining these target of opportunity data.
+
+\end{acknowledgements}
+
 \appendix

### 2017-10-26 document/XSGRB_sample.tex
Small edit

@@
 We additionally provide an estimate of the continuum for all the spectra
-presented here. We have developed an algorithm that attempts to automatically
-find the continuum placement along with the error on the continuum estimate. For
-each arm,
+presented here. For this, we have developed an algorithm that attempts to automatically
+finds the continuum placement along with the error on the continuum estimate.
 
@@
 
-We have
-adopted the following naming convention GRB*OB*.fits
-
-In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
+%We have
+%adopted the following naming convention GRB*OB*.fits
+%
+%
+%In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
+%
+%
 

### 2017-10-26 document/XSGRB_sample.tex
Adding Kaspers plot
fix

@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=16cm]{figures/XSGRB_redshift.pdf}
+	\caption{Redshift distribution of the XSGRB sample.}
+	\label{fig:z}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 UPDATE THIS.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+All spectra have been made publicly available in the reduced form used
+in this paper.
+
+Our sample serves the purpose to further characterize the environments of GRBs
+that was also much advanced by the previous surveys based primarily on
+lower-resolution spectographs. The most striking result found is that GRBs
+typically are located behind very large column densities of neutral hydrogen.
+Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+towards about 10$^5$ quasars have a handful such systems with
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
+in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006,
+	Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
+spectrum from X-rays to the sub-mm) we hence characterize the properties of
+star-forming galaxies over cosmic history in terms of redshifts, metallicities,
+molecular contents, ISM temperatures, UV-flux densities, extinction curves,
+etc.  A number of independent papers have been published or submitted for
+publication focusing on many of these specific issues of our sample such as
+extinction curves [][see also
+\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
+underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
+\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
+radiation \citep{Tanvir17}.
+
+It seems important to end this paper by stressing that the potential of 
+using GRB sightlines as probes is far from fully harvested. The sample of
+sightlines probed by our spectra is not representative for all GRB 
+sightlines as we have shown and consistent with earlier findings
+from samples based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009}
+and from studies of complete samples of GRB host galaxies 
+\citep{Hjorth2012,Covinto2013,Perley2016}. 
+\cite{Kruhler2013} convincingly argue, that 
+very rich sightlines like that probed by the remarkable GRB\,080607 
+\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
+sampled by our spectra. However, with current instrumentation these sightlines are
+out of reach except under very fortunate circumstances as in the case of 
+GRB\,080607 when the afterglow could be observed only a few minutes after the 
+burst with a 10-m class telescope. Observations of such sightlines with 
+X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
+is likely to be very rewarding.
 
 \begin{acknowledgements}
+	*** We should add an ode to Neil Gehrels here ***.
 	JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion and conclusions}\label{conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this paper we have presented the results of a dedicated effort over the
-years 2009 -- 2016 to use the X-shooter spectrograph on the ESO-VLT to secure
-spectroscopic observations of afterglows and host galaxies of GRBs detected by
-{\it Swift}. This work was initiated by a consortium using Guaranteed Time on
-X-shooter, but over the years the project continuued in open time and the team
-was opened to include any researchers interested in contribution to this
-effort.
-
-The sample we present here include spectroscopic observations of 86 
-systems fulfilling our sample criteria of which 72 are afterglow spectra 
-and 14 are observations of the underlying host galaxies. 
-UPDATE THIS.
-All spectra have been made publicly available in the reduced form used
-in this paper.
-
-Our sample serves the purpose to further characterize the environments of GRBs
-that was also much advanced by the previous surveys based primarily on
-lower-resolution spectographs. The most striking result found is that GRBs
-typically are located behind very large column densities of neutral hydrogen.
-Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ quasars have a handful such systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
-in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006,
-Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we hence characterize the properties of
-star-forming galaxies over cosmic history in terms of redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities, extinction curves,
-etc.  A number of independent papers have been published or submitted for
-publication focusing on many of these specific issues of our sample such as
-extinction curves [][see also
-\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
-underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
-\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
-radiation \citep{Tanvir17}.
-
-It seems important to end this paper by stressing that the potential of 
-using GRB sightlines as probes is far from fully harvested. The sample of
-sightlines probed by our spectra is not representative for all GRB 
-sightlines as we have shown and consistent with earlier findings
-from samples based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009}
-and from studies of complete samples of GRB host galaxies 
-\citep{Hjorth2012,Covinto2013,Perley2016}. 
-\cite{Kruhler2013} convincingly argue, that 
-very rich sightlines like that probed by the remarkable GRB\,080607 
-\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
-sampled by our spectra. However, with current instrumentation these sightlines are
-out of reach except under very fortunate circumstances as in the case of 
-GRB\,080607 when the afterglow could be observed only a few minutes after the 
-burst with a 10-m class telescope. Observations of such sightlines with 
-X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
-is likely to be very rewarding.
-
-\begin{acknowledgements}
-*** We should add an ode to Neil Gehrels here ***.
-JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
-Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
-acknowledges support by the European Commission under the Marie Curie
-Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
-European Commission under the Marie Curie Career Integration Grant programme
-(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
-{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
-acknowledge expert support from the ESO staff at the Paranal and La Silla
-observatories in obtaining these target of opportunity data.
-
-\end{acknowledgements}
-
 \appendix

### 2017-10-27 document/XSGRB_sample.tex
Intro itaration
A little
A little more work on abstract, intro and conclusions.

@@
 compared with literature values.
-We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and tentatively confirm previous
-results that larger optical darkness leads to increased X-ray absorption.
+We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and 
+tentatively confirm previous
+results that larger optical darkness leads is correlated with increased X-ray absorption.
 %This dataset provides a unique possibility to study the ISM across cosmic time.
@@
 afterglows and host galaxies of GRBs detected by {\it Swift}. 
-All data resulting from the survey are publicly availble in reduced form.
+Together with this paper we make all data resulting from the survey publicly 
+availble in reduced form.
 
@@
 publication focusing on many of these specific issues of our sample such as
-extinction curves [][see also
-\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
+extinction curves \citep[][see also
+\cite{Fynbo2014,Heintz2017}]{Japelj2015,Zafar17}, emission lines from the
 underlying host galaxies \citep{Kruhler2015}, the frequency of intervening

### 2017-11-22 document/XSGRB_sample.tex
Adding continuum estimate section

--- 2017-10-26 document/XSGRB_sample.tex
@@
 
-The {\it Swift} satellite \citep{Gehrels09}, which was launched in 2004, has
+The {\it Swift} satellite \citep{Gehrels2009}, which was launched in 2004, has
 made it possible to harvest much of the great potential in using GRBs as probes
 of galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.][]{Ricker04}.  A crucial aspect of the great
+from earlier missions \citep[e.g.][]{Ricker2004}.  A crucial aspect of the great
 success of the {\it Swift} mission has been the extensive ground based
@@
 conditions inside the host galaxies were extracted 
-\citep[e.g.,][]{Fiore,Thoene07,Prochaska07,Vreejswijk07,CastroTirado2010}.
+\citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007, Vreeswijk2007, Castro-Tirado2010}.
 
@@
 We additionally provide an estimate of the continuum for all the spectra
-presented here. For this, we have developed an algorithm that attempts to automatically
-finds the continuum placement along with the error on the continuum estimate.
+presented here. For this, we have developed an algorithmic approach that
+attempts to automatically finds the continuum placement along with the error on
+the continuum estimate through an iterative procedure. The method is entirely
+data-driven and does not rely on any physical assumptions. The method is used on
+each arm separately for each spectrum.
+
+To estimate the continuum, a number of points (typically on the order of 100)
+are inserted at random positions along the wavelength direction, the
+flux-density of each point determined by the median value of the spectrum in a
+small region ($\sim 1 \mathrm{\AA}$) surrounding each point. The points are
+fitted with a low order polynomial (we use \texttt{numpy.polynomial.chebyshev})
+and iteratively, the point furthest away are removed until the polynomial fit
+and points differ by less than $\sim$ 5 per cent. This filtering is used because
+the afterglow continuum of GRBs are power laws \citep{Piran2005}, and by
+removing points that differ significantly from a smooth continuum shape will
+guide the continuum estimate to a shape more reasonable for GRB afterglows.
+Additionally, points spaced less than around 1 per cent of the entire spectral
+coverage are pruned. The remaining filtered and pruned points are then spline
+interpolated using \texttt{scipy.interpolate.splrep;splev}, which serves as a
+first estimate of the continuum placement. An attempt to identify absorption and
+emission regions are then carried out, where they are marked as such if the
+difference between the estimated continuum and the observed spectrum is larger
+than 3 - 5 times the associated error spectrum. All regions marked as affected are then masked.
+
+Using the masked spectrum, the entire process is then repeated 500 times where
+the final continuum estimate is the mean of the continuum realizations and the
+associated error estimate is the standard deviation. This error reflects the
+stability of the algorithm across the spectrum. An example of the performance is
+shown in Fig. \ref{fig:spectrum}. In some cases, very little continuum is
+contained in a single arm and a manual continuum estimate is provided, similar
+to what is done in \citet{Lopez2016}. In these cased the continuum error is set
+to 10 per cent. The code for the continuum estimate is released along with the
+paper at \url{https://github.com/jselsing/XSGRB-sample-paper}.
+
+The continuum estimate and the associated error is distributed in a separate
+extension, as described in Sect. \ref{products}.
 
@@
 All the spectra are made available at
-\url{http://www.dark-cosmology.dk/~jselsing/XSGRB/} and additionally at , including the ones outside
+\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally at , including the ones outside
 the statistical sample. This release includes both prompt afterglow observations
@@
 have been repeated, a combined spectrum is also provided.  No attempt has been
-made to combine the arms, so for each observation, three spectra exists.
+made to join the arms, so for each observation, three spectra exists.
 
@@
 
+% Text for redshift distribution section
+One of the primary objectives of our follow-up campaign is to measure the
+redshift distribution for a well-defined, observationally unbiased and
+statistical useful sample of GRBs. The imposed selection criteria (see Sect.
+\ref{samplecrit}) ensures that the GRBs entering our homogenous sample is a good
+representation of the underlying population. The redshift distribution of such
+a sample holds valuable information about the occurance of GRBs through cosmic
+time, which can then be directly compared to the star-formation rate inferred
+from luminosity-selected samples of field galaxies.
+
+In Fig.~\ref{fig:z} we show the redshift distribution of all GRBs observed with
+X-shooter and for all the GRBs entering our sample. In the top panel we show a
+histogram of the two distributions and below we show the redshifts of the
+individual bursts as a function of GRB luminosity in the observed 15-150 keV
+band. To calculate the luminosity, $L_{\mathrm{BAT}}$, we follow a similar
+procedure as Lien(+16). We define this quantity as $L_{\mathrm{BAT}} =
+F_{\gamma}\,4 \pi\,d_L^2\,(1+z)^{-1}$, where $F_{\gamma}$ is the observed BAT
+fluence in the 15-150 keV band and $d_L$ is the luminosity distance to the
+burst at the given redshift. We clearly see a correlation between the minimum
+luminosity of the bursts and the redshift, mainly due to the Malmquist bias.
+The lower limit in luminosity of our sample as a function of redshift
+corresponds to a sensitivity of $10^{-7}$ erg cm$^{-2}$ s$^{-1}$ in the
+observed 15-150 keV band (contrary to the expected value of $10^{-8}$ erg
+cm$^{-2}$ s$^{-1}$; Swift/BAT+05). This is likely due to the more complex
+detection criteria defined by Swift \textbf{(expand by Daniele?)}. Due to the Malmquist
+bias the Swift luminosity detection limit is almost an order of magnitude
+brighter at $z=2$ than at $z=1$. At $z\geq4$ we are only able to observe GRBs
+that are hundred times brighter than the faintest bursts at $z=1$ and below.
+
+Except for the two intrinsically faintest bursts at $z\approx 1 - 1.5$, the
+statistical sample follows a constant lower limit in observed fluence as a
+function of redshift. This means that our selection is not biased at high
+redshifts except for the luminosity bias invoked by Swift. We find the
+following values for the redshift distribution. First for the statistical
+sample: $z_{\mathrm{avg}}=1.99$, $z_{\mathrm{med}}=1.64$, $z_{14\%}=0.60$ and
+$z_{86\%}=3.39$. For all GRBs observed with X-shooter we find
+$z_{\mathrm{avg}}=1.90$, $z_{\mathrm{med}}=1.52$, $z_{14\%}=0.55$ and
+$z_{86\%}=3.36$.
+
 Because the redshift completeness of our statistical sample is just 54 percent
@@
 	\centering
-	\includegraphics[width=16cm]{figures/XSGRB_redshift.pdf}
-	\caption{Redshift distribution of the XSGRB sample.}
+	\includegraphics[width=16cm]{figures/XSGRB_zBATlum.pdf}
+	\caption{Redshift distribution as a function of intrinsic BAT $\gamma$-ray
+		luminosity, $L_{\gamma,\mathrm{int}}$. The latter quantity is defined as
+		$L_{\gamma,\mathrm{int}} = F_{\gamma,\mathrm{obs}}\,4 \pi\,d_L^2\,(1+z)^{-1}$,
+		where $F_{\gamma,\mathrm{obs}}$ is the observed BAT fluence in the 15-150 keV
+		band and $d_L$ is the luminosity distance at the given redshift. Blue star
+		symbols denote the statistical sample and black dots show all GRBs observed
+		with X-shooter and by \textit{Swift}/BAT. Overplotted as the red star symbols
+		for comparison are the GRB, quasar, and galaxy with the highest
+		spectroscopically confirmed redshifts, the latter two shown at arbitrary
+		$L_{\gamma,\mathrm{int}}$. The blue solid line represents the BAT sensitivity
+		of $10^{-8}$ erg cm$^{-2}$ s$^{-1}$ in the 15-150 keV band. The blue dashed
+		line marks an observed sensitivity of $10^{-7}$ erg cm$^{-2}$ s$^{-1}$, which
+		represents our sample limit out to $z\approx 8$. The estimated epoch of
+		reionization is shown by the black dot-dashed line, with the uncertainty shown
+		as the black dashed lines, from the most recent measurement by Planck+16.}
 	\label{fig:z}
@@
 $\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
-in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006,
-	Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
+in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006b,
+	Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
 spectrum from X-rays to the sub-mm) we hence characterize the properties of
@@
 extinction curves [][see also
-\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
+\cite{Fynbo2014, Heintz2017}]\citep{Japelj2015, Zafar2017}, emission lines from the
 underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
 \ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
-radiation \citep{Tanvir17}.
+radiation \citep{Tanvir2017}.
 
@@
 and from studies of complete samples of GRB host galaxies 
-\citep{Hjorth2012,Covinto2013,Perley2016}. 
+\citep{Hjorth2012, Covino2013, Perley2016}. 
 \cite{Kruhler2013} convincingly argue, that 
 very rich sightlines like that probed by the remarkable GRB\,080607 
-\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
+\citep{Prochaska2009} are probably significantly more frequent than in the sightlines 
 sampled by our spectra. However, with current instrumentation these sightlines are
@@
 \subsection{GRB120119A (z = 1.728)}
-The data presented here has not been published before. Three epochs of
+The data presented here has been examined by Japelj+15 and Zafar+17 who both 
+find a significant amount of extinction, $A_V\approx 1$ mag. Three epochs of
 observations have been obtained, the first two immediately after the burst, and
@@
 \textit{Swift} trigger, the first epoch contains bright afterglow continuum.
-Rich in absorption features belonging to a multitude of ions, $z =	1.728$ is
+Rich in absorption features belonging to a multitude of ions at $z =    1.728$ is
 estimated for the host with intervening systems at $z = 1.476$, $z = 1.214$, $z
@@
 GRB in which emission lines from \hb~and \ha~are found at the redshift of the
-host, confirming the association of the absorption line system and the host.
+host, confirming the association of the absorption line system and the host. We
+also detect C\,\textsc{i} in absorption which indicates the presence of cold gas.
 
@@
 \subsection{GRB140506A  (z = 0.889)}	
+The data presented here has formed the basis of GCN \#
+16214\footnote{\url{http://gcn.gsfc.nasa.gov/other/140506A.gcn3}} and is
+published in Fynbo et al. (2014) and Heintz et al. (2017a). The observations
+consists of $4\times 600$ sec at 8.8 and 33 hours after the burst. We detect
+absorption features from a wide range of ions, together with molecular
+absorption from CH+, all at $z=0.889$. The optical/near-infrared afterglow
+reveal an unusual steep extinction curve which is found to be caused by dust
+immediate to the burst.
 
@@
 The data presented here also formed the basis of GCN \#
-19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, but is not
-published elsewhere. Observations started 22.37 hours after the BAT trigger and
-lasted for 2.4ks. The afterglow continuum is detected across the entire
+19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, and is
+published in Heintz et al. (2017b). Observations started 22.37 hours after the
+BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the entire
 spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii~and
@@
 3.6ks, is obtained after the afterglow has faded, confirming the emission line
-detections.
+detections. The host galaxy is found to have a roughly solar metallicity and is
+among the most luminous GRB hosts at $z < 1$.
 

### 2017-11-22 document/XSGRB_sample.tex
Adding data science format description

@@
 
+%All the raw data files along with the associated calibration files are available at the ESO archive \footnote{\url{http://archive.eso.org/wdb/wdb/eso/xshooter/form}}
+
 All the spectra are made available at
-\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally at , including the ones outside
-the statistical sample. This release includes both prompt afterglow observations
-as well as late time observations. An overview of all the spectra provided here
-are given in Table. \ref{sample_overview}. For each burst, each individual
-observation is provided in a separate reduction and in cases where observations
-have been repeated, a combined spectrum is also provided.  No attempt has been
-made to join the arms, so for each observation, three spectra exists.
-
-%We have
-%adopted the following naming convention GRB*OB*.fits
-%
-%
-%In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
-%
-%
+\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally through the
+ESO archive in the form of phase 3 material
+\footnote{\url{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}}. This
+release includes the spectra of GRBs outside the statistical sample and includes
+both prompt afterglow observations as well as late time observations of the
+associated hosts and represents \textit{all} observations of GRBs and their
+hosts carried out by VLT/X-shooter from the commissioning of the instrument,
+\startdate, and until the end of the last period of the program 098.A-0055,
+\termdate and thus constitutes 8 years of GRB observations with VLT/X-shooter.
+An overview of all the spectra provided here are given in Table.
+\ref{sample_overview}. For each burst, each individual observation is provided
+in a separate reduction and in cases where observations have been repeated for
+an increased S/N or to follow the temporal evolution, a combined spectrum is
+also provided. No attempt has been made to join the spectroscopic arms, so for
+each observation, three spectra are provided in separate files.
+
+All spectra are released in the ESO Science Data Product (SDP) format
+\citep{Micol2016}, and formatted as binary FITS files. The naming convention is
+based on the GRB name and the observation number, and follow the scheme
+\texttt{GRBxxxxxxx\_OBxarm.fits}. For example, the visual arm of the third
+observation of GRB~151021A, observed in RRM mode (see Sect. \ref{RRM}), is named
+\texttt{GRB151021A\_OB3VIS.fits}.
+
+Each file contains 7 columns with following contents and descriptions:
+\begin{itemize}
+	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion and drifts in the wavelength solution ($\mathrm{\AA}$).
+	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
+	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
+	\item QUAL - Bad pixel map, where a value different from zero indicates a bad pixel.
+	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
+	\item CONTINUUM\_ERR - Relative error on the continuum estimate 
+	\item TELL\_CORR - Inverse transmission spectrum. Multiply FLUX and ERR column with this column to correct for telluric absorption. 
+\end{itemize}
 
@@
 \todo{Get number} bursts has been triggered by \textit{Swift} in the period
-since the commissioning of VLT/X-shooter in November, 2009. This sample
+since the commissioning of VLT/X-shooter in March, 2009. This sample
 constitutes the \textit{statistical sample} on which we will try to address the

### 2017-11-22 py/Sample_completeness_old.py

@@ -71,14 +71,14 @@ def main():
     HI_c, HI_o, HI_s = HI_c[(HI_c > 0)], HI_o[(HI_o > 0)], HI_s[(HI_s > 0)]
 
     # Plot
-    fig, (ax1, ax2) = pl.subplots(ncols=2)
+    fig, (ax1, ax2, ax3) = pl.subplots(ncols=3)
 
 
 
     # BAT fluence
-    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
-    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
-    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=False, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
     print(len(BAT_c[~np.isnan(BAT_c)]))
     l, m, h = np.percentile(np.log10(1e-7*BAT_c), [16, 50, 84])
     print(m, m - l, h - m)
@@ -97,9 +97,9 @@ def main():
 
 
     # XRT flux
-    sns.distplot(np.log10(1e-11*XRT_c), ax=ax2, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
[…]

### 2017-11-23 document/XSGRB_sample.tex
Remove bad nhx

@@
 discussed as part of the investigation of the GRB population. A prime example of
-a burst outside the statistical sample is GRB161023A with it's rich
+a burst outside the statistical sample is GRB~161023A with it's rich
 spectrum containing 11 intervening absorption systems and pending further analysis.
@@
 continuum, which again depends on the afterglow brightness at time of follow-up.
-Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better references} a rapid
-follow-up is essential for the accurate designation of the GRB host \todo{Get
-reference}. In Fig. \ref{fig:timing} we plot the follow-up delay from the BAT
-trigger to the start of the spectroscopic observation. Visible as the
-shortest delays are the bursts observed in RRM-mode in which where the fastest follow-up
-between BAT trigger and start of spectroscopic observations is for the short, $z = 1.717$, GRB160410A which was
-followed up only 8.4 minutes after the BAT trigger.
+Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better
+	references} a rapid follow-up is essential for the accurate designation of the
+GRB host \todo{Get reference}. In Fig. \ref{fig:timing} we plot the follow-up
+delay from the BAT trigger to the start of the spectroscopic observation.
+Visible as the shortest delays are the bursts observed in RRM-mode in which
+where the fastest follow-up between BAT trigger and start of spectroscopic
+observations is for the short, $z = 1.717$, GRB160410A which was followed up
+only 8.4 minutes after the BAT trigger.
 

### 2017-11-23 py/Sample_completeness.py

@@ -76,9 +76,9 @@ def main():
 
 
     # BAT fluence
-    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
-    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
-    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
+    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Full Swift sample")
+    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Complete sample")
+    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Observed sample")
     print(len(BAT_c[~np.isnan(BAT_c)]))
     l, m, h = np.percentile(np.log10(1e-7*BAT_c), [16, 50, 84])
     print(m, m - l, h - m)
@@ -97,9 +97,9 @@ def main():
 
 
     # XRT flux
-    sns.distplot(np.log10(1e-11*XRT_c), ax=ax2, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
-    sns.distplot(np.log10(1e-11*XRT_s), ax=ax2, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"})
[…]

### 2017-11-24 py/Sample_completeness.py
Doing number to-dos and update sample comparison

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -18,8 +19,8 @@ params = {
    'legend.fontsize': 10,
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
-   'text.usetex': False,
-   'figure.figsize': [9, 4.5]
+   'text.usetex': True,
+   'figure.figsize': [7.281, 4.5]
    }
 mpl.rcParams.update(params)
 
@@ -32,7 +33,8 @@ def filt_nan(input_array):
         try:
             holder.append(float(ii))
         except:
-            pass
+            holder.append(np.nan)
+
     return np.array(holder)
 
 
@@ -44,7 +46,7 @@ def main():
 
     # Read in burst list
     # burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
-    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
+    burst_table = pd.read_csv("../data/Burst list - master.csv")
 
     name, z, observed, OA, BAT, XRT, HI = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values, burst_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values.astype("float"), burst_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values.astype("float"), burst_table["XRT Column Density (NH) [10^21 cm^-2]"].values.astype("float")
 
@@ -52,12 +54,14 @@ def main():
     XRT_o = XRT[observed == "Yes"]
     HI_o = HI[observed == "Yes"]
 
-    BAT_c = BAT[observed == "No"]
-    XRT_c = XRT[observed == "No"]
[…]

### 2017-11-24 py/Sample_completeness_nhx.py

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -19,7 +20,7 @@ params = {
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
    'text.usetex': False,
-   'figure.figsize': [9, 4.5]
+   #'figure.figsize': [4.5, 7.281]
    }
 mpl.rcParams.update(params)
 
@@ -44,135 +45,93 @@ def main():
 
     # Read in burst list
     # burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
-    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
+    burst_table = pd.read_csv("../data/Burst list - master.csv")
 
-    name, z, observed, OA, BAT, XRT, HI = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values, burst_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values.astype("float"), burst_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values.astype("float"), burst_table["XRT Column Density (NH) [10^21 cm^-2]"].values.astype("float")
+    name, z, observed, OA, NHX, NHXh, NHXl = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values, burst_table["Ave_NH_PC"].values.astype("float"), burst_table["Ave_dNH_PC+"].values.astype("float"), burst_table["Ave_dNH_PC-"].values.astype("float")
 
-    BAT_o = BAT[observed == "Yes"]
-    XRT_o = XRT[observed == "Yes"]
-    HI_o = HI[observed == "Yes"]
+    print(z)
 
[…]

### 2017-11-24 py/betaOX.py

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -72,6 +73,9 @@ def main():
     name = np.array([ii[3:] for ii in name])
     # print(name)
     acqmag = filt_nan(acqmag, fill_value=24)
+    idx_limit = (acqmag == 24)
+
+
     X_Fnu = filt_nan(X_Fnu)
     # Optical flux in microjansky
     O_Fnu = 10**((23.9 - acqmag)/2.5)
@@ -83,12 +87,13 @@ def main():
     betaOX = np.log10(O_Fnu/X_Fnu)/np.log10(X_nu0/O_nu0)
     # Mask bad values and reverse order
     mask = np.isnan(betaOX) | (sample == "No")
-
     betaOX = betaOX[~mask][::-1]
     name = name[~mask][::-1]
-    for ii, kk in enumerate(betaOX):
-        print(name[ii], kk)
-    exit()
+    idx_limit = idx_limit[~mask][::-1]
+
+    # for ii, kk in enumerate(betaOX):
+    #     print(name[ii], kk)
+    # exit()
     fynbo_table = np.genfromtxt("../data/Fynbo2009betaOX.dat", dtype=None)
     f_name = fynbo_table[:, 0]
     fynbo_table[:, 1] = [ii.replace('<', '') for ii in fynbo_table[:, 1]]
@@ -98,25 +103,43 @@ def main():
 
 
     # Get Swift HI values
-    swift_table = pd.read_table("../data/grb_table_1482495106.txt", delimiter="\t", dtype=None)
-    name_s, BAT_s, XRT_s, HI_s = swift_table["GRB"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values, swift_table["XRT Column Density (NH) [10^21 cm^-2]"].values
+    swift_table = pd.read_table("../data/allSwiftGRBs_NH.txt", delimiter="\t", dtype=None)
[…]

### 2017-12-05 document/tables/sample_properties.tex
Doing number to-dos and update sample comparison
Small update

@@
 		\hline\noalign{\smallskip}
-		N$_{BAT}$ & 1071 & 156 & 85\\
-		$\log(15-150$keV fluence)  & $-5.9_{-0.7}^{+0.6}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$   \\
-		N$_{XRT}$ & 1004 & 154 & 84\\
-		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.6}^{+0.8}$  \\
-		N$_{HI}$ & 1034 & 155 & 85\\
-		$\log(N_H)$ & $21.4_{-0.5}^{+0.5}$ &  $-21.3_{-0.6}^{+0.6}$ &  $21.4_{-0.7}^{+0.7}$  \\
+		N$_{BAT}$ & 987 & 156 & 89\\
+		$\log(15-150$keV fluence)  & $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.7}^{+0.7}$   \\
+		N$_{XRT}$ & 909 & 154 & 88\\
+		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.7}^{+0.9}$  \\
+		N$_{\mathrm{HI_x}}$ & 250 & 95 & 74\\
+		$\log(\mathrm{N}_{\mathrm{HI_x}})$ & $21.7_{-0.9}^{+0.6}$ &  $21.5_{-3.4}^{+0.7}$ &  $21.6_{-4.5}^{+0.7}$  \\
 
@@
 	\end{tabular}
-	\caption{Population properties for the \textit{Swift sample} and the subset of bursts fulfulling the sample criteria.
-The population characteristics for the three samples exhibit a large degree of similarity, signifying the ability of our selection criteria to effectively capture the underlying population.
-\label{tab:sample_properties}}
-\end{table*}
+	\caption{Population properties (median and 14th and 86th percentiles as the
+	error intervals) for the \textit{Swift sample} and the subset of bursts
+	fulfilling the sample criteria. The population characteristics of the three
+	samples are very similar, which shows that our selection criteria effectively
+	conserve the statistical properties of the underlying population, as least for these parameters.  \label{tab:sample_properties}} \end{table*}

### 2017-12-05 py/Sample_completeness.py

@@ -60,8 +60,8 @@ def main():
 
     swift_table = pd.read_table("../data/grb_table_1511519199.txt", delimiter="\t", dtype=None)
     name_s, BAT_s, XRT_s, HI_s = swift_table["GRB"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values, swift_table["XRT Column Density (NH) [10^21 cm^-2]"].values
-    print(len(name_s))
-    exit()
+    # print(len(name_s))
+    # exit()
     # Exclude sample bursts
     idx = [ii for ii, kk in enumerate(name_s) if kk not in name]
 
@@ -203,9 +203,9 @@ def main():
 
 
 
-    ax1.set_xlabel(r"log(15-150 keV Fluence) [erg/cm$^2$]")
-    ax2.set_xlabel(r"log(0.3-10 keV Flux) [erg/cm$^2$/s]")
-    ax3.set_xlabel(r"log(N$_H$) [cm$^2$]")
+    ax1.set_xlabel(r"log(15-150 keV Fluence) [$\mathrm{erg}~\mathrm{cm}^{-2}$]")
+    ax2.set_xlabel(r"log(0.3-10 keV Flux) [$\mathrm{erg}~\mathrm{cm}^{-2}\mathrm{s}^{-1}$]")
+    ax3.set_xlabel(r"log($N_{\mathrm{HI,X}}$) [$\mathrm{cm}^{-2}$]")
     # ax1.set_ylabel(r"N")
 
     ax2.set_xlim(-18, -9)

### 2017-12-05 py/follow_up_timing.py

@@ -5,6 +5,8 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
+
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -30,18 +32,15 @@ def main():
 
 
     # Read in burst list
-    # burst_table = np.array([[ii, kk, ll, pp, tt] for ii, kk, ll, pp, tt in np.genfromtxt("../data/burst_list.dat", dtype=None)])
-    burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
-
+    burst_table = pd.read_csv("../data/Burst list - observed.csv")
     name, z, delay, mag = burst_table["GRB"].values, burst_table["z"].values, burst_table["Follow-up delay"].values, burst_table["Acquisition mag"].values
-    
 
 
 
-    for kk, ll in list(zip(name, z)):
-        print(kk, cosmo.luminosity_distance(ll).to(u.cm))
 
-    exit()
+    # for kk, ll in list(zip(name, z)):
+    #     print(kk, cosmo.luminosity_distance(ll).to(u.cm))
+    # exit()
 
     for ii, ll in enumerate(mag):
         # print(ll)
@@ -56,6 +55,9 @@ def main():
     delay_sort = np.argsort(delay)
     sorted_delay = delay[delay_sort]
     sorted_z = z[delay_sort]
+    print(len(sorted_z))
+    print(sorted_z)
+    exit()
     fractional_completeness = (1 - np.cumsum(np.isnan(sorted_z).astype("int"))/len(sorted_z))*100
     # print(len(sorted_z), np.sum(np.isnan(sorted_z).astype("int")))
     # print(fractional_completeness[-1])
@@ -73,12 +75,11 @@ def main():
     # With redshift and acq mag
     sc = ax1.scatter(delay[idx_hasz & idx_limits], mag[idx_hasz & idx_limits], c=z[idx_hasz & idx_limits], cmap=cmap, s=25)
     # Without redshift and with acq mag
-    ax1.scatter(delay[~idx_hasz & idx_limits], mag[~idx_hasz & idx_limits], color="black", s=25)
[…]

### 2017-12-06 py/Get_z_bat.py
Redshift distilled updated

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+from matplotlib.ticker import FormatStrFormatter
+from astropy.cosmology import Planck15 as cosmo
+import astropy.units as u
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': True,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
+
+def filt_nan(input_array):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except:
+            holder.append(np.nan)
+
+    return np.array(holder)
+
+
+def main():
+    """
+    Small script to get Swift redshift BAT .
+    """
+    # swift_table = pd.read_table("../data/allSwiftGRBs_NH.txt", delimiter="\t", dtype=None)
+    # # print(swift_table)
+    # name_s, z_s, NHX_s, NHXh_s, NHXl_s = swift_table["#GRB"].values, swift_table["Redshift"].values, swift_table["Ave_NH_PC"].values, swift_table["Ave_dNH_PC+"].values, swift_table["Ave_dNH_PC-"].values
+
+    swift_table = pd.read_table("../data/grb_table_1511519199.txt", delimiter="\t", dtype=None)
+    name, BAT, z = swift_table["GRB"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["Redshift"].values
+    z_out = []
+    idxs = []
+    for ii, kk in enumerate(z):
+      try:
+        z_out.append(kk.split()[0])
+        idxs.append(ii)
+      except:
+        pass
+
+    name = name[idxs]
+    BAT = filt_nan(BAT[idxs])
[…]

### 2017-12-06 py/DarknessHI.py
redshift and darkness

@@ -98,10 +98,10 @@ def main():
     name_g_ot = name_g[idx][(OT[idx] == "y") & (inst[idx] == "Swift")]
     name_g_not = name_g[idx][~(OT[idx] == "y") & (inst[idx] == "Swift")]
 
-    print(idx)
+    # print(idx)
 
-    print((len(name_g_not))/(len(name_g_ot) + len(name_g_not)))
-    exit()
+    # print((len(name_g_not))/(len(name_g_ot) + len(name_g_not)))
+    # exit()
     print(len(BAT_ot))
     mask_ot = (BAT_ot > -8) & (BAT_ot < -3) & (HI_ot > 19) & (HI_ot < 24)
     g = sns.JointGrid(x=BAT_ot[mask_ot], y=HI_ot[mask_ot], xlim = (-8, -3), ylim = (19, 24), space=0)

### 2017-12-06 py/Get_z_bat.py

@@ -71,6 +71,7 @@ def main():
     # L_bat = np.array(L_bat)
 
 
+    
     pl.scatter(z, np.log10(L_bat))
     pl.show()
 

### 2017-12-06 py/betaOX.py

@@ -172,8 +172,11 @@ def main():
 
     # print(stats.ks_2samp(HI, f_HI))
     # print(len(HI))
-    # l, m, h = np.percentile(HI[betaOX[idx_2] < 0.5], [16, 50, 84])
-    # print(m, m - l, h - m)
+
+    l, m, h = np.percentile(HI[betaOX[idx_2] >= 0.5], [16, 50, 84])
+    print(m, m - l, h - m)
+    l, m, h = np.percentile(HI[betaOX[idx_2] < 0.5], [16, 50, 84])
+    print(m, m - l, h - m)
 
     #Fraction of dark bursts
     print(len(HI[betaOX[idx_2] < 0.5])/(len(HI[betaOX[idx_2] < 0.5]) + len(HI[betaOX[idx_2] >= 0.5])))

### 2017-12-06 py/betaOX.py
darkness

@@ -172,16 +172,20 @@ def main():
 
     # print(stats.ks_2samp(HI, f_HI))
     # print(len(HI))
-
+    print("Above 0.5")
     l, m, h = np.percentile(HI[betaOX[idx_2] >= 0.5], [16, 50, 84])
+    print(HI[betaOX[idx_2] >= 0.5])
     print(m, m - l, h - m)
+    print("Below 0.5")
     l, m, h = np.percentile(HI[betaOX[idx_2] < 0.5], [16, 50, 84])
+    print(HI[betaOX[idx_2] >= 0.5])
     print(m, m - l, h - m)
 
     #Fraction of dark bursts
     print(len(HI[betaOX[idx_2] < 0.5])/(len(HI[betaOX[idx_2] < 0.5]) + len(HI[betaOX[idx_2] >= 0.5])))
     print(stats.ks_2samp(HI[betaOX[idx_2] < 0.5], HI[betaOX[idx_2] >= 0.5]))
-    # exit()
+    print(len(HI))
+    exit()
     # print(len(f_HI))
     # l, m, h = np.percentile(f_HI, [16, 50, 84])
     # print(m, m - l, h - m)

### 2017-12-06 py/Get_z_bat.py
zbatl

@@ -71,9 +71,26 @@ def main():
     # L_bat = np.array(L_bat)
 
 
-    
-    pl.scatter(z, np.log10(L_bat))
-    pl.show()
+    np.savetxt("../data/zbatl.dat", list(zip(z, np.log10(L_bat))))
+
+
+
+
+
+    L_bat = []
+    z = np.arange(0, 15, 0.01)
+    for ll in z:
+        dL = cosmo.luminosity_distance(ll).to(u.cm)
+        BAT_flu = 3 * 1e-8 * u.erg * u.cm**(-2)
+        L_bat.append(BAT_flu * 4 * np.pi * dL**2 / (1 + ll)/u.erg)
+    # L_bat = np.array(L_bat)
+
+
+    np.savetxt("../data/zbatlswift.dat", list(zip(z, np.log10(L_bat))))
+
+
+
+
 
 if __name__ == '__main__':
   main()

### 2017-12-07 document/XSGRB_sample.tex
Adding RRM text

--- 2017-11-22 document/XSGRB_sample.tex
@@
 
-\abstract{
-In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
-afterglows that have \textit{ever} been observed at the VLT/X-shooter up to the end of 2016. This
-totals 137 spectroscopic observation of 118 individual bursts, of which
+\abstract{In this work we present the spectra of all $\gamma$-ray bursts (GRBs)
+	afterglows that have been observed at the VLT/X-shooter up to \termdate. This
+	totals 137 spectroscopic observations of 118 individual bursts, of which
 	redshifts have been secured for 91 \% covering a redshift range from 0.059 to
-$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
-has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
-of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
-We here provide a public release of all the spectra, including continuum
-estimates and telluric corrections. We provide an assessment of the degree of
-completeness with respect to the parent GRB population, in terms of the X-ray
-properties of the bursts in the sample and find that the sample presented here
-can not be significantly distinguished from the \textit{Swift} sample.
-For all bursts (39) for which it is possible, we provide a measurement of the
-optically derived hydrogen column density, increasing the total number of
-published HI column densities by $\sim$ 50 \% and characterize the distribution
-compared with literature values.
-We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and tentatively confirm previous
-results that larger optical darkness leads to increased X-ray absorption.
-%This dataset provides a unique possibility to study the ISM across cosmic time.
+	$\sim$ 8.5. Some of these observation are late-time observations of the
+	underlying host galaxy. Based on an set of selection criteria that are unbiased
+	with regards to intrinsic properties of the GRBs, the follow-up effort has been
+	guided by the \textit{Swift}-telescope to produce a homogeneous sample of
+	afterglows for 86 spectra. We here provide a public release of all the reduced
+	spectra, including continuum estimates and telluric corrections. We provide an
+	assessment of the degree of completeness with respect to the parent GRB
+	population, in terms of the X-ray properties of the bursts in the sample and
+	find that the sample presented here is consistent with being drawn from the
+	\textit{Swift} sample. For all bursts (39) for which it is possible, we provide
+	a measurement of the hydrogen column density, increasing the total number of
+	published HI column density measurements by $\sim$ 50 \%. We constrain the
+	fraction of dark bursts to be 27\% - 42\% \todo{Preliminary} and tentatively
+	confirm previous results that larger optical darkness is correlated with
+	increased X-ray absorption. This dataset provides a unique possibility to study
+	the ISM across cosmic time.
+
 }
 
-\keywords{Gamma-ray burst: individual: GRB~120815A --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
+\keywords{Gamma-ray burst: general --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
 
@@
 Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
-derive new insights into a very wide range of topics in astrophysics. Examples
-range from small scale phenomena relating to the magnetars, properties of highly
-relativistic jets, hyper/supernova explosions, the interstellar medium, dust extinction
-curves, starburst galaxies, chemical and molecular abundances, escape of ionizing 
-radiation, the ionization state of the intergalactic medium, intervening
-absoption systems to standard candles in cosmology.  In that sense GRB research
-is very multidisciplinary.
+derive new constraints on a very wide range of topics in astrophysics. Examples
+range from small scale phenomena relating to magnetars, properties of highly
+relativistic jets, hyper/supernova explosions, the interstellar medium, dust
+extinction curves, starburst galaxies, chemical and molecular abundances, escape
+of ionizing radiation, the ionization state of the intergalactic medium,
+intervening absoption systems to standard candles in cosmology
+\citep[e.g.,][]{Lyons2010, Molinari2007, HjorthBloom2012, Wijers1998,
+	Prochaska2009, Savaglio2006, Ghirlanda2007}. In that sense GRB research is very
+multidisciplinary.
 
@@
 of galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.][]{Ricker2004}.  A crucial aspect of the great
+from earlier missions \citep[e.g.,][]{Ricker2004}. A crucial aspect of the great
 success of the {\it Swift} mission has been the extensive ground based
@@
 
-In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
-was secured using low-resolution spectrographs 
-\citep[typically
+In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy was
+secured using low-resolution spectrographs \citep[typically
 R=$\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Low-resolution
-spectroscopy is powerful for a number of reasons, first of all is the ability
-to secure information for even very faint targets. This allows the measurements
-of a number of important parameters such as redshifts, spectral slopes,
-and extinction. For
-a handful of very bright afterglows high-resolution (typically R$>$20000)
-spectra were secured and for these events much more information about
-conditions inside the host galaxies were extracted 
-\citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007, Vreeswijk2007, Castro-Tirado2010}.
-
-The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
-generation instruments at the ESO Very Large Telescope (VLT). It is a
-high-throughput echelle spectrograph covering the full spectral range from the
-UV atmospheric cutoff around 300 nm to the near-infrared (near-IR) K-band 
-ending around 2500 nm with three spectral
-arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
-in mind as the fading luminosities of such sources makes in urgent to secure as
-extensive coverage as possible in the shortest possible time. At the same time
-the resolution was set to be in the range 4000--9000 in order to be able to get
-a large useful spectral range between the many sky-background emission lines in
-the red and near-IR spectral ranges. In this paper we present the results
-of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
-spectrograph to secure spectroscopic observations of
-afterglows and host galaxies of GRBs detected by {\it Swift}. 
-All data resulting from the survey are publicly availble in reduced form.
+spectroscopy is powerful for a number of reasons, first of all is the ability to
+secure information for even very faint targets. This allows the measurements of
+a number of important parameters such as redshifts, spectral slopes, and
+extinction. For a handful of very bright afterglows high-resolution (typically
+R$>$20000) spectra were secured, and for these events much more information
+about conditions inside the host galaxies were extracted
+\citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007, Vreeswijk2007,
+	Castro-Tirado2010}.
+
+The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
+instruments at the ESO Very Large Telescope (VLT). It is a high-throughput
+echelle spectrograph covering the full spectral range from the UV atmospheric
+cutoff around 300 nm to the near-infrared (near-IR) K-band ending around 2500 nm
+with three spectral arms (UVB, VIS, NIR). It was designed very much with
+transient source follow-up in mind as the fading luminosities of such sources
+makes it urgent to secure as extensive coverage as possible in the shortest
+possible time. At the same time, the resolution was designed to be in the range
+4000--9000 in order to be able to get a large useful spectral range between the
+many sky-background emission lines in the red and near-IR spectral ranges.
+
+In this paper, we present the results of a dedicated effort over the years 2009
+-- 2017 to use the X-shooter spectrograph to secure spectroscopic observations
+of afterglows and host galaxies of GRBs detected by {\it Swift}. We here make
+all data resulting from the survey, publicly available in reduced form.
 
@@
 the sample including the sample selection criteria and the observational
-strategy.  In Sec.t~\ref{results} we describe the results of the survey, i.e.
-the efficiency of the follow-up effort and the characteristics of the observed
-bursts. We also here assess the completeness of the finally realized sample.
-Finally, we offer our conclusions in Sect.~\ref{conclusions}.
+strategy. In Sect.~\ref{obs}, we describe the observations and the instrumental
+setups, and in Sect.~\ref{proc} we detail the methodological strategies adopted
+in the data reduction process and auxiliary material. In Sect.~\ref{results} we
+describe the results of the survey, i.e. the efficiency of the follow-up effort
+and the characteristics of the observed bursts. We also here assess the
+completeness of the finally realized sample. Finally, we offer our conclusions
+in Sect.~\ref{conclusions}.
 
@@
 of the telescope or poorly localized bursts as reported by the alerting
-telescope. To maximize the return of the follow-up campaign we have chosen a few
-selection criteria that attempts to provide an unbiased selection of bursts,
-while allowing for a high follow-up success rate. The importance of defining
-unbiasing selection criteria has been highlighted previously
-\citep{Salvaterra2012, Hjorth2012, Vergani2015, Perley2016a}, when trying to
-address the specific underlying distribution functions such as the redshift
-distribution, host metallicity distribution, or brightness distribution. When
-investigating a specific distribution function, a high degree of
-\textit{completeness} is desired meaning that in terms of the parameter of
-interest, the largest possible fraction of targets in the sample has
-determinations of that parameter \citep{Perley2016b}.
-
-We simultaneously want to minimize any biases against astrophysical conditions
-while at the same time maximizing likelihood of observations allowing us to
-obtain a higher degree of observing completeness. By restricting the selection
-criteria to conditions local to Galaxy and therefore irrespective of intrinsic
-GRB properties, the hope is that the sample collect represents the underlying
-distribution of GRBs in a fair way while at the same time, additional unbiasing
-cuts can be imposed that can improve the follow-up success. The selection
-criteria used here is based on previous successes \cite{Jakobsson2006b,
-	Fynbo2009, Hjorth2012}, which has shown to represent the underlying distribution
-to a high degree. We test the degree of completeness in Sect.
-\ref{completeness}.
-
-The selection criteria are defined as follows:
+telescope. To maximize the return of the follow-up campaign, we have chosen a
+few selection criteria that facilitates an unbiased selection of bursts, while
+at the same time allowing for a high follow-up success rate. The importance of
+defining unbiased selection criteria has been highlighted previously
+\citep{Jakobsson2006b, Salvaterra2012, Hjorth2012, Vergani2015, Perley2016a},
+when trying to address the intrinsic underlying distribution functions such as
+the redshift distribution, host metallicity distribution, or afterglow
+brightness distribution. When investigating a specific distribution function, a
+high degree of \textit{completeness} is desired \citep[e.g.,][]{Perley2016b}.
+
+In defining the selection criteria, we simultaneously aim to minimize any biases
+against intrinsic astrophysical conditions while at the same time maximizing the
+likelihood of successful observations, hence allowing us to obtain a higher
+degree of completeness. By restricting the selection criteria to conditions
+local to the Milky Way and therefore independent of intrinsic GRB properties,
+the aim is that the sample collected represents the underlying distribution of
+GRBs in a fair way. The selection criteria used here is based on previous,
+similar studies \citep{Jakobsson2006b, Fynbo2009, Hjorth2012}. We characterize
+the sample completeness in Sect. \ref{completeness}.
+
+Our selection criteria are:
 
 \begin{enumerate}
-	\item GRB triggered onboard by Swift.
-	\item XRT started observing within 10 minutes since the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12.
+	\item GRB triggered by \textit{Swift}.
+	\item XRT started observing within 10 minutes after the GRB; an XRT position must be distributed within 12 hr.
+	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12 degrees.
 	\footnote{Note that in the P84 proposal the criteria have
@@
 	defining the sample.}
-	\item No bright closeby stars.
+	\item No closeby bright stars.
 	\item Galactic $A_V \lesssim 0.5$ mag.
@@
 
-Because our ability to observe GRB afterglows is strongly correlated with the
-rapidness and precision of the target positions delivered by the triggering
-facilities, by selecting only bursts that have been triggered on board the
-\textit{Swift} space telescope \cite{Gehrels2004}, by the Burst Alert Telescope
-(BAT) \citep{Barthelmy2005}, we start out with a homogeneously selected sample
-where burst characteristics are delivered immediately, allowing for an informed
+Our ability to observe GRB afterglows is strongly dependent on the timing and
+the precision of the target positions delivered by the triggering facilities. By
+selecting only bursts that have been triggered on board the \textit{Swift} space
+telescope \citep{Gehrels2004}, based on the Burst Alert Telescope (BAT)
+\citep{Barthelmy2005}, we  start out with a homogeneously selected sample where
+burst characteristics are delivered immediately, allowing for an informed
 follow-up strategy. Despite the complexity of the triggering mechanism on board
-\textit{Swift} \citep{Band2006}, attempts at inferring properties of the
-underlying cosmic GRB population based on the detection thresholds and
-triggering algorithms have been made \citep{Lien2014, Graff2016} and thereby by
-restricting the follow-up effort to bursts detected by \textit{Swift}, the
-limitations of the parent sample is well studied.
+\textit{Swift} \citep{Band2006, Coward2013a}, attempts at inferring properties
+of the underlying GRB population based on the detection thresholds and
+triggering algorithms have been made \citep{Lien2014, Graff2016}. Restricting
+the follow-up effort to bursts detected by \textit{Swift}, we therefore ensure
+that the limitations of the parent sample is well studied.
 
 Because the localization accuracy of BAT is 1 - 4 arcminutes
-\citep{Barthelmy2005}, we additionally require an X-ray position from the X-ray
-Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
-\citep{Barthelmy2000} with 10 hours and to account for observing constraints on
-\textit{Swift} that XRT began observations within 10 minutes. The additional
-rapidness requirement of the XRT follow-up means that the number of detected
-X-ray afterglows for BAT triggers is 100 percent \citep{Burrows2007}. The
-\textit{Swift} constraints on the GRB follow-up means that the sample presented
-here inherits the \textit{Swift} sample properties, except the additional XRT
-constraint. Because the XRT completeness is very high\citep{Burrows2007}, this
-should change the parent sample significantly. This issue is addressed again in
-Sect. \ref{completeness}.
-
-To allow for observability we additionally require that the GRB is visible
-(duh!) from the telescope site at Cerro Paranal, Antofagasta, Región de
-Antofagasta, Chile, for a least 1 hour allowing time for the spectroscopic
+\citep{Barthelmy2005}, a secure host identification is in general impossible
+based on BAT alone. We therefore additionally require an X-ray position from the
+X-ray Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
+\citep{Barthelmy2000} within 10 hours and to account for observing constraints
+on \textit{Swift} that XRT began observations within 10 minutes. The additional
+timing requirement of the XRT follow-up means that all bursts in our sample have
+detected X-ray afterglows. Because the XRT completeness is very high
+\citep{Burrows2007}, this cut should not alter the parent sample significantly.
+
+To ensure a minimum of observability, we require that the GRB is visible from
+the telescope site at Cerro Paranal, Chile, for a least 1 hour after the trigger
+with the sun below $-12$ degrees. This secures time for the spectroscopic
 observations to completed. Since the GRB population is isotropically distributed
-seen from Earth and that GRB properties does not depend on position on the sky
-\citep{Briggs1996}, this cut does not influence our ability to sample the
-underlying GRB population. The same arguments applies to the requirement that
-there are no nearby bright stars, which is additionally supplemented by the
-requirement what the galactic dust content does not exert an extinction higher
-than $A_V \lesssim 0.5$ mag based on the extinction maps by
-\citet{Schlegel1998}. Because of  isotropy, these additional cuts should not
-influence the optical properties of the bursts themselves, only our ability to
-successfully secure observations that allows us to investigate the spectroscopic
-properties of GRBs.
-
-We address the succesfullness of these criteria and the actual follow-up effort
-in representing the full \textit{Swift} sample ion Sect. \ref{completeness}.
+on the sky as seen from Earth and because the GRB properties does not depend on
+position on the sky \citep{Briggs1996}, this cut does not influence our ability
+to fairly sample the underlying GRB population. The same arguments applies to
+the requirement that there are no nearby bright stars. We additionally require
+that the Galactic extinction is below $A_V \lesssim 0.5$ mag, based on the
+extinction maps by \citet{Schlegel1998}. Because of isotropy, these additional
+cuts should not influence the optical properties of the bursts themselves, only
+our ability to successfully secure observations that allows us to investigate
+the spectroscopic properties of GRBs.
+
+We further discuss the effect of these selection criteria and their implication
+for the compleness of the sample in Sect. \ref{completeness}.
 
@@
 
-\todo{Will be written by Daniele?}
+\todo{Will be written by Daniele}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The rapid-response mode is
+
+Under rare circumstances, the use of the rapid response mode (RRM) has been
+possible. In case of a promptly visible GRB, a robotic trigger was sent to
+the telescope if at the moment of the GCN notice the GRB is observable and
+fulfills the following criteria.
+
+\begin{enumerate}
+	\item GRB triggered by \textit{Swift}.
+	\item X-shooter mounted at the telescope
+	\item GRB promptly visible
+	\item Error radius < 60\arcsec.
+	\item No limit on declination. 
+	\item No constraint on Moon phase or distance. 
+	\item Elevation in the sky > 22$^\circ$ (both now and 15 min after the trigger).
+	\item Sun elevation < -12$^\circ$ (both now and 15 min after the trigger).
+	\item The X-ray position must be available less than 1 hr after the GRB.
+	\item The flags KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true).
+\end{enumerate}
+
+The use of RRM is unique in the sense that it allowes for the detailed study of
+the temporal variability of the GRB afterglow due to effect of the GRB itself on
+the sorrounding medium \citep[e.g., see][for examples]{Dessauges-Zavadsky2006,
+	Vreeswijk2007, Vreeswijk2013}
+
+For 8 of the observed GRBs in the spectra released here, RRM triggered. One of
+the RRM triggers is outside the statistical sample and two of the RRM trigger
+are on short GRBs. The fastest response between \swift~GRB trigger and the
+beginning of spectroscopic integration is for GRB~160410A, for which the delay
+was only 8.4 minutes.
 
@@
 
-Observations in this sample has been carried out using the cross-dispersed
-echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on two of the four
-Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the duration
-of this follow-up campaign, which covers the entire lifetime of X-shooter.
-Observations have been carried out from the ESO period P84 through P98 under the
-following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.A-0045,
-094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088, 089.A-0067,
-088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. A few additional
-bursts, not a part of the \textit{statistical sample}, have been included from
-092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI: Greiner), and Levan under
-095.B-0811(B) (PI: Levan). This sample represents \textit{all} GRBs that have
-ever been followed up by X-shooter.
-
-The first GRB followed up is GRB090313, observed 15th of March 2009, which was
-while X-shooter was mounted on UT3 during the commissioning of the instrument.
-The bursts observed during the commissioning period is not a part of our
-statistical sample, but are nonetheless published as part of the bursts observed
-by the X-shooter GRB team. The first burst observed after science verification
-was completed when X-shooter was moved to UT2 is GRB091018 which thereby
-constitute the first burst entering our statistical sample. For all bursts that
-fulfill our sample selection criteria, described in Sect. \ref{samplecrit},
-spectroscopic follow-up has been attempted with X-shooter. Various conditions
-can affect our ability to follow up a given burst, and a discussion of this
-effect is included in Sect. \ref{badbursts}.
-
-X-shooter can cover the spectral wavelength region from 300 nm to 2480 nm in a
-single exposure, by splitting the light in three separate spectroscopic arms
-using dichroics.  This way each arm work as a separate instrument, each
-functioning as its own echelle spectrograph.  The ultraviolet blue(UVB) arm
-covering 3000 - 550 nm, the visual(VIS) arm covering 550 - 1020 nm, and the
-near-infrared()NIR) arm covering 10200 - 2480 nm with the possibility of
-applying a k-band blocking filter cutting the coverage of the NIR arm at 1020 -
-2100 nm. For all observations, a nod-and-shuffle observing scheme has been
-employed, with a nodding throw of 5 \arcsec. The individual observations has
-been carried out in a standard ABBA pattern, which allows flexibility in the sky
-subtraction method for the different arms.
-
-For the large majority of the bursts we have observed with a slit width of
-1\farc0, 0\farc9, and 0\farc9 for the UVB, VIS, and NIR-arm respectively, which
-puts a lower limit on the delivered resolution of the spectra based on the
-tabulated values of the delivered resolutions, which is 4350, 7450, and 5300 for
-the UVB, VIS and NIR-arm respectively\footnote{\url{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}}.
-The slit width also sets the width of the atmospheric sky lines and determines
-the amount of light lost due to the wavelength-dependent seeing PSF extending
-outside the coverage of the slit, where the width of sky-lines is always set by
-the slit width whereas both the delivered resolution and the slitloss changes
-for the better as the seeing PSF drops below the slit width.
-For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9
-observed with a 0\farc9 slit only 76.1 percent of the light will enter the slit,
-meaning that for almost all observations a slitloss correction is required. We
-describe how slitlosses were corrected for in Sect. \ref{postproc}.
-For accurate measurements of velocity widths, a precise instrumental resolution
-is required and this becomes better when the delivered seeing is better than the
-slit width. We discuss this in Sect. \ref{resolution}.
-
-Due to mechanical issues, the atmospheric dispersion corrector (ADC) were
+The observations obtained for this sample have been secured with the
+cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
+one of the two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during
+the duration of this follow-up campaign. The observations have been taken during
+a period of 8 years corresponding to the ESO observing periods P84 through P98
+under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
+095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
+089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. We
+have included a few additional bursts, from the programmes 092.D-0056(A) (PI:
+Rau), 092.D-0633(E) (PI: Greiner), and 095.B-0811(B) (PI: Levan). The total
+collection of spectra represents \textit{all} GRBs that have been followed up by
+X-shooter up to \termdate.
+
+The first GRB followed up is GRB090313, observed on the 15th of March, 2009,
+during the commissioning of X-shooter UT3. The bursts observed during the
+commissioning or science verification (GRB~090313, GRB~090530, GRB~090809,
+GRB~090926) are not a part of our statistical sample. The first burst observed
+after science verification and X-shooter was moved to UT2, is GRB~091018, which
+thereby constitute the first burst entering our statistical sample. For all
+bursts that fulfill our sample selection criteria, described in Sect.
+\ref{samplecrit}, spectroscopic follow-up has been attempted with X-shooter.
+Various conditions can affect our ability to follow up a given burst, and a
+discussion of these conditions and their consequences for the sample is
+included in Sect. \ref{badbursts}.
+
+X-shooter covers the spectral wavelength region from 300 nm to 2480 nm in a
+single exposure, by separating the light into three separate spectroscopic arms
+through the use of two dichroics. The ultraviolet blue (UVB) arm covers 300 -
+550 nm, the visual (VIS) arm covers 550 - 1020 nm, and the near-infrared (NIR)
+arm covers 1020 - 2480 nm. For some of the observations, we have applied a
+$K$-band blocking filter, cutting the coverage of the NIR arm at 2100 nm. This
+is done to reduce the amount of scattered background light from the thermal
+infrared. For the majority of observations, a nod-and-shuffle observing scheme
+has been employed, with a nodding throw of 5\arcsec. Each nodding observation
+have typically been carried out in a standard ABBA pattern. For some cases,
+conditions during the follow-up, either technical or weather, have necessitated
+alterations to this scheme as described in App. \ref{notes}. For RRM triggers, a
+slightly different observing strategy was employed. Starting as rapidly as
+possible, a simple stare mode sequence was started, with 5 spectroscopic
+integrations with increasing exposure times.
+
+For the majority of the bursts, we have observed with a slit width of 1\farc0,
+0\farc9, and 0\farc9 for the UVB, VIS, and NIR-arm respectively. This sets a
+lower limit on the delivered resolution of the spectra based on the tabulated
+values of the delivered resolutions, which is 4350, 7450, and 5300 for the UVB,
+VIS and NIR-arm
+respectively\footnote{\url{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}}.
+%The slit width sets the width of the atmospheric sky lines and determines the amount of light lost due to the wavelength-dependent seeing PSF extending outside the coverage of the slit, where the width of sky-lines is always set by the slit width whereas both the delivered resolution and the slitloss changes for the better as the seeing PSF drops below the slit width.
+For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a 0\farc9 slit only 76.1 percent of the light will enter the slit, meaning that for almost all observations a slitloss correction is required.
+We describe how slitlosses were corrected for in Sect. \ref{postproc}. For
+accurate measurements involving line profiles, knowledge of the precise
+instrumental resolution is required. The resolution becomes better than the
+nominal one, listed above, when the delivered seeing is smaller than the
+projected width of the slit on the sky. We discuss how we determine the
+effective instrumental resolution in Sect. \ref{resolution}.
+
+Due to a mechanical failure, the atmospheric dispersion corrector (ADC) was
 disabled from 1st of August 2012. Only GRB~100728B was affected by the failing
@@
 chromatic slit losses due to atmospheric dispersion, all subsequent observations
-have been carried out at parallactic angle. A consequence of this is that for
-all observations following 1st of August 2012, the source moves across the
-spatial direction of the slit as a function of wavelength. This effect has been
-modeled in the extraction procedure, as described in \ref{extract}.
-
-We list the overview of all the observations in Tab. \ref{tab:sample_overview}.
-\input{tables/burst_overview.tex} and plot the positions off all the bursts on
-the night sky in Fig. \ref{fig:skymap}. The central zone of avoidance due to
-high galactic extinction cutoff is visible with an apparent isotropic
-distribution of the bursts away from the galactic center except in the upper
-right corner, where Earth casts a shadow on the sky as seen from the telescope.
-
-A subset of the bursts presented here  have already had their hosts investigated
-in \citet{Kruhler2015}, for which extractions of the hosts exist. The focus of
-the data presented here are on the afterglows themselves and in the absence of a
-clear afterglow, the host. We will not, however, investigate the hosts.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data processing}
+have been carried out at parallactic angle. A consequence of this, is that for
+all observations following 1st of August 2012, the centroid of spectrum changes
+position across the spatial direction of the slit as a function of wavelength.
+This effect has been modeled in the extraction procedure, as described in Sect. 
+\ref{extract}.
+
+\input{tables/burst_overview.tex}
+
+We provide an overview of all the observations in Table
+\ref{tab:sample_overview} and plot the positions of all the bursts on the
+celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from
+central zone of avoidance due to high galactic extinction cutoff (marked in
+white), the GRB positions have an apparent isotropic distribution except in the
+upper left quadrant, which cannot be observed from Paranal.
+
+A subset of the spectra presented here, primarily host observations, have
+already been published in \citet{Kruhler2015}. We include independent
+reductions of them here for completeness.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data processing} \label{proc}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this section we describe how the final data products are produced and
-subsequently post-processed. $\sim$ 30 percent of the data released here have
-been reduced in a manner similar to what is described in \citet{Kruhler2015}.
-The two reduction schemes have been tested and are in good agreement. All
-post-processing scripts developed for this dataset are made publicly available
-at \url{https://github.com/jselsing/XSGRB_reduction_scripts}, along with
-instructions of use.
-
-Before any reductions are done, the raw object images are run through the
+subsequently post-processed. %$\sim$ 30 percent of the data released here have
+% been reduced in a manner similar to what is described in \citet{Kruhler2015}.
+% The two reduction schemes have been tested and are in good agreement.
+All post-processing scripts developed for this dataset are made publicly
+available at \url{https://github.com/jselsing/XSGRB_reduction_scripts}, along
+with instructions of use.
+
+Before any reductions are initiated, the raw object images are run through the
 cosmic-ray removal algorithm \citep{VanDokkum2001} implementation,
@@
 sources. Because of the broken ADC, for some objects there is a lot of curvature
-in the object trace. This means that for some bursts, the initial sky-estimate
-has been made from a limited number of pixels in the spatial direction. By doing
-an initial subtracting the sky on the un-rectified image we ensure that bulk of
-the sky background is not redistributed by the rectification process.
+in the object trace along the dispersion axis of the slit. This means that for
+some bursts, the initial sky-estimate has been made from a limited number of
+pixels in the spatial direction. By doing an initial subtracting the sky on the
+un-rectified image we ensure that bulk of the sky background is not
+redistributed by the rectification process.
 
@@
 single image. Because this strategy is employed, we can reject outliers in the
-stack and weight by the average inverse variance of the background. When
-weighting images where the noise in each pixel is dominated by Poisson noise it
-important to estimate the background variance in a large enough region, so that
-to remove the correlation between the signal and the weights. To this end, the
-weight map is generated by a running median window over the variance map, where
-the trace as been masked and width of the window is chosen to be wide enough for
-median variance to be generated on the basis of several hundred pixels. This
-weighting scheme automatically also optimally combines images of different
-exposure times or images where the background is varying, which is often the
-case when a burst has been observed close to twilight
-
-Because the background is very bright and there is a high abundance of broad
-sky-lines in the NIR arm, when there are no contaminating sources in the slit,
-the sky has been put back on the images and they have been reduced in pairs of
-two, subtracting the two from each other, keeping the WCS static. This amounts
-to the regular nodding reduction, only we can reject outliers and weight by the
-averaged inverse variance map.
+stack and weight by an averaged measure of the inverse variance of the
+background. When weighting images where the noise in each pixel is dominated by
+Poisson noise it important to estimate the background variance in a large enough
+region, so that any correlation between the signal and the weights are removed.
+To this end, the weight map is generated by a running median window over the
+variance map, where the trace as been masked and width of the window is chosen
+to be wide enough for median variance to be generated on the basis of several
+hundred pixels. This weighting scheme automatically also optimally combines
+images of different exposure times or images where the background is varying,
+which is often the case when a burst has been observed close to twilight
+
+For the NIR arm, because the background is very bright and there is a high
+abundance of broad sky-lines, when there are no contaminating sources in the
+slit, the sky has been put back on the images and they have been reduced in
+pairs of two, subtracting the two from each other, keeping the WCS static. This
+amounts to the regular nodding reduction, only we can reject outliers and weight
+by the averaged inverse variance map.
 
 By STARE reducing all observations we additionally get a spectrum of the sky,
-which we can use to calibrate the wavelength solution in Sect. \ref{wavecal}.
+which we can use to calibrate the wavelength solution.
 
@@
 during operations. The flexures modify the projection of the slit on the
-detector with respect to the one obtained in daytime calibration. This requires a
-modification of the wavelength solution in order to process correctly the
+detector with respect to the one obtained in daytime calibration. This requires
+a modification of the wavelength solution in order to process correctly the
 night-time data. Part of this correction is performed by the pipeline using the
@@
 \footnote{X-shooter User Manual available at
-	https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}.
-
-The remaining offset is corrected by cross-correlation with a synthetic sky
-spectrum \citep{Noll2012, Jones2013} after the continuum, estimated as the mode
-of all flux values, has been subtrated. To get the correct seeing PSF with which
-to convolve the synthetic sky an initial refinement of the wavelength solution
-have been obtained by cross-correlating the observed sky with an unconvolved
-synthetic sky. This preliminary wavelength calibration is applied to the
-observed sky. The synthetic spectrum is then convolved with an increasing seeing
-PSF and the width that minimizes $\chi^2$ with the updated observed sky is
-chosen to be the effective sky-PSF. Using the synthetic sky with the matched
-resolution, a final wavelength calibration can then be calculated by
+	https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}. The
+remaining offset is corrected by cross-correlating the observed sky spectrum
+with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum,
+estimated as the mode of all flux values, has been subtracted. To get the correct
+seeing PSF with which to convolve the synthetic sky an initial refinement of the
+wavelength solution have been obtained by cross-correlating the observed sky
+with an unconvolved synthetic sky. This preliminary wavelength calibration is
+applied to the observed sky. The synthetic spectrum is then convolved with an
+increasing seeing PSF and the width that minimizes $\chi^2$ with the updated
+observed sky is chosen to be the effective sky-PSF. Using the synthetic sky with
+the matched resolution, a final wavelength calibration can then be calculated by
 cross-correlating the observed sky with the correctly broadened sky spectrum, as
@@
 time were applied to the corresponding spectra.
-
-Using the convolved synthetic sky, the $\gtrsim 3 \sigma$ sky brightness pixels
-have been added to the bad pixel map to avoid the cores of the brightest
-sky-lines.
+Using the convolved synthetic sky, the pixels containing the brightest sky lines
+have been added to the bad pixel map.
 
@@
 complexity of the objects entering the slit. Due to the malfunctioning ADC, see
-Sect. \ref{obs}, the spectral trace moves across the slit as a function of
-wavelength for a large fraction of the bursts observed meaning that using a
-single aperture for the spectral extraction is inadequate due to the large
-amount of background that would then enter the slit. To optimally select the
-extraction regions we therefore need to model the trace position.
+Sect. \ref{obs}, the spectral trace changes position across the slit in the
+spatial direction as a function of wavelength. For a large fraction of the
+bursts observed, using a single aperture for the spectral extraction is
+inadequate due to the large amount of background that would then enter the slit.
+To optimally select the extraction regions we therefore need to model the trace
+position.
 
@@
 \citep{Moffat1969} adequately describes an imaging PSF due to atmospheric
-turbulence, but due to the aberrations in the dispersion elements and the
-rectification process, the PSF we are trying to model different from this
+turbulence, but because of aberrations in the dispersion elements and the
+rectification process, the PSF we are trying to model is different from this
 profile. To allow for flexibility in the model, we have chosen the Voigt
@@
 observations have been queried for the ambient conditions which controls the
-degree to which the trace moves.
+degree to which the trace changes position in the spatial direction.
 
@@
 accurate tracing of the PSF, typically 200 bins for moderate signal-to-noise.
-For each of the bins and using the guess position, the spectral PSF has been fit
-using the unweighted chi-squared minimization algorithm implemented in
-\texttt{scipy.optimize.curve\_fit} \citep{scipy}. Since we know that the trace
-varies slowly as a function of wavelength, we have then fitted a low-order
-polynomial to the fit parameters as a function of wavelength, which allows us to
-evaluate the spectral PSF at all wavelengths and this way accurately model the
-entire spectral PSF.
-
-Equipped with a model for how the light falls on the entire image, we can then
-employ the optimal extraction algorithm \cite{Horne1986}, which weights the
-extraction aperture by the spectral profile, or alternatively sum all pixels
-within 1 FWHM of the modeled profile. Where possible, we have used the optimal
-extraction. In cases where the trace is very weak, even in the binned images, an
-aperture has been selected manually which covers emission lines, if present, and
-when nothing is immediately visible, the entire nodding window. The error- and
-bad pixel maps are in all cases propagated throughout the extraction.
+For each of the bins, using the analytically guided guess position, the spectral
+PSF has been fit using the unweighted chi-squared minimization algorithm
+implemented in \texttt{scipy.optimize.curve\_fit} \citep{scipy}. Since we know
+that the trace varies slowly as a function of wavelength, we have then fitted a
+low-order polynomial to the fit parameters as a function of wavelength, which
+allows us to evaluate the spectral PSF at all wavelengths and this way
+accurately model the entire spectral PSF.
+
+Equipped with a model for how the light is distributed across the entire
+dispersion direction, we can  employ the optimal extraction algorithm
+\cite{Horne1986}, which weights the extraction aperture by the spectral profile,
+or alternatively sum all pixels within 1 FWHM of the modeled profile. Where
+possible, we have used the optimal extraction. In cases where the trace is very
+weak, even in the binned images, an aperture has been selected manually which
+covers emission lines, if present, and when nothing is immediately visible, the
+entire nodding window. The error- and bad pixel maps are in all cases propagated
+throughout the extraction.
 
 In cases where multiple traces are visible in the slit, additional components
-for the profile are used in the optimal extraction. The additional components
-do not share the PSF parameters and in cases where the additional component is an
+for the profile are used in the optimal extraction. The additional components do
+not share the PSF parameters and in cases where the additional component is an
 extended object, the fits have been inspected to ensure that the additional
 component does now skew the fit towards a different PSF. The additional
-components are not used for the weights.
-
-The spectra are corrected for galactic extinction using the E(B-V) value from
+components are not used for the weights in the extractions.
+
+The spectra are corrected for Galactic extinction using the E(B-V) value from
 the dust maps of \citet{Schlegel1998} with the update in
@@
 total to selective extinction $R_V = 3.1$. The wavelengths of the extracted
-1D-spectra are moved to vacuum, moved to the barycentric frame, and the
+1D-spectra are moved to vacuum, corrected for barycentric motion, and the
 wavelength recalibration described in Sect. \ref{wavecal} is applied. Pixels
-with pixel-to-pixel variation large than $50 \sigma$ are added to the bad pixel
-map.
+with pixel-to-pixel variation larger than $50 \sigma$ are additionally added to
+the bad pixel map.
 
@@
 
-For all earth-based telescopes, the light of interest has to pass through Earths
+For all earth-based telescopes, the light first has to pass through Earths
 atmosphere, where the atmospheric content and conditions make an imprint on the
@@
 the continuum estimate through an iterative procedure. The method is entirely
-data-driven and does not rely on any physical assumptions. The method is used on
+data-driven and does not rely on any physical assumptions. The method is applied on
 each arm separately for each spectrum.
@@
 are inserted at random positions along the wavelength direction, the
-flux-density of each point determined by the median value of the spectrum in a
-small region ($\sim 1 \mathrm{\AA}$) surrounding each point. The points are
+flux-density of each point is determined by the median value of the spectrum in
+a small region ($\sim 1 \mathrm{\AA}$) surrounding each point. The points are
 fitted with a low order polynomial (we use \texttt{numpy.polynomial.chebyshev})
 and iteratively, the point furthest away are removed until the polynomial fit
-and points differ by less than $\sim$ 5 per cent. This filtering is used because
-the afterglow continuum of GRBs are power laws \citep{Piran2005}, and by
+differ from the points by less than $\sim$ 5 per cent. This filtering is used
+because the afterglow continuum of GRBs are power laws \citep{Piran2005}, and by
 removing points that differ significantly from a smooth continuum shape will
@@
 difference between the estimated continuum and the observed spectrum is larger
-than 3 - 5 times the associated error spectrum. All regions marked as affected are then masked.
+than 3 - 5 times the associated error spectrum. All regions marked as affected
+are then masked.
 
@@
 contained in a single arm and a manual continuum estimate is provided, similar
-to what is done in \citet{Lopez2016}. In these cased the continuum error is set
+to what is done in \citet{Lopez2016}. In these cases the continuum error is set
 to 10 per cent. The code for the continuum estimate is released along with the
 paper at \url{https://github.com/jselsing/XSGRB-sample-paper}.
-
-The continuum estimate and the associated error is distributed in a separate
-extension, as described in Sect. \ref{products}.
 
@@
 The afterglow spectra described in this paper are obtained in
-Target-of-Opportunity (override) mode.
-In most cases there is therefore little possibility to tweak slit widths to the
-seeing at the time of observations (i.e. to optimise spectral resolution and
-signal to noise), and almost all our data is therefore taken with a fixed set
-of slit widths and binning, described above.
-In a fair number of cases, the seeing full width at half maximum (FWHM) is
+Target-of-Opportunity (override) mode. In most cases there is therefore little
+possibility to tweak slit widths to the seeing at the time of observations (i.e.
+to optimize spectral resolution and signal to noise), and almost all our data is
+therefore taken with a fixed set of slit widths and binning, described above. In
+a fair number of cases, the seeing full width at half maximum (FWHM) is
 considerably smaller than the silt width, and the delivered spectral resolution
 will then be determined by the seeing rather than slit width, as afterglows are
-point sources (this is evidently not the case for extended sources, e.g for
-host galaxies).
-The delivered resolution for slit width dominated spectra post-reduction and
-extraction can easily be determined from the bright sky emission lines.
-For afterglow spectra with very high signal to noise, the delivered spectral
-resolution can at times be determined from the science data themselves.
-However, in the presence of multiple velocity components in absorption, other
-forms of line broadening, and a lack of lines at some redshifts, this is
-difficult to do at poorer signal to noise ratios (the majority of spectra in
-our sample).
-A broad starting value for the expect resolution will help fitting of these
-spectra, and can be important in upper limit determination, and for this reason
-we construct a aim to construct a crude relation between the seeing and the
-delivered resolution at our slit width, binning, and reduction pipeline
-settings.
+point sources (this is evidently not the case for extended sources, e.g for host
+galaxies). The delivered resolution for slit width dominated spectra
+post-reduction and extraction can easily be determined from the bright sky
+emission lines. For afterglow spectra with very high signal to noise, the
+delivered spectral resolution can at times be determined from the science data
+themselves. However, in the presence of multiple velocity components in
+absorption, other forms of line broadening, and a lack of lines at some
+redshifts, this is difficult to do at poorer signal to noise ratios (the
+majority of spectra in our sample). A broad starting value for the expect
+resolution will help fitting of these spectra, and can be important in upper
+limit determination, and for this reason we construct a crude relation between
+the seeing and the delivered resolution at our slit width, binning, and
+reduction pipeline settings.
+
 To this end we use observations of telluric standard stars that are taken with
 identical instrument settings as our afterglow spectra, usually just after the
-science data, as part of the ESO X-shooter calibration plan.
-These spectra have been reduced together with the afterglow spectra, using
-identical pipeline settings with the same version of the pipeline.
-First we fit a Gaussian function in the spatial direction of the trace of the
-standard star at 792 nm (i.e. in the VIS arm).
-After this, we fit a series of  20 telluric absorption lines in the telluric
-standard star spectra with Gaussians, taking care to select transitions that
-are not almost-resolved multiples, should be intrinsically unresolved, and are
-in areas with well defined continuum flux.
-We pick 34 telluric standard stars spanning a range of DIMM seeing values, with
-the majority between $0.5-1.5 arcsec$.
-The resulting distribution of spectral FWHM (km/s) as a function of spatial
-FWHM at 792 nm is fairly well described by a linear relation $a + b*x$, with
-$x$ the spatial FWHM in pixels (with 0.15 arc sec per pixel),  $a= 21.4 \pm
-1.3$ km/s, $b=1.4 \pm 0.2$.
-We use this linear relation as a way to estimate the spectral resolution for
-medium to poor signal to noise afterglow spectra in the VIS arm.
-To extend this to the UVB and NIR arm, we measured a series of lines in NIR arm
-spectra of a subset of 19 sources used for the VIS arm above, and find that the
-resulting distribution is consistent with a simple scaling of the VIS arm
-relation by the ratio of resolutions of the NIR and VIS arm for unresolved,
-slit filling, sources as given on the ESO instrument website.
-The UVB arm contains no suitable absorption lines to use, and we therefore use
-a scaled value as in the NIR arm.
-While this simple method is not terribly accurate (for one, the spatial profile
-of the trace is not a perfect Gaussian), but it gives a sufficiently accurate
-estimate for the analysis of these poor signal to noise science spectra.
+science data, as part of the ESO X-shooter calibration plan. These spectra have
+been reduced together with the afterglow spectra, using identical pipeline
+settings with the same version of the pipeline. First we fit a Gaussian function
+in the spatial direction of the trace of the standard star at 792 nm (i.e. in
+the VIS arm). After this, we fit a series of  20 telluric absorption lines in
+the telluric standard star spectra with Gaussians, taking care to select
+transitions that are not almost-resolved multiples, should be intrinsically
+unresolved, and are in areas with well defined continuum flux. We pick 34
+telluric standard stars spanning a range of DIMM seeing values, with the
+majority between $0.5-1.5 arcsec$. The resulting distribution of spectral FWHM
+(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
+linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
+sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$. We use this linear
+relation as a way to estimate the spectral resolution for medium to poor signal
+to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR
+arm, we measured a series of lines in NIR arm spectra of a subset of 19 sources
+used for the VIS arm above, and find that the resulting distribution is
+consistent with a simple scaling of the VIS arm relation by the ratio of
+resolutions of the NIR and VIS arm for unresolved, slit filling, sources as
+given on the ESO instrument website. The UVB arm contains no suitable absorption
+lines to use, and we therefore use a scaled value as in the NIR arm. While this
+simple method is not terribly accurate (for one, the spatial profile of the
+trace is not a perfect Gaussian), but it gives a sufficiently accurate estimate
+for the analysis of these poor signal to noise science spectra.
 
@@
 both prompt afterglow observations as well as late time observations of the
-associated hosts and represents \textit{all} observations of GRBs and their
-hosts carried out by VLT/X-shooter from the commissioning of the instrument,
-\startdate, and until the end of the last period of the program 098.A-0055,
-\termdate and thus constitutes 8 years of GRB observations with VLT/X-shooter.
-An overview of all the spectra provided here are given in Table.
-\ref{sample_overview}. For each burst, each individual observation is provided
+associated hosts and represents \textit{all} afterglow spectra of GRBs carried
+out by VLT/X-shooter from the commissioning of the instrument, \startdate, and
+until the end of the last period of the program 098.A-0055, \termdate and thus
+constitutes 8 years of GRB afterglow observations with VLT/X-shooter. An
+overview of all the spectra provided here are given in Table.
+\ref{tab:sample_overview}. For each burst, each individual observation is provided
 in a separate reduction and in cases where observations have been repeated for
@@
 
-In this section we describe the efficiency of the follow-up effort and the
+In this section, we describe the successfulness of the follow-up effort and the
 characteristics of the observed bursts. We also assess the degree to which the
-sample obtained is complete. An important note is that here we present
-\textit{all} GRBs that have ever been observed with X-shooter, while only a
-subset of these constitute our \textit{statistical} sample. The statistical
-sample is based on the selection criteria described in Sect. \ref{samplecrit}
-and should represent an unbiased sample of the underlying, parent GRB
-population. Single bursts outside the sample criteria have been followed up on
-the basis of the curiosity of their light curve or their brightness, but is not
-discussed as part of the investigation of the GRB population. A prime example of
-a burst outside the statistical sample is GRB161023A with it's rich
-spectrum containing 11 intervening absorption systems and pending further analysis.
+obtained sample is representative for the full \textit{Swift} sample. An
+important note is that here we include \textit{all} GRB after \startdate, that
+have been observed with X-shooter, while only a subset of these constitute our
+\textit{statistical} sample. The statistical sample is based on the selection
+criteria described in Sect. \ref{samplecrit}. Some bursts not fulfilling the
+sample criteria have been followed up due to interesting characteristics, e.g.
+curiosity properties of their light curves, their brightness, etc. These bursts
+are not discussed as part of the investigation of the statistical properties of
+the GRB population. A prime example of a burst outside the statistical sample is
+bright INTEGRAL burst GRB~161023A, that turned to be a rich spectrum with at
+least 11 intervening absorption systems.
 
@@
 
-Redshift determination of bursts where the host is too faint for a spectroscopic
-redshift relies on the absorption spectrum imprinted on the GRB afterglow
-continuum, which again depends on the afterglow brightness at time of follow-up.
-Because the afterglow rapidly fades \citep{Nousek2006, Vecchio2016}\todo{Better references} a rapid
-follow-up is essential for the accurate designation of the GRB host \todo{Get
-reference}. In Fig. \ref{fig:timing} we plot the follow-up delay from the BAT
-trigger to the start of the spectroscopic observation. Visible as the
-shortest delays are the bursts observed in RRM-mode in which where the fastest follow-up
-between BAT trigger and start of spectroscopic observations is for the short, $z = 1.717$, GRB160410A which was
-followed up only 8.4 minutes after the BAT trigger.
+Redshift determination of bursts for which the host is too faint for a
+spectroscopic redshift measurement relies on the detection of absorption line
+imprinted on the GRB afterglow continuum. Because the afterglow rapidly fades
+(typically as $\sim t^{-1}$) a rapid follow-up is essential. In Fig.
+\ref{fig:timing} we plot the delay from the BAT trigger to the start of the
+spectroscopic observation. The shortest delays are observed in RRM-mode. The
+fastest follow-up between BAT trigger and start of spectroscopic observations in
+our sample is for the short, $z = 1.717$, GRB160410A for which spectroscopic
+integration was initialed only 8.4 minutes after the BAT trigger. To illustrate
+importance of the follow-up delay for the redshift completeness, we plot the
+redshift completeness as a function of delay time in Fig. \ref{fig:timing} for
+all the bursts we have followed up, including the ones outside the statistical
+sample. Visible is the decreasing fraction of GRBs with a redshift determination
+for increasing delays. The overall redshift completeness for bursts that we have
+followed up is 92 per cent. This degree of completeness in the followed bursts,
+illustrates the efficiency of VLT/X-shooter in redshift determination. Not shown
+in the figure are an additional 12 bursts that have redshift determinations
+based on late-time host observations with delay times longer than $\sim$10 days.
 
 \begin{figure}
-	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}}
-	\caption{Follow-up timing and afterglow magnitude at the start of observation.
-	The points have been colored based on the redshift of the corresponding burst.
-	Points with black faces does not have a redshift determination and downarrows
-	indicate a non-detection in the acquisition image. In blue is shown the
-	fractional redshift completess as a function of delay-time.}
+	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}} \caption{afterglow
+	magnitude at the start of observation as a function of follow-up delay. The
+	points have been colored based on the redshift of the corresponding burst. Red
+	crosses indicate GRBs without a measured redshift and arrows indicate bursts
+	for which the afterglow was not detected in the acquisition image. In red is
+	shown the redshift completeness as a function of follow-up delay.}
 \label{fig:timing}
 \end{figure}
-
-To illustrate the effect of redshift completeness loss due to late follow-up, we
-overplot the redshift completeness fraction over the delay time. Visible is the
-increasing fraction of GRBs without a good redshift determination for increasing
-delays. That the redshift completeness of the bursts we have followed up are $>$
-90 percent \todo{Get number}, shows the ability of VLT/X-shooter to successfully
-get redshifts. Not shown in the figure is an additional 12 \todo{Get number}
-bursts that we have redshift determinations from late-time host observations
-with delay-times longer than $\sim$10 days.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
-\todo{Get number} bursts has been triggered by \textit{Swift} in the period
-since the commissioning of VLT/X-shooter in March, 2009. This sample
-constitutes the \textit{statistical sample} on which we will try to address the
-GRB population properties. Of this sample, 86 \todo{Get number} have been
-spectroscopically followed up with X-shooter. In order to assess whether our
-sample criteria and the subset of bursts followed up represent the underlying
-parent population of GRB, we compare some intrinsic GRB properties of our sample
-to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
-that \textit{Swift} randomly samples the underlying GRB population, if the
-sample statistics of our subsample are similar to the \textit{Swift}-sample, we
-can assume that the population properties should be conserved. There is some
-evidence that the burst duration does not scale with redshift as expected
-through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
-that we could be missing an increasing fraction of the bursts at higher
-redshifts, but \citet{Lien2016} argues that we should only be missing the bright
-"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
-the sample presented here is complete. We show the comparison between the burst
-properties in Fig. \ref{fig:swift_complete}.
+bursts have triggered \textit{Swift} in the period since the commissioning of
+VLT/X-shooter. This sample constitutes the \textit{statistical sample} from
+which we will derive statistical properties of the GRB population. The redshift
+completeness of the full statistical sample is 60 per cent. We return to the
+question of redshift completeness in in Sect. \ref{redshift}. From this sample,
+90 GRBs have been spectroscopically followed up with X-shooter. In order to
+assess whether the subset of bursts followed up are representative of the
+underlying GRB parent population, we compare intrinsic properties of GRBs in our
+sample to GRBs in the full sample followed up by \textit{Swift}. We show the
+comparison between the BAT (15-150 keV) fluence, the XRT flux (0.3 - 10 keV) at
+11 hours, and the intrinsic X-ray derived equivalent hydrogen column density,
+$N_{\mathrm{HI,X}}$, in Fig. \ref{fig:swift_complete}. For the latter, we can
+only use values of $N_{\mathrm{HI,X}}$ derived for bursts with a measured
+redshift, excluding $\sim 75$ per cent of the bursts. We return to this last
+point in Sect. \ref{darkness}.
 
@@
 	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
-	\caption{Comparison between bursts properties of all bursts observed with
-	\textit{Swift}, the properties of the ones that fullfill the sample criteria
-	specified in Sect. \ref{samplecrit}, and the subset that has been observed as
-	part of the statistical sample. The left plot shows the fluence in 15-150 keV
-	as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
-	after the bursts as measured by XRT. Based on the XRT spectrum,
-	\citep{Evans2009}, a hydrogen column density has been inferred which we show in
-	the right most panel.}
+\caption{Comparison between burst properties of all bursts observed with
+	\textit{Swift} (blue), the ones that fullfill the sample criteria specified in
+	Sect. \ref{samplecrit} (green), and the subset that has been observed as part of
+	the statistical sample (red). The left plot shows the fluence in in range 15-150
+	keV as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
+	after the bursts as measured by XRT. In the right most panel, we show
+	$N_{\mathrm{HI,X}}$ based on the XRT spectrum \citep{Evans2009}.}
 \label{fig:swift_complete}
@@
 
-Using the observational characteristics of the 1244 \todo{Get number} bursts
-observed by \textit{Swift} to date \cite{Gehrels2009}, and the X-ray derived
-hydrogen column densities \citep{Evans2009}, we can quantify the degree to which
-our sample is biased against the overall \textit{Swift} sample in terms of
-intrinsic GRB properties. The values are queried from the online \textit{Swift}
-database maintained\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
-Three main samples are of interest to compare to assess the completeness of the
-follow-up campaign; the full \textit{Swift} sample consisting of all the bursts
-observed by \textit{Swift}, excluding the bursts entering our sample (1), the
-burst that fulfill the selection criteria imposed in Sect. \ref{samplecrit}, but
-has not been followed up (2), and the bursts actually followed up with X-shooter
-(3).
+Using the observational characteristics of the 1322 bursts observed by
+24.10.2017 by \textit{Swift}, and $N_{\mathrm{HI,X}}$ \citep{Evans2009}, we can
+quantify the degree to which our sample is biased relative to the overall
+\textit{Swift} sample. The values are queried from the online \textit{Swift}
+database\footnote{\url{http://swift.gsfc.nasa.gov/archive/grb\_table/}}$^,$\footnote{\url{http://www.swift.ac.uk/xrt\_live\_cat/}}.
+Three samples are of interest in order to assess the completeness of the follow-up campaign (Fig.~\ref{fig:swift_complete}); the full \textit{Swift} sample consisting of all the bursts observed by \textit{Swift} (blue), all the bursts that fulfill the selection criteria imposed in Sect.
+\ref{samplecrit} (green), and the bursts actually followed up with X-shooter (red).
 
@@
 
-For each of the samples we calculate the median and 14th and 86th percentile
-which can be used as point estimates for the population distribution. These are
-shown in \ref{tab:sample_properties}. It can be seen from the values that the
-three samples have very similar distributions in terms of the point estimates
-chosen, suggesting that our selection criteria are unbiasing against the
-\textit{Swift}-sample and that additionally, the follow-up effort is unbiasing
-towards the intrinsic GRB properties. This is despite spectroscopic follow-up
-only being carried out in cases where either a detectable optical afterglow or a
-clear counterpart are seen, which naively should be biased against dark bursts
-occurring in more obscured galaxies, which is shown to exhibit different
-galactic properties \citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b,
-	Perley2015b}.
+For each of the samples, we calculate the median, 14th, and 86th percentiles of
+each of the distributions, which can be used as point estimates for the
+population distribution. These are provided in Table.
+\ref{tab:sample_properties}. It can be seen from the values that the three
+samples have very similar distributions in terms of the point estimates chosen.
+This suggests that our selection criteria are unbiased compared to the
+\textit{Swift}-sample and that additionally, the follow-up effort conserves the
+distributions of the intrinsic GRB properties (except perhaps for $N_{\mathrm{HI,X}}$, see Sect. \ref{darkness}).
 
 Additionally, using a 2-sided Kolmogorov-Smirnov test (KS-test), we can assess
-the degree to which the null that the two distribution are drawn from the same
-distribution can be rejected. We show a graphic representation of the test
-statistics in Fig. \ref{fig:p_values}. We have run the test on whether is can be
-rejected that the sample criteria is drawn from the same distribution as the
-\textit{Swift} sample, if the observed sample is dissimilar from the statistical
-sample, and again whether the observed sample differs from the full
-\textit{Swift} sample. A high p-value indicates little evidence against the null
-meaning for the parameters investigated here, only the X-ray derived column
-densities exhibits the highest degree of dissimilarity, but not to a significant
-level. A large part of the tension between the two samples in $N_H$ is driven by
-a small number of low column densities($\log(N_H) \lesssim 19 \mathrm{cm}^2$)
-bursts.
+the degree to which the null, that the two distributions are drawn from the same
+parent distribution, can be rejected. We show a graphic representation of the
+test statistics in Fig. \ref{fig:p_values}. A high p-value indicates little
+evidence against the null, meaning for the properties investigated here. The
+distribution of $N_{\mathrm{HI,X}}$ exhibits the highest degree of
+dissimilarity, but the two distribution are still consistent with being drawn
+from the same underlying distribution.
 
@@
 
-We therefore conclude that the sample presented here, to a high degree
-represents the intrinsic properties of the GRBs in the full \textit{Swift} sample.
-\todo{This doesn't quite feel exhaustive? Is there a better way of addressing the issue of completeness}
+We therefore conclude that the statistics of the sample presented here,
+conserves the intrinsic properties of the GRBs in the full \textit{Swift}
+sample. At least in terms of BAT fluence and X-ray afterglow flux at 11 hours.
 
@@
 
-GRB follow-up requires the immediate availability of the telescope in order to
-successfully carry out spectroscopic observations. Out of the 158 \todo{Get
-exact number} bursts in the statistical sample, 32($\sim$ 20 percent) \todo{Get
-exact number} have not been observed due to conditions, not relating the GRB or
-afterglow properties. The reasons includes unavailability of the telescope due
-to technical maintenance such af mirror re-coating, a visitor rejecting the TOO
-trigger, or bad weather at the observing site. Because this cut is irrespective
-of the GRB properties, it is not expected that this should change sample
-properties. Addressing whether this cut introduces a bias in the followed sample
-is not trivial, but in terms of intrinsic GRB properties, the sample
-distributions does not change significantly, as shown in Sect. \ref{completeness}.
-\todo{Should this be quantified somehow?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sample darkness}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A subset of all GRBs exhibit no detectable or very faint optical afterglow
-\citep{Groot1998, Djorgovski2001, Fynbo2001}, which had been parametrized in
-term of "darkness" \citep{Jakobsson2004, VanderHorst2009}. The X-ray properties
-of this subset has previously been investigated \citep{DePasquale2003, Fynbo2009,
-	Melandri2012}, finding a slightly higher X-ray luminosity compared to the
-optically brights bursts combined with similar prompt characteristics pointing
-towards a denser circumburst material. This indicates along with investigations
-of host galaxy extinction \citep{Greiner2011, Kruhler2011, Hjorth2012}, that the
-extinction of the optical afterglow is primarily driven by the presence of dust,
-local to the burst \todo{Local dust reference}. Because the dust content of a
-galaxy is correlated with the gas amount \citep{Bohlin1978, Guver2009}, we
-expect the X-ray derived column densities for the bursts without an optical
-afterglow to be higher than the ones without.
-
-For all bursts with follow-up within 100 hours we can calculate the
-"darkness"-parameter, $\beta_{OX}$ \citep{Jakobsson2004, Milvang-Jensen2012}. This required the
-simultaneous measurement of the X-ray flux density and the optical flux density
-which is in practice never possible. We use the measured acquisition camera
-magnitude reported in Tab. \ref{tab:sample_overview} to get the optical flux
-density at the beginning of the spectroscopic integration. Because we know the
-delay between the follow-up and the \textit{Swift} trigger, we can use the
-measured XRT burst lightcurve \citep{Evans2007,
-	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
-to infer the corresponding X-ray flux density. This is done by either linearly
-interpolating between temporally neighboring XRT measurements of by
-extrapolating the last few X-ray data points to the spectroscopic delay. When
-the afterglow is not detected in the acquisition camera, an upper limit of $> 24$
-mags is used, which propagates into an upper limit on $\beta_{OX}$.
-
-In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
-the one presented in \citet{Fynbo2009}. We have replaced limits with values,
-meaning that we artificially bias the distribution towards higher
-$\beta_{OX}$-values because darker bursts naturally fainter in the optical
-making the fraction of non-detections higher for lower $\beta_{OX}$. We
-marginally confirm the result by \citet{Fynbo2009} that dark bursts $\beta_{OX}
-< 0.5$) have slightly higher X-ray derived column densities, by calculating the
-median, 14th, and 86th percentile of the distributions arising by separating the
-sample at $\beta_{OX} = 0.5$. For bursts with $\beta_{OX} \geq 0.5$ we find
-$\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
-$21.8_{-0.9}^{+0.5}$, meaning that the median value for dark bursts is higher,
-but with a great deal of overlap. A 2-sided KS test fails to reject the null
-that they are drawn from the same distribution with $p = 0.43$, meaning there is
-no clear evidence for a discrepancy.
-
-\begin{figure}
-	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
-	\caption{Optical extinction relative to X-ray against X-ray derived hydrogen
-	column densities. In red is shown the sample presented in \citet{Fynbo2009}
-	where is lines indicate the kernel density estimate of the distribution where
-	limits have been replaced with values. Darker colors represented denser points.
-	The correcspoing marginal distributions are shown. In blue is the points for
-	the bursts presented here along with the marginal histograms. Limits on
-	$\beta_{OX}$ is shown by downwards facing arrows.}
-	\label{fig:betaOX}
-\end{figure}
-
-Because we only follow up targets with either a clear afterglow or counterpart,
-correctly associating a galaxy with a burst when there is no detected optical
-afterglow is more difficult, see \citet{Levesque2010} and \citet{Perley2017} for
-a recent example. Because the presence of optical transient is highly dependent
-on the "darkness", we expect that the $\beta_{OX}$ sample is biased against dark
-bursts. The larger relative fraction of upper limits for increased darkness is
-visible both in the sample presented here and the one in \citet{Fynbo2009}. The
-X-ray properties for the followed sample does not differ signicantly from the
-statistical sample, as shown in Sect. \ref{completeness}. Using the table
-maintained by Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html} \todo{How
-to reference this}, we can see how the presence of an optical afterglow affects
-the follow-up statistics. 50.5 percent of all \textit{Swift}-detected bursts do
-not have an detected optical afterglow, but this number also includes bursts
-where no optical telescopes were available, so is a lower limit. For the bursts
-entering our statistical sample, this number is $\sim$ 27 percent, close to the
-maximum fraction of dark bursts \citep{Melandri2012}. Of the subsample (3) that
-have actually been attempted, this number is $\sim$ 18 percent, confirming that
-we are indeed biased against bursts with an optical afterglow in the
-spectroscopic sample.
-
-Since \textit{all} GRBs should have an optical afterglow, the presence of an optical afterglow is more a statement about the sensitivity than optical properties of the bursts.
-\todo{Hmm, how do I finish this?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{On the redshift distribution of GRBs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Text for redshift distribution section
-One of the primary objectives of our follow-up campaign is to measure the
-redshift distribution for a well-defined, observationally unbiased and
-statistical useful sample of GRBs. The imposed selection criteria (see Sect.
-\ref{samplecrit}) ensures that the GRBs entering our homogenous sample is a good
-representation of the underlying population. The redshift distribution of such
-a sample holds valuable information about the occurance of GRBs through cosmic
-time, which can then be directly compared to the star-formation rate inferred
-from luminosity-selected samples of field galaxies.
-
-In Fig.~\ref{fig:z} we show the redshift distribution of all GRBs observed with
-X-shooter and for all the GRBs entering our sample. In the top panel we show a
-histogram of the two distributions and below we show the redshifts of the
-individual bursts as a function of GRB luminosity in the observed 15-150 keV
-band. To calculate the luminosity, $L_{\mathrm{BAT}}$, we follow a similar
-procedure as Lien(+16). We define this quantity as $L_{\mathrm{BAT}} =
-F_{\gamma}\,4 \pi\,d_L^2\,(1+z)^{-1}$, where $F_{\gamma}$ is the observed BAT
-fluence in the 15-150 keV band and $d_L$ is the luminosity distance to the
-burst at the given redshift. We clearly see a correlation between the minimum
-luminosity of the bursts and the redshift, mainly due to the Malmquist bias.
-The lower limit in luminosity of our sample as a function of redshift
-corresponds to a sensitivity of $10^{-7}$ erg cm$^{-2}$ s$^{-1}$ in the
-observed 15-150 keV band (contrary to the expected value of $10^{-8}$ erg
-cm$^{-2}$ s$^{-1}$; Swift/BAT+05). This is likely due to the more complex
-detection criteria defined by Swift \textbf{(expand by Daniele?)}. Due to the Malmquist
-bias the Swift luminosity detection limit is almost an order of magnitude
-brighter at $z=2$ than at $z=1$. At $z\geq4$ we are only able to observe GRBs
-that are hundred times brighter than the faintest bursts at $z=1$ and below.
-
-Except for the two intrinsically faintest bursts at $z\approx 1 - 1.5$, the
-statistical sample follows a constant lower limit in observed fluence as a
-function of redshift. This means that our selection is not biased at high
-redshifts except for the luminosity bias invoked by Swift. We find the
-following values for the redshift distribution. First for the statistical
-sample: $z_{\mathrm{avg}}=1.99$, $z_{\mathrm{med}}=1.64$, $z_{14\%}=0.60$ and
+Out of the 158 bursts in the statistical sample, 32($\sim$ 20 percent) have not
+been observed due to reasons unrelated to the GRB or afterglow properties. The
+reasons includes unavailability of the telescope due to technical maintenance
+(e.g., mirror re-coating), a visitor rejecting the TOO trigger, or bad weather.
+Because this cut is unrelated to the GRB properties, it will not change the
+statistical properties of the full sample. Removing all these bursts,
+drastically improves the redshift completeness from 60 per cent to 83 per cent.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{On the redshift distribution of GRBs} \label{redshift}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+One of the objectives of our follow-up campaign is to measure the redshift
+distribution for a well-defined, observationally unbiased and statistical useful
+sample of GRBs. The imposed selection criteria (see Sect. \ref{samplecrit})
+ensure that the GRBs entering our homogeneous sample fairly represent  the
+underlying population. The redshift distribution of such a sample holds valuable
+information about the occurrence of GRBs through cosmic time
+\citep{Jakobsson2012, Perley2016a}.
+
+In Fig.~\ref{fig:z} we show the redshift distribution of all  the observed GRBs.
+In the top panel we show a histogram for the full sample and the statistical
+sample and in the main panel, we show the redshifts of the individual bursts as
+a function of GRB luminosity in the observed 15-150 keV band. To calculate the
+luminosity, $L_{\mathrm{BAT}}$, we follow a similar procedure as
+\citet{Lien2016} and define $L_{\mathrm{BAT}} = F_{\gamma}\,4
+\pi\,d_L^2\,(1+z)^{-1}$, where $F_{\gamma}$ is the observed BAT fluence in the
+15-150 keV band and $d_L$ is the luminosity distance to the burst at the given
+redshift. Note that this measure of luminosity does not include any
+k-correction. As an indication of the effect of the \textit{Swift} sensitivity
+limit on the redshift distribution, also in the figure, we have shown the so
+called $\sim 1$ s flux BAT sensitivity limit ($\sim 3 \times
+10^{-8}~\mathrm{erg}~\mathrm{s}^{-1}~\mathrm{cm}^{-2}$;
+\citealt{Baumgartner2013, Lien2016}). Due to complex triggering mechanism of
+\swift, this sensitivity limit should be interpreted with some caution as the
+effective limit depends on the light curve of the prompt emission signal. Due to
+the dilution of light with distance, the \swift GRB luminosity detection limit
+is almost an order of magnitude brighter at $z=2$ than at $z=1$. At $z\geq4$
+we are only able to observe GRBs that are $\sim$ hundred times brighter than the
+faintest bursts at $z=1$ and below.
+
+We find the following mean and median and 14th and 86th percentiles for the
+redshift distribution. First for the statistical sample:
+$z_{\mathrm{avg}}=1.99$, $z_{\mathrm{med}}=1.64$, $z_{14\%}=0.60$ and
 $z_{86\%}=3.39$. For all GRBs observed with X-shooter we find
 $z_{\mathrm{avg}}=1.90$, $z_{\mathrm{med}}=1.52$, $z_{14\%}=0.55$ and
-$z_{86\%}=3.36$.
-
-Because the redshift completeness of our statistical sample is just 54 percent
-\todo{Get number}, making inference of the underlying redshift distribution of
-GRBs on this ample is difficult. Because a large fraction of the bursts for which we were
-unable to secure a redshift are missing due to terrestrial reasons, see Sect.
-\ref{badbursts}, imposing additional unbiasing cuts on the sample can increase
-the redshift completeness at the cost of sample size. As long as the cuts we
-impose are not selective in terms of the burst properties, the sample properties
-stay the same. The first cut we try is telescope availability such as bad
-weather, visitor rejected the TOO, or unavailability of the instrument. Because
-telescope availability is independent on the burst properties, by cutting away
-these from our sample, the underlying redshift distribution should be conserved.
-By trimming away the 30 bursts \todo{Get number} that can be rejected due to
-telescope availability, the redshift completeness increases to 63 percent
-\todo{Get number}.
-
-As is shown in Sect. \ref{timing}, the redshift completeness
-decreases with increasing follow-up delay, we can additionally increase the
-redshift completeness by imposing cuts on the follow-up delay. Our ability to
-rapidly follow up a burst is also not dependent on the burst properties and
-therefore we can additionally cut delays longer than 24 hours (all burst in the
-sample are in principle observable within 24 hours). \todo{Unfinished section}
-
-The biased introduced in the redshift distribution by the increased ability to
-secure redshift of optically brighter bursts has been investigated by
-\citet{Turpin2016} which find that we systematically find redshift for longer
-GRBs. Additionally only the brightest GRBs are seen above redshift $z \gtrsim
-1$. \todo{Unfinished section}
+$z_{86\%}=3.36$. For comparison, the corresponding values found in
+\citet{Fynbo2009} are z = 2.2 for both the mean and the median.
+
+%Because the redshift completeness of our statistical sample is 60 per cent,
+%making inference of the \textit{true} redshift distribution of GRBs on this sample
+%is difficult. A large fraction of the bursts for which we were unable to
+%secure a redshift are missing due to reasons unrelated to the GRBs (see Sect.
+%\ref{badbursts}). By trimming away the 33 bursts that can be rejected due to
+%telescope availability, the redshift completeness increases to 83 percent.
+%
+%As is shown in Sect. \ref{timing}, the redshift completeness
+%decreases with increasing follow-up delay, we can additionally increase the
+%redshift completeness by imposing cuts on the follow-up delay. Our ability to
+%rapidly follow up a burst is also not dependent on the burst properties and
+%therefore we can additionally cut delays longer than 24 hours (all burst in the
+%sample are in principle observable within 24 hours). \todo{Unfinished section}
+
+Because the redshift completeness of our statistical sample is 60 per cent,
+making inference of the \textit{true} redshift distribution of GRBs on this
+sample is difficult. The biases introduced in the redshift distribution by the
+increased ability to secure redshift of optically brighter bursts has been
+investigated by \citet{Turpin2016} which find that we systematically find
+redshift for longer GRBs. Additionally only the brightest GRBs are seen above
+redshift $z \gtrsim 1$ as shown in Fig. \ref{fig:z}. Additionally, as described
+in detail in \citet{Perley2016b}, bursts for which the redshift is measured from
+the afterglow are systematically found for host galaxies with a lower luminosity
+than burst for which the redshift is measured from the host galaxy. Only a few
+GRBs hosted in galaxies more massive than $10^{10}~\mathrm{M}_\odot$ have the
+redshift measured based on the afterglow continuum.
 
 \begin{figure*}
-	\centering
-	\includegraphics[width=16cm]{figures/XSGRB_zBATlum.pdf}
+	\centering \includegraphics[width=16cm]{figures/XSGRB_redshift_swiftdots.pdf}
 \caption{Redshift distribution as a function of intrinsic BAT $\gamma$-ray
-		luminosity, $L_{\gamma,\mathrm{int}}$. The latter quantity is defined as
-		$L_{\gamma,\mathrm{int}} = F_{\gamma,\mathrm{obs}}\,4 \pi\,d_L^2\,(1+z)^{-1}$,
-		where $F_{\gamma,\mathrm{obs}}$ is the observed BAT fluence in the 15-150 keV
-		band and $d_L$ is the luminosity distance at the given redshift. Blue star
-		symbols denote the statistical sample and black dots show all GRBs observed
-		with X-shooter and by \textit{Swift}/BAT. Overplotted as the red star symbols
-		for comparison are the GRB, quasar, and galaxy with the highest
-		spectroscopically confirmed redshifts, the latter two shown at arbitrary
-		$L_{\gamma,\mathrm{int}}$. The blue solid line represents the BAT sensitivity
-		of $10^{-8}$ erg cm$^{-2}$ s$^{-1}$ in the 15-150 keV band. The blue dashed
-		line marks an observed sensitivity of $10^{-7}$ erg cm$^{-2}$ s$^{-1}$, which
-		represents our sample limit out to $z\approx 8$. The estimated epoch of
-		reionization is shown by the black dot-dashed line, with the uncertainty shown
-		as the black dashed lines, from the most recent measurement by Planck+16.}
+	luminosity, $L_{\mathrm{BAT}}$. Bursts that are a part of the statistical
+	sample are marked by blue stars and black dots show all GRBs observed with
+	X-shooter. All \swift~GRBs with measured redshifts is shown in grey. For
+	comparison, we overplot with red stars the GRB \citep{Tanvir2009b,
+		Salvaterra2009a}, quasar \citep{Banados2017}, and galaxy \citep{Oesch2016} with
+	the highest spectroscopically confirmed redshifts, the latter two shown at
+	arbitrary $L_{\mathrm{BAT}}$. The blue solid line represents so-called 1 s BAT
+	sensitivity limit described in the text. The estimated epoch of reionization is
+	shown by the black dot-dashed line, with the uncertainty shown as the black
+	dashed lines, from the most recent measurement by \citet{Planck2015}.}
 \label{fig:z}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sample darkness} \label{darkness}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A large fraction of all GRBs exhibit no detectable or very faint optical
+afterglows \citep{Groot1998, Djorgovski2001, Fynbo2001}. The degree of optical
+extinction relative to the X-ray brightness has been parametrized in term of
+their optical darkness,  $\beta_{OX}$\citep{Jakobsson2004, Rol2005,
+	VanderHorst2009}. The X-ray properties of such bursts has previously been
+investigated \citep{DePasquale2003, Fynbo2009, Melandri2012}. There is some
+indications that dark bursts have somewhat higher X-ray luminosity and
+\nhx~compared to the optically brights bursts. This indicates along with
+investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
+	Hjorth2012, Perley2016b}, that the extinction of the optical afterglow is
+primarily driven by the presence of dust in the host galaxy.
+
+For all bursts with follow-up within 100 hours we calculate the
+"darkness"-parameter, $\beta_{OX}$ \citep{Jakobsson2004}. This requires the
+simultaneous measurement of the X-ray flux density and the optical flux density
+which is in practice never possible. As a proxy, we use the measured acquisition
+camera magnitude reported in Tab. \ref{tab:sample_overview} to get the optical
+flux density at the beginning of the spectroscopic integration. Because we know
+the delay between the follow-up and the \textit{Swift} trigger, we can use the
+measured XRT lightcurve \citep{Evans2007,
+	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
+to infer the corresponding X-ray flux density at the time of the optical
+observation. This is done by either linearly interpolating between temporally
+neighboring XRT measurements of by extrapolating the last few X-ray data points
+to the spectroscopic delay. When the afterglow is not detected in the
+acquisition camera, an upper limit of $> 24$ mags is used, which propagates into
+an upper limit on $\beta_{OX}$.
+
+In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
+the one presented in \citet{Fynbo2009}. The values from \citet{Fynbo2009} have
+been treated as detections, meaning that we artificially bias the distribution
+towards higher $\beta_{OX}$-values. We tentatively confirm the result by
+\citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher \nhx.
+Specifically, for bursts with measured redshift either from the afterglow or the
+host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$ we
+find $\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
+$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
+within the error intervals. A 2-sided KS test fails to reject the null that they
+are drawn from the same distribution with $p = 0.43$, meaning there is not
+strong evidence for a discrepancy.
+
+\begin{figure}
+	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}} \caption{Optical
+	extinction relative to X-ray against X-ray derived hydrogen column densities.
+	In red is shown the sample presented in \citet{Fynbo2009} where is lines
+	indicate the kernel density estimate of the distribution where limits have been
+	replaced with values. We note that contrary to what is presented in
+	\citet{Fynbo2009}, we exclude all bursts for which we do not have a redshift.
+	Darker colors represented higher density of points. The corresponding marginal
+	distributions are shown. In blue is the points for the bursts presented here
+	along with the marginal histograms. Limits on $\beta_{OX}$ is shown by
+	downwards facing arrows.} \label{fig:betaOX}
+\end{figure}
+
+It is clear, however, that the fraction of dark bursts for which we have 
+measured the redshift is lower than the intrinsic dark burst fraction. To measure the
+redshift as we either need a detected afterglow to obtain a spectrum or to 
+locate the host galaxy and determine the redshift from there. It is more difficuly to
+correctly associate a galaxy with a burst when there is no detected optical
+afterglow and hence a correct redshift measurement is more difficult to make,
+see \citet{Jakobsson2005, Levesque2010} and \citet{Perley2017} for a recent
+example.
+
+Using the table maintained by
+Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html}, we can see how the
+presence of an optical afterglow affects the follow-up statistics. 50.5 percent
+of all \textit{Swift}-detected bursts do not have an detected optical afterglow,
+but this number also includes bursts where no optical observations were
+available, so is a lower limit. For the bursts entering our statistical sample,
+the number of bursts without a detected optical afterglow is $\sim$ 27 percent,
+close to the upper limit on the fraction of dark bursts found in a complete
+sample with a very high degree of redshift completeness \citep{Melandri2012}. Of
+the bursts for which follow-up has actually been attempted, this number is
+$\sim$ 18 percent, confirming that we are indeed biased against bursts without a
+detected optical afterglow in the spectroscopic sample.
+
+Despite that the fraction of dark bursts is lower in the observed sample,
+compared to the statistical sample, the X-ray properties not differ
+significantly, as shown in Sect. \ref{completeness}. If the tendency for dark
+bursts to have a higher \nhx~is real, that the decreased fraction of dark bursts
+does not alter the observed prompt X-ray brightness distribution, potentially
+reflects the independence of the X-ray brightness on the density of the
+circumburst medium \citep{Freedman2001, Berger2003, Nysewander2009}, if the
+measured \nhx is primarily driven by a high column local to the burst. Because we
+only use values of \nhx~in the comparison for which the GRB has a measured
+redshift, this measure is likely biased toward optically brighter bursts.
+
+Since \textit{all} GRBs should have an optical afterglow, the presence of an
+optical afterglow is more a statement about the sensitivity than optical
+properties of the bursts.
+
+This is despite spectroscopic follow-up only being carried out in cases where
+either a detectable optical afterglow or a clear counterpart are seen, which
+naively should be biased against dark bursts occurring in more obscured
+galaxies, which is shown to exhibit different galactic properties
+\citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b, Perley2015b}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Hydrogen column densities}
@@
 
-Because the locations of GRBs are associated with the most intensely starforming
-regions \citep{Hogg1999, Bloom2002, Fruchter2006}, the GRB afterglow light has
-to propagate through the large amounts of hydrogen fueling the star formation.
+Because the locations of long GRBs are associated with intensely star forming regions
+\citep{Hogg1999, Bloom2002, Fruchter2006}, the GRB afterglow light has to
+propagate through the large amounts of hydrogen fueling the star formation.
 Because a significant fraction of the hydrogen has not been ionized yet, the
-optical depth at the wavelength of \lya~is very high, saturating at the line
-center. This causes a strong absorption system to appear in the afterglow
-continuum. For bursts with $z \gtrsim 1.7$, the position of \lya~moves into the
-spectroscopic coverage of X-shooter, meaning that we are able to detect this
-absorption trough due to \lya.
-
-Due of the stochastic nature of the \lya-forest, the blue wing of the Lyman
-alpha absorption line is randomly superposed with forest systems, along with
-strong absorption from \mnii~and \SIiii, making it notoriously difficult to
-model. Additionally, Similarly so, the red wing has got the ISM signatures
-imprinted on it, especially strong absorption due to \SIii, \sii~and \nv~which
-can exhibit significant velocity structure. Along with instrumental effects, the
-generative model for the data that we would use in a likelihood-based analysis
-would be almost boundless in complexity, making formal fitting of the column
-densities outside the scope of this paper.
-
-Using an analytic approximation to an absorption line profile and it's
-dependence on the column density \citep{TepperGarcia2006}, we overplot a
-synthetic absorption line with a specified column density on our observed
-spectrum. By tuning the value of the hydrogen column density until the synthetic
-absorption line matches the spectrum, we can thereby infer the actual column
-density of the GRB sight line in a manual way. In a similar fashion, the errors
-on the hydrogen column can be inferred. We show the results of this procedure
-for all bursts where possible in Fig. \ref{fig:HI1} and the inferred hydrogen
-column densities in Tab. \ref{tab:HI}
+optical depth at the wavelength of \lya~is very high, saturating not only the
+line center, but also the damping wings. This causes a strong absorption system
+to appear in the afterglow continuum. For bursts with $z \gtrsim 1.7$, the
+position of \lya~moves into the spectroscopic coverage of X-shooter, meaning
+that we are able to detect this absorption trough due to \lya.
+
+Due to the stochastic nature of the \lya-forest, the blue wing of the
+Lyman-$\alpha$ absorption line is randomly superposed with Lyman-$\alpha$
+forest systems, along with strong absorption from \mnii~and \SIiii, making it
+notoriously difficult to model. Additionally, the red wing has got the ISM
+signatures imprinted on it, especially strong absorption due to \SIii, \sii~and
+\nv, which can exhibit significant velocity structure. Along with instrumental
+effects, the generative model for the data that we would use in a
+likelihood-based analysis would be very complicated, we have decided not to
+make formal $\chi$-2 fitting of the hydrogen column densities, but instead use
+a more subjective visual fitting to the absorption profile.  Using an analytic
+approximation to the absorption profile from \citep{TepperGarcia2006}, we
+overplot a synthetic absorption line with a specified column density on our
+observed, normalised spectrum. By tuning the value of the hydrogen column
+density until the synthetic absorption line matches the spectrum, we can
+thereby infer the actual column density of the GRB sight line in a manual way.
+In a similar fashion, the errors on the hydrogen column can be estimated. We
+show the results of this procedure for all bursts where possible in Fig.
+\ref{fig:HI1} and the inferred hydrogen column densities in Table~\ref{tab:HI}
 
@@
 
-In a similar manner to how the X-ray derived column densities was used to assess
-the sample completeness in Sect. \ref{completeness}, we addresses the same
-question in terms of the optically derived hydrogen column densities. Using the
-compilation by \citet{Tanvir2017} we have all 81 published HI values previously,
-we can compare to the 39 new HI columns added by this sample. We note that this
-sample increase the number of optically derived hydrogen column densities with
-$\sim$ 50 percent. We compare the median, the 14th, and 86th percentiles of the
-two distributions where the sample presented here has $21.8_{0.8}^{0.3}$ and the
-rest of the litterature values has $21.5_{1.5}^{0.4}$. We see that the two
-distribution has a large degree of overlap due to the large width of the
-distributions, but find a slightly higher median value for the new sample
-presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
-relatively large evidence against the null that the two samples are drawn from
-the same underlying distribution. Because the bursts that have measurements of
-the hydrogen column density are selected sole based on our ability to infer a
-column, it is difficult to make any strong conclusions about the completeness in
-terms of gas content. We show the comparison between the two samples in Fig.
-\ref{fig:NH_dist}, where we additionally show the column-distribution for 12081
-quasar DLAs published in \citet{Noterdaeme2012b}. The ability of the GRBs to
-select the highest N$_H$ is immediately visible as is also noted previously
-\citep{Prochaska2007, Fynbo2009}. The reason for this is the selection of sight
-lines as random position of the QSO DLAs whereas the GRB forms in dense regions
-in the host. The N$_H$ distribution has also be used the infer the position of
-the gas in host \cite{Buchner2016} where a galaxy-wide gas reservoir is
-preferred.
+In the compilation of HI measurements towards GRBs in \citet{Tanvir2017} there
+are 81 published HI values. We here provide 39 new HI column density
+measurements, i.e. an increase of the number of optically derived hydrogen
+column densities of $\sim$ 50 percent. We compare the median, the 14th, and 86th
+percentiles of the two distributions where the sample presented here has
+$21.8_{0.8}^{0.3}$ and the rest of the litterature values has
+$21.5_{1.5}^{0.4}$. We see that the two distribution has a large degree of
+overlap due to the large width of the distributions, but we find a slightly
+higher median value for the new sample presented here. A 2-sided KS test gets us
+a p-value of p = 0.006, meaning relatively large evidence against the null that
+the two samples are drawn from the same underlying distribution. Because the
+bursts that have measurements of the hydrogen column density are selected solely
+based on our ability to infer a column, it is difficult to make any strong
+conclusions about the population statistics in terms of gas content.
+
+We show the
+comparison between the two samples in Fig. \ref{fig:NH_dist}. We additionally
+show the column-distribution for 12081 quasar DLAs from \citet{Noterdaeme2012b}.
+The fact that GRBs systematically are located behind the highest \nh,
+previously noted by \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is stricking. The
+reason for this is that quasars sampling of sightlines through galaxies is
+cross-section selected, whereas GRB sightlines probe more central, dense regions
+in their hosts.
 
@@
 In this paper we have presented the results of a dedicated effort over the
-years 2009 -- 2016 to use the X-shooter spectrograph on the ESO-VLT to secure
+years 2009 -- 2017 to use the X-shooter spectrograph on the ESO-VLT to secure
 spectroscopic observations of afterglows and host galaxies of GRBs detected by
 {\it Swift}. This work was initiated by a consortium using Guaranteed Time on
-X-shooter, but over the years the project continuued in open time and the team
+X-shooter, but over the years the project continued in open time and the team
 was opened to include any researchers interested in contribution to this
@@
 in this paper.
+
+SWIFT LIMITATIONS
+There is some evidence that the burst duration, $T_{90}$, does not scale with
+redshift as expected from time-dilation \citep{Kocevski2013, Littlejohns2013a},
+suggesting that we could be missing an increasing fraction of the bursts at
+higher redshifts, but \citet{Lien2016} argues that we should only be missing the
+bright "short" GRBs", but to the degree that \textit{Swift} is complete, we can
+test if the sample presented here is complete.
 
@@
 systems in the majority of cases \citep[][and this work]{Jakobsson2006b,
-	Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
-spectrum from X-rays to the sub-mm) we hence characterize the properties of
-star-forming galaxies over cosmic history in terms of redshifts, metallicities,
-molecular contents, ISM temperatures, UV-flux densities, extinction curves,
-etc.  A number of independent papers have been published or submitted for
-publication focusing on many of these specific issues of our sample such as
-extinction curves [][see also
-\cite{Fynbo2014, Heintz2017}]\citep{Japelj2015, Zafar2017}, emission lines from the
-underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
-\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
-radiation \citep{Tanvir2017}.
+	Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy (throughout the
+electromagnetic spectrum from X-rays to the sub-mm) we hence characterize the
+properties of star-forming galaxies over cosmic history in terms of redshifts,
+metallicities, molecular contents, ISM temperatures, UV-flux densities,
+extinction curves, etc.  A number of independent papers have been published or
+submitted for publication focusing on many of these specific issues of our
+sample such as extinction curves \citep[][see also \citealt{Fynbo2014,
+	Heintz2017}]{Japelj2015, Zafar17}, emission lines from the underlying host
+galaxies \citep{Kruhler2015}, the frequency of intervening \ion{Mg}{ii}
+absorbers \citep{Christensen2017}, and escape of ionizing radiation
+\citep{Tanvir2017}.
 
@@
 
-\section{Notes on Individual objects}
+\section{Notes on Individual objects} \label{notes}
 
@@
 \subsection{GRB120119A (z = 1.728)}
-The data presented here has been examined by Japelj+15 and Zafar+17 who both 
-find a significant amount of extinction, $A_V\approx 1$ mag. Three epochs of
-observations have been obtained, the first two immediately after the burst, and
-the last one long after the afterglow had faded. Starting 1.4 hours after the
-\textit{Swift} trigger, the first epoch contains bright afterglow continuum.
-Rich in absorption features belonging to a multitude of ions at $z =    1.728$ is
-estimated for the host with intervening systems at $z = 1.476$, $z = 1.214$, $z
-= 0.662$ and $z = 0.632$. The second epoch, obtained 4.5 hours after the burst
-contains the fading afterglow. A third epoch is obtained $>1$ year after the
-GRB in which emission lines from \hb~and \ha~are found at the redshift of the
-host, confirming the association of the absorption line system and the host. We
-also detect C\,\textsc{i} in absorption which indicates the presence of cold gas.
+
+The data presented here has been examined by \citet{Japelj2015} and
+\citet{Zafar2017} who both find a significant amount of extinction, $A_V\approx
+1$ mag. Three epochs of observations have been obtained, the first two
+immediately after the burst, and the last one long after the afterglow had
+faded. Starting 1.4 hours after the \textit{Swift} trigger, the first epoch
+contains bright afterglow continuum. Rich in absorption features belonging to a
+multitude of ions at $z =    1.728$ is estimated for the host with intervening
+systems at $z = 1.476$, $z = 1.214$, $z = 0.662$ and $z = 0.632$. The second
+epoch, obtained 4.5 hours after the burst contains the fading afterglow. A third
+epoch is obtained $>1$ year after the GRB in which emission lines from \hb~and
+\ha~are found at the redshift of the host, confirming the association of the
+absorption line system and the host. We also detect C\,\textsc{i} in absorption
+which indicates the presence of cold gas.
 
@@
 16214\footnote{\url{http://gcn.gsfc.nasa.gov/other/140506A.gcn3}} and is
-published in Fynbo et al. (2014) and Heintz et al. (2017a). The observations
+published in \citet{Fynbo2014} and \citet{Heintz2017a}. The observations
 consists of $4\times 600$ sec at 8.8 and 33 hours after the burst. We detect
@@
 19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, and is
-published in Heintz et al. (2017b). Observations started 22.37 hours after the
+published in \citet{Heintz2017b}. Observations started 22.37 hours after the
 BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the entire

### 2017-12-07 py/skymap.py

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import healpy as hp
 import seaborn as sns; sns.set_style('ticks')
@@ -58,7 +59,7 @@ def main():
     burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
     name_o = np.array([ii[3:] for ii in burst_table["GRB"].values])
     # Read in Swift table
-    swift_table = pd.read_table("../data/grb_table_1482495106.txt", delimiter="\t", dtype=None)
+    swift_table = pd.read_table("../data/grb_table_1511519199.txt", delimiter="\t", dtype=None)
     name_s, RA, DEC, redshift = swift_table["GRB"].values, pd.to_numeric(swift_table["BAT RA (J2000)"], errors='coerce').values, pd.to_numeric(swift_table["BAT Dec (J2000)"], errors='coerce').values, swift_table["Redshift"].values
 
     # Convert ra and dec to l, b
@@ -85,8 +86,10 @@ def main():
     l_z = l[mask_z]
     b_z = b[mask_z]
 
+
+    print(l_z, b_z)
     # Scatterplot with bursts position
-    hp.visufunc.projscatter(l_z, b_z, facecolors='none', marker = "*", s = 50, lonlat=True)
+    hp.visufunc.projscatter(l_z, b_z, facecolors='none', edgecolor="black", marker = "*", s = 50, lonlat=True)
 
     # Get indices of sample bursts and observed bursts
     idx = [ii for ii, kk in enumerate(name_s_z) if kk in name]

### 2018-01-20 document/XSGRB_sample.tex
Adding individual bursts

@@
 
-\title{XSGRB - The X-shooter GRB afterglow sample
+\title{XSGRB I - The X-shooter GRB afterglow sample
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	a measurement of the hydrogen column density, increasing the total number of
-	published HI column density measurements by $\sim$ 50 \%. We constrain the
-	fraction of dark bursts to be 27\% - 42\% \todo{Preliminary} and tentatively
-	confirm previous results that larger optical darkness is correlated with
-	increased X-ray absorption. This dataset provides a unique possibility to study
-	the ISM across cosmic time.
-
-}
+	published HI column density measurements by $\sim$ 50 per cent. We constrain the
+	fraction of dark bursts to be less than 28 per cent based on the detection of an
+	optical afterglow and tentatively confirm previous results that larger optical
+	darkness is correlated with increased X-ray absorption. This dataset provides a
+	unique possibility to study the ISM across cosmic time. }
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{RRM observations} \label{RRM}
+\subsection{RRM observations} \label{RRM}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-For 8 of the observed GRBs in the spectra released here, RRM triggered. One of
-the RRM triggers is outside the statistical sample and two of the RRM trigger
-are on short GRBs. The fastest response between \swift~GRB trigger and the
-beginning of spectroscopic integration is for GRB~160410A, for which the delay
-was only 8.4 minutes.
+For 8 of the observed GRBs released here, RRM triggered. One of the RRM triggers
+is outside the statistical sample and two of the RRM trigger are on short GRBs.
+The fastest response between \swift~GRB trigger and the beginning of
+spectroscopic integration is for GRB~160410A, for which the delay was only 8.4
+minutes.
 
@@
 \begin{figure*}
-	\centerline{\includegraphics{figures/skymap.pdf}}
-	\caption{Mollweide projection in galactic coordinates of the full sky showing
-		the positions on the sky of the bursts presented in this work. The empty
+	\centerline{\includegraphics{figures/skymap.pdf}} \caption{Mollweide projection
+	in galactic coordinates of the full sky showing the positions on the sky of the
+	bursts presented in this work. The equator is the Galactic plane. The empty
 	circles are the positions of all the \textit{Swift} bursts detected until the
@@
 	\citet{Schlegel1998} where the sample criteria cut with $A_V \lesssim 0.5$ mag
-		is removed.}
+	is removed. The colorbar below indicates the value of $A_V$ on the plot. The
+	dotted lines indicate intervals of 30$^\circ$ in longitude and latitude.}
 \label{fig:skymap}
@@
 
-The first GRB followed up is GRB090313, observed on the 15th of March, 2009,
+The first GRB followed up is GRB~090313, observed on the 15th of March, 2009,
 during the commissioning of X-shooter UT3. The bursts observed during the
@@
 For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a 0\farc9 slit only 76.1 percent of the light will enter the slit, meaning that for almost all observations a slitloss correction is required.
-We describe how slitlosses were corrected for in Sect. \ref{postproc}. For
-accurate measurements involving line profiles, knowledge of the precise
+%We describe how slitlosses were corrected for in Sect. \ref{postproc}.
+For accurate measurements involving line profiles, knowledge of the precise
 instrumental resolution is required. The resolution becomes better than the
@@
 	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
-	\caption{Telluric corrected, normalized spectrum of GRB121024A at $z = 2.300$,
-	meant to illustrate the typical data quality. The acquisition magnitude is r =
-	20, meaning it is in the brighter end of the sample presented here, but not the
-	brightest. The spectrum is rich in absorption, including molecular absorption
-	of $H_2$. The absorption trough visible at $\sim$ 4000 $\AA$ is due to \lya in
-	the host. We have overplot the most prominent lines seen in GRB afterglows from
-	\citet{Christensen2011a}. The continuum estimate is shown in dashed red.
-	Additionally, three intervening systems are seen in this sightline  This
-	spectrum has been analyzed in detail in \citet{Friis2015}.}
-	\label{fig:spectrum}
+\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$,
+	that illustrates the typical data quality. The continuum estimate is shown in
+	dashed red and the error spectrum in solid blue. The acquisition magnitude is
+	$r$ = 20, meaning it is in the brighter end of the sample presented here, but
+	not the brightest. The spectrum is rich in absorption, including absorption
+	from molecular $\mathrm{H_2}$. The absorption trough visible at $\sim$ 4000
+	$\AA$ is due to \lya~in the host. We have overplot the most prominent lines
+	seen in GRB afterglows from \citet{Christensen2011a}. Additionally, three
+	intervening systems are seen in this sightline This spectrum has been analyzed
+	in detail in \citet{Friis2015}.} \label{fig:spectrum}
 \end{figure*}
@@
 telluric standard stars spanning a range of DIMM seeing values, with the
-majority between $0.5-1.5 arcsec$. The resulting distribution of spectral FWHM
+majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
 (km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
@@
 fastest follow-up between BAT trigger and start of spectroscopic observations in
-our sample is for the short, $z = 1.717$, GRB160410A for which spectroscopic
+our sample is for the short, $z = 1.717$, GRB~160410A for which spectroscopic
 integration was initialed only 8.4 minutes after the BAT trigger. To illustrate
@@
 \begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/NH_dist.pdf}
+	\centering \includegraphics[width=9cm]{figures/NH_dist.pdf}
 \caption{Distributions of DLA hydrogen column densities for DLAs found in
-		quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in green is the kernel density estimate
-		is the similar distribution, only for DLAs in GRB sightlines. Values are taken
-		from the compilation in Tanvir2017 et al. (in prep)\citet{Tanvir2017}, along
-		with the new values presented in this sample. The marginal distributions for
-		the two distribution are also shown, where the different environments probed
-		are clearly visible in the hydrogen column densities, as previously also noted
-		in \citet{Fynbo2009}}
+	quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in
+	green is the kernel density estimate is the similar distribution, only for DLAs
+	in GRB sightlines. Values are taken from the compilation in Tanvir2017 et al.
+	(in prep), along with the new values presented in this sample. The marginal
+	distributions for the two distribution are also shown along the left side and
+	on the top, where the different environments probed are clearly visible in the
+	hydrogen column densities, as previously also noted in \citet{Fynbo2009}}
 \label{fig:NH_dist}
@@
 
-In this paper we have presented the results of a dedicated effort over the
-years 2009 -- 2017 to use the X-shooter spectrograph on the ESO-VLT to secure
+In this paper we have presented the results of a dedicated effort over the years
+2009 -- 2017 to use the X-shooter spectrograph on the ESO-VLT to secure
 spectroscopic observations of afterglows and host galaxies of GRBs detected by
-{\it Swift}. This work was initiated by a consortium using Guaranteed Time on
+\swift. This work was initiated by a consortium using Guaranteed Time on
 X-shooter, but over the years the project continued in open time and the team
-was opened to include any researchers interested in contribution to this
-effort.
-
-The sample we present here include spectroscopic observations of 86 
-systems fulfilling our sample criteria of which 72 are afterglow spectra 
-and 14 are observations of the underlying host galaxies. 
-UPDATE THIS.
-All spectra have been made publicly available in the reduced form used
-in this paper.
-
-SWIFT LIMITATIONS
-There is some evidence that the burst duration, $T_{90}$, does not scale with
-redshift as expected from time-dilation \citep{Kocevski2013, Littlejohns2013a},
-suggesting that we could be missing an increasing fraction of the bursts at
-higher redshifts, but \citet{Lien2016} argues that we should only be missing the
-bright "short" GRBs", but to the degree that \textit{Swift} is complete, we can
-test if the sample presented here is complete.
+was opened to include any researchers interested in contribution to this effort.
+
+The sample we present here include spectroscopic observations of 90 systems
+fulfilling our sample criteria including 17 spectra that are late time
+observations of the underlying host galaxies. All spectra have been made
+publicly available in the reduced form used in this paper.
 
@@
 	Heintz2017}]{Japelj2015, Zafar17}, emission lines from the underlying host
-galaxies \citep{Kruhler2015}, the frequency of intervening \ion{Mg}{ii}
-absorbers \citep{Christensen2017}, and escape of ionizing radiation
-\citep{Tanvir2017}.
-
-It seems important to end this paper by stressing that the potential of 
-using GRB sightlines as probes is far from fully harvested. The sample of
-sightlines probed by our spectra is not representative for all GRB 
-sightlines as we have shown and consistent with earlier findings
-from samples based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009}
-and from studies of complete samples of GRB host galaxies 
-\citep{Hjorth2012, Covino2013, Perley2016}. 
-\cite{Kruhler2013} convincingly argue, that 
-very rich sightlines like that probed by the remarkable GRB\,080607 
-\citep{Prochaska2009} are probably significantly more frequent than in the sightlines 
-sampled by our spectra. However, with current instrumentation these sightlines are
-out of reach except under very fortunate circumstances as in the case of 
-GRB\,080607 when the afterglow could be observed only a few minutes after the 
-burst with a 10-m class telescope. Observations of such sightlines with 
-X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
-is likely to be very rewarding.
+galaxies \citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
+\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al. in
+prep). %\citep{Tanvir2017}.
+
+%SWIFT LIMITATIONS
+%There is some evidence that the burst duration, $T_{90}$, does not scale with
+%redshift as expected from time-dilation \citep{Kocevski2013, Littlejohns2013a},
+%suggesting that we could be missing an increasing fraction of the bursts at
+%higher redshifts, but \citet{Lien2016} argues that we should only be missing the
+%bright "short" GRBs", but to the degree that \textit{Swift} is complete, we can
+%test if the sample presented here is complete.
+
+It seems important to end this paper by stressing that the potential of using
+GRB sightlines as probes is far from fully harvested. The sample of sightlines
+probed by our spectra is not representative for all GRB sightlines as we have
+shown and consistent with earlier findings from samples based on low-resolution
+spectroscopy \citep[e.g.,][]{Fynbo2009} and from studies of complete samples of
+GRB host galaxies \citep{Hjorth2012, Covino2013, Perley2016}. \cite{Kruhler2013}
+convincingly argue, that very rich sightlines like that probed by the remarkable
+GRB~080607 \citep{Prochaska2009} are probably significantly more frequent than
+in the sightlines sampled by our spectra. However, with current instrumentation
+these sightlines are out of reach except under very fortunate circumstances as
+in the case of GRB~080607 when the afterglow could be observed only a few
+minutes after the burst with a 10-m class telescope. Observations of such
+sightlines with X-shooter-like spectrographs on th enext generation of 20--40-m
+telescopes is likely to be very rewarding.
 
@@
 
-\subsection{GRB090313 (z = 3.373)}
+\subsection{GRB~090313 (z = 3.373)}
 The first GRB ever observed with X-shooter, during the commissioning of the
@@
 \#9015\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9015.gcn3}} and is published
-in \citet{DeUgartePostigo2010}. Due to the lingering brightness of GRB090313,
+in \citet{DeUgartePostigo2010}. Due to the lingering brightness of GRB~090313,
 6.9 ks spectroscopic integration starting 45 hours after the BAT trigger reveals
@@
 
-\subsection{GRB090530 (z = 1.266)}
+\subsection{GRB~090530 (z = 1.266)}
 Observed during paranalization of the instrument, this data forms the basis of
@@
 
-\subsection{GRB090809 (z = 2.737)}
+\subsection{GRB~090809 (z = 2.737)}
 Observed during the first science verification period and was used for GCN
@@
 
-\subsection{GRB090926 (z = 2.106)}
+\subsection{GRB~090926 (z = 2.106)}
 Starting during the second science verification period, this dataset forms the
@@
 
-\subsection{GRB091018 (z = 0.971)}
+\subsection{GRB~091018 (z = 0.971)}
 The first burst observed during normal operation after science verification was
@@
 
-\subsection{GRB091127 (z = 0.490)}
+\subsection{GRB~091127 (z = 0.490)}
 Obtained 4 days after the burst trigger, this data forms the basis for GCN \#
@@
 
-\subsection{GRB100205A  (z = na)}
+\subsection{GRB~100205A  (z = na)}
 Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected
@@
 
-\subsection{GRB100219A (z = 4.667)}
+\subsection{GRB~100219A (z = 4.667)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100316B (z = 1.180)}
+\subsection{GRB~100316B (z = 1.180)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100316D (z = 0.059)}
+\subsection{GRB~100316D (z = 0.059)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100418A (z=0.624)}
+\subsection{GRB~100418A (z=0.624)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100424A (z=2.465)}
+\subsection{GRB~100424A (z=2.465)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100425A (z=1.1755)}
+\subsection{GRB~100425A (z=1.1755)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100615A (z=1.398)}
+\subsection{GRB~100615A (z=1.398)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100621A (z=0.542)}
+\subsection{GRB~100621A (z=0.542)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB100625A (z=0.452)}
+\subsection{GRB~100625A (z=0.452)}
 The data presented here is of the candidate host galaxy, taken long after the
@@
 
-\subsection{GRB100724A* (z = 1.288)}
+\subsection{GRB~100724A* (z = 1.288)}
+
 The data presented here also formed the basis of GCN \#
-10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. The spectrum
-has not otherwise been published previously. The observations were carried out
-in RRM starting 11 min after the GRB trigger. See section \ref{RRM}, for a
+10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. The spectrum has
+not otherwise been published previously. The observations were carried out in
+RRM starting 11 min after the GRB trigger. See section \ref{RRM}, for a
 description of the RRM scheme. Absorption lines from several ionic species are
-detected in the afterglow continuum at a common redshift of $z = 1.288$. This
-is not a part of the statistical sample.
-
-\subsection{GRB100728B (z=2.106)}
+detected in the afterglow continuum at a common redshift of $z = 1.288$. This is
+not a part of the statistical sample.
+
+\subsection{GRB~100728B (z=2.106)}
+
 The data presented here also formed the basis of GCN \#
-11317\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11317.gcn3}}. The spectrum
-has not otherwise been published previously. Starting 22 hours after the burst
+11317\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11317.gcn3}}. The spectrum has
+not otherwise been published previously. Starting 22 hours after the burst
 trigger, 7.2 ks of observations reveals a faint afterglow continuum with \lya-
@@
 of X-shooter is depressed with respect to normal operations, resulting in a
-poorer throughout. Additionally, the position of the trace on the slit moves
-due to atmospheric differential refraction.
-
-\subsection{GRB100814A (z=1.439)}
-The spectra presented here has not been published previously. The
-observations consists of three visits, the first beginning only 0.9 hours
-after the \textit{Swift} trigger, the other two visits were 2.13 and 98.40
-hours after the trigger, respectively. A bright afterglow continuum is
-present in all visits, allowing identification of absorption features
-belonging to a wide range of ions at $z=1.439$. A complex velocity structure
-in the absorption features belonging to \mgii, shows several components,
-separated by as much as 500km/s, pointing to a likely merger scenario in
-the host.
-
-\subsection{GRB100816A (z=0.805)}
+poorer throughout. Additionally, the position of the trace on the slit moves due
+to atmospheric differential refraction.
+
+\subsection{GRB~100814A (z=1.439)}
+
+The spectra presented here has not been published previously. The observations
+consists of three visits, the first beginning only 0.9 hours after the
+\textit{Swift} trigger, the other two visits were 2.13 and 98.40 hours after the
+trigger, respectively. A bright afterglow continuum is present in all visits,
+allowing identification of absorption features belonging to a wide range of ions
+at $z=1.439$. A complex velocity structure in the absorption features belonging
+to \mgii, shows several components, separated by as much as 500km/s, pointing to
+a likely merger scenario in the host.
+
+\subsection{GRB~100816A (z=0.805)}
+
 The data presented here also formed the basis of GCN \#
-11123\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11123.gcn3}}. The spectrum
-has not otherwise been published previously. This short GRB was observed 28.4
-hours after the GRB trigger. 4 x 1200 s of exposure reveals two distinct sets
-of emission lines, spatially offset $\lesssim 1 \arcsec $, very close in
-redshift space, $z=0.8034$ and $z=0.8049$, indicating either an interacting
-host or some complex velocity structure of the host. Faint underlying continua
-are present under both sets of lines.
-
-\subsection{GRB100901A (z=1.408)}
+11123\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11123.gcn3}}. The spectrum has
+not otherwise been published previously. This short GRB was observed 28.4 hours
+after the GRB trigger. 4 x 1200 s of exposure reveals two distinct sets of
+emission lines, spatially offset $\lesssim 1 \arcsec $, very close in redshift
+space, $z=0.8034$ and $z=0.8049$, indicating either an interacting host or some
+complex velocity structure of the host. Faint underlying continua are present
+under both sets of lines.
+
+\subsection{GRB~100901A (z=1.408)}
+
 The data presented here has been published in \citet{Hartoog2013}. Because of
@@
 
-\subsection{GRB101219A (z=0.718)}
+\subsection{GRB~101219A (z=0.718)}
+
 This data has not been published before. Starting 3.7 hours after the GRB
@@
 
-\subsection{GRB101219B (z=0.552)}
+\subsection{GRB~101219B (z=0.552)}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB110128A (z=2.339)}
-
-\subsection{GRB110407A (z=na)}
-
-\subsection{GRB110709B (z=2.1092 (NEW!!!!))}
+\subsection{GRB~110128A (z=2.339)}
+
+These observations forms the basis of GCN \#
+11607\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11607.gcn3}}, but has not been
+published before. Spectroscopic integration started 6.55 hours after the
+\swift~trigger and lasted for a total of 7.2ks The afterglow continuum is
+detected across the entire spectral coverage at moderate S/N. Absorption lines
+in the continuum is detected from \lya, \oi, \cii, \SIiv, \civ, \SIii~and \feii,
+all at a common redshift of $z=2.339$. From the broad \lya~trough, a hydrogen
+column density $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 22.6 \pm 0.2$ is
+derived. An intervening system at $z=2.20$ is tentatively identified from an
+absorption feature, likely due to \civ.
+
+\subsection{GRB~110407A (z=na)}
+
+These observations have not been published before. Starting 12.36 hours after
+the BAT trigger, 4.8 ks spectroscopic integration yield a very faint trace down
+to $\sim$4300\AA, only visible after binning heavily. This could indicate a
+redshift, $z ~\sim 2.5$, but no emission lines or absorption lines are
+immediately visible to support this.
+
+\subsection{GRB~110709B (z=2.109 (NEW!!!!))}
+
 This is a late-time observation (> 1 year) and has previously been used in
@@
 along with low-significance detection of \ha~at the end of the spectral
-coverage, both at a consistent redshift, $z=2.1092$, securing it as the redshift
+coverage, both at a consistent redshift, $z=2.109$, securing it as the redshift
 of the GRB.
 
-\subsection{GRB110715A (z=0.82)}
-
-\subsection{GRB110721A (z=0.382)}
-
-\subsection{GRB110808A (z=1.348)}
-
-\subsection{GRB110818A (z=3.36)}
-
-\subsection{GRB111005A (z=0.013)}
-
-\subsection{GRB111008A  (z=4.989)}
-
-\subsection{GRB111107A (z=2.893)}
-
-\subsection{GRB111117A (z=2.211)}
-
-\subsection{GRB111123A  (z=3.151)}
-
-\subsection{GRB111129A (z=na)}
-
-\subsection{GRB111209A (z=0.677)}
-
-\subsection{GRB111211A (z=0.478)}
-
-\subsection{GRB111228A (z=0.716)}
-
-\subsection{GRB120118B (z = 2.943)}
+\subsection{GRB~110715A (z=0.823)}
+
+These observation, starting 12.3 hours after the trigger, have been published in
+\citet{Sanchez-Ramirez2017} and additionally formed the basis of GCN \#
+12164\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12164.gcn3}}. Only a single
+exposure of 600 s was obtained, before strong winds interrupted the
+observations. A red continuum is detected across all arms and a multitude of
+absorption lines are superposed on the afterglow continuum. We identify lines
+belonging to \alii, \aliii, \znii, \crii, \feii, \mgii, \mgi, \caii, and \caii,
+all at  $z=0.823$, marking it as the redshift of the GRB.
+
+\subsection{GRB~110721A (z=0.382)}
+
+This is a Fermi burst with a LAT detection and thus outside the statistical
+sample, but nonetheless followed up due to the extremely high peak energy
+\citep{Axelsson2012}. Starting 28.7 after the burst trigger, 2.4ks spectroscopic
+observation reveals after heavy binning, a wide, faint trace down to $\sim$ 5800
+\AA, offset by 2.5\arcsec~relative to the centering of the slit. No good
+redshift measurement can be inferred from this. We have adopted the redshift
+from GCN \# 12193\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12193.gcn3}}
+
+\subsection{GRB~110808A (z=1.348)}
+
+This spectrum has already formed the basis of GCN \#
+12258\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12258.gcn3}}, but is not
+published otherwise. Starting 3 hours after the \swift~trigger, a rich spectrum
+is obtained in 2.4ks spectroscopic integration. The GRB afterglow continuum is
+visible across all three spectrscopic arms of VLT/X-shooter with additionally,
+emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$. At the same
+redshift, we identify absorption lines superposed on the afterglow continuum
+from \mgii~and \feii.
+
+\subsection{GRB~110818A (z=3.36)}
+
+Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8ks
+reveals a moderate S/N GRB afterglow continuum, down to $\sim$ 5000 \AA. The
+simultaneous detection of absorption features identified as \lya, \SIii, \civ,
+\alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet, securely sets
+$z = 3.36$ as the redshift of the GRB. This data forms the basis of GCN \#
+12284\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12284.gcn3}}, but is not
+published elsewhere.
+
+\subsection{GRB~111005A (z=0.013)}
+
+The data presented here has previously been published in \citet{Michaowski2016}.
+2.4ks spectroscopic integration of the host galaxy, obtained long after the
+burst had faded, contains bright emission lines filling the entire slit on top
+of a broad, underlying stellar continuum. We identify emission lines from \oii,
+\hd, \hg, \hb, \oiii, \nii, \hb, \sii, \ariii, and \sii, all at $z=0.013$.
+Significant velocity structure of the lines across the spatial direction of the
+slit indicates a large degree of coherent motion relative to the line-of-sight.
+
+\subsection{GRB~111008A  (z=4.989)}
+
+This data formed the basis of GCN \#
+12431\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12431.gcn3}} and is
+additionally published in \citet{Sparre2014}. Observations of this GRB afterglow
+was initiated 7.71 hours after the BAT trigger and had a duration of 8.4ks. A
+second observational epoch started 20.1 hours and lasted for 6.6ks. The GRB
+afterglow continuum is well detected down to $\sim$7600 \AA, with several strong
+absorption features imprinted. All at a common $z = 4.990$ \lya~is clearly
+detected and we additionally detect lines identified as \SIii, \feii, \civ,
+\mgii, \SIii*, \sii*, \oi*. An intervening DLA systems is additionally detected
+at  $z = 4.61$ as seen from \lya~and \mgii~absorption.
+
+\subsection{GRB~111107A (z=2.893)}
+
+GCN \# 12542\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12542.gcn3}} is based
+on this spectrum, but it is not published elsewhere. Spectroscopic integration
+started 5.26 hours after the \swift~trigger and consists of $4 \times 1200$ s
+integration in the UVB and VIS and $16 \times 300$ s in NIR, the observations
+ending in twilight. The GRB afterglow continuum is well detected across the arms
+with absorption lines from \lya, \civ, \feii, and \mgii, all at a consistent
+redshift of $z = 2.893$. Additionally an intervening \mgii~system is detected at
+$z = 1.998$. From the \lya~absorption trough, we additionally infer $\log
+(N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.0 \pm 0.2$.
+
+\subsection{GRB~111117A (z=2.211)}
+
+This data has previously been used to form some of the basis of
+\citet{Selsing2017}. Starting 37.3 hours after the BAT trigger, 4.8ks of
+spectroscopic integration yields faint emission lines identified at \oii, \hb,
+\oiii~and \ha, all at a common $z = 2.211$, marking it as the redshift of the
+GRB. No afterglow continuum is detected.
+
+\subsection{GRB~111123A  (z=3.151)}
+
+This data formed the basis of GCN \#
+14273\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14273.gcn3}}, but is not
+published elsewhere. Observed twice, the first time shortly after the GRB and
+the second long after the burst had faded, securely sets the redshift of the
+host at $z = 3.151$ based on the detection of emission lines identified as
+\oii~and \oiii.
+
+\subsection{GRB~111129A (z=1.080)}
+
+Starting 8.26 hours after and lasting 3.6ks, these observations have previously
+been published in \citet{Kruhler2015}. A very faint continuum is visible after
+severe binning and a redshift is suggested in \citet{Kruhler2015}, based on the
+detection of \oii. At this redshift, \hb~and \oii~are located in the gap between
+the VIS and NIR arm and \ha is located in the middle of the $JH$-bandgap and
+there are therefore not detected.
+
+\subsection{GRB~111209A (z=0.677)}
+
+These spectra has previously been used in \citet{Levan2013, Greiner2015,
+	Kruhler2015, Kann2017} and additionally formed the basis for GCN \#
+12648\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12648.gcn3}}. The first epoch
+of spectroscopic observations was initiated 17.7 hours after the BAT trigger and
+lasted for 4.8 seconds. A very bright afterglow continuum is detected across the
+entire spectral coverage of X-shooter, with several absorption features
+imprinted. The absorption features are identified as \feii, \mgii, \mgi, \cah,
+and \cak - all at a common redshift of $z = 0.677$. The second epoch, taken 20
+days later, still contains a faint continuum detected across all arms. The
+detection of nebular emission lines identified as \oii, \oii~and \ha~at the same
+redshift, securely marks it at the redshift of this ultra-long GRB with
+accompanying GRB-SN.
+
+\subsection{GRB~111211A (z=0.478)}
+
+These data has formed the basis for GCN \#
+12677\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12677.gcn3}} and is also published in \cite{Kruhler2015}. Observations began 31 hours after the AGILE trigger
+and consisted of $4 \times 600$ s. A bright GRB afterglow continuum is detected
+across the entire spectral coverage of X-shooter with absorption and emission
+features visible. We identify absorption features due to \feii, \mgii, and \caii
+and emission lines from \oiii~and \ha, all at a common $z = 0.478$, which we
+suggest is the redshift of the GRB. Additionally detected in the GRB afterglow
+continuum is broad undulation, suggesting an accompanying SN.
+
+\subsection{GRB~111228A (z=0.716)}
+
+This data formed the basis of GCN \#
+12770\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12770.gcn3}} and is also published in \cite{Kruhler2015}.
+Observations began 15.9 hours after the BAT trigger and consist of $4 \times 600$ s.
+The GRB afterglow continuum is clearly detected in all the spectroscopic arms and superposed on the continuum are absorption features identified as due to \feii, \mnii, \mgii, \mgi, \cah, and \cak, all at $z = 0.716$.
+Supporting this redshift measurement as the redshift of the GRB is the detection of nebular emission line from \oiii.
+
+\subsection{GRB~120118B (z = 2.943)}
 The data presented here also formed the basis of GCN \#
 14225\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14225.gcn3}}, but is not
-published otherwise. This late-time observation of the host of GRB120118B
+published otherwise. This late-time observation of the host of GRB~120118B
 consists of 3.6 ks exposures and contains emission lines belonging to \oii and
@@
 
-\subsection{GRB120119A (z = 1.728)}
+\subsection{GRB~120119A (z = 1.728)}
 
@@
 
-\subsection{GRB120211A (z = 2.346)}
+\subsection{GRB~120211A (z = 2.346)}
 The data presented here has been published in \citet{Kruhler2015}. Two
-observations of the host of GRB120211A has been obtained, starting 2013.02.17,
+observations of the host of GRB~120211A has been obtained, starting 2013.02.17,
 $> 1 year$ after the burst has faded. A redshift for this object has been
@@
 
-\subsection{GRB120224A (z = 1.10 NEW!!!)}
+\subsection{GRB~120224A (z = 1.10 NEW!!!)}
 The data presented here has formed the basis of GCN \#
@@
 
-\subsection{GRB120311 (z = 0.350 NEW!!!)}
+\subsection{GRB~120311 (z = 0.350 NEW!!!)}
 The data presented here has formed the basis of GCN \#
@@
 
-\subsection{GRB120327A (z = 2.813)}
+\subsection{GRB~120327A (z = 2.813)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB120404A (z = 2.876)}
+\subsection{GRB~120404A (z = 2.876)}
 The data presented here has formed the basis of GCN \#
@@
 \textit{Swift}-trigger reveals a low-intensity afterglow continuum on which
-absorption from \lya is detected in two distinct regions at redshifts $z=2.876$
-and $z=.255$. These absorption systems are confirmed by ionic absorption
+absorption from \lya~is detected in two distinct regions at redshifts $z=2.876$
+and $z=2.55$. These absorption systems are confirmed by ionic absorption
 features at both of these redshifts.
 
-\subsection{GRB120422A (z = 0.283)}
+\subsection{GRB~120422A (z = 0.283)}
 The data presented here also formed the basis of GCN \# 
@@
 
-\subsection{GRB120712A (z = 4.175)}
-The data presented here also formed the basis of GCN \#
-13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
+\subsection{GRB~120712A (z = 4.175)}
+The data presented here also formed the basis of GCN \#13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
 published elsewhere. 4.8 ks integration time, starting 10.5 hours after the BAT
@@
 
-\subsection{GRB120714B (z = 0.398)}
-The data presented here also formed the basis of GCN \#
-13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
+\subsection{GRB~120714B (z = 0.398)}
+The data presented here also formed the basis of GCN \#13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
 published elsewhere. Observations of this burst started 7.8 hours after the GRB
@@
 
-\subsection{GRB120716A (z = 2.486)}
-The data presented here also formed the basis of GCN \#
-13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
+\subsection{GRB~120716A (z = 2.486)}
+The data presented here also formed the basis of GCN \#13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
 published elsewhere. Despite observations starting 62 hours after the
@@
 
-\subsection{GRB120722A (z = 0.959)}
-The data presented here also formed the basis of GCN \#
-13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
+\subsection{GRB~120722A (z = 0.959)}
+The data presented here also formed the basis of GCN \#13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
 published elsewhere. On 4.8 ks integration time, starting 10 hours after the
@@
 
-\subsection{GRB120805A (z $\sim$ 3.9 NEW!!!)}
+\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)}
 A separate reduction of this burst has been published in \citet{Kruhler2015},
@@
 
-\subsection{GRB120815A* (z = 2.358)} 
+\subsection{GRB~120815A* (z = 2.358)} 
 Not a part of the statistical sample, thisburst also formed the basis of  GCN
-\# 13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
+\#13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
 published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
@@
 
-\subsection{GRB120909A (z = 3.929)}
-The data presented here has formed the basis of GCN \#
-13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
+\subsection{GRB~120909A (z = 3.929)}
+The data presented here has formed the basis of GCN \#13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
 published otherwise. A very rapid follow-up, starting only 1.7 hours after the
@@
 
-\subsection{GRB120923A (z = $\gtrsim8$?)}
-
-\subsection{GRB121024A (z = 2.300)}
-The data presented here also formed the basis of GCN \#
-13890\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13890.gcn3}} and is published
-in \citet{Friis2015}. Also rapid, starting 1.8 hours after the \textit{Swift}
-trigger, a bright afterglow continuum is visible across all arms. A broad
-absorption feature from Lyman alpha, along with narrow lines from \civ, \SIii,
-\SIiv, \feii, \sii, and \alii, as well as fine structure lines associated with
-\SIii$^*$ are all detected at $z = 2.300$, securely setting it as the redshift
-of the GRB.
-
-\subsection{GRB121027A (z = 1.773)}		
-The data presented here has formed the basis of GCN \#
-13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
+\subsection{GRB~120923A (z = $\gtrsim8$?)}
+
+\subsection{GRB~121024A (z = 2.300)}
+
+The data presented here also formed the basis of GCN
+\#13890\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13890.gcn3}} and is
+published in \citet{Friis2015}. Also rapid, starting 1.8 hours after the
+\textit{Swift} trigger, a bright afterglow continuum is visible across all arms.
+A broad absorption feature from Lyman alpha, along with narrow lines from \civ,
+\SIii, \SIiv, \feii, \sii, and \alii, as well as fine structure lines associated
+with \SIii$^*$ are all detected at $z = 2.300$, securely setting it as the
+redshift of the GRB.
+
+\subsection{GRB~121027A (z = 1.773)}
+
+The data presented here has formed the basis of GCN
+\#13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
 published otherwise. Starting 69.6 hours after the GRB trigger, that we detect
@@
 
-\subsection{GRB121201A (z = 3.385)}
-
-\subsection{GRB121229A (z = 2.707)}
-
-\subsection{GRB130131B (z = 2.539)}
-
-\subsection{GRB130408A (z = 3.758)}
-The data presented here also formed the basis of GCN \#
-14365\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14365.gcn3}}. The spectrum
+\subsection{GRB~121201A (z = 3.385)}
+
+These data formed the basis for GCN
+\#14035\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14035.gcn3}} and is
+additionally published in \citet{Kruhler2015}. These observations started 12.9
+hours after the \swift~trigger and consists of 4.8ks spectroscopic integration
+under good atmospheric conditions. The GRB afterglow continuum is well detected
+at all arms. A broad absorption trough due to \lya is visible at $z = 3.385$,
+which along with the detection of absorption features identified as \SIiv, \civ,
+\alii, and \aliii, marks it as the redshift of the GRB. In the middle of the
+\lya~trough, we additionally detect \lya~emission. By modelling the
+\lya~absorption, we infer $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 22.0 \pm
+0.3$.
+
+\subsection{GRB~121209A (z = 2.1)}
+
+This data is published in \citet{Kruhler2015}, but is not used elsewhere. This
+short spectroscopic integration of 1.2ks reveals a very faint trace visible in
+the visual arm after binning. The spectrum is taken at high airmass and
+observations were discontinued due to the hardware limit of the telescope. No
+emission lines are detected in the spectrum. The redshift is adopted from
+\citet{Kruhler2015}.
+
+\subsection{GRB~121229A (z = 2.707)}
+
+These data formed the basis for GCN
+\#14120\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14120.gcn3}}, but is not
+published elsewhere. Taken under poor seeing conditions, a total of 4.8ks
+spectroscopic integration starting 2 hours after the \swift~trigger yields a low
+S/N GRB afterglow continuum is all arms. Binning the spectrum reveals broad
+absorption troughs, which we identify as \lyb~and \lya~at $z = 2.707$.
+Additionally, an intervening system at $z = 1.658$ is detected from absorption
+features of \mgii. From the absorption trough due to \lya, we infer $\log
+(N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.7 \pm 0.2$. Due to strong contamination
+in the slit, the background is slightly over subtracted, cause the background to
+go negative in the center of the \lya~trough.
+
+\subsection{GRB~130131B (z = 2.539)}
+
+These data formed the basis for GCN
+\#14286\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14286.gcn3}} and is
+additionally published in \citet{Kruhler2015}. This is a late-time observation,
+taken long after the GRB afterglow has faded. In 7.2ks spectroscopic
+integration, emission lines identified as \oii~and \oiii~are detected at a
+common $z = 2.539$, which we suggest is the redshift of the GRB.
+
+\subsection{GRB~130408A (z = 3.758)}
+
+The data presented here also formed the basis of GCN
+\#14365\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14365.gcn3}}. The spectrum
 has not otherwise been published previously. The observations consists of two
@@
 
-\subsection{GRB130418A (z = 1.218)}
-
-\subsection{GRB130427A (z = 0.34)}
-
-\subsection{GRB130427B (z = 2.78)}
-
-\subsection{GRB130603B (z = 0.356)}
-
-\subsection{GRB130606A (z = 5.913)}
-The data presented here also formed the basis of GCN \#
-14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is published
-in \citet{Hartoog2015}. The observations consists of three 2x600sec visits
-starting 7.1 hrs after the burst at fairly high airmass. We detect absorption
-features from a wide range of ions at $z=5.913$ as well as intervening
-absorption at $z=2.3103, 2.5207, 3.4515, 4.4660, 4.5309, 4.5427, 4.6497 $ and $
-4.7244$.
-
-\subsection{GRB130612A (z = 2.006)}
-
-\subsection{GRB130615A (z = $~3$?)}
-
-\subsection{GRB130701A (z = 1.155)}
-
-\subsection{GRB131011A (z = 1.874)}
-
-\subsection{GRB131030A (z = 1.296)}
-
-\subsection{GRB131103A (z = 0.599)}
-
-\subsection{GRB131105A (z = 1.686)}
-
-\subsection{GRB131117A (z = 4.042)}
-
-\subsection{GRB131231A (z =0.642)}
-
-\subsection{GRB140114A (z = $\sim2.8$?)}
-
-\subsection{GRB140213A (z = 1.208)}
-
-\subsection{GRB140301A (z = 1.416)}
-
-\subsection{GRB140311A (z = 4.95)}
-
-\subsection{GRB140430A (z = 1.601)}
-
-\subsection{GRB140506A  (z = 0.889)}	
+\subsection{GRB~130418A (z = 1.222)}
+
+GCN \#14390\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14390.gcn3}} is based on
+this spectrum, but it is not published elsewhere. Starting only 4.57 hours after
+the \swift~trigger, 1.2ks observations contains a bright GRB afterglow
+continuum, visible across the entire spectral coverage of X-shooter. Superposed
+on the afterglow continuum are absorption features which we identify as \civ,
+\feii, and \mgii, caused by an absorber at $z = 1.217$, and additional
+absorption from \civ~ at $z = 1.222$. The two systems are offset by $\sim
+1500~\mathrm{km}~\mathrm{s}^{-1}$ and the proximity of the two absorption
+systems in velocity space, suggests a possible association of the two systems
+with peculiar velocity affecting the measured redshift. We adopt $z = 1.222$ as
+the redshift of the GRB. Note, that this value is slightly different from the
+one reported in GCN \# 14390.
+
+\subsection{GRB~130427A (z = 0.340)}
+
+This spectrum is also published in \citet{Xu2013b} and \citet{Kruhler2015} and
+additionally has formed the basis for GCN
+\#14491\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14491.gcn3}}. Starting 16.5
+hours after the BAT trigger, these observations lasting $2 \times 600$ s
+contains a very bright GRB afterglow continuum across the total spectral
+coverage of X-shooter. In absorption we identify features from the following
+metal resonance lines: \feii, \mnii, \mgii, \mgi, \TIii~and additional line
+absorption from \caii~and \nai. Simultaneously, we find emission lines from \ha,
+\hb, \oiii, \oii - all at common redshift of $z = 0.340$, which is the redshift
+of the GRB. This is one of the most energetic GRBs observed, and it's proximity
+along with the associated broad-lined Type Ic SN, 2013cq has caused it to be one
+of the more well studies GRBs.
+
+\subsection{GRB~130427B (z = 2.780)}
+
+This spectrum formed the basis of GCN
+\#14493\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14493.gcn3}}, but is not
+published otherwise. A short, 2x600 s spectroscopic integration obtained before
+twilight, 20.6 hours after the BAT trigger, captures a faint afterglow continuum
+visible across the entire spectral coverage at low S/N. Due to the low S/N, the
+metal lines are weak, but the broad absorption though due to \lya~is detected.
+From this we measure the redshift to be $z = 2.780$ and provides a measure of
+the neutral hydrogen column density, \nh~$= 21.9 \pm 0.3$. The redshift is
+confirmed by the presence of \feii~at a consistent redshift.
+
+\subsection{GRB~130603B (z = 0.356)}
+
+This burst is the first sGRB observed with a potential associated
+kilonova\citep{Tanvir2013a, Berger2013a}.  GCN
+\#14757\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14757.gcn3}} was based on
+this spectrum, but it is not published elsewhere. Starting 8.2 hours after the
+\swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
+Continuum is clearly detected across all arms from both host and afterglow and
+superposed are both absorption (\cahk~and \mgii) and emission lines (\oii, \hb,
+\oiii, \ha, and \sii), all at a consistent redshift of $z = 0.356$, which is the
+redshift of the GRB.
+
+\subsection{GRB~130606A (z = 5.913)}
+
+The data presented here also formed the basis of GCN
+\#14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is
+published in \citet{Hartoog2015}. The observations consists of three
+$2\times600$ s visits starting 7.1 hrs after the burst at fairly high airmass.
+We detect absorption features from a wide range of ions at $z=5.913$ as well as
+intervening absorption at $z=2.3103, 2.5207, 3.4515, 4.4660, 4.5309, 4.5427,
+4.6497 $ and $ 4.7244$.
+
+\subsection{GRB~130612A (z = 2.006)}
+
+The spectral features of this spectrum has previously been reported in GCN
+\#14882\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14882.gcn3}}, but is not
+published elsewhere. Starting only 1.1 hour after the \swift~trigger,
+$2\times600$ s spectroscopic integration captures a moderate S/N afterglow
+continuum across the total spectral coverage of X-shooter. At a consistent
+redshift of $z = 2.006$, absorption from the metal resonance lines \feii, \mnii,
+\mgii, \mgi~are identified. Additionally, \lya~is visible as a broad absorption
+trough, from we can infer \nh~= $22.2 \pm 0.2$, which is in the upper end of the
+hydrogen column density distribution. The blue part of the GRB continuum
+exhibits a downturn in the conitnuum level which could indicate the presence of
+a significant amount of dust along the line-of-sight.
+
+\subsection{GRB~130615A (z $\sim 3$)}
+
+This spectrum has not previously been published. Starting only 45 minutes after
+the BAT trigger, $2\times600$ s spectroscopic integration carried on into the
+beginning twilight. Observed at very high airmass with a quickly varying
+background, a faint afterglow trace is visible across all arms of X-shooter,
+down to ~ 4800 \AA, confirming the approximate redshift suggested in GCN
+\#14898\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14898.gcn3}}.
+
+\subsection{GRB~130701A (z = 1.155)}
+
+These data formed the basis for GCN
+\#14956\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14956.gcn3}} and is
+additionally published in \citet{Kruhler2015}. Starting 5.5 hours after the GRB
+trigger, $2\times600$ s reveals a bright continuum visible across the entire
+spectral coverage of X-shooter. Superposed are absorption features which we
+identify as due to \feii, \mgii, \mgi, and \caii~- all at a consistent redshift
+of $z = 1.155$, which we take to be the redshift of the GRB.
+
+\subsection{GRB~131011A (z = 1.874)}
+
+This data formed the basis for
+\#15330\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15330.gcn3}}, but is not
+published elsewhere. Starting $\sim$1.5 days after the Fermi-GBM trigger, 4.5 ks
+spectroscopic integration captures a modest S/N GRB afterglow continuum all the
+way down to $\sim$3200 \AA. Imprinted on the continuum are absorption features,
+which we identify as due to \lya, \feii,  \mgii, \mgi~at the same redshift,
+which we measure to be $z = 1.874$. From the broad absorption trough due to
+\lya, we infer \nh~$= 22.0 \pm 0.3$. This spectrum is taken under ESO programme
+ID: 092.D-0056(A) (PI: Rau),
+
+\subsection{GRB~131030A (z = 1.296)}
+
+\subsection{GRB~131103A (z = 0.599)}
+
+\subsection{GRB~131105A (z = 1.686)}
+
+\subsection{GRB~131117A (z = 4.042)}
+
+\subsection{GRB~131231A (z =0.642)}
+
+\subsection{GRB~140114A (z = $\sim2.8$?)}
+
+\subsection{GRB~140213A (z = 1.208)}
+
+\subsection{GRB~140301A (z = 1.416)}
+
+\subsection{GRB~140311A (z = 4.95)}
+
+\subsection{GRB~140430A (z = 1.601)}
+
+\subsection{GRB~140506A  (z = 0.889)}	
 The data presented here has formed the basis of GCN \#
@@
 
-\subsection{GRB140515A (z = $\sim6.32$?)}
-
-\subsection{GRB140614A (z = 4.233)}
-
-\subsection{GRB140622A (z = 0.959)}
-
-\subsection{GRB141028A (z = 2.332)}
-
-\subsection{GRB141031A  (z = na)}
-
-\subsection{GRB141109A (z = 2.993)}
-
-\subsection{GRB150206A (z = 2.087)}
-
-\subsection{GRB150301B (z = 1.517)}
-
-\subsection{GRB150403A (z = 2.06)}
-
-\subsection{GRB150423A (z = 1.394)}
-
-\subsection{GRB150428A (z = na)}
-
-\subsection{GRB150514A (z = 0.807)}
-
-\subsection{GRB150518A (z = 0.256)}
-
-\subsection{GRB150616A (z = 1.188)}
-
-\subsection{GRB150727A (z = 0.313)}
-
-\subsection{GRB150821A (z = 0.755)}
-
-\subsection{GRB150831A (z = na)}
-
-\subsection{GRB150910A (z = 1.359)}
-
-\subsection{GRB150915A (z = 1.968)}
-
-\subsection{GRB151021A (z = 2.330)}
+\subsection{GRB~140515A (z = $\sim6.32$?)}
+
+\subsection{GRB~140614A (z = 4.233)}
+
+\subsection{GRB~140622A (z = 0.959)}
+
+\subsection{GRB~141028A (z = 2.332)}
+
+\subsection{GRB~141031A  (z = na)}
+
+\subsection{GRB~141109A (z = 2.993)}
+
+\subsection{GRB~150206A (z = 2.087)}
+
+\subsection{GRB~150301B (z = 1.517)}
+
+\subsection{GRB~150403A (z = 2.06)}
+
+\subsection{GRB~150423A (z = 1.394)}
+
+\subsection{GRB~150428A (z = na)}
+
+\subsection{GRB~150514A (z = 0.807)}
+
+\subsection{GRB~150518A (z = 0.256)}
+
+\subsection{GRB~150616A (z = 1.188)}
+
+\subsection{GRB~150727A (z = 0.313)}
+
+\subsection{GRB~150821A (z = 0.755)}
+
+\subsection{GRB~150831A (z = na)}
+
+\subsection{GRB~150910A (z = 1.359)}
+
+\subsection{GRB~150915A (z = 1.968)}
+
+\subsection{GRB~151021A (z = 2.330)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB151027B (z = 4.063)}
-
-\subsection{GRB151029A (z = 1.423)}
-
-\subsection{GRB151031A (z = 1.167)}
-
-\subsection{GRB160117B (z = 0.87)}
-
-\subsection{GRB160203A (z = 3.517)}
+\subsection{GRB~151027B (z = 4.063)}
+
+\subsection{GRB~151029A (z = 1.423)}
+
+\subsection{GRB~151031A (z = 1.167)}
+
+\subsection{GRB~160117B (z = 0.87)}
+
+\subsection{GRB~160203A (z = 3.517)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB160228A (z = 1.64)}
-
-\subsection{GRB160303A (z = na)}
-
-\subsection{GRB160314A (z = 0.726)}
-
-\subsection{GRB160410A (z = 1.717)}
-
-\subsection{GRB160425A (z = 0.555)}
-
-\subsection{GRB160625B (z = 1.406)}
-
-\subsection{GRB160804A (z = 0.736)}
+\subsection{GRB~160228A (z = 1.64)}
+
+\subsection{GRB~160303A (z = na)}
+
+\subsection{GRB~160314A (z = 0.726)}
+
+\subsection{GRB~160410A (z = 1.717)}
+
+\subsection{GRB~160425A (z = 0.555)}
+
+\subsection{GRB~160625B (z = 1.406)}
+
+\subsection{GRB~160804A (z = 0.736)}
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB161001A (z = 0.891?)}
-
-\subsection{GRB161007A (z =4.6 ??? NEW!!!)}
-This data has not been published elsewhere. Observations for GRB161007A started
+\subsection{GRB~161001A (z = 0.891?)}
+
+\subsection{GRB~161007A (z =4.6 ??? NEW!!!)}
+This data has not been published elsewhere. Observations for GRB~161007A started
 323 hours after the burst trigger and contains the potential host. 4 x 600 seconds
@@
 
-\subsection{GRB161014A (z =2.823)}
+\subsection{GRB~161014A (z =2.823)}
 The data presented here also formed the basis of GCN \#
@@
 along with metal absorption features from \mgii, \SIii, \cii, \civ, \alii,
-\aliii, and	\feii, all at $z =2.823$. Similar to GRB140506 \citep{Fynbo2014}, a
+\aliii, and	\feii, all at $z =2.823$. Similar to GRB~140506 \citep{Fynbo2014}, a
 break in the continuum shape is detected bluewards of 600 nm, possible
@@
 
-\subsection{GRB161023A (z = 2.710)}
-
-\subsection{GRB161117A (z = 1.549)}
+\subsection{GRB~161023A (z = 2.710)}
+
+\subsection{GRB~161117A (z = 1.549)}
 

### 2018-01-20 py/betaOX.py

@@ -125,6 +125,10 @@ def main():
     idx_limit = idx_limit[idx_2]
     g.x = HI[~idx_limit]
     g.y = betaOX[idx_2][~idx_limit]
+
+    np.savetxt("betaOX.dat", list(zip(name_s[idx][~idx_limit] ,HI[~idx_limit], betaOX[idx_2][~idx_limit])), fmt = '%s %s %s')
+    exit()
+
     g = g.plot_joint(pl.scatter, color=color, label="This work")
 
     g.x = HI[idx_limit]

### 2018-01-20 py/skymap.py

@@ -48,7 +48,8 @@ def main():
     ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
     A_Vmap = ebv_map*3.1
     A_Vmap[A_Vmap >= 0.5] = 0
-    ax = hp.mollview(A_Vmap, max=0.5, cbar=False, title="")
+    ax = hp.mollview(A_Vmap, max=0.5, title="")
+    # , cbar=False
     hp.graticule()
 
     # Overplot swift positions
@@ -87,7 +88,7 @@ def main():
     b_z = b[mask_z]
 
 
-    print(l_z, b_z)
+    # print(l_z, b_z)
     # Scatterplot with bursts position
     hp.visufunc.projscatter(l_z, b_z, facecolors='none', edgecolor="black", marker = "*", s = 50, lonlat=True)
 
@@ -106,7 +107,7 @@ def main():
     b_noz = b[mask_noz]
 
     # Scatterplot with bursts position
-    hp.visufunc.projscatter(l_noz, b_noz, facecolors='none', lonlat=True)
+    hp.visufunc.projscatter(l_noz, b_noz, facecolors='none', edgecolor="black", s = 30, lonlat=True)
 
     # Get indices of sample bursts and observed bursts
     idx = [ii for ii, kk in enumerate(name_s_noz) if kk in name]

### 2018-01-20 py/spec_plot.py

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -12,12 +13,12 @@ from scipy import stats, signal, interpolate
 import matplotlib as mpl
 
 params = {
-   'axes.labelsize': 10,
-   'font.size': 10,
-   'legend.fontsize': 10,
-   'xtick.labelsize': 10,
-   'ytick.labelsize': 10,
-   'text.usetex': False,
+   'axes.labelsize': 14,
+   'font.size': 14,
+   'legend.fontsize': 14,
+   'xtick.labelsize': 14,
+   'ytick.labelsize': 14,
+   'text.usetex': True,
    'figure.figsize': [8.27, 11.7]
    }
 mpl.rcParams.update(params)
@@ -36,14 +37,15 @@ def main():
     # Get linelist
     import lineid_plot
     fit_line_positions = np.genfromtxt("../data/grblines_Christensen2011.dat", dtype=None)
-    # fit_line_positions = np.genfromtxt("../data/grblines.dat", dtype=None)
-    linelist = []
-    linenames = []
-    for n in fit_line_positions:
-        linelist.append(float(n[0]))
-        linenames.append(str(n[1]))
+
+    linelist = [0]*len(fit_line_positions)
+    linenames = [0]*len(fit_line_positions)
+    for ii, nn in enumerate(fit_line_positions):
+        linelist[ii] = float(nn[0])
+        linenames[ii] = nn[1].decode("utf-8")
     linelist = np.array(linelist)
     linenames = np.array(linenames)
+    # print(linenames)
     # Redshifts of systems
     z = [2.300]
     ls = ["dashed"]

### 2018-01-21 document/XSGRB_sample.tex
Final notes on 2013 bursts

@@
 
+\subsection{GRB~130925A (z = 0.347)}
+
+This spectrum has already been used in GCN
+\#15250\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15250.gcn3}} and is
+additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
+Observations of this burst began with X-shooter 3.5 hours after the
+\swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
+afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
+primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
+\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, which we as the
+redshift of the GRB. This spectrum is taken under ESO programme ID:
+091.A-0877(A) (PI: Schady).
+
 \subsection{GRB~131011A (z = 1.874)}
@@
 \lya, we infer \nh~$= 22.0 \pm 0.3$. This spectrum is taken under ESO programme
-ID: 092.D-0056(A) (PI: Rau),
-
-\subsection{GRB~131030A (z = 1.296)}
+ID: 092.D-0056(A) (PI: Rau).
+
+\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}
+
+This data has not been published before. Starting 3.4 hours after the
+\swift~trigger, $6\times600$ s exposure taken under good conditions, contains a
+bright GRB afterglow continuum across the entire spectral coverage of X-shooter.
+A myriad of absorption features are superposed on the afterglow continuum which
+we identify as being caused by \SIiv, \SIii, \civ, \alii, \aliii, \znii, \crii,
+\NIii, \feii, \NIii$^*$, and \feii$^*$ at $z = 1.296$. A very strong
+\mgii-absorber is also detected, intervening the line-of-sight at $z = 1.164$
+with lines from \SIii, \civ, \aliii, \aliii, \feii, \mnii, and many more.
 
@@
 
+This spectrum has already been used to form the basis for GCN
+\#15451\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15451.gcn3}} and is
+additionally published in \citet{Kruhler2015}. Starting 5.8 hours after the BAT
+trigger, $4\times600$ s exposure captures a modest S/N continuum across all
+arms. Imprinted on the continuum are absorption features identifed as due to
+\feii~and \mgii~as well as emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
+and \nii. All the lines are measured at a consistent redshift of $z = 0.599$,
+which we take as the redshift of the GRB.
+
 \subsection{GRB~131105A (z = 1.686)}
 
+This spectrum has already been used in shaping GCN
+\#15450\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15450.gcn3}} and is
+additionally published in \citet{Kruhler2015}. Starting only 1.3 hours after the
+\swift~trigger, a total of 4.8 ks spectroscopic integration contains a low S/N
+GRB afterglow continuum across the entire spectral coverage of X-shooter. There
+are deviations from the continuum at both emission and absorption. We identify
+lines from \hb, \oiii, and \ha~in emission and \feii, and \mgii~in absorption.
+All lines are at a consistent $z = 1.686$, which is probably the redshift of the
+GRB. Absorption lines at shorter wavelengths are also detected, but at low
+significance due to an apparent downturn in the continuum caused by the presence
+of dust local to the burst.
+
 \subsection{GRB~131117A (z = 4.042)}
 
-\subsection{GRB~131231A (z =0.642)}
+This spectrum has previously been used in GCN
+\#15494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15494.gcn3}}, but is not
+published in a refereed paper. Starting only 68 minutes after the BAT trigger,
+4.8 ks spectroscopic integration secures afterglow continuum for this bursts,
+which is measured to be at $z = 4.042$. A moderate S/N GRB afterglow continuum
+is detected down to $\sim$ 6100 \AA, signifying the onset of the \lya~forest,
+with part of the forest also visible. Metal absorption lines from \SIii~and
+\SIiv~are detected at a consistent redshift.
+
+\subsection{GRB~131231A (z = 0.642)}
+
+This spectrum has previously been published in \citet{Kruhler2015} and
+addtionally forms the basis for GCN
+\#15645\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15645.gcn3}}. This spectrum,
+observed the following year (20.2 hours after the \swift~trigger), consists of
+$4\times600$ s exposures. A high S/N GRB afterglow continuum is detected all the
+way though the X-shooter arms. We identify absorption features imprinted in the
+continuum as caused by \feii, \mgii, and \cahk at a consistent $z = 0.642$. By
+subtracting off the bright afterglow continuum, we readily detect emission lines
+arising from \oii, \hg, \hb, \oiii, and \ha~in the GRB host galaxy.
 

### 2018-01-22 document/XSGRB_sample.tex
Comments by Jure

@@
 \abstract{In this work we present the spectra of all $\gamma$-ray bursts (GRBs)
-	afterglows that have been observed at the VLT/X-shooter up to \termdate. This
-	totals 137 spectroscopic observations of 118 individual bursts, of which
-	redshifts have been secured for 91 \% covering a redshift range from 0.059 to
-	$\sim$ 8.5. Some of these observation are late-time observations of the
-	underlying host galaxy. Based on an set of selection criteria that are unbiased
-	with regards to intrinsic properties of the GRBs, the follow-up effort has been
-	guided by the \textit{Swift}-telescope to produce a homogeneous sample of
-	afterglows for 86 spectra. We here provide a public release of all the reduced
-	spectra, including continuum estimates and telluric corrections. We provide an
-	assessment of the degree of completeness with respect to the parent GRB
-	population, in terms of the X-ray properties of the bursts in the sample and
-	find that the sample presented here is consistent with being drawn from the
-	\textit{Swift} sample. For all bursts (39) for which it is possible, we provide
-	a measurement of the hydrogen column density, increasing the total number of
-	published HI column density measurements by $\sim$ 50 per cent. We constrain the
-	fraction of dark bursts to be less than 28 per cent based on the detection of an
-	optical afterglow and tentatively confirm previous results that larger optical
-	darkness is correlated with increased X-ray absorption. This dataset provides a
-	unique possibility to study the ISM across cosmic time. }
+	afterglows that have been promptly observed at the VLT/X-shooter up to
+	\termdate. This totals spectroscopic observations of 102 individual bursts
+	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
+	\% of these, covering a redshift range from 0.059 to 7.84. Based on an set of
+	selection criteria that are unbiased with regards to intrinsic properties of the
+	GRBs, the follow-up effort has been guided by the \textit{Swift}-telescope to
+	produce a homogeneous sample of afterglows for 86 spectra. We here provide a
+	public release of all the reduced spectra, including continuum estimates and
+	telluric corrections. For completeness, we also provide reductions for the
+	late-time observations of the underlying host galaxies. We provide an assessment
+	of the degree of completeness with respect to the parent GRB population, in
+	terms of the X-ray properties of the bursts in the sample and find that the
+	sample presented here is consistent with being drawn from the \textit{Swift}
+	sample. For all bursts (39) for which it is possible, we provide a measurement
+	of the hydrogen column density, increasing the total number of published HI
+	column density measurements by $\sim$ 50 per cent. We constrain the fraction of
+	dark bursts to be less than 28 per cent based on the detection of an optical
+	afterglow and tentatively confirm previous results that larger optical darkness
+	is correlated with increased X-ray absorption. This dataset provides a unique
+	possibility to study the ISM across cosmic time. }
 
@@
 
-In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy was
-secured using low-resolution spectrographs \citep[typically
+In the beginning of the "{\it Swift} era" most of the follow-up afterglow
+spectroscopy was secured using low-resolution spectrographs \citep[typically
 R=$\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Low-resolution
-spectroscopy is powerful for a number of reasons, first of all is the ability to
-secure information for even very faint targets. This allows the measurements of
-a number of important parameters such as redshifts, spectral slopes, and
-extinction. For a handful of very bright afterglows high-resolution (typically
-R$>$20000) spectra were secured, and for these events much more information
-about conditions inside the host galaxies were extracted
-\citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007, Vreeswijk2007,
-	Castro-Tirado2010}.
+spectroscopy is powerful as it allows us to secure information even for very faint
+targets. This allows the measurements of a number of important parameters such
+as redshifts, spectral slopes, and extinction. For a handful of very bright
+afterglows high-resolution (typically R$>$20000) spectra were secured, and for
+these events much more information about conditions inside the host galaxies
+were extracted \citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007,
+	Vreeswijk2007, Castro-Tirado2010}.
 
@@
 4000--9000 in order to be able to get a large useful spectral range between the
-many sky-background emission lines in the red and near-IR spectral ranges.
+many sky-background emission lines in the red and near-IR spectral ranges. The
+large spectral coverage allows for the spectroscopic observations of the
+highest redshift GRBs.
 
@@
 the follow-up effort to bursts detected by \textit{Swift}, we therefore ensure
-that the limitations of the parent sample is well studied.
+that the limitations of the parent sample are well studied.
 
@@
 observations to completed. Since the GRB population is isotropically distributed
-on the sky as seen from Earth and because the GRB properties does not depend on
+on the sky as seen from Earth and because the GRB properties do not depend on
 position on the sky \citep{Briggs1996}, this cut does not influence our ability
-to fairly sample the underlying GRB population. The same arguments applies to
+to fairly sample the underlying GRB population. The same arguments apply to
 the requirement that there are no nearby bright stars. We additionally require
@@
 We further discuss the effect of these selection criteria and their implication
-for the compleness of the sample in Sect. \ref{completeness}.
+for the completeness of the sample in Sect. \ref{completeness}.
 
@@
 
-For 8 of the observed GRBs released here, RRM triggered. One of the RRM triggers
-is outside the statistical sample and two of the RRM trigger are on short GRBs.
-The fastest response between \swift~GRB trigger and the beginning of
-spectroscopic integration is for GRB~160410A, for which the delay was only 8.4
-minutes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Observations} \label{obs}
+There 8 GRBs in the sample that were being observed with the RRM mode. One of
+the RRM triggers is outside the statistical sample and two of the RRM trigger
+are on short GRBs. The fastest response between \swift~GRB trigger and the
+beginning of spectroscopic integration is for GRB~160410A, for which the delay
+was only 8.4 minutes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Observations} \label{obs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	in galactic coordinates of the full sky showing the positions on the sky of the
-	bursts presented in this work. The equator is the Galactic plane. The empty
-	circles are the positions of all the \textit{Swift} bursts detected until the
-	end of 2016. In blue is shown the position of the bursts fulfilling the sample
-	criteria specified in Sect. \ref{samplecrit}. The red circles are added around
-	GRBs which have been followed up by X-shooter since the commissioning of the
-	instrument. Blue points with red circles represent GRBs that enter our sample
-	and have X-shooter spectroscopy. The different samples are compared in Sect.
-	\ref{results}. The background shows the dust maps presented in
-	\citet{Schlegel1998} where the sample criteria cut with $A_V \lesssim 0.5$ mag
-	is removed. The colorbar below indicates the value of $A_V$ on the plot. The
-	dotted lines indicate intervals of 30$^\circ$ in longitude and latitude.}
-\label{fig:skymap}
+	bursts presented in this work. The equator is the Galactic plane. Star symbols
+	indicate bursts with measured redshifts and the circles indicate those without.
+	The empty stars/circles are the positions of all the \textit{Swift} bursts
+	detected until the end of 2016. A blue filling of the symbol indicates the
+	position of bursts fulfilling the sample criteria specified in Sect.
+	\ref{samplecrit}. The red circles are added around GRBs which have been
+	followed up by X-shooter since the commissioning of the instrument. Blue points
+	with red circles represent GRBs that enter our sample and have X-shooter
+	spectroscopy. The different samples are compared in Sect. \ref{results}. The
+	background shows the dust maps presented in \citet{Schlegel1998} where the
+	sample criteria cut with $A_V \lesssim 0.5$ mag is removed. The colorbar below
+	indicates the value of $A_V$ on the plot. The dotted lines indicate intervals
+	of 30$^\circ$ in longitude and latitude.} \label{fig:skymap}
 \end{figure*}
@@
 For a typical observation, each of the exposures in the nodding sequence has
-been reduced as single observation and then subsequently combined to form a
+been reduced as a single observation and then subsequently combined to form a
 single image. Because this strategy is employed, we can reject outliers in the
@@
 background. When weighting images where the noise in each pixel is dominated by
-Poisson noise it important to estimate the background variance in a large enough
+Poisson noise it is important to estimate the background variance in a large enough
 region, so that any correlation between the signal and the weights are removed.
 To this end, the weight map is generated by a running median window over the
-variance map, where the trace as been masked and width of the window is chosen
+variance map, where the trace has been masked and width of the window is chosen
 to be wide enough for median variance to be generated on the basis of several
@@
 with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum,
-estimated as the mode of all flux values, has been subtracted. To get the correct
-seeing PSF with which to convolve the synthetic sky an initial refinement of the
-wavelength solution have been obtained by cross-correlating the observed sky
-with an unconvolved synthetic sky. This preliminary wavelength calibration is
-applied to the observed sky. The synthetic spectrum is then convolved with an
-increasing seeing PSF and the width that minimizes $\chi^2$ with the updated
-observed sky is chosen to be the effective sky-PSF. Using the synthetic sky with
-the matched resolution, a final wavelength calibration can then be calculated by
-cross-correlating the observed sky with the correctly broadened sky spectrum, as
-a function of a velocity offset. Both a multiplicative and an additive offset to
-the wavelength calibration has been tested, but in terms of $\chi^2$, the model
-with only a multiplicative offset is preferred.
-
-The resulting offsets, which were smaller than 0.01 nm in the UVB and VIS data
-and smaller than 0.05 nm in the NIR spectra, but changing over short period of
-time were applied to the corresponding spectra.
-Using the convolved synthetic sky, the pixels containing the brightest sky lines
-have been added to the bad pixel map.
+estimated as the mode of all flux values, has been subtracted. To get the
+correct seeing PSF with which to convolve the synthetic sky an initial
+refinement of the wavelength solution have been obtained by cross-correlating
+the observed sky with an unconvolved synthetic sky. This preliminary wavelength
+calibration is applied to the observed sky. The synthetic spectrum is then
+convolved with an increasing seeing PSF and the width that minimizes $\chi^2$
+with the updated observed sky is chosen to be the effective sky-PSF. Using the
+synthetic sky with the matched resolution, a final wavelength calibration can
+then be calculated by cross-correlating the observed sky with the correctly
+broadened sky spectrum, as a function of a velocity offset. Both a
+multiplicative and an additive offset to the wavelength calibration has been
+tested, but in terms of $\chi^2$, the model with only a multiplicative offset is
+preferred. The resulting offsets, which were smaller than 0.01 nm in the UVB and
+VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short
+period of time were applied to the corresponding spectra. Using the convolved
+synthetic sky, the pixels containing the brightest sky lines have been added to
+the bad pixel map.
 
@@
 \citep{Smette2015, Kausch2015}, which can then be used to correct for the
-absorption. The accuracy of the correction depends on the S/N for the target
-spectrum, where we have chosen the requirement that the afterglow continuum
-spectrum has a median signal-to-noise higher than a value of 10.
+absorption. The accuracy of the correction depends on the S/N per pixel for the
+target spectrum, where we have chosen the requirement that the afterglow
+continuum spectrum has a median signal-to-noise higher than a value of 10.
 
@@
 model by \citep{Noll2012, Jones2013} for which we generate a synthetic
-transmission spectrum, which we then use, where the ambient parameters for the
+transmission spectrum,  where the ambient parameters for the
 observations have been used.
@@
 	$r$ = 20, meaning it is in the brighter end of the sample presented here, but
-	not the brightest. The spectrum is rich in absorption, including absorption
-	from molecular $\mathrm{H_2}$. The absorption trough visible at $\sim$ 4000
-	$\AA$ is due to \lya~in the host. We have overplot the most prominent lines
-	seen in GRB afterglows from \citet{Christensen2011a}. Additionally, three
-	intervening systems are seen in this sightline This spectrum has been analyzed
-	in detail in \citet{Friis2015}.} \label{fig:spectrum}
+	not the brightest. The spectrum is rich in absorption and emission, including
+	absorption from molecular $\mathrm{H_2}$. The absorption trough visible at
+	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have overplot the most
+	prominent lines seen in GRB afterglows from \citet{Christensen2011a}.
+	Additionally, three intervening systems are seen in this sightline This
+	spectrum has been analyzed in detail in \citet{Friis2015}.}
+\label{fig:spectrum}
 \end{figure*}
@@
 presented here. For this, we have developed an algorithmic approach that
-attempts to automatically finds the continuum placement along with the error on
+attempts to automatically estimate the continuum placement along with the error on
 the continuum estimate through an iterative procedure. The method is entirely
@@
 coverage are pruned. The remaining filtered and pruned points are then spline
-interpolated using \texttt{scipy.interpolate.splrep;splev}, which serves as a
+interpolated using \texttt{scipy.interpolate.splrep.splev}, which serves as a
 first estimate of the continuum placement. An attempt to identify absorption and
@@
 point sources (this is evidently not the case for extended sources, e.g for host
-galaxies). The delivered resolution for slit width dominated spectra
+galaxies). The delivered resolution for slit-width-dominated spectra
 post-reduction and extraction can easily be determined from the bright sky
@@
 simple method is not terribly accurate (for one, the spatial profile of the
-trace is not a perfect Gaussian), but it gives a sufficiently accurate estimate
-for the analysis of these poor signal to noise science spectra.
+trace is not a perfect Gaussian), it nevertheless gives a sufficiently accurate
+estimate for the analysis of these poor signal to noise science spectra.
 
@@
 out by VLT/X-shooter from the commissioning of the instrument, \startdate, and
-until the end of the last period of the program 098.A-0055, \termdate and thus
+until the end of the last period of the program 098.A-0055, \termdate~and thus
 constitutes 8 years of GRB afterglow observations with VLT/X-shooter. An
@@
 completeness of the full statistical sample is 60 per cent. We return to the
-question of redshift completeness in in Sect. \ref{redshift}. From this sample,
+question of redshift completeness in Sect. \ref{redshift}. From this sample,
 90 GRBs have been spectroscopically followed up with X-shooter. In order to
@@
 	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
-	\caption{Relational graph shown the respective p-values that the different
-	sample are drawn from the same underlying distribution. The arrows represent
+	\caption{Relational graph showing the respective p-values that the different
+	samples are drawn from the same underlying distribution. The arrows represent
 	the comparison with each of the p-values are listed next to. Only in the HI
@@
 been observed due to reasons unrelated to the GRB or afterglow properties. The
-reasons includes unavailability of the telescope due to technical maintenance
+reasons include unavailability of the telescope due to technical maintenance
 (e.g., mirror re-coating), a visitor rejecting the TOO trigger, or bad weather.
@@
 One of the objectives of our follow-up campaign is to measure the redshift
-distribution for a well-defined, observationally unbiased and statistical useful
+distribution for a well-defined, observationally unbiased and statistically useful
 sample of GRBs. The imposed selection criteria (see Sect. \ref{samplecrit})
@@
 effective limit depends on the light curve of the prompt emission signal. Due to
-the dilution of light with distance, the \swift GRB luminosity detection limit
+the dilution of light with distance, the \swift~GRB luminosity detection limit
 is almost an order of magnitude brighter at $z=2$ than at $z=1$. At $z\geq4$
@@
 	sample are marked by blue stars and black dots show all GRBs observed with
-	X-shooter. All \swift~GRBs with measured redshifts is shown in grey. For
+	X-shooter. All \swift~GRBs with measured redshifts are shown in grey. For
 	comparison, we overplot with red stars the GRB \citep{Tanvir2009b,
@@
 their optical darkness,  $\beta_{OX}$\citep{Jakobsson2004, Rol2005,
-	VanderHorst2009}. The X-ray properties of such bursts has previously been
-investigated \citep{DePasquale2003, Fynbo2009, Melandri2012}. There is some
+	VanderHorst2009}. The X-ray properties of such bursts have previously been
+investigated \citep{DePasquale2003, Fynbo2009, Melandri2012}. There are some
 indications that dark bursts have somewhat higher X-ray luminosity and
-\nhx~compared to the optically brights bursts. This indicates along with
+\nhx~compared to the optically bright bursts. This indicates along with
 investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
@@
 
-In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_H$ distribution with
-the one presented in \citet{Fynbo2009}. The values from \citet{Fynbo2009} have
-been treated as detections, meaning that we artificially bias the distribution
-towards higher $\beta_{OX}$-values. We tentatively confirm the result by
-\citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher \nhx.
-Specifically, for bursts with measured redshift either from the afterglow or the
-host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$ we
-find $\log(N_H) = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
-$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
-within the error intervals. A 2-sided KS test fails to reject the null that they
-are drawn from the same distribution with $p = 0.43$, meaning there is not
-strong evidence for a discrepancy.
+In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{HI, X}}$
+distribution with the one presented in \citet{Fynbo2009}. The values from
+\citet{Fynbo2009} have been treated as detections, meaning that we artificially
+bias the distribution towards higher $\beta_{OX}$-values. We tentatively confirm
+the result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have
+higher \nhx. Specifically, for bursts with measured redshift either from the
+afterglow or the host galaxy, we find the following: For bursts with $\beta_{OX}
+\geq 0.5$ we find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$
+we find \nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is
+contained within the error intervals. A 2-sided KS test fails to reject the null
+that they are drawn from the same distribution with $p = 0.43$, meaning there is
+no strong evidence for a discrepancy.
 
@@
 redshift as we either need a detected afterglow to obtain a spectrum or to 
-locate the host galaxy and determine the redshift from there. It is more difficuly to
+locate the host galaxy and determine the redshift from there. It is more difficult to
 correctly associate a galaxy with a burst when there is no detected optical
@@
 Due to the stochastic nature of the \lya-forest, the blue wing of the
-Lyman-$\alpha$ absorption line is randomly superposed with Lyman-$\alpha$
-forest systems, along with strong absorption from \mnii~and \SIiii, making it
+Lyman-$\alpha$ absorption line is randomly superposed with Lyman-$\alpha$ forest
+systems, along with strong absorption from \mnii~and \SIiii, making it
 notoriously difficult to model. Additionally, the red wing has got the ISM
@@
 effects, the generative model for the data that we would use in a
-likelihood-based analysis would be very complicated, we have decided not to
-make formal $\chi$-2 fitting of the hydrogen column densities, but instead use
-a more subjective visual fitting to the absorption profile.  Using an analytic
+likelihood-based analysis would be very complicated, we have decided not to make
+formal $\chi$-2 fitting of the hydrogen column densities, but instead use a more
+subjective visual measurement to the absorption profile.  Using an analytic
 approximation to the absorption profile from \citep{TepperGarcia2006}, we
@@
 observed, normalised spectrum. By tuning the value of the hydrogen column
-density until the synthetic absorption line matches the spectrum, we can
-thereby infer the actual column density of the GRB sight line in a manual way.
-In a similar fashion, the errors on the hydrogen column can be estimated. We
-show the results of this procedure for all bursts where possible in Fig.
+density until the synthetic absorption line matches the spectrum, we can thereby
+infer the actual column density of the GRB sight line in a manual way.
+Similarly, the uncertainty on the hydrogen column can be estimated by adjusting
+the error, until the confidence bounds contain the continuum variation. We show
+the results of this procedure for all bursts where possible in Fig.
 \ref{fig:HI1} and the inferred hydrogen column densities in Table~\ref{tab:HI}
@@
 percentiles of the two distributions where the sample presented here has
-$21.8_{0.8}^{0.3}$ and the rest of the litterature values has
+$21.8_{0.8}^{0.3}$ and the rest of the literature values has
 $21.5_{1.5}^{0.4}$. We see that the two distribution has a large degree of
@@
 \citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al. in
-prep). %\citep{Tanvir2017}.
+prep). A number of additional companion papers are also planned,
+investigating the detailed properties of the sample presented here, including
+equivalent width distributions (de Ugarte Postigo et al., in preparation),
+metallicities and kinematics (Th{\"o}ne et al., in preparation), high ionization
+lines (Heintz et al., in preparation), molecular lines (Bolmer et al., in
+preparation), fine-structure lines (Vreeswijk et al., in preparation), composite
+GRB afterglow spectrum (Selsing et al., in preparation).
 
@@
 GRB sightlines as probes is far from fully harvested. The sample of sightlines
-probed by our spectra is not representative for all GRB sightlines as we have
+probed by our spectra are not representative for all GRB sightlines as we have
 shown and consistent with earlier findings from samples based on low-resolution
@@
 minutes after the burst with a 10-m class telescope. Observations of such
-sightlines with X-shooter-like spectrographs on th enext generation of 20--40-m
+sightlines with X-shooter-like spectrographs on the next generation of 20--40-m
 telescopes is likely to be very rewarding.
@@
 \begin{acknowledgements}
-	*** We should add an ode to Neil Gehrels here ***.
-	JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
-	Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
-	acknowledges support by the European Commission under the Marie Curie
-	Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
-	European Commission under the Marie Curie Career Integration Grant programme
-	(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
-	{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
+	*** We should add an acknowledgment to Neil Gehrels here ***. JPUF, BMJ and DX acknowledge
+support from the ERC-StG grant EGGS-278202.  The Dark Cosmology Centre is
+funded by the Danish National Research Foundation.  TK acknowledges support by
+the European Commission under the Marie Curie Intra-European Fellowship
+Programme in FP7.  AdUP acknowledges support by the European Commission under
+the Marie Curie Career Integration Grant programme (FP7-PEOPLE-2012-CIG
+322307).  This work made use of data supplied by the UK {\it Swift} Science
+Data Centre at the University of Leicester. JJ acknowledges support from NOVA
+and NWO-FAPESP grant for advanced instrumentation in astronomy. Finally, we
 acknowledge expert support from the ESO staff at the Paranal and La Silla
@@
 
-\subsection{GRB~140114A (z = $\sim2.8$?)}
+\subsection{GRB~140114A (z = 3.0)}
+
+This spectrum has previously been published in \citet{Kruhler2015}. This is a
+late time observation, taken long after the GRB had faded. Despite a long
+integration time of 5.4 ks, no clear features stand out to clearly secure a
+redshift measurement. By heavily binning the spectrum, a faint trace is visible
+down to 4850 \AA, which if interpreted at the onset of \lya signifies z $\sim$
+3. We adopt the redshift inferred in \citet{Kruhler2015}.
 
@@
 
+The data presented here also formed the basis of GCN
+\#15831\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15831.gcn3}} and is
+additionally published in \citet{Kruhler2015}. Starting 5.8 after the
+\swift~trigger alert, $2\times600$ s spectroscopic integration contains a high
+S/N GRB afterglow continuum across the entirety of X-shooter. Imprinted on the
+afterglow continuum are absorption features, which we identify as metal
+resonance lines from \civ, \alii, \aliii, \feii, \mgii, and \mgi. These lines
+are likely formed by metals in the GRB host, which we measure to be at $z =
+1.208$.
+
 \subsection{GRB~140301A (z = 1.416)}
 
-\subsection{GRB~140311A (z = 4.95)}
+These data formed the basis for GCN
+\#15900\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15900.gcn3}} and is
+additionally published in \citet{Kruhler2015}. Spectroscopic follow-up began 9
+hours after the BAT trigger and lasted for $12\times600$ s. A low S/N, spatially
+extended continuum is visible across the entire spectral coverage of X-shooter.
+The GRB afterglow continuum is visible at moderate S/N on top of the underlying
+host continuum. The host exhibits usual nebular emission from \oii, \hb, \oiii,
+\ha, \nii, and \sii which puts it at $z = 1.416$. Supporting this as the
+redshift of the GRB are absorption features from the \mgii-doublet in the GRB
+afterglow continuum.
+
+\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}
+
+This spectrum has not been published previously. Starting 32.5 hours after the
+GRB trigger onboard \swift, this observation lasted $14\times600$ s for a total
+of 8.4 ks. Some loss occurred during the observations, which reduced integration
+time in the UVB and VIS arm slightly. The GRB afterglow continuum is clearly
+visible in the VIS and NIR arm of X-shooter. The continuum is a very rich
+absorption, with at least the following lines identified: \lyg, \lyb, \lya,
+\SIii, \SIiv, \civ, \alii, \aliii, \feii, \mgii, and \mgi. All of these lines
+are a $z = 4.954$, which we take as the redshift of the GRB. This spectrum is
+taken under ESO programme ID: 092.D-0633(E) (PI: Greiner).
 
 \subsection{GRB~140430A (z = 1.601)}
+
+%GCN Circular #16194 
+%T. Kruehler (ESO Santiago), D. Malesani (DARK/NBI), A. de Ugarte Postigo (IAA/CSIC and DARK/NBI), A. Melandri (INAF/OABr), J. P. U. Fynbo (DARK/NBI), report on behalf of a larger collaboration: 
+%We observed the optical afterglow of GRB 140430A (Siegel et al., GCN 16190) with the ESO VLT equipped with X-shooter.
+Observations started at 23:09 UT on 2014 April 30 (2.5 hr after the GRB), still in twilight, and two 10-minute spectra were secured.
+The covered wavelength range is 3000-25000 AA.
+%We detect a clear continuum, and several absorption features, which we interpret as due, among others, to Si II, C IV, Al II, Fe II, and Mg II at a common redshift z = 1.60.
+In the infrared arm, we also detect the [O III] 5008 nebular line in emission at the same redshift.
 

### 2018-01-22 py/betaOX.py

@@ -127,7 +127,7 @@ def main():
     g.y = betaOX[idx_2][~idx_limit]
 
     np.savetxt("betaOX.dat", list(zip(name_s[idx][~idx_limit] ,HI[~idx_limit], betaOX[idx_2][~idx_limit])), fmt = '%s %s %s')
-    exit()
+    # exit()
 
     g = g.plot_joint(pl.scatter, color=color, label="This work")
 

### 2018-01-22 document/XSGRB_sample.tex
Notes on individual bursts

@@
 A separate reduction of this burst has been published in \citet{Kruhler2015},
-but not otherwise. Starting 9 days after the burst trigger, this is host
+but is not used otherwise. Starting 9 days after the burst trigger, this is host
 observation and does not contain any afterglow continuum. In 3.6 ks integration
@@
 
-\subsection{GRB~120815A* (z = 2.358)} 
-Not a part of the statistical sample, thisburst also formed the basis of  GCN
+\subsection{GRB~120815A (z = 2.358)} 
+Not a part of the statistical sample, this burst also formed the basis of  GCN
 \#13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
@@
 primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
-\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, which we as the
-redshift of the GRB. This spectrum is taken under ESO programme ID:
-091.A-0877(A) (PI: Schady).
+\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, as well as those
+from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
+spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).
 
@@
 
-%GCN Circular #16194 
-%T. Kruehler (ESO Santiago), D. Malesani (DARK/NBI), A. de Ugarte Postigo (IAA/CSIC and DARK/NBI), A. Melandri (INAF/OABr), J. P. U. Fynbo (DARK/NBI), report on behalf of a larger collaboration: 
-%We observed the optical afterglow of GRB 140430A (Siegel et al., GCN 16190) with the ESO VLT equipped with X-shooter.
-Observations started at 23:09 UT on 2014 April 30 (2.5 hr after the GRB), still in twilight, and two 10-minute spectra were secured.
-The covered wavelength range is 3000-25000 AA.
-%We detect a clear continuum, and several absorption features, which we interpret as due, among others, to Si II, C IV, Al II, Fe II, and Mg II at a common redshift z = 1.60.
-In the infrared arm, we also detect the [O III] 5008 nebular line in emission at the same redshift.
+A separate reduction of this burst has been published in \citet{Kruhler2015},
+and additionally the spectrum has also been used in GCN
+\#16194\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16194.gcn3}}. Observations
+for this burst began in twilight, 2.5 hours after the BAT trigger and lasted for
+$2\times600$ s. The spectrum contains a moderate S/N GRB afterglow continuum all
+the way through the spectroscopic arms of X-shooter. We identify absorption
+features in the afterglow continuum from \SIii, \civ, \alii, \feii, and \mgii,
+and emission lines from \oii~and \oiii~- all at $z = 1.601$ which is likely the
+redshift of the GRB.
 
 \subsection{GRB~140506A  (z = 0.889)}
-The data presented here has formed the basis of GCN \#
-16214\footnote{\url{http://gcn.gsfc.nasa.gov/other/140506A.gcn3}} and is
-published in \citet{Fynbo2014} and \citet{Heintz2017a}. The observations
-consists of $4\times 600$ sec at 8.8 and 33 hours after the burst. We detect
-absorption features from a wide range of ions, together with molecular
-absorption from CH+, all at $z=0.889$. The optical/near-infrared afterglow
-reveal an unusual steep extinction curve which is found to be caused by dust
-immediate to the burst.
-
-\subsection{GRB~140515A (z = $\sim6.32$?)}
+
+The data presented here has formed the basis of GCN
+\#16217\footnote{\url{https://gcn.gsfc.nasa.gov/gcn3/16217.gcn3}} and is
+published in \citet{Fynbo2014}, \citet{Kruhler2015}, and \citet{Heintz2017a}.
+The observations consists of $4\times 600$ s at 8.8 and 33 hours after the
+burst. We detect absorption features from a wide range of ions, together with
+molecular absorption from CH+, all at $z=0.889$. The optical/near-infrared
+afterglow reveal an unusual steep extinction curve which is found to be caused
+by dust immediate to the burst.
+
+\subsection{GRB~140515A (z = 6.327)}
+
+This spectrum has previously been used in \citet{Melandri2015}, but is not
+published elsewhere. Starting 15.5 hours after the \swift~trigger, $8\times600$
+s spectroscopic integration captures this very high redshift GRB afterglow. The
+Gunn-Peterson trough is visible against the GRB afterglow continuum, starting at
+8900 \AA, which along with absorption lines from the \mgii~doublet securely sets
+the redshift of this GRB at $z = 6.327$. From the red wing of the \lya-profile
+we measure \nh~$=19.0 \pm 0.5$, which is very low compared to the measured
+distribution of $N_{\mathrm{HI}}$.
 
 \subsection{GRB~140614A (z = 4.233)}
+
+This spectrum forms the basis for GCN
+\#16401\footnote{\url{http://gcn.gsfc.nasa.gov/other/16401.gcn3}}, but is not
+published elsewhere. $4\times600$ s spectroscopic integration, starting 3.8
+hours after the BAT trigger catches this GRB afterglow. The continuum is
+detected at moderate S/N in both the VIS and NIR arms of X-shooter, heavily
+affected by absorption features. We identify the the lines belonging to \lya,
+\SIii, \cii, \cii$^*$, \alii, \aliii, \feii, and \mgii~at a consistent redshift
+of $z = 4.233$. From the \lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
 

### 2018-01-23 document/XSGRB_sample.tex
Addressing Kaspers comments and notes on 2014 bursts

@@
 
-\abstract{In this work we present the spectra of all $\gamma$-ray bursts (GRBs)
+\abstract{In this work we present the spectra of all $\gamma$-ray burst (GRB)
 	afterglows that have been promptly observed at the VLT/X-shooter up to
@@
 	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
-	\% of these, covering a redshift range from 0.059 to 7.84. Based on an set of
-	selection criteria that are unbiased with regards to intrinsic properties of the
+	\% of these, covering a redshift range from 0.059 to 7.84. Based on a set of observational selection criteria that are unbiased with regards to intrinsic properties of the
 	GRBs, the follow-up effort has been guided by the \textit{Swift}-telescope to
@@
 	sample. For all bursts (39) for which it is possible, we provide a measurement
-	of the hydrogen column density, increasing the total number of published HI
+	of the neutral hydrogen column density, increasing the total number of published HI
 	column density measurements by $\sim$ 50 per cent. We constrain the fraction of
@@
 
-Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
+$\gamma$-ray bursts (GRBs) are interesting as they provide the opportunity to
 derive new constraints on a very wide range of topics in astrophysics. Examples
@@
 
-The {\it Swift} satellite \citep{Gehrels2009}, which was launched in 2004, has
-made it possible to harvest much of the great potential in using GRBs as probes
-of galaxies in the intergalactic medium, which was already hinted at by results
+The \textit{Niel Gehrels Swift Observatory} ({\it Swift}) satellite
+\citep{Gehrels2004, Gehrels2009}, which was launched in 2004, has made it
+possible to harvest much of the great potential in using GRBs as probes of
+galaxies in the intergalactic medium, which was already hinted at by results
 from earlier missions \citep[e.g.,][]{Ricker2004}. A crucial aspect of the great
-success of the {\it Swift} mission has been the extensive ground based
-follow-up observations of the afterglows and of the host galaxies of the GRBs.
-This work has involved a large community of researchers both inside and outside
-the official {\it Swift} team. This fruitful collaboration has been greatly
+success of the {\it Swift} mission has been the extensive ground based follow-up
+observations of the afterglows and of the host galaxies of the GRBs. This work
+has involved a large community of researchers both inside and outside the
+official {\it Swift} team. This fruitful collaboration has been greatly
 facilitated by the open data access policy of the {\it Swift} mission.
@@
 based on BAT alone. We therefore additionally require an X-ray position from the
-X-ray Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
-\citep{Barthelmy2000} within 10 hours and to account for observing constraints
+X-Ray Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
+\citep{Barthelmy2000} within 12 hours and to account for observing constraints
 on \textit{Swift} that XRT began observations within 10 minutes. The additional
@@
 with the sun below $-12$ degrees. This secures time for the spectroscopic
-observations to completed. Since the GRB population is isotropically distributed
+observations to be completed. Since the GRB population is isotropically distributed
 on the sky as seen from Earth and because the GRB properties do not depend on
-position on the sky \citep{Briggs1996}, this cut does not influence our ability
+position on the sky \citep{Meegan1992, Briggs1996}, this cut does not influence our ability
 to fairly sample the underlying GRB population. The same arguments apply to
@@
 
-The use of RRM is unique in the sense that it allowes for the detailed study of
+The use of RRM is unique in the sense that it allows for the detailed study of
 the temporal variability of the GRB afterglow due to effect of the GRB itself on
@@
 The first GRB followed up is GRB~090313, observed on the 15th of March, 2009,
-during the commissioning of X-shooter UT3. The bursts observed during the
+during the commissioning of X-shooter on UT3. The bursts observed during the
 commissioning or science verification (GRB~090313, GRB~090530, GRB~090809,
@@
 %The slit width sets the width of the atmospheric sky lines and determines the amount of light lost due to the wavelength-dependent seeing PSF extending outside the coverage of the slit, where the width of sky-lines is always set by the slit width whereas both the delivered resolution and the slitloss changes for the better as the seeing PSF drops below the slit width.
-For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a 0\farc9 slit only 76.1 percent of the light will enter the slit, meaning that for almost all observations a slitloss correction is required.
+For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a 0\farc9 slit only 76.1 per cent of the light will enter the slit, meaning that for almost all observations a slitloss correction is required.
 %We describe how slitlosses were corrected for in Sect. \ref{postproc}.
@@
 celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from
-central zone of avoidance due to high galactic extinction cutoff (marked in
+central zone of avoidance due to high Galactic extinction cutoff (marked in
 white), the GRB positions have an apparent isotropic distribution except in the
@@
 some bursts, the initial sky-estimate has been made from a limited number of
-pixels in the spatial direction. By doing an initial subtracting the sky on the
-un-rectified image we ensure that bulk of the sky background is not
-redistributed by the rectification process.
+pixels in the spatial direction. By doing an initial subtraction of the sky
+background on the un-rectified image we ensure that the bulk of the sky
+background is not redistributed by the rectification process.
 
@@
 0.06 nm/pixel with a 0.21 "/pixel in the NIR arm.  Because the tabulated
-resolution is a lower limit, by choosing a sampling of 0.02 nm/pixel, we ensure
+resolution is a lower limit, we choose a sampling of 0.02 nm/pixel to ensure
 that the bluest part of neither of the arms have a sampling lower than the
@@
 been reduced as a single observation and then subsequently combined to form a
-single image. Because this strategy is employed, we can reject outliers in the
+single image. We employ this strategy so that we can reject outliers in the
 stack and weight by an averaged measure of the inverse variance of the
@@
 
-X-shooter, being installed at the VLT Cassegrain focus is  prone to flexures
-during operations. The flexures modify the projection of the slit on the
-detector with respect to the one obtained in daytime calibration. This requires
-a modification of the wavelength solution in order to process correctly the
-night-time data. Part of this correction is performed by the pipeline using the
-frames taken during X-shooter Active Flexure Compensation procedure
-\footnote{X-shooter User Manual available at
+Since X-shooter is installed at the VLT Cassegrain focus it is prone to
+flexures. The flexures modify the projection of the slit on the detector with
+respect to the one obtained in daytime calibration. This requires a modification
+of the wavelength solution in order to correctly process the night-time data.
+Part of this correction is performed by the pipeline using the frames taken
+during the X-shooter Active Flexure Compensation procedure \footnote{X-shooter
+	User Manual available at
 	https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}. The
@@
 correct seeing PSF with which to convolve the synthetic sky an initial
-refinement of the wavelength solution have been obtained by cross-correlating
+refinement of the wavelength solution has been obtained by cross-correlating
 the observed sky with an unconvolved synthetic sky. This preliminary wavelength
@@
 spatial direction as a function of wavelength. For a large fraction of the
-bursts observed, using a single aperture for the spectral extraction is
+observed bursts, using a single aperture for the spectral extraction is
 inadequate due to the large amount of background that would then enter the slit.
@@
 in App. \ref{voigt}. Since additionally, the host galaxy could also have a
-contribution the image profile, this choice allows for the required freedom if
+contribution to the image profile, this choice allows for the required freedom if
 additional flux is in the wings of the profile.
@@
 low-order polynomial to the fit parameters as a function of wavelength, which
-allows us to evaluate the spectral PSF at all wavelengths and this way
+allows us to evaluate the spectral PSF at all wavelengths and in this way
 accurately model the entire spectral PSF.
@@
 dispersion direction, we can  employ the optimal extraction algorithm
-\cite{Horne1986}, which weights the extraction aperture by the spectral profile,
+\citep{Horne1986}, which weights the extraction aperture by the spectral profile,
 or alternatively sum all pixels within 1 FWHM of the modeled profile. Where
@@
 
-The spectra are corrected for Galactic extinction using the E(B-V) value from
+The spectra are corrected for Galactic extinction using the $E(B-V)$ value from
 the dust maps of \citet{Schlegel1998} with the update in
@@
 of ways. We employ a prioritized list of methods here, depending on the
-availability of the different method. Since the observation are often taken at
+availability of the different methods. Since the observation are often taken at
 odd times under varying conditions, this prioritized list ensures that we are
@@
 reason have not been observed a telluric standard, we rely on the synthetic sky
-model by \citep{Noll2012, Jones2013} for which we generate a synthetic
-transmission spectrum,  where the ambient parameters for the
+model by \citet{Noll2012} and \citet{Jones2013} for which we generate a
+synthetic transmission spectrum,  where the ambient parameters for the
 observations have been used.
@@
 The afterglow spectra described in this paper are obtained in
-Target-of-Opportunity (override) mode. In most cases there is therefore little
+Target-of-Opportunity (override) mode. In most cases there are therefore little
 possibility to tweak slit widths to the seeing at the time of observations (i.e.
@@
 redshifts, this is difficult to do at poorer signal to noise ratios (the
-majority of spectra in our sample). A broad starting value for the expect
+majority of spectra in our sample). A broad starting value for the expected
 resolution will help fitting of these spectra, and can be important in upper
@@
 \footnote{\url{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}}. This
-release includes the spectra of GRBs outside the statistical sample and includes
-both prompt afterglow observations as well as late time observations of the
-associated hosts and represents \textit{all} afterglow spectra of GRBs carried
-out by VLT/X-shooter from the commissioning of the instrument, \startdate, and
-until the end of the last period of the program 098.A-0055, \termdate~and thus
-constitutes 8 years of GRB afterglow observations with VLT/X-shooter. An
-overview of all the spectra provided here are given in Table.
-\ref{tab:sample_overview}. For each burst, each individual observation is provided
-in a separate reduction and in cases where observations have been repeated for
-an increased S/N or to follow the temporal evolution, a combined spectrum is
-also provided. No attempt has been made to join the spectroscopic arms, so for
-each observation, three spectra are provided in separate files.
+release includes for completeness the spectra of GRBs outside the statistical
+sample and includes both prompt afterglow observations as well as late time
+observations of the associated hosts and represents \textit{all} afterglow
+spectra of GRBs carried out by the X-shooter spectrograph since the
+commissioning of the instrument, \startdate, and until the end of the last
+period of the program 098.A-0055, \termdate~and thus constitutes 8 years of GRB
+afterglow observations with X-shooter. An overview of all the spectra provided
+here are given in Table. \ref{tab:sample_overview}. For each burst, each
+individual observation is provided in a separate reduction and in cases where
+observations have been repeated for an increased S/N or to follow the temporal
+evolution, a combined spectrum is also provided. No attempt has been made to
+join the spectroscopic arms, so for each observation, three spectra are provided
+in separate files.
 
@@
 	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion and drifts in the wavelength solution ($\mathrm{\AA}$).
-	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
-	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
+	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
+	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
 	\item QUAL - Bad pixel map, where a value different from zero indicates a bad pixel.
-	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
+	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
 	\item CONTINUUM\_ERR - Relative error on the continuum estimate 
@@
 
-In this section, we describe the successfulness of the follow-up effort and the
+In this section, we describe the efficiency of the follow-up effort and the
 characteristics of the observed bursts. We also assess the degree to which the
 obtained sample is representative for the full \textit{Swift} sample. An
-important note is that here we include \textit{all} GRB after \startdate, that
-have been observed with X-shooter, while only a subset of these constitute our
+important note is that here we include \textit{all} GRBs after \startdate, that
+have been observed with X-shooter, while only a subset of these constitutes our
 \textit{statistical} sample. The statistical sample is based on the selection
@@
 are not discussed as part of the investigation of the statistical properties of
-the GRB population. A prime example of a burst outside the statistical sample is
-bright INTEGRAL burst GRB~161023A, that turned to be a rich spectrum with at
+the GRB population. A prime example of a burst outside the statistical sample is the
+bright \textit{INTEGRAL} burst GRB~161023A, that turned to be a rich spectrum with at
 least 11 intervening absorption systems.
@@
 \begin{figure}
-	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}} \caption{afterglow
+	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}} \caption{Afterglow
 	magnitude at the start of observation as a function of follow-up delay. The
@@
 	Sect. \ref{samplecrit} (green), and the subset that has been observed as part of
-	the statistical sample (red). The left plot shows the fluence in in range 15-150
+	the statistical sample (red). The left plot shows the fluence in 15-150
 	keV as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
@@
 	samples are drawn from the same underlying distribution. The arrows represent
-	the comparison with each of the p-values are listed next to. Only in the HI
-	column density distribution is there mild evidence against the null, but the
+	the comparison direction with each of the p-values listed next to. Only in the
+	HI column density distribution is there mild evidence against the null, but the
 	discrepancy is mainly driven by a relatively larger fractional contribution
-	from low-column hosts.}
-	\label{fig:p_values}
+	from low-column hosts.} \label{fig:p_values}
 \end{figure}
@@
 
-Out of the 158 bursts in the statistical sample, 32($\sim$ 20 percent) have not
+Out of the 158 bursts in the statistical sample, 32($\sim$ 20 per cent) have not
 been observed due to reasons unrelated to the GRB or afterglow properties. The
@@
 %\ref{badbursts}). By trimming away the 33 bursts that can be rejected due to
-%telescope availability, the redshift completeness increases to 83 percent.
+%telescope availability, the redshift completeness increases to 83 per cent.
 %
@@
 Because the redshift completeness of our statistical sample is 60 per cent,
-making inference of the \textit{true} redshift distribution of GRBs on this
-sample is difficult. The biases introduced in the redshift distribution by the
-increased ability to secure redshift of optically brighter bursts has been
-investigated by \citet{Turpin2016} which find that we systematically find
-redshift for longer GRBs. Additionally only the brightest GRBs are seen above
-redshift $z \gtrsim 1$ as shown in Fig. \ref{fig:z}. Additionally, as described
-in detail in \citet{Perley2016b}, bursts for which the redshift is measured from
-the afterglow are systematically found for host galaxies with a lower luminosity
-than burst for which the redshift is measured from the host galaxy. Only a few
-GRBs hosted in galaxies more massive than $10^{10}~\mathrm{M}_\odot$ have the
-redshift measured based on the afterglow continuum.
+making inference of the \textit{true} redshift distribution of GRBs based on
+this sample is difficult. The biases introduced in the measured redshift
+distribution by our increased ability to secure redshift of optically brighter
+bursts has been investigated by \citet{Turpin2016} which find that we
+systematically find redshift for longer GRBs. Additionally only the brightest
+GRBs are seen above redshift $z \gtrsim 1$ as shown in Fig. \ref{fig:z}. As
+described in detail in \citet{Perley2016b}, bursts for which the redshift is
+measured from the afterglow are systematically found for host galaxies with a
+lower luminosity than burst for which the redshift is measured from the host
+galaxy. Only a few GRBs hosted in galaxies, with stellar masses more than
+$10^{10}~\mathrm{M}_\odot$ have the redshift measured based on the afterglow
+continuum.
 
@@
 	the highest spectroscopically confirmed redshifts, the latter two shown at
-	arbitrary $L_{\mathrm{BAT}}$. The blue solid line represents so-called 1 s BAT
-	sensitivity limit described in the text. The estimated epoch of reionization is
-	shown by the black dot-dashed line, with the uncertainty shown as the black
-	dashed lines, from the most recent measurement by \citet{Planck2015}.}
-\label{fig:z}
+	arbitrary $L_{\mathrm{BAT}}$. The blue solid line represents the so-called 1 s
+	BAT sensitivity limit described in the text. The estimated epoch of
+	reionization is shown by the black dot-dashed line, with the uncertainty shown
+	as the black dashed lines, from the most recent measurement by
+	\citet{Planck2015}.} \label{fig:z}
 \end{figure*}
@@
 	extinction relative to X-ray against X-ray derived hydrogen column densities.
-	In red is shown the sample presented in \citet{Fynbo2009} where is lines
-	indicate the kernel density estimate of the distribution where limits have been
-	replaced with values. We note that contrary to what is presented in
-	\citet{Fynbo2009}, we exclude all bursts for which we do not have a redshift.
-	Darker colors represented higher density of points. The corresponding marginal
-	distributions are shown. In blue is the points for the bursts presented here
-	along with the marginal histograms. Limits on $\beta_{OX}$ is shown by
-	downwards facing arrows.} \label{fig:betaOX}
+	In red is shown the sample presented in \citet{Fynbo2009} where the lines
+	indicate the kernel density estimate of the distribution. For the kernel
+	density estimate, the limits have been replaced with values. We note that
+	contrary to what is presented in \citet{Fynbo2009}, we exclude all bursts for
+	which we do not have a redshift. Darker colors represented higher density of
+	points. The corresponding marginal distributions are shown. In blue is the
+	points for the bursts presented here along with the marginal histograms. Limits
+	on $\beta_{OX}$ is shown by downwards facing arrows.} \label{fig:betaOX}
 \end{figure}
@@
 It is clear, however, that the fraction of dark bursts for which we have
-measured the redshift is lower than the intrinsic dark burst fraction. To measure the
-redshift as we either need a detected afterglow to obtain a spectrum or to 
-locate the host galaxy and determine the redshift from there. It is more difficult to
-correctly associate a galaxy with a burst when there is no detected optical
-afterglow and hence a correct redshift measurement is more difficult to make,
-see \citet{Jakobsson2005, Levesque2010} and \citet{Perley2017} for a recent
-example.
+measured the redshift is lower than the intrinsic dark burst fraction. To
+measure the redshift we either need a detected afterglow to obtain a spectrum or
+to locate the host galaxy and determine the redshift from there. It is more
+difficult to correctly associate a galaxy with a burst when there is no detected
+optical afterglow and hence a correct redshift measurement is more difficult to
+make, see \citet{Jakobsson2005, Levesque2010} and \citet{Perley2017} for a
+recent example.
 
@@
 Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html}, we can see how the
-presence of an optical afterglow affects the follow-up statistics. 50.5 percent
+presence of an optical afterglow affects the follow-up statistics. 50.5 per cent
 of all \textit{Swift}-detected bursts do not have an detected optical afterglow,
@@
 available, so is a lower limit. For the bursts entering our statistical sample,
-the number of bursts without a detected optical afterglow is $\sim$ 27 percent,
+the number of bursts without a detected optical afterglow is $\sim$ 27 per cent,
 close to the upper limit on the fraction of dark bursts found in a complete
@@
 the bursts for which follow-up has actually been attempted, this number is
-$\sim$ 18 percent, confirming that we are indeed biased against bursts without a
+$\sim$ 18 per cent, confirming that we are indeed biased against bursts without a
 detected optical afterglow in the spectroscopic sample.
@@
 effects, the generative model for the data that we would use in a
-likelihood-based analysis would be very complicated, we have decided not to make
-formal $\chi$-2 fitting of the hydrogen column densities, but instead use a more
-subjective visual measurement to the absorption profile.  Using an analytic
-approximation to the absorption profile from \citep{TepperGarcia2006}, we
-overplot a synthetic absorption line with a specified column density on our
-observed, normalised spectrum. By tuning the value of the hydrogen column
-density until the synthetic absorption line matches the spectrum, we can thereby
-infer the actual column density of the GRB sight line in a manual way.
-Similarly, the uncertainty on the hydrogen column can be estimated by adjusting
-the error, until the confidence bounds contain the continuum variation. We show
-the results of this procedure for all bursts where possible in Fig.
-\ref{fig:HI1} and the inferred hydrogen column densities in Table~\ref{tab:HI}
+likelihood-based analysis would be very complicated. We have therefore decided
+not to make formal $\chi^2$ fitting of the hydrogen column densities, but
+instead use a more subjective visual measurement to the absorption profile. 
+Using an analytic approximation to the absorption profile from
+\citet{TepperGarcia2006}, we overplot a synthetic absorption line with a
+specified column density on our observed, normalised spectrum. By tuning the
+value of the hydrogen column density until the synthetic absorption line matches
+the spectrum, we can thereby infer the actual column density of the GRB sight
+line in a manual way. Similarly, the uncertainty on the hydrogen column can be
+estimated by adjusting the error, until the confidence bounds contain the
+continuum variation. We show the results of this procedure for all bursts where
+possible in Fig. \ref{fig:HI1} and the inferred hydrogen column densities in
+Table~\ref{tab:HI}.
 
@@
 
-In the compilation of HI measurements towards GRBs in \citet{Tanvir2017} there
-are 81 published HI values. We here provide 39 new HI column density
-measurements, i.e. an increase of the number of optically derived hydrogen
-column densities of $\sim$ 50 percent. We compare the median, the 14th, and 86th
-percentiles of the two distributions where the sample presented here has
-$21.8_{0.8}^{0.3}$ and the rest of the literature values has
-$21.5_{1.5}^{0.4}$. We see that the two distribution has a large degree of
+In the compilation of HI measurements towards GRBs in Tanvir et al. (submitted) %\citet{Tanvir2017} 
+there are 81 published HI values. We here provide 39 new HI
+column density measurements, i.e. an increase of the number of optically derived
+hydrogen column densities of $\sim$ 50 per cent. We compare the median, the 14th,
+and 86th percentiles of the two distributions where the sample presented here
+has $21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
+$21.5_{-1.5}^{+0.4}$. We see that the two distribution has a large degree of
 overlap due to the large width of the distributions, but we find a slightly
@@
 
-We show the
-comparison between the two samples in Fig. \ref{fig:NH_dist}. We additionally
-show the column-distribution for 12081 quasar DLAs from \citet{Noterdaeme2012b}.
-The fact that GRBs systematically are located behind the highest \nh,
-previously noted by \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is stricking. The
-reason for this is that quasars sampling of sightlines through galaxies is
-cross-section selected, whereas GRB sightlines probe more central, dense regions
-in their hosts.
+We show the comparison between the two samples in Fig. \ref{fig:NH_dist}. We
+additionally show the column density-distribution for the 12081 quasar absorbers
+with \nh~$> 20$ from \citet{Noterdaeme2012b}. The fact that GRBs systematically
+are located behind the highest \nh, previously noted by
+\citep[e.g.,][]{Prochaska2007, Fynbo2009}, is stricking. The reason for this is
+that quasar sample sight-lines through galaxies that are cross-section selected,
+whereas GRB sight-lines probe the dense, star-forming regions in their hosts.
 
@@
 	\centering \includegraphics[width=9cm]{figures/NH_dist.pdf}
-\caption{Distributions of DLA hydrogen column densities for DLAs found in
+\caption{Distributions of hydrogen column densities for absorbers found in
 	quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in
-	green is the kernel density estimate is the similar distribution, only for DLAs
-	in GRB sightlines. Values are taken from the compilation in Tanvir2017 et al.
-	(in prep), along with the new values presented in this sample. The marginal
-	distributions for the two distribution are also shown along the left side and
-	on the top, where the different environments probed are clearly visible in the
-	hydrogen column densities, as previously also noted in \citet{Fynbo2009}}
-\label{fig:NH_dist}
+	green is the kernel density estimate is the similar distribution, only for
+	absorbers in GRB sightlines. Values are taken from the compilation in Tanvir et
+	al. (submitted), along with the new values presented in this sample. The
+	marginal distributions for the two distribution are also shown along the left
+	side and on the top, where the different environments probed are clearly
+	visible in the hydrogen column densities, as previously also noted in
+	\citet{Fynbo2009}.} \label{fig:NH_dist}
 \end{figure}
@@
 X-shooter, but over the years the project continued in open time and the team
-was opened to include any researchers interested in contribution to this effort.
+was opened to include any researchers interested in contributing to this effort.
 
@@
 galaxies \citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
-\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al. in
-prep). A number of additional companion papers are also planned,
+\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al., submitted). A number of additional companion papers are also planned,
 investigating the detailed properties of the sample presented here, including
@@
 \begin{acknowledgements}
-	*** We should add an acknowledgment to Neil Gehrels here ***. JPUF, BMJ and DX acknowledge
-support from the ERC-StG grant EGGS-278202.  The Dark Cosmology Centre is
-funded by the Danish National Research Foundation.  TK acknowledges support by
-the European Commission under the Marie Curie Intra-European Fellowship
-Programme in FP7.  AdUP acknowledges support by the European Commission under
-the Marie Curie Career Integration Grant programme (FP7-PEOPLE-2012-CIG
-322307).  This work made use of data supplied by the UK {\it Swift} Science
-Data Centre at the University of Leicester. JJ acknowledges support from NOVA
-and NWO-FAPESP grant for advanced instrumentation in astronomy. Finally, we
-acknowledge expert support from the ESO staff at the Paranal and La Silla
-observatories in obtaining these target of opportunity data.
+	*** We should add an acknowledgment to Neil Gehrels here ***. JPUF, BMJ and DX
+acknowledge support from the ERC-StG grant EGGS-278202.  The Dark Cosmology
+Centre is funded by the Danish National Research Foundation.  TK acknowledges
+support by the European Commission under the Marie Curie Intra-European
+Fellowship Programme in FP7.  AdUP acknowledges support by the European
+Commission under the Marie Curie Career Integration Grant programme
+(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
+{\it Swift} Science Data Centre at the University of Leicester. JJ acknowledges
+support from NOVA and NWO-FAPESP grant for advanced instrumentation in
+astronomy. KEH and PJ acknowledge support by a Project Grant (162948--051) from
+The Icelandic Research Fund. Finally, we acknowledge expert support from the
+ESO staff at the Paranal and La Silla observatories in obtaining these target
+of opportunity data.
 
@@
 published elsewhere. $4\times600$ s spectroscopic integration, starting 3.8
-hours after the BAT trigger catches this GRB afterglow. The continuum is
-detected at moderate S/N in both the VIS and NIR arms of X-shooter, heavily
-affected by absorption features. We identify the the lines belonging to \lya,
-\SIii, \cii, \cii$^*$, \alii, \aliii, \feii, and \mgii~at a consistent redshift
-of $z = 4.233$. From the \lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
+hours after the BAT trigger catches this GRB afterglow, which turns out to be at
+very high redshift. The continuum is detected at moderate S/N in both the VIS
+and NIR arms of X-shooter, heavily affected by absorption features. We identify
+the the lines belonging to \lya, \SIii, \cii, \cii$^*$, \alii, \aliii, \feii,
+and \mgii~at a consistent redshift of $z = 4.233$. From the shape of the
+\lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
 
@@
 
+The characteristics of this short (T90 (15-350 keV ) = is $0.13 \pm 0.04$ s
+\citep{Lien2016}) GRB spectrum has been published in GCN
+\#16437\footnote{\url{http://gcn.gsfc.nasa.gov/other/16437.gcn3}}, but does not
+appear in the refereed literature. Spectroscopic observations began with
+X-shooter only 34 minutes after the BAT trigger and lasted for $2\times600$ s.
+In the spectrum, we detect continuum across all three arms of X-shooter. In the
+UVB arm, the continuum is only visible after heavily binning the spectrum in the
+dispersion direction. It is unclear how much of the continuum is from the host
+galaxy and how much is from the potential GRB afterglow. Superposed on the
+continuum, emission lines from \oii, \hb, \oiii, \ha~are all detected at $z =
+0.959$. We use this as the likely redshift of the GRB.
+
 \subsection{GRB~141028A (z = 2.332)}
 
+This spectrum has already been used to inform GCN
+\#16983\footnote{\url{http://gcn.gsfc.nasa.gov/other/16983.gcn3}}, but is not
+published elsewhere. Starting 15.4 hours after the \swift~trigger and lasting
+for a total of 2.4 ks, these observation nicely captures a bright GRB afterglow.
+Very rich in absorption, the continuum is detected across the entire spectral
+coverage of X-shooter, except the blue half of the UVB arm, where the
+\lya-forest absorbs the continuum. The redshift of the host of the GRB is
+measured based in the detection of features from \lya, \SIii, \civ, \cii, \feii,
+and \mgii - all at a consistent $z = 2.332$. From the \lya~trough we infer
+\nh~$=20.6 \pm 0.15$. Two intervening system at $z = 1.823$ and $z = 2.09$ are
+also found in the spectrum based on the detection of \civ.
+
 \subsection{GRB~141031A  (z = na)}
 
+This spectrum is an attempt at a late-time host redshift measurement for
+GRB~141031A. The spectrum is taken long after the burst has faded and consists
+of $4\times600$ s spectroscopic integration. In the 2D spectrum are two sources,
+which are both offset from the attempted host and are thus likely foreground
+objects. The spectrum does not contain anything that can be used to measure a
+redshift from.
+
 \subsection{GRB~141109A (z = 2.993)}
+
+The gross content of the spectrum has been issued in GCN
+\#17040\footnote{\url{http://gcn.gsfc.nasa.gov/other/17040.gcn3}}, but it is not
+used otherwise. Starting only 1.9 hours after the BAT trigger, these
+$4\times600$ s spectra contains a high S/N GRB afterglow. The afterglow
+continuum is readily detected all the way down to 3700 \AA, although the bluest
+part is affected by the \lya-forest. A broad absorption trough from neutral
+hydrogen is clearly visible, with additional, narrower absorption features
+identified as due to \SIii, \SIii$^*$, \cii, \cii$^*$, \SIiv, \civ, \feii,
+\feii$^*$, \oi$^*$, and \nii$^*$. These are all detected at a consistent $z =
+2.993$, which due to the detection of the locally exited fine-structure lines
+securely sets this as the redshift of the GRB. Additionally, two
+\mgii~absorption systems are detected at $z = 1.67$ and $z=2.5$. From the broad
+\lya-absorption, we measure a neutral hydrogen column density of \nh~$=22.1 \pm
+0.1$.
 
@@
 
-\subsection{GRB~161014A (z =2.823)}
-The data presented here also formed the basis of GCN \#
-20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}}, but is not
-published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8 ks of
-integration time captures the afterglow continuum across all three
+\subsection{GRB~161014A (z =2.823)} The data presented here also formed the
+basis of GCN \# 20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}},
+but is not published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8
+ks of integration time captures the afterglow continuum across all three
 spectroscopic arms. A broad absorption trough due to Lyman alpha is visible,
 along with metal absorption features from \mgii, \SIii, \cii, \civ, \alii,
-\aliii, and	\feii, all at $z =2.823$. Similar to GRB~140506 \citep{Fynbo2014}, a
-break in the continuum shape is detected bluewards of 600 nm, possible
-signifying some anomalous form of extinction.
+\aliii, and	\feii, all at $z =2.823$. Similar to GRB~140506 \citep{Fynbo2014,
+	Heintz2017b},  a break is tentatively detected in the continuum shape is
+detected bluewards of 600 nm, possible signifying some anomalous form of
+extinction.
 

### 2018-01-23 py/Sample_completeness.py

@@ -80,9 +80,9 @@ def main():
 
 
     # BAT fluence
-    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Full Swift sample")
-    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Complete sample")
-    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Observed sample")
+    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 2, "linestyle": "dashed"}, label="Full Swift sample")
+    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 2, "linestyle": "dashed"}, label="Complete sample")
+    sns.distplot(np.log10(1e-7*BAT_o), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 2, "linestyle": "dashed"}, label="Observed sample")
     print(len(BAT_c[~np.isnan(BAT_c)]))
     l, m, h = np.percentile(np.log10(1e-7*BAT_c), [16, 50, 84])
     print(m, m - l, h - m)
@@ -101,9 +101,9 @@ def main():
 
 
     # XRT flux
-    sns.distplot(np.log10(1e-11*XRT_s), ax=ax2, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 3, "linestyle": "dashed"}, label="Full Swift sample")
[…]

### 2018-01-23 py/betaOX.py

@@ -163,7 +163,7 @@ def main():
     ax = pl.gca()
     ax.axhline(0.5, color="black", linestyle="dashed", alpha=0.5)
     ax.annotate(r"$\beta_{OX} = 0.5$", (19.6, 0.45))
-    g.set_axis_labels(r"log(N$_H$) [cm$^2$]", r"Darkness [$\beta_{OX}$]")
+    g.set_axis_labels(r"$\log(N_{\mathrm{HI, X}}/\mathrm{cm}^{-2})$", r"Darkness [$\beta_{OX}$]")
     pl.tight_layout()
 
     # print(stats.ks_2samp(betaOX[idx_2], f_betaOX[f_idx_2]))
@@ -189,7 +189,7 @@ def main():
     print(len(HI[betaOX[idx_2] < 0.5])/(len(HI[betaOX[idx_2] < 0.5]) + len(HI[betaOX[idx_2] >= 0.5])))
     print(stats.ks_2samp(HI[betaOX[idx_2] < 0.5], HI[betaOX[idx_2] >= 0.5]))
     print(len(HI))
-    exit()
+    # exit()
     # print(len(f_HI))
     # l, m, h = np.percentile(f_HI, [16, 50, 84])
     # print(m, m - l, h - m)

### 2018-01-25 document/XSGRB_sample.tex
Comments from Giorgos and Jens + notes on individual bursts

@@
 
+\usepackage{float}
+
 \begin{document}
 
-\title{XSGRB I - The X-shooter GRB afterglow sample
+\title{XS-GRB I - The X-shooter GRB afterglow sample
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
@@
 	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
-	\% of these, covering a redshift range from 0.059 to 7.84. Based on a set of observational selection criteria that are unbiased with regards to intrinsic properties of the
-	GRBs, the follow-up effort has been guided by the \textit{Swift}-telescope to
-	produce a homogeneous sample of afterglows for 86 spectra. We here provide a
-	public release of all the reduced spectra, including continuum estimates and
-	telluric corrections. For completeness, we also provide reductions for the
-	late-time observations of the underlying host galaxies. We provide an assessment
-	of the degree of completeness with respect to the parent GRB population, in
-	terms of the X-ray properties of the bursts in the sample and find that the
-	sample presented here is consistent with being drawn from the \textit{Swift}
+	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
+	of observational selection criteria that are unbiased with regards to intrinsic
+	properties of the GRBs, the follow-up effort has been guided by the
+	\textit{Swift}-telescope to produce a homogeneous sample of spectra for 91 GRBs.
+	We here provide a public release of all the reduced spectra, including continuum
+	estimates and telluric corrections. For completeness, we also provide reductions
+	for the late-time observations of the underlying host galaxies. We provide an
+	assessment of the degree of completeness with respect to the parent GRB
+	population, in terms of the X-ray properties of the bursts in the sample and
+	find that the sample presented here is representative of the full \textit{Swift}
 	sample. For all bursts (39) for which it is possible, we provide a measurement
-	of the neutral hydrogen column density, increasing the total number of published HI
-	column density measurements by $\sim$ 50 per cent. We constrain the fraction of
-	dark bursts to be less than 28 per cent based on the detection of an optical
-	afterglow and tentatively confirm previous results that larger optical darkness
-	is correlated with increased X-ray absorption. This dataset provides a unique
-	possibility to study the ISM across cosmic time. }
+	of the neutral hydrogen column density, increasing the total number of published
+	HI column density measurements by $\sim$ 50 per cent. %	We constrain the
+	% fraction
+	%	of dark bursts to be less than 28 per cent based on the detection of an
+	% optical
+	%	afterglow and tentatively confirm previous results that larger optical
+	% darkness
+	%	is correlated with increased X-ray absorption.
+	This dataset provides a unique possibility to study the ISM across cosmic time,
+	from the local progenitor surroundings to the galaxy wide conditions}
 
@@
 
-$\gamma$-ray bursts (GRBs) are interesting as they provide the opportunity to
-derive new constraints on a very wide range of topics in astrophysics. Examples
-range from small scale phenomena relating to magnetars, properties of highly
-relativistic jets, hyper/supernova explosions, the interstellar medium, dust
-extinction curves, starburst galaxies, chemical and molecular abundances, escape
-of ionizing radiation, the ionization state of the intergalactic medium,
-intervening absoption systems to standard candles in cosmology
-\citep[e.g.,][]{Lyons2010, Molinari2007, HjorthBloom2012, Wijers1998,
-	Prochaska2009, Savaglio2006, Ghirlanda2007}. In that sense GRB research is very
-multidisciplinary.
+$\gamma$-ray bursts (GRBs) provide the opportunity to derive new constraints on
+a very wide range of topics in astrophysics. Examples range from small scale
+phenomena relating to magnetars, properties of highly relativistic jets,
+hyper/supernova explosions, the interstellar medium, dust extinction curves,
+starburst galaxies, chemical and molecular abundances, escape of ionizing
+radiation, the ionization state of the intergalactic medium, intervening
+absoption systems to standard candles in cosmology \citep[e.g.,][]{Lyons2010,
+	Molinari2007, HjorthBloom2012, Wijers1998, Prochaska2009, Savaglio2006,
+	Ghirlanda2007, Vergani2009, Christensen2017}.
 
@@
 galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.,][]{Ricker2004}. A crucial aspect of the great
-success of the {\it Swift} mission has been the extensive ground based follow-up
-observations of the afterglows and of the host galaxies of the GRBs. This work
-has involved a large community of researchers both inside and outside the
-official {\it Swift} team. This fruitful collaboration has been greatly
-facilitated by the open data access policy of the {\it Swift} mission.
+from earlier missions \citep[e.g.,][]{Ricker2004}. With the two instruments, the
+Burst Alert Telescope (BAT) \citep{Barthelmy2005} and the X-Ray Telescope (XRT)
+\citep{Burrows2005}, \swift~is an ideal starting point for GRB investigations. A
+crucial aspect of the success of the {\it Swift} mission has been the extensive
+ground based follow-up observations of the afterglows and of the host galaxies
+of the GRBs, involving a large community of researchers. This fruitful
+collaboration has been greatly facilitated by the open data access policy of the
+{\it Swift} mission.
 
@@
 	defining the sample.}
-	\item No closeby bright stars.
+	\item No bright, nearby stars within $ 1.8 + 0.4 \times exp[(20 - R)/2.05$] arcsec, where R is the USNO magnitude of the star)
 	\item Galactic $A_V \lesssim 0.5$ mag.
@@
 selecting only bursts that have been triggered on board the \textit{Swift} space
-telescope \citep{Gehrels2004}, based on the Burst Alert Telescope (BAT)
-\citep{Barthelmy2005}, we  start out with a homogeneously selected sample where
+telescope \citep{Gehrels2004}, based on the BAT, we  start out with a homogeneously selected sample where
 burst characteristics are delivered immediately, allowing for an informed
@@
 
-Because the localization accuracy of BAT is 1 - 4 arcminutes
-\citep{Barthelmy2005}, a secure host identification is in general impossible
-based on BAT alone. We therefore additionally require an X-ray position from the
-X-Ray Telescope (XRT) \citep{Burrows2005} to be distributed to the GCN network
-\citep{Barthelmy2000} within 12 hours and to account for observing constraints
-on \textit{Swift} that XRT began observations within 10 minutes. The additional
-timing requirement of the XRT follow-up means that all bursts in our sample have
-detected X-ray afterglows. Because the XRT completeness is very high
-\citep{Burrows2007}, this cut should not alter the parent sample significantly.
+Because the localization accuracy of BAT is 1 -- 4 arcminutes
+\citep{Barthelmy2005}, a secure host association is in general very hard based
+on BAT alone. We therefore additionally require an X-ray position from XRT to be
+distributed to the GCN network \citep{Barthelmy2000} within 12 hours and to
+account for observing constraints on \textit{Swift} that XRT began observations
+within 10 minutes. The additional timing requirement of the XRT follow-up means
+that all bursts in our sample have detected X-ray afterglows. Because the XRT
+completeness is very high \citep{Burrows2007}, this cut should not alter the
+parent sample significantly.
 
@@
 
-Under rare circumstances, the use of the rapid response mode (RRM) has been
-possible. In case of a promptly visible GRB, a robotic trigger was sent to
-the telescope if at the moment of the GCN notice the GRB is observable and
-fulfills the following criteria.
+Under rare circumstances, the use of the ESO rapid response mode (RRM) has been
+possible. In case of a promptly visible GRB, a robotic trigger was sent to the
+telescope if at the moment of the GCN notice the GRB is observable and fulfills
+the following criteria.
 
@@
 	\item No constraint on Moon phase or distance. 
-	\item Elevation in the sky > 22$^\circ$ (both now and 15 min after the trigger).
-	\item Sun elevation < -12$^\circ$ (both now and 15 min after the trigger).
+	\item Elevation in the sky > 22$^\circ$ (both at trigger time and 15 min after).
+	\item Sun elevation < -12$^\circ$ (both at trigger time and 15 min after).
 	\item The X-ray position must be available less than 1 hr after the GRB.
-	\item The flags KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true).
+	\item The flags \swift-circulated tags: KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true).
 \end{enumerate}
@@
 
-There 8 GRBs in the sample that were being observed with the RRM mode. One of
-the RRM triggers is outside the statistical sample and two of the RRM trigger
-are on short GRBs. The fastest response between \swift~GRB trigger and the
-beginning of spectroscopic integration is for GRB~160410A, for which the delay
-was only 8.4 minutes.
+There are 8 GRBs in the sample that have beem being observed with the RRM mode.
+One of the RRM triggers is outside the statistical sample and two of the RRM
+trigger are on short GRBs. The fastest response between \swift~GRB trigger and
+the beginning of spectroscopic integration is for GRB~160410A, for which the
+delay was only 8.4 minutes.
 
@@
 	position of bursts fulfilling the sample criteria specified in Sect.
-	\ref{samplecrit}. The red circles are added around GRBs which have been
-	followed up by X-shooter since the commissioning of the instrument. Blue points
-	with red circles represent GRBs that enter our sample and have X-shooter
-	spectroscopy. The different samples are compared in Sect. \ref{results}. The
-	background shows the dust maps presented in \citet{Schlegel1998} where the
-	sample criteria cut with $A_V \lesssim 0.5$ mag is removed. The colorbar below
-	indicates the value of $A_V$ on the plot. The dotted lines indicate intervals
-	of 30$^\circ$ in longitude and latitude.} \label{fig:skymap}
+	\ref{samplecrit}. The red outline is added around GRBs which have been followed
+	up by X-shooter in this work. Blue symbols with red outlines represent GRBs
+	that enter our sample and have X-shooter spectroscopy. The different samples
+	are compared in Sect. \ref{results}. The background shows the dust maps
+	presented in \citet{Schlegel1998} where the sample criteria cut with $A_V
+	\lesssim 0.5$ mag is replaced with a white color. The colorbar below indicates
+	the value of $A_V$ on the plot. The dotted lines indicate intervals of
+	30$^\circ$ in longitude and latitude.} \label{fig:skymap}
 \end{figure*}
@@
 095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260. We
-have included a few additional bursts, from the programmes 092.D-0056(A) (PI:
+089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260 (PI:
+Fynbo). We have included a few additional bursts, from the programmes
+084.D-0265(A) (PI: Benetti), 091.A-0877(A) (PI: Schady), 092.D-0056(A) (PI:
 Rau), 092.D-0633(E) (PI: Greiner), and 095.B-0811(B) (PI: Levan). The total
-collection of spectra represents \textit{all} GRBs that have been followed up by
-X-shooter up to \termdate.
+collection of spectra represents \textit{all} GRB afterglows that have been
+followed up by X-shooter up to \termdate.
 
@@
 central zone of avoidance due to high Galactic extinction cutoff (marked in
-white), the GRB positions have an apparent isotropic distribution except in the
+white), the GRB positions have an isotropic distribution except in the
 upper left quadrant, which cannot be observed from Paranal.
 
-A subset of the spectra presented here, primarily host observations, have
+30 percent of the spectra presented here, primarily host observations, have
 already been published in \citet{Kruhler2015}. We include independent
@@
 on the image, we need to chose a model which can represent how the light falls
-on the slit. We know from \citet{Trujillo2001} that the Moffat function
-\citep{Moffat1969} adequately describes an imaging PSF due to atmospheric
-turbulence, but because of aberrations in the dispersion elements and the
-rectification process, the PSF we are trying to model is different from this
-profile. To allow for flexibility in the model, we have chosen the Voigt
-function as a model for the spectral PSF and we describe how this is evaluated
-in App. \ref{voigt}. Since additionally, the host galaxy could also have a
-contribution to the image profile, this choice allows for the required freedom if
-additional flux is in the wings of the profile.
+on the slit. We know from \citet{Trujillo2001} that a Moffat function
+\citep{Moffat1969} with an index of $\beta = 4.765$ adequately describes an
+imaging PSF due to atmospheric turbulence, but because of aberrations in the
+dispersion elements and the rectification process, the PSF we are trying to
+model is different from this profile. To allow for flexibility in the model, we
+have chosen the Voigt function as a model for the spectral PSF and we describe
+how this is evaluated in App. \ref{voigt}. Since additionally, the host galaxy
+could also have a contribution to the image profile, this choice allows for the
+required freedom if additional flux is in the wings of the profile.
 
@@
 odd times under varying conditions, this prioritized list ensures that we are
-always doing the optimal correction.
+always doing the best possible correction.
 
@@
 	prominent lines seen in GRB afterglows from \citet{Christensen2011a}.
-	Additionally, three intervening systems are seen in this sightline This
+	Additionally, three intervening systems are seen in this sightline. This
 	spectrum has been analyzed in detail in \citet{Friis2015}.}
@@
 lines to use, and we therefore use a scaled value as in the NIR arm. While this
-simple method is not terribly accurate (for one, the spatial profile of the
-trace is not a perfect Gaussian), it nevertheless gives a sufficiently accurate
-estimate for the analysis of these poor signal to noise science spectra.
+simple method is perfect (the spatial profile of the trace is not a perfect
+Gaussian), it nevertheless gives a sufficiently accurate estimate for the
+analysis of these poor signal to noise science spectra.
 
@@
 are not discussed as part of the investigation of the statistical properties of
-the GRB population. A prime example of a burst outside the statistical sample is the
-bright \textit{INTEGRAL} burst GRB~161023A, that turned to be a rich spectrum with at
-least 11 intervening absorption systems.
+the GRB population. A prime example of a spectacular burst outside the
+statistical sample is the bright \textit{INTEGRAL} burst GRB~161023A, that
+contains at least 11 intervening absorption systems, see \ref{GRB161023A}.
 
@@
 our sample is for the short, $z = 1.717$, GRB~160410A for which spectroscopic
-integration was initialed only 8.4 minutes after the BAT trigger. To illustrate
-importance of the follow-up delay for the redshift completeness, we plot the
+integration was initiated only 8.4 minutes after the BAT trigger. To illustrate
+the importance of the follow-up delay for the redshift completeness, we plot the
 redshift completeness as a function of delay time in Fig. \ref{fig:timing} for
 all the bursts we have followed up, including the ones outside the statistical
-sample. Visible is the decreasing fraction of GRBs with a redshift determination
-for increasing delays. The overall redshift completeness for bursts that we have
-followed up is 92 per cent. This degree of completeness in the followed bursts,
-illustrates the efficiency of VLT/X-shooter in redshift determination. Not shown
-in the figure are an additional 12 bursts that have redshift determinations
-based on late-time host observations with delay times longer than $\sim$10 days.
+sample. As can be seen from the figure, the fraction of GRBs with a redshift
+determination decreases with follow-up delay. The overall redshift completeness
+for bursts that we have followed up is 92 per cent. This degree of completeness
+in the followed bursts, illustrates the efficiency of VLT/X-shooter in redshift
+determination. Not shown in the figure are an additional 12 bursts that have
+redshift determinations based on late-time host observations with delay times
+longer than $\sim$10 days.
 
 \begin{figure}
-	\centerline{\includegraphics[width=8cm]{figures/timing.pdf}} \caption{Afterglow
+	\centerline{\includegraphics[width=9cm]{figures/timing.pdf}} \caption{Afterglow
 	magnitude at the start of observation as a function of follow-up delay. The
@@
 which we will derive statistical properties of the GRB population. The redshift
-completeness of the full statistical sample is 60 per cent. We return to the
-question of redshift completeness in Sect. \ref{redshift}. From this sample,
-90 GRBs have been spectroscopically followed up with X-shooter. In order to
-assess whether the subset of bursts followed up are representative of the
-underlying GRB parent population, we compare intrinsic properties of GRBs in our
-sample to GRBs in the full sample followed up by \textit{Swift}. We show the
-comparison between the BAT (15-150 keV) fluence, the XRT flux (0.3 - 10 keV) at
-11 hours, and the intrinsic X-ray derived equivalent hydrogen column density,
+completeness of the full statistical sample is 61 per cent. We return to the
+question of redshift completeness in Sect. \ref{redshift}. From this sample, 91
+GRBs have been spectroscopically followed up with X-shooter. In order to assess
+whether the subset of bursts followed up are representative of the underlying
+GRB parent population, we compare intrinsic properties of GRBs in our sample to
+GRBs in the full sample followed up by \textit{Swift}. We show the comparison
+between the BAT (15 -- 150 keV) fluence, the XRT flux (0.3 -- 10 keV) at 11
+hours, and the intrinsic X-ray derived equivalent hydrogen column density,
 $N_{\mathrm{HI,X}}$, in Fig. \ref{fig:swift_complete}. For the latter, we can
 only use values of $N_{\mathrm{HI,X}}$ derived for bursts with a measured
-redshift, excluding $\sim 75$ per cent of the bursts. We return to this last
-point in Sect. \ref{darkness}.
+redshift, excluding $\sim 75$ per cent of the full \swift~sample. We return to
+the last point in Sect. \ref{darkness}.
 
@@
 \caption{Comparison between burst properties of all bursts observed with
-	\textit{Swift} (blue), the ones that fullfill the sample criteria specified in
-	Sect. \ref{samplecrit} (green), and the subset that has been observed as part of
-	the statistical sample (red). The left plot shows the fluence in 15-150
-	keV as observed by BAT. The middle panel shows the 0.3 - 10 keV flux, 11 hours
-	after the bursts as measured by XRT. In the right most panel, we show
-	$N_{\mathrm{HI,X}}$ based on the XRT spectrum \citep{Evans2009}.}
-\label{fig:swift_complete}
+	\textit{Swift} (blue), the statistical sample that fullfill the criteria
+	specified in Sect. \ref{samplecrit} (green), and the subset that has been
+	observed as part of the statistical sample (red). The left plot shows the
+	fluence in 15-150 keV as observed by BAT. The middle panel shows the 0.3 - 10
+	keV flux, 11 hours after the bursts as measured by XRT. In the right most
+	panel, we show $N_{\mathrm{HI,X}}$ based on the XRT spectrum
+	\citep{Evans2009}.} \label{fig:swift_complete}
 \end{figure*}
@@
 each of the distributions, which can be used as point estimates for the
-population distribution. These are provided in Table.
+population distribution. These are provided in Table
 \ref{tab:sample_properties}. It can be seen from the values that the three
@@
 \textit{Swift}-sample and that additionally, the follow-up effort conserves the
-distributions of the intrinsic GRB properties (except perhaps for $N_{\mathrm{HI,X}}$, see Sect. \ref{darkness}).
+distributions of the intrinsic GRB properties (except perhaps for
+$N_{\mathrm{HI,X}}$, see Sect. \ref{darkness}).
 
@@
 We therefore conclude that the statistics of the sample presented here,
-conserves the intrinsic properties of the GRBs in the full \textit{Swift}
-sample. At least in terms of BAT fluence and X-ray afterglow flux at 11 hours.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Properties of rejected triggers} \label{badbursts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Out of the 158 bursts in the statistical sample, 32($\sim$ 20 per cent) have not
-been observed due to reasons unrelated to the GRB or afterglow properties. The
-reasons include unavailability of the telescope due to technical maintenance
+conserves the intrinsic properties of the GRBs in the full \textit{Swift} sample
+- at least in terms of BAT fluence and X-ray flux at 11 hours.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Properties of rejected triggers} \label{badbursts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Out of the 158 bursts in the statistical sample, 33( $\sim$ 20 per cent) have
+not been observed due to reasons unrelated to the GRB or afterglow properties.
+The reasons include unavailability of the telescope due to technical maintenance
 (e.g., mirror re-coating), a visitor rejecting the TOO trigger, or bad weather.
@@
 statistical properties of the full sample. Removing all these bursts,
-drastically improves the redshift completeness from 60 per cent to 83 per cent.
+drastically improves the redshift completeness from 61 per cent to 84 per cent.
+The remaining burst not followed up already had redshift from other instruments
+or were very faint and thus observations unlikely to yield anything. In the
+remainder of the text, we consider the 125 bursts our statistical sample.
 
@@
 sample of GRBs. The imposed selection criteria (see Sect. \ref{samplecrit})
-ensure that the GRBs entering our homogeneous sample fairly represent  the
+ensure that the GRBs entering our homogeneous sample, fairly represent the
 underlying population. The redshift distribution of such a sample holds valuable
@@
 
+\input{tables/redshift_comparison}
+
 In Fig.~\ref{fig:z} we show the redshift distribution of all  the observed GRBs.
@@
 sample and in the main panel, we show the redshifts of the individual bursts as
-a function of GRB luminosity in the observed 15-150 keV band. To calculate the
+a function of GRB luminosity in the observed 15 -- 150 keV band. To calculate the
 luminosity, $L_{\mathrm{BAT}}$, we follow a similar procedure as
@@
 
-We find the following mean and median and 14th and 86th percentiles for the
-redshift distribution. First for the statistical sample:
-$z_{\mathrm{avg}}=1.99$, $z_{\mathrm{med}}=1.64$, $z_{14\%}=0.60$ and
-$z_{86\%}=3.39$. For all GRBs observed with X-shooter we find
-$z_{\mathrm{avg}}=1.90$, $z_{\mathrm{med}}=1.52$, $z_{14\%}=0.55$ and
-$z_{86\%}=3.36$. For comparison, the corresponding values found in
-\citet{Fynbo2009} are z = 2.2 for both the mean and the median.
-
-%Because the redshift completeness of our statistical sample is 60 per cent,
-%making inference of the \textit{true} redshift distribution of GRBs on this sample
-%is difficult. A large fraction of the bursts for which we were unable to
-%secure a redshift are missing due to reasons unrelated to the GRBs (see Sect.
-%\ref{badbursts}). By trimming away the 33 bursts that can be rejected due to
-%telescope availability, the redshift completeness increases to 83 per cent.
-%
-%As is shown in Sect. \ref{timing}, the redshift completeness
-%decreases with increasing follow-up delay, we can additionally increase the
-%redshift completeness by imposing cuts on the follow-up delay. Our ability to
-%rapidly follow up a burst is also not dependent on the burst properties and
-%therefore we can additionally cut delays longer than 24 hours (all burst in the
-%sample are in principle observable within 24 hours). \todo{Unfinished section}
-
-Because the redshift completeness of our statistical sample is 60 per cent,
+We compare the point estimates for the redshift distributions of previous
+complete samples of GRBs in Tab. \ref{tab:redshift_comparison}. We see that when
+we compare to other complete samples, the XS-GRB presented here has the lowest
+average redshift. However, the other samples also exhibit a large spread in the
+redshift distributions. A 2-sided KS test confirms that there is only mild
+evidence against the Fynbo09 (p-value = 0.05) and TOUGH (p-value = 0.05) samples
+being drawn from the same parent sample, while XS-GRB is consistent with the
+SHOALS (p-value = 0.08) and BAT6 (p-value = 0.86) samples.
+
+Because the redshift completeness of our statistical sample is 84 per cent,
 making inference of the \textit{true} redshift distribution of GRBs based on
-this sample is difficult. The biases introduced in the measured redshift
+this sample is impossible. The biases introduced in the measured redshift
 distribution by our increased ability to secure redshift of optically brighter
@@
 GRBs are seen above redshift $z \gtrsim 1$ as shown in Fig. \ref{fig:z}. As
-described in detail in \citet{Perley2016b}, bursts for which the redshift is
+described in detail in \citet{Hjorth2012} and \citet{Perley2016b}, bursts for which the redshift is
 measured from the afterglow are systematically found for host galaxies with a
@@
 	as the black dashed lines, from the most recent measurement by
-	\citet{Planck2015}.} \label{fig:z}
+	\citet{Planck2015}. On the top of the plot is shown the marginalization of the
+	redshift. Again the blue histogram represents the bursts than enter our sample
+	and the white histogram the full GRB population. } \label{fig:z}
 \end{figure*}
@@
 extinction relative to the X-ray brightness has been parametrized in term of
-their optical darkness,  $\beta_{OX}$\citep{Jakobsson2004, Rol2005,
+their optical darkness, $\beta_{OX}$ \citep{Jakobsson2004, Rol2005,
 	VanderHorst2009}. The X-ray properties of such bursts have previously been
-investigated \citep{DePasquale2003, Fynbo2009, Melandri2012}. There are some
+investigated \citep{DePasquale2003, Fynbo2009, Melandri2012} and there are some
 indications that dark bursts have somewhat higher X-ray luminosity and
-\nhx~compared to the optically bright bursts. This indicates along with
-investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
-	Hjorth2012, Perley2016b}, that the extinction of the optical afterglow is
-primarily driven by the presence of dust in the host galaxy.
+$N_{\mathrm{HI, X}}$ compared to the optically bright bursts. This indicates
+along with investigations of host galaxy properties \citep{Greiner2011,
+	Kruhler2011, Hjorth2012, Perley2016b}, that the extinction of the optical
+afterglow is primarily driven by the presence of dust in the host galaxy.
+\citet{Hjorth2012} find that systems with no optical afterglow have higher
+$N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which however
+also turn out to be redder.
 
@@
 simultaneous measurement of the X-ray flux density and the optical flux density
-which is in practice never possible. As a proxy, we use the measured acquisition
-camera magnitude reported in Tab. \ref{tab:sample_overview} to get the optical
-flux density at the beginning of the spectroscopic integration. Because we know
-the delay between the follow-up and the \textit{Swift} trigger, we can use the
-measured XRT lightcurve \citep{Evans2007,
+which is in practice is possible, but in reality extremely rarely available. As
+a proxy, we use the measured acquisition camera magnitude reported in Table
+\ref{tab:sample_overview} to get the optical flux density at the beginning of
+the spectroscopic integration. Because we know the delay between the follow-up
+and the \textit{Swift} trigger, we can use the measured XRT lightcurve
+\citep{Evans2007,
 	Evans2009}\footnote{http://www.swift.ac.uk/xrt\_curves/trigger\_numer/flux.qdp}
@@
 It is clear, however, that the fraction of dark bursts for which we have
-measured the redshift is lower than the intrinsic dark burst fraction. To
-measure the redshift we either need a detected afterglow to obtain a spectrum or
+measured redshifts for is lower than the intrinsic dark burst fraction. To
+measure a redshift, we either need a detected afterglow to obtain a spectrum or
 to locate the host galaxy and determine the redshift from there. It is more
@@
 the bursts for which follow-up has actually been attempted, this number is
-$\sim$ 18 per cent, confirming that we are indeed biased against bursts without a
-detected optical afterglow in the spectroscopic sample.
-
-Despite that the fraction of dark bursts is lower in the observed sample,
+$\sim$ 18 per cent, confirming that we are indeed biased against bursts without
+a detected optical afterglow in the spectroscopic sample. Since \textit{all}
+GRBs should have an optical afterglow, the presence of an optical afterglow is
+more a statement about the sensitivity than optical properties of the bursts.
+
+Regardless that the fraction of dark bursts is lower in the observed sample,
 compared to the statistical sample, the X-ray properties not differ
-significantly, as shown in Sect. \ref{completeness}. If the tendency for dark
-bursts to have a higher \nhx~is real, that the decreased fraction of dark bursts
-does not alter the observed prompt X-ray brightness distribution, potentially
-reflects the independence of the X-ray brightness on the density of the
-circumburst medium \citep{Freedman2001, Berger2003, Nysewander2009}, if the
-measured \nhx is primarily driven by a high column local to the burst. Because we
-only use values of \nhx~in the comparison for which the GRB has a measured
-redshift, this measure is likely biased toward optically brighter bursts.
-
-Since \textit{all} GRBs should have an optical afterglow, the presence of an
-optical afterglow is more a statement about the sensitivity than optical
-properties of the bursts.
-
-This is despite spectroscopic follow-up only being carried out in cases where
-either a detectable optical afterglow or a clear counterpart are seen, which
-naively should be biased against dark bursts occurring in more obscured
-galaxies, which is shown to exhibit different galactic properties
-\citep{Perley2009, Kruhler2011, Rossi2012, Perley2013b, Perley2015b}.
+significantly, as shown in Sect. \ref{completeness}. This is despite
+spectroscopic follow-up only being carried out in cases where either a
+detectable optical afterglow or a clear counterpart are seen, which naively
+should be biased against dark bursts occurring in more obscured galaxies, which
+is shown to exhibit different galactic properties \citep{Perley2009,
+	Kruhler2011, Rossi2012, Perley2013b, Perley2015b}. If the tendency for dark
+bursts to have a higher $N_{\mathrm{HI, X}}$ is real, that the decreased
+fraction of dark bursts does not alter the observed prompt X-ray brightness
+distribution, potentially reflects the independence of the X-ray brightness on
+the density of the circumburst medium \citep{Freedman2001, Berger2003,
+	Nysewander2009}, if the measured $N_{\mathrm{HI, X}}$ is primarily driven by a
+metal column local to the burst. Because we only use values for $N_{\mathrm{HI,
+		X}}$ in the comparison for which the GRB has a measured redshift, this measure
+is likely biased toward optically brighter bursts.
 
@@
 
-In the compilation of HI measurements towards GRBs in Tanvir et al. (submitted) %\citet{Tanvir2017} 
-there are 81 published HI values. We here provide 39 new HI
-column density measurements, i.e. an increase of the number of optically derived
-hydrogen column densities of $\sim$ 50 per cent. We compare the median, the 14th,
-and 86th percentiles of the two distributions where the sample presented here
-has $21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
-$21.5_{-1.5}^{+0.4}$. We see that the two distribution has a large degree of
-overlap due to the large width of the distributions, but we find a slightly
+In the compilation of $N_{\mathrm{HI}}$ measurements towards GRBs in Tanvir et
+al. (submitted) %\citet{Tanvir2017}
+there are 81 published $N_{\mathrm{HI}}$ values. We here provide 39 new neutral
+hydrogen column density measurements -- an increase of the number of optically
+derived hydrogen column densities of $\sim$ 50 per cent. We compare the median,
+the 14th, and 86th percentiles of the two distributions. The sample presented
+has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
+\nh~$= 21.5_{-1.5}^{+0.4}$. We see that the two distribution has a large degree
+of overlap due to the large width of the distributions, but we find a slightly
 higher median value for the new sample presented here. A 2-sided KS test gets us
@@
 
-The sample we present here include spectroscopic observations of 90 systems
-fulfilling our sample criteria including 17 spectra that are late time
+The sample presented here include spectroscopic observations of 91 systems
+fulfilling our sample criteria including 18 spectra that are late time
 observations of the underlying host galaxies. All spectra have been made
@@
 that was also much advanced by the previous surveys based primarily on
-lower-resolution spectographs. The most striking result found is that GRBs
-typically are located behind very large column densities of neutral hydrogen.
-Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
-towards about 10$^5$ quasars have a handful such systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
-in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006b,
-	Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy (throughout the
-electromagnetic spectrum from X-rays to the sub-mm) we hence characterize the
-properties of star-forming galaxies over cosmic history in terms of redshifts,
-metallicities, molecular contents, ISM temperatures, UV-flux densities,
-extinction curves, etc.  A number of independent papers have been published or
-submitted for publication focusing on many of these specific issues of our
-sample such as extinction curves \citep[][see also \citealt{Fynbo2014,
+lower-resolution spectographs. Only after observing more than 12000 damped
+Lyman-$\alpha$ absorbers (DLAs) towards about 10$^5$ quasars have a handful such
+systems with $\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
+\citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by
+contrast, reveal such systems in the majority of cases \citep[][and this
+work]{Jakobsson2006b, Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy
+(throughout the electromagnetic spectrum from X-rays to the sub-mm) we hence
+characterize the properties of star-forming galaxies over cosmic history in
+terms of redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
+densities, extinction curves, etc.  A number of independent papers have been
+published or submitted for publication focusing on many of these specific issues
+of our sample such as extinction curves \citep[][see also \citealt{Fynbo2014,
 	Heintz2017}]{Japelj2015, Zafar17}, emission lines from the underlying host
 galaxies \citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
-\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al., submitted). A number of additional companion papers are also planned,
+\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al.,
+submitted). A number of additional companion papers are also planned,
 investigating the detailed properties of the sample presented here, including
@@
 
-It seems important to end this paper by stressing that the potential of using
-GRB sightlines as probes is far from fully harvested. The sample of sightlines
-probed by our spectra are not representative for all GRB sightlines as we have
-shown and consistent with earlier findings from samples based on low-resolution
-spectroscopy \citep[e.g.,][]{Fynbo2009} and from studies of complete samples of
-GRB host galaxies \citep{Hjorth2012, Covino2013, Perley2016}. \cite{Kruhler2013}
-convincingly argue, that very rich sightlines like that probed by the remarkable
-GRB~080607 \citep{Prochaska2009} are probably significantly more frequent than
-in the sightlines sampled by our spectra. However, with current instrumentation
-these sightlines are out of reach except under very fortunate circumstances as
-in the case of GRB~080607 when the afterglow could be observed only a few
-minutes after the burst with a 10-m class telescope. Observations of such
-sightlines with X-shooter-like spectrographs on the next generation of 20--40-m
-telescopes is likely to be very rewarding.
+The potential of using GRB sightlines as probes is far from fully harvested. The
+sample of sightlines probed by our spectra are not representative for all GRB
+sightlines as we have shown and consistent with earlier findings from samples
+based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009} and from studies
+of complete samples of GRB host galaxies \citep{Hjorth2012, Covino2013,
+	Perley2016}. \cite{Kruhler2013} argue, that very rich sightlines like that
+probed by the remarkable GRB~080607 \citep{Prochaska2009} are probably
+significantly more frequent than in the sightlines sampled by our spectra.
+However, with current instrumentation these sightlines are out of reach except
+under very fortunate circumstances as in the case of GRB~080607 when the
+afterglow could be observed only a few minutes after the burst with a 10-m class
+telescope. Observations of such sightlines with X-shooter-like spectrographs on
+the next generation of 20--40-m telescopes is likely to be very rewarding, given
+that the a suitable GRB detector will be online.
 
@@
 acknowledge support from the ERC-StG grant EGGS-278202. The Dark Cosmology
-Centre is funded by the Danish National Research Foundation.  TK acknowledges
-support by the European Commission under the Marie Curie Intra-European
-Fellowship Programme in FP7.  AdUP acknowledges support by the European
-Commission under the Marie Curie Career Integration Grant programme
+Centre were funded by the Danish National Research Foundation. This work was
+supported by a VILLUM FONDEN Investigator grant to JH (project number 16599).
+TK acknowledges support by the European Commission under the Marie Curie
+Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
+European Commission under the Marie Curie Career Integration Grant programme
 (FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
@@
 
+This spectrum has previously been used to form the basis for GCN
+\#17420\footnote{\url{http://gcn.gsfc.nasa.gov/other/17420.gcn3}}, but it is not
+published in the refereed literature. Due to the short visibility of this burst,
+it does not enter into the statistical sample. Beginning in twilight,
+$4\times600$ s integration time contains a weak/moderate S/N GRB afterglow
+continuum throughout the X-shooter spectral coverage, down to $\sim$ 3750 \AA.
+In the continuum are absorption features, which we identify as due to metal
+lines from \znii, \feii, and \mgii. Additionally, \lya-absorption is seen at the
+end of the trace. The spectral position of the lines means that this GRB is at
+$z = 2.087$. From the \lya~trough, we infer \nh~$=21.7 \pm 0.4$. The afterglow
+continuum appears depressed in the blue end of the spectrum, suggestion dust
+extinction in the host.
+
 \subsection{GRB~150301B (z = 1.517)}
 
-\subsection{GRB~150403A (z = 2.06)}
+This spectrum has already been used to inform GCN
+\#17523\footnote{\url{http://gcn.gsfc.nasa.gov/other/17523.gcn3}}, but it is not
+used elsewhere. Starting 5.1 hours after the BAT trigger, this spectrum is based
+on $6 \times 600$ s spectroscopic integration. The GRB afterglow continuum is
+well detected across the entire spectral coverage of X-shooter at moderate S/N.
+Imprinted on the continuum are absorption features which we identify as being
+caused by \SIii, \civ, \alii, \feii, \mgii, and \mgi - all at a similar redshift
+of $z = 1.517$, which is likely the redshift of the GRB.
+
+\subsection{GRB~150403A (z = 2.057)}
+
+This spectrum has previously been used to form the basis for GCN
+\#17672\footnote{\url{http://gcn.gsfc.nasa.gov/other/17672.gcn3}}. $4 \times
+600$ s spectroscopic integration, starting 10 hours after the GRB trigger,
+captures a bright GRB afterglow. A broad absorption trough centered at $\sim$
+3700 \AA due to \lya signals the redshift of the GRB, which is refined to $z =
+2.057$ based on the additional detection of metal absorption lines. We readily
+identify features associated with \sii, \SIiv, \oi, \SIii, \SIii$^*$, \cii,
+\cii$^*$, \civ, \alii, \feii, \feii$^*$, \mni, \mgii, and \mgi~in the host of
+the GRB. An intervening \civ absorber is additionally detected at $z=1.76$. From
+the absorption trough due to \lya, we infer the amount of neutral hydrogen in
+the host along the line of sight to be \nh~$=21.8 \pm 0.2$.
 
 \subsection{GRB~150423A (z = 1.394)}
+
+The gross content of these observations have previously been presented in GCN
+\#17755\footnote{\url{http://gcn.gsfc.nasa.gov/other/17755.gcn3}}, but is not
+published as part of any refereed paper. This bona-fide short GRB (T90 (15-350
+keV ) = is $ 0.22 \pm 0.03$ s \citep{Lien2016}) was observed in RRM mode and
+spectroscopic integration started after only 22 minutes. A series of stare mode
+observations, increasing in exposure time, and ending with a nodding sequence
+totals $\sim$ 5000 s are combined to form this spectrum. A faint, almost
+featureless continuum is detected at low S/N all the way to the bluest part of
+the spectrum. An absorption doublet is detected in the VIS arm against the GRB
+afterglow continuum, which we identify as \mgii~at $z = 1.394$.
 
@@
 
-\subsection{GRB~161023A (z = 2.710)}
+\subsection{GRB~161023A (z = 2.710)}	 \label{GRB161023A}
 

### 2018-01-25 document/tables/redshift_comparison.tex
Comments from Giorgos and Jens + notes on individual bursts

\begin{table}[H]
	\centering
	\begin{tabular}{ccccc}
		\hline
		\hline\noalign{\smallskip}
		{Sample} & {N$_{\mathrm{bursts}}$} & {z$_{\mathrm{completeness}}$} &  {z$_{\mathrm{mean}}$} &  {$z_{\mathrm{med}}$} \\
		\hline\noalign{\smallskip}
		{\smallskip}
		XS-GRB & 125 & 84 \% & 1.86 & $1.46_{-0.86}^{+1.77}$ \\
		{\smallskip}
		SHOALS  & 119 &  99 \% &  2.18  & $2.06_{-1.20}^{+1.27}$ \\
		{\smallskip}
		BAT6 & 58 & 95 \% &  1.90 &  $1.71_{-1.04}^{+1.31}$ \\
		{\smallskip}
		TOUGH & 69 &  77 \% & 2.20 & $2.11_{-1.36}^{+1.42}$ \\
		{\smallskip}
		Fynbo09 & 146 &  49 \% &  2.2 & $2.1_{-1.23}^{+1.28}$ \\
		\hline\noalign{\smallskip}

\end{tabular}

\caption{Comparison between the redshift distributions of previous complete
	samples. The SHOALS redshift characteristics are taken from \citet{Perley2016a},
	the BAT6 redshifts are from \citet{Salvaterra2012}, TOUGH is from
	\cite{Hjorth2012}, and Fynbo09 are from \citet{Fynbo2009}. The errors shown on
	the median redshift contain 68 per cent of the probability mass.
	\label{tab:redshift_comparison}}

\end{table}

### 2018-01-25 document/tables/sample_properties.tex
Comments from Giorgos and Jens + notes on individual bursts

@@
 \begin{table*}[!ht]
-
 	\centering
@@
 		$\log(\mathrm{N}_{\mathrm{HI_x}})$ & $21.7_{-0.9}^{+0.6}$ &  $21.5_{-3.4}^{+0.7}$ &  $21.6_{-4.5}^{+0.7}$  \\
-
 		\hline\noalign{\smallskip}
@@
 \end{tabular}
-	\caption{Population properties (median and 14th and 86th percentiles as the
+
+\caption{
+	Population properties (median and 14th and 86th percentiles as the
 	error intervals) for the \textit{Swift sample} and the subset of bursts
@@
 	samples are very similar, which shows that our selection criteria effectively
-	conserve the statistical properties of the underlying population, as least for these parameters.  \label{tab:sample_properties}} \end{table*}
+	conserve the statistical properties of the underlying population, as least for
+	these parameters.  Notice that not all bursts have measurements of the
+	quantities we compare. \label{tab:sample_properties}
+	}
+
+\end{table*}

### 2018-01-25 py/Sample_completeness.py

@@ -36,7 +36,15 @@ def filt_nan(input_array):
             holder.append(np.nan)
 
     return np.array(holder)
+fbo = [3.2425, 2.8983, 1.2356, 3.9693, 1.38, 1.7102, 2.6147, 0.8278, 3.3467, 2.1995, 3.5328, 2.3, 4.0559, 3.9133, 1.5026, 2.1, 3.2213, 3.0749, 0.1257, 3.424, 1.92, 2.7108, 0.5428, 1.8836, 5.4636, 1.2622, 0.3463, 0.7578, 3.4344, 1.3145, 2.3521, 1.4965, 0.8397, 0.9705, 2.309, 2.0394, 3.6298, 2.4541, 2.1462, 5.2, 2.6918, 0.8227, 1.3308, 2.6419, 2.4274, 0.9382, 1.9492, 1.5119, 1.1014, 1.5457, 2.6892, 1.4171, 1.6403, 3.0368, 1.2322, 0.8454, 2.5914, 2.2045, 1.5042, 3.3604, 2.3739, 6.7, 0.6887, 1.6919, 0.7029, 3.6856, 3.2086, 1.5471, 2.9538, 1.0301, 2.433]
 
+tough = [1.95, 1.4436, 2.8983, 2.7, 0.6528, 5.2, 0.606, 3.96855, 1.38, 2.5043, 2.61469, 1.434, 0.8278, 6.295, 3.3467, 2.5273, 2.1992, 2.4296, 1.059, 0.9364, 0.481, 3.5328, 0.03351, 5.11, 3.2213, 2.1357, 3.773, 3.0749, 0.125, 3.424, 1.92, 2.7108, 1.532, 0.5428, 1.9229, 1.8836, 0.937, 5.4636, 1.2622, 0.3463, 0.7578, 3.4344, 1.3145, 2.6208, 2.3521, 2.3384, 1.49594, 0.8397, 1.9591, 2.309, 2.0394, 3.6298, 2.4541, 4.406, 0.0889, 0.787, 2.17, 2.0858]
+
[…]

### 2018-01-25 py/follow_up_timing.py

@@ -13,12 +13,12 @@ import numpy as np
 import matplotlib as mpl
 from matplotlib.ticker import FormatStrFormatter
 params = {
-   'axes.labelsize': 12,
-   'font.size': 12,
-   'legend.fontsize': 12,
-   'xtick.labelsize': 12,
-   'ytick.labelsize': 12,
-   'text.usetex': False,
+   'axes.labelsize': 15,
+   'font.size': 15,
+   'legend.fontsize': 15,
+   'xtick.labelsize': 15,
+   'ytick.labelsize': 15,
+   'text.usetex': True,
    'figure.figsize': [6, 6]
    }
 mpl.rcParams.update(params)
@@ -57,7 +57,7 @@ def main():
     sorted_z = z[delay_sort]
     print(len(sorted_z))
     print(sorted_z)
-    exit()
+    # exit()
     fractional_completeness = (1 - np.cumsum(np.isnan(sorted_z).astype("int"))/len(sorted_z))*100
     # print(len(sorted_z), np.sum(np.isnan(sorted_z).astype("int")))
     # print(fractional_completeness[-1])
@@ -67,19 +67,21 @@ def main():
     # Plot
     fig, ax1 = pl.subplots()
 
+
     color = sns.color_palette()[0]
     color_rgb = mpl.colors.colorConverter.to_rgb(color)
     colors = [sns.set_hls_values(color_rgb, l=l) for l in np.linspace(1, 0, 12)]
     cmap = sns.blend_palette(colors, as_cmap=True)
     # cmap = pl.get_cmap("plasma")
     # With redshift and acq mag
-    sc = ax1.scatter(delay[idx_hasz & idx_limits], mag[idx_hasz & idx_limits], c=z[idx_hasz & idx_limits], cmap=cmap, s=25)
+    sc = ax1.scatter(delay[idx_hasz & idx_limits], mag[idx_hasz & idx_limits], c=z[idx_hasz & idx_limits], cmap=cmap, s=35)
+
     # Without redshift and with acq mag
-    ax1.scatter(delay[~idx_hasz & idx_limits], mag[~idx_hasz & idx_limits], color=sns.color_palette()[2], s=35, marker="x")
+    ax1.scatter(delay[~idx_hasz & idx_limits], mag[~idx_hasz & idx_limits], color=sns.color_palette()[2], s=45, marker="x")
     # With redshift, but without acq mag
[…]

### 2018-01-25 py/skymap.py

@@ -48,8 +48,14 @@ def main():
     ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
     A_Vmap = ebv_map*3.1
     A_Vmap[A_Vmap >= 0.5] = 0
-    ax = hp.mollview(A_Vmap, max=0.5, title="")
+    ax = hp.mollview(A_Vmap, max=0.5, title="", cbar=False)
     # , cbar=False
+
+
+    fig = pl.gcf()
+    ax = pl.gca()
+    image = ax.get_images()[0]
+    cmap = fig.colorbar(image, ax=ax, location="bottom", pad = 0.02, label=r"$A_V$", shrink=0.5)
     hp.graticule()
 
     # Overplot swift positions

### 2018-01-26 document/XSGRB_sample.tex
Part Susanna and Paolo comments

@@
 galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.,][]{Ricker2004}. With the two instruments, the
-Burst Alert Telescope (BAT) \citep{Barthelmy2005} and the X-Ray Telescope (XRT)
-\citep{Burrows2005}, \swift~is an ideal starting point for GRB investigations. A
-crucial aspect of the success of the {\it Swift} mission has been the extensive
-ground based follow-up observations of the afterglows and of the host galaxies
-of the GRBs, involving a large community of researchers. This fruitful
-collaboration has been greatly facilitated by the open data access policy of the
-{\it Swift} mission.
+from earlier missions \citep[e.g.,][]{VanParadijs2000, Ricker2004}. With the two
+on-board instruments, the Burst Alert Telescope (BAT) \citep{Barthelmy2005} and
+the X-Ray Telescope (XRT) \citep{Burrows2005}, \swift~is an ideal telescope for
+GRB hunting. A crucial aspect of the success of the {\it Swift} mission has been
+the extensive ground based follow-up observations of the afterglows and of the
+host galaxies of the GRBs, involving a large community of researchers. This
+fruitful collaboration has been greatly facilitated by the open data access
+policy of the {\it Swift} mission. The close collaboration between detection
+facilities and electromagnetic  follow-up campaigns continue to be immensely
+rewarding, as recently highlighted by the simultaneous detection of
+gravitational waves and light from the neutron star merger in the shape of
+GW170817/GRB180817A/AT2017gfo \citep{LIGOScientificCollaboration2017a,
+	LIGOScientificCollaboration2017}.
 
@@
 restrictions exists already, being it weather conditions, pointing restrictions
-of the telescope or poorly localized bursts as reported by the alerting
+of the telescope or unconstrained burst localizations as reported by the alerting
 telescope. To maximize the return of the follow-up campaign, we have chosen a
@@
 	are compared in Sect. \ref{results}. The background shows the dust maps
-	presented in \citet{Schlegel1998} where the sample criteria cut with $A_V
-	\lesssim 0.5$ mag is replaced with a white color. The colorbar below indicates
-	the value of $A_V$ on the plot. The dotted lines indicate intervals of
-	30$^\circ$ in longitude and latitude.} \label{fig:skymap}
+	presented in \citet{Schlegel1998}. Please note that we removed the background
+	where the sample criteria cut ($A_V \lesssim 0.5$ mag) and replaced it with a
+	white color visible at the equator. The colorbar below indicates the value of
+	$A_V$ on the plot. The dotted lines indicate intervals of 30$^\circ$ in
+	longitude and latitude.} \label{fig:skymap}
 \end{figure*}
@@
 089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260 (PI:
-Fynbo). We have included a few additional bursts, from the programmes
-084.D-0265(A) (PI: Benetti), 091.A-0877(A) (PI: Schady), 092.D-0056(A) (PI:
-Rau), 092.D-0633(E) (PI: Greiner), and 095.B-0811(B) (PI: Levan). The total
-collection of spectra represents \textit{all} GRB afterglows that have been
-followed up by X-shooter up to \termdate.
+Fynbo). These proposals were initiated on Guaranteed Time. We have included a
+few additional bursts, from the programmes 084.D-0265(A) (PI: Benetti),
+091.A-0877(A) (PI: Schady), 092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI:
+Greiner), and 095.B-0811(B) (PI: Levan). The total collection of spectra
+represents \textit{all} GRB afterglows that have been followed up by X-shooter
+up to \termdate.
 
@@
 cosmic-ray removal algorithm \citep{VanDokkum2001} implementation,
-\textit{Astro-SCRAPPY} \footnote{\url{https://github.com/astropy/astroscrappy}},
+\textit{Astro-SCRAPPY}\footnote{\url{https://github.com/astropy/astroscrappy}},
 where a wide clipping radius have been used around detected cosmics to ensure that edge residuals are
@@
 spectrophotometric standards \citep{Vernet2010, Hamuy1994}, mirror flexure
-compensation, initial sky-subtraction and lastly the rectification and merging
-of the orders. For the initial sky-subtraction, the background has been
-estimated in regions adjacent to the object trace clear of contaminating
-sources. Because of the broken ADC, for some objects there is a lot of curvature
-in the object trace along the dispersion axis of the slit. This means that for
-some bursts, the initial sky-estimate has been made from a limited number of
-pixels in the spatial direction. By doing an initial subtraction of the sky
-background on the un-rectified image we ensure that the bulk of the sky
+compensation, sky-subtraction and lastly the rectification and merging
+of the orders. Errors and bad pixel maps are propagated throughout the
+extraction. For the initial sky-subtraction, the background has been estimated
+by a running median in regions adjacent to the object trace clear of
+contaminating sources. Due to the broken ADC, for some objects there is
+curvature in the object trace along the dispersion axis of the slit. This means
+that for some bursts, the initial sky-estimate has been made from a limited
+number of pixels in the spatial direction. By doing a subtraction of
+the sky background on the un-rectified image we ensure that the bulk of the sky
 background is not redistributed by the rectification process.
@@
 resolution is a lower limit, we choose a sampling of 0.02 nm/pixel to ensure
-that the bluest part of neither of the arms have a sampling lower than the
-Nyquist sampling rate of 2 pixels per resolution FWHM.
+that the lowest wavelength part of neither of the arms have a sampling lower
+than the Nyquist sampling rate of 2 pixels per resolution FWHM.
 
@@
 background. When weighting images where the noise in each pixel is dominated by
-Poisson noise it is important to estimate the background variance in a large enough
-region, so that any correlation between the signal and the weights are removed.
-To this end, the weight map is generated by a running median window over the
-variance map, where the trace has been masked and width of the window is chosen
-to be wide enough for median variance to be generated on the basis of several
-hundred pixels. This weighting scheme automatically also optimally combines
-images of different exposure times or images where the background is varying,
-which is often the case when a burst has been observed close to twilight
+Poisson noise it is important to estimate the background variance in a large
+enough region, so that any correlation between the signal and the weights are
+removed. To this end, the weight map is generated by a running median window
+over the variance map produced by the pipeline, where the trace has been masked
+and width of the window is chosen to be wide enough for median variance to be
+generated on the basis of several hundred pixels. This weighting scheme
+automatically also optimally combines images of different exposure times or
+images where the background is varying, which is often the case when a burst has
+been observed close to twilight
+
+An additional sky-subtraction is additionally run on all rectified 2D spectra.
+This is done to remove residual sky, still present. At each pixel in the
+dispersion direction, the spatial dimension is fit with a low-order polynomial,
+after all sources are masked out. This low order polynomial is then convolved
+with a few pixel wide Gaussian filter in the dispersion direction and subtracted
+of the entire 2D image.
 
@@
 
-By STARE reducing all observations we additionally get a spectrum of the sky,
-which we can use to calibrate the wavelength solution.
+Using this method for all observations we additionally get a spectrum of the
+sky, which we can use to calibrate the wavelength solution.
 
@@
 telluric features can directly be fit with an atmospheric model
-\citep{Smette2015, Kausch2015}, which can then be used to correct for the
-absorption. The accuracy of the correction depends on the S/N per pixel for the
-target spectrum, where we have chosen the requirement that the afterglow
-continuum spectrum has a median signal-to-noise higher than a value of 10.
+(\texttt{Molecfit} \citep{Smette2015,
+	Kausch2015}\footnote{\url{http://www.eso.org/sci/software/pipelines/skytools/molecfit}}), which can then be used to correct for the absorption.
+The accuracy of the correction depends on the S/N per pixel for the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
 
@@
 	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have overplot the most
-	prominent lines seen in GRB afterglows from \citet{Christensen2011a}.
-	Additionally, three intervening systems are seen in this sightline. This
-	spectrum has been analyzed in detail in \citet{Friis2015}.}
+	prominent lines seen in GRB afterglows from \citet{Christensen2011a}. The
+	regions of most severe telluric absorption is highlighted by a grey-shading the
+	background. Additionally, three intervening systems are seen in this sightline.
+	This spectrum has been analyzed in detail in \citet{Friis2015}.}
 \label{fig:spectrum}
@@
 absorption, other forms of line broadening, and a lack of lines at some
-redshifts, this is difficult to do at poorer signal to noise ratios (the
+redshifts, this is difficult to do at lower signal to noise ratios (the
 majority of spectra in our sample). A broad starting value for the expected
@@
 sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$. We use this linear
-relation as a way to estimate the spectral resolution for medium to poor signal
+relation as a way to estimate the spectral resolution for medium to low signal
 to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR
@@
 Gaussian), it nevertheless gives a sufficiently accurate estimate for the
-analysis of these poor signal to noise science spectra.
+analysis of these low signal to noise science spectra.
 
@@
 
+%SWIFT LIMITATIONS
+%There is some evidence that the burst duration, $T_{90}$, does not scale with
+%redshift as expected from time-dilation \citep{Kocevski2013, Littlejohns2013a},
+%suggesting that we could be missing an increasing fraction of the bursts at
+%higher redshifts, but \citet{Lien2016} argues that we should only be missing the
+%bright "short" GRBs", but to the degree that \textit{Swift} is complete, we can
+%test if the sample presented here is complete.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Because this cut is unrelated to the GRB properties, it will not change the
-statistical properties of the full sample. Removing all these bursts,
-drastically improves the redshift completeness from 61 per cent to 84 per cent.
-The remaining burst not followed up already had redshift from other instruments
-or were very faint and thus observations unlikely to yield anything. In the
-remainder of the text, we consider the 125 bursts our statistical sample.
+statistical properties of the full sample. Removing all these bursts for the
+statistical sample, drastically improves the redshift completeness from 61 per
+cent to 84 per cent. The remaining burst not followed up already had redshift
+from other instruments or were very faint and without a host association thus
+observations unlikely to yield a redshift measurement. In the remainder of the
+text, we consider the 125 bursts our statistical sample.
 
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}} \caption{Optical
@@
 
-It is clear, however, that the fraction of dark bursts for which we have
-measured redshifts for is lower than the intrinsic dark burst fraction. To
-measure a redshift, we either need a detected afterglow to obtain a spectrum or
-to locate the host galaxy and determine the redshift from there. It is more
-difficult to correctly associate a galaxy with a burst when there is no detected
-optical afterglow and hence a correct redshift measurement is more difficult to
-make, see \citet{Jakobsson2005, Levesque2010} and \citet{Perley2017} for a
-recent example.
-
 Using the table maintained by
@@
 presence of an optical afterglow affects the follow-up statistics. 50.5 per cent
-of all \textit{Swift}-detected bursts do not have an detected optical afterglow,
+of all \textit{Swift}-detected bursts do not have a detected optical afterglow,
 but this number also includes bursts where no optical observations were
-available, so is a lower limit. For the bursts entering our statistical sample,
-the number of bursts without a detected optical afterglow is $\sim$ 27 per cent,
-close to the upper limit on the fraction of dark bursts found in a complete
-sample with a very high degree of redshift completeness \citep{Melandri2012}. Of
-the bursts for which follow-up has actually been attempted, this number is
-$\sim$ 18 per cent, confirming that we are indeed biased against bursts without
-a detected optical afterglow in the spectroscopic sample. Since \textit{all}
-GRBs should have an optical afterglow, the presence of an optical afterglow is
-more a statement about the sensitivity than optical properties of the bursts.
+available, so is a lower limit. For the bursts that enter our statistical
+sample, the percentage of bursts without a detected optical afterglow is 28 per
+cent, close to the upper limit on the fraction of dark bursts found in a
+complete sample with a very high degree of redshift completeness
+\citep{Melandri2012}. Of the bursts for which follow-up has actually been
+attempted, this number is 23 per cent, suggesting a slight bias against
+bursts without a detected optical afterglow in the spectroscopic sample.
+
+However, the fraction of dark bursts for which we have measured redshifts for is
+lower than the for the ones with a detected optical afterglow. For the
+afterglows we have observed as part of the statistical sample that do not have a
+detected optical afterglow, the redshift completeness is 53 per cent. For
+comparison, for the afterglows in the statistical sample we have observed with
+an optical afterglow detected, the redshift completeness is 91 per cent. This
+also goes to show that the lack of redshift completeness in the sample presented
+here is in part due to the increased difficulty of measuring a redshift for
+bursts without an afterglow. To measure a redshift, we either need a detected
+afterglow to obtain a spectrum or to locate the host galaxy and determine the
+redshift from there. It is more difficult to correctly associate a galaxy with a
+burst when there is no detected optical afterglow and hence a correct redshift
+measurement is more difficult to make, see \citet{Jakobsson2005, Levesque2010}
+and \citet{Perley2017} for a recent example.
 
@@
 is shown to exhibit different galactic properties \citep{Perley2009,
-	Kruhler2011, Rossi2012, Perley2013b, Perley2015b}. If the tendency for dark
-bursts to have a higher $N_{\mathrm{HI, X}}$ is real, that the decreased
-fraction of dark bursts does not alter the observed prompt X-ray brightness
-distribution, potentially reflects the independence of the X-ray brightness on
-the density of the circumburst medium \citep{Freedman2001, Berger2003,
-	Nysewander2009}, if the measured $N_{\mathrm{HI, X}}$ is primarily driven by a
-metal column local to the burst. Because we only use values for $N_{\mathrm{HI,
-		X}}$ in the comparison for which the GRB has a measured redshift, this measure
-is likely biased toward optically brighter bursts.
+	Kruhler2011, Rossi2012, Perley2013b, Perley2015b}. That the decreased fraction
+of dark bursts in observed sample does not alter the observed prompt X-ray
+brightness distribution, potentially reflects the independence of the X-ray
+brightness on the density of the circumburst medium \citep{Freedman2001,
+	Berger2003, Nysewander2009}, if the measured $N_{\mathrm{HI, X}}$ is primarily
+driven by a metal column local to the burst. Because we only use values for
+$N_{\mathrm{HI, X}}$ in the comparison for which the GRB has a measured
+redshift, this measure is likely biased toward optically brighter bursts.
 
@@
 
-Because the locations of long GRBs are associated with intensely star forming regions
-\citep{Hogg1999, Bloom2002, Fruchter2006}, the GRB afterglow light has to
-propagate through the large amounts of hydrogen fueling the star formation.
-Because a significant fraction of the hydrogen has not been ionized yet, the
-optical depth at the wavelength of \lya~is very high, saturating not only the
-line center, but also the damping wings. This causes a strong absorption system
-to appear in the afterglow continuum. For bursts with $z \gtrsim 1.7$, the
-position of \lya~moves into the spectroscopic coverage of X-shooter, meaning
-that we are able to detect this absorption trough due to \lya.
+The locations of long GRBs are associated with intensely star forming regions
+\citep{Hogg1999, Bloom2002, Fruchter2006}. Because a significant fraction of the
+hydrogen in these regions has not been ionized yet, the optical depth at the
+wavelength of \lya~is very high, saturating not only the line center, but also
+the damping wings. This causes a strong absorption system from the
+\lya-transition to appear in the afterglow continuum. For bursts with $z \gtrsim
+1.7$, the position of \lya~moves into the spectroscopic coverage of X-shooter,
+meaning that we are able to detect this absorption trough due to \lya.
 
@@
 al. (submitted) %\citet{Tanvir2017}
-there are 81 published $N_{\mathrm{HI}}$ values. We here provide 39 new neutral
-hydrogen column density measurements -- an increase of the number of optically
-derived hydrogen column densities of $\sim$ 50 per cent. We compare the median,
-the 14th, and 86th percentiles of the two distributions. The sample presented
-has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
-\nh~$= 21.5_{-1.5}^{+0.4}$. We see that the two distribution has a large degree
-of overlap due to the large width of the distributions, but we find a slightly
-higher median value for the new sample presented here. A 2-sided KS test gets us
-a p-value of p = 0.006, meaning relatively large evidence against the null that
-the two samples are drawn from the same underlying distribution. Because the
-bursts that have measurements of the hydrogen column density are selected solely
-based on our ability to infer a column, it is difficult to make any strong
-conclusions about the population statistics in terms of gas content.
-
-We show the comparison between the two samples in Fig. \ref{fig:NH_dist}. We
-additionally show the column density-distribution for the 12081 quasar absorbers
-with \nh~$> 20$ from \citet{Noterdaeme2012b}. The fact that GRBs systematically
-are located behind the highest \nh, previously noted by
-\citep[e.g.,][]{Prochaska2007, Fynbo2009}, is stricking. The reason for this is
-that quasar sample sight-lines through galaxies that are cross-section selected,
-whereas GRB sight-lines probe the dense, star-forming regions in their hosts.
-
-\begin{figure}
+there are 81 published $N_{\mathrm{HI}}$ values, excluding the measurements
+provided here. We here provide 39 new neutral hydrogen column density
+measurements -- an increase of the number of optically derived hydrogen column
+densities of $\sim$ 50 per cent. We show the two distributions in  Fig.
+\ref{fig:NH_dist}. We compare the median, the 14th, and 86th percentiles of the
+two distributions. The sample presented has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the
+rest of the literature values has \nh~$= 21.5_{-1.5}^{+0.4}$. We see that the
+two distribution has a large degree of overlap due to the large width of the
+distributions, but we find a slightly higher median value for the new sample
+presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
+relatively large evidence against the null that the two samples are drawn from
+the same underlying distribution. Because the bursts that have measurements of
+the hydrogen column density are selected solely based on our ability to infer a
+column, it is difficult to make any strong conclusions about the population
+statistics in terms of gas content.
+
+In Fig. \ref{fig:NH_dist}, we additionally show the column density-distribution
+for the 12081 quasar absorbers with \nh~$> 20$ from \citet{Noterdaeme2012b}. The
+fact that GRBs systematically are located behind the highest \nh, previously
+noted by \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this
+figure. The reason for this is that quasar sample sight-lines through galaxies
+that are cross-section selected, whereas GRB sight-lines probe the dense,
+star-forming regions in their hosts.
+
+\begin{figure}[!ht]
 	\centering \includegraphics[width=9cm]{figures/NH_dist.pdf}
@@
 that was also much advanced by the previous surveys based primarily on
-lower-resolution spectographs. Only after observing more than 12000 damped
-Lyman-$\alpha$ absorbers (DLAs) towards about 10$^5$ quasars have a handful such
-systems with $\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
-\citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by
-contrast, reveal such systems in the majority of cases \citep[][and this
-work]{Jakobsson2006b, Fynbo2009, Cucchiara2015}.  With afterglow spectroscopy
-(throughout the electromagnetic spectrum from X-rays to the sub-mm) we hence
-characterize the properties of star-forming galaxies over cosmic history in
-terms of redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
+lower-resolution spectrographs. GRB afterglow sight-lines are unique in the
+sense that only after observing more than 12000 damped Lyman-$\alpha$ absorbers
+(DLAs) towards about 10$^5$ quasars, a handful systems with
+$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
+in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
+systems in the majority of cases \citep[][and this work]{Jakobsson2006b,
+	Fynbo2009, Cucchiara2015}. With afterglow spectroscopy (throughout the
+electromagnetic spectrum from X-rays to the sub-mm) we are able to characterize
+the properties of star-forming galaxies over cosmic history in terms of
+redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
 densities, extinction curves, etc.  A number of independent papers have been
@@
 GRB afterglow spectrum (Selsing et al., in preparation).
-
-%SWIFT LIMITATIONS
-%There is some evidence that the burst duration, $T_{90}$, does not scale with
-%redshift as expected from time-dilation \citep{Kocevski2013, Littlejohns2013a},
-%suggesting that we could be missing an increasing fraction of the bursts at
-%higher redshifts, but \citet{Lien2016} argues that we should only be missing the
-%bright "short" GRBs", but to the degree that \textit{Swift} is complete, we can
-%test if the sample presented here is complete.
 
@@
 in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the S/N of
-the afterglow continuum is poor especially in the UVB arm, why no clear
+the afterglow continuum is low especially in the UVB arm, why no clear
 absorption lines are detected against the afterglow continuum, although see
@@
 of X-shooter is depressed with respect to normal operations, resulting in a
-poorer throughout. Additionally, the position of the trace on the slit moves due
+lower throughout. Additionally, the position of the trace on the slit moves due
 to atmospheric differential refraction.
@@
 observations of the host of GRB~120211A has been obtained, starting 2013.02.17,
-$> 1 year$ after the burst has faded. A redshift for this object has been
+$> 1$ year after the burst has faded. A redshift for this object has been
 reported by \citet{Kruhler2015} and the features seen by those authors are
@@
 \subsection{GRB~120224A (z = 1.10 NEW!!!)}
-The data presented here has formed the basis of GCN \#
-12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, and has also
+
+The data presented here has formed the basis of GCN
+\#12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, and has also
 been published in \citet{Kruhler2015}. Starting 19.8 hours after the GRB
 trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting at
-$\sim$ 700 nm~and extending all the way through 2500 nm. We detect a $\sim
-2 \sigma$ emission line which, if interpreted as \ha, gives $z = 1.10$,
-supporting the redshift reported by \citet{Kruhler2015}.
+$\sim$ 700 nm~and extending all the way through 2500 nm. In the 2D-spectrum we
+detect a $\sim 2 \sigma$ emission line which, if interpreted as \ha, gives $z =
+1.10$, supporting the photometric redshift ($0.9 < z_\mathrm{phot} < 1.3$)
+derived by \citet{Kruhler2015}.
 
@@
 unless the GRB ocurred in a merging system. The extended emission in \ha,
-supports this interpretation. This burst is not apart of the statistical sample.
+supports this interpretation. This burst is not a part of the statistical sample.
 
@@
 \subsection{GRB~120815A (z = 2.358)}
+
 Not a part of the statistical sample, this burst also formed the basis of  GCN
@@
 include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
-\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
-fine structure lines from \NIii and \feii are exited local to the GRB.
-Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
+\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally fine
+structure lines from \NIii and \feii are exited local to the GRB. Intervening
+systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
 
 \subsection{GRB~120909A (z = 3.929)}
-The data presented here has formed the basis of GCN \#13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
+
+The data presented here has formed the basis of GCN
+\#13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
 published otherwise. A very rapid follow-up, starting only 1.7 hours after the
-BAT trigger, this 1.2 ks observation captures a very bright afterglow
-continuum, starting at 450 nm, signifying the onset of the Lyman limit for a 
-system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
-prominent absorption features in the form of \lya, \lyb, and \lyg, visible in
-the Lyman alpha forest. Metal absorption lines arising from \feii, \NIii,
-\SIii, \sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along
-with the corresponding fine structure lines from (\feii$^*$, \SIii$^*$,
-\oi$^*$, \oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
-
-\subsection{GRB~120923A (z = $\gtrsim8$?)}
+BAT trigger, this 1.2 ks observation captures a very bright afterglow continuum,
+starting at 450 nm, signifying the onset of the Lyman limit for a system at $z =
+3.929$. Absorption from high-column density hydrogen leaves very prominent
+absorption features in the form of \lya, \lyb, and \lyg, visible in the Lyman
+alpha forest. Metal absorption lines arising from \feii, \NIii, \SIii, \sii,
+\alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
+corresponding fine structure lines from (\feii$^*$, \SIii$^*$, \oi$^*$,
+\oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
+
+\subsection{GRB~120923A (z = 7.84)}
 
@@
 \#13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
-published otherwise. Starting 69.6 hours after the GRB trigger, that we detect
-the afterglow continuum a so high significance in all arms with 8.4 ks
-integration, testifies to the brightness of this burst. The concurrent
-identification of emission lines from \oiii~and absorption from \civ, \alii,
-\aliii, \mgi, \mgii, and \feii, tightly constrains the redshift of the burst to
-be $(z = 1.773$
+published otherwise. Starting 69.6 hours after the GRB trigger, we detect the
+afterglow continuum at high significance in all arms with 8.4 ks integration,
+testifying to the brightness of this burst. The concurrent identification of
+emission lines from \oiii~and absorption from \civ, \alii, \aliii, \mgi, \mgii,
+and \feii, tightly constrains the redshift of the burst to be $z = 1.773$.
 
@@
 afterglow reveal an unusual steep extinction curve which is found to be caused
-by dust immediate to the burst.
+by dust very close to the burst.
 

### 2018-01-26 py/NH_distribution_plot.py

@@ -40,9 +40,9 @@ def main():
     burst_table = pd.read_csv("../data/Burst list - HI columns.csv")
     name, z, NH, NHe, pub = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values, burst_table["Sample"].values
     XSGRB = (pub == "Selsing2017")
-    print(np.sum((~XSGRB).astype("int")))
-    print(len(name))
-    exit()
+    # print(np.sum((~XSGRB).astype("int")))
+    # print(len(name))
+    # exit()
 
 
     g = sns.jointplot(x=Noter_z, y=Noter_NH, kind="hex", stat_func=None, xlim = (1, 7), ylim = (20, 23), marginal_kws = {"kde": False, "norm_hist": True}, label="Noterdaeme et al. 2012b", space=0)

### 2018-01-26 py/spec_plot.py

@@ -66,8 +66,8 @@ def main():
             lineid_plot.plot_line_ids(wl[cut], flux[cut], linelist[idx]*(1 + pp), linenames[idx], arrow_tip=1.2, box_loc=1.6, ax = ax)
 
 
-        ax.plot(wl[cut][::10], signal.medfilt(error[cut][::10], 11), color=sns.color_palette()[0], alpha=0.9, zorder=1)
-        ax.axhline(1, linestyle="dashed", color=sns.color_palette()[2], alpha=0.9, zorder=10)
+        ax.plot(wl[cut][::10], signal.medfilt(error[cut][::10], 11), color=sns.color_palette("muted")[0], alpha=0.9, zorder=1)
+        ax.axhline(1, linestyle="dashed", color=sns.color_palette("muted")[2], alpha=0.9, zorder=10)
         ax.axhline(0, linestyle="dotted", color="black", alpha=0.7)
         ax.set_xlim(min(wl[cut]), max(wl[cut]))
         ax.set_ylim(-0.2, 2)
@@ -75,10 +75,14 @@ def main():
         # ax.spines['top'].set_visible(False)
         ax.yaxis.set_ticks_position('left')
         ax.xaxis.set_ticks_position('bottom')
+        # ax.axvspan(12600, 12800, color = "grey", alpha = 0.4)
+        ax.axvspan(13500, 14500, color = "grey", alpha = 0.4)
+        ax.axvspan(18000, 19500, color = "grey", alpha = 0.4)
     # Save figure for tex
     # pl.legend()
     ax4.set_ylabel(r"$F_{\lambda}\,\rm{(10^{-17}\,erg\,s^{-1}\,cm^{-2}\, \AA^{-1})}$")
-    ax7.set_xlabel(r"$\rm{Observed\,wavelength\, (\AA)}$")
+    # ax7.set_xlabel(r"$\rm{Observed\,wavelength\,}$ (\AA)")
+    ax7.set_xlabel(r"Observed Wavelength (\AA)")
     pl.tight_layout()
     pl.subplots_adjust(hspace=0.2)
     pl.savefig("../document/figures/GRB121024A.pdf", dpi="figure", rasterize=True)

### 2018-01-28 document/XSGRB_sample.tex
Jesper and notes on bursts

@@
 \thanks{Based on observations collected at the European Southern Observatory, Paranal, 
-	Chile, Program ID: 098.A-0055, 097.A-0036, 096.A-0079, 095.B-0811(B), 
+	Chile, Program IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.B-0811(B), 
 	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260}}
+	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.}}
 
@@
 
-\abstract{In this work we present the spectra of all $\gamma$-ray burst (GRB)
-	afterglows that have been promptly observed at the VLT/X-shooter up to
+\abstract{In this work we present spectra of all $\gamma$-ray burst (GRB)
+	afterglows that have been promptly observed with VLT/X-shooter up to
 	\termdate. This totals spectroscopic observations of 102 individual bursts
@@
 	find that the sample presented here is representative of the full \textit{Swift}
-	sample. For all bursts (39) for which it is possible, we provide a measurement
+	sample. For all bursts (41) for which it is possible, we provide a measurement
 	of the neutral hydrogen column density, increasing the total number of published
@@
 
-$\gamma$-ray bursts (GRBs) provide the opportunity to derive new constraints on
+$\gamma$-ray bursts (GRBs) provide the opportunity to derive constraints on
 a very wide range of topics in astrophysics. Examples range from small scale
@@
 \citep{Gehrels2004, Gehrels2009}, which was launched in 2004, has made it
-possible to harvest much of the great potential in using GRBs as probes of
-galaxies in the intergalactic medium, which was already hinted at by results
-from earlier missions \citep[e.g.,][]{VanParadijs2000, Ricker2004}. With the two
-on-board instruments, the Burst Alert Telescope (BAT) \citep{Barthelmy2005} and
-the X-Ray Telescope (XRT) \citep{Burrows2005}, \swift~is an ideal telescope for
-GRB hunting. A crucial aspect of the success of the {\it Swift} mission has been
-the extensive ground based follow-up observations of the afterglows and of the
-host galaxies of the GRBs, involving a large community of researchers. This
-fruitful collaboration has been greatly facilitated by the open data access
-policy of the {\it Swift} mission. The close collaboration between detection
-facilities and electromagnetic  follow-up campaigns continue to be immensely
-rewarding, as recently highlighted by the simultaneous detection of
-gravitational waves and light from the neutron star merger in the shape of
-GW170817/GRB180817A/AT2017gfo \citep{LIGOScientificCollaboration2017a,
-	LIGOScientificCollaboration2017}.
+possible to harvest much of the great potential in using GRBs as probes of the
+intergalactic medium, which was already hinted at by results from earlier
+missions \citep[e.g.,][]{Paradijs2000, Ricker2004}. With the two on-board
+instruments, the Burst Alert Telescope (BAT) \citep{Barthelmy2005} and the X-Ray
+Telescope (XRT) \citep{Burrows2005}, \swift~is an ideal telescope for GRB
+hunting. A crucial aspect of the success of the {\it Swift} mission has been the
+extensive ground based follow-up observations of the afterglows and of the host
+galaxies of the GRBs, involving a large community of researchers. This fruitful
+collaboration has been facilitated by the open data access policy of the
+{\it Swift} mission. The close collaboration between detection facilities and
+electromagnetic  follow-up campaigns continue to be immensely rewarding, as
+recently highlighted by the simultaneous detection of gravitational waves and
+light from the neutron star merger in the shape of GW170817/GRB180817A/AT2017gfo
+\citep{LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
 
@@
 
-The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
-instruments at the ESO Very Large Telescope (VLT). It is a high-throughput
-echelle spectrograph covering the full spectral range from the UV atmospheric
-cutoff around 300 nm to the near-infrared (near-IR) K-band ending around 2500 nm
-with three spectral arms (UVB, VIS, NIR). It was designed very much with
-transient source follow-up in mind as the fading luminosities of such sources
-makes it urgent to secure as extensive coverage as possible in the shortest
-possible time. At the same time, the resolution was designed to be in the range
-4000--9000 in order to be able to get a large useful spectral range between the
-many sky-background emission lines in the red and near-IR spectral ranges. The
-large spectral coverage allows for the spectroscopic observations of the
-highest redshift GRBs.
+The X-shooter spectrograph \citep{Vernet2011} was the first of the 2nd
+generation instruments at the ESO Very Large Telescope (VLT). It was designed
+very much with transient source follow-up in mind as the fading luminosities of
+such sources makes it urgent to secure as extensive coverage as possible in the
+shortest possible time. At the same time, the resolution was designed to be in
+the range 4000--9000 in order to be able to get a large useful spectral range
+between the many sky-background emission lines in the red and near-IR spectral
+ranges. The large spectral coverage allows for the spectroscopic observations of
+the highest redshift GRBs.
 
@@
 completeness of the finally realized sample. Finally, we offer our conclusions
-in Sect.~\ref{conclusions}.
+in Sect.~\ref{conclusions}. We use the $\Lambda$CDM cosmology parameters
+provided by \citet{Planck2015} in which the universe is flat with $H_0 =
+67.7$\,km\,s$^{-1}$ Mpc$^{-1}$ and $\Omega_\mathrm{m} = 0.307$.
 
@@
 	\item XRT started observing within 10 minutes after the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 minutes at least 30 degrees above the horizon, with the Sun below -12 degrees.
-	\footnote{Note that in the P84 proposal the criteria have
+	\item The target must be visible for at least 60 minutes, 30 degrees above the horizon, with the Sun below $-12$ degrees\footnote{Note that in the P84 proposal the criteria have
 	been stated a bit differently, the visibility constraint being replaced by a
-	declination + Sun angle constraint. The below criteria are however those
-	defining the sample.}
+	declination + Sun angle constraint. These criteria are, however, those
+	defining the sample.}.
 	\item No bright, nearby stars within $ 1.8 + 0.4 \times exp[(20 - R)/2.05$] arcsec, where R is the USNO magnitude of the star)
@@
 with the sun below $-12$ degrees. This secures time for the spectroscopic
-observations to be completed. Since the GRB population is isotropically distributed
-on the sky as seen from Earth and because the GRB properties do not depend on
-position on the sky \citep{Meegan1992, Briggs1996}, this cut does not influence our ability
-to fairly sample the underlying GRB population. The same arguments apply to
-the requirement that there are no nearby bright stars. We additionally require
-that the Galactic extinction is below $A_V \lesssim 0.5$ mag, based on the
-extinction maps by \citet{Schlegel1998}. Because of isotropy, these additional
-cuts should not influence the optical properties of the bursts themselves, only
-our ability to successfully secure observations that allows us to investigate
-the spectroscopic properties of GRBs.
-
-We further discuss the effect of these selection criteria and their implication
-for the completeness of the sample in Sect. \ref{completeness}.
+observations to be completed. Since the GRB population is isotropically
+distributed on the sky as seen from Earth and because the GRB properties do not
+depend on position on the sky \citep{Meegan1992, Briggs1996, Ukwatta2016}, this
+cut does not influence our ability to fairly sample the underlying GRB
+population. The same arguments apply to the requirement that there are no nearby
+bright stars. We additionally require that the Galactic extinction is below $A_V
+\lesssim 0.5$ mag, based on the extinction maps by \citet{Schlegel1998,
+	Schlafly2011}. Due to the isotropic distribution of GRBs on the sky, these
+additional cuts should not influence the optical properties of the bursts
+themselves, only our ability to successfully secure the observations that allows
+us to investigate the spectroscopic properties of GRBs.
+
+Bursts that fulfill these selection criteria are what we define as our
+\textit{statistical} sample. Using this sample we will be able to address
+population properties of \swift-detected bursts. We further discuss the effect
+of these selection criteria and their implication for the completeness of the
+sample in Sect. \ref{completeness}.
 
@@
 	\item Error radius < 60\arcsec.
-	\item No limit on declination. 
-	\item No constraint on Moon phase or distance. 
-	\item Elevation in the sky > 22$^\circ$ (both at trigger time and 15 min after).
-	\item Sun elevation < -12$^\circ$ (both at trigger time and 15 min after).
+%	\item No limit on declination. 
+%	\item No constraint on Moon phase or distance. 
+	\item Elevation in the sky $> 22^\circ$ (both at trigger time and 15 min after).
+	\item Sun elevation $< -12^\circ$ (both at trigger time and 15 min after).
 	\item The X-ray position must be available less than 1 hr after the GRB.
@@
 There are 8 GRBs in the sample that have beem being observed with the RRM mode.
-One of the RRM triggers is outside the statistical sample and two of the RRM
-trigger are on short GRBs. The fastest response between \swift~GRB trigger and
-the beginning of spectroscopic integration is for GRB~160410A, for which the
-delay was only 8.4 minutes.
+One of the RRM triggers is outside the sample we use to address the statistical
+properties of GRB afterglows and two of the RRM trigger are on short GRBs. The
+fastest response between \swift~GRB trigger and the beginning of spectroscopic
+integration is for GRB~160410A, for which the delay was only 8.4 minutes.
 
@@
 	indicate bursts with measured redshifts and the circles indicate those without.
-	The empty stars/circles are the positions of all the \textit{Swift} bursts
-	detected until the end of 2016. A blue filling of the symbol indicates the
+	The empty stars/circles are the positions of all the 1266 \textit{Swift} bursts
+	detected until \termdate. A blue filling of the symbol indicates the
 	position of bursts fulfilling the sample criteria specified in Sect.
@@
 	presented in \citet{Schlegel1998}. Please note that we removed the background
-	where the sample criteria cut ($A_V \lesssim 0.5$ mag) and replaced it with a
+	where the sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a
 	white color visible at the equator. The colorbar below indicates the value of
@@
 commissioning or science verification (GRB~090313, GRB~090530, GRB~090809,
-GRB~090926) are not a part of our statistical sample. The first burst observed
-after science verification and X-shooter was moved to UT2, is GRB~091018, which
-thereby constitute the first burst entering our statistical sample. For all
-bursts that fulfill our sample selection criteria, described in Sect.
-\ref{samplecrit}, spectroscopic follow-up has been attempted with X-shooter.
-Various conditions can affect our ability to follow up a given burst, and a
-discussion of these conditions and their consequences for the sample is
-included in Sect. \ref{badbursts}.
+GRB~090926) are not a part of the sample we use to investigate the statistical
+properties of GRB afterglows. The first burst observed after science
+verification and X-shooter was moved to UT2, is GRB~091018, which is the first
+burst entering our statistically homogeneous sample. For all bursts that fulfill
+our sample selection criteria, described in Sect. \ref{samplecrit},
+spectroscopic follow-up has been attempted with X-shooter. Various conditions
+can affect our ability to follow up a given burst, and a discussion of these
+conditions and their consequences for the sample is included in Sect.
+\ref{badbursts}.
 
@@
 \ref{tab:sample_overview} and plot the positions of all the bursts on the
-celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from
+celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from the
 central zone of avoidance due to high Galactic extinction cutoff (marked in
-white), the GRB positions have an isotropic distribution except in the
-upper left quadrant, which cannot be observed from Paranal.
+white), the GRB positions have an isotropic distribution, except in the upper
+left quadrant which cannot be observed from Paranal.
 
 30 percent of the spectra presented here, primarily host observations, have
-already been published in \citet{Kruhler2015}. We include independent
-reductions of them here for completeness.
+already been published in \citet{Kruhler2015}. Single bursts have additionally
+been published, based on unusual properties in their afterglows. We present
+individual notes on all the X-shooter spectra and their previous use in App.
+\ref{notes}. We include independent reductions of them here for completeness.
 
@@
 All post-processing scripts developed for this dataset are made publicly
-available at \url{https://github.com/jselsing/XSGRB_reduction_scripts}, along
-with instructions of use.
+available at \url{https://github.com/jselsing/XSGRB_reduction_scripts}.
 
@@
 spectrophotometric standards \citep{Vernet2010, Hamuy1994}, mirror flexure
-compensation, sky-subtraction and lastly the rectification and merging
-of the orders. Errors and bad pixel maps are propagated throughout the
-extraction. For the initial sky-subtraction, the background has been estimated
-by a running median in regions adjacent to the object trace clear of
-contaminating sources. Due to the broken ADC, for some objects there is
-curvature in the object trace along the dispersion axis of the slit. This means
-that for some bursts, the initial sky-estimate has been made from a limited
-number of pixels in the spatial direction. By doing a subtraction of
-the sky background on the un-rectified image we ensure that the bulk of the sky
-background is not redistributed by the rectification process.
+compensation, sky-subtraction and lastly the rectification and merging of the
+orders. Errors and bad pixel maps are propagated throughout the extraction. For
+the initial sky-subtraction, the background has been estimated by a running
+median in regions adjacent to the object trace clear of contaminating sources.
+Due to the broken ADC, for some objects there is curvature in the object trace
+along the dispersion axis of the slit. This means that for some bursts, the
+initial sky-estimate has been made from a limited number of pixels in the
+spatial direction. The subtraction of the sky background on the un-rectified
+image ensures that the bulk of the sky background is not redistributed by the
+rectification process.
 
 The image is rectified onto an equidistant grid with a dispersion sampling of
-0.02 nm/pixel and a 0.16 "/pixel spatial sampling for the UVB and VIS arm and
-0.06 nm/pixel with a 0.21 "/pixel in the NIR arm.  Because the tabulated
-resolution is a lower limit, we choose a sampling of 0.02 nm/pixel to ensure
-that the lowest wavelength part of neither of the arms have a sampling lower
-than the Nyquist sampling rate of 2 pixels per resolution FWHM.
+0.02 nm/pixel and a 0\farcs16 per pixel spatial sampling for the UVB and VIS
+arm and 0.06 nm/pixel with a 0\farcs21 per pixel in the NIR arm.  Because
+the tabulated resolution is a lower limit, we choose a sampling of 0.02 nm/pixel
+to ensure that the lowest wavelength part of neither of the arms have a sampling
+lower than the Nyquist sampling rate of 2 pixels per resolution FWHM.
 
@@
 background. When weighting images where the noise in each pixel is dominated by
-Poisson noise it is important to estimate the background variance in a large
+Poisson noise, it is important to estimate the background variance in a large
 enough region, so that any correlation between the signal and the weights are
@@
 over the variance map produced by the pipeline, where the trace has been masked
-and width of the window is chosen to be wide enough for median variance to be
-generated on the basis of several hundred pixels. This weighting scheme
+and the width of the window is chosen to be wide enough for median variance to
+be generated on the basis of several hundred pixels. This weighting scheme
 automatically also optimally combines images of different exposure times or
@@
 Using this method for all observations we additionally get a spectrum of the
-sky, which we can use to calibrate the wavelength solution.
+sky, which we can use to recalibrate the wavelength solution.
 
@@
 Part of this correction is performed by the pipeline using the frames taken
-during the X-shooter Active Flexure Compensation procedure \footnote{X-shooter
+during the X-shooter Active Flexure Compensation procedure\footnote{X-shooter
 	User Manual available at
@@
 correct seeing PSF with which to convolve the synthetic sky an initial
-refinement of the wavelength solution has been obtained by cross-correlating
-the observed sky with an unconvolved synthetic sky. This preliminary wavelength
+refinement of the wavelength solution has been obtained by cross-correlating the
+observed sky with an unconvolved synthetic sky. This preliminary wavelength
 calibration is applied to the observed sky. The synthetic spectrum is then
@@
 VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short
-period of time were applied to the corresponding spectra. Using the convolved
+periods of time were applied to the corresponding spectra. Using the convolved
 synthetic sky, the pixels containing the brightest sky lines have been added to
@@
 have chosen the Voigt function as a model for the spectral PSF and we describe
-how this is evaluated in App. \ref{voigt}. Since additionally, the host galaxy
-could also have a contribution to the image profile, this choice allows for the
-required freedom if additional flux is in the wings of the profile.
-
-To guide the guess position of the trace on the slit as a function of
+how this is evaluated in App. \ref{voigt}. Since the host galaxy could also give
+a contribution to the image profile, this choice allows for the required freedom
+if additional flux is in the wings of the profile.
+
+To guide the first guess position of the trace on the slit as a function of
 wavelength, we have used the analytic prescription for the trace position
@@
 \citet{Schlafly2011}\footnote{Queried from
-	http://irsa.ipac.caltech.edu/applications/DUST/index.html using
-	\citet{astroquery}}, and the extinction curve by \cite{Cardelli1989} with a
+	http://irsa.ipac.caltech.edu/applications/DUST/index.html using \texttt{astroquery}
+	\citep{astroquery}.}, and the extinction curve by \cite{Cardelli1989} with a
 total to selective extinction $R_V = 3.1$. The wavelengths of the extracted
@@
 	Kausch2015}\footnote{\url{http://www.eso.org/sci/software/pipelines/skytools/molecfit}}), which can then be used to correct for the absorption.
-The accuracy of the correction depends on the S/N per pixel for the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
+The accuracy of the correction depends on the S/N per pixel of the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
 
 If the afterglow is not sufficiently bright, telluric standard stars observed
-close in the time to the GRB can be used as a proxy for the atmospheric
+close in time to the GRB can be used as a proxy for the atmospheric
 condition during the GRB observation. Here we employ the telluric correction
@@
 In the last case, where the object is neither bright enough, or there for some
-reason have not been observed a telluric standard, we rely on the synthetic sky
-model by \citet{Noll2012} and \citet{Jones2013} for which we generate a
+reason have not been observed a telluric standard, we rely on a synthetic sky
+model \citep{Noll2012, Jones2013} for which we generate a
 synthetic transmission spectrum, where the ambient parameters for the
@@
 	absorption from molecular $\mathrm{H_2}$. The absorption trough visible at
-	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have overplot the most
-	prominent lines seen in GRB afterglows from \citet{Christensen2011a}. The
-	regions of most severe telluric absorption is highlighted by a grey-shading the
-	background. Additionally, three intervening systems are seen in this sightline.
-	This spectrum has been analyzed in detail in \citet{Friis2015}.}
+	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have marked the most prominent
+	lines seen in GRB afterglows from \citet{Christensen2011a}. The regions of most
+	severe telluric absorption is highlighted by a grey-shading the background.
+	Additionally, three intervening systems are seen in this sightline. This
+	spectrum has been analyzed in detail in \citet{Friis2015}.}
 \label{fig:spectrum}
@@
 flux-density of each point is determined by the median value of the spectrum in
-a small region ($\sim 1 \mathrm{\AA}$) surrounding each point. The points are
+a small region ($\sim 1~\mathrm{\AA}$) surrounding each point. The points are
 fitted with a low order polynomial (we use \texttt{numpy.polynomial.chebyshev})
-and iteratively, the point furthest away are removed until the polynomial fit
+and iteratively, the point furthest away is removed until the polynomial fit
 differ from the points by less than $\sim$ 5 per cent. This filtering is used
-because the afterglow continuum of GRBs are power laws \citep{Piran2005}, and by
-removing points that differ significantly from a smooth continuum shape will
-guide the continuum estimate to a shape more reasonable for GRB afterglows.
-Additionally, points spaced less than around 1 per cent of the entire spectral
-coverage are pruned. The remaining filtered and pruned points are then spline
-interpolated using \texttt{scipy.interpolate.splrep.splev}, which serves as a
-first estimate of the continuum placement. An attempt to identify absorption and
-emission regions are then carried out, where they are marked as such if the
-difference between the estimated continuum and the observed spectrum is larger
-than 3 - 5 times the associated error spectrum. All regions marked as affected
-are then masked.
+because the intrinsic afterglow continuum of GRBs can be modeled by power laws
+\citep{Piran2005}, and by removing points that differ significantly from a
+smooth continuum shape will guide the continuum estimate to a shape more
+reasonable for GRB afterglows. Additionally, points spaced closer than 1 per
+cent of the total spectral coverage are pruned. The remaining filtered and
+pruned points are then spline interpolated using
+\texttt{scipy.interpolate.splrep.splev}, which serves as a first estimate of the
+continuum placement. An attempt to identify absorption and emission regions are
+then carried out, where they are marked as such if the difference between the
+estimated continuum and the observed spectrum is larger than 3 - 5 times the
+associated error spectrum. All regions marked as affected are then masked.
 
@@
 	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-	\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric absorption lines in the VIS spectra, as a function of the FWHM of a Gaussian fit onto the trace in spatial direction at  792 nm.
-The lower horizontal axis is in units pixels, the top axis in arc seconds.
-The red datapoints show a subsample of NIR spectra.
-		The grey line shows a linear fit to the VIS datapoints. }
-	\label{fig:res}
+\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
+	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
+	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
+	horizontal axis is in units pixels, the top axis in arc seconds. The red
+	datapoints show a subsample of measurements obtained for NIR spectra. The grey
+	line shows a linear fit to the VIS datapoints. } \label{fig:res}
 \end{figure}
@@
 possibility to tweak slit widths to the seeing at the time of observations (i.e.
-to optimize spectral resolution and signal to noise), and almost all our data is
+to optimize spectral resolution and signal to noise), and almost all our data are
 therefore taken with a fixed set of slit widths and binning, described above. In
@@
 considerably smaller than the silt width, and the delivered spectral resolution
-will then be determined by the seeing rather than slit width, as afterglows are
+will then be determined by the seeing rather than by the slit width, as afterglows are
 point sources (this is evidently not the case for extended sources, e.g for host
@@
 given on the ESO instrument website. The UVB arm contains no suitable absorption
-lines to use, and we therefore use a scaled value as in the NIR arm. While this
-simple method is perfect (the spatial profile of the trace is not a perfect
-Gaussian), it nevertheless gives a sufficiently accurate estimate for the
-analysis of these low signal to noise science spectra.
+lines to use, and we therefore use a scaled value as in the NIR arm. This simple
+method is perfect (the spatial profile of the trace is not a perfect Gaussian),
+it nevertheless gives a sufficiently accurate estimate for the analysis of these
+low signal to noise science spectra.
 
@@
 \footnote{\url{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}}. This
-release includes for completeness the spectra of GRBs outside the statistical
-sample and includes both prompt afterglow observations as well as late time
+release includes both prompt afterglow observations as well as late time
 observations of the associated hosts and represents \textit{all} afterglow
@@
 period of the program 098.A-0055, \termdate~and thus constitutes 8 years of GRB
-afterglow observations with X-shooter. An overview of all the spectra provided
-here are given in Table. \ref{tab:sample_overview}. For each burst, each
-individual observation is provided in a separate reduction and in cases where
-observations have been repeated for an increased S/N or to follow the temporal
-evolution, a combined spectrum is also provided. No attempt has been made to
-join the spectroscopic arms, so for each observation, three spectra are provided
-in separate files.
+afterglow observations with X-shooter. An overview of all the spectra and their
+observational setups are given in Table. \ref{tab:sample_overview}. For each
+burst, each individual observation is provided in a separate reduction and in
+cases where observations have been repeated for an increased S/N or to follow
+the temporal evolution, a combined spectrum is also provided. No attempt has
+been made to join the spectroscopic arms, so for each observation, three spectra
+are provided in separate files.
 
@@
 
-Each file contains 7 columns with following contents and descriptions:
+Each file contains 7 columns with the following contents and descriptions:
 \begin{itemize}
@@
 sample criteria have been followed up due to interesting characteristics, e.g.
-curiosity properties of their light curves, their brightness, etc. These bursts
+curios properties of their light curves, their brightness, etc. These bursts
 are not discussed as part of the investigation of the statistical properties of
@@
 spectroscopic observation. The shortest delays are observed in RRM-mode. The
-fastest follow-up between BAT trigger and start of spectroscopic observations in
-our sample is for the short, $z = 1.717$, GRB~160410A for which spectroscopic
+fastest follow-up between BAT trigger and start of spectroscopic observations for any observation is for the short, $z = 1.717$, GRB~160410A for which spectroscopic
 integration was initiated only 8.4 minutes after the BAT trigger. To illustrate
@@
 sample. As can be seen from the figure, the fraction of GRBs with a redshift
-determination decreases with follow-up delay. The overall redshift completeness
+determination decreases with follow-up delay. The redshift completeness
 for bursts that we have followed up is 92 per cent. This degree of completeness
@@
 	\centerline{\includegraphics[width=9cm]{figures/timing.pdf}} \caption{Afterglow
-	magnitude at the start of observation as a function of follow-up delay. The
-	points have been colored based on the redshift of the corresponding burst. Red
-	crosses indicate GRBs without a measured redshift and arrows indicate bursts
-	for which the afterglow was not detected in the acquisition image. In red is
-	shown the redshift completeness as a function of follow-up delay.}
-\label{fig:timing}
+	magnitude at the start of observation as a function of follow-up delay for all
+	the afterglows that have been followed up. The points have been colored based
+	on the redshift of the corresponding burst. Red crosses indicate GRBs without a
+	measured redshift and arrows indicate bursts for which the afterglow was not
+	detected in the acquisition image. In red is shown the redshift completeness as
+	a function of follow-up delay.} \label{fig:timing}
 \end{figure}
@@
 
-Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
-bursts have triggered \textit{Swift} in the period since the commissioning of
-VLT/X-shooter. This sample constitutes the \textit{statistical sample} from
-which we will derive statistical properties of the GRB population. The redshift
-completeness of the full statistical sample is 61 per cent. We return to the
-question of redshift completeness in Sect. \ref{redshift}. From this sample, 91
-GRBs have been spectroscopically followed up with X-shooter. In order to assess
-whether the subset of bursts followed up are representative of the underlying
-GRB parent population, we compare intrinsic properties of GRBs in our sample to
-GRBs in the full sample followed up by \textit{Swift}. We show the comparison
-between the BAT (15 -- 150 keV) fluence, the XRT flux (0.3 -- 10 keV) at 11
-hours, and the intrinsic X-ray derived equivalent hydrogen column density,
-$N_{\mathrm{HI,X}}$, in Fig. \ref{fig:swift_complete}. For the latter, we can
-only use values of $N_{\mathrm{HI,X}}$ derived for bursts with a measured
-redshift, excluding $\sim 75$ per cent of the full \swift~sample. We return to
-the last point in Sect. \ref{darkness}.
+A total of 158 bursts fulfill the sample criteria specified in Sect.
+\ref{samplecrit},  since the commissioning of VLT/X-shooter. This sample
+constitutes the \textit{statistical sample} from which we will derive
+statistical properties of the GRB population. The redshift completeness of the
+full statistical sample is 61 per cent. We return to the question of redshift
+completeness in Sect. \ref{redshift}. From this sample, 91 GRBs have been
+spectroscopically followed up with X-shooter. In order to assess whether the
+subset of bursts followed up are representative of the underlying GRB parent
+population, we compare intrinsic properties of GRBs in our sample to GRBs in the
+full sample followed up by \textit{Swift}. We show the comparison between the
+BAT (15 -- 150 keV) fluence, the XRT flux (0.3 -- 10 keV) at 11 hours, and the
+intrinsic X-ray derived equivalent hydrogen column density, $N_{\mathrm{HI,X}}$,
+in Fig. \ref{fig:swift_complete}. For the latter, we can only use values of
+$N_{\mathrm{HI,X}}$ derived for bursts with a measured redshift, excluding $\sim
+75$ per cent of the full \swift~sample. We return to the last point in Sect.
+\ref{darkness}.
 
@@
 Additionally, using a 2-sided Kolmogorov-Smirnov test (KS-test), we can assess
-the degree to which the null, that the two distributions are drawn from the same
-parent distribution, can be rejected. We show a graphic representation of the
-test statistics in Fig. \ref{fig:p_values}. A high p-value indicates little
-evidence against the null, meaning for the properties investigated here. The
-distribution of $N_{\mathrm{HI,X}}$ exhibits the highest degree of
-dissimilarity, but the two distribution are still consistent with being drawn
-from the same underlying distribution.
+the degree to which the null hypothesis, that the two distributions are drawn
+from the same parent distribution, can be rejected. We show a graphic
+representation of the test statistics in Fig. \ref{fig:p_values}. A high p-value
+indicates little evidence against the null hypothesis . The distribution of
+$N_{\mathrm{HI,X}}$ exhibits the highest degree of dissimilarity, but the two
+distribution are still consistent with being drawn from the same underlying
+distribution.
 
@@
 	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
-\caption{Relational graph showing the respective p-values that the different
-	samples are drawn from the same underlying distribution. The arrows represent
-	the comparison direction with each of the p-values listed next to. Only in the
-	HI column density distribution is there mild evidence against the null, but the
-	discrepancy is mainly driven by a relatively larger fractional contribution
-	from low-column hosts.} \label{fig:p_values}
+\caption{Relational graph showing the respective p-values. They all represent
+	the degree that the different samples are drawn from the same underlying
+	distribution. The arrows represent the comparison direction with each of the
+	p-values they are listed next to. Only in the HI column density distribution is
+	there mild evidence against the null hypothesis, but the discrepancy is mainly
+	driven by a relatively larger fractional contribution from low-column hosts in
+	the statistical sample.} \label{fig:p_values}
 \end{figure}
@@
 10^{-8}~\mathrm{erg}~\mathrm{s}^{-1}~\mathrm{cm}^{-2}$;
-\citealt{Baumgartner2013, Lien2016}). Due to complex triggering mechanism of
+\citealt{Baumgartner2013, Lien2016}). Due to the complex triggering mechanism of
 \swift, this sensitivity limit should be interpreted with some caution as the
@@
 bursts has been investigated by \citet{Turpin2016} which find that we
-systematically find redshift for longer GRBs. Additionally only the brightest
-GRBs are seen above redshift $z \gtrsim 1$ as shown in Fig. \ref{fig:z}. As
-described in detail in \citet{Hjorth2012} and \citet{Perley2016b}, bursts for which the redshift is
-measured from the afterglow are systematically found for host galaxies with a
-lower luminosity than burst for which the redshift is measured from the host
-galaxy. Only a few GRBs hosted in galaxies, with stellar masses more than
+systematically are able to measure redshifts for longer duration GRBs.
+Additionally only the brightest GRBs are seen above redshift $z \gtrsim 1$ as
+shown in Fig. \ref{fig:z}. As described in detail in \citet{Hjorth2012} and
+\citet{Perley2016b}, bursts for which the redshift is measured from the
+afterglow are systematically found for host galaxies with a lower luminosity
+than burst for which the redshift is measured from the host galaxy. Only a few
+GRBs hosted in galaxies, with stellar masses more than
 $10^{10}~\mathrm{M}_\odot$ have the redshift measured based on the afterglow
@@
 	luminosity, $L_{\mathrm{BAT}}$. Bursts that are a part of the statistical
-	sample are marked by blue stars and black dots show all GRBs observed with
+	sample are marked by blue stars whereas black dots show all GRBs observed with
 	X-shooter. All \swift~GRBs with measured redshifts are shown in grey. For
@@
 
-A large fraction of all GRBs exhibit no detectable or very faint optical
-afterglows \citep{Groot1998, Djorgovski2001, Fynbo2001}. The degree of optical
-extinction relative to the X-ray brightness has been parametrized in term of
-their optical darkness, $\beta_{OX}$ \citep{Jakobsson2004, Rol2005,
-	VanderHorst2009}. The X-ray properties of such bursts have previously been
-investigated \citep{DePasquale2003, Fynbo2009, Melandri2012} and there are some
-indications that dark bursts have somewhat higher X-ray luminosity and
-$N_{\mathrm{HI, X}}$ compared to the optically bright bursts. This indicates
-along with investigations of host galaxy properties \citep{Greiner2011,
-	Kruhler2011, Hjorth2012, Perley2016b}, that the extinction of the optical
-afterglow is primarily driven by the presence of dust in the host galaxy.
-\citet{Hjorth2012} find that systems with no optical afterglow have higher
-$N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which however
-also turn out to be redder.
+A fraction of all GRBs exhibit no detectable or very faint optical afterglows
+\citep{Groot1998, Djorgovski2001, Fynbo2001}. The degree of optical extinction
+relative to the X-ray brightness has been parametrized in term of their optical
+darkness, $\beta_{OX}$ \citep{Jakobsson2004, Rol2005, VanderHorst2009}. The
+X-ray properties of such bursts have previously been investigated
+\citep{DePasquale2003, Fynbo2009, Melandri2012} and there are some indications
+that dark bursts have somewhat higher X-ray luminosity and $N_{\mathrm{HI, X}}$
+compared to the optically bright bursts. This indicates along with
+investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
+	Hjorth2012, Perley2016b}, that the extinction of the optical afterglows is
+primarily driven by the presence of dust in the host galaxies. \citet{Hjorth2012}
+find that systems with no optical afterglow have higher $N_{\mathrm{HI, X}}$,
+irrespective of the nature of the host -- which however also turn out to be
+redder.
 
@@
 neighboring XRT measurements of by extrapolating the last few X-ray data points
-to the spectroscopic delay. When the afterglow is not detected in the
-acquisition camera, an upper limit of $> 24$ mags is used, which propagates into
-an upper limit on $\beta_{OX}$.
+to the time of the spectroscopic observation. When the afterglow is not detected
+in the acquisition camera, an upper limit of $> 24$ mags is used, which
+propagates into an upper limit on $\beta_{OX}$.
 
@@
 \citet{Fynbo2009} have been treated as detections, meaning that we artificially
-bias the distribution towards higher $\beta_{OX}$-values. We tentatively confirm
-the result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have
-higher \nhx. Specifically, for bursts with measured redshift either from the
-afterglow or the host galaxy, we find the following: For bursts with $\beta_{OX}
-\geq 0.5$ we find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$
-we find \nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is
-contained within the error intervals. A 2-sided KS test fails to reject the null
-that they are drawn from the same distribution with $p = 0.43$, meaning there is
-no strong evidence for a discrepancy.
+bias the distribution towards higher $\beta_{OX}$-values. The two distributions
+exhibit a large degree of overlap. We cannot confirm the result by
+\citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher \nhx.
+Specifically, for bursts with measured redshift either from the afterglow or the
+host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$ we
+find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
+\nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
+within the error intervals. A 2-sided KS test fails to reject the null
+hypothesis that they are drawn from the same distribution with $p = 0.43$,
+meaning there is no strong evidence for a discrepancy. A Kendall's $\tau$ test
+suggest a non-significant, very low degree of negative correlation ($\Gamma =
+-0.15$ at a p-value = 0.10).
 
@@
 	which we do not have a redshift. Darker colors represented higher density of
-	points. The corresponding marginal distributions are shown. In blue is the
-	points for the bursts presented here along with the marginal histograms. Limits
-	on $\beta_{OX}$ is shown by downwards facing arrows.} \label{fig:betaOX}
+	points. In blue are the points for the bursts presented here along with the
+	marginal histograms. Limits on $\beta_{OX}$ are shown by downwards facing
+	arrows. The corresponding marginal distributions are shown along the edges of
+	the plot. There is no statistically significant correlation between the two
+	quantities.} \label{fig:betaOX}
 \end{figure}
@@
 measurement is more difficult to make, see \citet{Jakobsson2005, Levesque2010}
-and \citet{Perley2017} for a recent example.
+and \citet{Perley2017}.
 
@@
 rest of the literature values has \nh~$= 21.5_{-1.5}^{+0.4}$. We see that the
-two distribution has a large degree of overlap due to the large width of the
+two distributions have a large degree of overlap due to the large width of the
 distributions, but we find a slightly higher median value for the new sample
-presented here. A 2-sided KS test gets us a p-value of p = 0.006, meaning
+presented here. A 2-sided KS test gives a p-value of p = 0.006, meaning
 relatively large evidence against the null that the two samples are drawn from
@@
 
-In Fig. \ref{fig:NH_dist}, we additionally show the column density-distribution
+In Fig. \ref{fig:NH_dist}, we also show the column density-distribution
 for the 12081 quasar absorbers with \nh~$> 20$ from \citet{Noterdaeme2012b}. The
 fact that GRBs systematically are located behind the highest \nh, previously
-noted by \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this
+noted \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this
 figure. The reason for this is that quasar sample sight-lines through galaxies
@@
 \caption{Distributions of hydrogen column densities for absorbers found in
-	quasar absorption lines, from \citep{Noterdaeme2012b} in blue. Overplot in
-	green is the kernel density estimate is the similar distribution, only for
-	absorbers in GRB sightlines. Values are taken from the compilation in Tanvir et
-	al. (submitted), along with the new values presented in this sample. The
-	marginal distributions for the two distribution are also shown along the left
-	side and on the top, where the different environments probed are clearly
-	visible in the hydrogen column densities, as previously also noted in
+	quasar absorption lines, from \citet{Noterdaeme2012b} in blue. Overplotted in
+	green is the kernel density estimate of absorbers in GRB sightlines. Values are
+	taken from the compilation in Tanvir et al. (submitted), along with the new
+	values presented in this sample. We also show in the red the values derived in
+	this work. The marginal distributions for the three distributions are also shown
+	along the left side and on the top, where the different environments probed are
+	clearly visible in the hydrogen column densities, as previously also noted in
 	\citet{Fynbo2009}.} \label{fig:NH_dist}
@@
 (DLAs) towards about 10$^5$ quasars, a handful systems with
-$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
+\nh~$ > 22$ been identified \citep[e.g., five
 in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
@@
 the next generation of 20--40-m telescopes is likely to be very rewarding, given
-that the a suitable GRB detector will be online.
+that a suitable GRB detector will be available.
 
@@
 \end{equation}
-for any complex argument, $z = x + iy$. The complex probability function can be expressed as a sum of a real an imaginary part \citep{benner1995, abrarov2015b},
+for any complex argument, $z = x + iy$. The complex probability function can be expressed as a sum of a real and an imaginary part \citep{benner1995, abrarov2015b},
 \begin{equation} 
@@
 \end{equation}
-where is the real part, $\mathtt{Re}[W(x, y)] =  \sqrt{2 \pi} \sigma V(\lambda,\sigma, \gamma)$ if $x = \frac{\lambda}{\sqrt{2} \sigma}$ and $y = \frac{\gamma}{\sqrt{2} \sigma}$, which can be obtained by using the complex argument, $z = \frac{\lambda + i\gamma}{\sqrt{2} \sigma}$, in the complex probability function.
-If $\mathtt{Im}[z] \geq 0$, which is always guaranteed for the width of a spectral profile, the complex probability function equals the complex error function.
-The complex error function has numerous, fast, numerical approximations where in this work we use the \texttt{scipy.special.wofz} \citep{scipy} implementation.
+where the real part, $\mathtt{Re}[W(x, y)] =  \sqrt{2 \pi} \sigma
+V(\lambda,\sigma, \gamma)$ if $x = \frac{\lambda}{\sqrt{2} \sigma}$ and $y =
+\frac{\gamma}{\sqrt{2} \sigma}$, which can be obtained by using the complex
+argument, $z = \frac{\lambda + i\gamma}{\sqrt{2} \sigma}$, in the complex
+probability function. If $\mathtt{Im}[z] \geq 0$, which is always guaranteed for
+the width of a spectral profile, the complex probability function equals the
+complex error function. The complex error function has numerous, fast, numerical
+approximations where in this work we use the \texttt{scipy.special.wofz}
+\citep{scipy} implementation.
 
@@
 common redshift of $z = 3.373$. Two intervening systems at $z = 1.959$ and $z =
-1.800$ is identified based on strong \mgii-absorption. Because this burst is
+1.800$ are identified based on strong \mgii-absorption. Because this burst is
 observed before the instrument is science-verified, it does not enter into the
@@
 for a host at $z = 1.266$ from the detection of \mgii, \mgi, \SIii, \feii,
-\aliii. Because this burst is observed before the instrument is
+\aliii. Because this burst was observed before the instrument is
 science-verified, it does not enter into the statistical sample.
 
-\subsection{GRB~090809 (z = 2.737)}
-Observed during the first science verification period and was used for GCN
+\subsection{GRB~090809 (z = 2.737)} Observed during the first science
+verification period and was reported in GCN
 \#9761\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9761.gcn3}} and is
 additionally used as the basis for the master thesis by \'Asa Sk\'ulad\'ottir
-(2010). 7.2 ks integration starting 10.2 hours after the GRB trigger notice
-yields the clear afterglow continuum in all arms from with Lyman Limit located
-in the beginning of the UVB coverage. The simultaneous detection of absorption
+(2010). 7.2 ks integration starting 10.2 hours after the GRB trigger yields a
+clear afterglow continuum in all arms. The simultaneous detection of absorption
 lines identified as \lya, \SIii, \oi, \SIi$^*$, \SIiv, \civ, \feii, \alii,
-\aliii and \mgii at $z = 2.737$ sets it as the redshift of the GRB. Because this burst is observed before the instrument is
-science-verified, it does not enter into the statistical sample.
-
-\subsection{GRB~090926 (z = 2.106)}
-Starting during the second science verification period, this dataset forms the
-basis of GCN \#9942\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9942.gcn3}} and
-is additionally published in \citet{DElia2010}. Spectroscopic integration
-started 22 hours after the BAT trigger and from the acquistion camera the
-optical afterglow has R = 17.9 mag (vega) at the beginning of the observations
-which causes a very continuum to be seen in all arms. An absorption trough due
-to \lya is clearly visible along with numerous metal resonance lines \civ,
-\SIii, \SIii$^*$ \feii, \mgii, all at $z = 2.106$, marks this as the redshift of
-the GRB. Because this burst is observed before the instrument is
-science-verified, it does not enter into the statistical sample.
+\aliii~and \mgii~at $z = 2.737$ sets it as the redshift of the GRB. Because this
+burst is observed before the instrument was science-verified, it does not enter
+into the statistical sample.
+
+\subsection{GRB~090926 (z = 2.106)} Obtained during the second science
+verification period, this dataset forms the basis of GCN
+\#9942\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9942.gcn3}} and is
+additionally published in \citet{DElia2010}. Spectroscopic integration started
+22 hours after the BAT trigger and from the acquisition camera the optical
+afterglow is measured to R = 17.9 mag at the beginning of the observations which
+causes a strong continuum to be seen in all arms. An absorption trough due to
+\lya is clearly visible along with numerous metal resonance lines \civ, \SIii,
+\SIii$^*$ \feii, \mgii, all at $z = 2.106$. Because this burst was observed
+before the instrument is science-verified, it does not enter into the
+statistical sample.
 
@@
 hydrogen column density distribution. The blue part of the GRB continuum
-exhibits a downturn in the conitnuum level which could indicate the presence of
+exhibits a downturn in the continuum level which could indicate the presence of
 a significant amount of dust along the line-of-sight.
@@
 This spectrum forms the basis for GCN
-\#16401\footnote{\url{http://gcn.gsfc.nasa.gov/other/16401.gcn3}}, but is not
+\#16401\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16401.gcn3}}, but is not
 published elsewhere. $4\times600$ s spectroscopic integration, starting 3.8
@@
 \citep{Lien2016}) GRB spectrum has been published in GCN
-\#16437\footnote{\url{http://gcn.gsfc.nasa.gov/other/16437.gcn3}}, but does not
+\#16437\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16437.gcn3}}, but does not
 appear in the refereed literature. Spectroscopic observations began with
@@
 This spectrum has already been used to inform GCN
-\#16983\footnote{\url{http://gcn.gsfc.nasa.gov/other/16983.gcn3}}, but is not
+\#16983\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16983.gcn3}}, but is not
 published elsewhere. Starting 15.4 hours after the \swift~trigger and lasting
@@
 The gross content of the spectrum has been issued in GCN
-\#17040\footnote{\url{http://gcn.gsfc.nasa.gov/other/17040.gcn3}}, but it is not
+\#17040\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17040.gcn3}}, but it is not
 used otherwise. Starting only 1.9 hours after the BAT trigger, these
@@
 This spectrum has previously been used to form the basis for GCN
-\#17420\footnote{\url{http://gcn.gsfc.nasa.gov/other/17420.gcn3}}, but it is not
+\#17420\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17420.gcn3}}, but it is not
 published in the refereed literature. Due to the short visibility of this burst,
@@
 This spectrum has already been used to inform GCN
-\#17523\footnote{\url{http://gcn.gsfc.nasa.gov/other/17523.gcn3}}, but it is not
+\#17523\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17523.gcn3}}, but it is not
 used elsewhere. Starting 5.1 hours after the BAT trigger, this spectrum is based
@@
 This spectrum has previously been used to form the basis for GCN
-\#17672\footnote{\url{http://gcn.gsfc.nasa.gov/other/17672.gcn3}}. $4 \times
+\#17672\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17672.gcn3}}. $4 \times
 600$ s spectroscopic integration, starting 10 hours after the GRB trigger,
@@
 The gross content of these observations have previously been presented in GCN
-\#17755\footnote{\url{http://gcn.gsfc.nasa.gov/other/17755.gcn3}}, but is not
+\#17755\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17755.gcn3}}, but is not
 published as part of any refereed paper. This bona-fide short GRB (T90 (15-350
@@
 
+This spectrum is empty, but is included here for completeness. It is not
+published anywhere. $4\times600$ s spectroscopic observations, starting 3.7
+hours after the trigger does not reveal anything conclusive. The host
+association is additionally ambiguous for this, likely extincted (GCN
+\#17767\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17767.gcn3}}), GRB.
+
 \subsection{GRB~150514A (z = 0.807)}
 
+This spectrum has already been used in GCN
+\#17822\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17822.gcn3}}. Spectroscopic
+observations began 28.4 post trigger and consist of a $4 \times 600$ s nodding
+sequence. The GRB afterglow continuum is detected at moderate S/N across the
+entire spectral coverage of X-shooter. Narrow absorption features are imprinted
+in the continuum where we identify features from multiple \feii~transitions as
+well as the \mgii-doublet and the \mgi$\lambda2852.96$ resonance line. These
+lines are all found a position matching $z = 0.807$, suggesting it is the
+redshift of the GRB.
+
 \subsection{GRB~150518A (z = 0.256)}
 
+This gross content of this data has previously been issued in GCN
+\#17832\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17832.gcn3}}, but is not
+published. Starting $> 1$ day after the GRB trigger, $4 \times 600$ s
+spectroscopic integration securely allows us to measure the redshift of the
+host. A continuum is detected all the way through the spectral coverage with
+multiple emission lines superposed. It is not clear to what degree the GRB
+afterglow contributes to the continuum. We identify the emission lines as \oii,
+\hb, \oiii, \ha, \nii, \ha, and \sii~- all at $z = 0.256$. Due to the spatial
+proximity of this burst, it is a candidate for SN follow-up and indeed there are
+indications of re-brightening at the burst position (GCN
+\#17903\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17903.gcn3}}).
+
 \subsection{GRB~150616A (z = 1.188)}
 
+These host observations are taken long after the burst had faded  and have not
+been published before. They are included here for completeness. This bursts is
+excluded from the sample because an observing constraint delayed the
+\swift~slew, causing the XRT observations to begin 16 minutes post trigger. No
+continuum is detected, but emission lines from \oii, \oiii, and \ha~are all
+detected at $z = 1.188$, setting it as the redshift of the GRB.
+
 \subsection{GRB~150727A (z = 0.313)}
 
+The overall content of these observations have previously been reported in GCN
+\#18080\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18080.gcn3}}, but is not
+published. Starting 5 hours after the \swift~GRB trigger, these observations
+consists of a combined $2 \times 1200$ s and $2 \times 600$ s. A blue continuum
+is detected across the entire spectral coverage of X-shooter, suggesting a
+significant contribution from the GRB afterglow. Superposed on the continuum are
+emission lines which we identify as \hb, \oiii, and \ha~with a measured 
+redshift of z = 0.313. Supporting this redshift is the tentative detection of
+the \mgii~absorption doublet in the afterglow continuum. The relative strength
+of the lines suggest that the line-forming region is dust obscured, contrary to
+the story told be the blue afterglow continuum.
+
 \subsection{GRB~150821A (z = 0.755)}
 
-\subsection{GRB~150831A (z = na)}
+This gross content of this spectacular spectrum has already been issued in GCN
+\#18187\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18187.gcn3}}, but the
+spectrum has not been published. Starting just 12.4 minutes after the onboard
+trigger of \swift, a total of $4 \times 600$ s spectroscopic integration was
+obtained. The observations began just before dawn and the last two exposures are
+heavily affected by the brightening sky. A bright, high S/N GRB afterglow
+continuum is detected all across the the spectral window and imprinted on this
+are a myriad of absorption features. We identify individual lines from
+transitions in \aliii, \crii, \znii, \NIii$^*$, \feii, \feii$^*$, \scii, \mnii,
+\mgii, \mgi, \TIii, and  \caii~ - all at a consistent redshift of $z = 0.755$.
+The detection of fine-structure lines, exited local to the burst, clearly marks
+this as the redshift of the GRB.
 
@@
 
+This spectrum has not previously been published. Starting $\sim$ 20 hours after
+the BAT trigger, these observations were stopped after $2 \times 600$s, when it
+became apparent that only a modest S/N was obtainable, and the redshift had
+already been published( GCN
+\#18273\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18273.gcn3}}, GCN
+\#18274\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18274.gcn3}}).  The
+continuum is detected at low S/N all across X-shooter and \mgii~is detected at
+the suggested redshift, which we take as the redshift of the GRB.
+
 \subsection{GRB~150915A (z = 1.968)}
+
+The gross content of these spectra have already been issue in GCN \#
+18318\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18318.gcn3}}. Observations
+began 3.3 hours after the \swift~trigger and lasted for $4 \times 1200$ s. A
+moderate S/N afterglow continuum is detected across the entire spectral coverage
+of X-shooter, where however the bluest part is affected by \lya-forest
+absorption. Imprinted on the afterglow continuum are a wealth of both emission
+and absorption features, all caused in a system, for which we measure a redshift
+of $z = 1.968$. We identify in emission \oii, \hb, \oiii, and \ha and absorption
+due to \lya, \civ, \alii, \SIii, \feii, and \mgii. We additionally identify
+fine-structure absorption lines from \SIii$^*$ and \feii$^*$, at a similar
+redshift, which unequivocally sets the suggested redshift as the redshift of the
+GRB. The \lya-line is affected by an atmospheric transmission drop and
+\lya~emission in the trough, making the measurement of the neutral hydrogen
+column density difficult. We measure \nh~$21.2 \pm 0.3$.
 
@@
 \subsection{GRB~160203A (z = 3.517)}
+
 The data presented here also formed the basis of GCN \#

### 2018-01-28 py/NH_distribution_plot.py

@@ -1,13 +1,17 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 
-from __future__ import division, print_function
+# Plotting
+import matplotlib; matplotlib.use('TkAgg')
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import matplotlib as mpl
 
 from astropy.io import fits
 import pandas as pd
-import matplotlib.pyplot as pl
+# import matplotlib.pyplot as pl
 import healpy as hp
-import seaborn as sns; sns.set_style('ticks')
+# import seaborn as sns; sns.set_style('ticks')
 import numpy as np
 from scipy import stats
 
@@ -18,7 +22,7 @@ params = {
    'legend.fontsize': 10,
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
-   'text.usetex': False,
+   'text.usetex': True ,
    'figure.figsize': [7.281, 4.5]
    }
 
@@ -44,8 +48,9 @@ def main():
     # print(len(name))
     # exit()
 
-
-    g = sns.jointplot(x=Noter_z, y=Noter_NH, kind="hex", stat_func=None, xlim = (1, 7), ylim = (20, 23), marginal_kws = {"kde": False, "norm_hist": True}, label="Noterdaeme et al. 2012b", space=0)
+    # sns.set_palette("muted")
+    fig, ax1 = pl.subplots()
+    g = sns.jointplot(x=Noter_z, y=Noter_NH, kind="hex", stat_func=None, xlim = (1, 7), ylim = (20, 23), marginal_kws = {"kde": False, "norm_hist": True}, label="Noterdaeme et al. 2012b", space=0, color = "#4C72B0")
 
     # Plot old distribution
     NH_old = NH[~XSGRB]
@@ -66,26 +71,41 @@ def main():
     g.x = 1
     g.y = 1
     g = g.plot_joint(pl.plot, color="#55A868", label="Tanvir et al. 2017")
-    print(stats.ks_2samp(NH_new, NH_old))
-    print(stats.ks_2samp(z_new, z_old))
-    l, m, h = np.percentile(NH_new, [16, 50, 84])
-    print(m, m - l, h - m)
-    l, m, h = np.percentile(NH_old, [16, 50, 84])
-    print(m, m - l, h - m)
-    l, m, h = np.percentile(z_new, [16, 50, 84])
[…]

### 2018-01-28 py/betaOX.py

@@ -119,6 +119,16 @@ def main():
     f_HI[np.isnan(f_HI)] = 0
 
     # Plot new values
+
+
+    print(stats.pearsonr(HI, betaOX[idx_2]))
+    print(stats.spearmanr(HI, betaOX[idx_2]))
+    print(stats.pointbiserialr(HI, betaOX[idx_2]))
+    print(stats.kendalltau(HI, betaOX[idx_2]))
+    
+    exit()
+
+
     g = sns.JointGrid(x=HI, y=betaOX[idx_2], xlim = (19.5, 23), ylim = (0, 1.3), space=0)
     color = sns.color_palette()[0]
     g = g.plot_marginals(sns.distplot, hist=True, kde=False, norm_hist=True, color=color)
@@ -176,6 +186,9 @@ def main():
 
     # print(stats.ks_2samp(HI, f_HI))
     # print(len(HI))
+
+
+
     print("Above 0.5")
     l, m, h = np.percentile(HI[betaOX[idx_2] >= 0.5], [16, 50, 84])
     print(HI[betaOX[idx_2] >= 0.5])

### 2018-01-28 py/skymap.py

@@ -45,18 +45,18 @@ def main():
     """
     # Script to produce sky plots.
     """
-    ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
-    A_Vmap = ebv_map*3.1
-    A_Vmap[A_Vmap >= 0.5] = 0
-    ax = hp.mollview(A_Vmap, max=0.5, title="", cbar=False)
+    # ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
+    # A_Vmap = ebv_map*3.1
+    # A_Vmap[A_Vmap >= 0.5] = 0
+    # ax = hp.mollview(A_Vmap, max=0.5, title="", cbar=False)
     # , cbar=False
 
 
-    fig = pl.gcf()
-    ax = pl.gca()
-    image = ax.get_images()[0]
-    cmap = fig.colorbar(image, ax=ax, location="bottom", pad = 0.02, label=r"$A_V$", shrink=0.5)
-    hp.graticule()
+    # fig = pl.gcf()
+    # ax = pl.gca()
+    # image = ax.get_images()[0]
+    # cmap = fig.colorbar(image, ax=ax, location="bottom", pad = 0.02, label=r"$A_V$", shrink=0.5)
+    # hp.graticule()
 
     # Overplot swift positions
     # Read in master sample
@@ -66,9 +66,21 @@ def main():
     burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
     name_o = np.array([ii[3:] for ii in burst_table["GRB"].values])
     # Read in Swift table
-    swift_table = pd.read_table("../data/grb_table_1511519199.txt", delimiter="\t", dtype=None)
+    swift_table = pd.read_table("../data/grb_table_1517061198.txt", delimiter="\t", dtype=None)
     name_s, RA, DEC, redshift = swift_table["GRB"].values, pd.to_numeric(swift_table["BAT RA (J2000)"], errors='coerce').values, pd.to_numeric(swift_table["BAT Dec (J2000)"], errors='coerce').values, swift_table["Redshift"].values
+    count = 0
+    for ii, kk in enumerate(name_s):
+        hej = ''.join(filter(lambda x: x.isdigit(), kk))
+        print(hej, ii)
 
+        try:
+            if float(hej) < 170331:
+                count += 1
+        except:
+            pass
+    print(count)
+    # print(name_s)
+    exit()
[…]

### 2018-02-01 document/XSGRB_sample.tex
Many comments and notes on individual bursts

@@
 
-\title{XS-GRB I - The X-shooter GRB afterglow sample
-\thanks{Based on observations collected at the European Southern Observatory, Paranal, 
-	Chile, Program IDs: 098.A-0055, 097.A-0036, 096.A-0079, 095.B-0811(B), 
-	095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-	089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260.}}
+\title{XS-GRB I - The X-shooter GRB afterglow legacy sample \thanks{Based on
+		observations collected at the European Southern Observatory, Paranal, Chile,
+		Program IDs: 084.A-0260, 084.D-0265, 085.A-0009, 086.A-0073, 087.A-0055,
+		088.A-0051, 089.A-0067, 090.A-0088, 091.A-0877, 0091.C-0934, 092.A-0124,
+		092.D-0056, 093.A-0069, 094.A-0134, 095.A-0045, 095.B-0811, 096.A-0079,
+		097.A-0036, and 098.A-0055.}}
 
@@
 \abstract{In this work we present spectra of all $\gamma$-ray burst (GRB)
-	afterglows that have been promptly observed with VLT/X-shooter up to
-	\termdate. This totals spectroscopic observations of 102 individual bursts
+	afterglows that have been promptly observed with the X-shooter spectrograph up
+	to \termdate. This totals spectroscopic observations of 102 individual bursts
 	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
 	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
-	of observational selection criteria that are unbiased with regards to intrinsic
-	properties of the GRBs, the follow-up effort has been guided by the
-	\textit{Swift}-telescope to produce a homogeneous sample of spectra for 91 GRBs.
-	We here provide a public release of all the reduced spectra, including continuum
-	estimates and telluric corrections. For completeness, we also provide reductions
-	for the late-time observations of the underlying host galaxies. We provide an
-	assessment of the degree of completeness with respect to the parent GRB
-	population, in terms of the X-ray properties of the bursts in the sample and
-	find that the sample presented here is representative of the full \textit{Swift}
-	sample. For all bursts (41) for which it is possible, we provide a measurement
-	of the neutral hydrogen column density, increasing the total number of published
-	HI column density measurements by $\sim$ 50 per cent. %	We constrain the
+	of observational selection criteria that minimize biases with regards to
+	intrinsic properties of the GRBs, the follow-up effort has been focused on
+	producing a homogeneous sample of 91 afterglow spectra for GRBs discovered by
+	the {\em Swift} satellite. We here provide a public release of all the reduced
+	spectra, including continuum estimates and telluric absorption corrections. For
+	completeness, we also provide reductions for the 18 late-time observations of the
+	underlying host galaxies. We provide an assessment of the degree of completeness
+	with respect to the parent GRB population, in terms of the X-ray properties of
+	the bursts in the sample and find that the sample presented here is
+	representative of the full \textit{Swift} sample. For all bursts (41) for which
+	it is possible, we provide a measurement of the neutral hydrogen column density,
+	increasing the total number of published H{\sc I} column density measurements by
+	$\sim$ 50 per cent. %	We
+	% constrain the
 	% fraction
@@
 	%	is correlated with increased X-ray absorption.
-	This dataset provides a unique possibility to study the ISM across cosmic time,
-	from the local progenitor surroundings to the galaxy wide conditions}
+	This dataset provides a unique resource to study the ISM across cosmic time,
+	from the local progenitor surroundings to the galaxy-scale conditions.}
 
@@
 
-$\gamma$-ray bursts (GRBs) provide the opportunity to derive constraints on
-a very wide range of topics in astrophysics. Examples range from small scale
-phenomena relating to magnetars, properties of highly relativistic jets,
-hyper/supernova explosions, the interstellar medium, dust extinction curves,
-starburst galaxies, chemical and molecular abundances, escape of ionizing
-radiation, the ionization state of the intergalactic medium, intervening
-absoption systems to standard candles in cosmology \citep[e.g.,][]{Lyons2010,
-	Molinari2007, HjorthBloom2012, Wijers1998, Prochaska2009, Savaglio2006,
-	Ghirlanda2007, Vergani2009, Christensen2017}.
+$\gamma$-ray bursts (GRBs) provide constraints on a very wide range of topics in
+astrophysics. Examples range from small-scale phenomena relating to magnetars,
+properties of highly relativistic jets, hyper/supernova explosions, the
+interstellar medium, dust extinction curves, starburst galaxies, chemical and
+molecular abundances, escape of ionizing radiation, the ionization state of the
+intergalactic medium, intervening absoption systems to standard candles in
+cosmology \citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
+	Amati2008, Vergani2009, Prochaska2009, HjorthBloom2012, Rowlinson2017,
+	Christensen2017}.
 
@@
 hunting. A crucial aspect of the success of the {\it Swift} mission has been the
-extensive ground based follow-up observations of the afterglows and of the host
+extensive ground-based follow-up observations of the afterglows and of the host
 galaxies of the GRBs, involving a large community of researchers. This fruitful
@@
 
-In the beginning of the "{\it Swift} era" most of the follow-up afterglow
+In the beginning of the {\it Swift} era most of the follow-up afterglow
 spectroscopy was secured using low-resolution spectrographs \citep[typically
-R=$\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Low-resolution
-spectroscopy is powerful as it allows us to secure information even for very faint
-targets. This allows the measurements of a number of important parameters such
-as redshifts, spectral slopes, and extinction. For a handful of very bright
-afterglows high-resolution (typically R$>$20000) spectra were secured, and for
-these events much more information about conditions inside the host galaxies
-were extracted \citep[e.g.,][]{Fiore2005, Thone2007, Prochaska2007,
-	Vreeswijk2007, Castro-Tirado2010}.
-
-The X-shooter spectrograph \citep{Vernet2011} was the first of the 2nd
+$R=\lambda/\Delta\lambda$$<$1000, e.g.,][]{Fynbo2009}. Spectroscopy is powerful
+as it allows us to secure information even for very faint targets (24.5
+\citep{Kruhler2012}). This allows the measurements of a number of important
+parameters such as redshifts, spectral slopes, and extinction. For a handful of
+very bright afterglows high-resolution (typically R$>$20000) spectra have been
+secured, and for these events much more information about conditions inside the
+host galaxies were extracted \citep[e.g.,][]{Fiore2005, Thone2007,
+	Prochaska2007, Vreeswijk2007, DElia2009, Castro-Tirado2010}.
+
+The X-shooter spectrograph \citep{Vernet2011} was the first of the second
 generation instruments at the ESO Very Large Telescope (VLT). It was designed
@@
 and the characteristics of the observed bursts. We also here assess the
-completeness of the finally realized sample. Finally, we offer our conclusions
+completeness of the realized sample. Finally, we offer our conclusions
 in Sect.~\ref{conclusions}. We use the $\Lambda$CDM cosmology parameters
@@
 Being of transient nature, it is difficult to impose strong sample selection
-criteria on GRBs, without hampering the follow-up effort. Many natural follow-up
-restrictions exists already, being it weather conditions, pointing restrictions
+criteria on GRBs without hampering the follow-up effort. Many natural follow-up
+restrictions exist already, being it weather conditions, pointing restrictions
 of the telescope or unconstrained burst localizations as reported by the alerting
 telescope. To maximize the return of the follow-up campaign, we have chosen a
-few selection criteria that facilitates an unbiased selection of bursts, while
+few selection criteria that facilitate an unbiased selection of bursts, while
 at the same time allowing for a high follow-up success rate. The importance of
@@
 the aim is that the sample collected represents the underlying distribution of
-GRBs in a fair way. The selection criteria used here is based on previous,
+GRBs in a fair way. The selection criteria used here are based on previous,
 similar studies \citep{Jakobsson2006b, Fynbo2009, Hjorth2012}. We characterize
@@
 
-Our selection criteria are:
+The selection criteria that define a GRB as part of our \textit{statistical}
+	sample are:
 
 \begin{enumerate}
-	\item GRB triggered by \textit{Swift}.
+	\item GRB triggered by BAT onboard the \swift~satellite 
 	\item XRT started observing within 10 minutes after the GRB; an XRT position must be distributed within 12 hr.
@@
 selecting only bursts that have been triggered on board the \textit{Swift} space
-telescope \citep{Gehrels2004}, based on the BAT, we  start out with a homogeneously selected sample where
-burst characteristics are delivered immediately, allowing for an informed
-follow-up strategy. Despite the complexity of the triggering mechanism on board
-\textit{Swift} \citep{Band2006, Coward2013a}, attempts at inferring properties
-of the underlying GRB population based on the detection thresholds and
-triggering algorithms have been made \citep{Lien2014, Graff2016}. Restricting
-the follow-up effort to bursts detected by \textit{Swift}, we therefore ensure
-that the limitations of the parent sample are well studied.
+telescope \citep{Gehrels2004}, based on the BAT, we  start out with a sample
+where burst characteristics are delivered immediately, allowing for an informed
+follow-up strategy. However, the BAT sensitivity varies across its field of
+view, so selection is not entirely homogeneous. Despite the complexity of the
+triggering mechanism on board \textit{Swift} \citep{Band2006, Coward2013a},
+attempts at inferring properties of the underlying GRB population based on the
+detection thresholds and triggering algorithms have been made \citep{Lien2014,
+	Graff2016}. Restricting the follow-up effort to bursts detected by
+\textit{Swift}, we therefore ensure that the limitations of the parent sample
+are well studied.
 
@@
 observations to be completed. Since the GRB population is isotropically
-distributed on the sky as seen from Earth and because the GRB properties do not
+distributed on the sky as seen from Earth, and because the GRB properties do not
 depend on position on the sky \citep{Meegan1992, Briggs1996, Ukwatta2016}, this
@@
 Under rare circumstances, the use of the ESO rapid response mode (RRM) has been
-possible. In case of a promptly visible GRB, a robotic trigger was sent to the
-telescope if at the moment of the GCN notice the GRB is observable and fulfills
-the following criteria.
+possible. RRM is an automatic override of current observations at the telescope
+that allows the shortest possible delays between the GRB trigger and initiation
+of spectroscopic observations. This allows spectroscopic integration to begin on
+the rapidly fading, optical transient within minutes of the burst. In case of a
+promptly visible GRB, a robotic trigger was sent to the telescope if at the
+moment of the GCN notice the GRB is observable and fulfills the following
+criteria:
 
@@
 	\item GRB triggered by \textit{Swift}.
-	\item X-shooter mounted at the telescope
-	\item GRB promptly visible
+	\item X-shooter mounted at the telescope.
+	\item GRB promptly visible.
 	\item Error radius < 60\arcsec.
@@
 	\item The X-ray position must be available less than 1 hr after the GRB.
-	\item The flags \swift-circulated tags: KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true).
+	\item The \swift-circulated tags: KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true)\footnote{These tags are distributed as part of the GCN notices.
+Explanations for all tags are \href{https://gcn.gsfc.nasa.gov/sock_pkt_def_doc.html}{available online}.}.
 \end{enumerate}
@@
 the temporal variability of the GRB afterglow due to effect of the GRB itself on
-the sorrounding medium \citep[e.g., see][for examples]{Dessauges-Zavadsky2006,
-	Vreeswijk2007, Vreeswijk2013}
-
-There are 8 GRBs in the sample that have beem being observed with the RRM mode.
+the sorrounding medium \citep[e.g., see][]{Dessauges-Zavadsky2006,
+	Vreeswijk2007, DElia2009, Vreeswijk2013}
+
+There are nine GRBs in the sample that have been observed with the RRM mode.
 One of the RRM triggers is outside the sample we use to address the statistical
@@
 fastest response between \swift~GRB trigger and the beginning of spectroscopic
-integration is for GRB~160410A, for which the delay was only 8.4 minutes.
+integration is for GRB~160410A, for which the delay was only 8.4 minutes. In the
+cases where the RRM trigger was rejected, it was in most cases due to a wrong
+instrument being mounted at the focus.
+
+Potentially, the proximity of the telescope (Paranal, Chile) to the South
+Atlantic Anomaly (SSA) affects the rate of RRM triggers. BAT switches off when
+the satellite goes through the SSA and that means we have less triggers of GRBs
+that are immediately observable, compared to the rest of the \swift~orbit.
 
@@
 	in galactic coordinates of the full sky showing the positions on the sky of the
-	bursts presented in this work. The equator is the Galactic plane. Star symbols
-	indicate bursts with measured redshifts and the circles indicate those without.
-	The empty stars/circles are the positions of all the 1266 \textit{Swift} bursts
-	detected until \termdate. A blue filling of the symbol indicates the
-	position of bursts fulfilling the sample criteria specified in Sect.
-	\ref{samplecrit}. The red outline is added around GRBs which have been followed
-	up by X-shooter in this work. Blue symbols with red outlines represent GRBs
-	that enter our sample and have X-shooter spectroscopy. The different samples
-	are compared in Sect. \ref{results}. The background shows the dust maps
-	presented in \citet{Schlegel1998}. Please note that we removed the background
-	where the sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a
-	white color visible at the equator. The colorbar below indicates the value of
+	bursts presented in this work. The equator is the Galactic plane. The empty
+	stars/circles are the positions of all the 1266 \textit{Swift} bursts detected
+	until \termdate. Stars indicate bursts with measured redshifts and circles
+	indicate those without. Blue stars indicate the position of the 98 bursts
+	fulfilling the sample criteria specified in Sect. \ref{samplecrit} that have a
+	measured redshift. Red outlines are added to the 77 GRBs that enter our
+	statistical sample with both X-shooter spectroscopy and a measured redshift.
+	Red outlines of empty stars represent bursts which is observed with X-shooter
+	and has a measured redshift, but is not a part of our statistical sample. The
+	blue dots show the positions of the 63 GRBs of our statistical sample that lack
+	redshift measurements. A red outline is added around the 8 bursts in our
+	statistical sample which was observed with X-shooter, but did not yield a
+	redshift measurement. A single empty circle with a red outline indicate a burst
+	outside the statistical sample that was followed up with X-shooter, but without
+	a redshift measurement. The different samples are compared in Sect.
+	\ref{results}. The background shows the dust maps presented in
+	\citet{Schlegel1998}. Please note that we removed the background where the
+	sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a white
+	color visible at the equator. The grey-scale bar below indicates the value of
 	$A_V$ on the plot. The dotted lines indicate intervals of 30$^\circ$ in
@@
 the duration of this follow-up campaign. The observations have been taken during
-a period of 8 years corresponding to the ESO observing periods P84 through P98
-under the following programme IDs: 098.A-0055, 097.A-0036, 096.A-0079,
-095.A-0045, 094.A-0134, 093.A-0069, 092.A-0124, 0091.C-0934, 090.A-0088,
-089.A-0067, 088.A-0051, 087.A-0055, 086.A-0073, 085.A-0009 and 084.A-0260 (PI:
+a period of eight years corresponding to the ESO observing periods P84 through P98
+under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
+087.A-0055, 088.A-0051, 089.A-0067, 090.A-0088, 0091.C-0934, 092.A-0124,
+093.A-0069, 094.A-0134, 095.A-0045, 096.A-0079, 097.A-0036, and 098.A-0055(PI:
 Fynbo). These proposals were initiated on Guaranteed Time. We have included a
-few additional bursts, from the programmes 084.D-0265(A) (PI: Benetti),
-091.A-0877(A) (PI: Schady), 092.D-0056(A) (PI: Rau), 092.D-0633(E) (PI:
-Greiner), and 095.B-0811(B) (PI: Levan). The total collection of spectra
-represents \textit{all} GRB afterglows that have been followed up by X-shooter
-up to \termdate.
-
-The first GRB followed up is GRB~090313, observed on the 15th of March, 2009,
-during the commissioning of X-shooter on UT3. The bursts observed during the
-commissioning or science verification (GRB~090313, GRB~090530, GRB~090809,
-GRB~090926) are not a part of the sample we use to investigate the statistical
-properties of GRB afterglows. The first burst observed after science
-verification and X-shooter was moved to UT2, is GRB~091018, which is the first
+few additional bursts, from the programmes 084.D-0265 (PI: Benetti), 091.A-0877
+(PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633 (PI: Greiner), and 095.B-0811
+(PI: Levan). The total collection of spectra represents \textit{all} GRB
+afterglows that have been followed up by X-shooter up to \termdate, which marks
+the end of the XS-GRB legacy follow-up program.
+
+The first GRB followed up was GRB~090313, observed on the 15th of March, 2009,
+during the commissioning of X-shooter on the third Unit Telescope (UT3) of the
+VLT . The bursts observed during the commissioning or the science verification
+process (GRB~090313, GRB~090530, GRB~090809, GRB~090926) are not a part of the
+sample we use to investigate the statistical properties of GRB afterglows, due
+to the inhomogeneity of their selection. The first burst observed after science
+verification and X-shooter was moved to UT2, was GRB~091018, which is the first
 burst entering our statistically homogeneous sample. For all bursts that fulfill
 our sample selection criteria, described in Sect. \ref{samplecrit},
-spectroscopic follow-up has been attempted with X-shooter. Various conditions
+spectroscopic follow-up have been attempted with X-shooter. Various conditions
 can affect our ability to follow up a given burst, and a discussion of these
@@
 VIS and NIR-arm
-respectively\footnote{\url{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}}.
+respectively\footnote{See \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{this URL} for the nominal resolutions.}.
 %The slit width sets the width of the atmospheric sky lines and determines the amount of light lost due to the wavelength-dependent seeing PSF extending outside the coverage of the slit, where the width of sky-lines is always set by the slit width whereas both the delivered resolution and the slitloss changes for the better as the seeing PSF drops below the slit width.
@@
 ADC prior to disablement, resulting in a lower-than-nominal throughput. To avoid
-chromatic slit losses due to atmospheric dispersion, all subsequent observations
-have been carried out at parallactic angle. A consequence of this, is that for
-all observations following 1st of August 2012, the centroid of spectrum changes
-position across the spatial direction of the slit as a function of wavelength.
-This effect has been modeled in the extraction procedure, as described in Sect. 
-\ref{extract}.
+chromatic slit losses due to atmospheric dispersion, nearly all subsequent
+observations have been carried out at parallactic angle. A consequence of this,
+is that for all observations following 1st of August 2012, the centroid of the
+trace of the source changes position across the spatial direction of the slit as
+a function of wavelength. This effect has been modeled in the extraction
+procedure, as described in Sect. \ref{extract}.
 
@@
 celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from the
-central zone of avoidance due to high Galactic extinction cutoff (marked in
+central zone of avoidance, due to high Galactic extinction cutoff (marked in
 white), the GRB positions have an isotropic distribution, except in the upper
@@
 In this section we describe how the final data products are produced and
-subsequently post-processed. %$\sim$ 30 percent of the data released here have
-% been reduced in a manner similar to what is described in \citet{Kruhler2015}.
+subsequently post-processed. 
+%29 percent of the data released here have been
+% reduced in a manner similar to what is described in \citet{Kruhler2015}.
 % The two reduction schemes have been tested and are in good agreement.
@@
 \textit{Astro-SCRAPPY}\footnote{\url{https://github.com/astropy/astroscrappy}},
-where a wide clipping radius have been used around detected cosmics to ensure that edge residuals are
-robustly rejected.
+where a wide clipping radius was used around detected cosmic ray hits to ensure
+that edge residuals are robustly rejected.
 
@@
 spectrophotometric standards \citep{Vernet2010, Hamuy1994}, mirror flexure
-compensation, sky-subtraction and lastly the rectification and merging of the
-orders. Errors and bad pixel maps are propagated throughout the extraction. For
-the initial sky-subtraction, the background has been estimated by a running
-median in regions adjacent to the object trace clear of contaminating sources.
-Due to the broken ADC, for some objects there is curvature in the object trace
-along the dispersion axis of the slit. This means that for some bursts, the
-initial sky-estimate has been made from a limited number of pixels in the
-spatial direction. The subtraction of the sky background on the un-rectified
-image ensures that the bulk of the sky background is not redistributed by the
-rectification process.
-
-The image is rectified onto an equidistant grid with a dispersion sampling of
-0.02 nm/pixel and a 0\farcs16 per pixel spatial sampling for the UVB and VIS
-arm and 0.06 nm/pixel with a 0\farcs21 per pixel in the NIR arm.  Because
-the tabulated resolution is a lower limit, we choose a sampling of 0.02 nm/pixel
-to ensure that the lowest wavelength part of neither of the arms have a sampling
-lower than the Nyquist sampling rate of 2 pixels per resolution FWHM.
+compensation(see Sect. \ref{wavecal}), sky-subtraction and lastly the
+rectification and merging of the orders. Errors and bad pixel maps are
+propagated throughout the extraction. For the initial sky-subtraction, the
+background has been estimated by a running median in regions adjacent to the
+object trace clear of contaminating sources. Due to the broken ADC, for some
+objects there is curvature in the object trace along the dispersion axis of the
+slit(see Sect. \ref{extract}). This means that for these bursts, the initial
+sky-estimate was made from a limited number of pixels in the spatial direction.
+The subtraction of the sky background on the un-rectified image ensures that the
+bulk of the sky background is not redistributed by the rectification process.
+
+X-shooter is an echelle spectrograph, and therefore the individual echelle
+orders are curved across image space. The individual orders therefore need to be
+rectified. In order to transform the image space (pixels) into a physical
+(wavelength-slit) space, the image pixels are resampled onto a physical grid,
+while propagating the pixel uncertainties derived by the pipeline. This
+rectification  process correlates neighboring pixels and in order to minimize
+the degree of correlation, we need to choose a physical sampling that matches
+the pixel sampling. We rectified the image onto an equidistant grid with a
+dispersion sampling of 0.02 nm/pixel and a 0\farcs16 per pixel spatial sampling
+for the UVB and VIS arm and 0.06 nm/pixel with a 0\farcs21 per pixel in the NIR
+arm.  Because the tabulated resolution is a lower limit to the delivered
+resolution, we choose a sampling of 0.02 nm/pixel to ensure that the lowest
+wavelength part of neither of the arms have a sampling lower than the Nyquist
+sampling rate of 2 pixels per resolution FWHM.
 
@@
 stack and weight by an averaged measure of the inverse variance of the
-background. When weighting images where the noise in each pixel is dominated by
+background. When weighting images, where the noise in each pixel is dominated by
 Poisson noise, it is important to estimate the background variance in a large
@@
 over the variance map produced by the pipeline, where the trace has been masked
-and the width of the window is chosen to be wide enough for median variance to
-be generated on the basis of several hundred pixels. This weighting scheme
+and the width of the window is chosen to be wide enough for the median variance to
+be calculated on the basis of several hundred pixels. This weighting scheme
 automatically also optimally combines images of different exposure times or
@@
 
-For the NIR arm, because the background is very bright and there is a high
-abundance of broad sky-lines, when there are no contaminating sources in the
-slit, the sky has been put back on the images and they have been reduced in
-pairs of two, subtracting the two from each other, keeping the WCS static. This
-amounts to the regular nodding reduction, only we can reject outliers and weight
-by the averaged inverse variance map.
-
-Using this method for all observations we additionally get a spectrum of the
-sky, which we can use to recalibrate the wavelength solution.
+In the NIR arm where the background is very bright and there are a high number
+of broad sky-lines, an alternative approach to sky subtraction has been
+employed. When there are no contaminating sources in the slit, the sky has been
+put back on the images and the images are combined in pairs of two, subtracting
+the two from each other while keeping the WCS static. Due to the nod-and-shuffle
+offsets used between observations, this conserves the source flux while removing
+the sky at the expense of a decrease in signal-to-noise by a factor of
+$\sqrt{2}$. This amounts to the regular nodding reduction, only we can reject
+outliers and weight by the averaged inverse variance map.
+
+Reducing the images as single observations for all exposures, we additionally
+get a spectrum of the sky which we can use to recalibrate the wavelength
+solution.
 
@@
 
-Since X-shooter is installed at the VLT Cassegrain focus it is prone to
+Since X-shooter is installed at the VLT Cassegrain focus, it is prone to
 flexures. The flexures modify the projection of the slit on the detector with
@@
 during the X-shooter Active Flexure Compensation procedure\footnote{X-shooter
-	User Manual available at
-	https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}. The
+	User Manual available at \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}}. The
 remaining offset is corrected by cross-correlating the observed sky spectrum
@@
 estimated as the mode of all flux values, has been subtracted. To get the
-correct seeing PSF with which to convolve the synthetic sky an initial
+correct seeing point-spread-function (PSF) with which to convolve the synthetic sky, an initial
 refinement of the wavelength solution has been obtained by cross-correlating the
@@
 then be calculated by cross-correlating the observed sky with the correctly
-broadened sky spectrum, as a function of a velocity offset. Both a
+broadened sky spectrum as a function of a velocity offset. Both a
 multiplicative and an additive offset to the wavelength calibration has been
@@
 VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short
-periods of time were applied to the corresponding spectra. Using the convolved
-synthetic sky, the pixels containing the brightest sky lines have been added to
-the bad pixel map.
+periods of time were applied to the corresponding spectra\footnote{The
+	wavelength shifts have been extensively studied by
+	ESO \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
+Using the convolved synthetic sky, the pixels containing the brightest sky lines have been added to the bad pixel map.
 
@@
 
-To guide the first guess position of the trace on the slit as a function of
+To guide the estimated position of the trace on the slit as a function of
 wavelength, we have used the analytic prescription for the trace position
@@
 weak, even in the binned images, an aperture has been selected manually which
-covers emission lines, if present, and when nothing is immediately visible, the
+covers all emission lines, if present, and when nothing is immediately visible, the
 entire nodding window. The error- and bad pixel maps are in all cases propagated
@@
 of ways. We employ a prioritized list of methods here, depending on the
-availability of the different methods. Since the observation are often taken at
+availability of the chosen methods. Since the observations are often taken at
 odd times under varying conditions, this prioritized list ensures that we are
@@
 	Kausch2015}\footnote{\url{http://www.eso.org/sci/software/pipelines/skytools/molecfit}}), which can then be used to correct for the absorption.
-The accuracy of the correction depends on the S/N per pixel of the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
+The accuracy of the correction depends on the signal-to-noise per pixel of the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
 
@@
 presented here. For this, we have developed an algorithmic approach that
-attempts to automatically estimate the continuum placement along with the error on
-the continuum estimate through an iterative procedure. The method is entirely
-data-driven and does not rely on any physical assumptions. The method is applied on
-each arm separately for each spectrum.
+attempts to automatically estimate the continuum placement along with the error
+on the continuum estimate through an iterative procedure. The method is entirely
+data-driven and does not rely on any physical assumptions. The method is applied
+on each arm separately for each spectrum, to allow the widest possible
+wavelength range of the spectral shape to guide the normalization.
 
 To estimate the continuum, a number of points (typically on the order of 100)
-are inserted at random positions along the wavelength direction, the
+are inserted at random positions along the wavelength direction, and the
 flux-density of each point is determined by the median value of the spectrum in
@@
 and iteratively, the point furthest away is removed until the polynomial fit
-differ from the points by less than $\sim$ 5 per cent. This filtering is used
+differed from the points by less than $\sim$ 5 per cent. This filtering is used
 because the intrinsic afterglow continuum of GRBs can be modeled by power laws
-\citep{Piran2005}, and by removing points that differ significantly from a
-smooth continuum shape will guide the continuum estimate to a shape more
-reasonable for GRB afterglows. Additionally, points spaced closer than 1 per
-cent of the total spectral coverage are pruned. The remaining filtered and
-pruned points are then spline interpolated using
-\texttt{scipy.interpolate.splrep.splev}, which serves as a first estimate of the
-continuum placement. An attempt to identify absorption and emission regions are
-then carried out, where they are marked as such if the difference between the
-estimated continuum and the observed spectrum is larger than 3 - 5 times the
-associated error spectrum. All regions marked as affected are then masked.
+\citep{Piran2005}, and removing points that differ significantly from a smooth
+continuum shape will guide the continuum estimate to a shape more reasonable for
+GRB afterglows. The reason for the non-physical model for the normalization is
+that it has the flexibility to captures instrumental variations of the continuum
+level that are not easily contained in a more physically motivated model.
+Additionally, points spaced closer than 1 per cent of the total spectral
+coverage are pruned. The remaining filtered and pruned points are then spline
+interpolated using \texttt{scipy.interpolate.splrep.splev}, which serves as a
+first estimate of the continuum placement. To avoid the spline to diverge at the
+edges, the spline-based continuum is tapered with a low order polynomial. An
+attempt to identify absorption and emission regions are then carried out, where
+they are marked as such if the difference between the estimated continuum and
+the observed spectrum is larger than 3 - 5 times the associated error spectrum.
+All regions marked as affected are then masked.
 
@@
 stability of the algorithm across the spectrum. An example of the performance is
-shown in Fig. \ref{fig:spectrum}. In some cases, very little continuum is
-contained in a single arm and a manual continuum estimate is provided, similar
-to what is done in \citet{Lopez2016}. In these cases the continuum error is set
-to 10 per cent. The code for the continuum estimate is released along with the
-paper at \url{https://github.com/jselsing/XSGRB-sample-paper}.
+shown in Fig. \ref{fig:spectrum}. In the \lya~forest, where there is very little
+flux at the continuum level, the performance of the normalization algorithm
+depends on the continuum coverage redwards of the \lya~line. In some cases, very
+little continuum is contained in a single arm and a manual continuum estimate is
+provided, similar to what is done in \citet{Lopez2016}. In these cases the
+continuum error is set to 10 per cent. The code for the continuum estimate is
+released along with the paper at
+\url{https://github.com/jselsing/XSGRB-sample-paper}.
 
@@
 The afterglow spectra described in this paper are obtained in
-Target-of-Opportunity (override) mode. In most cases there are therefore little
+Target-of-Opportunity (override) mode. In most cases, there is little
 possibility to tweak slit widths to the seeing at the time of observations (i.e.
-to optimize spectral resolution and signal to noise), and almost all our data are
+to optimize spectral resolution and signal-to-noise), and almost all our data are
 therefore taken with a fixed set of slit widths and binning, described above. In
 a fair number of cases, the seeing full width at half maximum (FWHM) is
-considerably smaller than the silt width, and the delivered spectral resolution
+considerably smaller than the slit width, and the delivered spectral resolution
 will then be determined by the seeing rather than by the slit width, as afterglows are
 point sources (this is evidently not the case for extended sources, e.g for host
-galaxies). The delivered resolution for slit-width-dominated spectra
+galaxies). The delivered resolution for slit-width dominated spectra
 post-reduction and extraction can easily be determined from the bright sky
-emission lines. For afterglow spectra with very high signal to noise, the
+emission lines. For afterglow spectra with very high signal-to-noise, the
 delivered spectral resolution can at times be determined from the science data
@@
 absorption, other forms of line broadening, and a lack of lines at some
-redshifts, this is difficult to do at lower signal to noise ratios (the
+redshifts, this is difficult to do at lower signal-to-noise ratios (the
 majority of spectra in our sample). A broad starting value for the expected
@@
 to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR
-arm, we measured a series of lines in NIR arm spectra of a subset of 19 sources
+arm, we measured a series of lines in the NIR arm spectra of a subset of 19 sources
 used for the VIS arm above, and find that the resulting distribution is
@@
 lines to use, and we therefore use a scaled value as in the NIR arm. This simple
-method is perfect (the spatial profile of the trace is not a perfect Gaussian),
-it nevertheless gives a sufficiently accurate estimate for the analysis of these
-low signal to noise science spectra.
+analysis gives a sufficiently accurate estimate for the analysis of the low
+signal-to-noise science spectra.
 
@@
 
-All the spectra are made available at
-\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally through the
-ESO archive in the form of phase 3 material
-\footnote{\url{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}}. This
-release includes both prompt afterglow observations as well as late time
-observations of the associated hosts and represents \textit{all} afterglow
-spectra of GRBs carried out by the X-shooter spectrograph since the
+All the spectra are made available as a single
+\href{http://www.dark-cosmology.dk/~jselsing/XSGRB}{ZIP file}, through
+\url{http://grbspec.iaa.es}, and additionally through the ESO archive in the
+form of \href{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}{phase 3
+	material}. This release includes both prompt afterglow observations as well as
+late time observations of the associated hosts, and represents \textit{all}
+afterglow spectra of GRBs carried out by the X-shooter spectrograph since the
 commissioning of the instrument, \startdate, and until the end of the last
-period of the program 098.A-0055, \termdate~and thus constitutes 8 years of GRB
+period of the program 098.A-0055, \termdate~and thus constitutes eight years of GRB
 afterglow observations with X-shooter. An overview of all the spectra and their
-observational setups are given in Table. \ref{tab:sample_overview}. For each
-burst, each individual observation is provided in a separate reduction and in
-cases where observations have been repeated for an increased S/N or to follow
-the temporal evolution, a combined spectrum is also provided. No attempt has
-been made to join the spectroscopic arms, so for each observation, three spectra
-are provided in separate files.
+observational setups are given in Table \ref{tab:sample_overview}. For each
+burst, each individual observation is provided in a separate reduction, and in
+cases where observations have been repeated for an increased signal-to-noise or
+to follow the temporal evolution, a combined spectrum is also provided. No
+attempt has been made to join the spectroscopic arms, so for each observation,
+three spectra are provided in separate files.
 
@@
 	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion and drifts in the wavelength solution ($\mathrm{\AA}$).
-	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
-	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
+	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
 	\item QUAL - Bad pixel map, where a value different from zero indicates a bad pixel.
-	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$)
-	\item CONTINUUM\_ERR - Relative error on the continuum estimate 
+	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+	\item CONTINUUM\_ERR - Relative error on the continuum estimate.
 	\item TELL\_CORR - Inverse transmission spectrum. Multiply FLUX and ERR column with this column to correct for telluric absorption. 
@@
 criteria described in Sect. \ref{samplecrit}. Some bursts not fulfilling the
-sample criteria have been followed up due to interesting characteristics, e.g.
+sample criteria have been followed up due to interesting characteristics, e.g.,
 curios properties of their light curves, their brightness, etc. These bursts
@@
 statistical sample is the bright \textit{INTEGRAL} burst GRB~161023A, that
-contains at least 11 intervening absorption systems, see \ref{GRB161023A}.
+contains at least 11 intervening absorption systems (See \ref{161023}).
 
@@
 Redshift determination of bursts for which the host is too faint for a
-spectroscopic redshift measurement relies on the detection of absorption line
+spectroscopic redshift measurement relies on the detection of absorption lines
 imprinted on the GRB afterglow continuum. Because the afterglow rapidly fades
@@
 spectroscopic observation. The shortest delays are observed in RRM-mode. The
-fastest follow-up between BAT trigger and start of spectroscopic observations for any observation is for the short, $z = 1.717$, GRB~160410A for which spectroscopic
-integration was initiated only 8.4 minutes after the BAT trigger. To illustrate
-the importance of the follow-up delay for the redshift completeness, we plot the
-redshift completeness as a function of delay time in Fig. \ref{fig:timing} for
-all the bursts we have followed up, including the ones outside the statistical
-sample. As can be seen from the figure, the fraction of GRBs with a redshift
-determination decreases with follow-up delay. The redshift completeness
-for bursts that we have followed up is 92 per cent. This degree of completeness
-in the followed bursts, illustrates the efficiency of VLT/X-shooter in redshift
-determination. Not shown in the figure are an additional 12 bursts that have
-redshift determinations based on late-time host observations with delay times
-longer than $\sim$10 days.
+fastest follow-up between BAT trigger and start of spectroscopic observations
+for any observation is for the short, $z = 1.717$, GRB~160410A for which
+spectroscopic integration was initiated only 8.4 minutes after the BAT trigger.
+To illustrate the importance of the follow-up delay for the redshift
+completeness, we plot the redshift completeness as a function of delay time in
+Fig. \ref{fig:timing} for all the bursts we have followed up, including the ones
+outside the statistical sample. As can be seen from the figure, the fraction of
+GRBs with a redshift determination decreases with follow-up delay. The redshift
+completeness for bursts that we have followed up is 92 per cent. This degree of
+completeness in the followed bursts, illustrates the efficiency of VLT/X-shooter
+in redshift determination. Not shown in the figure are an additional 12 bursts
+that have redshift determinations based on late-time host observations with
+delay times longer than $\sim$10 days.
 
@@
 	\centerline{\includegraphics[width=9cm]{figures/timing.pdf}} \caption{Afterglow
-	magnitude at the start of observation as a function of follow-up delay for all
-	the afterglows that have been followed up. The points have been colored based
-	on the redshift of the corresponding burst. Red crosses indicate GRBs without a
-	measured redshift and arrows indicate bursts for which the afterglow was not
-	detected in the acquisition image. In red is shown the redshift completeness as
-	a function of follow-up delay.} \label{fig:timing}
+	magnitude at the start of observation and redshift completeness as a function
+	of follow-up delay for all the afterglows that have been followed up. The
+	points have been colored based on the redshift of the corresponding burst. Red
+	crosses indicate GRBs without a measured redshift and arrows indicate bursts
+	for which the afterglow was not detected in the acquisition image. In red is
+	shown the redshift completeness as a function of follow-up delay.}
+\label{fig:timing}
 \end{figure}
@@
 
-A total of 158 bursts fulfill the sample criteria specified in Sect.
-\ref{samplecrit},  since the commissioning of VLT/X-shooter. This sample
-constitutes the \textit{statistical sample} from which we will derive
-statistical properties of the GRB population. The redshift completeness of the
-full statistical sample is 61 per cent. We return to the question of redshift
-completeness in Sect. \ref{redshift}. From this sample, 91 GRBs have been
-spectroscopically followed up with X-shooter. In order to assess whether the
-subset of bursts followed up are representative of the underlying GRB parent
-population, we compare intrinsic properties of GRBs in our sample to GRBs in the
-full sample followed up by \textit{Swift}. We show the comparison between the
-BAT (15 -- 150 keV) fluence, the XRT flux (0.3 -- 10 keV) at 11 hours, and the
-intrinsic X-ray derived equivalent hydrogen column density, $N_{\mathrm{HI,X}}$,
-in Fig. \ref{fig:swift_complete}. For the latter, we can only use values of
-$N_{\mathrm{HI,X}}$ derived for bursts with a measured redshift, excluding $\sim
-75$ per cent of the full \swift~sample. We return to the last point in Sect.
-\ref{darkness}.
+Of all the BAT-triggered bursts, a total of 158 bursts fulfill the sample
+criteria specified in Sect. \ref{samplecrit}, since the commissioning of
+VLT/X-shooter. This sample constitutes the \textit{statistical sample} from
+which we will derive statistical properties of the GRB population. The redshift
+completeness of the full statistical sample is 61 per cent. We return to the
+question of redshift completeness in Sect. \ref{redshift}. From this sample, 91
+GRBs have been spectroscopically followed up with X-shooter. In order to assess
+whether the subset of bursts followed up are representative of the underlying
+GRB parent population, we compare intrinsic properties of GRBs in our sample to
+GRBs in the full sample followed up by \textit{Swift}. We show the comparison
+between the BAT (15 -- 150 keV) fluence, the XRT flux (0.3 -- 10 keV) at 11
+hours, and the intrinsic X-ray derived equivalent hydrogen column density at the
+redshift of the GRB, in excess of the Galactic X-ray absorption column,
+$N_{\mathrm{HI,X}}$, in Fig. \ref{fig:swift_complete}. For the latter, we can
+only use values of $N_{\mathrm{HI,X}}$ derived for bursts with a measured
+redshift, excluding $\sim 75$ per cent of the full \swift~sample. We return to
+the last point in Sect. \ref{darkness}.
 
@@
 	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
-\caption{Comparison between burst properties of all bursts observed with
+\caption{Comparison between the burst properties of all bursts observed with
 	\textit{Swift} (blue), the statistical sample that fullfill the criteria
@@
 
-Using the observational characteristics of the 1322 bursts observed by
-24.10.2017 by \textit{Swift}, and $N_{\mathrm{HI,X}}$ \citep{Evans2009}, we can
+Using the observational characteristics of the 1266 bursts observed until
+\termdate~by \textit{Swift}, and the derived $N_{\mathrm{HI,X}}$ \citep{Evans2009}, we can
 quantify the degree to which our sample is biased relative to the overall
@@
 	distribution. The arrows represent the comparison direction with each of the
-	p-values they are listed next to. Only in the HI column density distribution is
+	p-values they are listed next to. Only in the H{\sc I} column density distribution is
 	there mild evidence against the null hypothesis, but the discrepancy is mainly
@@
 
-Out of the 158 bursts in the statistical sample, 33( $\sim$ 20 per cent) have
-not been observed due to reasons unrelated to the GRB or afterglow properties.
-The reasons include unavailability of the telescope due to technical maintenance
-(e.g., mirror re-coating), a visitor rejecting the TOO trigger, or bad weather.
-Because this cut is unrelated to the GRB properties, it will not change the
-statistical properties of the full sample. Removing all these bursts for the
-statistical sample, drastically improves the redshift completeness from 61 per
-cent to 84 per cent. The remaining burst not followed up already had redshift
-from other instruments or were very faint and without a host association thus
-observations unlikely to yield a redshift measurement. In the remainder of the
-text, we consider the 125 bursts our statistical sample.
+Out of the 158 bursts in the statistical sample, 33 ( $\sim$ 20 per cent) were
+not observed due to reasons unrelated to the GRB or afterglow properties. The
+reasons include unavailability of the telescope due to technical maintenance
+(e.g., mirror re-coating), a visiting observer rejecting the ToO trigger, or bad
+weather. Because this cut is unrelated to the GRB properties, it will not change
+the statistical properties of the full sample. Removing these bursts for the
+statistical sample, dramatically improves the redshift completeness from 61 per cent to 84
+per cent. The remaining burst not followed up already had redshift from other
+instruments or were very faint and without a host association thus observations
+unlikely to yield a redshift measurement. In the remainder of the text, we
+consider the 125 bursts our statistical sample.
 
@@
 sample and in the main panel, we show the redshifts of the individual bursts as
-a function of GRB luminosity in the observed 15 -- 150 keV band. To calculate the
-luminosity, $L_{\mathrm{BAT}}$, we follow a similar procedure as
-\citet{Lien2016} and define $L_{\mathrm{BAT}} = F_{\gamma}\,4
-\pi\,d_L^2\,(1+z)^{-1}$, where $F_{\gamma}$ is the observed BAT fluence in the
-15-150 keV band and $d_L$ is the luminosity distance to the burst at the given
-redshift. Note that this measure of luminosity does not include any
-k-correction. As an indication of the effect of the \textit{Swift} sensitivity
-limit on the redshift distribution, also in the figure, we have shown the so
-called $\sim 1$ s flux BAT sensitivity limit ($\sim 3 \times
-10^{-8}~\mathrm{erg}~\mathrm{s}^{-1}~\mathrm{cm}^{-2}$;
+a function of GRB energy in the observed 15 -- 150 keV band. To calculate the
+energy, $E_{\mathrm{BAT}}$, we follow a similar procedure as \citet{Lien2016}
+and define $E_{\mathrm{BAT}} = F_{\gamma}\,4 \pi\,d_L^2\,(1+z)^{-1}$, where
+$F_{\gamma}$ is the observed BAT fluence in the 15-150 keV band and $d_L$ is the
+luminosity distance to the burst at the given redshift. Note that this measure
+of luminosity does not include any $k$-correction. As an indication of the
+effect of the \textit{Swift} sensitivity limit on the redshift distribution,
+also in the figure, we have shown the so called $\sim 1$ s flux BAT sensitivity
+limit ($\sim 3 \times 10^{-8}~\mathrm{erg}~\mathrm{s}^{-1}~\mathrm{cm}^{-2}$;
 \citealt{Baumgartner2013, Lien2016}). Due to the complex triggering mechanism of
@@
 the dilution of light with distance, the \swift~GRB luminosity detection limit
-is almost an order of magnitude brighter at $z=2$ than at $z=1$. At $z\geq4$
-we are only able to observe GRBs that are $\sim$ hundred times brighter than the
-faintest bursts at $z=1$ and below.
+is almost an order of magnitude brighter at $z=2$ than at $z=1$. At $z\geq4$ we
+are only able to observe GRBs that are $\sim$ hundred times brighter than the
+faintest bursts at $z=1$ and below. The effect of GRB redshift on the \swift~triggering criteria have previously been studied in detail \citep{Littlejohns2013a}.
 
@@
 making inference of the \textit{true} redshift distribution of GRBs based on
-this sample is impossible. The biases introduced in the measured redshift
-distribution by our increased ability to secure redshift of optically brighter
-bursts has been investigated by \citet{Turpin2016} which find that we
-systematically are able to measure redshifts for longer duration GRBs.
-Additionally only the brightest GRBs are seen above redshift $z \gtrsim 1$ as
+this sample is impossible. %The biases introduced in the measured redshift
+%distribution by our increased ability to secure redshifts for optically
+% brighter
+%bursts has been investigated by \citet{Turpin2016} who find that we
+%systematically are able to measure redshifts for longer duration GRBs.
+%Additionally o
+For instance, only the brightest GRBs are seen above redshift $z \gtrsim 1$ as
 shown in Fig. \ref{fig:z}. As described in detail in \citet{Hjorth2012} and
@@
 $10^{10}~\mathrm{M}_\odot$ have the redshift measured based on the afterglow
-continuum.
+continuum. This is likely related to the presence of higher contents of dust in
+more massive galaxies, leading to a larger fraction of extincted afterglows.
 
 \begin{figure*}
-	\centering \includegraphics[width=16cm]{figures/XSGRB_redshift_swiftdots.pdf}
+	\centering \includegraphics[width=16cm]{figures/XSGRB_redshift.pdf}
 \caption{Redshift distribution as a function of intrinsic BAT $\gamma$-ray
-	luminosity, $L_{\mathrm{BAT}}$. Bursts that are a part of the statistical
-	sample are marked by blue stars whereas black dots show all GRBs observed with
-	X-shooter. All \swift~GRBs with measured redshifts are shown in grey. For
+	energy, $E_{\mathrm{BAT}}$. Bursts that are a part of the statistical sample
+	are marked by blue stars whereas black dots show all GRBs observed with
+	X-shooter. All \swift~GRBs with measured redshifts are shown in gray. For
 	comparison, we overplot with red stars the GRB \citep{Tanvir2009b,
@@
 	the highest spectroscopically confirmed redshifts, the latter two shown at
-	arbitrary $L_{\mathrm{BAT}}$. The blue solid line represents the so-called 1 s
+	arbitrary $E_{\mathrm{BAT}}$. The blue solid line represents the so-called 1 s
 	BAT sensitivity limit described in the text. The estimated epoch of
@@
 relative to the X-ray brightness has been parametrized in term of their optical
-darkness, $\beta_{OX}$ \citep{Jakobsson2004, Rol2005, VanderHorst2009}. The
-X-ray properties of such bursts have previously been investigated
+darkness, using the measurement of, or limit on, the optical to X-ray spectral
+index $\beta_{OX}$ \citep{Jakobsson2004, Rol2005, VanderHorst2009}. The X-ray
+properties of such bursts have previously been investigated
 \citep{DePasquale2003, Fynbo2009, Melandri2012} and there are some indications
 that dark bursts have somewhat higher X-ray luminosity and $N_{\mathrm{HI, X}}$
-compared to the optically bright bursts. This indicates along with
+compared to the optically bright bursts \citep{Campana2012}. This indicates along with
 investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
 	Hjorth2012, Perley2016b}, that the extinction of the optical afterglows is
-primarily driven by the presence of dust in the host galaxies. \citet{Hjorth2012}
-find that systems with no optical afterglow have higher $N_{\mathrm{HI, X}}$,
-irrespective of the nature of the host -- which however also turn out to be
-redder.
+primarily driven by the presence of dust in the host galaxies and not solely by
+unfortunate placement of the synchrotron spectral break frequencies.
+\citet{Hjorth2012} find that systems with no optical afterglow have higher
+$N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which however
+also turn out to be redder.
 
@@
 In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{HI, X}}$
-distribution with the one presented in \citet{Fynbo2009}. The values from
-\citet{Fynbo2009} have been treated as detections, meaning that we artificially
-bias the distribution towards higher $\beta_{OX}$-values. The two distributions
-exhibit a large degree of overlap. We cannot confirm the result by
-\citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher \nhx.
-Specifically, for bursts with measured redshift either from the afterglow or the
-host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$ we
-find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
+distribution with the one presented in \citet{Fynbo2009}. We take the
+$N_{\mathrm{HI,X}}$ values from the
+\href{http://www.swift.ac.uk/xrt_spectra}{XRT spectral fits} \citep{Evans2009}.
+The values from \citet{Fynbo2009} have been treated as detections, meaning that
+we artificially bias the distribution towards higher $\beta_{OX}$-values. The
+two distributions exhibit a large degree of overlap. We cannot confirm the
+result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher
+\nhx. Specifically, for bursts with measured redshift either from the afterglow
+or the host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$
+we find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
 \nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
@@
 Using the table maintained by
-Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html}, we can see how the
+J. Greiner\footnote{http://www.mpe.mpg.de/~jcg/grbgen.html}, we can see how the
 presence of an optical afterglow affects the follow-up statistics. 50.5 per cent
-of all \textit{Swift}-detected bursts do not have a detected optical afterglow,
-but this number also includes bursts where no optical observations were
-available, so is a lower limit. For the bursts that enter our statistical
-sample, the percentage of bursts without a detected optical afterglow is 28 per
-cent, close to the upper limit on the fraction of dark bursts found in a
-complete sample with a very high degree of redshift completeness
-\citep{Melandri2012}. Of the bursts for which follow-up has actually been
-attempted, this number is 23 per cent, suggesting a slight bias against
-bursts without a detected optical afterglow in the spectroscopic sample.
+of all \textit{Swift}-triggered bursts in this list do not have a detected
+optical afterglow. This number also includes bursts where no optical
+observations were available, so the real number is likely to be lower. For the
+bursts that enter our statistical sample, the percentage of bursts without a
+detected optical afterglow is 28 per cent, close to the upper limit on the
+fraction of dark bursts found in a complete sample with a very high degree of
+redshift completeness \citep{Melandri2012}. Of the bursts for which follow-up
+has actually been attempted, this number is 23 per cent, suggesting a slight
+bias against bursts without a detected optical afterglow in the spectroscopic
+sample.
 
@@
 
-Regardless that the fraction of dark bursts is lower in the observed sample,
-compared to the statistical sample, the X-ray properties not differ
+Regardless the fraction of dark bursts being lower in the observed sample,
+compared to the statistical sample, the X-ray properties do not differ
 significantly, as shown in Sect. \ref{completeness}. This is despite
@@
 
-The locations of long GRBs are associated with intensely star forming regions
+The locations of long GRBs are associated with intensely star-forming regions
 \citep{Hogg1999, Bloom2002, Fruchter2006}. Because a significant fraction of the
-hydrogen in these regions has not been ionized yet, the optical depth at the
-wavelength of \lya~is very high, saturating not only the line center, but also
-the damping wings. This causes a strong absorption system from the
-\lya-transition to appear in the afterglow continuum. For bursts with $z \gtrsim
-1.7$, the position of \lya~moves into the spectroscopic coverage of X-shooter,
-meaning that we are able to detect this absorption trough due to \lya.
+hydrogen along the line of sight in these regions has not been ionized, the
+optical depth at the wavelength of \lya~is very high, saturating not only the
+line center, but also the damping wings. This causes a strong absorption system
+from the \lya-transition to appear in the afterglow continuum. For bursts with
+$z \gtrsim 1.7$, the position of \lya~moves into the spectroscopic coverage of
+X-shooter, meaning that we are able to detect this absorption trough due to
+\lya. The exceptionally good UV response of X-shooter, allows us to robustly
+measure H{\sc i}-columns - also at these very low redshifts.
 
@@
 \swift. This work was initiated by a consortium using Guaranteed Time on
-X-shooter, but over the years the project continued in open time and the team
-was opened to include any researchers interested in contributing to this effort.
-
-The sample presented here include spectroscopic observations of 91 systems
-fulfilling our sample criteria including 18 spectra that are late time
+X-shooter, but over the years the project continued in open time.
+
+The sample presented here includes spectroscopic observations of 91 systems
+fulfilling our sample criteria, including 18 spectra that are late-time
 observations of the underlying host galaxies. All spectra have been made
@@
 sense that only after observing more than 12000 damped Lyman-$\alpha$ absorbers
-(DLAs) towards about 10$^5$ quasars, a handful systems with
-\nh~$ > 22$ been identified \citep[e.g., five
-in][]{Noterdaeme2012b}. Long GRB afterglow spectra, by contrast, reveal such
-systems in the majority of cases \citep[][and this work]{Jakobsson2006b,
-	Fynbo2009, Cucchiara2015}. With afterglow spectroscopy (throughout the
-electromagnetic spectrum from X-rays to the sub-mm) we are able to characterize
-the properties of star-forming galaxies over cosmic history in terms of
-redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
+(DLAs) towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ been
+identified \citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra,
+by contrast, reveal such systems in the majority of cases \citep[][and this
+work]{Jakobsson2006b, Fynbo2009, Cucchiara2015}. With afterglow spectroscopy
+(throughout the electromagnetic spectrum from X-rays to the sub-mm) we are able
+to characterize the properties of star-forming galaxies over cosmic history in
+terms of redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
 densities, extinction curves, etc.  A number of independent papers have been
@@
 galaxies \citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
-\citep{Christensen2017}, and escape of ionizing radiation (Tanvir et al.,
-submitted). A number of additional companion papers are also planned,
-investigating the detailed properties of the sample presented here, including
-equivalent width distributions (de Ugarte Postigo et al., in preparation),
-metallicities and kinematics (Th{\"o}ne et al., in preparation), high ionization
-lines (Heintz et al., in preparation), molecular lines (Bolmer et al., in
-preparation), fine-structure lines (Vreeswijk et al., in preparation), composite
-GRB afterglow spectrum (Selsing et al., in preparation).
+\citep{Christensen2017}, \citet{Arabsalmani2018} on the metallicity-scaling
+relations, and escape of ionizing radiation (Tanvir et al., submitted). A number
+of additional companion papers are also planned, investigating the detailed
+properties of the sample presented here, including equivalent width
+distributions (de Ugarte Postigo et al., in preparation), metallicities and
+kinematics (Th{\"o}ne et al., in preparation), high ionization lines (Heintz et
+al., in preparation), molecular lines (Bolmer et al., in preparation),
+fine-structure lines (Vreeswijk et al., in preparation), composite GRB afterglow
+spectrum (Selsing et al., in preparation).
 
@@
 significantly more frequent than in the sightlines sampled by our spectra.
-However, with current instrumentation these sightlines are out of reach except
+However, with current instrumentation, these sightlines are out of reach except
 under very fortunate circumstances as in the case of GRB~080607 when the
@@
 \begin{acknowledgements}
-	*** We should add an acknowledgment to Neil Gehrels here ***. JPUF, BMJ and DX
-acknowledge support from the ERC-StG grant EGGS-278202. The Dark Cosmology
-Centre were funded by the Danish National Research Foundation. This work was
-supported by a VILLUM FONDEN Investigator grant to JH (project number 16599).
-TK acknowledges support by the European Commission under the Marie Curie
-Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
-European Commission under the Marie Curie Career Integration Grant programme
-(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
-{\it Swift} Science Data Centre at the University of Leicester. JJ acknowledges
+	*** We should add an acknowledgment to Neil Gehrels and Javier Gorosabel here
+***. 
+%
+JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.
+The Dark Cosmology Centre were funded by the Danish National Research
+Foundation. 
+%
+This work was supported by a VILLUM FONDEN Investigator grant to JH
+(project number 16599). 
+%
+TK acknowledges support by the European Commission
+under the Marie Curie Intra-European Fellowship Programme in FP7.  
+%
+JJ acknowledges
 support from NOVA and NWO-FAPESP grant for advanced instrumentation in
-astronomy. KEH and PJ acknowledge support by a Project Grant (162948--051) from
-The Icelandic Research Fund. Finally, we acknowledge expert support from the
-ESO staff at the Paranal and La Silla observatories in obtaining these target
-of opportunity data.
+astronomy. 
+%
+KEH and PJ acknowledge support by a Project Grant (162948--051) from
+The Icelandic Research Fund. 
+%
+AG acknowledges the financial support from the
+Slovenian Research Agency (research core funding No. P1-0031 and project grant
+No. J1-8136). 
+%
+CT acknowledges support from a Spanish National Research Grant of Excellence
+under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
+fellowship under grant number RyC-2012-09984.
+%
+AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
+Grant for Researchers and Cultural Creators, and the Spanish Ministry of Economy
+and Competitiveness through project AYA2014-58381-P.
+%
+ZC acknowledges support from the Spanish research project AYA 2014-58381-P and
+support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
+%
+AK acknowledges support from the Spanish research project AYA 2014-58381-P and
+support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2015- 26153.
+%
+RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
+Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
+%
+This work made use of data supplied by the UK {\it Swift} Science Data Centre at
+the University of Leicester. Finally, it is with pleasure that we acknowledge
+expert support from the ESO staff at the Paranal and La Silla observatories in
+obtaining these target of opportunity data.
 
@@
 
-\subsection{GRB~090313 (z = 3.373)}
+\subsection{GRB~090313 (z = 3.373)} \label{090313}
+
 The first GRB ever observed with X-shooter, during the commissioning of the
@@
 
-\subsection{GRB~090530 (z = 1.266)}
+\subsection{GRB~090530 (z = 1.266)}\label{090530}
+
 Observed during paranalization of the instrument, this data forms the basis of
@@
 
-\subsection{GRB~090809 (z = 2.737)} Observed during the first science
-verification period and was reported in GCN
+\subsection{GRB~090809 (z = 2.737)} \label{090809}
+
+Observed during the first science verification period and was reported in GCN
 \#9761\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9761.gcn3}} and is
@@
 
-\subsection{GRB~090926 (z = 2.106)} Obtained during the second science
+\subsection{GRB~090926 (z = 2.106)} \label{090926}
+
+Obtained during the second science
 verification period, this dataset forms the basis of GCN
@@
 
-\subsection{GRB~091018 (z = 0.971)}
+\subsection{GRB~091018 (z = 0.971)}\label{091018}
+
 The first burst observed during normal operation after science verification was
@@
 all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
-\mnii, \mgii, \mgi, and \caii al located at $z = 0.971$, setting is as the
+\mnii, \mgii, \mgi, and \caii~- al located at $z = 0.971$, setting is as the
 redshift of the host.
 
-\subsection{GRB~091127 (z = 0.490)}
+\subsection{GRB~091127 (z = 0.490)} \label{091127}
+
 Obtained 4 days after the burst trigger, this data forms the basis for GCN \#
 10233\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10233.gcn3}} and is published
-in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the S/N of
+in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the signal-to-noise of
 the afterglow continuum is low especially in the UVB arm, why no clear
@@
 
-\subsection{GRB~100205A  (z = na)}
+\subsection{GRB~100205A  (z = na)} \label{100205}
+
 Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected
@@
 
-\subsection{GRB~100219A (z = 4.667)}
+\subsection{GRB~100219A (z = 4.667)} \label{100219}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~100316B (z = 1.180)}
+\subsection{GRB~100316B (z = 1.180)} \label{100316}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~100316D (z = 0.059)}
+\subsection{GRB~100316D (z = 0.059)} \label{100316}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~100418A (z=0.624)}
+\subsection{GRB~100418A (z=0.624)} \label{100418}
+
 The data presented here also formed the basis of GCN \#
@@
 detection of emission features belonging to nebular lines, including \hi, \oii,
-\oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due
-to the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii,
-all at a consistent redshift. Temporal evolution of the fine structure lines
+\oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due to
+the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all
+at a consistent redshift. Temporal evolution of the fine structure lines
 belonging to \feii$^*$ is found between the epochs.
 
-\subsection{GRB~100424A (z=2.465)}
+\subsection{GRB~100424A (z=2.465)} \label{100424}
+
 The data presented here also formed the basis of GCN \#
-14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}. The spectrum
-has not otherwise been published previously. Observations carried out, long
-after the burst has faded.  Emission lines from the host are detected at
-$z=2.465$.
-
-\subsection{GRB~100425A (z=1.1755)}
+14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}. The spectrum has
+not otherwise been published previously. Observations carried out, long after
+the burst has faded.  Emission lines from the host are detected at $z=2.465$.
+
+\subsection{GRB~100425A (z=1.1755)} \label{100425}
+
 The data presented here also formed the basis of GCN \#
 10684\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10684.gcn3}} and is used in
-Skuladottir (2010), but not published elsewhere. Observations started 4
-hours after the \textit{Swift} trigger, totaling 2.4 ks. Absorption features
-from \mgii~and \feii~in the afterglow continuum are detected at $z=1.1755$.
-
-\subsection{GRB~100615A (z=1.398)}
+Skuladottir (2010), but not published elsewhere. Observations started 4 hours
+after the \textit{Swift} trigger, totaling 2.4 ks. Absorption features from
+\mgii~and \feii~in the afterglow continuum are detected at $z=1.1755$.
+
+\subsection{GRB~100615A (z=1.398)} \label{100615}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~100621A (z=0.542)}
+\subsection{GRB~100621A (z=0.542)} \label{100621}
+
 The data presented here also formed the basis of GCN \#
@@
 published elsewhere. Beginning 7.1 hours after the GRB, 2.4 ks observations
-reveal emission lines from \oii, \hb~and \oiii~at a common redshift of
-$z=0.542$ and a very weak afterglow continuum.
-
-\subsection{GRB~100625A (z=0.452)}
+reveal emission lines from \oii, \hb~and \oiii~at a common redshift of $z=0.542$
+and a very weak afterglow continuum.
+
+\subsection{GRB~100625A (z=0.452)} \label{100625}
+
 The data presented here is of the candidate host galaxy, taken long after the
@@
 
-\subsection{GRB~100724A* (z = 1.288)}
+\subsection{GRB~100724A (z = 1.288} \label{100724}
 
@@
 
-\subsection{GRB~100728B (z=2.106)}
+\subsection{GRB~100728B (z=2.106)} \label{100728}
 
@@
 lower throughout. Additionally, the position of the trace on the slit moves due
-to atmospheric differential refraction.
-
-\subsection{GRB~100814A (z=1.439)}
+to atmospheric differential refraction. The presence of the DLA is confirmed in
+the 2D image and despite the observational challenges that affects this burst,
+we measure \nh~$=21.2 \pm 0.15$.
+
+\subsection{GRB~100814A (z=1.439)} \label{100814}
 
@@
 
-\subsection{GRB~100816A (z=0.805)}
+\subsection{GRB~100816A (z=0.805)} \label{100816}
 
@@
 
-\subsection{GRB~100901A (z=1.408)}
+\subsection{GRB~100901A (z=1.408)} \label{100901}
 
@@
 
-\subsection{GRB~101219A (z=0.718)}
+\subsection{GRB~101219A (z=0.718)} \label{101219}
 
@@
 
-\subsection{GRB~101219B (z=0.552)}
+\subsection{GRB~101219B (z=0.552)} \label{101219}
 
@@
 
-\subsection{GRB~110128A (z=2.339)}
+\subsection{GRB~110128A (z=2.339)} \label{110128}
 
@@
 \swift~trigger and lasted for a total of 7.2ks The afterglow continuum is
-detected across the entire spectral coverage at moderate S/N. Absorption lines
-in the continuum is detected from \lya, \oi, \cii, \SIiv, \civ, \SIii~and \feii,
-all at a common redshift of $z=2.339$. From the broad \lya~trough, a hydrogen
-column density $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 22.6 \pm 0.2$ is
-derived. An intervening system at $z=2.20$ is tentatively identified from an
-absorption feature, likely due to \civ.
-
-\subsection{GRB~110407A (z=na)}
+detected across the entire spectral coverage at moderate signal-to-noise.
+Absorption lines in the continuum is detected from \lya, \oi, \cii, \SIiv, \civ,
+\SIii~and \feii, all at a common redshift of $z=2.339$. From the broad
+\lya~trough, a hydrogen column density $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2})
+= 22.6 \pm 0.2$ is derived. An intervening system at $z=2.20$ is tentatively
+identified from an absorption feature, likely due to \civ.
+
+\subsection{GRB~110407A (z=na)} \label{110407}
 
@@
 
-\subsection{GRB~110709B (z=2.109 (NEW!!!!))}
+\subsection{GRB~110709B (z=2.109 (NEW!!!!))} \label{110709}
 
@@
 
-\subsection{GRB~110715A (z=0.823)}
+\subsection{GRB~110715A (z=0.823)} \label{110715}
 
@@
 
-\subsection{GRB~110721A (z=0.382)}
+\subsection{GRB~110721A (z=0.382)} \label{110721}
 
@@
 
-\subsection{GRB~110808A (z=1.348)}
+\subsection{GRB~110808A (z=1.348)} \label{110808}
 
@@
 
-\subsection{GRB~110818A (z=3.36)}
+\subsection{GRB~110818A (z=3.36)} \label{110818}
 
 Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8ks
-reveals a moderate S/N GRB afterglow continuum, down to $\sim$ 5000 \AA. The
+reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 5000 \AA. The
 simultaneous detection of absorption features identified as \lya, \SIii, \civ,
@@
 
-\subsection{GRB~111005A (z=0.013)}
+\subsection{GRB~111005A (z=0.013)} \label{111005}
 
@@
 
-\subsection{GRB~111008A  (z=4.989)}
+\subsection{GRB~111008A  (z=4.989)} \label{111008}
 
@@
 
-\subsection{GRB~111107A (z=2.893)}
+\subsection{GRB~111107A (z=2.893)} \label{111107}
 
@@
 
-\subsection{GRB~111117A (z=2.211)}
+\subsection{GRB~111117A (z=2.211)} \label{111117}
 
@@
 
-\subsection{GRB~111123A  (z=3.151)}
+\subsection{GRB~111123A  (z=3.151)} \label{111123}
 
@@
 
-\subsection{GRB~111129A (z=1.080)}
+\subsection{GRB~111129A (z=1.080)} \label{111129}
 
@@
 
-\subsection{GRB~111209A (z=0.677)}
+\subsection{GRB~111209A (z=0.677)} \label{111209}
 
@@
 
-\subsection{GRB~111211A (z=0.478)}
+\subsection{GRB~111211A (z=0.478)} \label{111211}
 
@@
 
-\subsection{GRB~111228A (z=0.716)}
+\subsection{GRB~111228A (z=0.716)} \label{111228}
 
@@
 
-\subsection{GRB~120118B (z = 2.943)}
+\subsection{GRB~120118B (z = 2.943)} \label{120118}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~120119A (z = 1.728)}
+\subsection{GRB~120119A (z = 1.728)} \label{120119}
 
@@
 
-\subsection{GRB~120211A (z = 2.346)}
+\subsection{GRB~120211A (z = 2.346)} \label{120211}
+
 The data presented here has been published in \citet{Kruhler2015}. Two
@@
 
-\subsection{GRB~120224A (z = 1.10 NEW!!!)}
+\subsection{GRB~120224A (z = 1.10 NEW!!!)} \label{120224}
 
@@
 
-\subsection{GRB~120311 (z = 0.350 NEW!!!)}
+\subsection{GRB~120311 (z = 0.350 NEW!!!)} \label{120311}
+
 The data presented here has formed the basis of GCN \#
@@
 
-\subsection{GRB~120327A (z = 2.813)}
+\subsection{GRB~120327A (z = 2.813)} \label{120327}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~120404A (z = 2.876)}
+\subsection{GRB~120404A (z = 2.876)} \label{120404}
+
 The data presented here has formed the basis of GCN \#
@@
 
-\subsection{GRB~120422A (z = 0.283)}
+\subsection{GRB~120422A (z = 0.283)} \label{120422}
+
 The data presented here also formed the basis of GCN \# 
@@
 
-\subsection{GRB~120712A (z = 4.175)}
-The data presented here also formed the basis of GCN \#13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
+\subsection{GRB~120712A (z = 4.175)} \label{120712}
+
+The data presented here also formed the basis of GCN
+\#13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
 published elsewhere. 4.8 ks integration time, starting 10.5 hours after the BAT
@@
 
-\subsection{GRB~120714B (z = 0.398)}
-The data presented here also formed the basis of GCN \#13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
+\subsection{GRB~120714B (z = 0.398)} \label{120714}
+
+The data presented here also formed the basis of GCN
+\#13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
 published elsewhere. Observations of this burst started 7.8 hours after the GRB
@@
 coverage of X-shooter, with both emission lines from  \oii, \hb, \oiii~and \ha,
-as well as absorption from \mgii~detected at $z = 0.398$, securely setting it
-as the redshift of the GRB.
-
-\subsection{GRB~120716A (z = 2.486)}
-The data presented here also formed the basis of GCN \#13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
+as well as absorption from \mgii~detected at $z = 0.398$, securely setting it as
+the redshift of the GRB.
+
+\subsection{GRB~120716A (z = 2.486)} \label{120716}
+
+The data presented here also formed the basis of GCN
+\#13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
 published elsewhere. Despite observations starting 62 hours after the
@@
 
-\subsection{GRB~120722A (z = 0.959)}
-The data presented here also formed the basis of GCN \#13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
+\subsection{GRB~120722A (z = 0.959)} \label{120722}
+
+The data presented here also formed the basis of GCN
+\#13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
 published elsewhere. On 4.8 ks integration time, starting 10 hours after the
@@
 \mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
-\hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the
-redshift of the GRB.
-
-\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)}
+\hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the redshift
+of the GRB.
+
+\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)} \label{120805}
+
 A separate reduction of this burst has been published in \citet{Kruhler2015},
@@
 
-\subsection{GRB~120815A (z = 2.358)}
+\subsection{GRB~120815A (z = 2.358)} \label{120815}
 
@@
 
-\subsection{GRB~120909A (z = 3.929)}
+\subsection{GRB~120909A (z = 3.929)} \label{120909}
 
@@
 
-\subsection{GRB~120923A (z = 7.84)}
-
-\subsection{GRB~121024A (z = 2.300)}
+\subsection{GRB~120923A (z = 7.84)} \label{120923}
+
+\subsection{GRB~121024A (z = 2.300)} \label{121024}
 
@@
 
-\subsection{GRB~121027A (z = 1.773)}
+\subsection{GRB~121027A (z = 1.773)} \label{121027}
 
@@
 
-\subsection{GRB~121201A (z = 3.385)}
+\subsection{GRB~121201A (z = 3.385)} \label{121201}
 
@@
 
-\subsection{GRB~121209A (z = 2.1)}
+\subsection{GRB~121209A (z = 2.1)} \label{121209}
 
@@
 
-\subsection{GRB~121229A (z = 2.707)}
+\subsection{GRB~121229A (z = 2.707)} \label{121229}
 
@@
 spectroscopic integration starting 2 hours after the \swift~trigger yields a low
-S/N GRB afterglow continuum is all arms. Binning the spectrum reveals broad
+signal-to-noise GRB afterglow continuum is all arms. Binning the spectrum reveals broad
 absorption troughs, which we identify as \lyb~and \lya~at $z = 2.707$.
@@
 
-\subsection{GRB~130131B (z = 2.539)}
+\subsection{GRB~130131B (z = 2.539)} \label{130131}
 
@@
 
-\subsection{GRB~130408A (z = 3.758)}
+\subsection{GRB~130408A (z = 3.758)} \label{130408}
 
@@
 
-\subsection{GRB~130418A (z = 1.222)}
+\subsection{GRB~130418A (z = 1.222)} \label{130418}
 
@@
 
-\subsection{GRB~130427A (z = 0.340)}
+\subsection{GRB~130427A (z = 0.340)} \label{130427}
 
@@
 
-\subsection{GRB~130427B (z = 2.780)}
+\subsection{GRB~130427B (z = 2.780)} \label{130427}
 
@@
 twilight, 20.6 hours after the BAT trigger, captures a faint afterglow continuum
-visible across the entire spectral coverage at low S/N. Due to the low S/N, the
+visible across the entire spectral coverage at low signal-to-noise. Due to the low signal-to-noise, the
 metal lines are weak, but the broad absorption though due to \lya~is detected.
@@
 
-\subsection{GRB~130603B (z = 0.356)}
+\subsection{GRB~130603B (z = 0.356)}\label{130603}
 
@@
 
-\subsection{GRB~130606A (z = 5.913)}
+\subsection{GRB~130606A (z = 5.913)}\label{130606}
 
@@
 
-\subsection{GRB~130612A (z = 2.006)}
+\subsection{GRB~130612A (z = 2.006)}\label{130612}
 
@@
 published elsewhere. Starting only 1.1 hour after the \swift~trigger,
-$2\times600$ s spectroscopic integration captures a moderate S/N afterglow
+$2\times600$ s spectroscopic integration captures a moderate signal-to-noise afterglow
 continuum across the total spectral coverage of X-shooter. At a consistent
@@
 
-\subsection{GRB~130615A (z $\sim 3$)}
+\subsection{GRB~130615A (z $\sim 3$)} \label{130615}
 
@@
 
-\subsection{GRB~130701A (z = 1.155)}
+\subsection{GRB~130701A (z = 1.155)}\label{130701}
 
@@
 
-\subsection{GRB~130925A (z = 0.347)}
+\subsection{GRB~130925A (z = 0.347)}\label{130925}
 
@@
 
-\subsection{GRB~131011A (z = 1.874)}
+\subsection{GRB~131011A (z = 1.874)}\label{131011}
 
@@
 published elsewhere. Starting $\sim$1.5 days after the Fermi-GBM trigger, 4.5 ks
-spectroscopic integration captures a modest S/N GRB afterglow continuum all the
+spectroscopic integration captures a modest signal-to-noise GRB afterglow continuum all the
 way down to $\sim$3200 \AA. Imprinted on the continuum are absorption features,
@@
 
-\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}
+\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}	 \label{131030}
 
@@
 
-\subsection{GRB~131103A (z = 0.599)}
+\subsection{GRB~131103A (z = 0.599)}\label{131103}
 
@@
 additionally published in \citet{Kruhler2015}. Starting 5.8 hours after the BAT
-trigger, $4\times600$ s exposure captures a modest S/N continuum across all
+trigger, $4\times600$ s exposure captures a modest signal-to-noise continuum across all
 arms. Imprinted on the continuum are absorption features identifed as due to
@@
 
-\subsection{GRB~131105A (z = 1.686)}
+\subsection{GRB~131105A (z = 1.686)}\label{131105}
 
@@
 additionally published in \citet{Kruhler2015}. Starting only 1.3 hours after the
-\swift~trigger, a total of 4.8 ks spectroscopic integration contains a low S/N
+\swift~trigger, a total of 4.8 ks spectroscopic integration contains a low signal-to-noise
 GRB afterglow continuum across the entire spectral coverage of X-shooter. There
@@
 
-\subsection{GRB~131117A (z = 4.042)}
+\subsection{GRB~131117A (z = 4.042)}\label{131117}
 
@@
 4.8 ks spectroscopic integration secures afterglow continuum for this bursts,
-which is measured to be at $z = 4.042$. A moderate S/N GRB afterglow continuum
+which is measured to be at $z = 4.042$. A moderate signal-to-noise GRB afterglow continuum
 is detected down to $\sim$ 6100 \AA, signifying the onset of the \lya~forest,
@@
 
-\subsection{GRB~131231A (z = 0.642)}
+\subsection{GRB~131231A (z = 0.642)}\label{131231}
 
@@
 observed the following year (20.2 hours after the \swift~trigger), consists of
-$4\times600$ s exposures. A high S/N GRB afterglow continuum is detected all the
+$4\times600$ s exposures. A high signal-to-noise GRB afterglow continuum is detected all the
 way though the X-shooter arms. We identify absorption features imprinted in the
@@
 
-\subsection{GRB~140114A (z = 3.0)}
+\subsection{GRB~140114A (z = 3.0)}	\label{140114}
 
@@
 
-\subsection{GRB~140213A (z = 1.208)}
+\subsection{GRB~140213A (z = 1.208)}\label{140213}
 
@@
 \swift~trigger alert, $2\times600$ s spectroscopic integration contains a high
-S/N GRB afterglow continuum across the entirety of X-shooter. Imprinted on the
+signal-to-noise GRB afterglow continuum across the entirety of X-shooter. Imprinted on the
 afterglow continuum are absorption features, which we identify as metal
@@
 
-\subsection{GRB~140301A (z = 1.416)}
+\subsection{GRB~140301A (z = 1.416)}\label{140301}
 
@@
 additionally published in \citet{Kruhler2015}. Spectroscopic follow-up began 9
-hours after the BAT trigger and lasted for $12\times600$ s. A low S/N, spatially
+hours after the BAT trigger and lasted for $12\times600$ s. A low signal-to-noise, spatially
 extended continuum is visible across the entire spectral coverage of X-shooter.
-The GRB afterglow continuum is visible at moderate S/N on top of the underlying
+The GRB afterglow continuum is visible at moderate signal-to-noise on top of the underlying
 host continuum. The host exhibits usual nebular emission from \oii, \hb, \oiii,
@@
 
-\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}
+\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}	 \label{140311}
 
@@
 
-\subsection{GRB~140430A (z = 1.601)}
+\subsection{GRB~140430A (z = 1.601)}\label{140430}
 
@@
 for this burst began in twilight, 2.5 hours after the BAT trigger and lasted for
-$2\times600$ s. The spectrum contains a moderate S/N GRB afterglow continuum all
+$2\times600$ s. The spectrum contains a moderate signal-to-noise GRB afterglow continuum all
 the way through the spectroscopic arms of X-shooter. We identify absorption
@@
 
-\subsection{GRB~140506A  (z = 0.889)}
+\subsection{GRB~140506A  (z = 0.889)} \label{140506}
 
@@
 
-\subsection{GRB~140515A (z = 6.327)}
+\subsection{GRB~140515A (z = 6.327)}\label{140515}
 
@@
 
-\subsection{GRB~140614A (z = 4.233)}
+\subsection{GRB~140614A (z = 4.233)}\label{140614}
 
@@
 hours after the BAT trigger catches this GRB afterglow, which turns out to be at
-very high redshift. The continuum is detected at moderate S/N in both the VIS
+very high redshift. The continuum is detected at moderate signal-to-noise in both the VIS
 and NIR arms of X-shooter, heavily affected by absorption features. We identify
@@
 
-\subsection{GRB~140622A (z = 0.959)}
+\subsection{GRB~140622A (z = 0.959)}\label{140622}
 
@@
 
-\subsection{GRB~141028A (z = 2.332)}
+\subsection{GRB~141028A (z = 2.332)}\label{141028}
 
@@
 
-\subsection{GRB~141031A  (z = na)}
+\subsection{GRB~141031A  (z = na)}	\label{141031}
 
@@
 
-\subsection{GRB~141109A (z = 2.993)}
+\subsection{GRB~141109A (z = 2.993)}\label{141109}
 
@@
 used otherwise. Starting only 1.9 hours after the BAT trigger, these
-$4\times600$ s spectra contains a high S/N GRB afterglow. The afterglow
+$4\times600$ s spectra contains a high signal-to-noise GRB afterglow. The afterglow
 continuum is readily detected all the way down to 3700 \AA, although the bluest
@@
 
-\subsection{GRB~150206A (z = 2.087)}
+\subsection{GRB~150206A (z = 2.087)}\label{150206}
 
@@
 it does not enter into the statistical sample. Beginning in twilight,
-$4\times600$ s integration time contains a weak/moderate S/N GRB afterglow
+$4\times600$ s integration time contains a weak/moderate signal-to-noise GRB afterglow
 continuum throughout the X-shooter spectral coverage, down to $\sim$ 3750 \AA.
@@
 
-\subsection{GRB~150301B (z = 1.517)}
+\subsection{GRB~150301B (z = 1.517)}\label{150301}
 
@@
 on $6 \times 600$ s spectroscopic integration. The GRB afterglow continuum is
-well detected across the entire spectral coverage of X-shooter at moderate S/N.
+well detected across the entire spectral coverage of X-shooter at moderate signal-to-noise.
 Imprinted on the continuum are absorption features which we identify as being
@@
 
-\subsection{GRB~150403A (z = 2.057)}
+\subsection{GRB~150403A (z = 2.057)}\label{150403}
 
@@
 
-\subsection{GRB~150423A (z = 1.394)}
+\subsection{GRB~150423A (z = 1.394)}\label{150423}
 
@@
 totals $\sim$ 5000 s are combined to form this spectrum. A faint, almost
-featureless continuum is detected at low S/N all the way to the bluest part of
+featureless continuum is detected at low signal-to-noise all the way to the bluest part of
 the spectrum. An absorption doublet is detected in the VIS arm against the GRB
@@
 
-\subsection{GRB~150428A (z = na)}
+\subsection{GRB~150428A (z = na)}	\label{150428}
 
@@
 
-\subsection{GRB~150514A (z = 0.807)}
+\subsection{GRB~150514A (z = 0.807)}\label{150514}
 
@@
 observations began 28.4 post trigger and consist of a $4 \times 600$ s nodding
-sequence. The GRB afterglow continuum is detected at moderate S/N across the
+sequence. The GRB afterglow continuum is detected at moderate signal-to-noise across the
 entire spectral coverage of X-shooter. Narrow absorption features are imprinted
@@
 
-\subsection{GRB~150518A (z = 0.256)}
+\subsection{GRB~150518A (z = 0.256)}\label{150518}
 
@@
 
-\subsection{GRB~150616A (z = 1.188)}
+\subsection{GRB~150616A (z = 1.188)}\label{150616}
 
@@
 
-\subsection{GRB~150727A (z = 0.313)}
+\subsection{GRB~150727A (z = 0.313)}\label{150727}
 
@@
 
-\subsection{GRB~150821A (z = 0.755)}
+\subsection{GRB~150821A (z = 0.755)}\label{150821}
 
@@
 obtained. The observations began just before dawn and the last two exposures are
-heavily affected by the brightening sky. A bright, high S/N GRB afterglow
+heavily affected by the brightening sky. A bright, high signal-to-noise GRB afterglow
 continuum is detected all across the the spectral window and imprinted on this
@@
 
-\subsection{GRB~150910A (z = 1.359)}
+\subsection{GRB~150910A (z = 1.359)}\label{150910}
 
@@
 the BAT trigger, these observations were stopped after $2 \times 600$s, when it
-became apparent that only a modest S/N was obtainable, and the redshift had
+became apparent that only a modest signal-to-noise was obtainable, and the redshift had
 already been published( GCN
@@
 \#18274\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18274.gcn3}}).  The
-continuum is detected at low S/N all across X-shooter and \mgii~is detected at
+continuum is detected at low signal-to-noise all across X-shooter and \mgii~is detected at
 the suggested redshift, which we take as the redshift of the GRB.
 
-\subsection{GRB~150915A (z = 1.968)}
+\subsection{GRB~150915A (z = 1.968)}\label{150915}
 
@@
 began 3.3 hours after the \swift~trigger and lasted for $4 \times 1200$ s. A
-moderate S/N afterglow continuum is detected across the entire spectral coverage
+moderate signal-to-noise afterglow continuum is detected across the entire spectral coverage
 of X-shooter, where however the bluest part is affected by \lya-forest
@@
 
-\subsection{GRB~151021A (z = 2.330)}
+\subsection{GRB~151021A (z = 2.330)}\label{151021}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~151027B (z = 4.063)}
-
-\subsection{GRB~151029A (z = 1.423)}
-
-\subsection{GRB~151031A (z = 1.167)}
-
-\subsection{GRB~160117B (z = 0.87)}
-
-\subsection{GRB~160203A (z = 3.517)}
+\subsection{GRB~151027B (z = 4.063)}\label{151027}
+
+The content of these spectra has previously been issued in GCN \#
+18506\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18506.gcn3}}, but is not
+published. Beginning 5 hours after the BAT trigger, 4 spectroscopic integrations
+obtained in a nodding sequence, each lasting for 600s, securely allows us to
+measure a redshift for this GRB. Beginning at $\sim$ 4700 \AA, the \lya-forest
+is clearly detected at high signal-to-noise leading up to the broad
+\lya~absorption trough and the onset of a bright GRB afterglow continuum.
+Imprinted in the afterglow continuum are a range of absorption lines which we
+identify as due to \SIii, \SIii$^*$, \oi, \cii, \cii$^*$, \civ, \alii,  \alii,
+\feii, and \feii$^*$. These lines are all detected at $z = 4.063$ and the
+presence of fine-structure lines at a consistent redshift securely sets this as
+the redshift of the GRB. GRB~151027B was the 1000th GRB detected by \swift.
+
+\subsection{GRB~151029A (z = 1.423)}\label{151029}
+
+The gross content of these observations have previously been described in GCN \#
+18524\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18524.gcn3}}. Starting only 1
+hour after the BAT trigger, $2 \times 600$ s observations, extending into the
+morning twilight, were obtained of this GRB. The brightening sky affect the
+signal-to-noise of these spectra - especially in the blue end. The GRB afterglow
+is well detected over the entire spectral coverage of X-shooter, with spectral
+feaures from \feii~and \mgii~imprinted by an absorber at $z = 1.423$. The
+absorber er likely the host of the GRB and we therefore consider $z = 1.423$ the
+redshift of the GRB.
+
+\subsection{GRB~151031A (z = 1.167)}\label{151031}
+
+The spectral features in this spectrum has previously been reported in GCN
+\#18540\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18540.gcn3}}. Observed in
+RRM mode, these observations were initiated only 19 minutes after the BAT
+trigger. As per the standard RRM scheme, a series of stare observation,
+increasing in en exposure time and ending with a regular $4 \times 600$ s
+nodding sequence, were acquired for this burst. A moderate signal-to-noise
+continuum is detected all the way through the X-shooter spectral coverage. In
+the GRB afterglow continuum, we identify absorption features as \feii, \mgii,
+\mgi, and \caii~at a consistent redshift of $z = 1.167$. Supporting this as the
+redshift of the GRB host galaxy are the simultaneous detection of nebular
+emission lines from \oii, \hb, and \oiii.
+
+\subsection{GRB~160117B (z = 0.870)}	\label{160117}
+
+A preliminary analysis of these spectra have previously been presented in GCN
+\#18886\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18886.gcn3}}. Starting 13.5
+hours after the \swift~trigger, these spectra are formed on the basis of $4
+\times 1200$ s spectroscopic integration. The GRB afterglow is clearly detected
+at high signal-to-noise across the entire spectral coverage of X-shooter.
+Superposed on the continuum are absorption and emission features, which is
+identified as \feii, \mgii, \mgi~in absorption, and \oii, \hb, and \oiii. These
+spectral position of these lines all corrspond to a redshift of $z = 0.870$,
+which most likely is the redshift of the GRB.
+
+\subsection{GRB~160203A (z = 3.517)}\label{160203}
 
@@
 after the GRB trigger. We detect absorption features from a wide range of ions
-at $z=3.517$ as well as intervening absorption at $z=2.203$.
-
-\subsection{GRB~160228A (z = 1.64)}
-
-\subsection{GRB~160303A (z = na)}
-
-\subsection{GRB~160314A (z = 0.726)}
-
-\subsection{GRB~160410A (z = 1.717)}
-
-\subsection{GRB~160425A (z = 0.555)}
-
-\subsection{GRB~160625B (z = 1.406)}
-
-\subsection{GRB~160804A (z = 0.736)}
+at $z=3.517$ as well as intervening absorption at $z=2.203$. From the very high
+signal-to-noise spectrum we are able to measure \nh~$=21.75 \pm 0.10$, based on
+the shape of the \lya~profile. This burst is the subject of a forthcoming
+publication (Pugliese et al. in prep.).
+
+\subsection{GRB~160228A (z = 1.64)}	\label{160228}
+
+\subsection{GRB~160303A (z = na)}	\label{160303}
+
+\subsection{GRB~160314A (z = 0.726)}\label{160314}
+
+\subsection{GRB~160410A (z = 1.717)}\label{160410}
+
+\subsection{GRB~160425A (z = 0.555)}\label{160425}
+
+\subsection{GRB~160625B (z = 1.406)}\label{160625}
+
+\subsection{GRB~160804A (z = 0.736)}\label{160804}
+
 The data presented here also formed the basis of GCN \#
@@
 
-\subsection{GRB~161001A (z = 0.891?)}
-
-\subsection{GRB~161007A (z =4.6 ??? NEW!!!)}
+\subsection{GRB~161001A (z = 0.891?)}	\label{161001}
+
+\subsection{GRB~161007A (z =4.6 ???NEW!!!)} \label{161007}
 This data has not been published elsewhere. Observations for GRB~161007A started
@@
 
-\subsection{GRB~161014A (z =2.823)} The data presented here also formed the
+\subsection{GRB~161014A (z =2.823)} \label{161014}
+
+The data presented here also formed the
 basis of GCN \# 20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}},
@@
 
-\subsection{GRB~161023A (z = 2.710)}	 \label{GRB161023A}
-
-\subsection{GRB~161117A (z = 1.549)}
+\subsection{GRB~161023A (z = 2.710)}\label{161023}
+
+\subsection{GRB~161117A (z = 1.549)}\label{161117}
+
+\subsection{GRB~161219B (z = 1.549)}\label{161219}
 

### 2018-02-01 py/NH_distribution_plot.py

@@ -63,11 +63,33 @@ def main():
     # pl.legend()
     # Plot new distribution
     NH_new = NH[XSGRB]
+    NHe_new = NHe[XSGRB]
     z_new = z[XSGRB]
     g.x = z_new[NH_new > 20]
     g.y = NH_new[NH_new > 20]
     g = g.plot_marginals(sns.distplot, kde=False, color="#C44E52", bins = 25, norm_hist=True)
+
+
+
+
+
     g = g.plot_joint(pl.scatter, color="#C44E52", label="This work")
+
+
+
+    HIh = NH_new[NH_new > 20] + NHe_new[NH_new > 20]
+    HIl = NH_new[NH_new > 20] - NHe_new[NH_new > 20]
+
+    for ii, kk in enumerate(NH_new[NH_new > 20]):
+
+        print(z_new[NH_new > 20][ii], HIl[ii], HIh[ii])
+        g.x = np.array([z_new[NH_new > 20][ii], z_new[NH_new > 20][ii]])
+        g.y = np.array([HIl[ii], HIh[ii]])
+        g = g.plot_joint(pl.plot, color="#C44E52", lw=1)
+
+
+
+    # pl.show()
     g.x = 1
     g.y = 1
     g = g.plot_joint(pl.plot, color="#55A868", label="Tanvir et al. 2017")

### 2018-02-01 py/Sample_completeness.py

@@ -84,7 +84,7 @@ def main():
     print(stats.ks_2samp(fbo, filt_nan(z[pure == "Yes"])[~np.isnan(filt_nan(z[pure == "Yes"]))]))
     print(stats.ks_2samp(tough, filt_nan(z[pure == "Yes"])[~np.isnan(filt_nan(z[pure == "Yes"]))]))
     print(stats.ks_2samp(bat2, filt_nan(z[pure == "Yes"])[~np.isnan(filt_nan(z[pure == "Yes"]))]))
-    exit()
+    # exit()
     BAT_o = BAT[observed == "Yes"]
     XRT_o = XRT[observed == "Yes"]
     HI_o = HI[observed == "Yes"]
@@ -95,8 +95,17 @@ def main():
 
     swift_table = pd.read_table("../data/grb_table_1511519199.txt", delimiter="\t", dtype=None)
     name_s, BAT_s, XRT_s, HI_s = swift_table["GRB"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values, swift_table["XRT Column Density (NH) [10^21 cm^-2]"].values
-    # print(len(name_s))
+    # print(name_s)
     # exit()
+    counter = 0
+    for ii in name_s:
+        dat = ''.join(c for c in ii if c.isdigit())
+        print(dat)
+
+    exit()
+
+
+
     # Exclude sample bursts
     idx = [ii for ii, kk in enumerate(name_s) if kk not in name]
 
@@ -115,9 +124,9 @@ def main():
 
 
     # BAT fluence
-    sns.distplot(np.log10(1e-7*BAT_s), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 2, "linestyle": "dashed", 'cumulative': True}, label=r"\textit{Swift}")
-    sns.distplot(np.log10(1e-7*BAT_c), ax=ax1, kde=False, norm_hist=True, hist_kws={"histtype": "step", "alpha": 1, "linewidth": 2, "linestyle": "dashed", 'cumulative': True}, label="Statistical")
[…]

### 2018-02-01 py/betaOX.py

@@ -18,7 +18,7 @@ params = {
    'legend.fontsize': 10,
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
-   'text.usetex': False,
+   'text.usetex': True,
    'figure.figsize': [7.281, 4.5]
    }
 
@@ -105,12 +105,29 @@ def main():
     # Get Swift HI values
     swift_table = pd.read_table("../data/allSwiftGRBs_NH.txt", delimiter="\t", dtype=None)
     name_s, z_s, HI_s, HI_sh, HI_sl = swift_table["#GRB"].values, swift_table["Redshift"].values, swift_table["Ave_NH_PC"].values, swift_table["Ave_dNH_PC+"].values, -1*swift_table["Ave_dNH_PC-"].values
+    # print(HI_s)
+
+
+    # # Read in Buchner catalog
+    # grbcat = fits.open("../data/grbcat-annotated.fits")
+    # # print(grbcat[1].data.field)
+    # name_s = np.array(["".join(kk.split(" "))[3:] for kk in grbcat[1].data.field("Name")])
+    # HI_s, HI_sl, HI_sh, nH_std, z_s = grbcat[1].data.field("NH_sphere_mean"), grbcat[1].data.field("NH_sphere_hi"), grbcat[1].data.field("NH_sphere_lo"), grbcat[1].data.field("NH_sphere_std"), grbcat[1].data.field("z")
+ 
+
 
     # Find values with betaOX
     idx = [ii for ii, kk in enumerate(name_s) if kk in name and ~np.isnan(z_s[ii])]
     idx_2 = [ii for ii, kk in enumerate(name) if kk in name_s[idx]]
     f_idx = [ii for ii, kk in enumerate(name_s) if kk in f_name and ~np.isnan(z_s[ii])]
     f_idx_2 = [ii for ii, kk in enumerate(f_name) if kk in name_s[f_idx]]
+
+
+    # HI = filt_nan(HI_s[idx], fill_value=0)
+    # HIh = filt_nan(HI_sh[idx], fill_value=0)
+    # HIl = filt_nan(HI_sl[idx], fill_value=0)
+    # f_HI = filt_nan(HI_s[f_idx], fill_value=0)
+    # f_HI[np.isnan(f_HI)] = 0
 
     HI = np.log10(1e22*filt_nan(HI_s[idx], fill_value=0))
[…]

### 2018-02-01 py/skymap.py

@@ -19,7 +19,7 @@ params = {
    'legend.fontsize': 10,
    'xtick.labelsize': 10,
    'ytick.labelsize': 10,
-   'text.usetex': False,
+   'text.usetex': True,
    'figure.figsize': [9, 4.5]
    }
 matplotlib.rcParams.update(params)
@@ -45,33 +45,61 @@ def main():
     """
     # Script to produce sky plots.
     """
-    # ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
-    # A_Vmap = ebv_map*3.1
-    # A_Vmap[A_Vmap >= 0.5] = 0
-    # ax = hp.mollview(A_Vmap, max=0.5, title="", cbar=False)
+    ebv_map = hp.read_map("../data/lambda_sfd_ebv.fits")
+    A_Vmap = ebv_map*3.1
+    A_Vmap_cut = A_Vmap.copy()
+    A_Vmap_cut[A_Vmap >= 0.5] = 0
+    ax = hp.mollview(A_Vmap_cut, max=0.5, title="", cbar=False)
+
     # , cbar=False
 
 
-    # fig = pl.gcf()
-    # ax = pl.gca()
-    # image = ax.get_images()[0]
-    # cmap = fig.colorbar(image, ax=ax, location="bottom", pad = 0.02, label=r"$A_V$", shrink=0.5)
-    # hp.graticule()
+    fig = pl.gcf()
+    ax = pl.gca()
+    image = ax.get_images()[0]
+    cmap = fig.colorbar(image, ax=ax, location="bottom", pad = 0.02, label=r"$A_V$", shrink=0.5)
+    hp.graticule()
+
+    # A_Vmap_cut = A_Vmap.copy()
+    # A_Vmap_cut[A_Vmap <= 0.5] = np.nan
+    # ax = hp.mollview(A_Vmap_cut, title="", cbar=False, norm="hist", cmap="viridis", ax=ax)
+
+    # deep=["#4C72B0", "#55A868", "#C44E52",
+    #       "#8172B2", "#CCB974", "#64B5CD"],
 
     # Overplot swift positions
     # Read in master sample
     burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
     name, z, observed, OA, = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values
+
+
+    hasz = []
+    for ii, kk in enumerate(z):
+        try:
[…]

### 2018-02-01 py/spec_plot.py

@@ -80,7 +80,7 @@ def main():
         ax.axvspan(18000, 19500, color = "grey", alpha = 0.4)
     # Save figure for tex
     # pl.legend()
-    ax4.set_ylabel(r"$F_{\lambda}\,\rm{(10^{-17}\,erg\,s^{-1}\,cm^{-2}\, \AA^{-1})}$")
+    ax4.set_ylabel(r"$F_{\lambda}$")
     # ax7.set_xlabel(r"$\rm{Observed\,wavelength\,}$ (\AA)")
     ax7.set_xlabel(r"Observed Wavelength (\AA)")
     pl.tight_layout()

### 2018-02-02 document/XSGRB_sample.tex
GCN citation and last comments, individual notes, Daniele contrib

@@
 of afterglows and host galaxies of GRBs detected by {\it Swift}. We here make
-all data resulting from the survey, publicly available in reduced form.
+all data resulting from the survey, publicly available in reduced form (see Sect
+\ref{products}).
 
@@
 \begin{enumerate}
+
 	\item GRB triggered by BAT onboard the \swift~satellite
-	\item XRT started observing within 10 minutes after the GRB; an XRT position must be distributed within 12 hr.
-	\item The target must be visible for at least 60 minutes, 30 degrees above the horizon, with the Sun below $-12$ degrees\footnote{Note that in the P84 proposal the criteria have
-	been stated a bit differently, the visibility constraint being replaced by a
-	declination + Sun angle constraint. These criteria are, however, those
-	defining the sample.}.
-	\item No bright, nearby stars within $ 1.8 + 0.4 \times exp[(20 - R)/2.05$] arcsec, where R is the USNO magnitude of the star)
-	\item Galactic $A_V \lesssim 0.5$ mag.
+
+	\item XRT started observing within 10 minutes after the GRB; an XRT position
+	must be distributed within 12 hr.
+
+	\item The target must be visible for at least 60 minutes, 30 degrees above the
+	horizon, with the Sun below $-12$ degrees\footnote{Note that in the P84
+		proposal the criteria have been stated differently, the visibility constraint
+		being replaced by a declination + Sun angle constraint. The criteria listed
+		above are, however, those defining the sample.}.
+
+	\item No bright, nearby stars within $ 1.8 + 0.4 \times \exp[(20 - R)/2.05$]
+	arcsec, where $R$ is the USNO magnitude of the star)
+
+	\item Galactic $A_V \lesssim 0.5$ mag according to the maps of \citet{Schlegel1998}.
+
 \end{enumerate}
@@
 Because the localization accuracy of BAT is 1 -- 4 arcminutes
-\citep{Barthelmy2005}, a secure host association is in general very hard based
-on BAT alone. We therefore additionally require an X-ray position from XRT to be
-distributed to the GCN network \citep{Barthelmy2000} within 12 hours and to
-account for observing constraints on \textit{Swift} that XRT began observations
-within 10 minutes. The additional timing requirement of the XRT follow-up means
-that all bursts in our sample have detected X-ray afterglows. Because the XRT
-completeness is very high \citep{Burrows2007}, this cut should not alter the
+\citep{Barthelmy2005}, an afterglow identification based on BAT alone would be
+harder, and  host association impossible. We therefore additionally require an
+X-ray position from XRT to be distributed to the GCN network
+\citep{Barthelmy2000} within 12 hours and to account for observing constraints
+on \textit{Swift} that XRT began observations within 10 minutes. The additional
+timing requirement of the XRT follow-up means that all bursts in our sample have
+detected X-ray afterglows. Because the XRT completeness is very high for
+promptly-repointed GRBs \citep{Burrows2007}, this cut should not alter the
 parent sample significantly.
@@
 population. The same arguments apply to the requirement that there are no nearby
-bright stars. We additionally require that the Galactic extinction is below $A_V
-\lesssim 0.5$ mag, based on the extinction maps by \citet{Schlegel1998,
-	Schlafly2011}. Due to the isotropic distribution of GRBs on the sky, these
-additional cuts should not influence the optical properties of the bursts
+foreground bright stars. We additionally require that the Galactic extinction is
+below $A_V \lesssim 0.5$ mag, based on the extinction maps by
+\citet{Schlegel1998}\footnote{We use the updated values of \citet{Schlafly2011}
+	to correct for foreground extinction, but the sample criterion is based on the
+	old \citet{Schlegel1998} values for consistency with the first semesters of our
+	program.}. Choosing low-extinction sightlines also reduce the problem of field
+crowding and the contamination from Galactic outbursts posing as GRB impostors.
+These additional cuts should not influence the optical properties of the bursts
 themselves, only our ability to successfully secure the observations that allows
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Follow-up procedure and practicalities}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\todo{Will be written by Daniele}
+\subsection{Follow-up procedure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Our collaboration has set up a procedure to promptly react to alerts from
+\swift~and other satellites (such as \textit{Fermi}). We process automatically
+GCN notices in order to flag those events belonging to our statistical sample.
+Two people are permanently on alert to manually supervise each event, for
+example to recognize events of special interest beyond their inclusion in the
+sample. Members of our collaboration have access to numerous facilities spread
+throughout the world, among which the Nordic Optical Telescope (NOT), the Gran
+Telescopio Canarias (GTC), the Telescopio Nazionale Galileo (TNG), the Gamma-Ray
+Burst Optical Near-infrared Detector (GROND), the Xinlong observatory. Thanks to
+this network, we can often conduct searches for optical/NIR afterglows ahead of
+the time of the X-shooter spectroscopy, thus helping to plan the spectroscopic
+observations. If this is not possible, we use the built-in acquisition camera of
+X-shooter to search for counterparts, and adapt the observing strategy on the
+fly. In many cases we could interact directly with the staff at the telescope to
+aid the observation. Raw data are usually available within minutes in the ESO
+archive LINK!!! and are promptly reduced by members of our collaboration, often
+using archival calibration data which are readily available. This allows us in
+most cases to report the preliminary results (redshift, identification of the
+most prominent emission and absorption feature, etc.) within a few hours after
+the beginning of the observations.
 
@@
 
-Under rare circumstances, the use of the ESO rapid response mode (RRM) has been
-possible. RRM is an automatic override of current observations at the telescope
-that allows the shortest possible delays between the GRB trigger and initiation
-of spectroscopic observations. This allows spectroscopic integration to begin on
-the rapidly fading, optical transient within minutes of the burst. In case of a
-promptly visible GRB, a robotic trigger was sent to the telescope if at the
-moment of the GCN notice the GRB is observable and fulfills the following
-criteria:
+Under rare circumstances, the use of the ESO rapid response mode (RRM;
+\citealt{Vreeswijk2010}) has been possible. RRM is a system which automatically
+overrides current observations at the telescope. This allows the shortest
+possible delay between the GRB trigger and initiation of observations. This
+allows spectroscopic integration to begin on the rapidly fading, optical
+transient within minutes of the burst. In case of a promptly visible GRB, a
+robotic trigger was sent to the telescope if at the moment of the GCN notice the
+GRB is observable and fulfills the following criteria:
 
@@
 	\item GRB triggered by \textit{Swift}.
-	\item X-shooter mounted at the telescope.
-	\item GRB promptly visible.
+	\item X-shooter is the currently operating instrument at the telescope.
+	\item GRB promptly visible from Paranal.
 	\item Error radius < 60\arcsec.
@@
 
+The criteria to trigger RRM are looser than those to define the statistical
+sample. This is because of the expected larger brightness of any GRB counterpart
+minutes after the explosion, and because of the rarity of RRM triggers.
+
 The use of RRM is unique in the sense that it allows for the detailed study of
@@
 
-There are nine GRBs in the sample that have been observed with the RRM mode.
-One of the RRM triggers is outside the sample we use to address the statistical
-properties of GRB afterglows and two of the RRM trigger are on short GRBs. The
-fastest response between \swift~GRB trigger and the beginning of spectroscopic
-integration is for GRB~160410A, for which the delay was only 8.4 minutes. In the
-cases where the RRM trigger was rejected, it was in most cases due to a wrong
-instrument being mounted at the focus.
+There are nine GRBs that have been observed by us with the RRM mode. One of the
+RRM triggers is outside the statistical sample  and two of the RRM trigger are
+on short GRBs. The fastest response between \swift~GRB trigger and the beginning
+of spectroscopic integration was for GRB~160410A, for which the delay was only
+8.4 minutes. In many cases, unfortunately, we could not use RRM even for promptly visible events, because of the unavailability of X-shooter.
 
 Potentially, the proximity of the telescope (Paranal, Chile) to the South
-Atlantic Anomaly (SSA) affects the rate of RRM triggers. BAT switches off when
-the satellite goes through the SSA and that means we have less triggers of GRBs
-that are immediately observable, compared to the rest of the \swift~orbit.
+Atlantic Anomaly (SSA) affects the rate of RRM triggers \citep{Greiner2011}. BAT
+switches off when the satellite goes through the SSA and that means we have
+fewer GRB triggers immediately observable, compared to the rest of
+the \swift~orbit.
 
@@
 
-The first GRB followed up was GRB~090313, observed on the 15th of March, 2009,
-during the commissioning of X-shooter on the third Unit Telescope (UT3) of the
-VLT . The bursts observed during the commissioning or the science verification
-process (GRB~090313, GRB~090530, GRB~090809, GRB~090926) are not a part of the
-sample we use to investigate the statistical properties of GRB afterglows, due
-to the inhomogeneity of their selection. The first burst observed after science
-verification and X-shooter was moved to UT2, was GRB~091018, which is the first
-burst entering our statistically homogeneous sample. For all bursts that fulfill
-our sample selection criteria, described in Sect. \ref{samplecrit},
-spectroscopic follow-up have been attempted with X-shooter. Various conditions
-can affect our ability to follow up a given burst, and a discussion of these
-conditions and their consequences for the sample is included in Sect.
-\ref{badbursts}.
+The first GRB followed up was GRB~090313 \citep{DeUgartePostigo2010}, observed
+on the 15th of March, 2009, during the commissioning of X-shooter on the third
+Unit Telescope (UT3) of the VLT . The bursts observed during the commissioning
+or the science verification process (GRB~090313, GRB~090530, GRB~090809,
+GRB~090926) are not a part of the sample we use to investigate the statistical
+properties of GRB afterglows, due to the inhomogeneity of their selection. The
+first burst observed after science verification and X-shooter was moved to UT2,
+was GRB~091018, which is the first burst entering our statistically homogeneous
+sample. For all bursts that fulfill our sample selection criteria, described in
+Sect. \ref{samplecrit}, spectroscopic follow-up have been attempted with
+X-shooter. Various conditions can affect our ability to follow up a given burst,
+and a discussion of these conditions and their consequences for the sample is
+included in Sect. \ref{badbursts}.
 
@@
 images where the background is varying, which is often the case when a burst has
-been observed close to twilight
+been observed close to twilight.
 
@@
 employed. When there are no contaminating sources in the slit, the sky has been
-put back on the images and the images are combined in pairs of two, subtracting
-the two from each other while keeping the WCS static. Due to the nod-and-shuffle
-offsets used between observations, this conserves the source flux while removing
-the sky at the expense of a decrease in signal-to-noise by a factor of
-$\sqrt{2}$. This amounts to the regular nodding reduction, only we can reject
-outliers and weight by the averaged inverse variance map.
+put back on the images and the images are combined in pairs of two before
+extraction, subtracting the two from each other while keeping the WCS static.
+Due to the nod-and-shuffle offsets used between observations, this conserves the
+source flux while removing the sky at the expense of a decrease in
+signal-to-noise by a factor of $\sqrt{2}$. This amounts to the regular nodding
+reduction, only we can reject outliers and weight by the averaged inverse
+variance map.
 
@@
 get a spectrum of the sky which we can use to recalibrate the wavelength
-solution.
+solution in the post-processing steps.
 
@@
 during the X-shooter Active Flexure Compensation procedure\footnote{X-shooter
-	User Manual available at \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}}. The
+	User Manual available at \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}.}. The
 remaining offset is corrected by cross-correlating the observed sky spectrum
@@
 periods of time were applied to the corresponding spectra\footnote{The
-	wavelength shifts have been extensively studied by
-	ESO \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
-Using the convolved synthetic sky, the pixels containing the brightest sky lines have been added to the bad pixel map.
+	wavelength shifts have been extensively studied by ESO
+	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
+Using the convolved synthetic sky, the pixels containing the brightest sky lines have been flagged as such in the bad pixel map.
 
@@
 observations have been queried for the ambient conditions which controls the
-degree to which the trace changes position in the spatial direction.
+degree to which the trace changes position in the spatial direction. This
+analytic approach is only valid for a plane-parallel atmosphere, but because the
+final position is refined in the fit, it is adequate for our purposes.
 
 Based on the signal-to-noise of the afterglow continuum, the 2D-image has been
-binned down in the spectral direction to a number of elements that allows for an
+binned in the spectral direction to a number of elements that allows for an
 accurate tracing of the PSF, typically 200 bins for moderate signal-to-noise.
@@
 extended object, the fits have been inspected to ensure that the additional
-component does now skew the fit towards a different PSF. The additional
+component does not skew the fit towards a different PSF. The additional
 components are not used for the weights in the extractions.
@@
 \citet{Schlafly2011}\footnote{Queried from
-	http://irsa.ipac.caltech.edu/applications/DUST/index.html using \texttt{astroquery}
-	\citep{astroquery}.}, and the extinction curve by \cite{Cardelli1989} with a
-total to selective extinction $R_V = 3.1$. The wavelengths of the extracted
-1D-spectra are moved to vacuum, corrected for barycentric motion, and the
-wavelength recalibration described in Sect. \ref{wavecal} is applied. Pixels
-with pixel-to-pixel variation larger than $50 \sigma$ are additionally added to
-the bad pixel map.
+	http://irsa.ipac.caltech.edu/applications/DUST/index.html using
+	\texttt{astroquery} \citep{astroquery}.}, and the extinction curve by
+\cite{Cardelli1989} with a total to selective extinction $R_V = 3.1$. The
+wavelengths of the extracted 1D-spectra are wavelength recalibrated (described
+in Sect. \ref{wavecal}), moved to vacuum, and corrected for barycentric motion.
+Pixels with pixel-to-pixel variation larger than $50 \sigma$ are additionally
+added to the bad pixel map.
 
@@
 linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
-sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$. We use this linear
-relation as a way to estimate the spectral resolution for medium to low signal
-to noise afterglow spectra in the VIS arm. To extend this to the UVB and NIR
-arm, we measured a series of lines in the NIR arm spectra of a subset of 19 sources
-used for the VIS arm above, and find that the resulting distribution is
-consistent with a simple scaling of the VIS arm relation by the ratio of
-resolutions of the NIR and VIS arm for unresolved, slit filling, sources as
-given on the ESO instrument website. The UVB arm contains no suitable absorption
-lines to use, and we therefore use a scaled value as in the NIR arm. This simple
-analysis gives a sufficiently accurate estimate for the analysis of the low
-signal-to-noise science spectra.
+sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
+\ref{fig:res}). We use this linear relation as a way to estimate the spectral
+resolution for medium to low signal to noise afterglow spectra in the VIS arm.
+To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
+arm spectra of a subset of 19 sources used for the VIS arm above, and find that
+the resulting distribution is consistent with a simple scaling of the VIS arm
+relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
+filling, sources as given on the ESO instrument website. The UVB arm contains no
+suitable absorption lines to use, and we therefore use a scaled value as in the
+NIR arm. This simple analysis gives a sufficiently accurate estimate for the
+analysis of the low signal-to-noise science spectra.
 
@@
 afterglow observations with X-shooter. An overview of all the spectra and their
-observational setups are given in Table \ref{tab:sample_overview}. For each
+observational setups is given in Table \ref{tab:sample_overview}. For each
 burst, each individual observation is provided in a separate reduction, and in
@@
 \begin{itemize}
-	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion and drifts in the wavelength solution ($\mathrm{\AA}$).
-	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
-	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
-	\item QUAL - Bad pixel map, where a value different from zero indicates a bad pixel.
-	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+
+	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion
+	and drifts in the wavelength solution ($\mathrm{\AA}$).
+
+	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1}
+	\mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+
+	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1}
+	\mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+
+	\item QUAL - Bad pixel map, where a value different from zero indicates a bad
+	pixel.
+
+	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum}
+	($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2} \mathrm{\AA}^{-1}$).
+
 	\item CONTINUUM\_ERR - Relative error on the continuum estimate.
-	\item TELL\_CORR - Inverse transmission spectrum. Multiply FLUX and ERR column with this column to correct for telluric absorption. 
+
+	\item TELL\_CORR - Inverse transmission spectrum. Multiply FLUX and ERR column
+	with this column to correct for telluric absorption.
+
 \end{itemize}
@@
 statistical sample is the bright \textit{INTEGRAL} burst GRB~161023A, that
-contains at least 11 intervening absorption systems (See \ref{161023}).
+contains at least 15 intervening absorption systems (See \ref{161023}).
 
@@
 spectroscopic redshift measurement relies on the detection of absorption lines
-imprinted on the GRB afterglow continuum. Because the afterglow rapidly fades
-(typically as $\sim t^{-1}$) a rapid follow-up is essential. In Fig.
+imprinted on the GRB afterglow continuum. Because the optical afterglow rapidly
+fades (typically as $\sim t^{-1}$) a rapid follow-up is essential. In Fig.
 \ref{fig:timing} we plot the delay from the BAT trigger to the start of the
@@
 
-\begin{figure*}
+\begin{figure*}[!ht]
 	\centering \includegraphics[width=16cm]{figures/XSGRB_redshift.pdf}
@@
 	comparison, we overplot with red stars the GRB \citep{Tanvir2009b,
-		Salvaterra2009a}, quasar \citep{Banados2017}, and galaxy \citep{Oesch2016} with
-	the highest spectroscopically confirmed redshifts, the latter two shown at
-	arbitrary $E_{\mathrm{BAT}}$. The blue solid line represents the so-called 1 s
-	BAT sensitivity limit described in the text. The estimated epoch of
-	reionization is shown by the black dot-dashed line, with the uncertainty shown
-	as the black dashed lines, from the most recent measurement by
+		Salvaterra2009a}, quasar \citep{Banados2017}, and galaxies \citep{Zitrin2015,
+		Oesch2016} with the highest spectroscopically confirmed redshifts, the latter
+	three shown at arbitrary $E_{\mathrm{BAT}}$. The blue solid line represents the
+	so-called 1 s BAT sensitivity limit described in the text. The estimated epoch
+	of reionization is shown by the black dot-dashed line, with the uncertainty
+	shown as the black dashed lines, from the most recent measurement by
 	\citet{Planck2015}. On the top of the plot is shown the marginalization of the
@@
 that dark bursts have somewhat higher X-ray luminosity and $N_{\mathrm{HI, X}}$
-compared to the optically bright bursts \citep{Campana2012}. This indicates along with
-investigations of host galaxy properties \citep{Greiner2011, Kruhler2011,
-	Hjorth2012, Perley2016b}, that the extinction of the optical afterglows is
-primarily driven by the presence of dust in the host galaxies and not solely by
-unfortunate placement of the synchrotron spectral break frequencies.
-\citet{Hjorth2012} find that systems with no optical afterglow have higher
-$N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which however
-also turn out to be redder.
+compared to the optically bright bursts \citep{Campana2012}. This indicates
+along with investigations of host galaxy properties \citep{Greiner2011,
+	Kruhler2011, Hjorth2012, Perley2016b}, that the extinction of the optical
+afterglows is primarily driven by the presence of dust in the host galaxies and
+not solely by unfortunate placement of the synchrotron spectral break
+frequencies. \citet{Hjorth2012} find that systems with no optical afterglow have
+higher $N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which
+however also turn out to be redder. Additionally, the ISM absorption lines in
+dark sight-lines are found to be stronger compared to optically brighter bursts
+\citep{Christensen2011a}, which is consistent with the dark bursts being found
+in more metal-rich and dustier galaxies.
 
@@
 hypothesis that they are drawn from the same distribution with $p = 0.43$,
-meaning there is no strong evidence for a discrepancy. A Kendall's $\tau$ test
-suggest a non-significant, very low degree of negative correlation ($\Gamma =
--0.15$ at a p-value = 0.10).
+meaning that there is no strong evidence for a discrepancy. A Kendall's $\tau$
+test suggest a non-significant, very low degree of negative correlation ($\Gamma
+= -0.15$ at a p-value = 0.10).
 
@@
 The locations of long GRBs are associated with intensely star-forming regions
-\citep{Hogg1999, Bloom2002, Fruchter2006}. Because a significant fraction of the
-hydrogen along the line of sight in these regions has not been ionized, the
-optical depth at the wavelength of \lya~is very high, saturating not only the
-line center, but also the damping wings. This causes a strong absorption system
-from the \lya-transition to appear in the afterglow continuum. For bursts with
-$z \gtrsim 1.7$, the position of \lya~moves into the spectroscopic coverage of
-X-shooter, meaning that we are able to detect this absorption trough due to
+\citep{Hogg1999, Bloom2002, Fruchter2006, Lyman2017}. Because a significant fraction
+of the hydrogen along the line of sight in these regions has not been ionized,
+the optical depth at the wavelength of \lya~is very high, saturating not only
+the line center, but also the damping wings. This causes a strong absorption
+system from the \lya-transition to appear in the afterglow continuum. For bursts
+with $z \gtrsim 1.7$, the position of \lya~moves into the spectroscopic coverage
+of X-shooter, meaning that we are able to detect this absorption trough due to
 \lya. The exceptionally good UV response of X-shooter, allows us to robustly
@@
 systems, along with strong absorption from \mnii~and \SIiii, making it
-notoriously difficult to model. Additionally, the red wing has got the ISM
-signatures imprinted on it, especially strong absorption due to \SIii, \sii~and
-\nv, which can exhibit significant velocity structure. Along with instrumental
-effects, the generative model for the data that we would use in a
-likelihood-based analysis would be very complicated. We have therefore decided
-not to make formal $\chi^2$ fitting of the hydrogen column densities, but
-instead use a more subjective visual measurement to the absorption profile. 
-Using an analytic approximation to the absorption profile from
-\citet{TepperGarcia2006}, we overplot a synthetic absorption line with a
-specified column density on our observed, normalised spectrum. By tuning the
-value of the hydrogen column density until the synthetic absorption line matches
-the spectrum, we can thereby infer the actual column density of the GRB sight
-line in a manual way. Similarly, the uncertainty on the hydrogen column can be
-estimated by adjusting the error, until the confidence bounds contain the
-continuum variation. We show the results of this procedure for all bursts where
-possible in Fig. \ref{fig:HI1} and the inferred hydrogen column densities in
-Table~\ref{tab:HI}.
-
-\input{tables/HI_columns.tex}
+notoriously difficult to model. Additionally, the red wing has ISM signatures
+imprinted on it, especially strong absorption due to \SIii, \sii~and \nv, which
+can exhibit significant velocity structure. Along with instrumental effects, the
+generative model for the data that we would use in a likelihood-based analysis
+would be very complicated. We have therefore decided not to make formal $\chi^2$
+fitting of the hydrogen column densities, but instead use a more subjective
+visual measurement to the absorption profile. Using an analytic approximation to
+the absorption profile from \citet{TepperGarcia2006}, we overplot a synthetic
+absorption line with a specified column density on our observed, normalised
+spectrum. By tuning the value of the hydrogen column density until the synthetic
+absorption line matches the spectrum, we can thereby infer the actual column
+density of the GRB sight line in a manual way. Similarly, the uncertainty on the
+hydrogen column can be estimated by adjusting the error, until the confidence
+bounds contain the continuum variation. We show the results of this procedure
+for all bursts where possible in Fig. \ref{fig:HI1} and the inferred hydrogen
+column densities in Table~\ref{tab:HI}.
 
@@
 	values presented in this sample. We also show in the red the values derived in
-	this work. The marginal distributions for the three distributions are also shown
+	this work. The marginal distributions for the three samples are also shown
 	along the left side and on the top, where the different environments probed are
@@
 
+\input{tables/HI_columns.tex}
+
 \begin{acknowledgements}
-	*** We should add an acknowledgment to Neil Gehrels and Javier Gorosabel here
-***. 
+Much of the work done here would not have been done without Neil Gehrels, who has now passed away.
+We owe him a large debt for his invaluable work with the \swift~satellite.
+We also want to acknowledge Javier Gorosabel, who was taken from us too early.
 %
@@
 \bibliographystyle{mnras}
-\bibliography{XSGRB_sample}
+\bibliography{XSGRB_sample,GCN}
 
@@
 & = \int_{-\infty}^{\infty} \frac{1}{\sqrt{2 \pi} \sigma} e^{- \left( \frac{\xi}{\sqrt{2} \sigma}  \right)^2 } \frac{1}{\gamma \pi} \frac{\gamma^2}{(\lambda - \xi)^2 + \gamma^2} d\xi \\
-& = \frac{\gamma}{\sqrt{2} \sigma} \frac{1}{ \pi^{3/2}}   \int_{-\infty}^{\infty} \frac{e^{- \left( \frac{\xi}{\sqrt{2} \sigma}  \right)^2 }}{(\lambda - \xi)^2 + \gamma^2}.
+& = \frac{\gamma}{\sqrt{2} \sigma} \frac{1}{ \pi^{3/2}}   \int_{-\infty}^{\infty} \frac{e^{- \left( \frac{\xi}{\sqrt{2} \sigma}  \right)^2 }}{(\lambda - \xi)^2 + \gamma^2} d\xi.
 \end{split}
@@
 The first GRB ever observed with X-shooter, during the commissioning of the
-instrument, this data formed the basis of GCN
-\#9015\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9015.gcn3}} and is published
-in \citet{DeUgartePostigo2010}. Due to the lingering brightness of GRB~090313,
-6.9 ks spectroscopic integration starting 45 hours after the BAT trigger reveals
-a wealth of absorption features superposed on the afterglow continuum at a
-common redshift of $z = 3.373$. Two intervening systems at $z = 1.959$ and $z =
-1.800$ are identified based on strong \mgii-absorption. Because this burst is
-observed before the instrument is science-verified, it does not enter into the
+instrument, this data formed the basis of GCN 9015 \citep{GCN.9015} and is
+published in \citet{DeUgartePostigo2010}. Due to the lingering brightness of
+GRB~090313, 6.9 ks spectroscopic integration starting 45 hours after the BAT
+trigger reveals a wealth of absorption features superposed on the afterglow
+continuum at a common redshift of $z = 3.373$. Two intervening systems at $z =
+1.959$ and $z = 1.800$ are identified based on strong \mgii-absorption. Because
+this burst is observed before the instrument is science-verified, it does not
+enter into the statistical sample.
+
+\subsection{GRB~090530 (z = 1.266)}\label{090530}
+
+Observed during paranalization of the instrument, this data forms the basis of
+GCN 15571 \citep{GCN15571}, but is not published elsewhere. Observations began
+20.6 hours after the BAT trigger and 4.8 ks spectroscopic integration in all
+three arms reveals the absorption signature for a host at $z = 1.266$ from the
+detection of \mgii, \mgi, \SIii, \feii, \aliii. Because this burst was observed
+before the instrument is science-verified, it does not enter into the
 statistical sample.
 
-\subsection{GRB~090530 (z = 1.266)}\label{090530}
-
-Observed during paranalization of the instrument, this data forms the basis of
-GCN \#15571\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15571.gcn3}}, but is not
-published elsewhere. Observations began 20.6 hours after the BAT trigger and 4.8
-ks spectroscopic integration in all three arms reveals the absorption signature
-for a host at $z = 1.266$ from the detection of \mgii, \mgi, \SIii, \feii,
-\aliii. Because this burst was observed before the instrument is
-science-verified, it does not enter into the statistical sample.
-
 \subsection{GRB~090809 (z = 2.737)} \label{090809}
@@
 Observed during the first science verification period and was reported in GCN
-\#9761\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9761.gcn3}} and is
-additionally used as the basis for the master thesis by \'Asa Sk\'ulad\'ottir
-(2010). 7.2 ks integration starting 10.2 hours after the GRB trigger yields a
-clear afterglow continuum in all arms. The simultaneous detection of absorption
-lines identified as \lya, \SIii, \oi, \SIi$^*$, \SIiv, \civ, \feii, \alii,
-\aliii~and \mgii~at $z = 2.737$ sets it as the redshift of the GRB. Because this
-burst is observed before the instrument was science-verified, it does not enter
-into the statistical sample.
+9761 \citep{GCN.9761} and is additionally used as the basis for the master
+thesis by \'Asa Sk\'ulad\'ottir (2010). 7.2 ks integration starting 10.2 hours
+after the GRB trigger yields a clear afterglow continuum in all arms. The
+simultaneous detection of absorption lines identified as \lya, \SIii, \oi,
+\SIi$^*$, \SIiv, \civ, \feii, \alii, \aliii~and \mgii~at $z = 2.737$ sets it as
+the redshift of the GRB. Because this burst is observed before the instrument
+was science-verified, it does not enter into the statistical sample.
 
@@
 Obtained during the second science
-verification period, this dataset forms the basis of GCN
-\#9942\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/9942.gcn3}} and is
+verification period, this dataset forms the basis of GCN 9942 \citep{GCN.9942} and is
 additionally published in \citet{DElia2010}. Spectroscopic integration started
@@
 causes a strong continuum to be seen in all arms. An absorption trough due to
-\lya is clearly visible along with numerous metal resonance lines \civ, \SIii,
+\lya~is clearly visible along with numerous metal resonance lines \civ, \SIii,
 \SIii$^*$ \feii, \mgii, all at $z = 2.106$. Because this burst was observed
@@
 The first burst observed during normal operation after science verification was
-completed and there is the first burst that enter the statistical sample. This
-data is the basis for GCN \#
-10042\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10042.gcn3}} and is published
-in \citet{Wiersema2012}. With a bright afterglow and a rapid follow-up, this
-spectrum is of pristine quality. The afterglow continuum is bright throughout
-all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
-\mnii, \mgii, \mgi, and \caii~- al located at $z = 0.971$, setting is as the
-redshift of the host.
+completed and the first burst that enters the statistical sample. This data is
+the basis for GCN 10042 \citep{GCN10042} and is
+published in \citet{Wiersema2012}. With a bright afterglow and a rapid
+follow-up, this spectrum is of pristine quality. The afterglow continuum is
+bright throughout all spectroscopic arms which allows the ready detection of
+\alii, \aliii, \feii, \mnii, \mgii, \mgi, and \caii~- al located at $z = 0.971$,
+setting is as the redshift of the host.
 
@@
 
-Obtained 4 days after the burst trigger, this data forms the basis for GCN \#
-10233\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10233.gcn3}} and is published
-in \citet{Vergani2011}. Due to the late follow-up and a nearby moon, the signal-to-noise of
-the afterglow continuum is low especially in the UVB arm, why no clear
-absorption lines are detected against the afterglow continuum, although see
-\citet{Vergani2011} which report a tentative detection of \mgii. Emission lines
-from the underlying host is clearly visible with lines from \oii, \hb, \oiii,
-and \ha~all at $z = 0.490$. This bursts is additionally associated with
-SN2009nz.
+Obtained 4 days after the burst trigger, this data forms the basis for GCN 10233
+\citep{GCN10233} and is published in \citet{Vergani2011}. Due to the late
+follow-up and a nearby moon, the signal-to-noise of the afterglow continuum is
+low especially in the UVB arm, why no clear absorption lines are detected
+against the afterglow continuum, although see \citet{Vergani2011} which report a
+tentative detection of \mgii. Emission lines from the underlying host is clearly
+visible with lines from \oii, \hb, \oiii, and \ha~all at $z = 0.490$. This
+bursts is additionally associated with SN2009nz.
 
@@
 in 10.8 ks. GRB likely located at high
-redshift\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10399.gcn3}}. The spectrum
+redshift \citep[GCN GCN10399;][]{GCN10399}. The spectrum
 has not otherwise been published previously.
@@
 The data presented here also formed the basis of GCN \#
-10441\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10441.gcn3}} and is published
+10441 \citep{GCN10441} and is published
 in \citet{Thone2013}. Observations started 12.5 hours after the \textit{Swift}
@@
 
-The data presented here also formed the basis of GCN \#
-10495\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10495.gcn3}}. The spectrum
-has not otherwise been published previously. Observations started 44 minutes
+The data presented here also formed the basis of GCN 10495 \citep{GCN10495}. The spectrum
+has not otherwise been published. Observations started 44 minutes
 after the \textit{Swift} trigger and has a total exposure time of 2.4 ks.
@@
 
-The data presented here also formed the basis of GCN \#
-10512\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10512.gcn3}}, GCN \#
-10513\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10513.gcn3}}, GCN \#
-10543\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10543.gcn3}} and is published
-in \citet{Bufano2012} and \citet{Starling2011}. This GRB is very close by and
-has an associated SN, SN2010bh, and has therefore undergone intense follow-up. The data
-presented here consists of a subset of the entire VLT/X-shooter campaign,
+The data presented here also formed the basis of GCN 10512 \citep{GCN10512}, GCN
+\# 10513 \citep{GCN10513}, GCN 10543 \citep{GCN10543} and is published in
+\citet{Bufano2012} and \citet{Starling2011}. This GRB is very close by and has
+an associated SN, SN2010bh, and has therefore undergone intense follow-up. The
+data presented here consists of a subset of the entire VLT/X-shooter campaign,
 covering the four first observing days while the afterglow still contributes
@@
 after the burst, before the SN was discovered, and targeted the star-forming
-'A'-region\citep{Starling2011}, not the GRB. A very rich spectrum containing a
+'A'-region \citep{Starling2011}, not the GRB. A very rich spectrum containing a
 multitude of emission lines puts the host at $z = 0.059$. For three consequtive
-nights, 58, 79 and 101 hours after the \textit{Swift} trigger, the afterglow
-was observed as it transitioned into the spectrum of a high-velocity Ic-BL SN.
-The observations taken 79 and 101 hours after the burst are taken under
-programme 084.D-0265(A) (PI: Benetti), but with an identical setup to the first
-two observations.
+nights, 58, 79 and 101 hours after the \textit{Swift} trigger, the afterglow was
+observed as it transitioned into the spectrum of a high-velocity Ic-BL SN. The
+observations taken 79 and 101 hours after the burst are taken under programme
+084.D-0265(A) (PI: Benetti), but with an identical setup to the first two
+observations.
 
@@
 
-The data presented here also formed the basis of GCN \#
-10620\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10620.gcn3}} and GCN \#
-10631\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10631.gcn3}} and is published
+The data presented here also formed the basis of GCN 10620 \citep{GCN10620} and GCN 10631 \citep{GCN10631} and is published
 in \citet{DeUgartePostigo2011}. The burst have been followed up in three epochs
@@
 
-The data presented here also formed the basis of GCN \#
-14291\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14291.gcn3}}. The spectrum has
-not otherwise been published previously. Observations carried out, long after
-the burst has faded.  Emission lines from the host are detected at $z=2.465$.
+The data presented here also formed the basis of GCN 
+14291 \citep{GCN14291}. The spectrum has
+not otherwise been published. Observations carried out, long after the burst
+afterglow had faded.  Emission lines from the host are detected at $z=2.465$.
 
@@
 
-The data presented here also formed the basis of GCN \#
-10684\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10684.gcn3}} and is used in
+The data presented here also formed the basis of GCN 
+10684 \citep{GCN10684} and is used in
 Skuladottir (2010), but not published elsewhere. Observations started 4 hours
@@
 
-The data presented here also formed the basis of GCN \#
-14264\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14264.gcn3}}, but not
-published elsewhere. Host observation of a dark burst\citep{DElia2011} taken
+The data presented here also formed the basis of GCN 
+14264 \citep{GCN14264}, but not
+published elsewhere. Host observation of a dark burst \citep{DElia2011} taken
 long after the afterglow has faded. Emission lines from the host belonging to
@@
 
-The data presented here also formed the basis of GCN \#
-10876\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10876.gcn3}}, but not
+The data presented here also formed the basis of GCN 
+10876 \citep{GCN10876}, but not
 published elsewhere. Beginning 7.1 hours after the GRB, 2.4 ks observations
@@
 reveals a weak continuum present in all arms, but an absence of emission lines.
-This could indicate that the host primarily contains a older stellar
+This could indicate that the host primarily contains an older stellar
 population. The redshift, $z=0.452$, is taken from \citet{Fong2013}.
@@
 
-The data presented here also formed the basis of GCN \#
-10971\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/10971.gcn3}}. The spectrum has
-not otherwise been published previously. The observations were carried out in
-RRM starting 11 min after the GRB trigger. See section \ref{RRM}, for a
-description of the RRM scheme. Absorption lines from several ionic species are
-detected in the afterglow continuum at a common redshift of $z = 1.288$. This is
-not a part of the statistical sample.
+The data presented here also formed the basis of GCN 10971 \citep{GCN10971}. The
+spectrum has not otherwise been published previously. The observations were
+carried out in RRM starting 11 min after the GRB trigger. See section \ref{RRM},
+for a description of the RRM scheme. Absorption lines from several ionic species
+are detected in the afterglow continuum at a common redshift of $z = 1.288$.
+This is not a part of the statistical sample.
 
@@
 
-The data presented here also formed the basis of GCN \#
-11317\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11317.gcn3}}. The spectrum has
-not otherwise been published previously. Starting 22 hours after the burst
-trigger, 7.2 ks of observations reveals a faint afterglow continuum with \lya-
-and \mgii-absorption at $z=2.106$. Due to a malfunctioning ADC, the sensitivity
-of X-shooter is depressed with respect to normal operations, resulting in a
-lower throughout. Additionally, the position of the trace on the slit moves due
-to atmospheric differential refraction. The presence of the DLA is confirmed in
-the 2D image and despite the observational challenges that affects this burst,
-we measure \nh~$=21.2 \pm 0.15$.
+The data presented here also formed the basis of GCN 11317 \citep{GCN11317}. The
+spectrum has not otherwise been published previously. Starting 22 hours after
+the burst trigger, 7.2 ks of observations reveals a faint afterglow continuum
+with \lya- and \mgii-absorption at $z=2.106$. Due to a malfunctioning ADC, the
+sensitivity of X-shooter is depressed with respect to normal operations,
+resulting in a lower throughout. Additionally, the position of the trace on the
+slit moves due to atmospheric differential refraction. The presence of the DLA
+is confirmed in the 2D image and despite the observational challenges that
+affects this burst, we measure \nh~$=21.2 \pm 0.15$.
 
@@
 to \mgii, shows several components, separated by as much as 500km/s, pointing to
-a likely merger scenario in the host.
+a likely merger scenario in the host or starburst driven outflows.
 
@@
 
-The data presented here also formed the basis of GCN \#
-11123\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11123.gcn3}}. The spectrum has
-not otherwise been published previously. This short GRB was observed 28.4 hours
-after the GRB trigger. 4 x 1200 s of exposure reveals two distinct sets of
-emission lines, spatially offset $\lesssim 1 \arcsec $, very close in redshift
-space, $z=0.8034$ and $z=0.8049$, indicating either an interacting host or some
-complex velocity structure of the host. Faint underlying continua are present
-under both sets of lines.
+The data presented here also formed the basis of GCN 11123 \citep{GCN11123}. The
+spectrum has not otherwise been published previously. This short GRB was
+observed 28.4 hours after the GRB trigger. 4 x 1200 s of exposure reveals two
+distinct sets of emission lines, spatially offset $\lesssim 1 \arcsec $, very
+close in redshift space, $z=0.8034$ and $z=0.8049$, indicating either an
+interacting host or some complex velocity structure of the host. Faint
+underlying continua are present under both sets of emission lines.
 
@@
 The data presented here has been published in \citet{Hartoog2013}. Because of
-the unusual lingering brightness of this GRB, 2.4s of observations taken 65.98
+the unusual lingering brightness of this GRB, 2.4 ks of observations taken 65.98
 hours after the GRB trigger still reveals an afterglow continuum visible across
-the entire spectral coverage of X-shooter. Absorption lines from a wide range
-ion put the redshift at $z=1.408$, with intervening absorption systems at $z =
+the entire spectral coverage of X-shooter. Absorption lines from a wide range of
+ions sets the redshift at $z=1.408$, with intervening absorption systems at $z =
 1.3147$ and $z = 1.3179$.
@@
 This data has not been published before. Starting 3.7 hours after the GRB
-trigger, 7.2 ks of exposure time reveals a very faint continuum in the visual
-and near-infrared, only visible when heavily binning the images. No redshift
+trigger, 7.2 ks of exposure time reveal a very faint continuum in the visual and
+near-infrared, only visible when heavily binning the images. No redshift
 estimate is available from these observations.  Late-time Gemini-North
-observations reveal emission lines from the host at
-$z=0.718$\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11518.gcn3}}.
+observations reveal emission lines from the host at $z=0.718$ \citep{GCN11518}.
 
@@
 
-The data presented here also formed the basis of GCN \#
-11579\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11579.gcn3}} and is published
-in \citet{Sparre2011}.	The first observation, taken 11.6 hours after the burst
-trigger and lasting 4.8ks, reveals absorption from \mgii~and \mgi~in the host
-located at $z = 0.552$ on a featureless continuum visible across the entire
-coverage of X-shooter.  Subsequent observations taken 16 and 37 days after the
-trigger shows the fading spectral signature of a SN, SN2010ma.
+The data presented here also formed the basis of GCN 11579 \citep{GCN11579} and
+is published in \citet{Sparre2011}.	The first observation, taken 11.6 hours
+after the burst trigger and lasting 4.8ks, reveals absorption from \mgii~and
+\mgi~in the host located at $z = 0.552$ on a featureless continuum visible
+across the entire coverage of X-shooter.  Subsequent observations taken 16 and
+37 days after the trigger shows the fading spectral signature of a SN, SN2010ma.
 
@@
 
-These observations forms the basis of GCN \#
-11607\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/11607.gcn3}}, but has not been
-published before. Spectroscopic integration started 6.55 hours after the
-\swift~trigger and lasted for a total of 7.2ks The afterglow continuum is
+These observations form the basis of GCN 11607 \citep{GCN11607}, but have not
+been published before. Spectroscopic integration started 6.55 hours after the
+\swift~trigger and lasted for a total of 7.2ks. The afterglow continuum is
 detected across the entire spectral coverage at moderate signal-to-noise.
@@
 These observations have not been published before. Starting 12.36 hours after
-the BAT trigger, 4.8 ks spectroscopic integration yield a very faint trace down
+the BAT trigger, 4.8 ks spectroscopic integration yields a very faint trace down
 to $\sim$4300\AA, only visible after binning heavily. This could indicate a
@@
 
-These observation, starting 12.3 hours after the trigger, have been published in
-\citet{Sanchez-Ramirez2017} and additionally formed the basis of GCN \#
-12164\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12164.gcn3}}. Only a single
-exposure of 600 s was obtained, before strong winds interrupted the
-observations. A red continuum is detected across all arms and a multitude of
-absorption lines are superposed on the afterglow continuum. We identify lines
-belonging to \alii, \aliii, \znii, \crii, \feii, \mgii, \mgi, \caii, and \caii,
-all at  $z=0.823$, marking it as the redshift of the GRB.
+These observations, starting 12.3 hours after the trigger, have been published
+in \citet{Sanchez-Ramirez2017} and additionally formed the basis of GCN 12164
+\citep{GCN12164}. Only a single exposure of 600 s was obtained, before strong
+winds interrupted the observations. A red continuum is detected across all arms
+and a multitude of absorption lines are superposed on the afterglow continuum.
+We identify lines belonging to \alii, \aliii, \znii, \crii, \feii, \mgii, \mgi,
+\caii, and \caii, all at  $z=0.823$, marking it as the redshift of the GRB.
 
@@
 sample, but nonetheless followed up due to the extremely high peak energy
-\citep{Axelsson2012}. Starting 28.7 after the burst trigger, 2.4ks spectroscopic
-observation reveals after heavy binning, a wide, faint trace down to $\sim$ 5800
-\AA, offset by 2.5\arcsec~relative to the centering of the slit. No good
-redshift measurement can be inferred from this. We have adopted the redshift
-from GCN \# 12193\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12193.gcn3}}
+\citep{Axelsson2012}. Starting 28.7 hours after the burst trigger, 2.4ks
+spectroscopic observation reveals after heavy binning, a wide, faint trace down
+to $\sim$ 5800 \AA, offset by 2.5\arcsec~relative to the centering of the slit.
+No good redshift measurement can be inferred from this. We have adopted the
+redshift from GCN  12193 \citep{GCN12193}
 
@@
 
-This spectrum has already formed the basis of GCN \#
-12258\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12258.gcn3}}, but is not
-published otherwise. Starting 3 hours after the \swift~trigger, a rich spectrum
-is obtained in 2.4ks spectroscopic integration. The GRB afterglow continuum is
-visible across all three spectrscopic arms of VLT/X-shooter with additionally,
-emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$. At the same
-redshift, we identify absorption lines superposed on the afterglow continuum
-from \mgii~and \feii.
+This spectrum has already formed the basis of GCN 12258 \citep{GCN12258}, but is
+not published otherwise. Starting 3 hours after the \swift~trigger, a rich
+spectrum is obtained in 2.4ks spectroscopic integration. The GRB afterglow
+continuum is visible across all three spectrscopic arms of VLT/X-shooter with
+additionally, emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$.
+At the same redshift, we identify absorption lines superposed on the afterglow
+continuum from \mgii~and \feii.
 
@@
 Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8ks
-reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 5000 \AA. The
-simultaneous detection of absorption features identified as \lya, \SIii, \civ,
-\alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet, securely sets
-$z = 3.36$ as the redshift of the GRB. This data forms the basis of GCN \#
-12284\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12284.gcn3}}, but is not
-published elsewhere.
+reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 5000
+\AA. The simultaneous detection of absorption features identified as \lya,
+\SIii, \civ, \alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet,
+securely sets $z = 3.36$ as the redshift of the GRB. This data forms the basis
+of GCN 12284 \citep{GCN12284}, but is not published elsewhere.
 
@@
 
-This data formed the basis of GCN \#
-12431\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12431.gcn3}} and is
-additionally published in \citet{Sparre2014}. Observations of this GRB afterglow
-was initiated 7.71 hours after the BAT trigger and had a duration of 8.4ks. A
-second observational epoch started 20.1 hours and lasted for 6.6ks. The GRB
-afterglow continuum is well detected down to $\sim$7600 \AA, with several strong
-absorption features imprinted. All at a common $z = 4.990$ \lya~is clearly
+This data formed the basis of GCN 12431 \citep{GCN12431} and is additionally
+published in \citet{Sparre2014}. Observations of this GRB afterglow was
+initiated 7.71 hours after the BAT trigger and had a duration of 8.4ks. A second
+observational epoch started 20.1 hours and lasted for 6.6ks. The GRB afterglow
+continuum is well detected down to $\sim$7600 \AA, with several strong
+absorption features imprinted, all at a common $z = 4.990$. \lya~is clearly
 detected and we additionally detect lines identified as \SIii, \feii, \civ,
@@
 
-GCN \# 12542\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12542.gcn3}} is based
-on this spectrum, but it is not published elsewhere. Spectroscopic integration
-started 5.26 hours after the \swift~trigger and consists of $4 \times 1200$ s
-integration in the UVB and VIS and $16 \times 300$ s in NIR, the observations
-ending in twilight. The GRB afterglow continuum is well detected across the arms
-with absorption lines from \lya, \civ, \feii, and \mgii, all at a consistent
-redshift of $z = 2.893$. Additionally an intervening \mgii~system is detected at
-$z = 1.998$. From the \lya~absorption trough, we additionally infer $\log
-(N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.0 \pm 0.2$.
+GCN  12542 \citep{GCN12542} is based on this spectrum, but it is not published
+elsewhere. Spectroscopic integration started 5.26 hours after the \swift~trigger
+and consists of $4 \times 1200$ s integration in the UVB and VIS and $16 \times
+300$ s in NIR, the observations ending in twilight. The GRB afterglow continuum
+is well detected across the arms with absorption lines from \lya, \civ, \feii,
+and \mgii, all at a consistent redshift of $z = 2.893$. Additionally an
+intervening \mgii~system is detected at $z = 1.998$. From the \lya~absorption
+trough, we additionally infer $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.0
+\pm 0.2$.
 
@@
 
-This data formed the basis of GCN \#
-14273\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14273.gcn3}}, but is not
-published elsewhere. Observed twice, the first time shortly after the GRB and
-the second long after the burst had faded, securely sets the redshift of the
-host at $z = 3.151$ based on the detection of emission lines identified as
-\oii~and \oiii.
+This data formed the basis of GCN 14273 \citep{GCN14273}, but is not published
+elsewhere. Observed twice, the first time shortly after the GRB and the second
+long after the burst had faded, securely sets the redshift of the host at $z =
+3.151$ based on the detection of emission lines identified as \oii~and \oiii.
 
@@
 detection of \oii. At this redshift, \hb~and \oii~are located in the gap between
-the VIS and NIR arm and \ha is located in the middle of the $JH$-bandgap and
-there are therefore not detected.
+the VIS and NIR arm and \ha~is located in the middle of the $JH$-bandgap and is
+therefore not detected.
 
@@
 
-These spectra has previously been used in \citet{Levan2013, Greiner2015,
-	Kruhler2015, Kann2017} and additionally formed the basis for GCN \#
-12648\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12648.gcn3}}. The first epoch
-of spectroscopic observations was initiated 17.7 hours after the BAT trigger and
-lasted for 4.8 seconds. A very bright afterglow continuum is detected across the
-entire spectral coverage of X-shooter, with several absorption features
-imprinted. The absorption features are identified as \feii, \mgii, \mgi, \cah,
-and \cak - all at a common redshift of $z = 0.677$. The second epoch, taken 20
-days later, still contains a faint continuum detected across all arms. The
-detection of nebular emission lines identified as \oii, \oii~and \ha~at the same
-redshift, securely marks it at the redshift of this ultra-long GRB with
-accompanying GRB-SN.
+These spectra have previously been used in \citet{Levan2013, Greiner2015,
+	Kruhler2015, Kann2017} and additionally formed the basis for GCN 12648
+\citep{GCN12648}. The first epoch of spectroscopic observations was initiated
+17.7 hours after the BAT trigger and lasted for 4.8 seconds. A very bright
+afterglow continuum is detected across the entire spectral coverage of
+X-shooter, with several absorption features imprinted. The absorption features
+are identified as \feii, \mgii, \mgi, \cah, and \cak - all at a common redshift
+of $z = 0.677$. The second epoch, taken 20 days later, still contains a faint
+continuum detected across all arms. The detection of nebular emission lines
+identified as \oii, \oii~and \ha~at the same redshift, securely marks it at the
+redshift of this ultra-long GRB with accompanying GRB-SN.
 
@@
 
-These data has formed the basis for GCN \#
-12677\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12677.gcn3}} and is also published in \cite{Kruhler2015}. Observations began 31 hours after the AGILE trigger
-and consisted of $4 \times 600$ s. A bright GRB afterglow continuum is detected
-across the entire spectral coverage of X-shooter with absorption and emission
-features visible. We identify absorption features due to \feii, \mgii, and \caii
-and emission lines from \oiii~and \ha, all at a common $z = 0.478$, which we
-suggest is the redshift of the GRB. Additionally detected in the GRB afterglow
-continuum is broad undulation, suggesting an accompanying SN.
+These data has formed the basis for GCN 12677 \citep{GCN12677} and is also
+published in \cite{Kruhler2015}. Observations began 31 hours after the AGILE
+trigger and consisted of $4 \times 600$ s. A bright GRB afterglow continuum is
+detected across the entire spectral coverage of X-shooter with absorption and
+emission features visible. We identify absorption features due to \feii, \mgii,
+and \caii~and emission lines from \oiii~and \ha, all at a common $z = 0.478$,
+which we suggest is the redshift of the GRB. Additionally detected in the GRB
+afterglow continuum is broad undulation, suggesting an accompanying SN.
 
@@
 
-This data formed the basis of GCN \#
-12770\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12770.gcn3}} and is also published in \cite{Kruhler2015}.
-Observations began 15.9 hours after the BAT trigger and consist of $4 \times 600$ s.
-The GRB afterglow continuum is clearly detected in all the spectroscopic arms and superposed on the continuum are absorption features identified as due to \feii, \mnii, \mgii, \mgi, \cah, and \cak, all at $z = 0.716$.
-Supporting this redshift measurement as the redshift of the GRB is the detection of nebular emission line from \oiii.
+This data formed the basis of GCN 12770 \citep{GCN12770} and is also published
+in \cite{Kruhler2015}. Observations began 15.9 hours after the BAT trigger and
+consist of $4 \times 600$ s. The GRB afterglow continuum is clearly detected in
+all the spectroscopic arms and superposed on the continuum are absorption
+features identified as due to \feii, \mnii, \mgii, \mgi, \cah, and \cak, all at
+$z = 0.716$. Supporting this redshift measurement as the redshift of the GRB is
+the detection of nebular emission line from \oiii.
 
@@
 
-The data presented here also formed the basis of GCN \#
-14225\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14225.gcn3}}, but is not
-published otherwise. This late-time observation of the host of GRB~120118B
-consists of 3.6 ks exposures and contains emission lines belonging to \oii and
-\oiii at $z = 2.943$, suggested to be redshift of the host.
+The data presented here also formed the basis of GCN 14225 \citep{GCN14225}, but
+is not published otherwise. This late-time observation of the host of
+GRB~120118B consists of 3.6 ks exposures and contains emission lines belonging
+to \oii~and \oiii~at $z = 2.943$, suggested to be redshift of the host.
 
@@
 
-The data presented here has formed the basis of GCN
-\#12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, and has also
-been published in \citet{Kruhler2015}. Starting 19.8 hours after the GRB
-trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting at
-$\sim$ 700 nm~and extending all the way through 2500 nm. In the 2D-spectrum we
-detect a $\sim 2 \sigma$ emission line which, if interpreted as \ha, gives $z =
-1.10$, supporting the photometric redshift ($0.9 < z_\mathrm{phot} < 1.3$)
+The data presented here has formed the basis of GCN 12991 \citep{GCN12991}, and
+has also been published in \citet{Kruhler2015}. Starting 19.8 hours after the
+GRB trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting
+at $\sim$ 700 nm~and extending all the way through 2500 nm. In the 2D-spectrum
+we detect a $\sim 2 \sigma$ emission line which, if interpreted as \ha, gives $z
+= 1.10$, supporting the photometric redshift ($0.9 < z_\mathrm{phot} < 1.3$)
 derived by \citet{Kruhler2015}.
@@
 
-The data presented here has formed the basis of GCN \#
-12991\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/12991.gcn3}}, but is not
-published otherwise. Starting just before twilight, 3.65 hours after the burst,
-a faint afterglow continuum is detected at all wavelengths. Due to the
-faintness of the afterglow, no absorption features are discernible superposed
-on the continuum. Displaced from the afterglow continuum by 1\farc4, emission
-lines belonging to \hb, \oiii~and \ha~are detected at $z = 0.350$. The line
-belonging to \ha~shows some extended emission toward the afterglow continuum.
-The angular distance between the two sources correspond to a projected distance
-in the host plane of 6 kpc, posing a potential problem for the host redshift,
-unless the GRB ocurred in a merging system. The extended emission in \ha,
-supports this interpretation. This burst is not a part of the statistical sample.
+The data presented here has formed the basis of GCN 12991 \citep{GCN12991}, but
+is not published otherwise. Starting just before twilight, 3.65 hours after the
+burst, a faint afterglow continuum is detected at all wavelengths. Due to the
+faintness of the afterglow, no absorption features are discernible superposed on
+the continuum. Displaced from the afterglow continuum by 1\farc4, emission lines
+belonging to \hb, \oiii~and \ha~are detected at $z = 0.350$. The line belonging
+to \ha~shows some extended emission toward the afterglow continuum. The angular
+distance between the two sources correspond to a projected distance in the host
+plane of 6 kpc, posing a potential problem for the host redshift, unless the GRB
+ocurred in a merging system. The extended emission in \ha, supports this
+interpretation. This burst is not a part of the statistical sample.
 
@@
 
-The data presented here also formed the basis of GCN \#
-13134\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13134.gcn3}} and is published
-in \citet{DElia2014}. The observation consists of two visits, 2.13 hrs and
-29.98 hrs after the burst, with an afteglow continuum visible in all arms for
-both visits. We detect absorption features from Ly-limit, \lya, \cii/\cii$^*$,
-\SIii/\SIii$^*$, \ali, \feii ~and \mgii ~are detected at a consistent redshift,
+The data presented here also formed the basis of GCN 13134 \citep{GCN13134} and
+is published in \citet{DElia2014}. The observation consists of two visits, 2.13
+hrs and 29.98 hrs after the burst, with an afteglow continuum visible in all
+arms for both visits. We detect absorption features from Ly-limit, \lya,
+\cii/\cii$^*$, \SIii/\SIii$^*$, \ali, \feii ~and \mgii~at a consistent redshift,
 $z = 2.813$.
@@
 
-The data presented here has formed the basis of GCN \#
-13227\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13227.gcn3}}, but is not
-published otherwise. 9.6 ks integration, starting 15.7 hours after the
+The data presented here has formed the basis of GCN 13227 \citep{GCN13227}, but
+is not published otherwise. 9.6 ks integration, starting 15.7 hours after the
 \textit{Swift}-trigger reveals a low-intensity afterglow continuum on which
@@
 
-The data presented here also formed the basis of GCN \# 
-13257\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13257.gcn3}} and is published
-in \citet{Schulze2014}. A GRB-SN, this burst has been followed up multiple
-times. The data presented here only contain the first epoch in which the
-afterglow is still visible and before the rise of SN2012bz. Starting 16.5 hours
-after the burst, 4.8 ks integration time captures both the host and the burst
-in emission. A blue afterglow continuum is detected at all wavelengths covered
-by X-shooter, on which \mgii absorption at $z = 0.283$ is found. Offset by
-1\farc75, the host is clearly detected at a consistent redshift with a rich
+The data presented here also formed the basis of GCN 13257 \citep{GCN13257} and
+is published in \citet{Schulze2014}. A GRB-SN, this burst has been followed up
+multiple times. The data presented here only contains the first epoch in which
+the afterglow is still visible and before the rise of SN2012bz. Starting 16.5
+hours after the burst, 4.8 ks integration time captures both the host and the
+burst in emission. A blue afterglow continuum is detected at all wavelengths
+covered by X-shooter, on which \mgii~absorption at $z = 0.283$ is found. Offset
+by 1\farc75, the host is clearly detected at a consistent redshift with a rich
 emission line spectrum, the lines extending towards to burst.
@@
 
-The data presented here also formed the basis of GCN
-\#13460\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13460.gcn3}} and is not
-published elsewhere. 4.8 ks integration time, starting 10.5 hours after the BAT
-trigger, shows a bright afterglow continuum starting at $\sim$ 472 nm,
+The data presented here also formed the basis of GCN 13460 \citep{GCN13460}, but
+is not published elsewhere. 4.8 ks integration time, starting 10.5 hours after
+the BAT trigger, shows a bright afterglow continuum starting at $\sim$ 472 nm,
 signifying the onset of the Lyman alpha forest, for a GRB located at $z =
@@
 
-The data presented here also formed the basis of GCN
-\#13477\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13477.gcn3}}, but is not
-published elsewhere. Observations of this burst started 7.8 hours after the GRB
-trigger, lasting 4.8 ks. A continuum is visible across the entire spectral
-coverage of X-shooter, with both emission lines from  \oii, \hb, \oiii~and \ha,
-as well as absorption from \mgii~detected at $z = 0.398$, securely setting it as
-the redshift of the GRB.
+The data presented here also formed the basis of GCN 13477 \citep{GCN13477}, but
+is not published elsewhere. Observations of this burst started 7.8 hours after
+the GRB trigger, lasting 4.8 ks. A continuum is visible across the entire
+spectral coverage of X-shooter, with both emission lines from  \oii, \hb,
+\oiii~and \ha, as well as absorption from \mgii~detected at $z = 0.398$,
+securely setting it as the redshift of the GRB.
 
@@
 
-The data presented here also formed the basis of GCN
-\#13494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13494.gcn3}}, but is not
-published elsewhere. Despite observations starting 62 hours after the
+The data presented here also formed the basis of GCN 13494 \citep{GCN13494}, but
+is not published elsewhere. Despite observations starting 62 hours after the
 \textit{Swift} trigger and lasting 3.6 ks, a bright afterglow is clearly seen,
@@
 
-The data presented here also formed the basis of GCN
-\#13507\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13507.gcn3}}, but is not
-published elsewhere. On 4.8 ks integration time, starting 10 hours after the
-burst trigger, the simultaneous detection of absorption features belonging to
-\mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
+The data presented here also formed the basis of GCN 13507 \citep{GCN13507}, but
+is not published elsewhere. In 4.8 ks integration time, starting 10 hours after
+the burst trigger, the simultaneous detection of absorption features belonging
+to \mgii~and \feii~superposed on a blue continuum, and emission lines from \oii,
 \hg, \hb, \oiii~and \ha, all at $z = 0.959$, confidently sets it as the redshift
@@
 time, we detect a faint continuum visible from 450 nm and all the way through
-2100 nm, in contrast to what is found previously. The continuum from 4500 -
-600 nm is detected at very low significance. If the drop at 450 nm is the
-Lyman limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim
-3.9$. The absence of nebular lines if due to \oii~falling in a telluric
-absorption band and the rest being shifted out of the wavelength coverage.
+2100 nm, in contrast to what is found previously. The continuum from 4500 - 600
+nm is detected at very low significance. If the drop at 450 nm is the Lyman
+limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim 3.9$. The
+absence of nebular lines if due to \oii~falling in a telluric absorption band
+and the rest being shifted out of the wavelength coverage.
 
@@
 Not a part of the statistical sample, this burst also formed the basis of  GCN
-\#13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
-published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
-trigger and consist of 2.4 ks integration. A bright afterglow continuum is
-detected across the entire spectral coverage of X-shooter, with a multitude of
-absorption lines superposed. Absorption features from the host at $z = 2.358$
-include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
-\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally fine
-structure lines from \NIii and \feii are exited local to the GRB. Intervening
-systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.
+13649 \citep{GCN13649} and is published in \citet{Kruhler2013}. Observations
+started 1.69 hours after the BAT trigger and consist of 2.4 ks integration. A
+bright afterglow continuum is detected across the entire spectral coverage of
+X-shooter, with a multitude of absorption lines superposed. Absorption features
+from the host at $z = 2.358$ include a DLA as well as metal absorption lines
+from \nv, \sii, \SIii, \oi, \civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and
+\mgi. Additionally fine structure lines from \NIii~and \feii~are exited local to
+the GRB. Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z =
+2.00$.
 
@@
 
-The data presented here has formed the basis of GCN
-\#13730\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13730.gcn3}}, but is not
-published otherwise. A very rapid follow-up, starting only 1.7 hours after the
-BAT trigger, this 1.2 ks observation captures a very bright afterglow continuum,
-starting at 450 nm, signifying the onset of the Lyman limit for a system at $z =
-3.929$. Absorption from high-column density hydrogen leaves very prominent
-absorption features in the form of \lya, \lyb, and \lyg, visible in the Lyman
-alpha forest. Metal absorption lines arising from \feii, \NIii, \SIii, \sii,
-\alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
+The data presented here has formed the basis of GCN 13730 \citep{GCN13730}, but
+is not published otherwise. A very rapid follow-up, starting only 1.7 hours
+after the BAT trigger, this 1.2 ks observation captures a very bright afterglow
+continuum, starting at 450 nm, signifying the onset of the Lyman limit for a
+system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
+prominent absorption features in the form of \lya, \lyb, and \lyg, visible in
+the Lyman alpha forest. Metal absorption lines arising from \feii, \NIii, \SIii,
+\sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
 corresponding fine structure lines from (\feii$^*$, \SIii$^*$, \oi$^*$,
@@
 
-The data presented here also formed the basis of GCN
-\#13890\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13890.gcn3}} and is
-published in \citet{Friis2015}. Also rapid, starting 1.8 hours after the
+The data presented here also formed the basis of GCN 13890 \citep{GCN13890} and
+is published in \citet{Friis2015}. Also rapid, starting 1.8 hours after the
 \textit{Swift} trigger, a bright afterglow continuum is visible across all arms.
@@
 
-The data presented here has formed the basis of GCN
-\#13930\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13930.gcn3}}, but is not
-published otherwise. Starting 69.6 hours after the GRB trigger, we detect the
-afterglow continuum at high significance in all arms with 8.4 ks integration,
-testifying to the brightness of this burst. The concurrent identification of
-emission lines from \oiii~and absorption from \civ, \alii, \aliii, \mgi, \mgii,
-and \feii, tightly constrains the redshift of the burst to be $z = 1.773$.
+The data presented here has formed the basis of GCN 13930 \citep{GCN13930}, but
+is not published otherwise. Starting 69.6 hours after the GRB trigger, we detect
+the afterglow continuum at high significance in all arms with 8.4 ks
+integration, testifying to the brightness of this burst. The concurrent
+identification of emission lines from \oiii~and absorption from \civ, \alii,
+\aliii, \mgi, \mgii, and \feii, tightly constrains the redshift of the burst to
+be $z = 1.773$.
 
@@
 
-These data formed the basis for GCN
-\#14035\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14035.gcn3}} and is
-additionally published in \citet{Kruhler2015}. These observations started 12.9
-hours after the \swift~trigger and consists of 4.8ks spectroscopic integration
-under good atmospheric conditions. The GRB afterglow continuum is well detected
-at all arms. A broad absorption trough due to \lya is visible at $z = 3.385$,
-which along with the detection of absorption features identified as \SIiv, \civ,
+These data formed the basis for GCN 14035 \citep{GCN14035} and is additionally
+published in \citet{Kruhler2015}. These observations started 12.9 hours after
+the \swift~trigger and consists of 4.8ks spectroscopic integration under good
+atmospheric conditions. The GRB afterglow continuum is well detected at all
+arms. A broad absorption trough due to \lya is visible at $z = 3.385$, which
+along with the detection of absorption features identified as \SIiv, \civ,
 \alii, and \aliii, marks it as the redshift of the GRB. In the middle of the
@@
 
-These data formed the basis for GCN
-\#14120\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14120.gcn3}}, but is not
-published elsewhere. Taken under poor seeing conditions, a total of 4.8ks
-spectroscopic integration starting 2 hours after the \swift~trigger yields a low
-signal-to-noise GRB afterglow continuum is all arms. Binning the spectrum reveals broad
-absorption troughs, which we identify as \lyb~and \lya~at $z = 2.707$.
-Additionally, an intervening system at $z = 1.658$ is detected from absorption
-features of \mgii. From the absorption trough due to \lya, we infer $\log
-(N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.7 \pm 0.2$. Due to strong contamination
-in the slit, the background is slightly over subtracted, cause the background to
-go negative in the center of the \lya~trough.
+These data formed the basis for GCN 14120 \citep{GCN14120}, but is not published
+elsewhere. Taken under poor seeing conditions, a total of 4.8ks spectroscopic
+integration starting 2 hours after the \swift~trigger yields a low
+signal-to-noise GRB afterglow continuum is all arms. Binning the spectrum
+reveals broad absorption troughs, which we identify as \lyb~and \lya~at $z =
+2.707$. Additionally, an intervening system at $z = 1.658$ is detected from
+absorption features of \mgii. From the absorption trough due to \lya, we infer
+$\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.7 \pm 0.2$. Due to strong
+contamination in the slit, the background is slightly over subtracted, cause the
+background to go negative in the center of the \lya~trough.
 
@@
 
-These data formed the basis for GCN
-\#14286\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14286.gcn3}} and is
-additionally published in \citet{Kruhler2015}. This is a late-time observation,
-taken long after the GRB afterglow has faded. In 7.2ks spectroscopic
-integration, emission lines identified as \oii~and \oiii~are detected at a
-common $z = 2.539$, which we suggest is the redshift of the GRB.
+These data formed the basis for GCN 14286 \citep{GCN14286} and is additionally
+published in \citet{Kruhler2015}. This is a late-time observation, taken long
+after the GRB afterglow has faded. In 7.2ks spectroscopic integration, emission
+lines identified as \oii~and \oiii~are detected at a common $z = 2.539$, which
+we suggest is the redshift of the GRB.
 
@@
 
-The data presented here also formed the basis of GCN
-\#14365\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14365.gcn3}}. The spectrum
-has not otherwise been published previously. The observations consists of two
-600sec spectra taken 1.9hrs after the burst. We detect absorption features from
-a wide range of ions. We also detect intervening absorption at $z=1.255$ and
-$z=3.248$.
+The data presented here also formed the basis of GCN 14365 \citep{GCN14365}. The
+spectrum has not otherwise been published previously. The observations consists
+of two 600sec spectra taken 1.9hrs after the burst. We detect absorption
+features from a wide range of ions. We also detect intervening absorption at
+$z=1.255$ and $z=3.248$.
 
@@
 
-GCN \#14390\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14390.gcn3}} is based on
-this spectrum, but it is not published elsewhere. Starting only 4.57 hours after
-the \swift~trigger, 1.2ks observations contains a bright GRB afterglow
-continuum, visible across the entire spectral coverage of X-shooter. Superposed
-on the afterglow continuum are absorption features which we identify as \civ,
-\feii, and \mgii, caused by an absorber at $z = 1.217$, and additional
-absorption from \civ~ at $z = 1.222$. The two systems are offset by $\sim
-1500~\mathrm{km}~\mathrm{s}^{-1}$ and the proximity of the two absorption
-systems in velocity space, suggests a possible association of the two systems
-with peculiar velocity affecting the measured redshift. We adopt $z = 1.222$ as
-the redshift of the GRB. Note, that this value is slightly different from the
-one reported in GCN \# 14390.
+GCN 14390 \citep{GCN14390} is based on this spectrum, but it is not published
+elsewhere. Starting only 4.57 hours after the \swift~trigger, 1.2ks observations
+contains a bright GRB afterglow continuum, visible across the entire spectral
+coverage of X-shooter. Superposed on the afterglow continuum are absorption
+features which we identify as \civ, \feii, and \mgii, caused by an absorber at
+$z = 1.217$, and additional absorption from \civ~ at $z = 1.222$. The two
+systems are offset by $\sim 1500~\mathrm{km}~\mathrm{s}^{-1}$ and the proximity
+of the two absorption systems in velocity space, suggests a possible association
+of the two systems with peculiar velocity affecting the measured redshift. We
+adopt $z = 1.222$ as the redshift of the GRB. Note, that this value is slightly
+different from the one reported in GCN  14390.
 
@@
 This spectrum is also published in \citet{Xu2013b} and \citet{Kruhler2015} and
-additionally has formed the basis for GCN
-\#14491\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14491.gcn3}}. Starting 16.5
+additionally has formed the basis for GCN 14491 \citep{GCN14491}. Starting 16.5
 hours after the BAT trigger, these observations lasting $2 \times 600$ s
@@
 
-This spectrum formed the basis of GCN
-\#14493\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14493.gcn3}}, but is not
+This spectrum formed the basis of GCN 14493 \citep{GCN14493}, but is not
 published otherwise. A short, 2x600 s spectroscopic integration obtained before
 twilight, 20.6 hours after the BAT trigger, captures a faint afterglow continuum
-visible across the entire spectral coverage at low signal-to-noise. Due to the low signal-to-noise, the
-metal lines are weak, but the broad absorption though due to \lya~is detected.
-From this we measure the redshift to be $z = 2.780$ and provides a measure of
-the neutral hydrogen column density, \nh~$= 21.9 \pm 0.3$. The redshift is
-confirmed by the presence of \feii~at a consistent redshift.
+visible across the entire spectral coverage at low signal-to-noise. Due to the
+low signal-to-noise, the metal lines are weak, but the broad absorption though
+due to \lya~is detected. From this we measure the redshift to be $z = 2.780$ and
+provides a measure of the neutral hydrogen column density, \nh~$= 21.9 \pm 0.3$.
+The redshift is confirmed by the presence of \feii~at a consistent redshift.
 
@@
 This burst is the first sGRB observed with a potential associated
-kilonova\citep{Tanvir2013a, Berger2013a}.  GCN
-\#14757\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14757.gcn3}} was based on
-this spectrum, but it is not published elsewhere. Starting 8.2 hours after the
-\swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
+kilonova \citep{Tanvir2013a, Berger2013a}.  GCN 14757 \citep{GCN14757} was based
+on this spectrum, but it is not published elsewhere. Starting 8.2 hours after
+the \swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
 Continuum is clearly detected across all arms from both host and afterglow and
@@
 
-The data presented here also formed the basis of GCN
-\#14816\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14816.gcn3}} and is
-published in \citet{Hartoog2015}. The observations consists of three
+The data presented here also formed the basis of GCN 14816 \citep{GCN14816} and
+is published in \citet{Hartoog2015}. The observations consists of three
 $2\times600$ s visits starting 7.1 hrs after the burst at fairly high airmass.
@@
 
-The spectral features of this spectrum has previously been reported in GCN
-\#14882\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14882.gcn3}}, but is not
-published elsewhere. Starting only 1.1 hour after the \swift~trigger,
-$2\times600$ s spectroscopic integration captures a moderate signal-to-noise afterglow
-continuum across the total spectral coverage of X-shooter. At a consistent
-redshift of $z = 2.006$, absorption from the metal resonance lines \feii, \mnii,
-\mgii, \mgi~are identified. Additionally, \lya~is visible as a broad absorption
-trough, from we can infer \nh~= $22.2 \pm 0.2$, which is in the upper end of the
-hydrogen column density distribution. The blue part of the GRB continuum
-exhibits a downturn in the continuum level which could indicate the presence of
-a significant amount of dust along the line-of-sight.
+The spectral features of this spectrum has previously been reported in GCN 14882
+\citep{GCN14882}, but is not published elsewhere. Starting only 1.1 hour after
+the \swift~trigger, $2\times600$ s spectroscopic integration captures a moderate
+signal-to-noise afterglow continuum across the total spectral coverage of
+X-shooter. At a consistent redshift of $z = 2.006$, absorption from the metal
+resonance lines \feii, \mnii, \mgii, \mgi~are identified. Additionally, \lya~is
+visible as a broad absorption trough, from we can infer \nh~= $22.2 \pm 0.2$,
+which is in the upper end of the hydrogen column density distribution. The blue
+part of the GRB continuum exhibits a downturn in the continuum level which could
+indicate the presence of a significant amount of dust along the line-of-sight.
 
@@
 background, a faint afterglow trace is visible across all arms of X-shooter,
-down to ~ 4800 \AA, confirming the approximate redshift suggested in GCN
-\#14898\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14898.gcn3}}.
+down to ~ 4800 \AA, confirming the approximate redshift suggested in GCN 14898
+\citep{GCN14898}.
 
@@
 
-These data formed the basis for GCN
-\#14956\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/14956.gcn3}} and is
-additionally published in \citet{Kruhler2015}. Starting 5.5 hours after the GRB
-trigger, $2\times600$ s reveals a bright continuum visible across the entire
-spectral coverage of X-shooter. Superposed are absorption features which we
-identify as due to \feii, \mgii, \mgi, and \caii~- all at a consistent redshift
-of $z = 1.155$, which we take to be the redshift of the GRB.
+These data formed the basis for GCN 14956 \citep{GCN14956} and is additionally
+published in \citet{Kruhler2015}. Starting 5.5 hours after the GRB trigger,
+$2\times600$ s reveals a bright continuum visible across the entire spectral
+coverage of X-shooter. Superposed are absorption features which we identify as
+due to \feii, \mgii, \mgi, and \caii~- all at a consistent redshift of $z =
+1.155$, which we take to be the redshift of the GRB.
 
@@
 
-This spectrum has already been used in GCN
-\#15250\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15250.gcn3}} and is
+This spectrum has already been used in GCN 15250 \citep{GCN15250} and is
 additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
@@
 
-This data formed the basis for
-\#15330\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15330.gcn3}}, but is not
-published elsewhere. Starting $\sim$1.5 days after the Fermi-GBM trigger, 4.5 ks
-spectroscopic integration captures a modest signal-to-noise GRB afterglow continuum all the
-way down to $\sim$3200 \AA. Imprinted on the continuum are absorption features,
-which we identify as due to \lya, \feii,  \mgii, \mgi~at the same redshift,
-which we measure to be $z = 1.874$. From the broad absorption trough due to
-\lya, we infer \nh~$= 22.0 \pm 0.3$. This spectrum is taken under ESO programme
-ID: 092.D-0056(A) (PI: Rau).
-
-\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}	 \label{131030}
+This data formed the basis for 15330 \citep{GCN15330}, but is not published
+elsewhere. Starting $\sim$1.5 days after the Fermi-GBM trigger, 4.5 ks
+spectroscopic integration captures a modest signal-to-noise GRB afterglow
+continuum all the way down to $\sim$3200 \AA. Imprinted on the continuum are
+absorption features, which we identify as due to \lya, \feii,  \mgii, \mgi~at
+the same redshift, which we measure to be $z = 1.874$. From the broad absorption
+trough due to \lya, we infer \nh~$= 22.0 \pm 0.3$. This spectrum is taken under
+ESO programme ID: 092.D-0056(A) (PI: Rau).
+
+\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}	
+\label{131030}
 
@@
 
-This spectrum has already been used to form the basis for GCN
-\#15451\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15451.gcn3}} and is
-additionally published in \citet{Kruhler2015}. Starting 5.8 hours after the BAT
-trigger, $4\times600$ s exposure captures a modest signal-to-noise continuum across all
-arms. Imprinted on the continuum are absorption features identifed as due to
-\feii~and \mgii~as well as emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
-and \nii. All the lines are measured at a consistent redshift of $z = 0.599$,
-which we take as the redshift of the GRB.
+This spectrum has already been used to form the basis for GCN 15451
+\citep{GCN15451} and is additionally published in \citet{Kruhler2015}. Starting
+5.8 hours after the BAT trigger, $4\times600$ s exposure captures a modest
+signal-to-noise continuum across all arms. Imprinted on the continuum are
+absorption features identifed as due to \feii~and \mgii~as well as emission
+lines from \oii, \hd, \hg, \hb, \oiii, \ha, and \nii. All the lines are measured
+at a consistent redshift of $z = 0.599$, which we take as the redshift of the
+GRB.
 
@@
 
-This spectrum has already been used in shaping GCN
-\#15450\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15450.gcn3}} and is
+This spectrum has already been used in shaping GCN 15450 \citep{GCN15450} and is
 additionally published in \citet{Kruhler2015}. Starting only 1.3 hours after the
-\swift~trigger, a total of 4.8 ks spectroscopic integration contains a low signal-to-noise
-GRB afterglow continuum across the entire spectral coverage of X-shooter. There
-are deviations from the continuum at both emission and absorption. We identify
-lines from \hb, \oiii, and \ha~in emission and \feii, and \mgii~in absorption.
-All lines are at a consistent $z = 1.686$, which is probably the redshift of the
-GRB. Absorption lines at shorter wavelengths are also detected, but at low
-significance due to an apparent downturn in the continuum caused by the presence
-of dust local to the burst.
+\swift~trigger, a total of 4.8 ks spectroscopic integration contains a low
+signal-to-noise GRB afterglow continuum across the entire spectral coverage of
+X-shooter. There are deviations from the continuum at both emission and
+absorption. We identify lines from \hb, \oiii, and \ha~in emission and \feii,
+and \mgii~in absorption. All lines are at a consistent $z = 1.686$, which is
+probably the redshift of the GRB. Absorption lines at shorter wavelengths are
+also detected, but at low significance due to an apparent downturn in the
+continuum caused by the presence of dust local to the burst.
 
@@
 
-This spectrum has previously been used in GCN
-\#15494\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15494.gcn3}}, but is not
+This spectrum has previously been used in GCN 15494 \citep{GCN15494}, but is not
 published in a refereed paper. Starting only 68 minutes after the BAT trigger,
 4.8 ks spectroscopic integration secures afterglow continuum for this bursts,
-which is measured to be at $z = 4.042$. A moderate signal-to-noise GRB afterglow continuum
-is detected down to $\sim$ 6100 \AA, signifying the onset of the \lya~forest,
-with part of the forest also visible. Metal absorption lines from \SIii~and
-\SIiv~are detected at a consistent redshift.
+which is measured to be at $z = 4.042$. A moderate signal-to-noise GRB afterglow
+continuum is detected down to $\sim$ 6100 \AA, signifying the onset of the
+\lya~forest, with part of the forest also visible. Metal absorption lines from
+\SIii~and \SIiv~are detected at a consistent redshift.
 
@@
 This spectrum has previously been published in \citet{Kruhler2015} and
-addtionally forms the basis for GCN
-\#15645\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15645.gcn3}}. This spectrum,
+addtionally forms the basis for GCN 15645 \citep{GCN15645}. This spectrum,
 observed the following year (20.2 hours after the \swift~trigger), consists of
-$4\times600$ s exposures. A high signal-to-noise GRB afterglow continuum is detected all the
-way though the X-shooter arms. We identify absorption features imprinted in the
-continuum as caused by \feii, \mgii, and \cahk at a consistent $z = 0.642$. By
-subtracting off the bright afterglow continuum, we readily detect emission lines
-arising from \oii, \hg, \hb, \oiii, and \ha~in the GRB host galaxy.
+$4\times600$ s exposures. A high signal-to-noise GRB afterglow continuum is
+detected all the way though the X-shooter arms. We identify absorption features
+imprinted in the continuum as caused by \feii, \mgii, and \cahk at a consistent
+$z = 0.642$. By subtracting off the bright afterglow continuum, we readily
+detect emission lines arising from \oii, \hg, \hb, \oiii, and \ha~in the GRB
+host galaxy.
 
@@
 
-The data presented here also formed the basis of GCN
-\#15831\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15831.gcn3}} and is
-additionally published in \citet{Kruhler2015}. Starting 5.8 after the
+The data presented here also formed the basis of GCN 15831 \citep{GCN15831} and
+is additionally published in \citet{Kruhler2015}. Starting 5.8 after the
 \swift~trigger alert, $2\times600$ s spectroscopic integration contains a high
-signal-to-noise GRB afterglow continuum across the entirety of X-shooter. Imprinted on the
-afterglow continuum are absorption features, which we identify as metal
-resonance lines from \civ, \alii, \aliii, \feii, \mgii, and \mgi. These lines
-are likely formed by metals in the GRB host, which we measure to be at $z =
-1.208$.
+signal-to-noise GRB afterglow continuum across the entirety of X-shooter.
+Imprinted on the afterglow continuum are absorption features, which we identify
+as metal resonance lines from \civ, \alii, \aliii, \feii, \mgii, and \mgi. These
+lines are likely formed by metals in the GRB host, which we measure to be at $z
+= 1.208$.
 
@@
 
-These data formed the basis for GCN
-\#15900\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15900.gcn3}} and is
-additionally published in \citet{Kruhler2015}. Spectroscopic follow-up began 9
-hours after the BAT trigger and lasted for $12\times600$ s. A low signal-to-noise, spatially
+These data formed the basis for GCN 15900 \citep{GCN15900} and is additionally
+published in \citet{Kruhler2015}. Spectroscopic follow-up began 9 hours after
+the BAT trigger and lasted for $12\times600$ s. A low signal-to-noise, spatially
 extended continuum is visible across the entire spectral coverage of X-shooter.
-The GRB afterglow continuum is visible at moderate signal-to-noise on top of the underlying
-host continuum. The host exhibits usual nebular emission from \oii, \hb, \oiii,
-\ha, \nii, and \sii which puts it at $z = 1.416$. Supporting this as the
-redshift of the GRB are absorption features from the \mgii-doublet in the GRB
-afterglow continuum.
-
-\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}	 \label{140311}
+The GRB afterglow continuum is visible at moderate signal-to-noise on top of the
+underlying host continuum. The host exhibits usual nebular emission from \oii,
+\hb, \oiii, \ha, \nii, and \sii which puts it at $z = 1.416$. Supporting this as
+the redshift of the GRB are absorption features from the \mgii-doublet in the
+GRB afterglow continuum.
+
+\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}	
+\label{140311}
 
@@
 A separate reduction of this burst has been published in \citet{Kruhler2015},
-and additionally the spectrum has also been used in GCN
-\#16194\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16194.gcn3}}. Observations
-for this burst began in twilight, 2.5 hours after the BAT trigger and lasted for
-$2\times600$ s. The spectrum contains a moderate signal-to-noise GRB afterglow continuum all
-the way through the spectroscopic arms of X-shooter. We identify absorption
-features in the afterglow continuum from \SIii, \civ, \alii, \feii, and \mgii,
-and emission lines from \oii~and \oiii~- all at $z = 1.601$ which is likely the
-redshift of the GRB.
+and additionally the spectrum has also been used in GCN 16194 \citep{GCN16194}.
+Observations for this burst began in twilight, 2.5 hours after the BAT trigger
+and lasted for $2\times600$ s. The spectrum contains a moderate signal-to-noise
+GRB afterglow continuum all the way through the spectroscopic arms of X-shooter.
+We identify absorption features in the afterglow continuum from \SIii, \civ,
+\alii, \feii, and \mgii, and emission lines from \oii~and \oiii~- all at $z =
+1.601$ which is likely the redshift of the GRB.
 
@@
 
-The data presented here has formed the basis of GCN
-\#16217\footnote{\url{https://gcn.gsfc.nasa.gov/gcn3/16217.gcn3}} and is
-published in \citet{Fynbo2014}, \citet{Kruhler2015}, and \citet{Heintz2017a}.
+The data presented here has formed the basis of GCN 16217 \citep{GCN16217} and
+is published in \citet{Fynbo2014}, \citet{Kruhler2015}, and \citet{Heintz2017a}.
 The observations consists of $4\times 600$ s at 8.8 and 33 hours after the
@@
 
-This spectrum forms the basis for GCN
-\#16401\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16401.gcn3}}, but is not
+This spectrum forms the basis for GCN 16401 \citep{GCN16401}, but is not
 published elsewhere. $4\times600$ s spectroscopic integration, starting 3.8
 hours after the BAT trigger catches this GRB afterglow, which turns out to be at
-very high redshift. The continuum is detected at moderate signal-to-noise in both the VIS
-and NIR arms of X-shooter, heavily affected by absorption features. We identify
-the the lines belonging to \lya, \SIii, \cii, \cii$^*$, \alii, \aliii, \feii,
-and \mgii~at a consistent redshift of $z = 4.233$. From the shape of the
-\lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
+very high redshift. The continuum is detected at moderate signal-to-noise in
+both the VIS and NIR arms of X-shooter, heavily affected by absorption features.
+We identify the the lines belonging to \lya, \SIii, \cii, \cii$^*$, \alii,
+\aliii, \feii, and \mgii~at a consistent redshift of $z = 4.233$. From the shape
+of the \lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
 
@@
 The characteristics of this short (T90 (15-350 keV ) = is $0.13 \pm 0.04$ s
-\citep{Lien2016}) GRB spectrum has been published in GCN
-\#16437\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16437.gcn3}}, but does not
-appear in the refereed literature. Spectroscopic observations began with
-X-shooter only 34 minutes after the BAT trigger and lasted for $2\times600$ s.
-In the spectrum, we detect continuum across all three arms of X-shooter. In the
-UVB arm, the continuum is only visible after heavily binning the spectrum in the
-dispersion direction. It is unclear how much of the continuum is from the host
-galaxy and how much is from the potential GRB afterglow. Superposed on the
+\citep{Lien2016}) GRB spectrum has been published in GCN 16437 \citep{GCN16437},
+but does not appear in the refereed literature. Spectroscopic observations began
+with X-shooter only 34 minutes after the BAT trigger and lasted for $2\times600$
+s. In the spectrum, we detect continuum across all three arms of X-shooter. In
+the UVB arm, the continuum is only visible after heavily binning the spectrum in
+the dispersion direction. It is unclear how much of the continuum is from the
+host galaxy and how much is from the potential GRB afterglow. Superposed on the
 continuum, emission lines from \oii, \hb, \oiii, \ha~are all detected at $z =
@@
 
-This spectrum has already been used to inform GCN
-\#16983\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16983.gcn3}}, but is not
-published elsewhere. Starting 15.4 hours after the \swift~trigger and lasting
-for a total of 2.4 ks, these observation nicely captures a bright GRB afterglow.
-Very rich in absorption, the continuum is detected across the entire spectral
-coverage of X-shooter, except the blue half of the UVB arm, where the
+This spectrum has already been used to inform GCN 16983 \citep{GCN16983}, but is
+not published elsewhere. Starting 15.4 hours after the \swift~trigger and
+lasting for a total of 2.4 ks, these observation nicely captures a bright GRB
+afterglow. Very rich in absorption, the continuum is detected across the entire
+spectral coverage of X-shooter, except the blue half of the UVB arm, where the
 \lya-forest absorbs the continuum. The redshift of the host of the GRB is
@@
 
-The gross content of the spectrum has been issued in GCN
-\#17040\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17040.gcn3}}, but it is not
-used otherwise. Starting only 1.9 hours after the BAT trigger, these
-$4\times600$ s spectra contains a high signal-to-noise GRB afterglow. The afterglow
-continuum is readily detected all the way down to 3700 \AA, although the bluest
-part is affected by the \lya-forest. A broad absorption trough from neutral
-hydrogen is clearly visible, with additional, narrower absorption features
-identified as due to \SIii, \SIii$^*$, \cii, \cii$^*$, \SIiv, \civ, \feii,
-\feii$^*$, \oi$^*$, and \nii$^*$. These are all detected at a consistent $z =
-2.993$, which due to the detection of the locally exited fine-structure lines
-securely sets this as the redshift of the GRB. Additionally, two
+The gross content of the spectrum has been issued in GCN 17040 \citep{GCN17040},
+but it is not used otherwise. Starting only 1.9 hours after the BAT trigger,
+these $4\times600$ s spectra contains a high signal-to-noise GRB afterglow. The
+afterglow continuum is readily detected all the way down to 3700 \AA, although
+the bluest part is affected by the \lya-forest. A broad absorption trough from
+neutral hydrogen is clearly visible, with additional, narrower absorption
+features identified as due to \SIii, \SIii$^*$, \cii, \cii$^*$, \SIiv, \civ,
+\feii, \feii$^*$, \oi$^*$, and \nii$^*$. These are all detected at a consistent
+$z = 2.993$, which due to the detection of the locally exited fine-structure
+lines securely sets this as the redshift of the GRB. Additionally, two
 \mgii~absorption systems are detected at $z = 1.67$ and $z=2.5$. From the broad
@@
 
-This spectrum has previously been used to form the basis for GCN
-\#17420\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17420.gcn3}}, but it is not
-published in the refereed literature. Due to the short visibility of this burst,
-it does not enter into the statistical sample. Beginning in twilight,
-$4\times600$ s integration time contains a weak/moderate signal-to-noise GRB afterglow
-continuum throughout the X-shooter spectral coverage, down to $\sim$ 3750 \AA.
-In the continuum are absorption features, which we identify as due to metal
-lines from \znii, \feii, and \mgii. Additionally, \lya-absorption is seen at the
-end of the trace. The spectral position of the lines means that this GRB is at
-$z = 2.087$. From the \lya~trough, we infer \nh~$=21.7 \pm 0.4$. The afterglow
-continuum appears depressed in the blue end of the spectrum, suggestion dust
-extinction in the host.
+This spectrum has previously been used to form the basis for GCN 17420
+\citep{GCN17420}, but it is not published in the refereed literature. Due to the
+short visibility of this burst, it does not enter into the statistical sample.
+Beginning in twilight, $4\times600$ s integration time contains a weak/moderate
+signal-to-noise GRB afterglow continuum throughout the X-shooter spectral
+coverage, down to $\sim$ 3750 \AA. In the continuum are absorption features,
+which we identify as due to metal lines from \znii, \feii, and \mgii.
+Additionally, \lya-absorption is seen at the end of the trace. The spectral
+position of the lines means that this GRB is at $z = 2.087$. From the
+\lya~trough, we infer \nh~$=21.7 \pm 0.4$. The afterglow continuum appears
+depressed in the blue end of the spectrum, suggestion dust extinction in the
+host.
 
@@
 
-This spectrum has already been used to inform GCN
-\#17523\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17523.gcn3}}, but it is not
-used elsewhere. Starting 5.1 hours after the BAT trigger, this spectrum is based
-on $6 \times 600$ s spectroscopic integration. The GRB afterglow continuum is
-well detected across the entire spectral coverage of X-shooter at moderate signal-to-noise.
-Imprinted on the continuum are absorption features which we identify as being
-caused by \SIii, \civ, \alii, \feii, \mgii, and \mgi - all at a similar redshift
-of $z = 1.517$, which is likely the redshift of the GRB.
+This spectrum has already been used to inform GCN 17523 \citep{GCN17523}, but it
+is not used elsewhere. Starting 5.1 hours after the BAT trigger, this spectrum
+is based on $6 \times 600$ s spectroscopic integration. The GRB afterglow
+continuum is well detected across the entire spectral coverage of X-shooter at
+moderate signal-to-noise. Imprinted on the continuum are absorption features
+which we identify as being caused by \SIii, \civ, \alii, \feii, \mgii, and \mgi
+- all at a similar redshift of $z = 1.517$, which is likely the redshift of the
+GRB.
 
@@
 
-This spectrum has previously been used to form the basis for GCN
-\#17672\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17672.gcn3}}. $4 \times
-600$ s spectroscopic integration, starting 10 hours after the GRB trigger,
-captures a bright GRB afterglow. A broad absorption trough centered at $\sim$
-3700 \AA due to \lya signals the redshift of the GRB, which is refined to $z =
-2.057$ based on the additional detection of metal absorption lines. We readily
-identify features associated with \sii, \SIiv, \oi, \SIii, \SIii$^*$, \cii,
-\cii$^*$, \civ, \alii, \feii, \feii$^*$, \mni, \mgii, and \mgi~in the host of
-the GRB. An intervening \civ absorber is additionally detected at $z=1.76$. From
-the absorption trough due to \lya, we infer the amount of neutral hydrogen in
-the host along the line of sight to be \nh~$=21.8 \pm 0.2$.
+This spectrum has previously been used to form the basis for GCN 17672
+\citep{GCN17672}. $4 \times 600$ s spectroscopic integration, starting 10 hours
+after the GRB trigger, captures a bright GRB afterglow. A broad absorption
+trough centered at $\sim$ 3700 \AA due to \lya signals the redshift of the GRB,
+which is refined to $z = 2.057$ based on the additional detection of metal
+absorption lines. We readily identify features associated with \sii, \SIiv, \oi,
+\SIii, \SIii$^*$, \cii, \cii$^*$, \civ, \alii, \feii, \feii$^*$, \mni, \mgii,
+and \mgi~in the host of the GRB. An intervening \civ absorber is additionally
+detected at $z=1.76$. From the absorption trough due to \lya, we infer the
+amount of neutral hydrogen in the host along the line of sight to be \nh~$=21.8
+\pm 0.2$.
 
@@
 The gross content of these observations have previously been presented in GCN
-\#17755\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17755.gcn3}}, but is not
-published as part of any refereed paper. This bona-fide short GRB (T90 (15-350
-keV ) = is $ 0.22 \pm 0.03$ s \citep{Lien2016}) was observed in RRM mode and
-spectroscopic integration started after only 22 minutes. A series of stare mode
-observations, increasing in exposure time, and ending with a nodding sequence
-totals $\sim$ 5000 s are combined to form this spectrum. A faint, almost
-featureless continuum is detected at low signal-to-noise all the way to the bluest part of
-the spectrum. An absorption doublet is detected in the VIS arm against the GRB
-afterglow continuum, which we identify as \mgii~at $z = 1.394$.
+17755 \citep{GCN17755}, but is not published as part of any refereed paper. This
+bona-fide short GRB (T90 (15-350 keV ) = is $ 0.22 \pm 0.03$ s \citep{Lien2016})
+was observed in RRM mode and spectroscopic integration started after only 22
+minutes. A series of stare mode observations, increasing in exposure time, and
+ending with a nodding sequence totals $\sim$ 5000 s are combined to form this
+spectrum. A faint, almost featureless continuum is detected at low
+signal-to-noise all the way to the bluest part of the spectrum. An absorption
+doublet is detected in the VIS arm against the GRB afterglow continuum, which we
+identify as \mgii~at $z = 1.394$.
 
@@
 hours after the trigger does not reveal anything conclusive. The host
-association is additionally ambiguous for this, likely extincted (GCN
-\#17767\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17767.gcn3}}), GRB.
+association is additionally ambiguous for this, likely extincted (GCN 17767
+\citep{GCN17767}), GRB.
 
@@
 
-This spectrum has already been used in GCN
-\#17822\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17822.gcn3}}. Spectroscopic
+This spectrum has already been used in GCN 17822 \citep{GCN17822}. Spectroscopic
 observations began 28.4 post trigger and consist of a $4 \times 600$ s nodding
-sequence. The GRB afterglow continuum is detected at moderate signal-to-noise across the
-entire spectral coverage of X-shooter. Narrow absorption features are imprinted
-in the continuum where we identify features from multiple \feii~transitions as
-well as the \mgii-doublet and the \mgi$\lambda2852.96$ resonance line. These
-lines are all found a position matching $z = 0.807$, suggesting it is the
-redshift of the GRB.
+sequence. The GRB afterglow continuum is detected at moderate signal-to-noise
+across the entire spectral coverage of X-shooter. Narrow absorption features are
+imprinted in the continuum where we identify features from multiple
+\feii~transitions as well as the \mgii-doublet and the \mgi$\lambda2852.96$
+resonance line. These lines are all found a position matching $z = 0.807$,
+suggesting it is the redshift of the GRB.
 
@@
 
-This gross content of this data has previously been issued in GCN
-\#17832\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17832.gcn3}}, but is not
-published. Starting $> 1$ day after the GRB trigger, $4 \times 600$ s
-spectroscopic integration securely allows us to measure the redshift of the
-host. A continuum is detected all the way through the spectral coverage with
-multiple emission lines superposed. It is not clear to what degree the GRB
-afterglow contributes to the continuum. We identify the emission lines as \oii,
-\hb, \oiii, \ha, \nii, \ha, and \sii~- all at $z = 0.256$. Due to the spatial
-proximity of this burst, it is a candidate for SN follow-up and indeed there are
-indications of re-brightening at the burst position (GCN
-\#17903\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/17903.gcn3}}).
+This gross content of this data has previously been issued in GCN 17832
+\citep{GCN17832}, but is not published. Starting $> 1$ day after the GRB
+trigger, $4 \times 600$ s spectroscopic integration securely allows us to
+measure the redshift of the host. A continuum is detected all the way through
+the spectral coverage with multiple emission lines superposed. It is not clear
+to what degree the GRB afterglow contributes to the continuum. We identify the
+emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and \sii~- all at $z =
+0.256$. Due to the spatial proximity of this burst, it is a candidate for SN
+follow-up and indeed there are indications of re-brightening at the burst
+position (GCN 17903 \citep{GCN17903}).
 
@@
 The overall content of these observations have previously been reported in GCN
-\#18080\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18080.gcn3}}, but is not
-published. Starting 5 hours after the \swift~GRB trigger, these observations
-consists of a combined $2 \times 1200$ s and $2 \times 600$ s. A blue continuum
-is detected across the entire spectral coverage of X-shooter, suggesting a
-significant contribution from the GRB afterglow. Superposed on the continuum are
-emission lines which we identify as \hb, \oiii, and \ha~with a measured 
-redshift of z = 0.313. Supporting this redshift is the tentative detection of
-the \mgii~absorption doublet in the afterglow continuum. The relative strength
-of the lines suggest that the line-forming region is dust obscured, contrary to
-the story told be the blue afterglow continuum.
+18080 \citep{GCN18080}, but is not published. Starting 5 hours after the
+\swift~GRB trigger, these observations consists of a combined $2 \times 1200$ s
+and $2 \times 600$ s. A blue continuum is detected across the entire spectral
+coverage of X-shooter, suggesting a significant contribution from the GRB
+afterglow. Superposed on the continuum are emission lines which we identify as
+\hb, \oiii, and \ha~with a measured redshift of z = 0.313. Supporting this
+redshift is the tentative detection of the \mgii~absorption doublet in the
+afterglow continuum. The relative strength of the lines suggest that the
+line-forming region is dust obscured, contrary to the story told be the blue
+afterglow continuum.
 
@@
 This gross content of this spectacular spectrum has already been issued in GCN
-\#18187\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18187.gcn3}}, but the
-spectrum has not been published. Starting just 12.4 minutes after the onboard
-trigger of \swift, a total of $4 \times 600$ s spectroscopic integration was
-obtained. The observations began just before dawn and the last two exposures are
-heavily affected by the brightening sky. A bright, high signal-to-noise GRB afterglow
-continuum is detected all across the the spectral window and imprinted on this
-are a myriad of absorption features. We identify individual lines from
-transitions in \aliii, \crii, \znii, \NIii$^*$, \feii, \feii$^*$, \scii, \mnii,
-\mgii, \mgi, \TIii, and  \caii~ - all at a consistent redshift of $z = 0.755$.
-The detection of fine-structure lines, exited local to the burst, clearly marks
-this as the redshift of the GRB.
+18187 \citep{GCN18187}, but the spectrum has not been published. Starting just
+12.4 minutes after the onboard trigger of \swift, a total of $4 \times 600$ s
+spectroscopic integration was obtained. The observations began just before dawn
+and the last two exposures are heavily affected by the brightening sky. A
+bright, high signal-to-noise GRB afterglow continuum is detected all across the
+the spectral window and imprinted on this are a myriad of absorption features.
+We identify individual lines from transitions in \aliii, \crii, \znii,
+\NIii$^*$, \feii, \feii$^*$, \scii, \mnii, \mgii, \mgi, \TIii, and  \caii~ - all
+at a consistent redshift of $z = 0.755$. The detection of fine-structure lines,
+exited local to the burst, clearly marks this as the redshift of the GRB.
 
@@
 the BAT trigger, these observations were stopped after $2 \times 600$s, when it
-became apparent that only a modest signal-to-noise was obtainable, and the redshift had
-already been published( GCN
-\#18273\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18273.gcn3}}, GCN
-\#18274\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18274.gcn3}}).  The
-continuum is detected at low signal-to-noise all across X-shooter and \mgii~is detected at
-the suggested redshift, which we take as the redshift of the GRB.
+became apparent that only a modest signal-to-noise was obtainable, and the
+redshift had already been published( GCN 18273 \citep{GCN18273}, GCN 18274
+\citep{GCN18274}).  The continuum is detected at low signal-to-noise all across
+X-shooter and \mgii~is detected at the suggested redshift, which we take as the
+redshift of the GRB.
 
@@
 
-The gross content of these spectra have already been issue in GCN \#
-18318\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18318.gcn3}}. Observations
-began 3.3 hours after the \swift~trigger and lasted for $4 \times 1200$ s. A
-moderate signal-to-noise afterglow continuum is detected across the entire spectral coverage
-of X-shooter, where however the bluest part is affected by \lya-forest
-absorption. Imprinted on the afterglow continuum are a wealth of both emission
-and absorption features, all caused in a system, for which we measure a redshift
-of $z = 1.968$. We identify in emission \oii, \hb, \oiii, and \ha and absorption
-due to \lya, \civ, \alii, \SIii, \feii, and \mgii. We additionally identify
-fine-structure absorption lines from \SIii$^*$ and \feii$^*$, at a similar
-redshift, which unequivocally sets the suggested redshift as the redshift of the
-GRB. The \lya-line is affected by an atmospheric transmission drop and
-\lya~emission in the trough, making the measurement of the neutral hydrogen
-column density difficult. We measure \nh~$21.2 \pm 0.3$.
+The gross content of these spectra have already been issue in GCN 18318
+\citep{GCN18318}. Observations began 3.3 hours after the \swift~trigger and
+lasted for $4 \times 1200$ s. A moderate signal-to-noise afterglow continuum is
+detected across the entire spectral coverage of X-shooter, where however the
+bluest part is affected by \lya-forest absorption. Imprinted on the afterglow
+continuum are a wealth of both emission and absorption features, all caused in a
+system, for which we measure a redshift of $z = 1.968$. We identify in emission
+\oii, \hb, \oiii, and \ha and absorption due to \lya, \civ, \alii, \SIii, \feii,
+and \mgii. We additionally identify fine-structure absorption lines from
+\SIii$^*$ and \feii$^*$, at a similar redshift, which unequivocally sets the
+suggested redshift as the redshift of the GRB. The \lya-line is affected by an
+atmospheric transmission drop and \lya~emission in the trough, making the
+measurement of the neutral hydrogen column density difficult. We measure
+\nh~$21.2 \pm 0.3$.
 
@@
 
-The data presented here also formed the basis of GCN \#
-18426\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
-published elsewhere. The observation was carried out in RRM starting 44 minutes
-after the GRB trigger. We detect absorption features from a wide range of ions
-at $z=2.330$ as well as intervening absorption at $z=1.49$.
+The data presented here also formed the basis of GCN 18426 \citep{GCN18982} and
+is not published elsewhere. The observation was carried out in RRM starting 44
+minutes after the GRB trigger. We detect absorption features from a wide range
+of ions at $z=2.330$ as well as intervening absorption at $z=1.49$.
 
@@
 
-The content of these spectra has previously been issued in GCN \#
-18506\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18506.gcn3}}, but is not
-published. Beginning 5 hours after the BAT trigger, 4 spectroscopic integrations
-obtained in a nodding sequence, each lasting for 600s, securely allows us to
-measure a redshift for this GRB. Beginning at $\sim$ 4700 \AA, the \lya-forest
-is clearly detected at high signal-to-noise leading up to the broad
-\lya~absorption trough and the onset of a bright GRB afterglow continuum.
-Imprinted in the afterglow continuum are a range of absorption lines which we
-identify as due to \SIii, \SIii$^*$, \oi, \cii, \cii$^*$, \civ, \alii,  \alii,
-\feii, and \feii$^*$. These lines are all detected at $z = 4.063$ and the
+The content of these spectra has previously been issued in GCN 18506
+\citep{GCN18506}, but is not published. Beginning 5 hours after the BAT trigger,
+4 spectroscopic integrations obtained in a nodding sequence, each lasting for
+600s, securely allows us to measure a redshift for this GRB. Beginning at $\sim$
+4700 \AA, the \lya-forest is clearly detected at high signal-to-noise leading up
+to the broad \lya~absorption trough and the onset of a bright GRB afterglow
+continuum. Imprinted in the afterglow continuum are a range of absorption lines
+which we identify as due to \SIii, \SIii$^*$, \oi, \cii, \cii$^*$, \civ, \alii, 
+\alii, \feii, and \feii$^*$. These lines are all detected at $z = 4.063$ and the
 presence of fine-structure lines at a consistent redshift securely sets this as
@@
 
-The gross content of these observations have previously been described in GCN \#
-18524\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18524.gcn3}}. Starting only 1
-hour after the BAT trigger, $2 \times 600$ s observations, extending into the
-morning twilight, were obtained of this GRB. The brightening sky affect the
-signal-to-noise of these spectra - especially in the blue end. The GRB afterglow
-is well detected over the entire spectral coverage of X-shooter, with spectral
-feaures from \feii~and \mgii~imprinted by an absorber at $z = 1.423$. The
-absorber er likely the host of the GRB and we therefore consider $z = 1.423$ the
-redshift of the GRB.
+The gross content of these observations have previously been described in GCN
+18524 \citep{GCN18524}. Starting only 1 hour after the BAT trigger, $2 \times
+600$ s observations, extending into the morning twilight, were obtained of this
+GRB. The brightening sky affect the signal-to-noise of these spectra -
+especially in the blue end. The GRB afterglow is well detected over the entire
+spectral coverage of X-shooter, with spectral feaures from \feii~and
+\mgii~imprinted by an absorber at $z = 1.423$. The absorber er likely the host
+of the GRB and we therefore consider $z = 1.423$ the redshift of the GRB.
 
@@
 
-The spectral features in this spectrum has previously been reported in GCN
-\#18540\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18540.gcn3}}. Observed in
-RRM mode, these observations were initiated only 19 minutes after the BAT
-trigger. As per the standard RRM scheme, a series of stare observation,
-increasing in en exposure time and ending with a regular $4 \times 600$ s
-nodding sequence, were acquired for this burst. A moderate signal-to-noise
-continuum is detected all the way through the X-shooter spectral coverage. In
-the GRB afterglow continuum, we identify absorption features as \feii, \mgii,
-\mgi, and \caii~at a consistent redshift of $z = 1.167$. Supporting this as the
-redshift of the GRB host galaxy are the simultaneous detection of nebular
-emission lines from \oii, \hb, and \oiii.
+The spectral features in this spectrum has previously been reported in GCN 18540
+\citep{GCN18540}. Observed in RRM mode, these observations were initiated only
+19 minutes after the BAT trigger. As per the standard RRM scheme, a series of
+stare observation, increasing in en exposure time and ending with a regular $4
+\times 600$ s nodding sequence, were acquired for this burst. A moderate
+signal-to-noise continuum is detected all the way through the X-shooter spectral
+coverage. In the GRB afterglow continuum, we identify absorption features as
+\feii, \mgii, \mgi, and \caii~at a consistent redshift of $z = 1.167$.
+Supporting this as the redshift of the GRB host galaxy are the simultaneous
+detection of nebular emission lines from \oii, \hb, and \oiii.
 
@@
 A preliminary analysis of these spectra have previously been presented in GCN
-\#18886\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18886.gcn3}}. Starting 13.5
-hours after the \swift~trigger, these spectra are formed on the basis of $4
-\times 1200$ s spectroscopic integration. The GRB afterglow is clearly detected
-at high signal-to-noise across the entire spectral coverage of X-shooter.
-Superposed on the continuum are absorption and emission features, which is
-identified as \feii, \mgii, \mgi~in absorption, and \oii, \hb, and \oiii. These
-spectral position of these lines all corrspond to a redshift of $z = 0.870$,
-which most likely is the redshift of the GRB.
+18886 \citep{GCN18886}. Starting 13.5 hours after the \swift~trigger, these
+spectra are formed on the basis of $4 \times 1200$ s spectroscopic integration.
+The GRB afterglow is clearly detected at high signal-to-noise across the entire
+spectral coverage of X-shooter. Superposed on the continuum are absorption and
+emission features, which is identified as \feii, \mgii, \mgi~in absorption, and
+\oii, \hb, and \oiii. These spectral position of these lines all corrspond to a
+redshift of $z = 0.870$, which most likely is the redshift of the GRB.
 
@@
 
-The data presented here also formed the basis of GCN \#
-18982\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/18982.gcn3}} and is not
-published elsewhere. The observation was carried out in RRM starting 18 minutes
-after the GRB trigger. We detect absorption features from a wide range of ions
-at $z=3.517$ as well as intervening absorption at $z=2.203$. From the very high
-signal-to-noise spectrum we are able to measure \nh~$=21.75 \pm 0.10$, based on
-the shape of the \lya~profile. This burst is the subject of a forthcoming
-publication (Pugliese et al. in prep.).
-
-\subsection{GRB~160228A (z = 1.64)}	\label{160228}
+The data presented here also formed the basis of GCN 18982 \citep{GCN18982} and
+is not published elsewhere. The observation was carried out in RRM starting 18
+minutes after the GRB trigger. We detect absorption features from a wide range
+of ions at $z=3.517$ as well as intervening absorption at $z=2.203$. From the
+very high signal-to-noise spectrum we are able to measure \nh~$=21.75 \pm 0.10$,
+based on the shape of the \lya~profile. This burst is the subject of a
+forthcoming publication (Pugliese et al. in prep.).
+
+\subsection{GRB~160228A (z = 1.640)} \label{160228}
+
+The gross content of these spectra have previously been presented in GCN 19186
+\citep{GCN19186}. Taken long after the burst had faded, the target is the likely
+host galaxy (See GCN 19186 \citep{GCN19186}). In $4\times1200$ s spectroscopic
+integration we detected emission lines from the host. We find \oiii~and \ha~at a
+consistent $z = 1.640$.  We therefore suggest that this is the redshift of the
+GRB.
 
@@
 
-\subsection{GRB~160314A (z = 0.726)}\label{160314}
+This is potentially a short burst with extended emission (GCN 19148
+\citep{GCN19148}). The content of the spectra have previously been reported in
+GCN 19154 \citep{GCN19154}. Observations began 19.1 hours after the BAT trigger
+and consists of $4\times 1200$ s spectroscopic integration. A very faint
+continuum is detected in the UVB and VIS arm after binning in the dispersion
+direction, but no clear redshift can be inferred from the continuum shape. No
+emission lines are readily apparent, suggesting little or no star formation in
+the host.
+
+\subsection{GRB~160314A (z = 0.726)} \label{160314}
+
+The gross structure of these spectra have previously been issued in GCN 19192
+\citep{GCN19192}. Spectroscopic integration began 13 hours post GRB trigger and
+lasted for $4 \times 1200$ s. In the data we see a moderate signal-to-noise
+continuum across the entire spectral coverage of X-shooter. Clearly visible on
+top of the continuum are emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
+\nii, and \sii~at a consistent redshift of $z = 0.726$. It is unclear how much
+of the continuum is due to the host and how much is from the GRB afterglow. At
+the same redshift, we see the tentative detection of \feii-absorption,
+supporting the redshift measurement and suggesting a significant afterglow
+contribution to the continuum.
 
@@
 
+This short GRB with extended emission (GCN 19276 \citep{GCN19276}) was observed
+in RRM mode, with observations starting after only 7.7 minutes. The gross
+content has already been presented in GCN 19274 \citep{GCN19274}. These
+observations have the shortest delay between trigger time and start of
+observation. Spectroscopic integration began after 8.4 minutes and lasted for $3
+\times 600$. The last image in the nodding sequence was aborted because the
+object was setting and the telescope reached the hardware limit. A blue,
+moderate signal-to-noise continuum is detected across the entire spectral
+coverage of X-shooter with multiple absorption features imprinted on it. At $z =
+1.717$ we detected the broad absorption trough due to \lya~and we additionally
+find \alii~and \feii~at a consistent redshift, setting it as the redshift of the
+GRB. Additionally, two intervening \civ~absorption systems are detected at  $z =
+1.444$ and $z = 1.581$. From the \lya~absorption trough, we measure a neutral
+hydrogen column density of \nh$= 21.2 \pm 0.2$. This is the only short GRB for
+which we have a measurement of $N_{\mathrm{HI}}$ and is the subject of an
+forthcoming paper (Selsing et al. in prep).
+
 \subsection{GRB~160425A (z = 0.555)}\label{160425}
 
+These spectra were already used to form the basis for GCN 19350
+\citep{GCN19350}, but are not published anywhere else. Spectroscopic integration
+started after 7.2 hours and lasted for $4 \times 1200$ s. A low signal-to-noise
+continuum is detected across the spectral coverage of X-shooter, but no clear
+absorption lines are easily discernible. It is therefore unclear how much of the
+continuum is due to the host and how much is due to the GRB. Superposed on the
+continuum are two sets of emission lines, offset spatially, but at very similar
+redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and \sii~in both
+systems -- all at $z = 0.555$. As per GCN 19350 \citep{GCN19350}, the GRB
+probably occurred in an interacting pair of galaxies.
+
 \subsection{GRB~160625B (z = 1.406)}\label{160625}
 
+GRB~160625B is the first GRB for which a signicant linear polarization was
+measured \cite{Troja2017}. Because this burst was detected by \textit{Fermi} and
+is therefore not a part of the statistical sample. The gross content of these
+spectra has previously been presented in GCN 19600 \citep{GCN19600}. Starting 30
+hours after the trigger, these spectra are obtained in $4 \times 600$ s. The GRB
+afterglow continuum is detected a high signal-to-noise throughout the spectral
+coverage of X-shooter. Imprinted on the afterglow continuum are a multitude of
+absorption features, from which we identify the responsible elements as \SIii,
+\oi, \SIiv, \civ, \alii, \aliii, \feii, \znii, \mgii, and \mgi~at a consistent
+redshift of $z = 1.406$. An additional intervening \mgii~absorber is detected at
+$z = 1.319$.
+
 \subsection{GRB~160804A (z = 0.736)}\label{160804}
 
-The data presented here also formed the basis of GCN \#
-19773\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/19773.gcn3}}, and is
-published in \citet{Heintz2017b}. Observations started 22.37 hours after the
-BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the entire
-spectral coverage of X-shooter and absorption lines from \mgi, \mgii, \feii~and
-\alii~are found at $z = 0.736$. At the same redshift, emission lines from \oii,
-\oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch, lasting
-3.6ks, is obtained after the afterglow has faded, confirming the emission line
-detections. The host galaxy is found to have a roughly solar metallicity and is
-among the most luminous GRB hosts at $z < 1$.
-
-\subsection{GRB~161001A (z = 0.891?)}	\label{161001}
+The data presented here also formed the basis of GCN 19773 \citep{GCN19773}, and
+is published in \citet{Heintz2017b}. Observations started 22.37 hours after the
+BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the
+entire spectral coverage of X-shooter and absorption lines from \mgi, \mgii,
+\feii~and \alii~are found at $z = 0.736$. At the same redshift, emission lines
+from \oii, \oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch,
+lasting 3.6ks, is obtained after the afterglow has faded, confirming the
+emission line detections. The host galaxy is found to have a roughly solar
+metallicity and is among the most luminous GRB hosts at $z < 1$.
+
+\subsection{GRB~161001A (z = 0.891)}	\label{161001}
+
+The content and scope of these spectra has previously been issued in GCN 19971
+\citep{GCN19971}. $4 \times 600$ s spectroscopic integration was obtained of
+this potentially short GRB (GCN 19974\citep{GCN19974}), starting 6.1 hours after
+the GRB trigger. The main source in the slit is likely of stellar origin and
+thus a foreground object unrelated to the GRB. Superposed, but slightly offset,
+from the stellar source are emission lines from \oii, \hb, and \ha~at a
+consistent redshift of $z = 0.891$. As stated in GCN 19971 \citep{GCN19971}, the
+association between the GRB and the source at $z = 0.891$ is uncertain. An
+alternative host association is suggestion in GCN 19975 \citep{GCN19975} just
+outside the XRT error circle. The fading of the source in GCN 19975
+\citep{GCN19975} is, however, not confirmed (Kruehler, private communication)
+and thus a clear localization unclear. Due to the proximity of the $z =
+0.891$-host suggested here to the XRT localization, we consider this redshift as
+the most likely redshift of the GRB.
 
 \subsection{GRB~161007A (z =4.6 ???NEW!!!)} \label{161007}
+
 This data has not been published elsewhere. Observations for GRB~161007A started
-323 hours after the burst trigger and contains the potential host. 4 x 600 seconds
-of observations reveals a faint continuum rising abruptly above the noise at
-$\sim$ 685 nm and continuing through 2100 nm. A very low significance
-continuum is detected at shorter wavelengths, down to $\sim$ 600 nm. If the
-host is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha
-break and the absence of nebular emission lines is due to \oii being shifted
-out of the wavelength coverage. Alternatively, an early-type host at $z = 0.71$
-could exhibit the 400 nm break at 600 nm, but due to the preference of
-long-duration GRBs for star-forming galaxies, this is the least likely
-explanation, why we believe the high-z solution.
+323 hours after the burst trigger and contains the potential host. 4 x 600
+seconds of observations reveals a faint continuum rising abruptly above the
+noise at $\sim$ 685 nm and continuing through 2100 nm. A very low significance
+continuum is detected at shorter wavelengths, down to $\sim$ 600 nm. If the host
+is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha break
+and the absence of nebular emission lines is due to \oii~being shifted out of
+the wavelength coverage. Alternatively, an early-type host at $z = 0.71$ could
+exhibit the 400 nm break at 600 nm, but due to the preference of long-duration
+GRBs for star-forming galaxies, this is the least likely explanation, why we
+prefer the high-z solution.
 
@@
 
-The data presented here also formed the
-basis of GCN \# 20061\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/20061.gcn3}},
+The data presented here also formed the basis of GCN  20061 \citep{GCN20061},
 but is not published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8
@@
 
+The gross content of these spectra have previously been reported in GCN 20104
+\citep{GCN20104}. This extremely bright \textit{INTEGRAL} burst is not a part of
+the statistical sample. Starting 3 hours after the trigger, the spectra are
+based on $2 \times 600$ s. The GRB afterglow continuum is detected with the
+highest signal-to-noise of all the spectra presented here. A myriad of
+absorption features are visible against the continuum, many which would not be
+visible against lower signal-to-noise spectra. We identify at least 15
+intervening absorbers between $z = 1.243$ and $z = 2.710$. In the host system we
+clearly see absorption from \lyb, \lya, \sii, \SIii, \oi, \SIiv, \civ, \alii,
+\aliii, \feii, \mgii, and \mgi. Several fine-structure transitions are
+additionally seen, clearly associating the $z = 2.710$ system with the GRB. From
+the \lya-line we infer \nh$=20.96 \pm 0.05$.
+
 \subsection{GRB~161117A (z = 1.549)}\label{161117}

### 2018-02-02 py/Sample_completeness.py

@@ -102,7 +102,7 @@ def main():
         dat = ''.join(c for c in ii if c.isdigit())
         print(dat)
 
-    exit()
+    # exit()
 
 
 
@@ -231,7 +231,7 @@ def main():
     # exit()
 
     for ax in (ax1, ax2, ax3):
-        ax.spines['top'].set_visible(False)
+        # ax.spines['top'].set_visible(False)
         ax.get_xaxis().tick_bottom()
         ax.get_yaxis().tick_left()
         ax.tick_params(axis='x', direction='out')

### 2018-02-04 document/XSGRB_sample.tex
Last bursts and edits

@@
 	afterglows that have been promptly observed with the X-shooter spectrograph up
-	to \termdate. This totals spectroscopic observations of 102 individual bursts
-	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
+	to \termdate. This totals spectroscopic observations of 103 individual bursts
+	observed with 48 hours of the GRB trigger. Redshifts have been measured for 97
 	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
@@
 	intrinsic properties of the GRBs, the follow-up effort has been focused on
-	producing a homogeneous sample of 91 afterglow spectra for GRBs discovered by
+	producing a homogeneous sample of 93 afterglow spectra for GRBs discovered by
 	the {\em Swift} satellite. We here provide a public release of all the reduced
@@
 	the bursts in the sample and find that the sample presented here is
-	representative of the full \textit{Swift} sample. For all bursts (41) for which
+	representative of the full \textit{Swift} sample. For all bursts (42) for which
 	it is possible, we provide a measurement of the neutral hydrogen column density,
@@
 	until \termdate. Stars indicate bursts with measured redshifts and circles
-	indicate those without. Blue stars indicate the position of the 98 bursts
+	indicate those without. Blue stars indicate the position of the 104 bursts
 	fulfilling the sample criteria specified in Sect. \ref{samplecrit} that have a
-	measured redshift. Red outlines are added to the 77 GRBs that enter our
+	measured redshift. Red outlines are added to the 84 GRBs that enter our
 	statistical sample with both X-shooter spectroscopy and a measured redshift.
@@
 	and has a measured redshift, but is not a part of our statistical sample. The
-	blue dots show the positions of the 63 GRBs of our statistical sample that lack
-	redshift measurements. A red outline is added around the 8 bursts in our
+	blue dots show the positions of the 61 GRBs of our statistical sample that lack
+	redshift measurements. A red outline is added around the six bursts in our
 	statistical sample which was observed with X-shooter, but did not yield a
-	redshift measurement. A single empty circle with a red outline indicate a burst
+	redshift measurement. Two empty circles with a red outlines indicate bursts
 	outside the statistical sample that was followed up with X-shooter, but without
@@
 
+%\begin{figure}[!ht]
+%	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+%\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
+%	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
+%	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
+%	horizontal axis is in units pixels, the top axis in arc seconds. The red
+%	datapoints show a subsample of measurements obtained for NIR spectra. The grey
+%	line shows a linear fit to the VIS datapoints. } \label{fig:res}
+%\end{figure}
+
 \begin{figure}[!ht]
-	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
-	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
-	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
-	horizontal axis is in units pixels, the top axis in arc seconds. The red
-	datapoints show a subsample of measurements obtained for NIR spectra. The grey
-	line shows a linear fit to the VIS datapoints. } \label{fig:res}
+	\centerline{\includegraphics[width=8cm]{figures/resolutions}} \caption{Blue
+	datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric
+	absorption lines in the VIS spectra, as a function of the airmass corrected
+	ambient FWHM. The red datapoints show a subsample of measurements obtained for
+	NIR spectra. The colored lines show the corresponding linear fits to the
+	datapoints. The colored bounds contain 68 per cent of the best-fit probability
+	mass. } \label{fig:res}
 \end{figure}
@@
 been reduced together with the afterglow spectra, using identical pipeline
-settings with the same version of the pipeline. First we fit a Gaussian function
-in the spatial direction of the trace of the standard star at 792 nm (i.e. in
-the VIS arm). After this, we fit a series of  20 telluric absorption lines in
-the telluric standard star spectra with Gaussians, taking care to select
-transitions that are not almost-resolved multiples, should be intrinsically
-unresolved, and are in areas with well defined continuum flux. We pick 34
-telluric standard stars spanning a range of DIMM seeing values, with the
-majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
-(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
-linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
-sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
-\ref{fig:res}). We use this linear relation as a way to estimate the spectral
-resolution for medium to low signal to noise afterglow spectra in the VIS arm.
-To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
-arm spectra of a subset of 19 sources used for the VIS arm above, and find that
-the resulting distribution is consistent with a simple scaling of the VIS arm
-relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
-filling, sources as given on the ESO instrument website. The UVB arm contains no
-suitable absorption lines to use, and we therefore use a scaled value as in the
-NIR arm. This simple analysis gives a sufficiently accurate estimate for the
-analysis of the low signal-to-noise science spectra.
+settings with the same version of the pipeline.
+
+%First we fit a Gaussian function
+%in the spatial direction of the trace of the standard star at 792 nm (i.e. in
+%the VIS arm). After this, we fit a series of  20 telluric absorption lines in
+%the telluric standard star spectra with Gaussians, taking care to select
+%transitions that are not almost-resolved multiples, should be intrinsically
+%unresolved, and are in areas with well defined continuum flux. We pick 34
+%telluric standard stars spanning a range of DIMM seeing values, with the
+%majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
+%(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
+%linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
+%sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
+%\ref{fig:res}). We use this linear relation as a way to estimate the spectral
+%resolution for medium to low signal to noise afterglow spectra in the VIS arm.
+%To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
+%arm spectra of a subset of 19 sources used for the VIS arm above, and find that
+%the resulting distribution is consistent with a simple scaling of the VIS arm
+%relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
+%filling, sources as given on the ESO instrument website. The UVB arm contains no
+%suitable absorption lines to use, and we therefore use a scaled value as in the
+%NIR arm. This simple analysis gives a sufficiently accurate estimate for the
+%analysis of the low signal-to-noise science spectra.
+
+%fit parameter 1-sigma error
+%———————————-
+%6.96733427249 +- 1.35492333728
+%36.1113577532 +- 1.36818568362
+%fit parameter 1-sigma error
+%———————————-
+%3.64773146238 +- 0.563740522948
+%24.2308362925 +- 0.792580030602
 
@@
 average redshift. However, the other samples also exhibit a large spread in the
-redshift distributions. A 2-sided KS test confirms that there is only mild
-evidence against the Fynbo09 (p-value = 0.05) and TOUGH (p-value = 0.05) samples
-being drawn from the same parent sample, while XS-GRB is consistent with the
-SHOALS (p-value = 0.08) and BAT6 (p-value = 0.86) samples.
+redshift distributions. A 2-sided KS test reveals that the XS-GRB sample is consistent with
+being drawn from the same parent sample with the following p-values: SHOALS (p-value = 0.13), BAT6 (p-value = 0.95), Fynbo09 (p-value = 0.08) and TOUGH (p-value = 0.09).
 
@@
 we artificially bias the distribution towards higher $\beta_{OX}$-values. The
-two distributions exhibit a large degree of overlap. We cannot confirm the
+two distributions exhibit a large degree of overlap. We confirm the
 result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher
@@
 within the error intervals. A 2-sided KS test fails to reject the null
-hypothesis that they are drawn from the same distribution with $p = 0.43$,
+hypothesis that they are drawn from the same distribution with $p = 0.11$,
 meaning that there is no strong evidence for a discrepancy. A Kendall's $\tau$
-test suggest a non-significant, very low degree of negative correlation ($\Gamma
-= -0.15$ at a p-value = 0.10).
+test, however, suggest a statistically significant, low degree of negative
+correlation ($\Gamma = -0.21$ at a p-value = 0.01).
 
@@
 \clearpage
+\begin{figure*}[!h]
+	\centering
+	\includegraphics[page=6, width=16cm]{figures/HI_measurements.pdf}
+	\caption*{Fig. \ref{fig:HI1}. continued.}
+	\label{fig:HI6}
+\end{figure*}
+\clearpage
 
@@
 detection of emission features belonging to nebular lines, including \hi, \oii,
-\oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due to
+\oiii, \neiii, \nii, [\sii], [\siii], and \hei~as well as absorption features due to
 the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all
@@
 of a broad, underlying stellar continuum. We identify emission lines from \oii,
-\hd, \hg, \hb, \oiii, \nii, \hb, \sii, \ariii, and \sii, all at $z=0.013$.
+\hd, \hg, \hb, \oiii, \nii, \hb, [\sii], \ariii, and [\siii], all at $z=0.013$.
 Significant velocity structure of the lines across the spatial direction of the
@@
 
-\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)} \label{120805}
+\subsection{GRB~120805A (z = 3.9 NEW!!!)} \label{120805}
 
@@
 time, we detect a faint continuum visible from 450 nm and all the way through
-2100 nm, in contrast to what is found previously. The continuum from 4500 - 600
+2100 nm, in contrast to what is found previously. The continuum from 450 - 600
 nm is detected at very low significance. If the drop at 450 nm is the Lyman
 limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim 3.9$. The
-absence of nebular lines if due to \oii~falling in a telluric absorption band
-and the rest being shifted out of the wavelength coverage.
+absence of nebular lines is due to \oii~falling in a telluric absorption band
+and the rest of the nebular lines being shifted out of the wavelength coverage.
 
@@
 
+This spectrum has previously been published in \citet{Tanvir2017a}. Starting
+18.5 hours after the BAT trigger, the final spectrum is based on $2 \times 4
+\times 1200$ s spectroscopic integration for a total of 160 minutes. Nothing is
+immediately visible in the 2D-spectra, however after severe binning in the
+dispersion direction, a faint trace shows up in the NIR arm, suggesting a very
+high redshift. We here adopt the redshift ($z = 7.84$) suggested by
+\citet{Tanvir2017a}.
+
 \subsection{GRB~121024A (z = 2.300)} \label{121024}
@@
 
-\subsection{GRB~121209A (z = 2.1)} \label{121209}
-
-This data is published in \citet{Kruhler2015}, but is not used elsewhere. This
-short spectroscopic integration of 1.2ks reveals a very faint trace visible in
-the visual arm after binning. The spectrum is taken at high airmass and
-observations were discontinued due to the hardware limit of the telescope. No
-emission lines are detected in the spectrum. The redshift is adopted from
-\citet{Kruhler2015}.
+%\subsection{GRB~121209A (z = 2.1)} \label{121209}
+%
+%This data is published in \citet{Kruhler2015}, but is not used elsewhere. This
+%short spectroscopic integration of 1.2ks reveals a very faint trace visible in
+%the visual arm after binning. The spectrum is taken at high airmass and
+%observations were discontinued due to the hardware limit of the telescope. No
+%emission lines are detected in the spectrum. The redshift is adopted from
+%\citet{Kruhler2015}.
 
@@
 superposed are both absorption (\cahk~and \mgii) and emission lines (\oii, \hb,
-\oiii, \ha, and \sii), all at a consistent redshift of $z = 0.356$, which is the
+\oiii, \ha, and [\sii]), all at a consistent redshift of $z = 0.356$, which is the
 redshift of the GRB.
@@
 
-\subsection{GRB~130615A (z $\sim 3$)} \label{130615}
+\subsection{GRB~130615A (z = 2.9)} \label{130615}
 
@@
 background, a faint afterglow trace is visible across all arms of X-shooter,
-down to ~ 4800 \AA, confirming the approximate redshift suggested in GCN 14898
+down to ~ 4800 \AA, which if interpreted as the break due to \lya suggests $z =
+2.9$. This supports the approximate redshift suggested in GCN 14898
 \citep{GCN14898}.
@@
 primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
-\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, as well as those
+\hb, \oii, \ha, \nii, [\sii]) are well detected at $z = 0.347$, as well as those
 from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
@@
 underlying host continuum. The host exhibits usual nebular emission from \oii,
-\hb, \oiii, \ha, \nii, and \sii which puts it at $z = 1.416$. Supporting this as
+\hb, \oiii, \ha, \nii, and [\sii] which puts it at $z = 1.416$. Supporting this as
 the redshift of the GRB are absorption features from the \mgii-doublet in the
@@
 to what degree the GRB afterglow contributes to the continuum. We identify the
-emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and \sii~- all at $z =
+emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and [\sii]~- all at $z =
 0.256$. Due to the spatial proximity of this burst, it is a candidate for SN
@@
 top of the continuum are emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
-\nii, and \sii~at a consistent redshift of $z = 0.726$. It is unclear how much
+\nii, and [\sii]~at a consistent redshift of $z = 0.726$. It is unclear how much
 of the continuum is due to the host and how much is from the GRB afterglow. At
@@
 continuum are two sets of emission lines, offset spatially, but at very similar
-redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and \sii~in both
+redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and [\sii]~in both
 systems -- all at $z = 0.555$. As per GCN 19350 \citep{GCN19350}, the GRB
@@
 \feii~and \alii~are found at $z = 0.736$. At the same redshift, emission lines
-from \oii, \oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch,
+from \oii, \oiii, \ha, \hb, \hg, \nii, [\sii], [\siii]~are found. A second epoch,
 lasting 3.6ks, is obtained after the afterglow has faded, confirming the
@@
 \citep{GCN19971}. $4 \times 600$ s spectroscopic integration was obtained of
-this potentially short GRB (GCN 19974\citep{GCN19974}), starting 6.1 hours after
-the GRB trigger. The main source in the slit is likely of stellar origin and
-thus a foreground object unrelated to the GRB. Superposed, but slightly offset,
-from the stellar source are emission lines from \oii, \hb, and \ha~at a
+this potentially short GRB \citep[GCN 19974;][]{GCN19974}, starting 6.1 hours
+after the GRB trigger. The main source in the slit is likely of stellar origin
+and thus a foreground object unrelated to the GRB. Superposed, but slightly
+offset, from the stellar source are emission lines from \oii, \hb, and \ha~at a
 consistent redshift of $z = 0.891$. As stated in GCN 19971 \citep{GCN19971}, the
@@
 
-\subsection{GRB~161219B (z = 1.549)}\label{161219}
+The majority of the spectral features of these spectra have previously been
+reported in GCN 20180 \citep{GCN20180}. Starting only $\sim$ 40 minutes after
+the BAT trigger, these observations consist of $4 \times 600$ s, taken under
+relatively poor observing conditions. The GRB afterglow continuum is detected
+throughout most of the X-shooter spectral coverage at low-to-moderate
+signal-to-noise, however the continuum level drops below the noise in the bluest
+part of the UVB arm at $\sim$ 3800 \AA. Imprinted in the continuum are
+absorption features, which we identify as due to \feii, \mgii, and \mgi~in an
+absorption system at $z = 1.549$. We consider this the likely redshift of the
+GRB.
+
+\subsection{GRB~161219B (z = 0.148)}\label{161219}
+
+The core signatures in these spectra have already been issued in GCN 20321
+\citep{GCN20321}, and are additionally used in both \citet{Ashall2017} and
+\citet{Cano2017}. This GRB is very close-by and is associated width the SN
+Ic-BL, SN2016jca \citep[GCN 20342;][]{GCN20342}. The observations consists of $4
+\times 600$ s spectroscopic integration, starting 1.5 days after the GRB
+trigger. The GRB afterglow continuum is detected at high signal-to-noise across
+the entire spectral coverage of X-shooter. Superposed on the continuum are
+absorption features, which we identify as being formed by the \mgii, \mgi, and
+the \cahk-lines at $z = 0.148$. Underneath the glaring afterglow continuum we
+additionally detect nebular emission from \oii, \hb, \oiii, \ha, and [\sii] at a
+consistent redshift, which is the likely redshift of the GRB.
+
+\subsection{GRB~170113A (z=1.968)}\label{170113}
+
+These gross content of these data have previously been issued in GCN 20458
+\citep{GCN20458}. Starting 15.2 hours after the BAT trigger, these observations
+consist of $4 \times 1200$ s spectroscopic integration. The GRB afterglow
+continuum is detected at moderate signal-to-noise across across the entire
+spectral coverage of X-shooter. In the bluest end og the spectral coverage, the
+background noise increases, this lowering the continuum below detection.
+Superposed on the afterglow continuum are both absorption lines and emission
+lines. In absorption we identify lines from \SIii, \feii, and \mgii, and in
+emission we detect \oii, \hb, and \oiii. All the lines are detected at a
+consistent redshift of $z=1.968$, which we take as the redshift of the GRB.
+
+\subsection{GRB~170202A (z=3.645)}\label{170202}
+
+GRB~170202A is the last GRB afterglow observed in this work. It concludes the
+follow-up effort that has lasted for 8 years. The majority of the spectral
+features in this spectrum has previously been reported in GCN 20589
+\citep{GCN20589}. Spectroscopy began 9.7 hours after the BAT trigger, and lasted
+for $4 \times 600$ s. The GRB afterglow is well detected at moderate-to-high
+signal-to-noise throughout the spectral coverage of X-shooter, down to 4300 \AA.
+The lowest wavelength continuum from 4300 \AA to 5600 \AA is absorbed by the
+\lya~forest, with the broad \lya~absorption trough centered at 5650 \AA. In the
+continuum, we additionally identify metal absorption features from \SIii, \civ,
+\SIiv, \feii, and \mgii, with the additionally detection of
+\SIii$^*$~fine-structure lines. With an additional detection of \oiii~and all
+lines detected at $z=3.645$, this unequivocally marks this as the redshift of
+the GRB. There is an additional \civ~absorber at $3.077$. From the
+\lya~absorption we measure \nh$= 21.55 \pm 0.10$.
+
+%\subsection{GRB~170214A (z = 2.53)}\label{170214A}
+%
+%
+% Greiner program
 

### 2018-02-04 py/betaOX.py

@@ -68,7 +68,7 @@ def main():
     # Script to produce beta OX.
     """
     # Read in burst list
-    burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
+    burst_table = pd.read_csv("../data/Burst list - observed.csv")
     name, sample, dt, acqmag, X_Fnu  = burst_table["GRB"].values, burst_table["In Sample"].values, burst_table["Follow-up delay"].values, burst_table["Acquisition mag"].values, burst_table["XRT flux density"].values
     name = np.array([ii[3:] for ii in name])
     # print(name)

### 2018-02-04 py/follow_up_timing.py

@@ -43,12 +43,12 @@ def main():
     # exit()
 
     for ii, ll in enumerate(mag):
-        # print(ll)
+        print(name[ii], ll)
         if "R" in str(ll):
             print(ll)
             mag[ii] = float(ll.split("=")[1]) + 0.21
-
-
+    # print(mag)
+    # exit()
     idx_limits = ~(mag == ">24")
     magnondet = 24 * (~idx_limits).astype("int")
 
@@ -68,7 +68,7 @@ def main():
     fig, ax1 = pl.subplots()
 
 
-    color = sns.color_palette()[0]
+    color = sns.color_palette("muted")[0]
     color_rgb = mpl.colors.colorConverter.to_rgb(color)
     colors = [sns.set_hls_values(color_rgb, l=l) for l in np.linspace(1, 0, 12)]
     cmap = sns.blend_palette(colors, as_cmap=True)
@@ -77,13 +77,13 @@ def main():
     sc = ax1.scatter(delay[idx_hasz & idx_limits], mag[idx_hasz & idx_limits], c=z[idx_hasz & idx_limits], cmap=cmap, s=35)
 
     # Without redshift and with acq mag
-    ax1.scatter(delay[~idx_hasz & idx_limits], mag[~idx_hasz & idx_limits], color=sns.color_palette()[2], s=45, marker="x")
+    ax1.scatter(delay[~idx_hasz & idx_limits], mag[~idx_hasz & idx_limits], color=sns.color_palette("muted")[2], s=45, marker="x")
     # With redshift, but without acq mag
     ax1.scatter(delay[idx_hasz & ~idx_limits], magnondet[idx_hasz & ~idx_limits], c=z[idx_hasz & ~idx_limits], cmap=cmap, s=175, marker=r'$\downarrow$')
     # No redshift and without acq mag
-    ax1.scatter(delay[~idx_hasz & ~idx_limits], magnondet[~idx_hasz & ~idx_limits], color=sns.color_palette()[2], s=175, marker=r'$\downarrow$')
+    ax1.scatter(delay[~idx_hasz & ~idx_limits], magnondet[~idx_hasz & ~idx_limits], color=sns.color_palette("muted")[2], s=175, marker=r'$\downarrow$')
     ax2 = pl.twinx()
-    ax2.plot(sorted_delay, fractional_completeness, color=sns.color_palette()[2])
[…]

### 2018-02-04 py/skymap.py

@@ -69,7 +69,7 @@ def main():
 
     # Overplot swift positions
     # Read in master sample
-    burst_table = pd.read_csv("../data/Burst list - CSV_master.csv")
+    burst_table = pd.read_csv("../data/Burst list - master.csv")
     name, z, observed, OA, = burst_table["GRB"].values, burst_table["Redshift"].values, burst_table["Observed"].values, burst_table["Afterglow"].values
 
 
@@ -82,7 +82,7 @@ def main():
             pass
 
     # Read in observed bursts
-    burst_table = pd.read_csv("../data/Burst list - CSV_observed.csv")
+    burst_table = pd.read_csv("../data/Burst list - observed.csv")
     name_o = np.array([ii[3:] for ii in burst_table["GRB"].values])
 
 
@@ -95,7 +95,22 @@ def main():
 
     # Read in Swift table
     swift_table = pd.read_table("../data/grb_table_1517061198.txt", delimiter="\t", dtype=None)
-    name_s, RA, DEC, redshift = swift_table["GRB"].values, pd.to_numeric(swift_table["BAT RA (J2000)"], errors='coerce').values, pd.to_numeric(swift_table["BAT Dec (J2000)"], errors='coerce').values, swift_table["Redshift"].values
+    name_s, RA, DEC, redshift, bat, xrt, nh = swift_table["GRB"].values, pd.to_numeric(swift_table["BAT RA (J2000)"], errors='coerce').values, pd.to_numeric(swift_table["BAT Dec (J2000)"], errors='coerce').values, swift_table["Redshift"].values, swift_table["BAT Fluence (15-150 keV) [10^-7 erg/cm^2]"].values, swift_table["XRT 11 Hour Flux (0.3-10 keV) [10^-11 erg/cm^2/s]"].values, swift_table["XRT Column Density (NH) [10^21 cm^-2]"].values
+
+    # nammmm = ["170111A", "170113A", "170127A", "170127B", "170202A", "170208A", "170208B"]
+
+    # idx = [ii for ii, kk in enumerate(name_s) if kk in nammmm]
+
+    # for ii, kk in enumerate(nammmm):
+    #     print(bat[idx][ii])
+
+    # for ii, kk in enumerate(nammmm):
+    #     print(xrt[idx][ii])
+
+    # for ii, kk in enumerate(nammmm):
+    #     print(nh[idx][ii])
+
+    # exit()
     count = 0
     for ii, kk in enumerate(name_s):
         hej = ''.join(filter(lambda x: x.isdigit(), kk))

### 2018-02-05 document/XSGRB_sample.tex
Final draft sent around

@@
 \abstract{In this work we present spectra of all $\gamma$-ray burst (GRB)
-	afterglows that have been promptly observed with the X-shooter spectrograph up
-	to \termdate. This totals spectroscopic observations of 103 individual bursts
-	observed with 48 hours of the GRB trigger. Redshifts have been measured for 97
+	afterglows that have been promptly observed with the X-shooter spectrograph
+	until \termdate. This totals spectroscopic observations of 103 individual GRBs
+	observed within 48 hours of the GRB trigger. Redshifts have been measured for 97
 	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
@@
 	spectra, including continuum estimates and telluric absorption corrections. For
-	completeness, we also provide reductions for the 18 late-time observations of the
-	underlying host galaxies. We provide an assessment of the degree of completeness
-	with respect to the parent GRB population, in terms of the X-ray properties of
-	the bursts in the sample and find that the sample presented here is
-	representative of the full \textit{Swift} sample. For all bursts (42) for which
-	it is possible, we provide a measurement of the neutral hydrogen column density,
-	increasing the total number of published H{\sc I} column density measurements by
-	$\sim$ 50 per cent. %	We
-	% constrain the
-	% fraction
-	%	of dark bursts to be less than 28 per cent based on the detection of an
-	% optical
-	%	afterglow and tentatively confirm previous results that larger optical
-	% darkness
-	%	is correlated with increased X-ray absorption.
-	This dataset provides a unique resource to study the ISM across cosmic time,
-	from the local progenitor surroundings to the galaxy-scale conditions.}
+	completeness, we also provide reductions for the 18 late-time observations of
+	the underlying host galaxies. We provide an assessment of the degree of
+	completeness with respect to the parent GRB population, in terms of the X-ray
+	properties of the bursts in the sample and find that the sample presented here
+	is representative of the full \textit{Swift} sample. We constrain the fraction
+	of dark bursts to be < 28 per cent and we confirm previous results, that higher
+	optical darkness is correlated with increased X-ray absorption. For the 42
+	bursts for which it is possible, we provide a measurement of the neutral
+	hydrogen column density, increasing the total number of published H{\sc I}
+	column density measurements by $\sim$ 33 per cent. This dataset provides a
+	unique resource to study the ISM across cosmic time, from the local progenitor
+	surroundings to the intervening universe.}
 
@@
 
-$\gamma$-ray bursts (GRBs) provide constraints on a very wide range of topics in
-astrophysics. Examples range from small-scale phenomena relating to magnetars,
-properties of highly relativistic jets, hyper/supernova explosions, the
-interstellar medium, dust extinction curves, starburst galaxies, chemical and
-molecular abundances, escape of ionizing radiation, the ionization state of the
-intergalactic medium, intervening absoption systems to standard candles in
-cosmology \citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
+$\gamma$-ray bursts (GRBs) are bright glimpses of electromagnetic radiation that
+pierce through the universe, all the way from the beginning of time. They
+provide constraints on a very wide range of topics in astrophysics. Examples
+range from small-scale phenomena relating to magnetars, properties of highly
+relativistic jets, hyper/supernova explosions, the interstellar medium, dust
+extinction curves, starburst galaxies, chemical and molecular abundances, escape
+of ionizing radiation, the ionization state of the intergalactic medium,
+intervening absoption systems to standard candles in cosmology
+\citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
 	Amati2008, Vergani2009, Prochaska2009, HjorthBloom2012, Rowlinson2017,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Spectral resolution} \label{resolution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{figure}[!ht]
+%	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
+%\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
+%	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
+%	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
+%	horizontal axis is in units pixels, the top axis in arc seconds. The red
+%	datapoints show a subsample of measurements obtained for NIR spectra. The grey
+%	line shows a linear fit to the VIS datapoints. } \label{fig:res}
+%\end{figure}
+
+\begin{figure}[!ht]
+	\centerline{\includegraphics[width=8cm]{figures/resolutuions}} \caption{Blue
+		datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric
+		absorption lines in the VIS spectra, as a function of the airmass corrected
+		ambient FWHM. The red datapoints show a subsample of measurements obtained for
+		NIR spectra. The colored lines show the corresponding linear fits to the
+		datapoints. The colored bounds contain 68 per cent of the best-fit probability
+		mass. The red and blue dashed line indicates the nominal resolution for the
+		slit-widths used for the telluric standard observations. As can be see, the
+		delivered resolution is in many cases superior to the nominal one. }
+	\label{fig:res}
+\end{figure}
+
+The afterglow spectra described in this paper are obtained in
+Target-of-Opportunity (override) mode. In most cases, there is little
+possibility to tweak slit widths to the seeing at the time of observations (i.e.
+to optimize spectral resolution and signal-to-noise), and almost all our data are
+therefore taken with a fixed set of slit widths and binning, described above. In
+a fair number of cases, the seeing full width at half maximum (FWHM) is
+considerably smaller than the slit width, and the delivered spectral resolution
+will then be determined by the seeing rather than by the slit width, as afterglows are
+point sources (this is evidently not the case for extended sources, e.g for host
+galaxies). The delivered resolution for slit-width dominated spectra
+post-reduction and extraction can easily be determined from the bright sky
+emission lines. For afterglow spectra with very high signal-to-noise, the
+delivered spectral resolution can at times be determined from the science data
+themselves. However, in the presence of multiple velocity components in
+absorption, other forms of line broadening, and a lack of lines at some
+redshifts, this is difficult to do at lower signal-to-noise ratios (the
+majority of spectra in our sample). A broad starting value for the expected
+resolution will help fitting of these spectra, and can be important in upper
+limit determination, and for this reason we construct a crude relation between
+the seeing and the delivered resolution at our slit width, binning, and
+reduction pipeline settings.
+
+To this end we use observations of telluric standard stars that are taken with
+identical instrument settings as our afterglow spectra, usually just after the
+science data, as part of the ESO X-shooter calibration plan. These spectra have
+been reduced together with the afterglow spectra, using identical pipeline
+settings with the same version of the pipeline. To get the resolutions, we
+select a series of atmospheric transitions that are resolved multiples which
+should be intrinsically unresolved, and are in areas with well defined continuum
+flux. We then fit the lines with Voigt-profiles and calculate a corresponding
+FWHM, which we can then convert to a delivered resolution using the wavelength
+of the chosen transitions. To get the resolution as a function of ambient
+conditions, for each observations we also calculate the airmass-corrected
+\href{http://www.eso.org/asm/ui/publicLog}{DIMM} seeing, which is measured at
+500 nm.
+
+The resulting distribution of spectral FWHM (km/s) as a function of spatial
+FWHM at 500 nm is fairly well described in the VIS arm by a linear relation $a
++ b*x$, with $x$ the spatial FWHM in arcseconds, $a= 25.2 \pm 0.8$ km/s, $b=3.8
+\pm 0.7$ (see Fig. \ref{fig:res}). For the NIR arm the corresponding relation
+is $a= 36.0 \pm 1.8$ km/s, $b=8.4 \pm 1.8$. We use these linear relations as a
+way to estimate the spectral resolution for medium to low signal-to-noise
+afterglow spectra. To extend this to the UVB arm, we calculate the ratio
+between the VIS and the NIR arm resolutions and find that the resulting
+distribution is consistent with a simple scaling of the VIS arm relation by the
+ratio of resolutions of the NIR and VIS arm for unresolved, slit filling,
+sources as given on the
+\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{ESO instrument website}.
+The UVB arm contains no suitable absorption lines to use, and we therefore use a scaled value.
+This simple analysis gives a sufficiently accurate estimate for the analysis of the low signal-to-noise science spectra.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spectral extraction}    \label{extract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Spectral resolution} \label{resolution}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{figure}[!ht]
-%	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
-%\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
-%	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
-%	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
-%	horizontal axis is in units pixels, the top axis in arc seconds. The red
-%	datapoints show a subsample of measurements obtained for NIR spectra. The grey
-%	line shows a linear fit to the VIS datapoints. } \label{fig:res}
-%\end{figure}
-
-\begin{figure}[!ht]
-	\centerline{\includegraphics[width=8cm]{figures/resolutions}} \caption{Blue
-	datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric
-	absorption lines in the VIS spectra, as a function of the airmass corrected
-	ambient FWHM. The red datapoints show a subsample of measurements obtained for
-	NIR spectra. The colored lines show the corresponding linear fits to the
-	datapoints. The colored bounds contain 68 per cent of the best-fit probability
-	mass. } \label{fig:res}
-\end{figure}
-
-The afterglow spectra described in this paper are obtained in
-Target-of-Opportunity (override) mode. In most cases, there is little
-possibility to tweak slit widths to the seeing at the time of observations (i.e.
-to optimize spectral resolution and signal-to-noise), and almost all our data are
-therefore taken with a fixed set of slit widths and binning, described above. In
-a fair number of cases, the seeing full width at half maximum (FWHM) is
-considerably smaller than the slit width, and the delivered spectral resolution
-will then be determined by the seeing rather than by the slit width, as afterglows are
-point sources (this is evidently not the case for extended sources, e.g for host
-galaxies). The delivered resolution for slit-width dominated spectra
-post-reduction and extraction can easily be determined from the bright sky
-emission lines. For afterglow spectra with very high signal-to-noise, the
-delivered spectral resolution can at times be determined from the science data
-themselves. However, in the presence of multiple velocity components in
-absorption, other forms of line broadening, and a lack of lines at some
-redshifts, this is difficult to do at lower signal-to-noise ratios (the
-majority of spectra in our sample). A broad starting value for the expected
-resolution will help fitting of these spectra, and can be important in upper
-limit determination, and for this reason we construct a crude relation between
-the seeing and the delivered resolution at our slit width, binning, and
-reduction pipeline settings.
-
-To this end we use observations of telluric standard stars that are taken with
-identical instrument settings as our afterglow spectra, usually just after the
-science data, as part of the ESO X-shooter calibration plan. These spectra have
-been reduced together with the afterglow spectra, using identical pipeline
-settings with the same version of the pipeline.
-
-%First we fit a Gaussian function
-%in the spatial direction of the trace of the standard star at 792 nm (i.e. in
-%the VIS arm). After this, we fit a series of  20 telluric absorption lines in
-%the telluric standard star spectra with Gaussians, taking care to select
-%transitions that are not almost-resolved multiples, should be intrinsically
-%unresolved, and are in areas with well defined continuum flux. We pick 34
-%telluric standard stars spanning a range of DIMM seeing values, with the
-%majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
-%(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
-%linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
-%sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
-%\ref{fig:res}). We use this linear relation as a way to estimate the spectral
-%resolution for medium to low signal to noise afterglow spectra in the VIS arm.
-%To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
-%arm spectra of a subset of 19 sources used for the VIS arm above, and find that
-%the resulting distribution is consistent with a simple scaling of the VIS arm
-%relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
-%filling, sources as given on the ESO instrument website. The UVB arm contains no
-%suitable absorption lines to use, and we therefore use a scaled value as in the
-%NIR arm. This simple analysis gives a sufficiently accurate estimate for the
-%analysis of the low signal-to-noise science spectra.
-
-%fit parameter 1-sigma error
-%———————————-
-%6.96733427249 +- 1.35492333728
-%36.1113577532 +- 1.36818568362
-%fit parameter 1-sigma error
-%———————————-
-%3.64773146238 +- 0.563740522948
-%24.2308362925 +- 0.792580030602
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Science data products} \label{products}
@@
 GRBs with a redshift determination decreases with follow-up delay. The redshift
-completeness for bursts that we have followed up is 92 per cent. This degree of
+completeness for bursts that we have followed up is 94 per cent. This degree of
 completeness in the followed bursts, illustrates the efficiency of VLT/X-shooter
@@
 
-Of all the BAT-triggered bursts, a total of 158 bursts fulfill the sample
+Of all the BAT-triggered bursts, a total of 165 bursts fulfill the sample
 criteria specified in Sect. \ref{samplecrit}, since the commissioning of
@@
 completeness of the full statistical sample is 61 per cent. We return to the
-question of redshift completeness in Sect. \ref{redshift}. From this sample, 91
+question of redshift completeness in Sect. \ref{redshift}. From this sample, 93
 GRBs have been spectroscopically followed up with X-shooter. In order to assess
@@
 Using the observational characteristics of the 1266 bursts observed until
-\termdate~by \textit{Swift}, and the derived $N_{\mathrm{HI,X}}$ \citep{Evans2009}, we can
-quantify the degree to which our sample is biased relative to the overall
-\textit{Swift} sample. The values are queried from the online \textit{Swift}
+\termdate~by \textit{Swift}, and the derived $N_{\mathrm{HI,X}}$
+\citep{Evans2009}, we can quantify the degree to which our sample is biased
+relative to the overall \textit{Swift} sample. The values are queried from the
+online \textit{Swift}
 database\footnote{\url{http://swift.gsfc.nasa.gov/archive/grb\_table/}}$^,$\footnote{\url{http://www.swift.ac.uk/xrt\_live\_cat/}}.
@@
 
-Out of the 158 bursts in the statistical sample, 33 ( $\sim$ 20 per cent) were
+Out of the 165 bursts in the statistical sample, 36 (22 per cent) were
 not observed due to reasons unrelated to the GRB or afterglow properties. The
@@
 the statistical properties of the full sample. Removing these bursts for the
-statistical sample, dramatically improves the redshift completeness from 61 per cent to 84
+statistical sample, dramatically improves the redshift completeness from 61 per cent to 88
 per cent. The remaining burst not followed up already had redshift from other
@@
 unlikely to yield a redshift measurement. In the remainder of the text, we
-consider the 125 bursts our statistical sample.
+consider the 129 bursts our statistical sample.
 
@@
 are only able to observe GRBs that are $\sim$ hundred times brighter than the
-faintest bursts at $z=1$ and below. The effect of GRB redshift on the \swift~triggering criteria have previously been studied in detail \citep{Littlejohns2013a}.
+faintest bursts at $z=1$ and below. The effect of GRB redshift on the
+\swift~triggering criteria have previously been studied in detail
+\citep{Littlejohns2013a}.
 
@@
 average redshift. However, the other samples also exhibit a large spread in the
-redshift distributions. A 2-sided KS test reveals that the XS-GRB sample is consistent with
-being drawn from the same parent sample with the following p-values: SHOALS (p-value = 0.13), BAT6 (p-value = 0.95), Fynbo09 (p-value = 0.08) and TOUGH (p-value = 0.09).
-
-Because the redshift completeness of our statistical sample is 84 per cent,
+redshift distributions. A 2-sided KS test reveals that the XS-GRB sample is
+consistent with being drawn from the same parent sample with the following
+p-values: SHOALS (p-value = 0.13), BAT6 (p-value = 0.95), Fynbo09 (p-value =
+0.08) and TOUGH (p-value = 0.09).
+
+Because the redshift completeness of our statistical sample is 88 per cent,
 making inference of the \textit{true} redshift distribution of GRBs based on
@@
 comparison, for the afterglows in the statistical sample we have observed with
-an optical afterglow detected, the redshift completeness is 91 per cent. This
+an optical afterglow detected, the redshift completeness is 92 per cent. This
 also goes to show that the lack of redshift completeness in the sample presented
@@
 
-In the compilation of $N_{\mathrm{HI}}$ measurements towards GRBs in Tanvir et
-al. (submitted) %\citet{Tanvir2017}
-there are 81 published $N_{\mathrm{HI}}$ values, excluding the measurements
-provided here. We here provide 39 new neutral hydrogen column density
-measurements -- an increase of the number of optically derived hydrogen column
-densities of $\sim$ 50 per cent. We show the two distributions in  Fig.
-\ref{fig:NH_dist}. We compare the median, the 14th, and 86th percentiles of the
-two distributions. The sample presented has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the
-rest of the literature values has \nh~$= 21.5_{-1.5}^{+0.4}$. We see that the
-two distributions have a large degree of overlap due to the large width of the
-distributions, but we find a slightly higher median value for the new sample
-presented here. A 2-sided KS test gives a p-value of p = 0.006, meaning
-relatively large evidence against the null that the two samples are drawn from
-the same underlying distribution. Because the bursts that have measurements of
-the hydrogen column density are selected solely based on our ability to infer a
-column, it is difficult to make any strong conclusions about the population
-statistics in terms of gas content.
+12 of the $N_{\mathrm{HI}}$ measurements for these spectra have previously been
+presented in \citet{Cucchiara2015} (See Table \ref{tab:HI}). We provide new
+measurements here for completeness. In the compilation of $N_{\mathrm{HI}}$
+measurements towards GRBs in Tanvir et al. (submitted) there are 93 published
+$N_{\mathrm{HI}}$ values, excluding the measurements provided here. We here
+provide 30 new neutral hydrogen column density measurements -- an increase of
+the number of optically derived hydrogen column densities of $\sim$ 33 per cent.
+We show the two distributions in  Fig. \ref{fig:NH_dist}. We compare the median,
+the 14th, and 86th percentiles of the two distributions. The sample presented
+has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
+\nh~$= 21.5_{-1.5}^{+0.4}$. We see that the two distributions have a large
+degree of overlap due to the large width of the distributions, but we find a
+slightly higher median value for the new sample presented here. A 2-sided KS
+test gives a p-value of p = 0.006, meaning relatively large evidence against the
+null that the two samples are drawn from the same underlying distribution.
+Because the bursts that have measurements of the hydrogen column density are
+selected solely based on our ability to infer a column, it is difficult to make
+any strong conclusions about the population statistics in terms of gas content.
 
@@
 
-The sample presented here includes spectroscopic observations of 91 systems
+The sample presented here includes spectroscopic observations of 93 systems
 fulfilling our sample criteria, including 18 spectra that are late-time
@@
 published or submitted for publication focusing on many of these specific issues
-of our sample such as extinction curves \citep[][see also \citealt{Fynbo2014,
-	Heintz2017}]{Japelj2015, Zafar17}, emission lines from the underlying host
-galaxies \citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
-\citep{Christensen2017}, \citet{Arabsalmani2018} on the metallicity-scaling
-relations, and escape of ionizing radiation (Tanvir et al., submitted). A number
-of additional companion papers are also planned, investigating the detailed
-properties of the sample presented here, including equivalent width
-distributions (de Ugarte Postigo et al., in preparation), metallicities and
-kinematics (Th{\"o}ne et al., in preparation), high ionization lines (Heintz et
-al., in preparation), molecular lines (Bolmer et al., in preparation),
-fine-structure lines (Vreeswijk et al., in preparation), composite GRB afterglow
-spectrum (Selsing et al., in preparation).
+of our sample such as extinction curves (\citealt{Japelj2015}, Zafar et al.
+submitted, see also \citealt{Fynbo2014, Heintz2017}]), emission lines from the
+underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
+\mgii~ absorbers \citep{Christensen2017}, \citet{Arabsalmani2018} on the
+metallicity-scaling relations, and escape of ionizing radiation (Tanvir et al.,
+submitted). A number of additional companion papers are also planned,
+investigating the detailed properties of the sample presented here, including
+equivalent width distributions (de Ugarte Postigo et al., in preparation),
+metallicities and kinematics (Th{\"o}ne et al., in preparation), high ionization
+lines (Heintz et al., in submitted), molecular lines (Bolmer et al., in
+preparation), fine-structure lines (Vreeswijk et al., in preparation), and
+composite GRB afterglow spectrum (Selsing et al., in preparation).
 
@@
 of complete samples of GRB host galaxies \citep{Hjorth2012, Covino2013,
-	Perley2016}. \cite{Kruhler2013} argue, that very rich sightlines like that
+	Perley2016a}. \cite{Kruhler2013} argue, that very rich sightlines like that
 probed by the remarkable GRB~080607 \citep{Prochaska2009} are probably
@@
 the BAT trigger, 4.8 ks spectroscopic integration yields a very faint trace down
-to $\sim$4300\AA, only visible after binning heavily. This could indicate a
+to $\sim$430 nm, only visible after binning heavily. This could indicate a
 redshift, $z ~\sim 2.5$, but no emission lines or absorption lines are
@@
 
-\subsection{GRB~110709B (z=2.109 (NEW!!!!))} \label{110709}
+\subsection{GRB~110709B (z=2.109)} \label{110709}
 
@@
 spectroscopic observation reveals after heavy binning, a wide, faint trace down
-to $\sim$ 5800 \AA, offset by 2.5\arcsec~relative to the centering of the slit.
+to $\sim$ 580 nm, offset by 2.5\arcsec~relative to the centering of the slit.
 No good redshift measurement can be inferred from this. We have adopted the
@@
 Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8ks
-reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 5000
-\AA. The simultaneous detection of absorption features identified as \lya,
+reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 500 nm. The simultaneous detection of absorption features identified as \lya,
 \SIii, \civ, \alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet,
@@
 observational epoch started 20.1 hours and lasted for 6.6ks. The GRB afterglow
-continuum is well detected down to $\sim$7600 \AA, with several strong
+continuum is well detected down to $\sim$760 nm, with several strong
 absorption features imprinted, all at a common $z = 4.990$. \lya~is clearly
@@
 The data presented here has been examined by \citet{Japelj2015} and
-\citet{Zafar2017} who both find a significant amount of extinction, $A_V\approx
+Zafar et al. (submitted), who both find a significant amount of extinction, $A_V\approx
 1$ mag. Three epochs of observations have been obtained, the first two
@@
 
-\subsection{GRB~120224A (z = 1.10 NEW!!!)} \label{120224}
+\subsection{GRB~120224A (z = 1.10)} \label{120224}
 
@@
 
-\subsection{GRB~120311 (z = 0.350 NEW!!!)} \label{120311}
+\subsection{GRB~120311 (z = 0.350)} \label{120311}
 
@@
 
-\subsection{GRB~120805A (z = 3.9 NEW!!!)} \label{120805}
+\subsection{GRB~120805A (z = 3.9)} \label{120805}
 
@@
 background, a faint afterglow trace is visible across all arms of X-shooter,
-down to ~ 4800 \AA, which if interpreted as the break due to \lya suggests $z =
+down to ~ 480 nm, which if interpreted as the break due to \lya suggests $z =
 2.9$. This supports the approximate redshift suggested in GCN 14898
@@
 spectroscopic integration captures a modest signal-to-noise GRB afterglow
-continuum all the way down to $\sim$3200 \AA. Imprinted on the continuum are
+continuum all the way down to $\sim$320 nm. Imprinted on the continuum are
 absorption features, which we identify as due to \lya, \feii,  \mgii, \mgi~at
@@
 
-\subsection{(CHECK ME - Bright un-used spectrum)GRB~131030A (z = 1.296)}	
+\subsection{GRB~131030A (z = 1.296)}	
 \label{131030}
@@
 which is measured to be at $z = 4.042$. A moderate signal-to-noise GRB afterglow
-continuum is detected down to $\sim$ 6100 \AA, signifying the onset of the
+continuum is detected down to $\sim$ 610 nm, signifying the onset of the
 \lya~forest, with part of the forest also visible. Metal absorption lines from
@@
 redshift measurement. By heavily binning the spectrum, a faint trace is visible
-down to 4850 \AA, which if interpreted at the onset of \lya signifies z $\sim$
+down to 485 nm, which if interpreted at the onset of \lya signifies z $\sim$
 3. We adopt the redshift inferred in \citet{Kruhler2015}.
@@
 
-\subsection{CHECK ME - Bright un-used spectrum - GRB~140311A (z = 4.954)}	
+\subsection{GRB~140311A (z = 4.954)}	
 \label{140311}
@@
 Gunn-Peterson trough is visible against the GRB afterglow continuum, starting at
-8900 \AA, which along with absorption lines from the \mgii~doublet securely sets
+890 nm, which along with absorption lines from the \mgii~doublet securely sets
 the redshift of this GRB at $z = 6.327$. From the red wing of the \lya-profile
@@
 these $4\times600$ s spectra contains a high signal-to-noise GRB afterglow. The
-afterglow continuum is readily detected all the way down to 3700 \AA, although
+afterglow continuum is readily detected all the way down to 370 nm, although
 the bluest part is affected by the \lya-forest. A broad absorption trough from
@@
 signal-to-noise GRB afterglow continuum throughout the X-shooter spectral
-coverage, down to $\sim$ 3750 \AA. In the continuum are absorption features,
+coverage, down to $\sim$ 375 nm. In the continuum are absorption features,
 which we identify as due to metal lines from \znii, \feii, and \mgii.
@@
 after the GRB trigger, captures a bright GRB afterglow. A broad absorption
-trough centered at $\sim$ 3700 \AA due to \lya signals the redshift of the GRB,
+trough centered at $\sim$ 370 nm due to \lya signals the redshift of the GRB,
 which is refined to $z = 2.057$ based on the additional detection of metal
@@
 became apparent that only a modest signal-to-noise was obtainable, and the
-redshift had already been published( GCN 18273 \citep{GCN18273}, GCN 18274
-\citep{GCN18274}).  The continuum is detected at low signal-to-noise all across
+redshift had already been published in GCN 18273 \citep{GCN18273} and GCN 18274
+\citep{GCN18274}. The continuum is detected at low signal-to-noise all across
 X-shooter and \mgii~is detected at the suggested redshift, which we take as the
@@
 600s, securely allows us to measure a redshift for this GRB. Beginning at $\sim$
-4700 \AA, the \lya-forest is clearly detected at high signal-to-noise leading up
+470 nm, the \lya-forest is clearly detected at high signal-to-noise leading up
 to the broad \lya~absorption trough and the onset of a bright GRB afterglow
@@
 \citep{GCN19971}. $4 \times 600$ s spectroscopic integration was obtained of
-this potentially short GRB \citep[GCN 19974;][]{GCN19974}, starting 6.1 hours
-after the GRB trigger. The main source in the slit is likely of stellar origin
-and thus a foreground object unrelated to the GRB. Superposed, but slightly
-offset, from the stellar source are emission lines from \oii, \hb, and \ha~at a
-consistent redshift of $z = 0.891$. As stated in GCN 19971 \citep{GCN19971}, the
-association between the GRB and the source at $z = 0.891$ is uncertain. An
-alternative host association is suggestion in GCN 19975 \citep{GCN19975} just
-outside the XRT error circle. The fading of the source in GCN 19975
-\citep{GCN19975} is, however, not confirmed (Kruehler, private communication)
-and thus a clear localization unclear. Due to the proximity of the $z =
-0.891$-host suggested here to the XRT localization, we consider this redshift as
-the most likely redshift of the GRB.
-
-\subsection{GRB~161007A (z =4.6 ???NEW!!!)} \label{161007}
+this potentially short GRB (GCN 19974; \citealt{GCN19974}), starting 6.1 hr
+after the GRB trigger. Close to the
+\href{http://www.swift.ac.uk/xrt_positions/00020702/}{XRT error circle}
+\citep{GCN19969} there is a bright, point-like source, also visible in archival
+images. Our spectrum reveals that this is an M-type star, based on the presence
+of TiO bands, and is thus unrelated to the GRB. Two more objects lie within or
+in the near proximity of the X-ray error circle, as visible in our images from
+the acquisition camera. One is blended with the above mentioned star, and is
+serendipitously covered by the X-shooter slit, revealing emission lines from
+\oii, \hb, and \ha~at a consistent redshift of $z = 0.891$. The second source,
+first noted by \citet{GCN19975}, is extended and lies just outside the XRT error
+circle. Based on GROND imaging, \citep{GCN19975} report moderate variability,
+and suggest this object as an afterglow candidate. Following our original
+imaging, we thus secured further late-time photometry with the X-shooter
+acquisition camera, starting on 2016 Oct 5.26 UT (4.22 days after the GRB),
+securing $3 \times 120$~s in each of the $g$ and $r$ filters. No variability is
+apparent in either bands. In offline analysis of the GROND data (T. Kr\"uhler,
+priv. comm.), the evidence for fading is less significant than originally
+reported. The situation is thus left ambiguous, but due to the consistency of
+the $z = 0.891$ object with the XRT localization, we consider this as the most
+likely redshift of GRB\,161001A.
+
+\subsection{GRB~161007A (z =4.7)} \label{161007}
 
 This data has not been published elsewhere. Observations for GRB~161007A started
-323 hours after the burst trigger and contains the potential host. 4 x 600
-seconds of observations reveals a faint continuum rising abruptly above the
-noise at $\sim$ 685 nm and continuing through 2100 nm. A very low significance
-continuum is detected at shorter wavelengths, down to $\sim$ 600 nm. If the host
-is located at $z \sim 4.6$, the drop in continuum flux is the Lyman alpha break
-and the absence of nebular emission lines is due to \oii~being shifted out of
-the wavelength coverage. Alternatively, an early-type host at $z = 0.71$ could
+323 hours after the burst trigger and contains the potential host. $4 \times
+600$ seconds of observations reveals a faint continuum rising abruptly above the
+noise at $\sim$ 690 nm and continuing through 2100 nm. If the host is located at
+$z \sim 4.7$, the drop in continuum flux is the Lyman and the
+absence of nebular emission lines is due to \oii~being shifted out of the
+wavelength coverage. Alternatively, an early-type host at $z = 0.71$ could
 exhibit the 400 nm break at 600 nm, but due to the preference of long-duration
@@
 signal-to-noise, however the continuum level drops below the noise in the bluest
-part of the UVB arm at $\sim$ 3800 \AA. Imprinted in the continuum are
+part of the UVB arm at $\sim$ 380 nm. Imprinted in the continuum are
 absorption features, which we identify as due to \feii, \mgii, and \mgi~in an
@@
 for $4 \times 600$ s. The GRB afterglow is well detected at moderate-to-high
-signal-to-noise throughout the spectral coverage of X-shooter, down to 4300 \AA.
-The lowest wavelength continuum from 4300 \AA to 5600 \AA is absorbed by the
-\lya~forest, with the broad \lya~absorption trough centered at 5650 \AA. In the
+signal-to-noise throughout the spectral coverage of X-shooter, down to 430 nm.
+The lowest wavelength continuum from 430 nmto 560 nm is absorbed by the
+\lya~forest, with the broad \lya~absorption trough centered at 565 nm. In the
 continuum, we additionally identify metal absorption features from \SIii, \civ,

### 2018-02-05 document/tables/redshift_comparison.tex
Many comments and notes on individual bursts
Final draft sent around

@@
 		\hline\noalign{\smallskip}
-		{Sample} & {N$_{\mathrm{bursts}}$} & {z$_{\mathrm{completeness}}$} &  {z$_{\mathrm{mean}}$} &  {$z_{\mathrm{med}}$} \\
+		{Sample} & {N$_{\mathrm{bursts}}$} & {$z_{\mathrm{completeness}}$} &  {$z_{\mathrm{mean}}$} &  {$z_{\mathrm{med}}$} \\
 		\hline\noalign{\smallskip}
 		{\smallskip}
-		XS-GRB & 125 & 84 \% & 1.86 & $1.46_{-0.86}^{+1.77}$ \\
+		XS-GRB & 129 & 88 \% & 1.89 & $1.52_{-0.91}^{+1.83}$ \\
 		{\smallskip}

### 2018-02-05 py/Sample_completeness.py

@@ -62,7 +62,7 @@ def main():
     l, m, h = np.nanpercentile(filt_nan(z[pure == "Yes"]), (14, 50, 86))
     mm = np.nanmean(filt_nan(z[pure == "Yes"]))
     print(mm, m, l - m, h - m)
-
+    exit()
     l, m, h = np.nanpercentile(fbo, (14, 50, 86))
     mm = np.nanmean(fbo)
     print(mm, m, l - m, h - m)

### 2018-02-15 document/XSGRB_sample.tex
Last updates before submission

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
+\documentclass[longauth]{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
+%\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
 %\documentclass[referee]{aa}
@@
 \usepackage{natbib}
-\usepackage{caption}
+%\usepackage{caption}
 
@@
 \usepackage{float}
+%\usepackage[symbol]{footmisc}
 
@@
 
-\title{XS-GRB I - The X-shooter GRB afterglow legacy sample \thanks{Based on
+\title{The X-shooter GRB afterglow legacy sample (XS-GRB). \thanks{Based on
 		observations collected at the European Southern Observatory, Paranal, Chile,
@@
 
-\titlerunning{The X-shooter sample of GRBs}
+\titlerunning{The X-shooter GRB afterglow legacy sample}
 
@@
 	afterglows that have been promptly observed with the X-shooter spectrograph
-	until \termdate. This totals spectroscopic observations of 103 individual GRBs
+	until \termdate. In total, we obtained spectroscopic observations of 103 individual GRBs
 	observed within 48 hours of the GRB trigger. Redshifts have been measured for 97
@@
 	is representative of the full \textit{Swift} sample. We constrain the fraction
-	of dark bursts to be < 28 per cent and we confirm previous results, that higher
+	of dark bursts to be < 28 per cent and we confirm previous results that higher
 	optical darkness is correlated with increased X-ray absorption. For the 42
@@
 
-\keywords{Gamma-ray burst: general --- galaxies: high-redshift --- ISM: molecules --- dust, extinction }
+\keywords{Gamma-ray burst: general --- Galaxies: high-redshift --- ISM: general --- Techniques: spectroscopic}
 
@@
 
-$\gamma$-ray bursts (GRBs) are bright glimpses of electromagnetic radiation that
-pierce through the universe, all the way from the beginning of time. They
-provide constraints on a very wide range of topics in astrophysics. Examples
-range from small-scale phenomena relating to magnetars, properties of highly
-relativistic jets, hyper/supernova explosions, the interstellar medium, dust
-extinction curves, starburst galaxies, chemical and molecular abundances, escape
-of ionizing radiation, the ionization state of the intergalactic medium,
-intervening absoption systems to standard candles in cosmology
-\citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
+Gamma-ray bursts (GRBs) are bright glimpses of electromagnetic radiation that
+pierce through the universe, all the way from the edges of the observable
+universe. They provide constraints on a very wide range of topics in
+astrophysics. Examples range from small-scale phenomena relating to magnetars,
+properties of highly relativistic jets, hyper/supernova explosions, the
+interstellar medium, dust extinction curves, starburst galaxies, chemical and
+molecular abundances, escape of ionizing radiation, the ionization state of the
+intergalactic medium, intervening absorption systems to standard candles in
+cosmology \citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
 	Amati2008, Vergani2009, Prochaska2009, HjorthBloom2012, Rowlinson2017,
@@
 
-The \textit{Niel Gehrels Swift Observatory} ({\it Swift}) satellite
+The \textit{Neil Gehrels Swift Observatory} ({\it Swift}) satellite
 \citep{Gehrels2004, Gehrels2009}, which was launched in 2004, has made it
@@
 intergalactic medium, which was already hinted at by results from earlier
-missions \citep[e.g.,][]{Paradijs2000, Ricker2004}. With the two on-board
-instruments, the Burst Alert Telescope (BAT) \citep{Barthelmy2005} and the X-Ray
-Telescope (XRT) \citep{Burrows2005}, \swift~is an ideal telescope for GRB
-hunting. A crucial aspect of the success of the {\it Swift} mission has been the
-extensive ground-based follow-up observations of the afterglows and of the host
-galaxies of the GRBs, involving a large community of researchers. This fruitful
-collaboration has been facilitated by the open data access policy of the
-{\it Swift} mission. The close collaboration between detection facilities and
+missions \citep[e.g.,][]{Paradijs2000, Ricker2004}. With the three on-board
+instruments, the Burst Alert Telescope \citep[BAT;][]{Barthelmy2005}, the X-Ray
+Telescope \citep[XRT;][]{Burrows2005}, and the UltraViolet and Optical Telescope
+\citep[UVOT;][]{Roming2005}, \swift~is an ideal observatory for GRB hunting. A
+crucial aspect of the success of the {\it Swift} mission has been the extensive
+ground-based follow-up observations of the afterglows and of the host galaxies
+of the GRBs, involving a large community of researchers. This fruitful
+collaboration has been facilitated by the open data access policy of the {\it
+	Swift} mission. The close collaboration between detection facilities and
 electromagnetic  follow-up campaigns continue to be immensely rewarding, as
 recently highlighted by the simultaneous detection of gravitational waves and
-light from the neutron star merger in the shape of GW170817/GRB180817A/AT2017gfo
+light from the neutron star merger in the shape of GW170817/GRB170817A/AT2017gfo
 \citep{LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
@@
 $R=\lambda/\Delta\lambda$$<$1000, e.g.,][]{Fynbo2009}. Spectroscopy is powerful
-as it allows us to secure information even for very faint targets (24.5
-\citep{Kruhler2012}). This allows the measurements of a number of important
+as it allows us to secure information even for very faint targets
+\citep{Kruhler2012}. This allows the measurement of a number of important
 parameters such as redshifts, spectral slopes, and extinction. For a handful of
@@
 generation instruments at the ESO Very Large Telescope (VLT). It was designed
-very much with transient source follow-up in mind as the fading luminosities of
-such sources makes it urgent to secure as extensive coverage as possible in the
+very much with transient follow-up in mind as the fading luminosities of
+such sources makes it urgent to secure as extensive wavelength coverage as possible in the
 shortest possible time. At the same time, the resolution was designed to be in
@@
 between the many sky-background emission lines in the red and near-IR spectral
-ranges. The large spectral coverage allows for the spectroscopic observations of
+ranges. The near-IR spectral coverage allows for spectroscopic observations of
 the highest redshift GRBs.
@@
 of afterglows and host galaxies of GRBs detected by {\it Swift}. We here make
-all data resulting from the survey, publicly available in reduced form (see Sect
+all the data resulting from the survey publicly available in reduced form (see Sect
 \ref{products}).
@@
 describe the results of the survey, i.e. the efficiency of the follow-up effort
-and the characteristics of the observed bursts. We also here assess the
+and the characteristics of the observed bursts. We also assess the
 completeness of the realized sample. Finally, we offer our conclusions
 in Sect.~\ref{conclusions}. We use the $\Lambda$CDM cosmology parameters
-provided by \citet{Planck2015} in which the universe is flat with $H_0 =
+provided by the \citet{Planck2015} in which the universe is flat with $H_0 =
 67.7$\,km\,s$^{-1}$ Mpc$^{-1}$ and $\Omega_\mathrm{m} = 0.307$.
@@
 of the telescope or unconstrained burst localizations as reported by the alerting
-telescope. To maximize the return of the follow-up campaign, we have chosen a
+facility. To maximize the return of the follow-up campaign, we have chosen a
 few selection criteria that facilitate an unbiased selection of bursts, while
@@
 local to the Milky Way and therefore independent of intrinsic GRB properties,
-the aim is that the sample collected represents the underlying distribution of
+the aim is that the collected sample represents the underlying distribution of
 GRBs in a fair way. The selection criteria used here are based on previous,
@@
 
-The selection criteria that define a GRB as part of our \textit{statistical}
+The selection criteria that define a GRB as part of our initial \textit{statistical}
 	sample are:
@@
 
-	\item GRB triggered by BAT onboard the \swift~satellite
+	\item GRB trigger by BAT onboard the \swift~satellite
 
@@
 
-	\item The target must be visible for at least 60 minutes, 30 degrees above the
-	horizon, with the Sun below $-12$ degrees\footnote{Note that in the P84
-		proposal the criteria have been stated differently, the visibility constraint
-		being replaced by a declination + Sun angle constraint. The criteria listed
-		above are, however, those defining the sample.}.
+	\item The target must be visible from Cerro Paranal for at least 60 minutes, 30
+degrees above the horizon, with the Sun below $-12$ degrees\footnote{Note that
+	in the P84 proposal the criteria have been stated differently, the visibility
+	constraint being replaced by a declination + Sun angle constraint. The criteria
+	listed above are, however, those defining the sample.}.
+
+	\item Galactic $A_V \leq 0.5$ mag according to the maps of \citet{Schlegel1998}.
 
@@
 arcsec, where $R$ is the USNO magnitude of the star)
-
-	\item Galactic $A_V \lesssim 0.5$ mag according to the maps of \citet{Schlegel1998}.
 
@@
 	old \citet{Schlegel1998} values for consistency with the first semesters of our
-	program.}. Choosing low-extinction sightlines also reduce the problem of field
+	program.}. Choosing low-extinction sightlines also reduces the problem of field
 crowding and the contamination from Galactic outbursts posing as GRB impostors.
 These additional cuts should not influence the optical properties of the bursts
-themselves, only our ability to successfully secure the observations that allows
+themselves, only our ability to successfully secure the observations that allow
 us to investigate the spectroscopic properties of GRBs.
 
-Bursts that fulfill these selection criteria are what we define as our
-\textit{statistical} sample. Using this sample we will be able to address
-population properties of \swift-detected bursts. We further discuss the effect
-of these selection criteria and their implication for the completeness of the
-sample in Sect. \ref{completeness}.
+Bursts that fulfill these selection criteria are what we define as our initial
+\textit{statistical} sample. To reach a higher degree of sample completeness, we
+impose additional cuts that increase our ability to infer population properties.
+These are presented in Sect. \ref{badbursts}. Using this sample we will be able
+to address population properties of \swift-detected bursts. We further discuss
+the effect of these selection criteria and their implication for the
+completeness of the sample in Sect. \ref{completeness}.
 
@@
 sample. Members of our collaboration have access to numerous facilities spread
-throughout the world, among which the Nordic Optical Telescope (NOT), the Gran
-Telescopio Canarias (GTC), the Telescopio Nazionale Galileo (TNG), the Gamma-Ray
-Burst Optical Near-infrared Detector (GROND), the Xinlong observatory. Thanks to
-this network, we can often conduct searches for optical/NIR afterglows ahead of
-the time of the X-shooter spectroscopy, thus helping to plan the spectroscopic
+throughout the world, among which the William Herschel Telescope (WHT), Nordic
+Optical Telescope (NOT), the Gran Telescopio Canarias (GTC), the Telescopio
+Nazionale Galileo (TNG), the Gamma-Ray Burst Optical Near-infrared Detector
+(GROND \citealt{Greiner2008}), and the Xinlong observatory. Thanks to this
+network, we can often conduct searches for optical/NIR afterglows ahead of the
+time of the X-shooter spectroscopy, thus helping to plan the spectroscopic
 observations. If this is not possible, we use the built-in acquisition camera of
@@
 fly. In many cases we could interact directly with the staff at the telescope to
-aid the observation. Raw data are usually available within minutes in the ESO
-archive LINK!!! and are promptly reduced by members of our collaboration, often
-using archival calibration data which are readily available. This allows us in
-most cases to report the preliminary results (redshift, identification of the
-most prominent emission and absorption feature, etc.) within a few hours after
-the beginning of the observations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{RRM observations} \label{RRM}
+aid the observation. Raw data are usually available within minutes in the
+\href{http://archive.eso.org/wdb/wdb/eso/xshooter/form}{ESO archive} and are
+promptly reduced by members of our collaboration, often using archival
+calibration data which are readily available. This allows us in most cases to
+report the preliminary results (redshift, identification of the most prominent
+emission and absorption feature, etc.) within a few hours after the beginning of
+the observations and plan additional X-shooter observations if deemed necessary.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Rapid response mode (RRM) observations} \label{RRM}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Under rare circumstances, the use of the ESO rapid response mode (RRM;
-\citealt{Vreeswijk2010}) has been possible. RRM is a system which automatically
-overrides current observations at the telescope. This allows the shortest
-possible delay between the GRB trigger and initiation of observations. This
-allows spectroscopic integration to begin on the rapidly fading, optical
-transient within minutes of the burst. In case of a promptly visible GRB, a
-robotic trigger was sent to the telescope if at the moment of the GCN notice the
-GRB is observable and fulfills the following criteria:
+\citealt{Vreeswijk2010}) has been possible. The RRM is a system to automatically
+override ongoing observations at the telescope. This allows the shortest
+possible delay between the GRB trigger and the initiation of observations, where
+spectroscopic integration of the rapidly fading, optical transient can commence
+within minutes of the burst. One limitation is that no instrument change is
+allowed in RRM, which lowers the number of successful triggers (X-shooter shares the telescope with two other instruments). In case of a promptly visible GRB, a robotic
+trigger is sent to the telescope if at the time of the GCN notice the GRB
+fulfills the following criteria:
 
 \begin{enumerate}
-	\item GRB triggered by \textit{Swift}.
-	\item X-shooter is the currently operating instrument at the telescope.
-	\item GRB promptly visible from Paranal.
-	\item Error radius < 60\arcsec.
+	\item The GRB triggered \textit{Swift} onboard.
+	\item The X-ray position must be available less than 1 hr after the GRB.
+	%    \item X-shooter is the currently operating instrument at the telescope.
+	% DM I removed the line above because we do send the trigger even if X-shooter is not mounted (though, the trigger is rejected, as explained in the text above)
+	%    \item GRB promptly visible from Cerro Paranal.
+	%    \item Error radius < 60\arcsec.
 	%    \item No limit on declination.
 	%    \item No constraint on Moon phase or distance.
-	\item Elevation in the sky $> 22^\circ$ (both at trigger time and 15 min after).
-	\item Sun elevation $< -12^\circ$ (both at trigger time and 15 min after).
-	\item The X-ray position must be available less than 1 hr after the GRB.
+	\item The elevation of the source in the sky is $> 22^\circ$ (both at trigger time and 15 min after).
+	\item The Sun elevation from Paranl is $< -12^\circ$ (both at trigger time and 15 min after).
+	\item Galactic $A_V \leq 1.0$~mag according to the maps of \citet{Schlegel1998}.
 	\item The \swift-circulated tags: KNOWN\_SOURCE, COSMIC\_RAY, DEF\_NOT\_GRB are set to false (PROB\_NOT\_GRB can be true)\footnote{These tags are distributed as part of the GCN notices.
@@
 
-The criteria to trigger RRM are looser than those to define the statistical
-sample. This is because of the expected larger brightness of any GRB counterpart
-minutes after the explosion, and because of the rarity of RRM triggers.
-
-The use of RRM is unique in the sense that it allows for the detailed study of
-the temporal variability of the GRB afterglow due to effect of the GRB itself on
-the sorrounding medium \citep[e.g., see][]{Dessauges-Zavadsky2006,
-	Vreeswijk2007, DElia2009, Vreeswijk2013}
-
-There are nine GRBs that have been observed by us with the RRM mode. One of the
-RRM triggers is outside the statistical sample  and two of the RRM trigger are
-on short GRBs. The fastest response between \swift~GRB trigger and the beginning
-of spectroscopic integration was for GRB~160410A, for which the delay was only
-8.4 minutes. In many cases, unfortunately, we could not use RRM even for promptly visible events, because of the unavailability of X-shooter.
-
-Potentially, the proximity of the telescope (Paranal, Chile) to the South
-Atlantic Anomaly (SSA) affects the rate of RRM triggers \citep{Greiner2011}. BAT
-switches off when the satellite goes through the SSA and that means we have
-fewer GRB triggers immediately observable, compared to the rest of
-the \swift~orbit.
+The criteria to trigger RRM are looser than those that define the statistical
+sample. This is both because of the expected larger brightness of GRB
+counterparts soon after the explosion, and because of the rarity of RRM
+triggers.
+
+The use of RRM is unique as it allows to sample a long logarithmic time span in
+the GRB lifetime, and it exploits the extreme brightness of early afterglows. A
+pivotal exampl is the study of temporal variability of GRB afterglow absorption
+systems due to effect of the GRB itself on the sorrounding medium \citep[e.g.,
+see][]{Dessauges-Zavadsky2006, Vreeswijk2007, DElia2009, Vreeswijk2013}.
+
+There are nine GRBs that have been observed with X-shooter in RRM mode. One of
+the RRM triggers is outside the statistical sample, and two of the RRM triggers
+are on short GRBs. Our fastest response (between the \swift~GRB trigger time and
+the beginning of spectroscopic integration) was for GRB~160410A, for which the
+delay was only 8.4 minutes.
+
+In many cases, unfortunately, we could not use RRM even for promptly visible
+events, e.g.{} because of the unavailability of X-shooter. Potentially, the
+proximity of the telescope (Cerro Paranal, Chile) to the South Atlantic Anomaly
+(SAA) affects the rate of RRM triggers \citep{Greiner2011}. BAT switches off
+when the satellite goes through the SSA and that means we have fewer events
+immediately observable, compared to the rest of the \swift~orbit.
 
@@
 	\centerline{\includegraphics{figures/skymap.pdf}} \caption{Mollweide projection
-	in galactic coordinates of the full sky showing the positions on the sky of the
+	in Galactic coordinates of the full sky showing the positions on the sky of the
 	bursts presented in this work. The equator is the Galactic plane. The empty
@@
 	statistical sample with both X-shooter spectroscopy and a measured redshift.
-	Red outlines of empty stars represent bursts which is observed with X-shooter
-	and has a measured redshift, but is not a part of our statistical sample. The
-	blue dots show the positions of the 61 GRBs of our statistical sample that lack
-	redshift measurements. A red outline is added around the six bursts in our
-	statistical sample which was observed with X-shooter, but did not yield a
-	redshift measurement. Two empty circles with a red outlines indicate bursts
-	outside the statistical sample that was followed up with X-shooter, but without
-	a redshift measurement. The different samples are compared in Sect.
+	Red outlines of empty stars represent bursts which have been observed with
+	X-shooter and has a measured redshift, but is not a part of our statistical
+	sample. The blue dots show the positions of the 61 GRBs of our statistical
+	sample that lack redshift measurements. A red outline is added around the six
+	bursts in our statistical sample which were observed with X-shooter, but did
+	not yield a redshift measurement. Two empty circles with red outlines indicate
+	bursts outside the statistical sample that were followed up with X-shooter, but
+	without a redshift measurement. The different samples are compared in Sect.
 	\ref{results}. The background shows the dust maps presented in
 	\citet{Schlegel1998}. Please note that we removed the background where the
-	sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a white
-	color visible at the equator. The grey-scale bar below indicates the value of
-	$A_V$ on the plot. The dotted lines indicate intervals of 30$^\circ$ in
-	longitude and latitude.} \label{fig:skymap}
+	sample criterion is violated ($A_V > 0.5$ mag) and replaced it with a white
+	background. The grey scale bar below indicates the value of $A_V$ on the plot.
+	The dotted lines indicate intervals of 30$^\circ$ in longitude and latitude.}
+\label{fig:skymap}
 \end{figure*}
@@
 cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
-one of the two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during
+one of two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during
 the duration of this follow-up campaign. The observations have been taken during
-a period of eight years corresponding to the ESO observing periods P84 through P98
-under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
-087.A-0055, 088.A-0051, 089.A-0067, 090.A-0088, 0091.C-0934, 092.A-0124,
-093.A-0069, 094.A-0134, 095.A-0045, 096.A-0079, 097.A-0036, and 098.A-0055(PI:
-Fynbo). These proposals were initiated on Guaranteed Time. We have included a
-few additional bursts, from the programmes 084.D-0265 (PI: Benetti), 091.A-0877
-(PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633 (PI: Greiner), and 095.B-0811
-(PI: Levan). The total collection of spectra represents \textit{all} GRB
-afterglows that have been followed up by X-shooter up to \termdate, which marks
-the end of the XS-GRB legacy follow-up program.
+a period of eight years corresponding to the ESO observing periods P84 through
+P98 under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
+087.A-0055, 088.A-0051, 089.A-0067, 090.A-0088, 092.A-0124, 093.A-0069,
+094.A-0134, 095.A-0045, 096.A-0079, 097.A-0036, and 098.A-0055 (PI: Fynbo) and
+0091.C-0934 (PI: Kaper). These proposals were initiated on Guaranteed Time. We
+have included a few additional bursts, from the programmes 084.D-0265 (PI:
+Benetti), 091.A-0877 (PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633 (PI:
+Greiner), and 095.B-0811 (PI: Levan). The total collection of spectra represents
+\textit{all} GRB afterglows that have been followed up by X-shooter up to
+\termdate, which marks the end of the XS-GRB legacy follow-up program.
 
 The first GRB followed up was GRB~090313 \citep{DeUgartePostigo2010}, observed
-on the 15th of March, 2009, during the commissioning of X-shooter on the third
-Unit Telescope (UT3) of the VLT . The bursts observed during the commissioning
+on the 15th of March, 2009, during the commissioning of X-shooter on the second
+Unit Telescope (UT2) of the VLT . The bursts observed during the commissioning
 or the science verification process (GRB~090313, GRB~090530, GRB~090809,
-GRB~090926) are not a part of the sample we use to investigate the statistical
-properties of GRB afterglows, due to the inhomogeneity of their selection. The
-first burst observed after science verification and X-shooter was moved to UT2,
+GRB~090926A) are not a part of the sample we use to investigate the statistical
+properties of GRB afterglows, due to the different criteria for their selection. The
+first burst observed after science verification and the mounting of X-shooter on UT2,
 was GRB~091018, which is the first burst entering our statistically homogeneous
@@
 X-shooter covers the spectral wavelength region from 300 nm to 2480 nm in a
-single exposure, by separating the light into three separate spectroscopic arms
+single exposure, by splitting the light into three separate spectroscopic arms
 through the use of two dichroics. The ultraviolet blue (UVB) arm covers 300 -
@@
 arm covers 1020 - 2480 nm. For some of the observations, we have applied a
-$K$-band blocking filter, cutting the coverage of the NIR arm at 2100 nm. This
+$K$-band blocking filter, cutting the coverage of the NIR arm at 2100 nm. The
+$K$-band blocking filter is only used after 2012, where it was installed. This
 is done to reduce the amount of scattered background light from the thermal
-infrared. For the majority of observations, a nod-and-shuffle observing scheme
+infrared. For the majority of observations, a nodding observing scheme
 has been employed, with a nodding throw of 5\arcsec. Each nodding observation
-have typically been carried out in a standard ABBA pattern. For some cases,
-conditions during the follow-up, either technical or weather, have necessitated
+has typically been carried out in a standard ABBA pattern. For some cases,
+conditions during the follow-up (either technical or weather), have necessitated
 alterations to this scheme as described in App. \ref{notes}. For RRM triggers, a
@@
 0\farc9, and 0\farc9 for the UVB, VIS, and NIR-arm respectively. This sets a
-lower limit on the delivered resolution of the spectra based on the tabulated
-values of the delivered resolutions, which is 4350, 7450, and 5300 for the UVB,
-VIS and NIR-arm
-respectively\footnote{See \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{this URL} for the nominal resolutions.}.
+lower limit on the delivered resolving power of the spectra based on the
+tabulated values of the delivered resolutions, which is 4350, 7450, and 5300 for
+the UVB, VIS and NIR-arm respectively\footnote{See
+	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{this URL} for the nominal spectral resolutions.}.
 %The slit width sets the width of the atmospheric sky lines and determines the amount of light lost due to the wavelength-dependent seeing PSF extending outside the coverage of the slit, where the width of sky-lines is always set by the slit width whereas both the delivered resolution and the slitloss changes for the better as the seeing PSF drops below the slit width.
@@
 For accurate measurements involving line profiles, knowledge of the precise
-instrumental resolution is required. The resolution becomes better than the
-nominal one, listed above, when the delivered seeing is smaller than the
-projected width of the slit on the sky. We discuss how we determine the
-effective instrumental resolution in Sect. \ref{resolution}.
+instrumental resolution is required. The spectral resolution becomes better than
+the nominal one, when the delivered seeing is smaller than the projected width
+of the slit on the sky. We discuss how we determine the effective instrumental
+resolution in Sect. \ref{resolution}.
 
 Due to a mechanical failure, the atmospheric dispersion corrector (ADC) was
-disabled from 1st of August 2012. Only GRB~100728B was affected by the failing
-ADC prior to disablement, resulting in a lower-than-nominal throughput. To avoid
-chromatic slit losses due to atmospheric dispersion, nearly all subsequent
-observations have been carried out at parallactic angle. A consequence of this,
-is that for all observations following 1st of August 2012, the centroid of the
-trace of the source changes position across the spatial direction of the slit as
-a function of wavelength. This effect has been modeled in the extraction
-procedure, as described in Sect. \ref{extract}.
+disabled from 1st of August 2012 until the end of this program. Only GRB~100728B
+was affected by the failing ADC prior to disablement, resulting in a
+lower-than-nominal throughput. To avoid chromatic slit losses due to atmospheric
+dispersion, nearly all subsequent observations have been carried out at
+parallactic angle. A consequence of this is that for all observations following
+1st of August 2012, the centroid of the trace of the source changes position
+across the spatial direction of the slit as a function of wavelength. This
+effect has been modeled in the extraction procedure, as described in Sect.
+\ref{extract}.
 
@@
 \ref{tab:sample_overview} and plot the positions of all the bursts on the
-celestial sphere in galactic coordinates in Fig. \ref{fig:skymap}. Away from the
+celestial sphere in Galactic coordinates in Fig. \ref{fig:skymap}. Away from the
 central zone of avoidance, due to high Galactic extinction cutoff (marked in
 white), the GRB positions have an isotropic distribution, except in the upper
-left quadrant which cannot be observed from Paranal.
-
-30 percent of the spectra presented here, primarily host observations, have
+left quadrant which cannot be observed from Paranal due to the declination
+constraints of the telescope.
+
+Thirty percent of the spectra presented here, primarily host observations, have
 already been published in \citet{Kruhler2015}. Single bursts have additionally
@@
 
-For a typical observation, each of the exposures in the nodding sequence has
+For a typical observation, each of the exposures in the nodding sequence have
 been reduced as a single observation and then subsequently combined to form a
@@
 
-An additional sky-subtraction is additionally run on all rectified 2D spectra.
+An additional sky-subtraction procedure is run on all rectified 2D spectra.
 This is done to remove residual sky, still present. At each pixel in the
@@
 with a few pixel wide Gaussian filter in the dispersion direction and subtracted
-of the entire 2D image.
-
-In the NIR arm where the background is very bright and there are a high number
-of broad sky-lines, an alternative approach to sky subtraction has been
+from the entire 2D image.
+
+In the NIR arm, where the background is very bright and there are a high number
+of bright sky-lines, an alternative approach to sky subtraction has been
 employed. When there are no contaminating sources in the slit, the sky has been
@@
 extraction, subtracting the two from each other while keeping the WCS static.
-Due to the nod-and-shuffle offsets used between observations, this conserves the
+Due to the nodding offsets used between observations, this conserves the
 source flux while removing the sky at the expense of a decrease in
@@
 during the X-shooter Active Flexure Compensation procedure\footnote{X-shooter
-	User Manual available at \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}.}. The
-remaining offset is corrected by cross-correlating the observed sky spectrum
-with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum,
-estimated as the mode of all flux values, has been subtracted. To get the
-correct seeing point-spread-function (PSF) with which to convolve the synthetic sky, an initial
-refinement of the wavelength solution has been obtained by cross-correlating the
-observed sky with an unconvolved synthetic sky. This preliminary wavelength
-calibration is applied to the observed sky. The synthetic spectrum is then
-convolved with an increasing seeing PSF and the width that minimizes $\chi^2$
-with the updated observed sky is chosen to be the effective sky-PSF. Using the
-synthetic sky with the matched resolution, a final wavelength calibration can
-then be calculated by cross-correlating the observed sky with the correctly
-broadened sky spectrum as a function of a velocity offset. Both a
-multiplicative and an additive offset to the wavelength calibration has been
-tested, but in terms of $\chi^2$, the model with only a multiplicative offset is
-preferred. The resulting offsets, which were smaller than 0.01 nm in the UVB and
-VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short
-periods of time were applied to the corresponding spectra\footnote{The
-	wavelength shifts have been extensively studied by ESO
-	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
+	User Manual available at
+	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}.}.
+The remaining offset is corrected by cross-correlating the observed sky spectrum with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum, estimated as the mode of all flux values, has been subtracted.
+To get the correct seeing point-spread-function (PSF) with which to convolve the synthetic sky, an initial refinement of the wavelength solution has been obtained by cross-correlating the observed sky with an unconvolved synthetic sky.
+This preliminary wavelength calibration is applied to the observed sky.
+The synthetic spectrum is then convolved with an increasing seeing PSF and the width that minimizes $\chi^2$ with the updated observed sky is chosen to be the effective sky-PSF.
+Using the synthetic sky with the matched resolution, a final wavelength calibration can then be calculated by cross-correlating the observed sky with the correctly broadened sky spectrum as a function of a velocity offset.
+Both a multiplicative and an additive offset to the wavelength calibration has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
+The resulting offsets, which were smaller than 0.01 nm in the UVB and VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short periods of time were applied to the corresponding spectra\footnote{The wavelength shifts have been extensively studied by ESO staff \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
 Using the convolved synthetic sky, the pixels containing the brightest sky lines have been flagged as such in the bad pixel map.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!t]
 	\centerline{\includegraphics[width=8cm]{figures/resolutuions}} \caption{Blue
-		datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric
+		data points show the FWHM (km/s) of Gaussian fits to unresolved telluric
 		absorption lines in the VIS spectra, as a function of the airmass corrected
-		ambient FWHM. The red datapoints show a subsample of measurements obtained for
+		DIMM seeing. The red data points show a subsample of measurements obtained for
 		NIR spectra. The colored lines show the corresponding linear fits to the
-		datapoints. The colored bounds contain 68 per cent of the best-fit probability
-		mass. The red and blue dashed line indicates the nominal resolution for the
-		slit-widths used for the telluric standard observations. As can be see, the
-		delivered resolution is in many cases superior to the nominal one. }
+		data points. The colored bounds contain 68 per cent of the best-fit probability
+		mass. The red and blue dashed lines indicate the nominal resolution for the
+		slit-widths used for the telluric standard observations. As can be seen, the
+		effective resolution is in many cases superior to the nominal one.}
 	\label{fig:res}
@@
 The afterglow spectra described in this paper are obtained in
-Target-of-Opportunity (override) mode. In most cases, there is little
-possibility to tweak slit widths to the seeing at the time of observations (i.e.
-to optimize spectral resolution and signal-to-noise), and almost all our data are
-therefore taken with a fixed set of slit widths and binning, described above. In
-a fair number of cases, the seeing full width at half maximum (FWHM) is
-considerably smaller than the slit width, and the delivered spectral resolution
-will then be determined by the seeing rather than by the slit width, as afterglows are
-point sources (this is evidently not the case for extended sources, e.g for host
+Target-of-Opportunity mode. In most cases, there is little possibility to tweak
+slit widths to the seeing at the time of observations (i.e. to optimize spectral
+resolution and signal-to-noise), and almost all our data are therefore taken
+with a fixed set of slit widths and binning, described above. In a fair number
+of cases, the seeing full width at half maximum (FWHM) is considerably smaller
+than the slit width, and the delivered spectral resolution will then be
+determined by the seeing rather than by the slit width, as afterglows are point
+sources (this is evidently not the case for extended sources, e.g for host
 galaxies). The delivered resolution for slit-width dominated spectra
@@
 absorption, other forms of line broadening, and a lack of lines at some
-redshifts, this is difficult to do at lower signal-to-noise ratios (the
-majority of spectra in our sample). A broad starting value for the expected
-resolution will help fitting of these spectra, and can be important in upper
-limit determination, and for this reason we construct a crude relation between
-the seeing and the delivered resolution at our slit width, binning, and
-reduction pipeline settings.
+redshifts, this is difficult to do at lower signal-to-noise ratios (the majority
+of spectra in our sample). A broad starting value for the expected resolution
+will help fitting of these spectra, and can be important in upper limit
+determination, and for this reason we construct a crude relation between the
+seeing and the delivered resolution at our slit width, binning, and reduction
+pipeline settings.
 
@@
 been reduced together with the afterglow spectra, using identical pipeline
-settings with the same version of the pipeline. To get the resolutions, we
-select a series of atmospheric transitions that are resolved multiples which
-should be intrinsically unresolved, and are in areas with well defined continuum
-flux. We then fit the lines with Voigt-profiles and calculate a corresponding
-FWHM, which we can then convert to a delivered resolution using the wavelength
-of the chosen transitions. To get the resolution as a function of ambient
-conditions, for each observations we also calculate the airmass-corrected
-\href{http://www.eso.org/asm/ui/publicLog}{DIMM} seeing, which is measured at
-500 nm.
-
-The resulting distribution of spectral FWHM (km/s) as a function of spatial
-FWHM at 500 nm is fairly well described in the VIS arm by a linear relation $a
-+ b*x$, with $x$ the spatial FWHM in arcseconds, $a= 25.2 \pm 0.8$ km/s, $b=3.8
-\pm 0.7$ (see Fig. \ref{fig:res}). For the NIR arm the corresponding relation
-is $a= 36.0 \pm 1.8$ km/s, $b=8.4 \pm 1.8$. We use these linear relations as a
-way to estimate the spectral resolution for medium to low signal-to-noise
-afterglow spectra. To extend this to the UVB arm, we calculate the ratio
-between the VIS and the NIR arm resolutions and find that the resulting
-distribution is consistent with a simple scaling of the VIS arm relation by the
-ratio of resolutions of the NIR and VIS arm for unresolved, slit filling,
-sources as given on the
+settings with the same version of the pipeline. To get the resolution for each
+observation, we select a series of atmospheric transitions that are resolved
+multiples which should be intrinsically unresolved, and are in areas with well
+defined continuum flux. We then fit the lines with Voigt-profiles and calculate
+a corresponding FWHM, which we can then convert to a delivered resolution using
+the wavelength of the chosen transitions. To get the resolution as a function of
+ambient conditions, for each observations we also calculate the
+airmass-corrected \href{http://www.eso.org/asm/ui/publicLog}{DIMM} seeing, which
+is measured at 500 nm.
+
+The resulting distribution of spectral FWHM (km/s) as a function of spatial FWHM
+at 500 nm is fairly well described in the VIS arm by a linear relation $a +
+b*x$, with $x$ the spatial FWHM in arcseconds, $a= 25.2 \pm 0.8$ km/s, $b=3.8
+\pm 0.7$ (see Fig. \ref{fig:res}). For the NIR arm the corresponding relation is
+$a= 36.0 \pm 1.8$ km/s, $b=8.4 \pm 1.8$. We use these linear relations as a way
+to estimate the spectral resolution for medium to low signal-to-noise afterglow
+spectra. To extend this to the UVB arm, we calculate the ratio between the VIS
+and the NIR arm resolutions and find that the resulting distribution is
+consistent with a simple scaling of the VIS arm relation by the ratio of
+resolutions of the NIR and VIS arm for unresolved, slit filling, sources as
+given on the
 \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{ESO instrument website}.
@@
 This simple analysis gives a sufficiently accurate estimate for the analysis of the low signal-to-noise science spectra.
+In all cases the determined resolution is written to the header with the "RESOLUTION" keyword.
 
@@
 techniques have been employed based on the brightness of the afterglow and the
-complexity of the objects entering the slit. Due to the malfunctioning ADC, see
-Sect. \ref{obs}, the spectral trace changes position across the slit in the
+complexity of the objects entering the slit. Due to the malfunctioning ADC (see
+Sect. \ref{obs}), the spectral trace changes position across the slit in the
 spatial direction as a function of wavelength. For a large fraction of the
@@
 imaging PSF due to atmospheric turbulence, but because of aberrations in the
-dispersion elements and the rectification process, the PSF we are trying to
+optical dispersion elements and the rectification process, the PSF we are trying to
 model is different from this profile. To allow for flexibility in the model, we
@@
 
-For all earth-based telescopes, the light first has to pass through Earths
+For all Earth-based telescopes, the light first has to pass through Earth's
 atmosphere, where the atmospheric content and conditions make an imprint on the
@@
 the atmospheric conditions have directly been imprinted on the spectrum. The
-telluric features can directly be fit with an atmospheric model
-(\texttt{Molecfit} \citep{Smette2015,
-	Kausch2015}\footnote{\url{http://www.eso.org/sci/software/pipelines/skytools/molecfit}}), which can then be used to correct for the absorption.
+telluric features can directly be fit with an atmospheric model \citep[\texttt{Molecfit};][]{Smette2015,
+	Kausch2015}\footnote{\url{http://www.eso.org/sci/software/pipelines/skytools/molecfit}}, which can then be used to correct for the absorption.
 The accuracy of the correction depends on the signal-to-noise per pixel of the target spectrum, where we have chosen the requirement that the afterglow continuum spectrum has a median signal-to-noise higher than a value of 10.
@@
 In the last case, where the object is neither bright enough, or there for some
-reason have not been observed a telluric standard, we rely on a synthetic sky
+reason a telluric standard have not been observed, we rely on a synthetic sky
 model \citep{Noll2012, Jones2013} for which we generate a
@@
 	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
-\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$,
+\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
 	that illustrates the typical data quality. The continuum estimate is shown in
 	dashed red and the error spectrum in solid blue. The acquisition magnitude is
-	$r$ = 20, meaning it is in the brighter end of the sample presented here, but
-	not the brightest. The spectrum is rich in absorption and emission, including
+	$R$ = 20, meaning it is in the brighter end of the sample presented here, but
+	not the brightest. The spectrum is rich in absorption and emission lines, including
 	absorption from molecular $\mathrm{H_2}$. The absorption trough visible at
@@
 	lines seen in GRB afterglows from \citet{Christensen2011a}. The regions of most
-	severe telluric absorption is highlighted by a grey-shading the background.
+	severe telluric absorption are highlighted by grey-shading the background.
 	Additionally, three intervening systems are seen in this sightline. This
@@
 GRB afterglows. The reason for the non-physical model for the normalization is
-that it has the flexibility to captures instrumental variations of the continuum
+that it has the flexibility to capture instrumental variations of the continuum
 level that are not easily contained in a more physically motivated model.
@@
 interpolated using \texttt{scipy.interpolate.splrep.splev}, which serves as a
-first estimate of the continuum placement. To avoid the spline to diverge at the
-edges, the spline-based continuum is tapered with a low order polynomial. An
-attempt to identify absorption and emission regions are then carried out, where
-they are marked as such if the difference between the estimated continuum and
-the observed spectrum is larger than 3 - 5 times the associated error spectrum.
-All regions marked as affected are then masked.
+first estimate of the continuum placement. To prevent the spline from diverging
+at the edges, the spline-based continuum is tapered with a low order polynomial.
+An attempt to identify absorption and emission regions is then carried out,
+where they are marked as such if the difference between the estimated continuum
+and the observed spectrum is larger than 3 - 5 times the associated error
+spectrum. All regions marked as affected are then masked.
 
@@
 All the spectra are made available as a single
-\href{http://www.dark-cosmology.dk/~jselsing/XSGRB}{ZIP file}, through
+\href{https://sid.erda.dk/cgi-sid/ls.py?share_id=DBuNORk1lI}{ZIP file}, through
 \url{http://grbspec.iaa.es}, and additionally through the ESO archive in the
@@
 obtained sample is representative for the full \textit{Swift} sample. An
-important note is that here we include \textit{all} GRBs after \startdate, that
-have been observed with X-shooter, while only a subset of these constitutes our
-\textit{statistical} sample. The statistical sample is based on the selection
-criteria described in Sect. \ref{samplecrit}. Some bursts not fulfilling the
-sample criteria have been followed up due to interesting characteristics, e.g.,
-curios properties of their light curves, their brightness, etc. These bursts
-are not discussed as part of the investigation of the statistical properties of
-the GRB population. A prime example of a spectacular burst outside the
-statistical sample is the bright \textit{INTEGRAL} burst GRB~161023A, that
-contains at least 15 intervening absorption systems (See \ref{161023}).
+important note is that here we provide a release for \textit{all} GRBs after
+\startdate, that have been observed with X-shooter, while only a subset of these
+constitutes our \textit{statistical} sample. The statistical sample is based on
+the selection criteria described in Sect. \ref{samplecrit}. Some bursts not
+fulfilling the sample criteria have been followed up due to interesting
+characteristics, e.g., curios properties of their light curves, their
+brightness, etc. These bursts are not discussed as part of the investigation of
+the statistical properties of the GRB population. A prime example of a
+spectacular burst outside the statistical sample is the bright \textit{INTEGRAL}
+burst GRB~161023A (de Ugarte Postigo et al., in prep), that contains at least 15
+intervening absorption systems (See \ref{161023}).
 
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 	\centerline{\includegraphics[width=9cm]{figures/timing.pdf}} \caption{Afterglow
@@
 	points have been colored based on the redshift of the corresponding burst. Red
-	crosses indicate GRBs without a measured redshift and arrows indicate bursts
+	symbols indicate GRBs without a measured redshift and arrows indicate bursts
 	for which the afterglow was not detected in the acquisition image. In red is
@@
 \end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	observed as part of the statistical sample (red). The left plot shows the
-	fluence in 15-150 keV as observed by BAT. The middle panel shows the 0.3 - 10
+	fluence in the 15-150 keV band as observed by BAT. The middle panel shows the 0.3 - 10
 	keV flux, 11 hours after the bursts as measured by XRT. In the right most
@@
 
-For each of the samples, we calculate the median, 14th, and 86th percentiles of
+For each of the samples, we calculate the median, 16th, and 84th percentiles of
 each of the distributions, which can be used as point estimates for the
@@
 from the same parent distribution, can be rejected. We show a graphic
-representation of the test statistics in Fig. \ref{fig:p_values}. A high p-value
+representation of the test statistics in Fig. \ref{fig:p_values}. A high $p$-value
 indicates little evidence against the null hypothesis . The distribution of
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
-\caption{Relational graph showing the respective p-values. They all represent
+\caption{Relational graph showing the respective $p$-values. They all represent
 	the degree that the different samples are drawn from the same underlying
 	distribution. The arrows represent the comparison direction with each of the
-	p-values they are listed next to. Only in the H{\sc I} column density distribution is
+	$p$-values they are listed next to. Only in the H{\sc I} column density distribution is
 	there mild evidence against the null hypothesis, but the discrepancy is mainly
@@
 
-Out of the 165 bursts in the statistical sample, 36 (22 per cent) were
-not observed due to reasons unrelated to the GRB or afterglow properties. The
-reasons include unavailability of the telescope due to technical maintenance
-(e.g., mirror re-coating), a visiting observer rejecting the ToO trigger, or bad
-weather. Because this cut is unrelated to the GRB properties, it will not change
-the statistical properties of the full sample. Removing these bursts for the
-statistical sample, dramatically improves the redshift completeness from 61 per cent to 88
-per cent. The remaining burst not followed up already had redshift from other
-instruments or were very faint and without a host association thus observations
-unlikely to yield a redshift measurement. In the remainder of the text, we
-consider the 129 bursts our statistical sample.
+Out of the 165 bursts meeting our initial selection criteria for the statistical
+sample, 36 (22 per cent) were not observed due to reasons unrelated to the GRB
+or afterglow properties. The reasons include unavailability of the telescope due
+to technical maintenance (e.g., mirror re-coating), a visiting observer
+rejecting the ToO trigger, or bad weather. Because this cut is unrelated to the
+GRB properties, it will not change the statistical properties of the full
+sample. Removing these bursts from the statistical sample, dramatically improves
+the redshift completeness from 61 per cent to 88 per cent. The remaining burst
+not followed up already had a redshift from other instruments or were very faint
+and without a host association, thus observations were unlikely to yield a redshift
+measurement. In the remainder of the text, we consider the 129 bursts our
+statistical sample.
 
@@
 consistent with being drawn from the same parent sample with the following
-p-values: SHOALS (p-value = 0.13), BAT6 (p-value = 0.95), Fynbo09 (p-value =
-0.08) and TOUGH (p-value = 0.09).
+$p$-values: SHOALS ($p$-value = 0.13), BAT6 ($p$-value = 0.95), Fynbo09
+($p$-value = 0.08) and TOUGH ($p$-value = 0.09). As a small note, the redshift
+distribution of BAT6 are not expected to be identical to the other complete
+samples due to the additional cut on the GRB peak flux in the BAT band.
 
 Because the redshift completeness of our statistical sample is 88 per cent,
-making inference of the \textit{true} redshift distribution of GRBs based on
+making an inference of the \textit{true} redshift distribution of GRBs based on
 this sample is impossible. %The biases introduced in the measured redshift
@@
 \citet{Perley2016b}, bursts for which the redshift is measured from the
-afterglow are systematically found for host galaxies with a lower luminosity
-than burst for which the redshift is measured from the host galaxy. Only a few
+afterglow are systematically found in host galaxies with a lower luminosity
+than bursts for which the redshift is measured from the host galaxy. Only a few
 GRBs hosted in galaxies, with stellar masses more than
@@
 \citep{Groot1998, Djorgovski2001, Fynbo2001}. The degree of optical extinction
-relative to the X-ray brightness has been parametrized in term of their optical
+relative to the X-ray brightness has been parametrized in terms of their optical
 darkness, using the measurement of, or limit on, the optical to X-ray spectral
@@
 \citep{DePasquale2003, Fynbo2009, Melandri2012} and there are some indications
-that dark bursts have somewhat higher X-ray luminosity and $N_{\mathrm{HI, X}}$
-compared to the optically bright bursts \citep{Campana2012}. This indicates
-along with investigations of host galaxy properties \citep{Greiner2011,
-	Kruhler2011, Hjorth2012, Perley2016b}, that the extinction of the optical
-afterglows is primarily driven by the presence of dust in the host galaxies and
-not solely by unfortunate placement of the synchrotron spectral break
-frequencies. \citet{Hjorth2012} find that systems with no optical afterglow have
-higher $N_{\mathrm{HI, X}}$, irrespective of the nature of the host -- which
-however also turn out to be redder. Additionally, the ISM absorption lines in
-dark sight-lines are found to be stronger compared to optically brighter bursts
-\citep{Christensen2011a}, which is consistent with the dark bursts being found
-in more metal-rich and dustier galaxies.
+that dark bursts have somewhat higher X-ray luminosity and $N_{\mathrm{H, X}}$
+compared to the optically bright bursts \citep{Campana2012, Watson2012}. The
+X-ray column density has been shown to be roughly correlated with the gas column
+density, which, for a given metallicity, also correlates with the dust
+extinction \citep{Watson2013, Covino2013a}, though the range in metallicity
+introduces a large additional scatter in the correlation between the extinction
+and the $N_{\mathrm{H, X}}$. This indicates along with investigations of host
+galaxy properties \citep{Greiner2011, Kruhler2011, Hjorth2012, Perley2016b},
+that the extinction of the optical afterglows is primarily driven by the
+presence of dust in the host galaxies and not solely by unfortunate placement of
+the synchrotron spectral break frequencies. \citet{Hjorth2012} find that systems
+with no optical afterglow have higher $N_{\mathrm{H, X}}$, irrespective of the
+nature of the host -- which however also turn out to be redder. Additionally,
+the ISM absorption lines in dark sight-lines are found to be stronger compared
+to optically brighter bursts \citep{Christensen2011a}, which is consistent with
+the dark bursts being found in more metal-rich and dustier galaxies.
 
@@
 simultaneous measurement of the X-ray flux density and the optical flux density
-which is in practice is possible, but in reality extremely rarely available. As
+which is in practice possible, but in reality extremely rarely available. As
 a proxy, we use the measured acquisition camera magnitude reported in Table
@@
 observation. This is done by either linearly interpolating between temporally
-neighboring XRT measurements of by extrapolating the last few X-ray data points
+neighboring XRT measurements or by extrapolating the last few X-ray data points
 to the time of the spectroscopic observation. When the afterglow is not detected
@@
 
-In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{HI, X}}$
+In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{H, X}}$
 distribution with the one presented in \citet{Fynbo2009}. We take the
@@
 test, however, suggest a statistically significant, low degree of negative
-correlation ($\Gamma = -0.21$ at a p-value = 0.01).
+correlation ($\Gamma = -0.21$ at a $p$-value = 0.01).
 
 \begin{figure}[!ht]
-	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}} \caption{Optical
-	extinction relative to X-ray against X-ray derived hydrogen column densities.
-	In red is shown the sample presented in \citet{Fynbo2009} where the lines
-	indicate the kernel density estimate of the distribution. For the kernel
-	density estimate, the limits have been replaced with values. We note that
-	contrary to what is presented in \citet{Fynbo2009}, we exclude all bursts for
-	which we do not have a redshift. Darker colors represented higher density of
-	points. In blue are the points for the bursts presented here along with the
-	marginal histograms. Limits on $\beta_{OX}$ are shown by downwards facing
-	arrows. The corresponding marginal distributions are shown along the edges of
-	the plot. There is no statistically significant correlation between the two
-	quantities.} \label{fig:betaOX}
+	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
+\caption{$\beta_{OX}$ against X-ray derived hydrogen column density. In red is
+	shown the sample presented in \citet{Fynbo2009} where the lines indicate the
+	kernel density estimate of the distribution. For the kernel density estimate,
+	the limits have been replaced with values. We note that contrary to what is
+	presented in \citet{Fynbo2009}, we exclude all bursts for which we do not have
+	a redshift. Darker colors represent a higher density of points. In blue are the
+	points for the bursts presented here along with the marginal histograms. Limits
+	on $\beta_{OX}$ are shown by downwards facing arrows. The corresponding
+	marginal distributions are shown along the edges of the plot. A Kendall's
+	$\tau$ test, indicates a statistically significant, low degree of negative
+	correlation ($\Gamma = -0.21$ at a $p$-value = 0.01)} \label{fig:betaOX}
 \end{figure}
@@
 
-However, the fraction of dark bursts for which we have measured redshifts for is
-lower than the for the ones with a detected optical afterglow. For the
-afterglows we have observed as part of the statistical sample that do not have a
-detected optical afterglow, the redshift completeness is 53 per cent. For
-comparison, for the afterglows in the statistical sample we have observed with
-an optical afterglow detected, the redshift completeness is 92 per cent. This
-also goes to show that the lack of redshift completeness in the sample presented
-here is in part due to the increased difficulty of measuring a redshift for
-bursts without an afterglow. To measure a redshift, we either need a detected
-afterglow to obtain a spectrum or to locate the host galaxy and determine the
-redshift from there. It is more difficult to correctly associate a galaxy with a
-burst when there is no detected optical afterglow and hence a correct redshift
-measurement is more difficult to make, see \citet{Jakobsson2005, Levesque2010}
-and \citet{Perley2017}.
+However, the fraction of dark bursts for which we have measured redshifts is
+lower than the ones with a detected optical afterglow. For the afterglows we
+have observed as part of the statistical sample that do not have a detected
+optical afterglow, the redshift completeness is 53 per cent. For comparison, for
+the afterglows in the statistical sample we have observed with an optical
+afterglow detected, the redshift completeness is 92 per cent. This also shows
+that the lack of redshift completeness in the sample presented here is in part
+due to the increased difficulty of measuring a redshift for bursts without an
+afterglow. To measure a redshift, we either need a detected afterglow to obtain
+a spectrum or to locate the host galaxy and determine the redshift from there.
+It is more difficult to correctly associate a galaxy with a burst when there is
+no detected optical afterglow and hence a correct redshift measurement is more
+difficult to make, see \citet{Jakobsson2005, Levesque2010} and
+\citet{Perley2017}.
 
@@
 brightness on the density of the circumburst medium \citep{Freedman2001,
-	Berger2003, Nysewander2009}, if the measured $N_{\mathrm{HI, X}}$ is primarily
-driven by a metal column local to the burst. Because we only use values for
-$N_{\mathrm{HI, X}}$ in the comparison for which the GRB has a measured
-redshift, this measure is likely biased toward optically brighter bursts.
+	Berger2003, Nysewander2009}, if the measured $N_{\mathrm{H, X}}$ is primarily
+driven by the gas column in the neighborhood of the burst. Because we only use
+values for $N_{\mathrm{H, X}}$ in the comparison for which the GRB has a
+measured redshift, this measure is likely biased toward optically brighter
+bursts \citep[e.g.][]{Watson2012}.
 
@@
 \lya. The exceptionally good UV response of X-shooter, allows us to robustly
-measure H{\sc i}-columns - also at these very low redshifts.
+measure H{\sc i}-columns - also at these relatively low redshifts.
 
@@
 We show the two distributions in  Fig. \ref{fig:NH_dist}. We compare the median,
-the 14th, and 86th percentiles of the two distributions. The sample presented
+the 16th, and 84th percentiles of the two distributions. The sample presented
 has a \nh~$= 21.8_{-0.8}^{+0.3}$ and the rest of the literature values has
@@
 slightly higher median value for the new sample presented here. A 2-sided KS
-test gives a p-value of p = 0.006, meaning relatively large evidence against the
-null that the two samples are drawn from the same underlying distribution.
+test gives a $p$-value of p = 0.006, meaning relatively strong evidence against the
+null hypothesis that the two samples are drawn from the same underlying distribution.
 Because the bursts that have measurements of the hydrogen column density are
@@
 
-In Fig. \ref{fig:NH_dist}, we also show the column density-distribution
+In Fig. \ref{fig:NH_dist}, we also show the column density distribution
 for the 12081 quasar absorbers with \nh~$> 20$ from \citet{Noterdaeme2012b}. The
-fact that GRBs systematically are located behind the highest \nh, previously
+fact that GRBs are systematically located behind the highest \nh, previously
 noted \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!t]
 	\centering \includegraphics[width=9cm]{figures/NH_dist.pdf}
@@
 spectroscopic observations of afterglows and host galaxies of GRBs detected by
-\swift. This work was initiated by a consortium using Guaranteed Time on
-X-shooter, but over the years the project continued in open time.
+\swift. This work was initiated by the consortium that built X-shooter and
+included this project as a key part of the GTO program, but over the years the
+project continued in open time.
 
@@
 that was also much advanced by the previous surveys based primarily on
-lower-resolution spectrographs. GRB afterglow sight-lines are unique in the
+lower-resolution spectroscopy. GRB afterglow sight-lines are unique in the
 sense that only after observing more than 12000 damped Lyman-$\alpha$ absorbers
-(DLAs) towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ been
+(DLAs) towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ have been
 identified \citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra,
-by contrast, reveal such systems in the majority of cases \citep[][and this
+by contrast, reveal such systems routinely \citep[][and this
 work]{Jakobsson2006b, Fynbo2009, Cucchiara2015}. With afterglow spectroscopy
@@
 of our sample such as extinction curves (\citealt{Japelj2015}, Zafar et al.
-submitted, see also \citealt{Fynbo2014, Heintz2017}]), emission lines from the
+submitted, see also \citealt{Fynbo2014, Heintz2017a}), emission lines from the
 underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
@@
 	Perley2016a}. \cite{Kruhler2013} argue, that very rich sightlines like that
-probed by the remarkable GRB~080607 \citep{Prochaska2009} are probably
-significantly more frequent than in the sightlines sampled by our spectra.
-However, with current instrumentation, these sightlines are out of reach except
-under very fortunate circumstances as in the case of GRB~080607 when the
-afterglow could be observed only a few minutes after the burst with a 10-m class
-telescope. Observations of such sightlines with X-shooter-like spectrographs on
-the next generation of 20--40-m telescopes is likely to be very rewarding, given
-that a suitable GRB detector will be available.
+probed by the remarkable GRB~080607 \citep{Prochaska2009, Sheffer2009, Perley2011} are
+probably significantly more frequent than in the sightlines sampled by our
+spectra. However, with current instrumentation, these sightlines are out of
+reach except under very fortunate circumstances as in the case of GRB~080607
+when the afterglow could be observed only a few minutes after the burst with a
+10-m class telescope. Observations of such sightlines with X-shooter-like
+spectrographs on the next generation of 20--40-m telescopes is likely to be very
+rewarding, given that a suitable GRB detector will be available.
 
@@
 We owe him a large debt for his invaluable work with the \swift~satellite.
-We also want to acknowledge Javier Gorosabel, who was taken from us too early.
+We also want to remember Javier Gorosabel, who was taken from us too early.
 %
 JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.
-The Dark Cosmology Centre were funded by the Danish National Research
+The Dark Cosmology Centre was funded by the Danish National Research
 Foundation. 
@@
 %
-JJ acknowledges
+LK and JJ acknowledges
 support from NOVA and NWO-FAPESP grant for advanced instrumentation in
@@
 %
-AK acknowledges support from the Spanish research project AYA 2014-58381-P and
+DAK acknowledges support from the Spanish research project AYA 2014-58381-P and
 support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2015- 26153.
@@
 %
+GL is supported by a research grant (19054) from VILLUM FONDEN.
+%
+SDV acknowledges the support of the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro
+%
+DM acknowledges support from the Instrument Center for Danish Astrophysics (IDA).
+%
 This work made use of data supplied by the UK {\it Swift} Science Data Centre at
@@
 
-\bibliographystyle{mnras}
+\bibliographystyle{aa}
 \bibliography{XSGRB_sample,GCN}
@@
 	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
-	\caption{Measurements of the hydrogen column-densities for all bursts with a
+	\caption{Measurements of the hydrogen column densities for all bursts with a
 		clear Lyman alpha absorption system. In solid black is shown the spectrum with
-		black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
+		black dotted giving the corresponding 1-$\sigma$ error. The black dashed line shows zero
 		flux density. The solid red line is the absorption of column density equal to
@@
 W(z)  
-& = \frac{i}{\pi} \int_{-\infty}^{\infty} \frac{e^{-t^2}}{z - t}  
+& = \frac{i}{\pi} \int_{-\infty}^{\infty} \frac{e^{-t^2}}{z - t}  dt
 \end{split}
@@
 The first GRB ever observed with X-shooter, during the commissioning of the
-instrument, this data formed the basis of GCN 9015 \citep{GCN.9015} and is
+instrument, these data formed the basis of GCN 9015 \citep{GCN.9015} and are
 published in \citet{DeUgartePostigo2010}. Due to the lingering brightness of
@@
 
-Observed during paranalization of the instrument, this data forms the basis of
-GCN 15571 \citep{GCN15571}, but is not published elsewhere. Observations began
+Observed during paranalization of the instrument, these data forms the basis of
+GCN 15571 \citep{GCN15571}, but are not published elsewhere. Observations began
 20.6 hours after the BAT trigger and 4.8 ks spectroscopic integration in all
@@
 detection of \mgii, \mgi, \SIii, \feii, \aliii. Because this burst was observed
-before the instrument is science-verified, it does not enter into the
+before the instrument was science-verified, it does not enter into the
 statistical sample.
@@
 
-\subsection{GRB~090926 (z = 2.106)} \label{090926}
+\subsection{GRB~090926A (z = 2.106)} \label{090926}
 
@@
 \SIii$^*$ \feii, \mgii, all at $z = 2.106$. Because this burst was observed
-before the instrument is science-verified, it does not enter into the
+before the instrument was science-verified, it does not enter into the
 statistical sample.
@@
 The first burst observed during normal operation after science verification was
-completed and the first burst that enters the statistical sample. This data is
-the basis for GCN 10042 \citep{GCN10042} and is
-published in \citet{Wiersema2012}. With a bright afterglow and a rapid
-follow-up, this spectrum is of pristine quality. The afterglow continuum is
-bright throughout all spectroscopic arms which allows the ready detection of
-\alii, \aliii, \feii, \mnii, \mgii, \mgi, and \caii~- al located at $z = 0.971$,
-setting is as the redshift of the host.
+completed and the first burst that enters the statistical sample. These data are
+the basis for GCN 10042 \citep{GCN10042} and is published in
+\citet{Wiersema2012}. With a bright afterglow and a rapid follow-up, this
+spectrum is of pristine quality. The afterglow continuum is bright throughout
+all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
+\mnii, \mgii, \mgi, and \caii~- all located at $z = 0.971$, setting it as the
+redshift of the GRB.
 
@@
 
-Obtained 4 days after the burst trigger, this data forms the basis for GCN 10233
-\citep{GCN10233} and is published in \citet{Vergani2011}. Due to the late
+Obtained 4 days after the burst trigger, these data forms the basis for GCN 10233
+\citep{GCN10233} and are published in \citet{Vergani2011}. Due to the late
 follow-up and a nearby moon, the signal-to-noise of the afterglow continuum is
-low especially in the UVB arm, why no clear absorption lines are detected
+low, especially in the UVB arm. This is why no clear absorption lines are detected
 against the afterglow continuum, although see \citet{Vergani2011} which report a
-tentative detection of \mgii. Emission lines from the underlying host is clearly
+tentative detection of \mgii. Emission lines from the underlying host are clearly
 visible with lines from \oii, \hb, \oiii, and \ha~all at $z = 0.490$. This
-bursts is additionally associated with SN2009nz.
+bursts is additionally associated with SN2009nz \citep{Cobb2010, Berger2011, Olivares2015}.
 
@@
 Observed 3 days after the \textit{Swift} trigger. No afterglow or host detected
-in 10.8 ks. GRB likely located at high
-redshift \citep[GCN GCN10399;][]{GCN10399}. The spectrum
-has not otherwise been published previously.
+is in 10.8 ks. The GRB is likely located at high redshift \citep[GCN
+GCN10399;][]{GCN10399}. The spectrum has not otherwise been published
+previously.
 
@@
 
-The data presented here also formed the basis of GCN \#
-10441 \citep{GCN10441} and is published
-in \citet{Thone2013}. Observations started 12.5 hours after the \textit{Swift}
-trigger and has a total exposure time of 4.8 ks. Absorption features, including
-those of \lya~and from a multitude of ions are detected against the afterglow
-continuum at $z = 4.667$. Additionally, absorption from an intervening system
-is found at $z = 2.181$.
-
-\subsection{GRB~100316B (z = 1.180)} \label{100316}
-
-The data presented here also formed the basis of GCN 10495 \citep{GCN10495}. The spectrum
-has not otherwise been published. Observations started 44 minutes
-after the \textit{Swift} trigger and has a total exposure time of 2.4 ks.
-Absorption features from \feii, \alii, \aliii,	\mgii~and \mgi~are well detected
-against the afterglow continuum at $z = 1.180$. Additionally, strong absorption
-lines from \feii~and \mgii~from an intervening system are found at $z = 1.063$.
-
-\subsection{GRB~100316D (z = 0.059)} \label{100316}
+These data presented here have also formed the basis of GCN 10441 \citep{GCN10441} and
+are published in \citet{Thone2013}. Observations started 12.5 hours after the
+\textit{Swift} trigger and has a total exposure time of 4.8 ks. Absorption
+features, including those of \lya~and from a multitude of ions are detected
+against the afterglow continuum at $z = 4.667$. Additionally, absorption from an
+intervening system is found at $z = 2.181$.
+
+\subsection{GRB~100316B (z = 1.180)} \label{100316B}
+
+The data presented here also formed the basis of GCN 10495 \citep{GCN10495}. The
+spectrum has not otherwise been published. Observations started 44 minutes after
+the \textit{Swift} trigger and have a total exposure time of 2.4 ks. Absorption
+features from \feii, \alii, \aliii,	\mgii~and \mgi~are well detected against the
+afterglow continuum at $z = 1.180$. Additionally, strong absorption lines from
+\feii~and \mgii~from an intervening system are found at $z = 1.063$.
+
+\subsection{GRB~100316D (z = 0.059)} \label{100316D}
 
 The data presented here also formed the basis of GCN 10512 \citep{GCN10512}, GCN
-\# 10513 \citep{GCN10513}, GCN 10543 \citep{GCN10543} and is published in
+10513 \citep{GCN10513}, GCN 10543 \citep{GCN10543} and are published in
 \citet{Bufano2012} and \citet{Starling2011}. This GRB is very close by and has
-an associated SN, SN2010bh, and has therefore undergone intense follow-up. The
-data presented here consists of a subset of the entire VLT/X-shooter campaign,
-covering the four first observing days while the afterglow still contributes
-significantly to the total emission. The first observations started 10 hours
-after the burst, before the SN was discovered, and targeted the star-forming
-'A'-region \citep{Starling2011}, not the GRB. A very rich spectrum containing a
-multitude of emission lines puts the host at $z = 0.059$. For three consequtive
-nights, 58, 79 and 101 hours after the \textit{Swift} trigger, the afterglow was
-observed as it transitioned into the spectrum of a high-velocity Ic-BL SN. The
-observations taken 79 and 101 hours after the burst are taken under programme
-084.D-0265(A) (PI: Benetti), but with an identical setup to the first two
-observations.
+an associated SN, SN2010bh, and has therefore undergone intense follow-up
+\citep{Olivares2011, Cano2011a, Izzo2017}. The spectra presented here consists
+of a subset of the entire VLT/X-shooter campaign, covering the four first
+observing days while the afterglow still contributes significantly to the total
+emission. The first observations started 10 hours after the burst, before the SN
+was discovered, and targeted the star-forming 'A'-region \citep{Starling2011},
+not the GRB. A very rich spectrum containing a multitude of emission lines puts
+the host at $z = 0.059$. For three consecutive nights, 58, 79 and 101 hours
+after the \textit{Swift} trigger, the afterglow was observed as it transitioned
+into the spectrum of a high-velocity Ic-BL SN. The observations taken 79 and 101
+hours after the burst are taken under programme 084.D-0265(A) (PI: Benetti), but
+with an identical setup to the first two observations.
 
@@
 
-The data presented here also formed the basis of GCN 10620 \citep{GCN10620} and GCN 10631 \citep{GCN10631} and is published
-in \citet{DeUgartePostigo2011}. The burst have been followed up in three epochs
-of observations, 0.4, 1.4, and 2.4 days after the burst, each lasting 4.8 ks.
-The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous
-detection of emission features belonging to nebular lines, including \hi, \oii,
-\oiii, \neiii, \nii, [\sii], [\siii], and \hei~as well as absorption features due to
-the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all
-at a consistent redshift. Temporal evolution of the fine structure lines
-belonging to \feii$^*$ is found between the epochs.
+The data presented here also formed the basis of GCN 10620 \citep{GCN10620} and
+GCN 10631 \citep{GCN10631} and are published in \citet{DeUgartePostigo2011}. The
+burst has been followed up in three epochs of observations, 0.4, 1.4, and 2.4
+days after the burst, each lasting 4.8 ks. The unambiguous redshift of the host,
+$z=0.624$, is found from the simultaneous detection of emission features
+belonging to nebular lines, including \hi, \oii, \oiii, \neiii, \nii, [\sii],
+[\siii], and \hei~as well as absorption features due to the presence of \znii,
+\crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all at a consistent
+redshift. Temporal evolution of the fine structure lines belonging to \feii$^*$
+is found between the epochs.
 
@@
 14291 \citep{GCN14291}. The spectrum has
-not otherwise been published. Observations carried out, long after the burst
+not otherwise been published. Observations were carried out long after the burst
 afterglow had faded. Emission lines from the host are detected at $z=2.465$.
@@
 
+The spectra presented here also formed the basis of GCN 10684 \citep{GCN10684}
+and are used in Skuladottir (2010), but not published elsewhere. Observations
+started 4 hours after the \textit{Swift} trigger, totaling 2.4 ks. Absorption
+features from \mgii~and \feii~in the afterglow continuum are detected at
+$z=1.1755$.
+
+\subsection{GRB~100615A (z=1.398)} \label{100615}
+
 The data presented here also formed the basis of GCN 
-10684 \citep{GCN10684} and is used in
-Skuladottir (2010), but not published elsewhere. Observations started 4 hours
-after the \textit{Swift} trigger, totaling 2.4 ks. Absorption features from
-\mgii~and \feii~in the afterglow continuum are detected at $z=1.1755$.
-
-\subsection{GRB~100615A (z=1.398)} \label{100615}
+14264 \citep{GCN14264}, but are not
+published elsewhere. Host observation of a dark burst \citep{DElia2011} taken
+long after the afterglow had faded. Emission lines from the host belonging to
+\oii, \neiii, \oiii~and \ha~are detected at a common redshift of $z=1.398$.
+
+\subsection{GRB~100621A (z=0.542)} \label{100621}
 
 The data presented here also formed the basis of GCN 
-14264 \citep{GCN14264}, but not
-published elsewhere. Host observation of a dark burst \citep{DElia2011} taken
-long after the afterglow has faded. Emission lines from the host belonging to
-\oii, \neiii, \oiii~and \ha~are detected at a common redshift of $z=1.398$.
-
-\subsection{GRB~100621A (z=0.542)} \label{100621}
-
-The data presented here also formed the basis of GCN 
-10876 \citep{GCN10876}, but not
+10876 \citep{GCN10876}, but are not
 published elsewhere. Beginning 7.1 hours after the GRB, 2.4 ks observations
@@
 The data presented here is of the candidate host galaxy, taken long after the
-burst has faded and have not previously been published. 4.8 ks of exposure
+burst had faded and have not previously been published. 4.8 ks of exposure
 reveals a weak continuum present in all arms, but an absence of emission lines.
@@
 are detected in the afterglow continuum at a common redshift of $z = 1.288$.
-This is not a part of the statistical sample.
 
@@
 is confirmed in the 2D image and despite the observational challenges that
-affects this burst, we measure \nh~$=21.2 \pm 0.15$.
+affects this spectrum, we measure \nh~$=21.2 \pm 0.15$.
 
@@
 The spectra presented here has not been published previously. The observations
-consists of three visits, the first beginning only 0.9 hours after the
+consist of three visits, the first beginning only 0.9 hours after the
 \textit{Swift} trigger, the other two visits were 2.13 and 98.40 hours after the
@@
 at $z=1.439$. A complex velocity structure in the absorption features belonging
-to \mgii, shows several components, separated by as much as 500km/s, pointing to
+to \mgii, shows several components, separated by as much as 500 km/s, pointing to
 a likely merger scenario in the host or starburst driven outflows.
@@
 
-The data presented here has been published in \citet{Hartoog2013}. Because of
+The data presented here have been published in \citet{Hartoog2013}. Because of
 the unusual lingering brightness of this GRB, 2.4 ks of observations taken 65.98
@@
 
-\subsection{GRB~101219A (z=0.718)} \label{101219}
-
-This data has not been published before. Starting 3.7 hours after the GRB
+\subsection{GRB~101219A (z=0.718)} \label{101219A}
+
+These data have not been published before. Starting 3.7 hours after the GRB
 trigger, 7.2 ks of exposure time reveal a very faint continuum in the visual and
@@
 
-\subsection{GRB~101219B (z=0.552)} \label{101219}
+\subsection{GRB~101219B (z=0.552)} \label{101219B}
 
 The data presented here also formed the basis of GCN 11579 \citep{GCN11579} and
-is published in \citet{Sparre2011}.	The first observation, taken 11.6 hours
-after the burst trigger and lasting 4.8ks, reveals absorption from \mgii~and
+are published in \citet{Sparre2011}. The first observation, taken 11.6 hours
+after the burst trigger and lasting 4.8 ks, reveals absorption from \mgii~and
 \mgi~in the host located at $z = 0.552$ on a featureless continuum visible
@@
 been published before. Spectroscopic integration started 6.55 hours after the
-\swift~trigger and lasted for a total of 7.2ks. The afterglow continuum is
+\swift~trigger and lasted for a total of 7.2 ks. The afterglow continuum is
 detected across the entire spectral coverage at moderate signal-to-noise.
-Absorption lines in the continuum is detected from \lya, \oi, \cii, \SIiv, \civ,
+Absorption lines in the continuum are detected from \lya, \oi, \cii, \SIiv, \civ,
 \SIii~and \feii, all at a common redshift of $z=2.339$. From the broad
@@
 coverage, both at a consistent redshift, $z=2.109$, securing it as the redshift
-of the GRB.
+of the GRB host.
 
@@
 sample, but nonetheless followed up due to the extremely high peak energy
-\citep{Axelsson2012}. Starting 28.7 hours after the burst trigger, 2.4ks
+\citep{Axelsson2012}. Starting 28.7 hours after the burst trigger, 2.4 ks
 spectroscopic observation reveals after heavy binning, a wide, faint trace down
@@
 No good redshift measurement can be inferred from this. We have adopted the
-redshift from GCN  12193 \citep{GCN12193}
+redshift from GCN 12193 \citep{GCN12193}.
 
@@
 not published otherwise. Starting 3 hours after the \swift~trigger, a rich
-spectrum is obtained in 2.4ks spectroscopic integration. The GRB afterglow
+spectrum is obtained in 2.4 ks spectroscopic integration. The GRB afterglow
 continuum is visible across all three spectrscopic arms of VLT/X-shooter with
-additionally, emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$.
-At the same redshift, we identify absorption lines superposed on the afterglow
-continuum from \mgii~and \feii.
+emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$. At the same
+redshift, we identify absorption lines superposed on the afterglow continuum
+from \mgii~and \feii.
 
@@
 
-Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8ks
+Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8 ks
 reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 500 nm. The simultaneous detection of absorption features identified as \lya,
 \SIii, \civ, \alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet,
-securely sets $z = 3.36$ as the redshift of the GRB. This data forms the basis
+securely sets $z = 3.36$ as the redshift of the GRB. These data form the basis
 of GCN 12284 \citep{GCN12284}, but is not published elsewhere.
@@
 
-The data presented here has previously been published in \citet{Michaowski2016}.
-2.4ks spectroscopic integration of the host galaxy, obtained long after the
+The data presented here have previously been published in \citet{Michaowski2016}.
+2.4 ks spectroscopic integration of the host galaxy, obtained long after the
 burst had faded, contains bright emission lines filling the entire slit on top
@@
 
-This data formed the basis of GCN 12431 \citep{GCN12431} and is additionally
-published in \citet{Sparre2014}. Observations of this GRB afterglow was
-initiated 7.71 hours after the BAT trigger and had a duration of 8.4ks. A second
-observational epoch started 20.1 hours and lasted for 6.6ks. The GRB afterglow
-continuum is well detected down to $\sim$760 nm, with several strong
-absorption features imprinted, all at a common $z = 4.990$. \lya~is clearly
-detected and we additionally detect lines identified as \SIii, \feii, \civ,
-\mgii, \SIii*, \sii*, \oi*. An intervening DLA systems is additionally detected
-at  $z = 4.61$ as seen from \lya~and \mgii~absorption.
+These data formed the basis of GCN 12431 \citep{GCN12431} and are additionally
+published in \citet{Sparre2014}. Observations of this GRB afterglow were
+initiated 7.71 hours after the BAT trigger and had a duration of 8.4 ks. A second
+observational epoch started 20.1 hours after the GRB trigger and lasted for
+6.6 ks. The GRB afterglow continuum is well detected down to $\sim$760 nm, with
+several strong absorption features imprinted, all at a common $z = 4.990$.
+\lya~is clearly detected and we additionally detect lines identified as \SIii,
+\feii, \civ, \mgii, \SIii*, \sii*, \oi*. An intervening DLA system is
+additionally detected at  $z = 4.61$ as seen from \lya~and \mgii~absorption.
 
@@
 
-This data has previously been used to form some of the basis of
-\citet{Selsing2017}. Starting 37.3 hours after the BAT trigger, 4.8ks of
+These data have previously been used to form some of the basis of
+\citet{Selsing2017}. Starting 37.3 hours after the BAT trigger, 4.8 ks of
 spectroscopic integration yields faint emission lines identified at \oii, \hb,
 \oiii~and \ha, all at a common $z = 2.211$, marking it as the redshift of the
-GRB. No afterglow continuum is detected.
+GRB host. No afterglow continuum is detected. This is the highest redshift short
+GRB detected to date.
 
@@
 
-This data formed the basis of GCN 14273 \citep{GCN14273}, but is not published
+These data formed the basis of GCN 14273 \citep{GCN14273}, but are not published
 elsewhere. Observed twice, the first time shortly after the GRB and the second
@@
 
-Starting 8.26 hours after and lasting 3.6ks, these observations have previously
+Starting 8.26 hours after the GRB trigger and lasting 3.6 ks, these observations have previously
 been published in \citet{Kruhler2015}. A very faint continuum is visible after
@@
 \citep{GCN12648}. The first epoch of spectroscopic observations was initiated
-17.7 hours after the BAT trigger and lasted for 4.8 seconds. A very bright
+17.7 hours after the BAT trigger and lasted for 4.8 ks. A very bright
 afterglow continuum is detected across the entire spectral coverage of
 X-shooter, with several absorption features imprinted. The absorption features
-are identified as \feii, \mgii, \mgi, \cah, and \cak - all at a common redshift
+are identified as \feii, \mgii, \mgi, \cah, and \cak~- all at a common redshift
 of $z = 0.677$. The second epoch, taken 20 days later, still contains a faint
@@
 
-These data has formed the basis for GCN 12677 \citep{GCN12677} and is also
+These data formed the basis for GCN 12677 \citep{GCN12677} and are also
 published in \cite{Kruhler2015}. Observations began 31 hours after the AGILE
@@
 which we suggest is the redshift of the GRB. Additionally detected in the GRB
-afterglow continuum is broad undulation, suggesting an accompanying SN.
+afterglow continuum are broad undulation, suggesting an accompanying SN.
 
@@
 
-This data formed the basis of GCN 12770 \citep{GCN12770} and is also published
+These data formed the basis of GCN 12770 \citep{GCN12770} and are also published
 in \cite{Kruhler2015}. Observations began 15.9 hours after the BAT trigger and
@@
 $z = 0.716$. Supporting this redshift measurement as the redshift of the GRB is
-the detection of nebular emission line from \oiii.
+the detection of nebular emission from \oiii.
 
@@
 The data presented here also formed the basis of GCN 14225 \citep{GCN14225}, but
-is not published otherwise. This late-time observation of the host of
+are not published otherwise. This late-time observation of the host of
 GRB~120118B consists of 3.6 ks exposures and contains emission lines belonging
@@
 
-The data presented here has been examined by \citet{Japelj2015} and
-Zafar et al. (submitted), who both find a significant amount of extinction, $A_V\approx
-1$ mag. Three epochs of observations have been obtained, the first two
-immediately after the burst, and the last one long after the afterglow had
-faded. Starting 1.4 hours after the \textit{Swift} trigger, the first epoch
-contains bright afterglow continuum. Rich in absorption features belonging to a
-multitude of ions at $z =    1.728$ is estimated for the host with intervening
-systems at $z = 1.476$, $z = 1.214$, $z = 0.662$ and $z = 0.632$. The second
-epoch, obtained 4.5 hours after the burst contains the fading afterglow. A third
-epoch is obtained $>1$ year after the GRB in which emission lines from \hb~and
-\ha~are found at the redshift of the host, confirming the association of the
-absorption line system and the host. We also detect C\,\textsc{i} in absorption
-which indicates the presence of cold gas.
+The data presented here have been examined by \citet{Japelj2015} and Zafar et al.
+(submitted), who both find a significant amount of extinction, $A_V\approx 1$
+mag, also supported by previous measurements \citep{Morgan2014a}. Three epochs
+of observations have been obtained, the first two immediately after the burst,
+and the last one long after the afterglow had faded. Starting 1.4 hours after
+the \textit{Swift} trigger, the first epoch contains bright afterglow continuum.
+Rich in absorption features belonging to a multitude of ions at $z = 1.728$ is
+estimated for the host with intervening systems at $z = 1.476$, $z = 1.214$, $z
+= 0.662$ and $z = 0.632$. The second epoch, obtained 4.5 hours after the burst
+contains the fading afterglow. A third epoch is obtained $>1$ year after the GRB
+in which emission lines from \hb~and \ha~are found at the redshift of the host,
+confirming the association of the absorption line system and the host. We also
+detect C\,\textsc{i} in absorption which indicates the presence of cold gas.
 
@@
 
-The data presented here has been published in \citet{Kruhler2015}. Two
+The data presented here have been published in \citet{Kruhler2015}. Two
 observations of the host of GRB~120211A has been obtained, starting 2013.02.17,
@@
 
-The data presented here has formed the basis of GCN 12991 \citep{GCN12991}, and
-has also been published in \citet{Kruhler2015}. Starting 19.8 hours after the
+The data presented here have formed the basis of GCN 12991 \citep{GCN12991}, and
+have also been published in \citet{Kruhler2015}. Starting 19.8 hours after the
 GRB trigger, a total exposure time of 2.4 ks reveals a faint continuum, starting
@@
 
-The data presented here has formed the basis of GCN 12991 \citep{GCN12991}, but
-is not published otherwise. Starting just before twilight, 3.65 hours after the
+The data presented here have formed the basis of GCN 12991 \citep{GCN12991}, but
+are not published otherwise. Starting just before twilight, 3.65 hours after the
 burst, a faint afterglow continuum is detected at all wavelengths. Due to the
@@
 plane of 6 kpc, posing a potential problem for the host redshift, unless the GRB
-ocurred in a merging system. The extended emission in \ha, supports this
+occurred in a merging system. The extended emission in \ha, supports this
 interpretation. This burst is not a part of the statistical sample.
@@
 The data presented here also formed the basis of GCN 13134 \citep{GCN13134} and
-is published in \citet{DElia2014}. The observation consists of two visits, 2.13
-hrs and 29.98 hrs after the burst, with an afteglow continuum visible in all
+are published in \citet{DElia2014}. The observation consists of two visits, 2.13
+hrs and 29.98 hrs after the burst, with an afterglow continuum visible in all
 arms for both visits. We detect absorption features from Ly-limit, \lya,
@@
 
-The data presented here has formed the basis of GCN 13227 \citep{GCN13227}, but
-is not published otherwise. 9.6 ks integration, starting 15.7 hours after the
+The data presented here have formed the basis of GCN 13227 \citep{GCN13227}, but
+are not published otherwise. 9.6 ks integration, starting 15.7 hours after the
 \textit{Swift}-trigger reveals a low-intensity afterglow continuum on which
@@
 The data presented here also formed the basis of GCN 13257 \citep{GCN13257} and
-is published in \citet{Schulze2014}. A GRB-SN, this burst has been followed up
-multiple times. The data presented here only contains the first epoch in which
+are published in \citet{Schulze2014}. Being a GRB-SN, this burst has been followed up
+multiple times. The data presented here only contain the first epoch in which
 the afterglow is still visible and before the rise of SN2012bz. Starting 16.5
@@
 The data presented here also formed the basis of GCN 13460 \citep{GCN13460}, but
-is not published elsewhere. 4.8 ks integration time, starting 10.5 hours after
+are not published elsewhere. 4.8 ks integration time, starting 10.5 hours after
 the BAT trigger, shows a bright afterglow continuum starting at $\sim$ 472 nm,
@@
 
-The data presented here also formed the basis of GCN 13477 \citep{GCN13477}, but
-is not published elsewhere. Observations of this burst started 7.8 hours after
-the GRB trigger, lasting 4.8 ks. A continuum is visible across the entire
-spectral coverage of X-shooter, with both emission lines from  \oii, \hb,
-\oiii~and \ha, as well as absorption from \mgii~detected at $z = 0.398$,
-securely setting it as the redshift of the GRB.
+The data presented here also formed the basis of GCN 13477 \citep{GCN13477} and
+are discussed in Klose et al. (submitted), but are not published anywhere.
+Observations of this burst started 7.8 hours after the GRB trigger, lasting 4.8
+ks. A continuum is visible across the entire spectral coverage of X-shooter,
+with both emission lines from  \oii, \hb, \oiii~and \ha, as well as absorption
+from \mgii~detected at $z = 0.398$, securely setting it as the redshift of the
+GRB.
 
@@
 The data presented here also formed the basis of GCN 13494 \citep{GCN13494}, but
-is not published elsewhere. Despite observations starting 62 hours after the
+are not published elsewhere. Despite observations starting 62 hours after the
 \textit{Swift} trigger and lasting 3.6 ks, a bright afterglow is clearly seen,
 along with a plethora of absorption features. Absorption of \lya-photons in the
-host leaves a broad trough, from which the Lyman alpha forest is visible
+host leaves a broad trough, beyond which the Lyman alpha forest is visible
 bluewards, all the way down to the Lyman limit. Metal absorption lines from
@@
 The data presented here also formed the basis of GCN 13507 \citep{GCN13507}, but
-is not published elsewhere. In 4.8 ks integration time, starting 10 hours after
+are not published elsewhere. In 4.8 ks integration time, starting 10 hours after
 the burst trigger, the simultaneous detection of absorption features belonging
@@
 A separate reduction of this burst has been published in \citet{Kruhler2015},
-but is not used otherwise. Starting 9 days after the burst trigger, this is host
+but is not used otherwise. Starting 9 days after the burst trigger, this is a host
 observation and does not contain any afterglow continuum. In 3.6 ks integration
@@
 absence of nebular lines is due to \oii~falling in a telluric absorption band
-and the rest of the nebular lines being shifted out of the wavelength coverage.
+and the rest of the strong nebular lines being shifted out of the wavelength coverage.
 
@@
 from \nv, \sii, \SIii, \oi, \civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and
-\mgi. Additionally fine structure lines from \NIii~and \feii~are exited local to
-the GRB. Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z =
-2.00$.
+\mgi. Also fine structure lines from \NIii~and \feii~are excited in the local
+environment of the GRB. Additionally, this spectrum is one of the rare cases in
+which we detect lines from molecular hydrogen, \h2. From the \lya-line we
+measure \nh=$22.1\pm0.10$. Intervening systems are found at $z = 1.539$, $z =
+1.693$, and $z = 2.00$.
 
@@
 
-The data presented here has formed the basis of GCN 13730 \citep{GCN13730}, but
-is not published otherwise. A very rapid follow-up, starting only 1.7 hours
-after the BAT trigger, this 1.2 ks observation captures a very bright afterglow
-continuum, starting at 450 nm, signifying the onset of the Lyman limit for a
-system at $z = 3.929$. Absorption from high-column density hydrogen leaves very
-prominent absorption features in the form of \lya, \lyb, and \lyg, visible in
-the Lyman alpha forest. Metal absorption lines arising from \feii, \NIii, \SIii,
-\sii, \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
-corresponding fine structure lines from (\feii$^*$, \SIii$^*$, \oi$^*$,
-\oi$^**$, \cii$^*$), securely anchoring the redshift of the host.
+The data presented here have formed the basis of GCN 13730 \citep{GCN13730}, but
+are not published otherwise. Follow-up, started only 1.7 hours after the BAT
+trigger. This 1.2 ks observation captures a very bright afterglow continuum,
+starting at 450 nm, signifying the onset of the Lyman limit for a system at $z =
+3.929$. Absorption from high-column density hydrogen leaves very prominent
+absorption features in the form of \lya, \lyb, and \lyg, visible in the Lyman
+alpha forest. Metal absorption lines arising from \feii, \NIii, \SIii, \sii,
+\alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
+corresponding fine structure lines (\feii$^*$, \SIii$^*$, \oi$^*$, \oi$^**$,
+\cii$^*$), securely anchoring the redshift of the host.
 
@@
 The data presented here also formed the basis of GCN 13890 \citep{GCN13890} and
-is published in \citet{Friis2015}. Also rapid, starting 1.8 hours after the
+are published in \citet{Friis2015}. Starting 1.8 hours after the
 \textit{Swift} trigger, a bright afterglow continuum is visible across all arms.
@@
 
-The data presented here has formed the basis of GCN 13930 \citep{GCN13930}, but
-is not published otherwise. Starting 69.6 hours after the GRB trigger, we detect
+The data presented here have formed the basis of GCN 13930 \citep{GCN13930}, but
+are not published otherwise. Starting 69.6 hours after the GRB trigger, we detect
 the afterglow continuum at high significance in all arms with 8.4 ks
@@
 
-These data formed the basis for GCN 14035 \citep{GCN14035} and is additionally
+These data formed the basis for GCN 14035 \citep{GCN14035} and are additionally
 published in \citet{Kruhler2015}. These observations started 12.9 hours after
-the \swift~trigger and consists of 4.8ks spectroscopic integration under good
+the \swift~trigger and consist of 4.8 ks spectroscopic integration under good
 atmospheric conditions. The GRB afterglow continuum is well detected at all
-arms. A broad absorption trough due to \lya is visible at $z = 3.385$, which
+arms. A broad absorption trough due to \lya~is visible at $z = 3.385$, which
 along with the detection of absorption features identified as \SIiv, \civ,
@@
 
-These data formed the basis for GCN 14120 \citep{GCN14120}, but is not published
-elsewhere. Taken under poor seeing conditions, a total of 4.8ks spectroscopic
-integration starting 2 hours after the \swift~trigger yields a low
+These data formed the basis for GCN 14120 \citep{GCN14120}, but are not
+published elsewhere. Taken under poor seeing conditions, a total of 4.8 ks
+spectroscopic integration starting 2 hours after the \swift~trigger yields a low
 signal-to-noise GRB afterglow continuum is all arms. Binning the spectrum
@@
 $\log (N_{\mathrm{HI}}/\mathrm{cm}^{-2}) = 21.7 \pm 0.2$. Due to strong
-contamination in the slit, the background is slightly over subtracted, cause the
-background to go negative in the center of the \lya~trough.
+contamination in the slit, the background is slightly over subtracted, causing
+the background to be negative in the center of the \lya~trough.
 
@@
 
-These data formed the basis for GCN 14286 \citep{GCN14286} and is additionally
+These data formed the basis for GCN 14286 \citep{GCN14286} and are additionally
 published in \citet{Kruhler2015}. This is a late-time observation, taken long
-after the GRB afterglow has faded. In 7.2ks spectroscopic integration, emission
+after the GRB afterglow had faded. In 7.2 ks spectroscopic integration, emission
 lines identified as \oii~and \oiii~are detected at a common $z = 2.539$, which
@@
 spectrum has not otherwise been published previously. The observations consists
-of two 600sec spectra taken 1.9hrs after the burst. We detect absorption
+of two 600 s spectra taken 1.9 hours after the burst. We detect absorption
 features from a wide range of ions. We also detect intervening absorption at
@@
 GCN 14390 \citep{GCN14390} is based on this spectrum, but it is not published
-elsewhere. Starting only 4.57 hours after the \swift~trigger, 1.2ks observations
-contains a bright GRB afterglow continuum, visible across the entire spectral
+elsewhere. Starting only 4.57 hours after the \swift~trigger, 1.2 ks observations
+contain a bright GRB afterglow continuum, visible across the entire spectral
 coverage of X-shooter. Superposed on the afterglow continuum are absorption
@@
 
-\subsection{GRB~130427A (z = 0.340)} \label{130427}
+\subsection{GRB~130427A (z = 0.340)} \label{130427A}
 
@@
 additionally has formed the basis for GCN 14491 \citep{GCN14491}. Starting 16.5
-hours after the BAT trigger, these observations lasting $2 \times 600$ s
-contains a very bright GRB afterglow continuum across the total spectral
-coverage of X-shooter. In absorption we identify features from the following
-metal resonance lines: \feii, \mnii, \mgii, \mgi, \TIii~and additional line
-absorption from \caii~and \nai. Simultaneously, we find emission lines from \ha,
-\hb, \oiii, \oii - all at common redshift of $z = 0.340$, which is the redshift
-of the GRB. This is one of the most energetic GRBs observed, and it's proximity
-along with the associated broad-lined Type Ic SN, 2013cq has caused it to be one
-of the more well studies GRBs.
-
-\subsection{GRB~130427B (z = 2.780)} \label{130427}
-
-This spectrum formed the basis of GCN 14493 \citep{GCN14493}, but is not
+hours after the BAT trigger, these observations lasting $2 \times 600$ s contain
+a very bright GRB afterglow continuum across the total spectral coverage of
+X-shooter. In absorption we identify features from the following metal resonance
+lines: \feii, \mnii, \mgii, \mgi, \TIii~and additional line absorption from
+\caii~and \nai. Simultaneously, we find emission lines from \ha, \hb, \oiii,
+\oii - all at common redshift of $z = 0.340$, which is the redshift of the GRB.
+This is one of the most energetic GRBs observed, and its proximity along with
+the associated broad-lined Type Ic SN, 2013cq has caused it to be one of the
+more well-studied GRBs \citep{Maselli2014, Perley2014, Ackermann2014}.
+
+\subsection{GRB~130427B (z = 2.780)} \label{130427B}
+
+This spectrum formed the basis of GCN 14493 \citep{GCN14493}, but are not
 published otherwise. A short, 2x600 s spectroscopic integration obtained before
@@
 visible across the entire spectral coverage at low signal-to-noise. Due to the
-low signal-to-noise, the metal lines are weak, but the broad absorption though
-due to \lya~is detected. From this we measure the redshift to be $z = 2.780$ and
-provides a measure of the neutral hydrogen column density, \nh~$= 21.9 \pm 0.3$.
-The redshift is confirmed by the presence of \feii~at a consistent redshift.
+low signal-to-noise, the metal lines are weak, but the broad absorption trough
+due to \lya~is detected. From the \lya~line we measure the redshift to be $z =
+2.780$ and provide a measure of the neutral hydrogen column density, \nh~$= 21.9
+\pm 0.3$. The redshift is confirmed by the presence of \feii~absorption at a
+consistent redshift.
 
@@
 
-This burst is the first sGRB observed with a potential associated
-kilonova \citep{Tanvir2013a, Berger2013a}.  GCN 14757 \citep{GCN14757} was based
-on this spectrum, but it is not published elsewhere. Starting 8.2 hours after
-the \swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
-Continuum is clearly detected across all arms from both host and afterglow and
-superposed are both absorption (\cahk~and \mgii) and emission lines (\oii, \hb,
-\oiii, \ha, and [\sii]), all at a consistent redshift of $z = 0.356$, which is the
-redshift of the GRB.
+This burst is the first short GRB observed with a potential associated kilonova
+\citep{Tanvir2013a, Berger2013a}.  GCN 14757 \citep{GCN14757} was based on this
+spectrum and it is additionally published in \citet{DeUgartePostigo2014e}.
+Starting 8.2 hours after the \swift~trigger, a total of 2.4 ks spectroscopic
+integration was obtained. Spectral continuum is clearly detected across all arms
+from both host and afterglow and superposed are both absorption (\cahk~and
+\mgii) and emission lines (\oii, \hb, \oiii, \ha, and [\sii]), all at a
+consistent redshift of $z = 0.356$, which is the redshift of the GRB.
 
@@
 The data presented here also formed the basis of GCN 14816 \citep{GCN14816} and
-is published in \citet{Hartoog2015}. The observations consists of three
-$2\times600$ s visits starting 7.1 hrs after the burst at fairly high airmass.
+are published in \citet{Hartoog2015}. The observations consist of three
+$2\times600$ s visits starting 7.1 hours after the burst at fairly high airmass.
 We detect absorption features from a wide range of ions at $z=5.913$ as well as
@@
 
-The spectral features of this spectrum has previously been reported in GCN 14882
-\citep{GCN14882}, but is not published elsewhere. Starting only 1.1 hour after
+The spectral features of this spectrum have previously been reported in GCN 14882
+\citep{GCN14882}, but are not published elsewhere. Starting only 1.1 hours after
 the \swift~trigger, $2\times600$ s spectroscopic integration captures a moderate
@@
 resonance lines \feii, \mnii, \mgii, \mgi~are identified. Additionally, \lya~is
-visible as a broad absorption trough, from we can infer \nh~= $22.2 \pm 0.2$,
+visible as a broad absorption trough, from which we can infer \nh~= $22.2 \pm 0.2$,
 which is in the upper end of the hydrogen column density distribution. The blue
@@
 background, a faint afterglow trace is visible across all arms of X-shooter,
-down to ~ 480 nm, which if interpreted as the break due to \lya suggests $z =
+down to ~ 480 nm, which if interpreted as the break due to \lya~suggests $z =
 2.9$. This supports the approximate redshift suggested in GCN 14898
@@
 
-These data formed the basis for GCN 14956 \citep{GCN14956} and is additionally
+These data formed the basis for GCN 14956 \citep{GCN14956} and are additionally
 published in \citet{Kruhler2015}. Starting 5.5 hours after the GRB trigger,
@@
 \swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
-afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
-primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
-\hb, \oii, \ha, \nii, [\sii]) are well detected at $z = 0.347$, as well as those
-from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
-spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).
+afterglow \citep[A$_\mathrm{V} = 5.9 \pm 0.7$;][]{Greiner2014}, with the
+spectrum primarily dominated by host emission lines. All the nebular lines
+(\oii, \hg, \hb, \oii, \ha, \nii, [\sii]) are well detected at $z = 0.347$, as
+well as those from \pad, \pag, and \pab. We take this as the redshift of the
+GRB. This spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).
 
@@
 
-This data formed the basis for 15330 \citep{GCN15330}, but is not published
+These data formed the basis for GCN 15330 \citep{GCN15330}, but are not published
 elsewhere. Starting $\sim$1.5 days after the Fermi-GBM trigger, 4.5 ks
@@
 
-This data has not been published before. Starting 3.4 hours after the
-\swift~trigger, $6\times600$ s exposure taken under good conditions, contains a
-bright GRB afterglow continuum across the entire spectral coverage of X-shooter.
-A myriad of absorption features are superposed on the afterglow continuum which
-we identify as being caused by \SIiv, \SIii, \civ, \alii, \aliii, \znii, \crii,
-\NIii, \feii, \NIii$^*$, and \feii$^*$ at $z = 1.296$. A very strong
-\mgii-absorber is also detected, intervening the line-of-sight at $z = 1.164$
-with lines from \SIii, \civ, \aliii, \aliii, \feii, \mnii, and many more.
+These data has not been published before. Starting 3.4 hours after the
+\swift~trigger, $6\times600$ s exposure were taken under good conditions,
+containing a bright GRB afterglow continuum across the entire spectral coverage
+of X-shooter. A myriad of absorption features are superposed on the afterglow
+continuum which we identify as being caused by \SIiv, \SIii, \civ, \alii,
+\aliii, \znii, \crii, \NIii, \feii, \NIii$^*$, and \feii$^*$ at $z = 1.296$. A
+very strong \mgii-absorber is also detected, intervening the line-of-sight at $z
+= 1.164$ with lines from \SIii, \civ, \aliii, \aliii, \feii, \mnii, and many
+more.
 
@@
 absorption features identifed as due to \feii~and \mgii~as well as emission
-lines from \oii, \hd, \hg, \hb, \oiii, \ha, and \nii. All the lines are measured
+lines from \oii, \hd, \hg, \hb, \oiii, \ha, and \niil. All the lines are measured
 at a consistent redshift of $z = 0.599$, which we take as the redshift of the
@@
 
-This spectrum has already been used in shaping GCN 15450 \citep{GCN15450} and is
+This spectrum has already been used in GCN 15450 \citep{GCN15450} and is
 additionally published in \citet{Kruhler2015}. Starting only 1.3 hours after the
@@
 published in a refereed paper. Starting only 68 minutes after the BAT trigger,
-4.8 ks spectroscopic integration secures afterglow continuum for this bursts,
+4.8 ks spectroscopic integration secures afterglow continuum for this burst,
 which is measured to be at $z = 4.042$. A moderate signal-to-noise GRB afterglow
@@
 $4\times600$ s exposures. A high signal-to-noise GRB afterglow continuum is
-detected all the way though the X-shooter arms. We identify absorption features
-imprinted in the continuum as caused by \feii, \mgii, and \cahk at a consistent
+detected all the way through the X-shooter arms. We identify absorption features
+imprinted in the continuum as caused by \feii, \mgii, and \cahk~at a consistent
 $z = 0.642$. By subtracting off the bright afterglow continuum, we readily
 detect emission lines arising from \oii, \hg, \hb, \oiii, and \ha~in the GRB
-host galaxy.
+host galaxy. This GRB is the subject of a forthcoming work Kann et al. (in
+prep).
 
@@
 redshift measurement. By heavily binning the spectrum, a faint trace is visible
-down to 485 nm, which if interpreted at the onset of \lya signifies z $\sim$
-3. We adopt the redshift inferred in \citet{Kruhler2015}.
+down to 485 nm, which if interpreted at the onset of \lya~signifies $z \sim
+3$. We adopt the redshift inferred in \citet{Kruhler2015}.
 
@@
 The data presented here also formed the basis of GCN 15831 \citep{GCN15831} and
-is additionally published in \citet{Kruhler2015}. Starting 5.8 after the
+are additionally published in \citet{Kruhler2015}. Starting 5.8 after the
 \swift~trigger alert, $2\times600$ s spectroscopic integration contains a high
@@
 
-These data formed the basis for GCN 15900 \citep{GCN15900} and is additionally
+These data formed the basis for GCN 15900 \citep{GCN15900} and are additionally
 published in \citet{Kruhler2015}. Spectroscopic follow-up began 9 hours after
@@
 time in the UVB and VIS arm slightly. The GRB afterglow continuum is clearly
-visible in the VIS and NIR arm of X-shooter. The continuum is a very rich
-absorption, with at least the following lines identified: \lyg, \lyb, \lya,
-\SIii, \SIiv, \civ, \alii, \aliii, \feii, \mgii, and \mgi. All of these lines
-are a $z = 4.954$, which we take as the redshift of the GRB. This spectrum is
-taken under ESO programme ID: 092.D-0633(E) (PI: Greiner).
+visible in the VIS and NIR arm of X-shooter. The continuum is very rich in
+absorption features, with at least the following lines identified: \lyg, \lyb,
+\lya, \SIii, \SIiv, \civ, \alii, \aliii, \feii, \mgii, and \mgi. All of these
+lines are at $z = 4.954$, which we take as the redshift of the GRB. This spectrum
+is taken under ESO programme ID: 092.D-0633(E) (PI: Greiner).
 
@@
 molecular absorption from CH+, all at $z=0.889$. The optical/near-infrared
-afterglow reveal an unusual steep extinction curve which is found to be caused
+afterglow reveals an unusual steep extinction curve which is found to be caused
 by dust very close to the burst.
@@
 
-This spectrum has previously been used in \citet{Melandri2015}, but is not
-published elsewhere. Starting 15.5 hours after the \swift~trigger, $8\times600$
-s spectroscopic integration captures this very high redshift GRB afterglow. The
-Gunn-Peterson trough is visible against the GRB afterglow continuum, starting at
-890 nm, which along with absorption lines from the \mgii~doublet securely sets
-the redshift of this GRB at $z = 6.327$. From the red wing of the \lya-profile
-we measure \nh~$=19.0 \pm 0.5$, which is very low compared to the measured
-distribution of $N_{\mathrm{HI}}$.
+This spectrum has previously been used in \citet{Melandri2015}. Starting 15.5
+hours after the \swift~trigger, $8\times600$ s spectroscopic integration
+captures this very high redshift GRB afterglow. The Gunn-Peterson trough is
+visible against the GRB afterglow continuum, starting at 890 nm, which along
+with absorption lines from the \mgii~doublet securely sets the redshift of this
+GRB at $z = 6.327$. From the red wing of the \lya-profile we measure \nh~$=19.0
+\pm 0.5$, which is very low compared to the measured distribution of
+$N_{\mathrm{HI}}$.
 
@@
 
-The characteristics of this short (T90 (15-350 keV ) = is $0.13 \pm 0.04$ s
-\citep{Lien2016}) GRB spectrum has been published in GCN 16437 \citep{GCN16437},
+The characteristics of this short \citep[T90 = $0.13 \pm 0.04$
+s;][]{Lien2016} GRB spectrum has been published in GCN 16437 \citep{GCN16437},
 but does not appear in the refereed literature. Spectroscopic observations began
@@
 
-This spectrum has already been used to inform GCN 16983 \citep{GCN16983}, but is
+This spectrum has already been used in GCN 16983 \citep{GCN16983}, but is
 not published elsewhere. Starting 15.4 hours after the \swift~trigger and
-lasting for a total of 2.4 ks, these observation nicely captures a bright GRB
+lasting for a total of 2.4 ks, these observations captures a bright GRB
 afterglow. Very rich in absorption, the continuum is detected across the entire
@@
 of $4\times600$ s spectroscopic integration. In the 2D spectrum are two sources,
-which are both offset from the attempted host and are thus likely foreground
+which are both offset from the targeted host position and are thus likely foreground
 objects. The spectrum does not contain anything that can be used to measure a
@@
 but it is not used otherwise. Starting only 1.9 hours after the BAT trigger,
-these $4\times600$ s spectra contains a high signal-to-noise GRB afterglow. The
+these $4\times600$ s spectra contain a high signal-to-noise GRB afterglow. The
 afterglow continuum is readily detected all the way down to 370 nm, although
@@
 \feii, \feii$^*$, \oi$^*$, and \nii$^*$. These are all detected at a consistent
-$z = 2.993$, which due to the detection of the locally exited fine-structure
+$z = 2.993$, which due to the detection of the locally excited fine-structure
 lines securely sets this as the redshift of the GRB. Additionally, two
@@
 
-This spectrum has already been used to inform GCN 17523 \citep{GCN17523}, but it
+This spectrum has already been used in GCN 17523 \citep{GCN17523}, but it
 is not used elsewhere. Starting 5.1 hours after the BAT trigger, this spectrum
@@
 after the GRB trigger, captures a bright GRB afterglow. A broad absorption
-trough centered at $\sim$ 370 nm due to \lya signals the redshift of the GRB,
+trough centered at $\sim$ 370 nm due to \lya~signals the redshift of the GRB,
 which is refined to $z = 2.057$ based on the additional detection of metal
@@
 \SIii, \SIii$^*$, \cii, \cii$^*$, \civ, \alii, \feii, \feii$^*$, \mni, \mgii,
-and \mgi~in the host of the GRB. An intervening \civ absorber is additionally
+and \mgi~in the host of the GRB. An intervening \civ~absorber is additionally
 detected at $z=1.76$. From the absorption trough due to \lya, we infer the
@@
 17755 \citep{GCN17755}, but is not published as part of any refereed paper. This
-bona-fide short GRB (T90 (15-350 keV ) = is $ 0.22 \pm 0.03$ s \citep{Lien2016})
-was observed in RRM mode and spectroscopic integration started after only 22
+bona-fide short GRB \citep[T90 = is $ 0.22 \pm 0.03$ s;][]{Lien2016} was
+observed in RRM mode and spectroscopic integration started after only 22
 minutes. A series of stare mode observations, increasing in exposure time, and
-ending with a nodding sequence totals $\sim$ 5000 s are combined to form this
-spectrum. A faint, almost featureless continuum is detected at low
+ending with a nodding sequence, totaling $\sim$ 5000 s, are combined to form
+this spectrum. A faint, almost featureless continuum is detected at low
 signal-to-noise all the way to the bluest part of the spectrum. An absorption
@@
 This spectrum is empty, but is included here for completeness. It is not
-published anywhere. $4\times600$ s spectroscopic observations, starting 3.7
+published anywhere. $4\times600$ s spectroscopic observation, starting 3.7
 hours after the trigger does not reveal anything conclusive. The host
-association is additionally ambiguous for this, likely extincted (GCN 17767
-\citep{GCN17767}), GRB.
+association is additionally ambiguous for this, likely extincted \citep[GCN
+17767;][]{GCN17767}, GRB.
 
@@
 This spectrum has already been used in GCN 17822 \citep{GCN17822}. Spectroscopic
-observations began 28.4 post trigger and consist of a $4 \times 600$ s nodding
-sequence. The GRB afterglow continuum is detected at moderate signal-to-noise
-across the entire spectral coverage of X-shooter. Narrow absorption features are
-imprinted in the continuum where we identify features from multiple
-\feii~transitions as well as the \mgii-doublet and the \mgi$\lambda2852.96$
-resonance line. These lines are all found a position matching $z = 0.807$,
-suggesting it is the redshift of the GRB.
+observations began 28.4 hours post trigger and consist of a $4 \times 600$ s
+nodding sequence. The GRB afterglow continuum is detected at moderate
+signal-to-noise across the entire spectral coverage of X-shooter. Narrow
+absorption features are imprinted in the continuum where we identify features
+from multiple \feii~transitions as well as the \mgii-doublet and the
+\mgi$\lambda2852.96$ resonance line. These lines are all found at a position
+matching $z = 0.807$, suggesting it is the redshift of the GRB.
 
@@
 to what degree the GRB afterglow contributes to the continuum. We identify the
-emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and [\sii]~- all at $z =
+emission lines as \oii, \hb, \oiii, \ha, \niil, \ha, and [\sii]~- all at $z =
 0.256$. Due to the spatial proximity of this burst, it is a candidate for SN
 follow-up and indeed there are indications of re-brightening at the burst
-position (GCN 17903 \citep{GCN17903}).
+position \citep[GCN 17903;][]{GCN17903}.
 
@@
 continuum is detected, but emission lines from \oii, \oiii, and \ha~are all
-detected at $z = 1.188$, setting it as the redshift of the GRB.
+detected at $z = 1.188$, setting it as the redshift of the GRB host.
 
@@
 18080 \citep{GCN18080}, but is not published. Starting 5 hours after the
-\swift~GRB trigger, these observations consists of a combined $2 \times 1200$ s
-and $2 \times 600$ s. A blue continuum is detected across the entire spectral
+\swift~GRB trigger, these observations consist of a combined $2 \times 1200$ s
+and $2 \times 600$ s integration. A blue continuum is detected across the entire spectral
 coverage of X-shooter, suggesting a significant contribution from the GRB
@@
 afterglow continuum. The relative strength of the lines suggest that the
-line-forming region is dust obscured, contrary to the story told be the blue
+line-forming region is dust obscured, contrary to the story told by the blue
 afterglow continuum.
@@
 
-This gross content of this spectacular spectrum has already been issued in GCN
+The gross content of this spectacular spectrum has already been issued in GCN
 18187 \citep{GCN18187}, but the spectrum has not been published. Starting just
@@
 at a consistent redshift of $z = 0.755$. The detection of fine-structure lines,
-exited local to the burst, clearly marks this as the redshift of the GRB.
+excited local to the burst, clearly marks this as the redshift of the GRB.
 
@@
 
-The gross content of these spectra have already been issue in GCN 18318
+The gross content of these spectra have already been issued in GCN 18318
 \citep{GCN18318}. Observations began 3.3 hours after the \swift~trigger and
 lasted for $4 \times 1200$ s. A moderate signal-to-noise afterglow continuum is
-detected across the entire spectral coverage of X-shooter, where however the
+detected across the entire spectral coverage of X-shooter, where the
 bluest part is affected by \lya-forest absorption. Imprinted on the afterglow
 continuum are a wealth of both emission and absorption features, all caused in a
-system, for which we measure a redshift of $z = 1.968$. We identify in emission
-\oii, \hb, \oiii, and \ha and absorption due to \lya, \civ, \alii, \SIii, \feii,
+system for which we measure a redshift of $z = 1.968$. We identify in emission
+\oii, \hb, \oiii, and \ha~and absorption due to \lya, \civ, \alii, \SIii, \feii,
 and \mgii. We additionally identify fine-structure absorption lines from
@@
 The data presented here also formed the basis of GCN 18426 \citep{GCN18982} and
-is not published elsewhere. The observation was carried out in RRM starting 44
+are not published elsewhere. The observation was carried out in RRM starting 44
 minutes after the GRB trigger. We detect absorption features from a wide range
@@
 presence of fine-structure lines at a consistent redshift securely sets this as
-the redshift of the GRB. GRB~151027B was the 1000th GRB detected by \swift.
+the redshift of the GRB. GRB~151027B was the 1000th GRB detected by \swift. The
+afterglow of this burst is the subject of a forthcoming publication
+\citep{Greiner2018}.
 
@@
 600$ s observations, extending into the morning twilight, were obtained of this
-GRB. The brightening sky affect the signal-to-noise of these spectra -
+GRB. The brightening sky affects the signal-to-noise of these spectra -
 especially in the blue end. The GRB afterglow is well detected over the entire
-spectral coverage of X-shooter, with spectral feaures from \feii~and
-\mgii~imprinted by an absorber at $z = 1.423$. The absorber er likely the host
+spectral coverage of X-shooter, with spectral features from \feii~and
+\mgii~imprinted by an absorber at $z = 1.423$. The absorber is likely the host
 of the GRB and we therefore consider $z = 1.423$ the redshift of the GRB.
@@
 spectral coverage of X-shooter. Superposed on the continuum are absorption and
-emission features, which is identified as \feii, \mgii, \mgi~in absorption, and
-\oii, \hb, and \oiii. These spectral position of these lines all corrspond to a
+emission features, which are identified as \feii, \mgii, \mgi~in absorption, and
+\oii, \hb, and \oiii. The spectral positions of these lines all correspond to a
 redshift of $z = 0.870$, which most likely is the redshift of the GRB.
 
-\subsection{GRB~160203A (z = 3.517)}\label{160203}
+\subsection{GRB~160203A (z = 3.518)}\label{160203}
 
 The data presented here also formed the basis of GCN 18982 \citep{GCN18982} and
-is not published elsewhere. The observation was carried out in RRM starting 18
+are not published elsewhere. The observation was carried out in RRM starting 18
 minutes after the GRB trigger. We detect absorption features from a wide range
-of ions at $z=3.517$ as well as intervening absorption at $z=2.203$. From the
+of ions at $z=3.518$ as well as intervening absorption at $z=2.203$. From the
 very high signal-to-noise spectrum we are able to measure \nh~$=21.75 \pm 0.10$,
@@
 \citep{GCN19186}. Taken long after the burst had faded, the target is the likely
-host galaxy (See GCN 19186 \citep{GCN19186}). In $4\times1200$ s spectroscopic
-integration we detected emission lines from the host. We find \oiii~and \ha~at a
-consistent $z = 1.640$.  We therefore suggest that this is the redshift of the
-GRB.
+host galaxy. In $4\times1200$ s spectroscopic integration we detected emission
+lines from the host. We find \oiii~and \ha~at a consistent $z = 1.640$.  We
+therefore suggest that this is the redshift of the GRB.
 
@@
 
-This is potentially a short burst with extended emission (GCN 19148
-\citep{GCN19148}). The content of the spectra have previously been reported in
+This is potentially a short burst with extended emission \citep[GCN
+19148;][]{GCN19148}. The content of the spectra have previously been reported in
 GCN 19154 \citep{GCN19154}. Observations began 19.1 hours after the BAT trigger
-and consists of $4\times 1200$ s spectroscopic integration. A very faint
+and consist of $4 \times 1200$ s spectroscopic integration. A very faint
 continuum is detected in the UVB and VIS arm after binning in the dispersion
@@
 
-The gross structure of these spectra have previously been issued in GCN 19192
+The gross content of these spectra have previously been issued in GCN 19192
 \citep{GCN19192}. Spectroscopic integration began 13 hours post GRB trigger and
@@
 top of the continuum are emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
-\nii, and [\sii]~at a consistent redshift of $z = 0.726$. It is unclear how much
+\niil, and [\sii]~at a consistent redshift of $z = 0.726$. It is unclear how much
 of the continuum is due to the host and how much is from the GRB afterglow. At
@@
 
-This short GRB with extended emission (GCN 19276 \citep{GCN19276}) was observed
-in RRM mode, with observations starting after only 7.7 minutes. The gross
-content has already been presented in GCN 19274 \citep{GCN19274}. These
+This short GRB with extended emission \citep[GCN 19276;][]{GCN19276} was
+observed in RRM mode, with observations starting after only 7.7 minutes. The
+gross content has already been presented in GCN 19274 \citep{GCN19274}. These
 observations have the shortest delay between trigger time and start of
 observation. Spectroscopic integration began after 8.4 minutes and lasted for $3
-\times 600$. The last image in the nodding sequence was aborted because the
+\times 600$ s. The last image in the nodding sequence was aborted because the
 object was setting and the telescope reached the hardware limit. A blue,
@@
 continuum is detected across the spectral coverage of X-shooter, but no clear
-absorption lines are easily discernible. It is therefore unclear how much of the
+absorption lines are discernible. It is therefore unclear how much of the
 continuum is due to the host and how much is due to the GRB. Superposed on the
 continuum are two sets of emission lines, offset spatially, but at very similar
-redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and [\sii]~in both
+redshifts. We identify lines from \oii, \hb, \oiii, \ha, \niil, and [\sii]~in both
 systems -- all at $z = 0.555$. As per GCN 19350 \citep{GCN19350}, the GRB
@@
 GRB~160625B is the first GRB for which a signicant linear polarization was
-measured \cite{Troja2017}. Because this burst was detected by \textit{Fermi} and
+measured \citep{Troja2017}. Because this burst was detected by \textit{Fermi}, it
 is therefore not a part of the statistical sample. The gross content of these
@@
 hours after the trigger, these spectra are obtained in $4 \times 600$ s. The GRB
-afterglow continuum is detected a high signal-to-noise throughout the spectral
+afterglow continuum is detected at high signal-to-noise throughout the spectral
 coverage of X-shooter. Imprinted on the afterglow continuum are a multitude of
@@
 The data presented here also formed the basis of GCN 19773 \citep{GCN19773}, and
-is published in \citet{Heintz2017b}. Observations started 22.37 hours after the
+are published in \citet{Heintz2017b}. Observations started 22.37 hours after the
 BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the
@@
 \feii~and \alii~are found at $z = 0.736$. At the same redshift, emission lines
-from \oii, \oiii, \ha, \hb, \hg, \nii, [\sii], [\siii]~are found. A second epoch,
-lasting 3.6ks, is obtained after the afterglow has faded, confirming the
+from \oii, \oiii, \ha, \hb, \hg, \niil, [\sii], and [\siii]~are found. A second epoch,
+lasting 3.6ks, is obtained after the afterglow had faded, confirming the
 emission line detections. The host galaxy is found to have a roughly solar
@@
 first noted by \citet{GCN19975}, is extended and lies just outside the XRT error
-circle. Based on GROND imaging, \citep{GCN19975} report moderate variability,
+circle. Based on GROND imaging, \citet{GCN19975} report moderate variability,
 and suggest this object as an afterglow candidate. Following our original
@@
 
-\subsection{GRB~161007A (z =4.7)} \label{161007}
-
-This data has not been published elsewhere. Observations for GRB~161007A started
-323 hours after the burst trigger and contains the potential host. $4 \times
-600$ seconds of observations reveals a faint continuum rising abruptly above the
-noise at $\sim$ 690 nm and continuing through 2100 nm. If the host is located at
-$z \sim 4.7$, the drop in continuum flux is the Lyman and the
-absence of nebular emission lines is due to \oii~being shifted out of the
-wavelength coverage. Alternatively, an early-type host at $z = 0.71$ could
-exhibit the 400 nm break at 600 nm, but due to the preference of long-duration
-GRBs for star-forming galaxies, this is the least likely explanation, why we
-prefer the high-z solution.
+\subsection{GRB~161007A (z = na)} \label{161007}
+
+These data have not been published elsewhere. Observations for GRB~161007A
+started 323 hr after the burst trigger and were centered on the potential host
+galaxy (see GCN 20014 \citep{GCN20014} and GCN 20020 \citep{GCN20020}). $4
+\times 600$ s of observations were obtained revealing a faint continuum rising
+abruptly above the noise at $\sim 685$ nm and extending up to 2100 nm. A very
+low significance continuum is detected at shorter wavelengths, down to $\sim
+600$ nm. No significant emission features could be identified. Two possibilities
+can explain the observed continuum break, but neither are fully satisfactory.
+The first option is the Lyman alpha cutoff at $z \sim 4.6$; at this redshift, the
+usually detected strong nebular lines would be shifted out of the covered
+wavelength range. The host would be however exceptionally bright, with an
+absolute magnitude at $\approx 150$~nm{} $M = -23.8$~AB, that is about 2.8~mag
+brighter than $M^*$ at that redshift (e.g. \citealt{Bouwens2015}). The inferred
+X-ray hydrogen-equivalent column density would also be a record-high $4 \times
+10^{21}$~cm$^{-2}$ (see e.g. \citealt{Campana2012}). Alternatively, the
+continuum discontinuity could be the 400~nm{} break at $z = 0.71$; such
+feature is however prominent only in early-type galaxies, which are unlikely
+hosts of long-duration GRBs (but see \citealt{Rossi2014}). Given the lack of
+optical variability, it cannot be excluded that the object within the XRT error
+circle is a chance association, as the probability of a random alignment is
+small but non negligible (of the order of 1\%). In conclusion, we cannot provide
+a secure redshift for GRB 161007A.
 
@@
 The data presented here also formed the basis of GCN  20061 \citep{GCN20061},
-but is not published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8
+but are not published elsewhere. Starting 11.6 hours after the GRB trigger, 4.8
 ks of integration time captures the afterglow continuum across all three
@@
 \aliii, and	\feii, all at $z =2.823$. Similar to GRB~140506 \citep{Fynbo2014,
-	Heintz2017b},  a break is tentatively detected in the continuum shape is
-detected bluewards of 600 nm, possible signifying some anomalous form of
-extinction.
+	Heintz2017b}, a break in the continuum shape is tentatively detected bluewards
+of 600 nm, possible signifying some anomalous form of extinction.
 
@@
 highest signal-to-noise of all the spectra presented here. A myriad of
-absorption features are visible against the continuum, many which would not be
+absorption features are visible against the continuum, many of which would not be
 visible against lower signal-to-noise spectra. We identify at least 15
@@
 additionally seen, clearly associating the $z = 2.710$ system with the GRB. From
-the \lya-line we infer \nh$=20.96 \pm 0.05$.
+the \lya-line we infer \nh$=20.96 \pm 0.05$. This spectrum will be further analyzed in de Ugarte Postigo et al. (in prep).
 
@@
 \citep{GCN20321}, and are additionally used in both \citet{Ashall2017} and
-\citet{Cano2017}. This GRB is very close-by and is associated width the SN
-Ic-BL, SN2016jca \citep[GCN 20342;][]{GCN20342}. The observations consists of $4
+\citet{Cano2017}. This GRB is very close-by and is associated with the SN
+Ic-BL SN2016jca \citep[GCN 20342;][]{GCN20342}. The observations consist of $4
 \times 600$ s spectroscopic integration, starting 1.5 days after the GRB
@@
 continuum is detected at moderate signal-to-noise across across the entire
-spectral coverage of X-shooter. In the bluest end og the spectral coverage, the
-background noise increases, this lowering the continuum below detection.
+spectral coverage of X-shooter. In the bluest end of the spectral coverage, the
+background noise increases, thus lowering the continuum below detection.
 Superposed on the afterglow continuum are both absorption lines and emission
@@
 follow-up effort that has lasted for 8 years. The majority of the spectral
-features in this spectrum has previously been reported in GCN 20589
+features in this spectrum have previously been reported in GCN 20589
 \citep{GCN20589}. Spectroscopy began 9.7 hours after the BAT trigger, and lasted
@@
 signal-to-noise throughout the spectral coverage of X-shooter, down to 430 nm.
-The lowest wavelength continuum from 430 nmto 560 nm is absorbed by the
+The lowest wavelength continuum from 430 nm to 560 nm is absorbed by the
 \lya~forest, with the broad \lya~absorption trough centered at 565 nm. In the
 continuum, we additionally identify metal absorption features from \SIii, \civ,
-\SIiv, \feii, and \mgii, with the additionally detection of
+\SIiv, \feii, and \mgii, with the additional detection of
 \SIii$^*$~fine-structure lines. With an additional detection of \oiii~and all
 lines detected at $z=3.645$, this unequivocally marks this as the redshift of
-the GRB. There is an additional \civ~absorber at $3.077$. From the
+the GRB. There is an additional \civ~absorber at $z = 3.077$. From the
 \lya~absorption we measure \nh$= 21.55 \pm 0.10$.

### 2018-02-15 document/authorlist.tex
Last bursts and edits
Final draft sent around
Last updates before submission

@@
-\author{J.~Selsing\inst{1}
-\and XSGRB collaboration
-%\and D.~Malesani\inst{1}
-%\and T.~Kr\"{u}hler\inst{1}
-%\and J.~P.~U. Fynbo\inst{1}
-%\and P.~Goldoni\inst{14}
-%\and A.~de~Ugarte~Postigo\inst{11}
-%\and J.~Japelj\inst{20}
-%\and P.~D'Avanzo
-%\and Z.~Cano
-%\and S.~Covino\inst{10}
-%\and V.~D'Elia\inst{712}
-%\and H.~Flores
-%\and O.~E.~Hartoog\inst{6}
-%\and J.~Hjorth\inst{1}
-%\and P.~Jakobsson\inst{5}
-%\and A.~Levan
-%\and A.~Melandri
-%\and S.~Piranomonte \inst{7}
-%\and R.~S\'anchez-Ram\'irez\inst{11}
-%\and S.~Schulze\inst{1718}
-%\and N.~R.~Tanvir\inst{19}
-%\and C.~Th{\"o}ne
-%\and S.~D.~Vergani\inst{78}
-%\and P.~M.~Vreeswijk\inst{3}
-%\and D.~J.~Watson\inst{1}
-%\and K.~Wiersema\inst{19}
-%\and D.~Xu\inst{1}
-%\and L.~Christensen\inst{1}
-%\and A.~De~Cia\inst{3}
-%\and L.~Kaper\inst{6}
-%\and L.~A.~Antonelli
-%\and F.~Fiore
-%\and A.~Gomboc
-%\and P.~Groot
-%\and F.~Hammer
-%\and C.~Ledoux\inst{2}
-%\and E.~Maiorano
-%\and B.~Milvang-Jensen\inst{1}
-%\and E.~Palazzi
-%\and E.~Pian
-%\and J.~Schaye
-%\and G.~Tagliaferri\inst{7},
-%\and R.~A.~M.~J.~Wijers\inst{6}
+\author{
+J.~Selsing \inst{1}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}
+\addtocounter{footnote}{2}
+\and D.~Malesani \inst{1, 2, 3}\fnmsep\thanks{On-call observer}
+\and P.~Goldoni \inst{4}\fnmsep$^\dagger$
+\and J.~P.~U. Fynbo \inst{1, 2}\fnmsep$^\dagger$
+\and T.~Kr\"{u}hler \inst{5}\fnmsep$^\dagger$
+\and L.~A.~Antonelli \inst{6}\fnmsep$^\dagger$
+\and M.~Arabsalmani \inst{7, 8}
+\and J.~Bolmer \inst{5, 9}\fnmsep$^\dagger$
+\and Z.~Cano \inst{10}\fnmsep$^\dagger$
+\and L.~Christensen \inst{1}
+\and S.~Covino \inst{11}\fnmsep$^\dagger$
+\and P.~D'Avanzo \inst{11}\fnmsep$^\dagger$
+\and V.~D'Elia \inst{12}\fnmsep$^\dagger$
+\and A.~De~Cia \inst{13}
+\and A.~de~Ugarte~Postigo \inst{1, 10}\fnmsep$^\dagger$
+\and H.~Flores \inst{14}\fnmsep$^\dagger$
+\and M.~Friis \inst{15, 16}
+\and A.~Gomboc \inst{17}
+\and J.~Greiner \inst{5}
+\and P.~Groot \inst{18}
+\and F.~Hammer \inst{14}
+\and O.E.~Hartoog \inst{19}\fnmsep$^\dagger$
+\and K.~E.~Heintz \inst{1, 2, 20}\fnmsep$^\dagger$
+\and J.~Hjorth \inst{1}\fnmsep$^\dagger$
+\and P.~Jakobsson \inst{20}\fnmsep$^\dagger$
+\and J.~Japelj \inst{19}\fnmsep$^\dagger$
+\and D.~A.~Kann \inst{10}\fnmsep$^\dagger$
+\and L.~Kaper \inst{19}
+\and C.~Ledoux \inst{9}
+\and G.~Leloudas \inst{1}
+\and A.J.~Levan \inst{21}\fnmsep$^\dagger$
+\and E.~Maiorano \inst{22}
+\and A.~Melandri \inst{11}\fnmsep$^\dagger$
+\and B.~Milvang-Jensen \inst{1, 2}
+\and E.~Palazzi \inst{22}
+\and J.~T.~Palmerio \inst{23}\fnmsep$^\dagger$
+\and D.~A.~Perley \inst{24}\fnmsep$^\dagger$
+\and E.~Pian \inst{22}
+\and S. ~Piranomonte \inst{6}\fnmsep$^\dagger$
+\and G.~Pugliese \inst{19}\fnmsep$^\dagger$
+\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{25}\fnmsep$^\dagger$
+\and S.~Savaglio \inst{26}
+\and P.~Schady \inst{5}
+\and S.~Schulze \inst{27}\fnmsep$^\dagger$
+\and J.~Sollerman \inst{28}
+\and M.~Sparre \inst{29}\fnmsep$^\dagger$
+\and G.~Tagliaferri \inst{11}
+\and N.~R.~Tanvir \inst{30}\fnmsep$^\dagger$
+\and C.~C.~Th\"{o}ne \inst{10}
+\and S.D.~Vergani \inst{14}\fnmsep$^\dagger$
+\and P.~Vreeswijk \inst{18, 26}\fnmsep$^\dagger$
+\and D.~Watson \inst{1, 2}\fnmsep$^\dagger$
+\and K.~Wiersema \inst{21, 30}\fnmsep$^\dagger$
+\and R.~Wijers \inst{19}
+\and D.~Xu \inst{31}\fnmsep$^\dagger$
+\and T.~Zafar \inst{32}
 }
@@
 \institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
-%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
-%\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
-%\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
-%\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
-%\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
-%\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
-%\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
-%\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
-%\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
-%\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
-%\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
-%\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
-%\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et Leonie Duquet, 75205 Paris Cedex 13, France
-%\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
-%\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3, 20126, Milano, Italy
-%\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astronom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
-%\and Millennium Center for Supernova Science
-%\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
-%\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI
+%\and Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+\and The Cosmic Dawn Center, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, 2800 Lyngby, Denmark
+\and APC, Astroparticule et Cosmologie, Universit\'e Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cit\'e, 10, Rue Alice Domon et L\'eonie Duquet, 75205, Paris Cedex 13, France
+\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
+\and INAF - Osservatorio Astronomico di Roma, Via Frascati 33, 00078, Monte Porzio Catone (Roma), Italy
+\and IRFU, CEA, Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France
+\and Universit\'e Paris Diderot, AIM, Sorbonne Paris Cit\'e, CEA, CNRS, F-91191 Gif-sur-Yvette, France
+\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
+\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain. % 10
+%
+\and INAF - Osservatorio Astronomico di Brera, via Bianchi 46, 23807, Merate (LC), Italy
+\and Space Science Data Center - Agenzia Spaziale Italiana, via del Politecnico, s.n.c., I-00133, Roma, Italy
+\and European Southern Observatory, Karl-Schwarzschild Str. 2, 85748 Garching bei M\”unchen, Germany
+\and GEPI, Observatoire de Paris, PSL University, CNRS, 5 place Jules Janssen 92195, Meudon, France
+\and KTH Royal Institute of Technology, Department of Physics, 106 91 Stockholm, Sweden
+\and The Oskar Klein Centre for Cosmoparticle Physics, AlbaNova University Centre, 106 91 Stockholm, Sweden
+\and Center for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 11c, 5270 Ajdov\v s\v cina, Slovenia
+\and Department of Astrophysics, IMAPP, Radboud University Nijmegen, PO Box 9010, 6500 GL Nijmegen, the Netherlands
+\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, the Netherlands
+\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107, Reykjavik, Iceland
+\and Department of Physics, University of Warwick, Coventry, CV4 7AL, UK % 20
+%
+\and IASF/INAF Bologna, via Piero Gobetti 101, 40129 Bologna, Italy
+\and Sorbonne Universit\'e, CNRS, UMR7095, Institut d'Astrophysique de Paris, F-75014, Paris, France
+\and Astrophysics Research Institute, Liverpool John Moores University, IC2, Liverpool Science Park, 146 Brownlow Hill, Liverpool L3 5RF, UK
+\and INAF, Istituto Astrofisica e Planetologia Spaziali, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
+\and Physics Department, University of Calabria, I-87036 Arcavacata di Rende, Italy
+\and Department of Particle Physics and Astrophysics, Weizmann Institute of Science, 234 Herzl Street, Rehovot, 761000, Israel
+\and Department of Astronomy, Stockholm University, AlbaNova, 10691 Stockholm, Sweden 
+\and Institut f\"ur Physik und Astronomie, Universit\"at Potsdam, Karl-Liebknecht-Str.\,24/25, 14476 Golm, Germany
+\and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, United Kingdom
+\and CAS Key Laboratory of Space Astronomy and Technology, National Astronomical Observatories, Chinese Academy of Sciences, Beijing, 100012, P.R. China % 30
+%
+\and Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.
 }

### 2018-02-15 document/tables/sample_properties.tex
Last bursts and edits
Last updates before submission

@@
 		\hline\noalign{\smallskip}
-		N$_{BAT}$ & 987 & 156 & 89\\
-		$\log(15-150$keV fluence)  & $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.7}^{+0.7}$   \\
-		N$_{XRT}$ & 909 & 154 & 88\\
-		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.7}^{+0.9}$  \\
-		N$_{\mathrm{HI_x}}$ & 250 & 95 & 74\\
+		N$_{BAT}$ & 981 & 163 & 92\\
+		$\log(15-150$ keV fluence)  & $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.7}^{+0.7}$   \\
+		N$_{XRT}$ & 902 & 160 & 90\\
+		$\log(0.3-10$ keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.7}^{+0.9}$  \\
+		N$_{\mathrm{HI_x}}$ & 248 & 99 & 79\\
 		$\log(\mathrm{N}_{\mathrm{HI_x}})$ & $21.7_{-0.9}^{+0.6}$ &  $21.5_{-3.4}^{+0.7}$ &  $21.6_{-4.5}^{+0.7}$  \\
@@
 \caption{
-	Population properties (median and 14th and 86th percentiles as the
+	Population properties (median and 16th and 84th percentiles as the
 	error intervals) for the \textit{Swift sample} and the subset of bursts

### 2018-02-15 py/NH_distribution_plot.py

@@ -59,7 +59,7 @@ def main():
     g.y = NH_old[NH_old > 20]
     # g = g.plot_joint(pl.scatter, color="#55A868")
     g = g.plot_marginals(sns.distplot, hist=False, color="#55A868", bins = 25, norm_hist=True, kde_kws={"shade": True, "bw": 0.2})
-    g = g.plot_joint(sns.kdeplot, color="#55A868", label="Tanvir et al. 2017")
+    g = g.plot_joint(sns.kdeplot, color="#55A868")
     # pl.legend()
     # Plot new distribution
     NH_new = NH[XSGRB]
@@ -115,8 +115,8 @@ def main():
     ax = pl.gca()
     handle, label = ax.get_legend_handles_labels()
     # from matplotlib.lines import Line2D     
-    l1 = mpl.lines.Line2D([1], [1], color = "#4C72B0")
-    l2 = mpl.lines.Line2D([1], [1], color = "#55A868")
+    l1 = mpl.lines.Line2D([1], [1], color = "#55A868")
+    l2 = mpl.lines.Line2D([1], [1], color = "#4C72B0")
     l3 = mpl.lines.Line2D([1], [1], color = "#C44E52")
     # print(handle, label)
     # pl.legend([l1, l2, l3], label)
@@ -128,7 +128,7 @@ def main():
         legobj.set_linewidth(5.0)
         # legobj.handlelength(2.0)
     pl.savefig("../document/figures/NH_dist.pdf")
-    # pl.show()
+    pl.show()
 
 if __name__ == '__main__':
     main()

### 2018-02-15 py/Sample_completeness.py

@@ -62,7 +62,7 @@ def main():
     l, m, h = np.nanpercentile(filt_nan(z[pure == "Yes"]), (14, 50, 86))
     mm = np.nanmean(filt_nan(z[pure == "Yes"]))
     print(mm, m, l - m, h - m)
-    exit()
+    # exit()
     l, m, h = np.nanpercentile(fbo, (14, 50, 86))
     mm = np.nanmean(fbo)
     print(mm, m, l - m, h - m)
@@ -249,7 +249,7 @@ def main():
 
     ax1.set_xlabel(r"$\log(15-150~\mathrm{keV}~\mathrm{Fluence}/\mathrm{erg}~\mathrm{cm}^{-2})$")
     ax2.set_xlabel(r"$\log(0.3-10~\mathrm{keV}~\mathrm{Flux}/\mathrm{erg}~\mathrm{cm}^{-2}~\mathrm{s}^{-1})$")
-    ax3.set_xlabel(r"$\log(N_{\mathrm{HI, X}}/\mathrm{cm}^{-2})$")
+    ax3.set_xlabel(r"$\log(N_{\mathrm{H, X}}/\mathrm{cm}^{-2})$")
     # ax1.set_ylabel(r"N")
 
     ax1.set_xlim(-8, -4)

### 2018-02-15 py/betaOX.py

@@ -190,7 +190,7 @@ def main():
     ax = pl.gca()
     ax.axhline(0.5, color="black", linestyle="dashed", alpha=0.5)
     ax.annotate(r"$\beta_{OX} = 0.5$", (19.6, 0.45))
-    g.set_axis_labels(r"$\log(N_{\mathrm{HI, X}}/\mathrm{cm}^{-2})$", r"Darkness [$\beta_{OX}$]")
+    g.set_axis_labels(r"$\log(N_{\mathrm{H, X}}/\mathrm{cm}^{-2})$", r"Darkness [$\beta_{OX}$]")
     pl.tight_layout()
 
     # print(stats.ks_2samp(betaOX[idx_2], f_betaOX[f_idx_2]))

### 2018-02-21 document/XSGRB_sample.tex
Last commit before Arxiv submission

@@
 \documentclass[longauth]{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
-%\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
 %\documentclass[referee]{aa}
@@
 
-\title{The X-shooter GRB afterglow legacy sample (XS-GRB). \thanks{Based on
+\title{The X-shooter GRB afterglow legacy sample (XS-GRB)\thanks{Based on
 		observations collected at the European Southern Observatory, Paranal, Chile,
@@
 		088.A-0051, 089.A-0067, 090.A-0088, 091.A-0877, 0091.C-0934, 092.A-0124,
-		092.D-0056, 093.A-0069, 094.A-0134, 095.A-0045, 095.B-0811, 096.A-0079,
-		097.A-0036, and 098.A-0055.}}
+		092.D-0056, 092.D-0633, 093.A-0069, 094.A-0134, 095.A-0045, 095.B-0811,
+		096.A-0079, 097.A-0036, 098.A-0136, and 098.A-0055.}}
 
@@
 
-\keywords{Gamma-ray burst: general --- Galaxies: high-redshift --- ISM: general --- Techniques: spectroscopic}
+\keywords{Gamma-ray burst: general --- Galaxies: high-redshift --- ISM: general --- Techniques: spectroscopic --- Galaxies: star formation --- catalogs}
 
@@
 	\item No bright, nearby stars within $ 1.8 + 0.4 \times \exp[(20 - R)/2.05$]
-arcsec, where $R$ is the USNO magnitude of the star)
+arcsec, where $R$ is the USNO magnitude of the star.
 
@@
 \begin{figure*}
-	\centerline{\includegraphics{figures/skymap.pdf}} \caption{Mollweide projection
+	\centerline{\includegraphics[width=\linewidth]{figures/skymap.pdf}} \caption{Mollweide projection
 	in Galactic coordinates of the full sky showing the positions on the sky of the
@@
 cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
-one of two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during
-the duration of this follow-up campaign. The observations have been taken during
-a period of eight years corresponding to the ESO observing periods P84 through
-P98 under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
+one of two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the
+duration of this follow-up campaign. The observations have been taken during a
+period of eight years corresponding to the ESO observing periods P84 through P98
+under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
 087.A-0055, 088.A-0051, 089.A-0067, 090.A-0088, 092.A-0124, 093.A-0069,
@@
 have included a few additional bursts, from the programmes 084.D-0265 (PI:
-Benetti), 091.A-0877 (PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633 (PI:
-Greiner), and 095.B-0811 (PI: Levan). The total collection of spectra represents
-\textit{all} GRB afterglows that have been followed up by X-shooter up to
-\termdate, which marks the end of the XS-GRB legacy follow-up program.
+Benetti), 091.A-0877 (PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633,
+098.A-0136 (PI: Greiner), and 095.B-0811 (PI: Levan). The total collection of
+spectra represents \textit{all} GRB afterglows that have been followed up by
+X-shooter up to \termdate, which marks the end of the XS-GRB legacy follow-up
+program.
 
@@
 \begin{figure}[!t]
-	\centerline{\includegraphics[width=8cm]{figures/resolutuions}} \caption{Blue
+	\centerline{\includegraphics[width=\columnwidth]{figures/resolutuions}} \caption{Blue
 		data points show the FWHM (km/s) of Gaussian fits to unresolved telluric
@@
 \begin{figure*}
-	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
+	\centerline{\includegraphics[width=\linewidth]{figures/GRB121024A.pdf}}
 \caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
@@
 \href{https://sid.erda.dk/cgi-sid/ls.py?share_id=DBuNORk1lI}{ZIP file}, through
-\url{http://grbspec.iaa.es}, and additionally through the ESO archive in the
+\url{http://grbspec.iaa.es} \citep{DeUgartePostigo2014f}, and additionally through the ESO archive in the
 form of \href{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}{phase 3
@@
 \begin{figure}[!ht]
-	\centerline{\includegraphics[width=9cm]{figures/timing.pdf}} \caption{Afterglow
+	\centerline{\includegraphics[width=\columnwidth]{figures/timing.pdf}} \caption{Afterglow
 	magnitude at the start of observation and redshift completeness as a function
@@
 \begin{figure*}
-	\centerline{\includegraphics[width=18cm]{figures/completeness_BAT.pdf}}
+	\centerline{\includegraphics[width=\linewidth]{figures/completeness_BAT.pdf}}
 \caption{Comparison between the burst properties of all bursts observed with
@@
 \begin{figure}[!ht]
-	\centerline{\includegraphics[width=9cm]{figures/XSGRB_p_values.pdf}}
+	\centerline{\includegraphics[width=\columnwidth]{figures/XSGRB_p_values.pdf}}
 \caption{Relational graph showing the respective $p$-values. They all represent
@@
 \begin{figure*}[!ht]
-	\centering \includegraphics[width=16cm]{figures/XSGRB_redshift.pdf}
+	\centering \includegraphics[width=\linewidth]{figures/XSGRB_redshift.pdf}
 \caption{Redshift distribution as a function of intrinsic BAT $\gamma$-ray
@@
 \begin{figure}[!ht]
-	\centerline{\includegraphics[width=9cm]{figures/betaOX.pdf}}
+	\centerline{\includegraphics[width=\columnwidth]{figures/betaOX.pdf}}
 \caption{$\beta_{OX}$ against X-ray derived hydrogen column density. In red is
@@
 \begin{figure}[!t]
-	\centering \includegraphics[width=9cm]{figures/NH_dist.pdf}
+	\centering \includegraphics[width=\columnwidth]{figures/NH_dist.pdf}
 \caption{Distributions of hydrogen column densities for absorbers found in
@@
 obtaining these target of opportunity data.
+%
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}.
+The analysis and plotting was achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy, and Scipy \citep{scipy, VanderWalt2011}, along with other community-developed packages.
+%
 
@@
 	\centering
-	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=1, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption{Measurements of the hydrogen column densities for all bursts with a
@@
 	\centering
-	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=2, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=3, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=4, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=5, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=6, width=16cm]{figures/HI_measurements.pdf}
+	\includegraphics[page=6, width=\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 X-ray hydrogen-equivalent column density would also be a record-high $4 \times
-10^{21}$~cm$^{-2}$ (see e.g. \citealt{Campana2012}). Alternatively, the
+10^{23}$~cm$^{-2}$ (see e.g. \citealt{Campana2012}). Alternatively, the
 continuum discontinuity could be the 400~nm{} break at $z = 0.71$; such

### 2018-02-21 py/NH_distribution_plot.py

@@ -17,14 +17,15 @@ from scipy import stats
 
 
 params = {
-   'axes.labelsize': 10,
-   'font.size': 10,
-   'legend.fontsize': 10,
-   'xtick.labelsize': 10,
-   'ytick.labelsize': 10,
+   'axes.labelsize': 14,
+   'font.size': 14,
+   'legend.fontsize': 14,
+   'xtick.labelsize': 14,
+   'ytick.labelsize': 14,
    'text.usetex': True ,
-   'figure.figsize': [7.281, 4.5]
+   # 'figure.figsize': [7.281, 4.5]
    }
+mpl.rcParams.update(params)
 
 def main():
     """
@@ -113,8 +114,9 @@ def main():
 
     # Save figure for tex
     ax = pl.gca()
+    ax.set_ylim((20, 24))
     handle, label = ax.get_legend_handles_labels()
-    # from matplotlib.lines import Line2D     
+    # from matplotlib.lines import Line2D
     l1 = mpl.lines.Line2D([1], [1], color = "#55A868")
     l2 = mpl.lines.Line2D([1], [1], color = "#4C72B0")
     l3 = mpl.lines.Line2D([1], [1], color = "#C44E52")

### 2018-02-21 py/betaOX.py

@@ -13,15 +13,15 @@ from scipy import stats
 import matplotlib as mpl
 
 params = {
-   'axes.labelsize': 10,
-   'font.size': 10,
-   'legend.fontsize': 10,
-   'xtick.labelsize': 10,
-   'ytick.labelsize': 10,
+   'axes.labelsize': 14,
+   'font.size': 14,
+   'legend.fontsize': 14,
+   'xtick.labelsize': 14,
+   'ytick.labelsize': 14,
    'text.usetex': True,
-   'figure.figsize': [7.281, 4.5]
+   # 'figure.figsize': [7.281, 4.5]
    }
-
+mpl.rcParams.update(params)
 
 def iqr(a):
     """Calculate the IQR for an array of numbers."""
@@ -189,8 +189,8 @@ def main():
     g = g.plot_joint(pl.plot, color=color, label="Fynbo et al. 2009")
     ax = pl.gca()
     ax.axhline(0.5, color="black", linestyle="dashed", alpha=0.5)
-    ax.annotate(r"$\beta_{OX} = 0.5$", (19.6, 0.45))
-    g.set_axis_labels(r"$\log(N_{\mathrm{H, X}}/\mathrm{cm}^{-2})$", r"Darkness [$\beta_{OX}$]")
+    ax.annotate(r"$\beta_{OX} = 0.5$", (19.6, 0.42))
+    g.set_axis_labels(r"$\log(N_{\mathrm{H, X}}/\mathrm{cm}^{-2})$", r"$\beta_{OX}$")
     pl.tight_layout()
 
     # print(stats.ks_2samp(betaOX[idx_2], f_betaOX[f_idx_2]))
@@ -226,7 +226,7 @@ def main():
 
 
     # Save figure for tex
-    pl.legend()
+    pl.legend(loc = 1)
     pl.savefig("../document/figures/betaOX.pdf", dpi="figure")
     pl.show()
 

### 2018-02-21 py/spec_plot.py

@@ -19,7 +19,7 @@ params = {
    'xtick.labelsize': 14,
    'ytick.labelsize': 14,
    'text.usetex': True,
-   'figure.figsize': [8.27, 11.7]
+   'figure.figsize': [8, 8*1.4142]
    }
 mpl.rcParams.update(params)
 

### 2018-08-28 document/referee_response.txt
Arxiv submission

This report presents a review of the article "The X-shooter GRB 
afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
authors present and make available to the community a complete sample 
of GRB afterglows observed with the X-Shooter spectrograph on the 
VLT. This dataset is remarkable in several respect (>100 spectra 
available taken over 8 years) and some of these have not be published 
elsewhere. However, there are a number of important issues regarding 
both the scientific measurements and their presentation which are 
detailed below. These concerns need to be addressed before the paper 
can be considered for publication. More details on these points 
together with other comments are provided below.

Major Concerns:

# - One aspect of these data that needs more emphasis is the 
# intervening absorbers. The data presented in this paper present a 
# remarkable number of metal and HI intervening absorbers as can be 
# judged from reading the appendix. These deserve much more attention, 
# including a table listing their redshifts, column density 
# measurements when available, metal lines detected, etc. A figure 
# with their redshift distribution would also be very welcome. In 
# addition, it would be interesting to compare the properties of this 
# population with the absorbers towards background quasars.

While we agree that the subject of intervening absorbers is very interesting and
important, the main focus of this paper is to present the technicalities of the
sample and the data reduction. In a series of fourth-coming papers, we will
investigate sub-themes independently. On the subject of intervening absorbers,
already published is Christensen et al., A&A, 608A, 84C, 2017,
(arXiv:1709.01084) and we also have de Ugarte Postigo et al. (in prep).

# - The manuscript states that the sample is homogeneous which is a bit 
# misleading. The spectrograph used to make the follow-up is the same 
# for all afterglows, but the SNR (due to a combination of brightness 
# of the source and time of observations after the burst) and the 
# rest-frame wavelength coverage vary enormously. While remarkable in 
# many respects, it is not appropriate to qualify this sample as 
# homogeneous.

We completely agree and have rephrased the manuscript to reflect this change.

# - There are a great many details provided about the data reduction, some of 
# which the reader fails to relate to the final scientific goals. For 
# example, a correction is applied to correct offsets in the 
# wavelength calibration, which presumably mostly impact the redshift 
# determination. How does the improvement translate into the redshift 
# determination accuracy? Similarly, for the telluric correction. What 
# is the impact of this on the final science? Conversely, a 
# description of the flux calibration error estimate would have been 
# welcome.

# - The comparison of N(HI) distributions between intervening quasar 
# absorbers and intrinsic GRB absorbers is not direct. The measurement 
# of the column density of a system in the case of a quasar absorber 
# integrate across the whole system. Conversely, in the case of GRBs, 
# the burst event takes place at a random position in the host galaxy 
# and in fact, on average, the column density will measure over only 
# half of the system. This geometric effect should be taken into 
# account.

# - Another concern is to do with using absorption features to measure 
# the redshift. Indeed, one cannot exclude that the absorption is not 
# at the position of host galaxy but lying in front of it. For this 
# reason, the redshifts provided should be seen as upper limits.

Other Concerns:

- the links to the actual data seem to be inactive. It is important to 
be able to review the material which is made available to the 
community.

- caption of Fig. 3 is cut-out

- Fig. 5: please label the y-axis

- Fig. 10: caption is cut-out again 
GRB150403A: could the continuum be lower? 
GRB150915A: is this fit significant?

- appendix A: refer to the spectral PSF in the title?

- appendix B: science-verified, Paranalization: 
perhaps avoid this ESO jargon?

typos: 
- pt. 4 p. 3: Paranl 
- p. 3: exampl 
- p. 3: sorrounding 
- p. 9: curios 
- there are numerous typos in the appendix which I will not review 
here. Please make sure to double check spelling there.

I would be happy to review an updated version of the manuscript where 
the above comments are taken into account.

### 2018-08-28 document/XSGRB_sample_arxiv.tex
Arxiv submission

--- 2018-02-21 document/XSGRB_sample.tex
@@
 
+%\usepackage{hyperref}
+
 \usepackage{graphicx,url,twoopt,natbib}
@@
 
-\usepackage{url}
-\usepackage{color,hyperref}
+\usepackage{color}
 \definecolor{linkcolor}{rgb}{0,0.3,0.7}
@@
 
-\usepackage{natbib}
+\usepackage{url}
 %\usepackage{caption}
@@
 	\item No bright, nearby stars within $ 1.8 + 0.4 \times \exp[(20 - R)/2.05$]
-arcsec, where $R$ is the USNO magnitude of the star.
+arcsec, where $R$ is the USNO magnitude of the star).
 
@@
 cross-dispersed echelle spectrograph, X-shooter \citep{Vernet2011}, mounted on
-one of two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during the
-duration of this follow-up campaign. The observations have been taken during a
-period of eight years corresponding to the ESO observing periods P84 through P98
-under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
+one of two Unit Telescopes at ESO/VLT, UT2 (Kueyen) and UT3 (Melipal) during
+the duration of this follow-up campaign. The observations have been taken during
+a period of eight years corresponding to the ESO observing periods P84 through
+P98 under the following programme IDs: 084.A-0260, 085.A-0009, 086.A-0073,
 087.A-0055, 088.A-0051, 089.A-0067, 090.A-0088, 092.A-0124, 093.A-0069,
@@
 have included a few additional bursts, from the programmes 084.D-0265 (PI:
-Benetti), 091.A-0877 (PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633,
-098.A-0136 (PI: Greiner), and 095.B-0811 (PI: Levan). The total collection of
-spectra represents \textit{all} GRB afterglows that have been followed up by
-X-shooter up to \termdate, which marks the end of the XS-GRB legacy follow-up
-program.
+Benetti), 091.A-0877 (PI: Schady), 092.D-0056 (PI: Rau), 092.D-0633, 098.A-0136 (PI:
+Greiner), and 095.B-0811 (PI: Levan). The total collection of spectra represents
+\textit{all} GRB afterglows that have been followed up by X-shooter up to
+\termdate, which marks the end of the XS-GRB legacy follow-up program.
 
@@
 \begin{figure*}
-	\centerline{\includegraphics[width=\linewidth]{figures/GRB121024A.pdf}}
+	\centerline{\includegraphics[width=16cm]{figures/GRB121024A.pdf}}
 \caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
@@
 \href{https://sid.erda.dk/cgi-sid/ls.py?share_id=DBuNORk1lI}{ZIP file}, through
-\url{http://grbspec.iaa.es} \citep{DeUgartePostigo2014f}, and additionally through the ESO archive in the
+\url{http://grbspec.iaa.es} \citep{deUgartePostigo2014f}, and additionally through the ESO archive in the
 form of \href{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}{phase 3
@@
 
-\bibliographystyle{aa}
+\bibliographystyle{aa_arxiv}
 \bibliography{XSGRB_sample,GCN}

### 2018-08-28 py/database/timediff.py

@@ -12,11 +12,11 @@ def timediff(t1, t2):
     time2 = datetime.datetime(int(t2[0:4]), int(t2[5:7]), int(t2[8:10]), hour=int(round(float(t2[11:13]))), minute=int(round(float(t2[14:16]))), second=int(round(float(t2[17:19]))))
 
     dt = time2 - time1
-    return dt.total_seconds()/(3600)
+    return dt.total_seconds()/(60)
 
 if __name__ == '__main__':
 
-    t_trigger = "2016:12:19T18:48:39"
-    t_obs = "2016-12-21T06:33:30.354"
+    t_trigger = "2018:04:04T00:45:35"
+    t_obs = "2018:04:04T01:01:33"
 
     print(timediff(t_trigger, t_obs))

### 2018-09-21 document/referee_response2.txt
Addressing referee comments

--- 2018-08-28 document/referee_response.txt
@@
-This report presents a review of the article "The X-shooter GRB 
-afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
-authors present and make available to the community a complete sample 
-of GRB afterglows observed with the X-Shooter spectrograph on the 
-VLT. This dataset is remarkable in several respect (>100 spectra 
-available taken over 8 years) and some of these have not be published 
-elsewhere. However, there are a number of important issues regarding 
-both the scientific measurements and their presentation which are 
-detailed below. These concerns need to be addressed before the paper 
-can be considered for publication. More details on these points 
-together with other comments are provided below.
-
-Major Concerns:
-
+# 
+# ------------------- 
+# Referee report 
+# 
+# This report presents a review of the article "The X-shooter GRB 
+# afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
+# authors present and make available to the community a complete sample 
+# of GRB afterglows observed with the X-Shooter spectrograph on the 
+# VLT. This dataset is remarkable in several respect (>100 spectra 
+# available taken over 8 years) and some of these have not be published 
+# elsewhere. However, there are a number of important issues regarding 
+# both the scientific measurements and their presentation which are 
+# detailed below. These concerns need to be addressed before the paper 
+# can be considered for publication. More details on these points 
+# together with other comments are provided below. 
+# 
+# 
+# 
+# Major Concerns: 
+# 
 # - One aspect of these data that needs more emphasis is the 
@@
 
-While we agree that the subject of intervening absorbers is very interesting and
-important, the main focus of this paper is to present the technicalities of the
-sample and the data reduction. In a series of fourth-coming papers, we will
-investigate sub-themes independently. On the subject of intervening absorbers,
-already published is Christensen et al., A&A, 608A, 84C, 2017,
-(arXiv:1709.01084) and we also have de Ugarte Postigo et al. (in prep).
+This paper is meant as a general introduction of the XSGRB legacy sample
+and constitutes the first in a series of papers. While we certainly
+agree that the issue of intervening absorbers is an interesting one, we
+feel that it is best treated in a dedicated publication, which we are
+already working on (de Ugarte Postigo et al. in prep.). This paper will
+contain the complete list of intervening systems, as well as linelists,
+redshifts, EWs, etc..
 
@@
 
-We completely agree and have rephrased the manuscript to reflect this change.
+The referee is correct here. Our statement about homogeneity concerns
+only the selection, not the outcome of our observations. We have
+attempted to reword anywhere relevant and we added an explicit caveat on
+the resulting S/N homogeneity in paragraph §2.1.
 
@@
 
+Concerning the wavelength calibration improvement, we have quantified
+the magnitude of the improvement, and explained its relevance (§3.2). We
+have also added a short paragraph to §3.5 explaining the impact of the
+telluric correction on the inferred results.
+
+Regarding the flux calibration accuracy, we have added a new section ($3.7). This includes
+
 # - The comparison of N(HI) distributions between intervening quasar 
@@
 
+This is certainly true, when comparing QSO DLAs to GRB DLA. To accounted for the geometric effect in Fig.
+
+#
 # - Another concern is to do with using absorption features to measure 
@@
 # reason, the redshifts provided should be seen as upper limits. 
+#
 
-Other Concerns:
-
-- the links to the actual data seem to be inactive. It is important to 
-be able to review the material which is made available to the 
-community.
-
-- caption of Fig. 3 is cut-out
-
-- Fig. 5: please label the y-axis
-
-- Fig. 10: caption is cut-out again 
-GRB150403A: could the continuum be lower? 
-GRB150915A: is this fit significant?
-
-- appendix A: refer to the spectral PSF in the title?
-
-- appendix B: science-verified, Paranalization: 
-perhaps avoid this ESO jargon?
-
-typos: 
-- pt. 4 p. 3: Paranl 
-- p. 3: exampl 
-- p. 3: sorrounding 
-- p. 9: curios 
-- there are numerous typos in the appendix which I will not review 
-here. Please make sure to double check spelling there.
-
-I would be happy to review an updated version of the manuscript where 
-the above comments are taken into account.
+# 
+# Other Concerns: 
+# 
+# - the links to the actual data seem to be inactive. It is important to 
+# be able to review the material which is made available to the 
+# community. 
+# 
+# - caption of Fig. 3 is cut-out 
+# 
+# - Fig. 5: please label the y-axis 
+# 
+# - Fig. 10: caption is cut-out again 
+# GRB150403A: could the continuum be lower? 
+# GRB150915A: is this fit significant? 
+# 
+# - appendix A: refer to the spectral PSF in the title? 
+# 
+# - appendix B: science-verified, Paranalization: 
+# perhaps avoid this ESO jargon? 
+# 
+# typos: 
+# - pt. 4 p. 3: Paranl 
+# - p. 3: exampl 
+# - p. 3: sorrounding 
+# - p. 9: curios 
+# - there are numerous typos in the appendix which I will not review 
+# here. Please make sure to double check spelling there. 
+# 
+# I would be happy to review an updated version of the manuscript where 
+# the above comments are taken into account. 
+# 
+# -----------------------

### 2018-09-21 document/XSGRB_sample_rev1.tex
Addressing referee comments

--- 2018-02-21 document/XSGRB_sample.tex
@@
 	intrinsic properties of the GRBs, the follow-up effort has been focused on
-	producing a homogeneous sample of 93 afterglow spectra for GRBs discovered by
+	producing a homogeneously selected sample of 93 afterglow spectra for GRBs discovered by
 	the {\em Swift} satellite. We here provide a public release of all the reduced
@@
 completeness of the sample in Sect. \ref{completeness}.
+
+We remark, however, that while the selection of the afterglows has been carried
+out in a homogeneous way, the resulting quality of the spectra is not uniform.
+Individual bursts vary significantly in brightness, due to intrinsic
+differences in luminosity, dust extinction, and time of observation. The covered
+rest-frame wavelength range is also different for each case. Given that our
+observations are targets-of-opportunity, this leads to a varying S/N, and in
+particular the sensitivity limit to detecting absorption features varies greatly
+from burst to burst and as a function of wavelength.
 
@@
 first burst observed after science verification and the mounting of X-shooter on UT2,
-was GRB~091018, which is the first burst entering our statistically homogeneous
+was GRB~091018, which is the first burst entering our statistically homogeneously selected
 sample. For all bursts that fulfill our sample selection criteria, described in
@@
 	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/inst.html}{this URL} for the nominal spectral resolutions.}. 
-%The slit width sets the width of the atmospheric sky lines and determines the amount of light lost due to the wavelength-dependent seeing PSF extending outside the coverage of the slit, where the width of sky-lines is always set by the slit width whereas both the delivered resolution and the slitloss changes for the better as the seeing PSF drops below the slit width.
-For atmospheric conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a 0\farc9 slit only 76.1 per cent of the light will enter the slit, meaning that for almost all observations a slitloss correction is required.
 %We describe how slitlosses were corrected for in Sect. \ref{postproc}.
@@
 	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc.html}{ESO website}.}.
-The remaining offset is corrected by cross-correlating the observed sky spectrum with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum, estimated as the mode of all flux values, has been subtracted.
-To get the correct seeing point-spread-function (PSF) with which to convolve the synthetic sky, an initial refinement of the wavelength solution has been obtained by cross-correlating the observed sky with an unconvolved synthetic sky.
-This preliminary wavelength calibration is applied to the observed sky.
-The synthetic spectrum is then convolved with an increasing seeing PSF and the width that minimizes $\chi^2$ with the updated observed sky is chosen to be the effective sky-PSF.
-Using the synthetic sky with the matched resolution, a final wavelength calibration can then be calculated by cross-correlating the observed sky with the correctly broadened sky spectrum as a function of a velocity offset.
-Both a multiplicative and an additive offset to the wavelength calibration has been tested, but in terms of $\chi^2$, the model with only a multiplicative offset is preferred.
-The resulting offsets, which were smaller than 0.01 nm in the UVB and VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short periods of time were applied to the corresponding spectra\footnote{The wavelength shifts have been extensively studied by ESO staff \href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in this document}.}.
-Using the convolved synthetic sky, the pixels containing the brightest sky lines have been flagged as such in the bad pixel map.
+
+The remaining offset is corrected by cross-correlating the observed sky spectrum
+with a synthetic sky spectrum \citep{Noll2012, Jones2013} after the continuum,
+estimated as the mode of all flux values, has been subtracted. To get the
+correct seeing point-spread-function (PSF) with which to convolve the synthetic
+sky, an initial refinement of the wavelength solution has been obtained by
+cross-correlating the observed sky with an unconvolved synthetic sky. This
+preliminary wavelength calibration is applied to the observed sky. The synthetic
+spectrum is then convolved with an increasing seeing PSF and the width that
+minimizes $\chi^2$ with the updated observed sky is chosen to be the effective
+sky-PSF. Using the synthetic sky with the matched resolution, a final wavelength
+calibration can then be calculated by cross-correlating the observed sky with
+the correctly broadened sky spectrum as a function of a velocity offset. Both a
+multiplicative and an additive offset to the wavelength calibration has been
+tested, but in terms of $\chi^2$, the model with only a multiplicative offset is
+preferred. The resulting offsets, which were smaller than 0.01 nm in the UVB and
+VIS data and smaller than 0.05 nm in the NIR spectra, but changing over short
+periods of time were applied to the corresponding spectra\footnote{The
+	wavelength shifts have been extensively studied by ESO staff
+	\href{https://www.eso.org/sci/facilities/paranal/instruments/xshooter/doc/XS_wlc_shift_150615.pdf}{in
+		this document}.}. Using the convolved synthetic sky, the pixels containing the
+brightest sky lines have been flagged as such in the bad pixel map.
+
+The resulting offsets are generally small, on the order of $\sim 5$~km~s$^{-1}$,
+and thus do not influence significantly the redshift measurements, whose
+uncertainties are primarily driven by the S/N of the spectra, the line-shapes,
+and the choice of ion species. Such a correction, however, can improve the
+accuracy of line fits, especially when multiple components are tied together,
+across a wide wavelength range, and allows a better comparison with
+high-resolution data. The scale of the correction additionally serves to
+solidify the accuracy of the wavelength solution.
 
@@
 observations have been used.
+
+The effect of telluric correction varies greatly depending on the wavelength
+region. In the red and ner-infrared areas, it allows to identify features in a
+credible manner. which would be otherwise hidden among or blended with
+atmospheric features. The telluric correction allows us to reliably recover the
+continuum level even in regions of moderately low atmospheric transmission (see,
+e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}). We urge however to use
+caution when deriving fitting lines in regions significantly affected by
+telluric absorption.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Slitloss correction and flux calibration accuracy} \label{fluxes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For the majority of the observation presented in this sample, the slit width of
+the target observations is less than the seeing PSF. This means that a fraction
+of the incoming light is lost outside the slit. As an example, for atmospheric
+conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a
+0\farc9 slit, only 76.1 per cent of the light will enter the slit. This slitloss
+is not corrected by the
+flux calibration, as flux standard stars are always observed with a 5 \arcsec
+slit. As a consequence, almost all spectra will, without slitloss correction,
+underestimate the flux by some fraction. Additionally, slitlosses are wavelength
+dependent due to the wavelength dependence of seeing \citep{Boyd1978} and as the
+seeing changes across the slit in wavelength, so does the slitloss.
+
+To correct for the slitlosses, we develop an algorithm that can, based on the
+atmospheric conditions at time of observation and the wavelength dependence of
+seeing, estimate the losses at each wavelength bin. We start with the DIMM
+seeing \citep{Sarazin1990}, which measures the seeing FWHM at zenith. We correct
+the zenith seeing by the airmass of the observational sightline, which gives us
+the effective seeing, delivered outside the slit at 5000 \AA. We can then use
+that the seeing $\sim \lambda^(1/5)$ to obtain a measure of the seeing across
+the entire wavelength coverage of X-shooter. At each wavelength bin of the
+spectra, we simulate a synthetic 2D seeing PSF based on a normalised, 2D Moffat
+profile which we integrate over the used slit width. The inverse of the fraction
+of the light that enters the slit, can then be used to correct for the estimated
+slit losses at each wavelength bin.
+
+To assess the accuracy of the flux-calibration, after the slit loss correction,
+we compare a subset of afterglow spectra with photometry derived from the
+acquisition camera. The acquisition camera takes a short image before
+integration and this the delay between the image and the spectroscopic
+integration is negligible. We select GRBs at DEC > 30~$^\circ$ to ensure
+SDSS coverage, which allows good calibration of the photometric zero point,
+acquisition camera magnitude brighter than 21.5 mag, and delay-time between GRB
+trigger and spectroscopic observation < 15 hours. These cuts leaves 27
+afterglow spectra covering the entire temporal span of the programme. Using this
+set we can make a test of the final flux-calibration accuracy.
+
+To compare the acquisition camera photometry with the spectra, we calculate the
+synthetic magnitude of the spectrum, in the used passband, along with the
+associated error. These two measures should be the same, modulo systematic
+differences affecting the spectra - i.e. bag pixels/regions etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Science data products} \label{products}
@@
 sample of GRBs. The imposed selection criteria (see Sect. \ref{samplecrit})
-ensure that the GRBs entering our homogeneous sample, fairly represent the
+ensure that the GRBs entering our homogeneously selected sample, fairly represent the
 underlying population. The redshift distribution of such a sample holds valuable
@@
 \begin{acknowledgements}
-Much of the work done here would not have been done without Neil Gehrels, who has now passed away.
-We owe him a large debt for his invaluable work with the \swift~satellite.
-We also want to remember Javier Gorosabel, who was taken from us too early.
+Much of the work done here would not have been done without Neil Gehrels, who
+has now passed away. We owe him a large debt for his invaluable work with the
+\swift~satellite. We also want to remember Javier Gorosabel, who was taken from
+us too early. 
 %
@@
 RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
-Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
+Creators and support from ASI (Italian Space Agency) through the Contract n.
+2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD
+project (grant agreement n. 654215).
 %
@@
 %
-SDV acknowledges the support of the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro
+SDV acknowledges the support of the French National Research Agency (ANR) under
+contract ANR-16-CE31-0003 BEaPro
 %
-DM acknowledges support from the Instrument Center for Danish Astrophysics (IDA).
+DM acknowledges support from the Instrument Center for Danish Astrophysics
+(IDA).
 %
@@
 %
-This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}.
-The analysis and plotting was achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy, and Scipy \citep{scipy, VanderWalt2011}, along with other community-developed packages.
+This research made use of Astropy, a community-developed core Python package for
+Astronomy \citep{TheAstropyCollaboration2013}. The analysis and plotting was
+achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy,
+and Scipy \citep{scipy, VanderWalt2011}, along with other community-developed
+packages.
 %

### 2018-10-04 document/XSGRB_sample_rev1.tex
Adding some referee response

@@
 
-Using the masked spectrum, the entire process is then repeated 500 times where
-the final continuum estimate is the mean of the continuum realizations and the
-associated error estimate is the standard deviation. This error reflects the
-stability of the algorithm across the spectrum. An example of the performance is
-shown in Fig. \ref{fig:spectrum}. In the \lya~forest, where there is very little
-flux at the continuum level, the performance of the normalization algorithm
-depends on the continuum coverage redwards of the \lya~line. In some cases, very
-little continuum is contained in a single arm and a manual continuum estimate is
-provided, similar to what is done in \citet{Lopez2016}. In these cases the
-continuum error is set to 10 per cent. The code for the continuum estimate is
-released along with the paper at
-\url{https://github.com/jselsing/XSGRB-sample-paper}.
+%Using the masked spectrum, the entire process is then repeated 500 times where
+%the final continuum estimate is the mean of the continuum realizations and the
+%associated error estimate is the standard deviation. This error reflects the
+%stability of the algorithm across the spectrum. An example of the performance is
+%shown in Fig. \ref{fig:spectrum}. In the \lya~forest, where there is very little
+%flux at the continuum level, the performance of the normalization algorithm
+%depends on the continuum coverage redwards of the \lya~line. In some cases, very
+%little continuum is contained in a single arm and a manual continuum estimate is
+%provided, similar to what is done in \citet{Lopez2016}. In these cases the
+%continuum error is set to 10 per cent. The code for the continuum estimate is
+%released along with the paper at
+%\url{https://github.com/jselsing/XSGRB-sample-paper}.
 
@@
 0\farc9 slit, only 76.1 per cent of the light will enter the slit. This slitloss
-is not corrected by the
-flux calibration, as flux standard stars are always observed with a 5 \arcsec
-slit. As a consequence, almost all spectra will, without slitloss correction,
-underestimate the flux by some fraction. Additionally, slitlosses are wavelength
-dependent due to the wavelength dependence of seeing \citep{Boyd1978} and as the
-seeing changes across the slit in wavelength, so does the slitloss.
+is not corrected by the flux calibration, as flux standard stars are always
+observed with a 5~\arcsec slit. As a consequence, almost all spectra will,
+without slitloss correction, underestimate the flux by some fraction.
+Additionally, slitlosses are wavelength dependent due to the wavelength
+dependence of seeing \citep{Boyd1978} and as the seeing changes across the slit
+in wavelength, so does the slitloss. This artificially introduces a change in
+continuum slope.
 
@@
 the effective seeing, delivered outside the slit at 5000 \AA. We can then use
-that the seeing $\sim \lambda^(1/5)$ to obtain a measure of the seeing across
-the entire wavelength coverage of X-shooter. At each wavelength bin of the
-spectra, we simulate a synthetic 2D seeing PSF based on a normalised, 2D Moffat
-profile which we integrate over the used slit width. The inverse of the fraction
-of the light that enters the slit, can then be used to correct for the estimated
-slit losses at each wavelength bin.
+that the seeing scales as $\sim \lambda^{-1/5}$ to obtain a measure of the
+seeing across the entire wavelength coverage of X-shooter. At each wavelength
+bin of the spectra, we simulate a synthetic 2D seeing PSF based on a normalised,
+2D Moffat profile which we integrate over the used slit width. The inverse of
+the fraction of the light that enters the slit, can then be used to correct for
+the estimated slit losses at each wavelength bin.
 
@@
 acquisition camera. The acquisition camera takes a short image before
-integration and this the delay between the image and the spectroscopic
-integration is negligible. We select GRBs at DEC > 30~$^\circ$ to ensure
-SDSS coverage, which allows good calibration of the photometric zero point,
-acquisition camera magnitude brighter than 21.5 mag, and delay-time between GRB
-trigger and spectroscopic observation < 15 hours. These cuts leaves 27
-afterglow spectra covering the entire temporal span of the programme. Using this
-set we can make a test of the final flux-calibration accuracy.
+integration causing the delay between the image and the spectroscopic
+integration is negligible. The acquisition camera suffer from varying imaging
+quality, lacking calibration files, and regions of reduced thoughput due to
+contaminants on the CCD. The derived photometry therefore suffers from
+systematics that can be hard to correct for. Because the afterglow brightness
+changes rapidly with time, no other imaging are sufficiently close in time to
+use.
+
+We select all GRBs for which the aquisition camera used a Sloan filter
+\citep{Gunn1998} and is covered by either SDSS \citep{Abolfathi2018} or
+Pan-STARRS \citep{Chambers2016}. Additionally we select acquisition camera
+magnitude brighter than 21.5 mag, and delay-time between GRB trigger and
+spectroscopic observation < 10 hours. These cuts leaves 7 afterglow spectra, for
+which a decent calibration of the photometric zero point can be derived. Using
+this set we can make a test of the final flux-calibration accuracy.
 

### 2018-10-12 document/referee_response.txt
Finishing referee response

@@
-This report presents a review of the article "The X-shooter GRB 
-afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
-authors present and make available to the community a complete sample 
-of GRB afterglows observed with the X-Shooter spectrograph on the 
-VLT. This dataset is remarkable in several respect (>100 spectra 
-available taken over 8 years) and some of these have not be published 
-elsewhere. However, there are a number of important issues regarding 
-both the scientific measurements and their presentation which are 
-detailed below. These concerns need to be addressed before the paper 
-can be considered for publication. More details on these points 
-together with other comments are provided below.
+# 
+# ------------------- Referee report
+# 
+# This report presents a review of the article "The X-shooter GRB
+# afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The
+# authors present and make available to the community a complete sample
+# of GRB afterglows observed with the X-Shooter spectrograph on the VLT.
+# This dataset is remarkable in several respect (>100 spectra available
+# taken over 8 years) and some of these have not be published elsewhere.
+# However, there are a number of important issues regarding both the
+# scientific measurements and their presentation which are detailed
+# below. These concerns need to be addressed before the paper can be
+# considered for publication. More details on these points together with
+# other comments are provided below.
+# 
+# # 
+# # # 
+# Major Concerns:
+# 
+# - One aspect of these data that needs more emphasis is the intervening
+# absorbers. The data presented in this paper present a remarkable
+# number of metal and HI intervening absorbers as can be judged from
+# reading the appendix. These deserve much more attention, including a
+# table listing their redshifts, column density measurements when
+# available, metal lines detected, etc. A figure with their redshift
+# distribution would also be very welcome. In addition, it would be
+# interesting to compare the properties of this population with the
+# absorbers towards background quasars. 
+#
 
-Major Concerns:
-
-# - One aspect of these data that needs more emphasis is the 
-# intervening absorbers. The data presented in this paper present a 
-# remarkable number of metal and HI intervening absorbers as can be 
-# judged from reading the appendix. These deserve much more attention, 
-# including a table listing their redshifts, column density 
-# measurements when available, metal lines detected, etc. A figure 
-# with their redshift distribution would also be very welcome. In 
-# addition, it would be interesting to compare the properties of this 
-# population with the absorbers towards background quasars.
-
-While we agree that the subject of intervening absorbers is very interesting and
-important, the main focus of this paper is to present the technicalities of the
-sample and the data reduction. In a series of fourth-coming papers, we will
-investigate sub-themes independently. On the subject of intervening absorbers,
-already published is Christensen et al., A&A, 608A, 84C, 2017,
-(arXiv:1709.01084) and we also have de Ugarte Postigo et al. (in prep).
+This paper is meant as a general introduction of the XSGRB legacy sample
+and constitutes the first in a series of papers. While we certainly
+agree that the issue of intervening absorbers is an interesting one, we
+feel that it is best treated in a dedicated publication, which we have
+already published Christensen et al., A&A, 608A, 84C, 2017,
+(arXiv:1709.01084) and are working on de Ugarte Postigo et al. in
+prep.). This paper will contain the complete list of intervening
+systems, as well as linelists, redshifts, EWs, etc..
 
@@
 # misleading. The spectrograph used to make the follow-up is the same
-# for all afterglows, but the SNR (due to a combination of brightness 
-# of the source and time of observations after the burst) and the 
+# for all afterglows, but the SNR (due to a combination of brightness of
+# the source and time of observations after the burst) and the
 # rest-frame wavelength coverage vary enormously. While remarkable in
@@
 
-We completely agree and have rephrased the manuscript to reflect this change.
+The referee is correct here. Our statement about homogeneity concerns
+only the selection, not the outcome of our observations. We have
+attempted to reword anywhere relevant and we added an explicit caveat on
+the resulting S/N homogeneity in paragraph §2.1.
 
-# - There are a great many details provided about the data reduction, some of 
-# which the reader fails to relate to the final scientific goals. For 
-# example, a correction is applied to correct offsets in the 
+# - There are a great many details provided about the data reduction,
+# some of which the reader fails to relate to the final scientific
+# goals. For example, a correction is applied to correct offsets in the
 # wavelength calibration, which presumably mostly impact the redshift
@@
 # determination accuracy? Similarly, for the telluric correction. What
-# is the impact of this on the final science? Conversely, a 
-# description of the flux calibration error estimate would have been 
-# welcome.
+# is the impact of this on the final science? Conversely, a description
+# of the flux calibration error estimate would have been welcome.
+
+Concerning the wavelength calibration improvement, we have quantified
+the magnitude of the improvement, and explained its relevance (§3.2). We
+have also added a short paragraph to §3.6 explaining the impact of the
+telluric correction on the inferred results.
+
+Regarding the flux calibration accuracy, we have added a new section
+($3.5). This includes a new way to correct for slit losses and tries to
+assess the flux accuracy of the spectra presented. The paper has
+significantly benefited from this suggestion.
 
@@
 
+This is certainly true, when comparing QSO DLAs to GRB DLA. We have
+updated the figure and text surrounding Fig. 10 to reflect this. For
+Fig. 9, where we are comparing GRB DLAs to GRB DLA, we have retained the
+original values for N_HI.
+
 # - Another concern is to do with using absorption features to measure
-# the redshift. Indeed, one cannot exclude that the absorption is not 
-# at the position of host galaxy but lying in front of it. For this 
-# reason, the redshifts provided should be seen as upper limits.
+# the redshift. Indeed, one cannot exclude that the absorption is not at
+# the position of host galaxy but lying in front of it. For this reason,
+# the redshifts provided should be seen as upper limits. 
+# 
+# 
+# 
+# 
+# # 
+# Other Concerns:
+# 
+# - the links to the actual data seem to be inactive. It is important to
+# be able to review the material which is made available to the
+# community.
 
-Other Concerns:
+Completely true. These are now active. Especially the ESO phase 3
+material have taken some time to get through. Apologies.
 
-- the links to the actual data seem to be inactive. It is important to 
-be able to review the material which is made available to the 
-community.
+# - caption of Fig. 3 is cut-out
 
-- caption of Fig. 3 is cut-out
+This is now fixed.
 
-- Fig. 5: please label the y-axis
+# - Fig. 5: please label the y-axis
 
-- Fig. 10: caption is cut-out again 
-GRB150403A: could the continuum be lower? 
-GRB150915A: is this fit significant?
+Done, thanks.
 
-- appendix A: refer to the spectral PSF in the title?
+# - Fig. 10: caption is cut-out again
 
-- appendix B: science-verified, Paranalization: 
-perhaps avoid this ESO jargon?
+Fixed, thanks.
 
-typos: 
-- pt. 4 p. 3: Paranl 
-- p. 3: exampl 
-- p. 3: sorrounding 
-- p. 9: curios 
-- there are numerous typos in the appendix which I will not review 
-here. Please make sure to double check spelling there.
+# GRB150403A: could the continuum be lower?
 
-I would be happy to review an updated version of the manuscript where 
-the above comments are taken into account.
+We have remeasured the N_HI for this burst. It does not change the
+inferred N_HI value, but the continuum now better matches the
+measurement.
+
+# GRB150915A: is this fit significant?
+
+We agree this could be misinterpreted. This bursts has Lyman alpha
+emission in the trough, as indicated in Tab. 4. We have now marked all
+the bursts affect by Lyman alpha emission.
+
+# - appendix A: refer to the spectral PSF in the title?
+
+This has now been reworded.
+
+# - appendix B: science-verified, Paranalization: perhaps avoid this ESO
+# jargon?
+
+We have tried to weed out the use of these types of words.
+
+# typos: - pt. 4 p. 3: Paranl - p. 3: exampl - p. 3: sorrounding - p. 9:
+# curios - there are numerous typos in the appendix which I will not
+# review here. Please make sure to double check spelling there.
+
+Thanks for noting these. They have been fixed.
+
+# I would be happy to review an updated version of the manuscript where
+# the above comments are taken into account.
+# 
+# -----------------------

### 2018-10-12 document/XSGRB_sample_rev1.tex
Finishing referee response

@@
 \input{definitions.tex}
+
+\usepackage{caption}
 
@@
 light from the neutron star merger in the shape of GW170817/GRB170817A/AT2017gfo
-\citep{LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
+\citep{2017ApJ...848L..14G, 2017ApJ...848L..15S, 2017ApJ...848L..16S, 2017ApJ...848L..17C, 2017ApJ...848L..18N, 2017ApJ...848L..19C, 2017ApJ...848L..27T, 2017ApJ...850L...1L, 2017Natur.551...64A, 2017Natur.551...67P, 2017Natur.551...71T, 2017Natur.551...75S, 2017Sci...358.1583K, LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
 
@@
 	\item The elevation of the source in the sky is $> 22^\circ$ (both at trigger time and 15 min after).
-	\item The Sun elevation from Paranl is $< -12^\circ$ (both at trigger time and 15 min after).
+	\item The Sun elevation from Paranal is $< -12^\circ$ (both at trigger time and 15 min after).
 	\item Galactic $A_V \leq 1.0$~mag according to the maps of \citet{Schlegel1998}.
@@
 the GRB lifetime, and it exploits the extreme brightness of early afterglows. A
-pivotal exampl is the study of temporal variability of GRB afterglow absorption
-systems due to effect of the GRB itself on the sorrounding medium \citep[e.g.,
+pivotal example is the study of temporal variability of GRB afterglow absorption
+systems due to effect of the GRB itself on the surrounding medium \citep[e.g.,
 see][]{Dessauges-Zavadsky2006, Vreeswijk2007, DElia2009, Vreeswijk2013}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Slit loss correction and flux calibration accuracy} \label{fluxes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For the majority of the observations presented in this sample, the slit width of
+the target observations is less than the seeing PSF. This means that a fraction
+of the incoming light is lost outside the slit. As an example, for atmospheric
+conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a
+0\farc9 slit, only 76.1 per cent of the light will enter the slit. This slit loss
+is not corrected by the flux calibration, as flux standard stars are always
+observed with a 5\arcsec{} slit. As a consequence, almost all spectra will,
+without slit loss correction, underestimate the flux by some fraction.
+Additionally, slit losses are wavelength dependent due to the wavelength
+dependence of seeing \citep{Boyd1978} and as the seeing changes across the slit
+in wavelength, so does the slit loss. This artificially introduces a change in
+continuum slope.
+
+\begin{figure}[!t]
+	\centerline{\includegraphics[width=\columnwidth]{figures/flux_accuracy}}
+\caption{The left panel shows the acquisition camera magnitudes as a function
+	of synthetic photometry derived from the spectra. The blue, dashed line
+	represent equality between the two measures. The right panel shows a histogram
+	of the differences between the two. The vertical lines in the histogram
+	indicates the 16th, 50th, and 84th percentiles of the distribution ($\Delta
+	\mathrm{mag} = 0.07 ^{+0.51} _{-0.45} $). This shows the accuracy of the flux
+	calibration. The colouring of the data-points shows the photometric
+	passband used by the acquisition camera, as indicated by the labels. The $R$- and $I$- band magnitudes have been calibrated against the USNO-B1 catalog,
+	which has a poor photometric accuracy; this contributes significantly to the  observed scatter, and for this reason we indicate these data with a light color.} \label{fig:flux_acq}
+\end{figure}
+
+To correct for the slitlosses, we develop an algorithm that can, based on the
+atmospheric conditions at time of observation and the wavelength dependence of
+seeing, estimate the losses at each wavelength bin. We start with the DIMM
+seeing \citep{Sarazin1990}, which measures the seeing FWHM at zenith. We correct
+the zenith seeing for the airmass of the observational sightline, which gives us
+the effective seeing, delivered outside the slit at 5000 \AA. We can then use
+that the seeing scales as $\sim \lambda^{-1/5}$ to obtain a measure of the
+seeing across the entire wavelength coverage of X-shooter. At each wavelength
+bin of the spectra, we simulate a synthetic 2D seeing PSF based on a normalised,
+2D Moffat profile which we integrate over the used slit width. This can be used to correct for the estimated slit losses at each wavelength bin.
+
+To assess the accuracy of the flux-calibration, after the slit loss correction,
+we compare a subset of afterglow spectra with photometry derived from the
+acquisition camera. The acquisition camera takes a short image just before
+the spectroscopic integration, which only introduces a minimal delay and allows us to secure photometry for the vast majority of events.
+As photometric flux standards are rarely acquired, we resort to sky surveys such as the USNO catalog (for the $R$ and $I$ bands; \citealt{Monet2003}) or SDSS and Pan-STARRS ($griz$; \citealt{Abolfathi2018,Chambers2016}), where available.
+
+To compare the photometry with the spectral flux calibration,
+we calculate synthetic magnitudes from the spectrum, in the used passbands,
+along with the associated error. Using this set we can make a test of the final flux-calibration accuracy, which
+we show in Fig.~\ref{fig:flux_acq}. As can be seen, there is a general agreement
+between the flux-calibration and the derived acquisition magnitudes, although
+with significant scatter. The right panel in Fig.~\ref{fig:flux_acq} shows a
+histogram of the differences, which can be summarized with $\Delta \mathrm{mag}
+= 0.07 ^{+0.51} _{-0.45} $, where the errors are the 16th and 84th percentiles of
+the distribution.
+The scatter is smaller if we exclude measurements calibrated against the USNO catalog, which is known to have a poor accuracy, also because, given the small field-of-view, we often need to rely of faint reference stars.
+
+Similar to what is found in \citet{Lopez2016}, there is in some cases a mismatch
+in the flux-calibration between the spectroscopic arms. It is outside the scope
+of this work to correct for these offsets on an individual burst basis.The flux
+calibration of the spectra presented here shows a general good agreement with
+photometry although with some scatter. The spectral shape within the
+individual arms is correct, but the absolute scaling might in some cases be
+discrepant.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Telluric correction} \label{tell_corr}
@@
 \begin{figure*}
-	\centerline{\includegraphics[width=\linewidth]{figures/GRB121024A.pdf}}
+	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
 	\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Slitloss correction and flux calibration accuracy} \label{fluxes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For the majority of the observation presented in this sample, the slit width of
-the target observations is less than the seeing PSF. This means that a fraction
-of the incoming light is lost outside the slit. As an example, for atmospheric
-conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a
-0\farc9 slit, only 76.1 per cent of the light will enter the slit. This slitloss
-is not corrected by the flux calibration, as flux standard stars are always
-observed with a 5~\arcsec slit. As a consequence, almost all spectra will,
-without slitloss correction, underestimate the flux by some fraction.
-Additionally, slitlosses are wavelength dependent due to the wavelength
-dependence of seeing \citep{Boyd1978} and as the seeing changes across the slit
-in wavelength, so does the slitloss. This artificially introduces a change in
-continuum slope.
-
-To correct for the slitlosses, we develop an algorithm that can, based on the
-atmospheric conditions at time of observation and the wavelength dependence of
-seeing, estimate the losses at each wavelength bin. We start with the DIMM
-seeing \citep{Sarazin1990}, which measures the seeing FWHM at zenith. We correct
-the zenith seeing by the airmass of the observational sightline, which gives us
-the effective seeing, delivered outside the slit at 5000 \AA. We can then use
-that the seeing scales as $\sim \lambda^{-1/5}$ to obtain a measure of the
-seeing across the entire wavelength coverage of X-shooter. At each wavelength
-bin of the spectra, we simulate a synthetic 2D seeing PSF based on a normalised,
-2D Moffat profile which we integrate over the used slit width. The inverse of
-the fraction of the light that enters the slit, can then be used to correct for
-the estimated slit losses at each wavelength bin.
-
-To assess the accuracy of the flux-calibration, after the slit loss correction,
-we compare a subset of afterglow spectra with photometry derived from the
-acquisition camera. The acquisition camera takes a short image before
-integration causing the delay between the image and the spectroscopic
-integration is negligible. The acquisition camera suffer from varying imaging
-quality, lacking calibration files, and regions of reduced thoughput due to
-contaminants on the CCD. The derived photometry therefore suffers from
-systematics that can be hard to correct for. Because the afterglow brightness
-changes rapidly with time, no other imaging are sufficiently close in time to
-use.
-
-We select all GRBs for which the aquisition camera used a Sloan filter
-\citep{Gunn1998} and is covered by either SDSS \citep{Abolfathi2018} or
-Pan-STARRS \citep{Chambers2016}. Additionally we select acquisition camera
-magnitude brighter than 21.5 mag, and delay-time between GRB trigger and
-spectroscopic observation < 10 hours. These cuts leaves 7 afterglow spectra, for
-which a decent calibration of the photometric zero point can be derived. Using
-this set we can make a test of the final flux-calibration accuracy.
-
-To compare the acquisition camera photometry with the spectra, we calculate the
-synthetic magnitude of the spectrum, in the used passband, along with the
-associated error. These two measures should be the same, modulo systematic
-differences affecting the spectra - i.e. bag pixels/regions etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Science data products} \label{products}
@@
 fulfilling the sample criteria have been followed up due to interesting
-characteristics, e.g., curios properties of their light curves, their
+characteristics, e.g., curious properties of their light curves, their
 brightness, etc. These bursts are not discussed as part of the investigation of
@@
 \citep{Jakobsson2012, Perley2016a}.
+
+A possible complication is that a large fraction of the redshifts are 
+measured from absorption systems, which in principle only provide lower 
+limits to the GRB redshift. While for individual systems we cannot rule 
+out a low-column density host which is missed, especially in low S/N 
+spectra (see for example the discussion by 
+\citealt{2008ApJ...688..470P, Perley2017a}), we argue that this issue is very 
+unlikely to affect significantly our sample.
+
+First, numerous works, including this one, have found that GRB 
+absorption systems have much higher column densities than randomly 
+selected sight lines 
+\citep{2001ApJ...549L.209G,Prochaska2007, Fynbo2009}. 
+This fact itself shows that the bulk of GRB absorption lines are not due 
+to random intervening systems, but are peculiar to the GRB host. One 
+second line of evidence comes from the systematic detection of nebular 
+emission lines at the same redshift as the absorption system. 
+\citet{Kruhler2015} report a 94\% detection rate of emission 
+lines at $z < 1$ (where sensitivity is not an issue, even in the 
+presence of a bright afterglow), always in agreement with absorption. 
+Besides, it is extremely rare to detect emission from QSO intervening 
+systems (see e.g. \citealt{2010MNRAS.408.2128F} for a discussion), which 
+again indicates that GRB absorption systems are native to their hosts. 
+An even more direct determination comes from the detection, in several 
+cases, of fine-structure and metastable lines (for example from Fe and 
+Ni), which are exclusively detected in GRB environments (and have been 
+directly connected to the UV field generated by the GRB itself; e.g. 
+\citealt{2006ApJ...648...95P, Vreeswijk2007, DElia2009}). 
+Finally, at $z \gtrsim 2$ (where Ly$\alpha$ enters the observed range) 
+the onset of the Ly$\alpha$ forest also provides an \textit{upper} limit 
+on the GRB redshift, which again is always consistent with the 
+absorption line redshift. The whole set of these arguments indicates 
+that in only very few cases, if any at all, is a misidentification of 
+the redshift with the intervening system a concrete possibility.
 
@@
 star-forming regions in their hosts.
+When comparing the GRB DLAs and and QSO DLAs we correct for the fact that, on average, the GRB light only goes though half the DLA.
+In Fig.
+\ref{fig:NH_dist}, we have therefore multiplied the inferred $N_{\mathrm{HI}}$ values by 2 \citep{Prochaska2008}.
 
@@
 	clearly visible in the hydrogen column densities, as previously also noted in
-	\citet{Fynbo2009}.} \label{fig:NH_dist}
+	\citet{Fynbo2009}. To correct for geometric differences between GRB DLAs and QSO DLAs, we have multiplied the inferred GRB DLA $N_{\mathrm{HI}}$ by 2.} \label{fig:NH_dist}
 \end{figure}
@@
 	\centering
-	\includegraphics[page=1, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=1, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption{Measurements of the hydrogen column densities for all bursts with a
@@
 	\centering
-	\includegraphics[page=2, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=2, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=3, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=3, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=4, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=4, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=5, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=5, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 	\centering
-	\includegraphics[page=6, width=\linewidth]{figures/HI_measurements.pdf}
+	\includegraphics[page=6, width=0.85\linewidth]{figures/HI_measurements.pdf}
 	\caption*{Fig. \ref{fig:HI1}. continued.}
@@
 
-\section{The complex error function and the Voigt profile} \label{voigt}
+\section{The complex error function and the spectral PSF} \label{voigt}
 When modeling the spectral PSF, we need to evaluate the Voigt-profile.
@@
 
-Observed during paranalization of the instrument, these data forms the basis of
+Observed during the verification of the instrument, these data forms the basis of
 GCN 15571 \citep{GCN15571}, but are not published elsewhere. Observations began
@@
 detection of \mgii, \mgi, \SIii, \feii, \aliii. Because this burst was observed
-before the instrument was science-verified, it does not enter into the
+before the instrument was finished completely, it does not enter into the
 statistical sample.
@@
 
-Observed during the first science verification period and was reported in GCN
+Observed during the first instrument testing period and was reported in GCN
 9761 \citep{GCN.9761} and is additionally used as the basis for the master
@@
 
-Obtained during the second science
-verification period, this dataset forms the basis of GCN 9942 \citep{GCN.9942} and is
-additionally published in \citet{DElia2010}. Spectroscopic integration started
-22 hours after the BAT trigger and from the acquisition camera the optical
-afterglow is measured to R = 17.9 mag at the beginning of the observations which
-causes a strong continuum to be seen in all arms. An absorption trough due to
-\lya~is clearly visible along with numerous metal resonance lines \civ, \SIii,
-\SIii$^*$ \feii, \mgii, all at $z = 2.106$. Because this burst was observed
-before the instrument was science-verified, it does not enter into the
-statistical sample.
+Obtained during the second instrument testing period, this dataset forms the
+basis of GCN 9942 \citep{GCN.9942} and is additionally published in
+\citet{DElia2010}. Spectroscopic integration started 22 hours after the BAT
+trigger and from the acquisition camera the optical afterglow is measured to R =
+17.9 mag at the beginning of the observations which causes a strong continuum to
+be seen in all arms. An absorption trough due to \lya~is clearly visible along
+with numerous metal resonance lines \civ, \SIii, \SIii$^*$ \feii, \mgii, all at
+$z = 2.106$. Because this burst was observed before the instrument was
+science-verified, it does not enter into the statistical sample.
 
@@
 
-The first burst observed during normal operation after science verification was
-completed and the first burst that enters the statistical sample. These data are
-the basis for GCN 10042 \citep{GCN10042} and is published in
-\citet{Wiersema2012}. With a bright afterglow and a rapid follow-up, this
-spectrum is of pristine quality. The afterglow continuum is bright throughout
-all spectroscopic arms which allows the ready detection of \alii, \aliii, \feii,
-\mnii, \mgii, \mgi, and \caii~- all located at $z = 0.971$, setting it as the
-redshift of the GRB.
+The first burst observed during normal operation and the first burst that enters
+the statistical sample. These data are the basis for GCN 10042 \citep{GCN10042}
+and is published in \citet{Wiersema2012}. With a bright afterglow and a rapid
+follow-up, this spectrum is of pristine quality. The afterglow continuum is
+bright throughout all spectroscopic arms which allows the ready detection of
+\alii, \aliii, \feii, \mnii, \mgii, \mgi, and \caii~- all located at $z =
+0.971$, setting it as the redshift of the GRB.
 

### 2018-10-12 py/NH_distribution_plot.py

@@ -10,7 +10,7 @@ import matplotlib as mpl
 from astropy.io import fits
 import pandas as pd
 # import matplotlib.pyplot as pl
-import healpy as hp
+# import healpy as hp
 # import seaborn as sns; sns.set_style('ticks')
 import numpy as np
 from scipy import stats
@@ -43,7 +43,7 @@ def main():
 
     # Read in GRB NH
     burst_table = pd.read_csv("../data/Burst list - HI columns.csv")
-    name, z, NH, NHe, pub = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values, burst_table["N_H_err"].values, burst_table["Sample"].values
+    name, z, NH, NHe, pub = np.array([ii[3:] for ii in burst_table["GRB"].values]), burst_table["z"].values, burst_table["N_H"].values+ np.log10(2), burst_table["N_H_err"].values, burst_table["Sample"].values
     XSGRB = (pub == "Selsing2017")
     # print(np.sum((~XSGRB).astype("int")))
     # print(len(name))

### 2018-10-12 py/Sample_completeness.py

@@ -246,7 +246,7 @@ def main():
         # ax.set_ylim(0, 1)
 
 
-
+    ax1.set_ylabel("Cumulative Fraction")
     ax1.set_xlabel(r"$\log(15-150~\mathrm{keV}~\mathrm{Fluence}/\mathrm{erg}~\mathrm{cm}^{-2})$")
     ax2.set_xlabel(r"$\log(0.3-10~\mathrm{keV}~\mathrm{Flux}/\mathrm{erg}~\mathrm{cm}^{-2}~\mathrm{s}^{-1})$")
     ax3.set_xlabel(r"$\log(N_{\mathrm{H, X}}/\mathrm{cm}^{-2})$")

### 2018-10-12 document/XSGRB_sample_rev1.tex
Adding ref

@@
 light from the neutron star merger in the shape of GW170817/GRB170817A/AT2017gfo
-\citep{2017ApJ...848L..14G, 2017ApJ...848L..15S, 2017ApJ...848L..16S, 2017ApJ...848L..17C, 2017ApJ...848L..18N, 2017ApJ...848L..19C, 2017ApJ...848L..27T, 2017ApJ...850L...1L, 2017Natur.551...64A, 2017Natur.551...67P, 2017Natur.551...71T, 2017Natur.551...75S, 2017Sci...358.1583K, LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
+\citep{2017ApJ...848L..14G, 2017ApJ...848L..15S, 2017ApJ...848L..16S, 2017ApJ...848L..17C, 2017ApJ...848L..18N, 2017ApJ...848L..19C, 2017ApJ...848L..27T, 2017ApJ...850L...1L, 2017Natur.551...64A, 2017Natur.551...67P, 2017Natur.551...71T, 2017Natur.551...75S, 2017Sci...358.1583K, 2017Sci...358.1556C, LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
 

### 2018-10-16 document/referee_response.txt
Adding co-author suggestions

@@
 (arXiv:1709.01084) and are working on de Ugarte Postigo et al. in
-prep.). This paper will contain the complete list of intervening
+prep.). These papers contain the complete list of intervening
 systems, as well as linelists, redshifts, EWs, etc..
@@
 # the redshifts provided should be seen as upper limits.
-# 
-# 
-# 
-# 
+
+We agree with the referee that, in principle, our values should be
+considered as lower limits. We contend, though, that in the large
+majority of cases the values we measured correspond to the actual
+redshift of the GRB. We have added a paragraph in the text explaining
+why this is.
+
 # #

### 2018-10-16 document/authorlist.tex
Last commit before Arxiv submission
Adding co-author suggestions

@@
 \author{
-J.~Selsing \inst{1}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}
-\addtocounter{footnote}{2}
+J.~Selsing \inst{1}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}\addtocounter{footnote}{2}
 \and D.~Malesani \inst{1, 2, 3}\fnmsep\thanks{On-call observer}
@@
 }
-
 \institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
-%\and Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
 \and The Cosmic Dawn Center, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
@@
 \and APC, Astroparticule et Cosmologie, Universit\'e Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cit\'e, 10, Rue Alice Domon et L\'eonie Duquet, 75205, Paris Cedex 13, France
-\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
+\and Max-Planck Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e 1, 85748 Garching, Germany
 \and INAF - Osservatorio Astronomico di Roma, Via Frascati 33, 00078, Monte Porzio Catone (Roma), Italy
@@
 \and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain. % 10
-%
 \and INAF - Osservatorio Astronomico di Brera, via Bianchi 46, 23807, Merate (LC), Italy
 \and Space Science Data Center - Agenzia Spaziale Italiana, via del Politecnico, s.n.c., I-00133, Roma, Italy
-\and European Southern Observatory, Karl-Schwarzschild Str. 2, 85748 Garching bei M\”unchen, Germany
+\and European Southern Observatory, Karl-Schwarzschild Str. 2, 85748 Garching bei M\"unchen, Germany
 \and GEPI, Observatoire de Paris, PSL University, CNRS, 5 place Jules Janssen 92195, Meudon, France
@@
 \and Department of Physics, University of Warwick, Coventry, CV4 7AL, UK % 20
-%
 \and IASF/INAF Bologna, via Piero Gobetti 101, 40129 Bologna, Italy
@@
 \and CAS Key Laboratory of Space Astronomy and Technology, National Astronomical Observatories, Chinese Academy of Sciences, Beijing, 100012, P.R. China % 30
-%
 \and Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.

### 2018-10-19 document/XSGRB_sample_rev1.tex
Last revisions before resubmission

--- 2018-10-12 document/XSGRB_sample_rev1.tex
@@
 light from the neutron star merger in the shape of GW170817/GRB170817A/AT2017gfo
-\citep{2017ApJ...848L..14G, 2017ApJ...848L..15S, 2017ApJ...848L..16S, 2017ApJ...848L..17C, 2017ApJ...848L..18N, 2017ApJ...848L..19C, 2017ApJ...848L..27T, 2017ApJ...850L...1L, 2017Natur.551...64A, 2017Natur.551...67P, 2017Natur.551...71T, 2017Natur.551...75S, 2017Sci...358.1583K, LIGOScientificCollaboration2017a, LIGOScientificCollaboration2017}.
+\citep{2017ApJ...848L..14G, 2017ApJ...848L..15S, 2017ApJ...848L..16S,
+	2017ApJ...848L..17C, 2017ApJ...848L..18N, 2017ApJ...848L..19C,
+	2017ApJ...848L..27T, 2017ApJ...850L...1L, 2017Natur.551...64A,
+	2017Natur.551...67P, 2017Natur.551...71T, 2017Natur.551...75S,
+	2017Sci...358.1583K, 2017Sci...358.1556C, Burgess2017, LIGOScientificCollaboration2017a,
+	LIGOScientificCollaboration2017}.
 
@@
 
-	\item GRB trigger by BAT onboard the \swift~satellite
+	\item GRB trigger by BAT on-board the \swift~satellite
 
@@
 Individual bursts vary significantly in brightness, due to intrinsic
-differences in luminosity, dust extinction, and time of observation. The covered
+differences in luminosity, dust extinction, time of observation, etc. The covered
 rest-frame wavelength range is also different for each case. Given that our
@@
 \begin{enumerate}
-	\item The GRB triggered \textit{Swift} onboard.
+	\item The GRB triggered \textit{Swift}.
 	\item The X-ray position must be available less than 1 hr after the GRB.
@@
 was affected by the failing ADC prior to disablement, resulting in a
-lower-than-nominal throughput. To avoid chromatic slit losses due to atmospheric
+lower-than-nominal throughput. To avoid chromatic slit-loss due to atmospheric
 dispersion, nearly all subsequent observations have been carried out at
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Slit loss correction and flux calibration accuracy} \label{fluxes}
+\subsection{Slit-loss correction and flux calibration accuracy} \label{fluxes}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 the target observations is less than the seeing PSF. This means that a fraction
-of the incoming light is lost outside the slit. As an example, for atmospheric
-conditions delivering a seeing PSF with a FWHM of 0\farc9 observed with a
-0\farc9 slit, only 76.1 per cent of the light will enter the slit. This slit loss
-is not corrected by the flux calibration, as flux standard stars are always
-observed with a 5\arcsec{} slit. As a consequence, almost all spectra will,
-without slit loss correction, underestimate the flux by some fraction.
-Additionally, slit losses are wavelength dependent due to the wavelength
-dependence of seeing \citep{Boyd1978} and as the seeing changes across the slit
-in wavelength, so does the slit loss. This artificially introduces a change in
-continuum slope.
+of the incoming light is lost outside the slit. This slit-loss is not corrected
+by the flux calibration, as flux standard stars are always observed with a
+5\arcsec{} slit. As a consequence, almost all spectra will, without slit-loss
+correction, underestimate the flux by some fraction. Additionally, slit-loss
+is wavelength dependent due to the wavelength dependence of seeing
+\citep{Boyd1978} and as the seeing changes across the slit in wavelength, so
+does the slit-loss. This artificially introduces a change in continuum slope.
 
@@
 	of synthetic photometry derived from the spectra. The blue, dashed line
-	represent equality between the two measures. The right panel shows a histogram
-	of the differences between the two. The vertical lines in the histogram
-	indicates the 16th, 50th, and 84th percentiles of the distribution ($\Delta
-	\mathrm{mag} = 0.07 ^{+0.51} _{-0.45} $). This shows the accuracy of the flux
-	calibration. The colouring of the data-points shows the photometric
-	passband used by the acquisition camera, as indicated by the labels. The $R$- and $I$- band magnitudes have been calibrated against the USNO-B1 catalog,
-	which has a poor photometric accuracy; this contributes significantly to the  observed scatter, and for this reason we indicate these data with a light color.} \label{fig:flux_acq}
+	represent equality between the two measures. The colouring of the data-points
+	shows the photometric passband used by the acquisition camera, as indicated by
+	the labels. The $R$- and $I$- band magnitudes have been calibrated against the
+	USNO-B1 catalog, which has a poor photometric accuracy; this contributes
+	significantly to the observed scatter, and for this reason we indicate these
+	data with a light color.  The right panel shows a histogram of the differences
+	between the two. The vertical lines in the histogram indicates the 16th, 50th,
+	and 84th percentiles of the distribution ($\Delta \mathrm{mag} = 0.07 ^{+0.51}
+	_{-0.45} $). This shows the accuracy of the flux calibration.}
+\label{fig:flux_acq}
 \end{figure}
 
-To correct for the slitlosses, we develop an algorithm that can, based on the
+To correct for the slit-loss, we develop an algorithm that can, based on the
 atmospheric conditions at time of observation and the wavelength dependence of
@@
 bin of the spectra, we simulate a synthetic 2D seeing PSF based on a normalised,
-2D Moffat profile which we integrate over the used slit width. This can be used to correct for the estimated slit losses at each wavelength bin.
-
-To assess the accuracy of the flux-calibration, after the slit loss correction,
+2D Moffat profile which we integrate over the used slit width. This can be used to correct for the estimated slit-loss at each wavelength bin.
+
+To assess the accuracy of the flux-calibration, after the slit-loss correction,
 we compare a subset of afterglow spectra with photometry derived from the
-acquisition camera. The acquisition camera takes a short image just before
-the spectroscopic integration, which only introduces a minimal delay and allows us to secure photometry for the vast majority of events.
-As photometric flux standards are rarely acquired, we resort to sky surveys such as the USNO catalog (for the $R$ and $I$ bands; \citealt{Monet2003}) or SDSS and Pan-STARRS ($griz$; \citealt{Abolfathi2018,Chambers2016}), where available.
+acquisition camera. The acquisition camera takes a short image just before the
+spectroscopic integration, which only introduces a minimal delay and allows us
+to secure photometry for the vast majority of events. As photometric flux
+standards are rarely acquired, we resort to sky surveys such as the USNO catalog
+(for the $R$ and $I$ bands; \citealt{Monet2003}), SDSS, or Pan-STARRS ($griz$;
+\citealt{Abolfathi2018,Chambers2016}), where available. The acquisition camera
+is equipped with similar filters, so we can avoid transformation between filter
+curves when comparing catalogues.
 
@@
 The effect of telluric correction varies greatly depending on the wavelength
-region. In the red and ner-infrared areas, it allows to identify features in a
+region. In the red and near-infrared areas, it allows to identify features in a
 credible manner. which would be otherwise hidden among or blended with
@@
 continuum level even in regions of moderately low atmospheric transmission (see,
-e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}). We urge however to use
-caution when deriving fitting lines in regions significantly affected by
-telluric absorption.
+e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}). Nonetheless, we recommend
+caution when fitting lines in regions significantly affected by telluric
+absorption.
 
@@
 
-First, numerous works, including this one, have found that GRB 
-absorption systems have much higher column densities than randomly 
-selected sight lines 
-\citep{2001ApJ...549L.209G,Prochaska2007, Fynbo2009}. 
-This fact itself shows that the bulk of GRB absorption lines are not due 
-to random intervening systems, but are peculiar to the GRB host. One 
-second line of evidence comes from the systematic detection of nebular 
-emission lines at the same redshift as the absorption system. 
-\citet{Kruhler2015} report a 94\% detection rate of emission 
-lines at $z < 1$ (where sensitivity is not an issue, even in the 
-presence of a bright afterglow), always in agreement with absorption. 
-Besides, it is extremely rare to detect emission from QSO intervening 
-systems (see e.g. \citealt{2010MNRAS.408.2128F} for a discussion), which 
-again indicates that GRB absorption systems are native to their hosts. 
-An even more direct determination comes from the detection, in several 
-cases, of fine-structure and metastable lines (for example from Fe and 
-Ni), which are exclusively detected in GRB environments (and have been 
-directly connected to the UV field generated by the GRB itself; e.g. 
-\citealt{2006ApJ...648...95P, Vreeswijk2007, DElia2009}). 
-Finally, at $z \gtrsim 2$ (where Ly$\alpha$ enters the observed range) 
-the onset of the Ly$\alpha$ forest also provides an \textit{upper} limit 
-on the GRB redshift, which again is always consistent with the 
-absorption line redshift. The whole set of these arguments indicates 
-that in only very few cases, if any at all, is a misidentification of 
-the redshift with the intervening system a concrete possibility.
+First, numerous works, including this one, have found that GRB absorption
+systems have much higher column densities than randomly selected sight lines
+\citep{2001ApJ...549L.209G, Savaglio2003,  Prochaska2007, Fynbo2009}. This fact
+itself shows that the bulk of GRB absorption lines are not due to random
+intervening systems, but are peculiar to the GRB host. A second line of evidence
+comes from the systematic detection of nebular emission lines at the same
+redshift as the absorption system. \citet{Kruhler2015} report a 94\% detection
+rate of emission lines at $z < 1$ (where sensitivity is not an issue, even in
+the presence of a bright afterglow), always in agreement with absorption.
+Besides, it is rare to detect emission from QSO intervening systems
+(see e.g. \citealt{2010MNRAS.408.2128F} for a discussion), which again indicates
+that GRB absorption systems are native to their hosts. An even more direct
+determination comes from the detection, in several cases, of fine-structure and
+metastable lines (for example from Fe and Ni), which are exclusively detected in
+GRB environments (and have been directly connected to the UV field generated by
+the GRB itself; e.g. \citealt{2006ApJ...648...95P, Vreeswijk2007, DElia2009}).
+Finally, at $z \gtrsim 2$ (where Ly$\alpha$ enters the observed range) the onset
+of the Ly$\alpha$ forest also provides an \textit{upper} limit on the GRB
+redshift, which again is always consistent with the absorption line redshift.
+The whole set of these arguments indicates that in only very few cases, if any
+at all, is a misidentification of the redshift with the intervening system a
+concrete possibility.
 
@@
 measurements here for completeness. In the compilation of $N_{\mathrm{HI}}$
-measurements towards GRBs in Tanvir et al. (submitted) there are 93 published
+measurements towards GRBs in \citet{Tanvir2017} there are 93 published
 $N_{\mathrm{HI}}$ values, excluding the measurements provided here. We here
@@
 
-In Fig. \ref{fig:NH_dist}, we also show the column density distribution
-for the 12081 quasar absorbers with \nh~$> 20$ from \citet{Noterdaeme2012b}. The
-fact that GRBs are systematically located behind the highest \nh, previously
-noted \citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this
-figure. The reason for this is that quasar sample sight-lines through galaxies
-that are cross-section selected, whereas GRB sight-lines probe the dense,
-star-forming regions in their hosts.
-When comparing the GRB DLAs and and QSO DLAs we correct for the fact that, on average, the GRB light only goes though half the DLA.
-In Fig.
-\ref{fig:NH_dist}, we have therefore multiplied the inferred $N_{\mathrm{HI}}$ values by 2 \citep{Prochaska2008}.
+In Fig. \ref{fig:NH_dist}, we also show the column density distribution for the
+12081 quasar absorbers with \nh~$> 20$ from \citet{Noterdaeme2012b}. The fact
+that GRBs are systematically located behind the highest \nh, previously noted
+\citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this figure. The
+reason for this is that quasar sample sight-lines through galaxies that are
+cross-section selected, whereas GRB sight-lines probe the dense, star-forming
+regions in their hosts. When comparing the GRB DLAs and QSO DLAs we correct for
+the fact that, on average the GRB light only goes through half of the full ISM
+gas column. In Fig. \ref{fig:NH_dist}, we have therefore multiplied the inferred
+$N_{\mathrm{HI}}$ values by 2 \citep{Prochaska2008}.
 
@@
 	green is the kernel density estimate of absorbers in GRB sightlines. Values are
-	taken from the compilation in Tanvir et al. (submitted), along with the new
+	taken from the compilation in \citet{Tanvir2017}, along with the new
 	values presented in this sample. We also show in the red the values derived in
@@
 that was also much advanced by the previous surveys based primarily on
-lower-resolution spectroscopy. GRB afterglow sight-lines are unique in the
-sense that only after observing more than 12000 damped Lyman-$\alpha$ absorbers
-(DLAs) towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ have been
+lower-resolution spectroscopy. GRB afterglow sight-lines are unique in the sense
+that only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
+towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ have been
 identified \citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra,
@@
 \mgii~ absorbers \citep{Christensen2017}, \citet{Arabsalmani2018} on the
-metallicity-scaling relations, and escape of ionizing radiation (Tanvir et al.,
-submitted). A number of additional companion papers are also planned,
+metallicity-scaling relations, and escape of ionizing radiation
+\citep{Tanvir2017} . A number of additional companion papers are also planned,
 investigating the detailed properties of the sample presented here, including
@@
 metallicities and kinematics (Th{\"o}ne et al., in preparation), high ionization
-lines (Heintz et al., in submitted), molecular lines (Bolmer et al., in
-preparation), fine-structure lines (Vreeswijk et al., in preparation), and
-composite GRB afterglow spectrum (Selsing et al., in preparation).
+lines (Heintz et al., in submitted), molecular lines \citep{Bolmer2018},
+fine-structure lines (Vreeswijk et al., in preparation), and composite GRB
+afterglow spectrum (Selsing et al., in preparation).
 
@@
 		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
-		with dashed lines.}
+		with dashed lines. The pressence of \lya, as noted in Tab. \ref{tab:HI}, is indicated with a black arrow.}
 	\label{fig:HI1}
@@
 \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all at a consistent
-redshift. Temporal evolution of the fine structure lines belonging to \feii$^*$
+redshift. Temporal evolution of the fine-structure lines belonging to \feii$^*$
 is found between the epochs.
@@
 and are used in Skuladottir (2010), but not published elsewhere. Observations
-started 4 hours after the \textit{Swift} trigger, totaling 2.4 ks. Absorption
+started 4 hours after the \textit{Swift} trigger, totalling 2.4 ks. Absorption
 features from \mgii~and \feii~in the afterglow continuum are detected at
@@
 spectrum has not otherwise been published previously. The observations were
-carried out in RRM starting 11 min after the GRB trigger. See section \ref{RRM},
+carried out in RRM starting 11 minutes after the GRB trigger. See section \ref{RRM},
 for a description of the RRM scheme. Absorption lines from several ionic species
@@
 spectroscopic observation reveals after heavy binning, a wide, faint trace down
-to $\sim$ 580 nm, offset by 2.5\arcsec~relative to the centering of the slit.
-No good redshift measurement can be inferred from this. We have adopted the
+to $\sim$ 580 nm, offset by 2.5\arcsec~relative to the centring of the slit. No
+good redshift measurement can be inferred from this. We have adopted the
 redshift from GCN 12193 \citep{GCN12193}.
@@
 spectrum is obtained in 2.4 ks spectroscopic integration. The GRB afterglow
-continuum is visible across all three spectrscopic arms of VLT/X-shooter with
+continuum is visible across all three spectroscopic arms of VLT/X-shooter with
 emission lines identified as \oii, \oiii, \ha~all at $z = 1.348$. At the same
@@
 Starting 6.15 hours after the BAT trigger, spectroscopic integration for 4.8 ks
-reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 500 nm. The simultaneous detection of absorption features identified as \lya,
-\SIii, \civ, \alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet,
+reveals a moderate signal-to-noise GRB afterglow continuum, down to $\sim$ 500
+nm. The simultaneous detection of absorption features identified as \lya, \SIii,
+\civ, \alii, \cah, \cak, and \mgii, and emission from the \oiii-doublet,
 securely sets $z = 3.36$ as the redshift of the GRB. These data form the basis
@@
 \oiii~and \ha, all at a common $z = 2.211$, marking it as the redshift of the
-GRB host. No afterglow continuum is detected. This is the highest redshift short
-GRB detected to date.
+GRB host. This is the highest redshift short GRB detected to date. The afterglow
+was detected in X-rays, but no optical- or radio GRB afterglow was detected.
 
@@
 detection of \oii. At this redshift, \hb~and \oii~are located in the gap between
-the VIS and NIR arm and \ha~is located in the middle of the $JH$-bandgap and is
+the VIS and NIR arm and \ha~is located in the middle of the $JH$-band-gap and is
 therefore not detected.
@@
 bluewards, all the way down to the Lyman limit. Metal absorption lines from
-\cii, \SIii, \oi, \feii, \civ, \SIiv, including fine structure transitions
+\cii, \SIii, \oi, \feii, \civ, \SIiv, including fine-structure transitions
 identified as \cii$^*$, \SIii$^*$, \feii$^*$ and metastable \NIii~lines are all
@@
 from \nv, \sii, \SIii, \oi, \civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and
-\mgi. Also fine structure lines from \NIii~and \feii~are excited in the local
+\mgi. Also fine-structure lines from \NIii~and \feii~are excited in the local
 environment of the GRB. Additionally, this spectrum is one of the rare cases in
@@
 \alii, \aliii, \cii, \oi, \civ, and \znii~are all detected along with the
-corresponding fine structure lines (\feii$^*$, \SIii$^*$, \oi$^*$, \oi$^**$,
+corresponding fine-structure lines (\feii$^*$, \SIii$^*$, \oi$^*$, \oi$^**$,
 \cii$^*$), securely anchoring the redshift of the host.
@@
 
-This spectrum has previously been published in \citet{Tanvir2017a}. Starting
+This spectrum has previously been published in \citet{Tanvir2018}. Starting
 18.5 hours after the BAT trigger, the final spectrum is based on $2 \times 4
@@
 high redshift. We here adopt the redshift ($z = 7.84$) suggested by
-\citet{Tanvir2017a}.
+\citet{Tanvir2018}.
 
@@
 A broad absorption feature from Lyman alpha, along with narrow lines from \civ,
-\SIii, \SIiv, \feii, \sii, and \alii, as well as fine structure lines associated
+\SIii, \SIiv, \feii, \sii, and \alii, as well as fine-structure lines associated
 with \SIii$^*$ are all detected at $z = 2.300$, securely setting it as the
@@
 contamination in the slit, the background is slightly over subtracted, causing
-the background to be negative in the center of the \lya~trough.
+the background to be negative in the centre of the \lya~trough.
 
@@
 This spectrum formed the basis of GCN 14493 \citep{GCN14493}, but are not
-published otherwise. A short, 2x600 s spectroscopic integration obtained before
+published otherwise. A short, 2 x 600 s spectroscopic integration obtained before
 twilight, 20.6 hours after the BAT trigger, captures a faint afterglow continuum
@@
 signal-to-noise continuum across all arms. Imprinted on the continuum are
-absorption features identifed as due to \feii~and \mgii~as well as emission
+absorption features identified as due to \feii~and \mgii~as well as emission
 lines from \oii, \hd, \hg, \hb, \oiii, \ha, and \niil. All the lines are measured
@@
 This spectrum has previously been published in \citet{Kruhler2015} and
-addtionally forms the basis for GCN 15645 \citep{GCN15645}. This spectrum,
+additionally forms the basis for GCN 15645 \citep{GCN15645}. This spectrum,
 observed the following year (20.2 hours after the \swift~trigger), consists of
@@
 This spectrum has not been published previously. Starting 32.5 hours after the
-GRB trigger onboard \swift, this observation lasted $14\times600$ s for a total
+GRB trigger on-board \swift, this observation lasted $14\times600$ s for a total
 of 8.4 ks. Some loss occurred during the observations, which reduced integration
@@
 after the GRB trigger, captures a bright GRB afterglow. A broad absorption
-trough centered at $\sim$ 370 nm due to \lya~signals the redshift of the GRB,
+trough centred at $\sim$ 370 nm due to \lya~signals the redshift of the GRB,
 which is refined to $z = 2.057$ based on the additional detection of metal
@@
 minutes. A series of stare mode observations, increasing in exposure time, and
-ending with a nodding sequence, totaling $\sim$ 5000 s, are combined to form
+ending with a nodding sequence, totalling $\sim$ 5000 s, are combined to form
 this spectrum. A faint, almost featureless continuum is detected at low
@@
 hours after the trigger does not reveal anything conclusive. The host
-association is additionally ambiguous for this, likely extincted \citep[GCN
+association is additionally ambiguous for this, likely reddened \citep[GCN
 17767;][]{GCN17767}, GRB.
@@
 18187 \citep{GCN18187}, but the spectrum has not been published. Starting just
-12.4 minutes after the onboard trigger of \swift, a total of $4 \times 600$ s
+12.4 minutes after the on-board trigger of \swift, a total of $4 \times 600$ s
 spectroscopic integration was obtained. The observations began just before dawn
@@
 4 spectroscopic integrations obtained in a nodding sequence, each lasting for
-600s, securely allows us to measure a redshift for this GRB. Beginning at $\sim$
+600 s, securely allows us to measure a redshift for this GRB. Beginning at $\sim$
 470 nm, the \lya-forest is clearly detected at high signal-to-noise leading up
@@
 
-GRB~160625B is the first GRB for which a signicant linear polarization was
+GRB~160625B is the first GRB for which a significant linear polarization was
 measured \citep{Troja2017}. Because this burst was detected by \textit{Fermi}, it
@@
 are published in \citet{Heintz2017b}. Observations started 22.37 hours after the
-BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the
+BAT trigger and lasted for 2.4 ks. The afterglow continuum is detected across the
 entire spectral coverage of X-shooter and absorption lines from \mgi, \mgii,
@@
 These data have not been published elsewhere. Observations for GRB~161007A
-started 323 hr after the burst trigger and were centered on the potential host
+started 323 hr after the burst trigger and were centred on the potential host
 galaxy (see GCN 20014 \citep{GCN20014} and GCN 20020 \citep{GCN20020}). $4
@@
 additionally seen, clearly associating the $z = 2.710$ system with the GRB. From
-the \lya-line we infer \nh$=20.96 \pm 0.05$. This spectrum will be further analyzed in de Ugarte Postigo et al. (in prep).
+the \lya-line we infer \nh$=20.96 \pm 0.05$. This spectrum will be further analysed in de Ugarte Postigo et al. (in prep).
 
@@
 The lowest wavelength continuum from 430 nm to 560 nm is absorbed by the
-\lya~forest, with the broad \lya~absorption trough centered at 565 nm. In the
+\lya~forest, with the broad \lya~absorption trough centred at 565 nm. In the
 continuum, we additionally identify metal absorption features from \SIii, \civ,

### 2019-02-07 document/XSGRB_sample_rev1.tex
Accepted version

@@
 	afterglows that have been promptly observed with the X-shooter spectrograph
-	until \termdate. In total, we obtained spectroscopic observations of 103 individual GRBs
+	until \termdate. In total, we have obtained spectroscopic observations of 103 individual GRBs
 	observed within 48 hours of the GRB trigger. Redshifts have been measured for 97
@@
 	is representative of the full \textit{Swift} sample. We constrain the fraction
-	of dark bursts to be < 28 per cent and we confirm previous results that higher
+	of dark bursts to be < 28 per cent and confirm previous results that higher
 	optical darkness is correlated with increased X-ray absorption. For the 42
@@
 	unique resource to study the ISM across cosmic time, from the local progenitor
-	surroundings to the intervening universe.}
+	surroundings to the intervening Universe.}
 
@@
 out in a homogeneous way, the resulting quality of the spectra is not uniform.
-Individual bursts vary significantly in brightness, due to intrinsic
-differences in luminosity, dust extinction, time of observation, etc. The covered
-rest-frame wavelength range is also different for each case. Given that our
-observations are targets-of-opportunity, this leads to a varying S/N, and in
-particular the sensitivity limit to detecting absorption features varies greatly
-from burst to burst and as a function of wavelength.
+Individual bursts vary significantly in brightness, due to intrinsic differences
+in luminosity, dust extinction, time of observation, etc. The covered rest-frame
+wavelength range is also different for each case due to the different redshifts
+of each GRB. Given that our observations are targets-of-opportunity, this leads
+to a varying S/N, and in particular the sensitivity limit to detect absorption
+features varies greatly from burst to burst and as a function of wavelength.
 
@@
 GRB~090926A) are not a part of the sample we use to investigate the statistical
-properties of GRB afterglows, due to the different criteria for their selection. The
-first burst observed after science verification and the mounting of X-shooter on UT2,
-was GRB~091018, which is the first burst entering our statistically homogeneously selected
-sample. For all bursts that fulfill our sample selection criteria, described in
-Sect. \ref{samplecrit}, spectroscopic follow-up have been attempted with
-X-shooter. Various conditions can affect our ability to follow up a given burst,
-and a discussion of these conditions and their consequences for the sample is
-included in Sect. \ref{badbursts}.
+properties of GRB afterglows, due to the different criteria for their selection.
+The first burst observed after science verification and the mounting of
+X-shooter on UT2, was GRB~091018, which is the first burst entering our
+homogeneously selected statistical sample. For all bursts that fulfill our
+sample selection criteria, described in Sect. \ref{samplecrit}, spectroscopic
+follow-up have been attempted with X-shooter. Various conditions can affect our
+ability to follow up a given burst, and a discussion of these conditions and
+their consequences for the sample is included in Sect. \ref{badbursts}.
 
@@
 across a wide wavelength range, and allows a better comparison with
-high-resolution data. The scale of the correction additionally serves to
-solidify the accuracy of the wavelength solution.
+high-resolution data. The modest size of the correction additionally serves to
+confirm the accuracy of the wavelength solution.
 
@@
 
-To correct for the slit-loss, we develop an algorithm that can, based on the
-atmospheric conditions at time of observation and the wavelength dependence of
-seeing, estimate the losses at each wavelength bin. We start with the DIMM
+To correct for the slit-loss, we developed an algorithm that can, based on the
+atmospheric conditions at the time of observation and the wavelength dependence
+of seeing, estimate the losses at each wavelength bin. We start with the DIMM
 seeing \citep{Sarazin1990}, which measures the seeing FWHM at zenith. We correct
-the zenith seeing for the airmass of the observational sightline, which gives us
-the effective seeing, delivered outside the slit at 5000 \AA. We can then use
-that the seeing scales as $\sim \lambda^{-1/5}$ to obtain a measure of the
-seeing across the entire wavelength coverage of X-shooter. At each wavelength
-bin of the spectra, we simulate a synthetic 2D seeing PSF based on a normalised,
-2D Moffat profile which we integrate over the used slit width. This can be used to correct for the estimated slit-loss at each wavelength bin.
+the zenith seeing for the median airmass of the observational sightline, which
+gives us the effective seeing, delivered at 5000 \AA. We then use the fact that the
+seeing scales as $\sim \lambda^{-1/5}$ to obtain a measure of the seeing across
+the entire wavelength coverage of X-shooter. At each wavelength bin of the
+spectra, we simulate a synthetic 2D seeing PSF based on a normalised, 2D Moffat
+profile which we integrate over the used slit width. This can be used to correct
+for the estimated slit-loss at each wavelength bin.
 
@@
 the distribution.
-The scatter is smaller if we exclude measurements calibrated against the USNO catalog, which is known to have a poor accuracy, also because, given the small field-of-view, we often need to rely of faint reference stars.
+The scatter is smaller if we exclude measurements calibrated against the USNO catalog, which is known to have a poor accuracy, also because, given the small field-of-view, we often need to rely on faint reference stars.
 
@@
 The effect of telluric correction varies greatly depending on the wavelength
-region. In the red and near-infrared areas, it allows to identify features in a
-credible manner. which would be otherwise hidden among or blended with
-atmospheric features. The telluric correction allows us to reliably recover the
-continuum level even in regions of moderately low atmospheric transmission (see,
-e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}). Nonetheless, we recommend
-caution when fitting lines in regions significantly affected by telluric
-absorption.
+region. In the red and near-infrared ranges, telluric correction allows to
+identify features in a credible manner, which would be otherwise hidden among or
+blended with atmospheric features. The telluric correction also allows us to reliably
+recover the continuum level even in regions of moderately low atmospheric
+transmission (see, e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}).
+Nonetheless, we recommend caution when fitting lines in regions significantly
+affected by telluric absorption.
 
@@
 the presence of a bright afterglow), always in agreement with absorption.
-Besides, it is rare to detect emission from QSO intervening systems
-(see e.g. \citealt{2010MNRAS.408.2128F} for a discussion), which again indicates
-that GRB absorption systems are native to their hosts. An even more direct
-determination comes from the detection, in several cases, of fine-structure and
-metastable lines (for example from Fe and Ni), which are exclusively detected in
-GRB environments (and have been directly connected to the UV field generated by
-the GRB itself; e.g. \citealt{2006ApJ...648...95P, Vreeswijk2007, DElia2009}).
+Besides, it is rare to detect emission from QSO intervening systems (see e.g.
+\citealt{2010MNRAS.408.2128F} for a discussion), which again indicates that GRB
+absorption systems are native to their hosts. An even more direct determination
+comes from the detection, in several cases, of fine-structure and metastable
+lines (for example from Fe and Ni), which are usually detected in GRB
+environments (and have been directly connected to the UV field generated by the
+GRB itself; e.g. \citealt{2006ApJ...648...95P, Vreeswijk2007, DElia2009}).
 Finally, at $z \gtrsim 2$ (where Ly$\alpha$ enters the observed range) the onset
-of the Ly$\alpha$ forest also provides an \textit{upper} limit on the GRB
+of the Ly$\alpha$ forest also provides confirming evidence of the GRB
 redshift, which again is always consistent with the absorption line redshift.
@@
 	clearly visible in the hydrogen column densities, as previously also noted in
-	\citet{Fynbo2009}. To correct for geometric differences between GRB DLAs and QSO DLAs, we have multiplied the inferred GRB DLA $N_{\mathrm{HI}}$ by 2.} \label{fig:NH_dist}
+	\citet{Fynbo2009}. To correct for geometric differences between GRB DLAs and QSO DLAs, we have multiplied the inferred GRB DLA $N_{\mathrm{HI}}$ by a factor of two} \label{fig:NH_dist}
 \end{figure}
@@
 us too early. 
+%
+We thank Andrea Rossi (INAF/OAS) for the help with providing a photometric
+constraint on the redshift of GRB~161007A.
 %
@@
 %
+This publication makes use of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of Technology, funded by the National Aeronautics and Space Administration.
+%
+The Pan-STARRS1 Surveys (PS1) and the PS1 public science archive have been made possible through contributions by the Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, the Max-Planck Society and its participating institutes, the Max Planck Institute for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, the Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las Cumbres Observatory Global Telescope Network Incorporated, the National Central University of Taiwan, the Space Telescope Science Institute, the National Aeronautics and Space Administration under Grant No.
+NNX08AR22G issued through the Planetary Science Division of the NASA Science Mission Directorate, the National Science Foundation Grant No.
+AST-1238877, the University of Maryland, Eotvos Lorand University (ELTE), the Los Alamos National Laboratory, and the Gordon and Betty Moore Foundation.
+%
 
@@
 
+\newpage
 \bibliographystyle{aa}
@@
 		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
-		with dashed lines. The pressence of \lya, as noted in Tab. \ref{tab:HI}, is indicated with a black arrow.}
+		with dashed lines. The presence of \lya, as noted in Tab. \ref{tab:HI}, is indicated with a black arrow.}
 	\label{fig:HI1}
@@
 
-These data have not been published elsewhere. Observations for GRB~161007A
-started 323 hr after the burst trigger and were centred on the potential host
-galaxy (see GCN 20014 \citep{GCN20014} and GCN 20020 \citep{GCN20020}). $4
-\times 600$ s of observations were obtained revealing a faint continuum rising
-abruptly above the noise at $\sim 685$ nm and extending up to 2100 nm. A very
-low significance continuum is detected at shorter wavelengths, down to $\sim
-600$ nm. No significant emission features could be identified. Two possibilities
-can explain the observed continuum break, but neither are fully satisfactory.
-The first option is the Lyman alpha cutoff at $z \sim 4.6$; at this redshift, the
-usually detected strong nebular lines would be shifted out of the covered
-wavelength range. The host would be however exceptionally bright, with an
-absolute magnitude at $\approx 150$~nm{} $M = -23.8$~AB, that is about 2.8~mag
-brighter than $M^*$ at that redshift (e.g. \citealt{Bouwens2015}). The inferred
-X-ray hydrogen-equivalent column density would also be a record-high $4 \times
-10^{23}$~cm$^{-2}$ (see e.g. \citealt{Campana2012}). Alternatively, the
-continuum discontinuity could be the 400~nm{} break at $z = 0.71$; such
-feature is however prominent only in early-type galaxies, which are unlikely
-hosts of long-duration GRBs (but see \citealt{Rossi2014}). Given the lack of
-optical variability, it cannot be excluded that the object within the XRT error
-circle is a chance association, as the probability of a random alignment is
-small but non negligible (of the order of 1\%). In conclusion, we cannot provide
-a secure redshift for GRB 161007A.
+These data have not been published elsewhere. Observations for GRB 161007A
+started 323 hr after the burst trigger and targeted the potential host galaxy
+(see GCN 20014 \citep{GCN20014} and GCN 20020 \citep{GCN20020}). $4 \times 600$s
+of observations were obtained, revealing a faint continuum rising abruptly above
+the noise at $\sim 685$~nm and extending up to 2100 nm. A very low significance
+continuum is detected at shorter wavelengths, down to $\sim  600 $ nm. No
+significant emission features could be identified. Two possibilities can explain
+the observed continuum break, but neither is fully satisfactory. The first
+option is the Ly$\alpha$ cutoff at $z \sim 4.6$; at this redshift, the usually detected
+strong nebular lines would be shifted out of the covered wavelength range. The
+host would be however exceptionally bright, with an absolute magnitude at
+$\approx 150$~nm{} $M = -23.8$~AB, that is about 2.8 mag brighter than $M^*$ at
+that redshift (e.g. \citealt{Bouwens2015}). In order to test this option, we
+secured $B$-band imaging (using the ESO/VLT UT1 equipped with the FORS2 camera)
+of the host, yielding a faint but significant detection ($B = 25.48 \pm 0.13$
+Vega). As the $B$ band would lie entirely bluewards of the Lyman limit at $z =
+4.6$, where negligible flux is expected to be transmitted by the IGM, our
+detection rules out this solution. A second solution is that the continuum
+discontinuity is the 400 nm break at $z = 0.71$, such feature is however more
+prominent in evolved galaxies, which are unlikely hosts of long-duration GRBs
+(but see \citealt{Rossi2014}). A fit to the sparse available photometry,
+including our FORS2 datum, the GTC $i$-band value \citep{GCN20014}, Pan-STARRS
+measurements \citep{Chambers2016}, and WISE \citep{Wright2010} W1 detection
+($\mbox{W1}=19.72\pm0.15$), is unable to tightly constrain the redshift ($0.47 <
+z < 1.34$). Fixing $z = 0.7$ yields however a significant star formation rate
+(25--200 $M_\odot$~yr$^{-1}$), which is in contrast with the lack of detected
+emission lines in the spectrum (the expected H$\alpha$ flux would be
+$10^{-15}$--$10^{-16}$ erg cm$^{-2}$ s$^{-1}$, easily detectable). As such we
+cannot identify a self-consistent solution. Given the lack of optical
+variability of the candidate counterpart, it cannot be excluded that the object
+within the XRT error circle is a chance association, as the probability of a
+random alignment is small but non negligible (of the order of 1\%). In
+conclusion, we cannot provide a secure redshift for GRB 161007A.
 

### 2019-02-07 document/final/XSGRB_sample_final.tex
Accepted version

--- 2019-02-07 document/XSGRB_sample_rev1.tex
@@
 		observations collected at the European Southern Observatory, Paranal, Chile,
-		Program IDs: 084.A-0260, 084.D-0265, 085.A-0009, 086.A-0073, 087.A-0055,
+		Programme IDs: 084.A-0260, 084.D-0265, 085.A-0009, 086.A-0073, 087.A-0055,
 		088.A-0051, 089.A-0067, 090.A-0088, 091.A-0877, 0091.C-0934, 092.A-0124,
@@
 	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
-	of observational selection criteria that minimize biases with regards to
+	of observational selection criteria that minimise biases with regards to
 	intrinsic properties of the GRBs, the follow-up effort has been focused on
@@
 	properties of the bursts in the sample and find that the sample presented here
-	is representative of the full \textit{Swift} sample. We constrain the fraction
+	is representative of the full \textit{Swift} sample. We have constrained the fraction
 	of dark bursts to be < 28 per cent and confirm previous results that higher
 	optical darkness is correlated with increased X-ray absorption. For the 42
-	bursts for which it is possible, we provide a measurement of the neutral
+	bursts for which it is possible, we have provided a measurement of the neutral
 	hydrogen column density, increasing the total number of published H{\sc I}
@@
 astrophysics. Examples range from small-scale phenomena relating to magnetars,
-properties of highly relativistic jets, hyper/supernova explosions, the
+properties of highly relativistic jets, hypernova and supernova explosions, the
 interstellar medium, dust extinction curves, starburst galaxies, chemical and
-molecular abundances, escape of ionizing radiation, the ionization state of the
+molecular abundances, escape of ionising radiation, the ionization state of the
 intergalactic medium, intervening absorption systems to standard candles in
-cosmology \citep[e.g.,][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
+cosmology \citep[e.g.][]{Wijers1998, Savaglio2006, Ghirlanda2007, Molinari2007,
 	Amati2008, Vergani2009, Prochaska2009, HjorthBloom2012, Rowlinson2017,
@@
 possible to harvest much of the great potential in using GRBs as probes of the
-intergalactic medium, which was already hinted at by results from earlier
-missions \citep[e.g.,][]{Paradijs2000, Ricker2004}. With the three on-board
+intergalactic medium, which has already been hinted at by results from earlier
+missions \citep[e.g.][]{Paradijs2000, Ricker2004}. With the three on-board
 instruments, the Burst Alert Telescope \citep[BAT;][]{Barthelmy2005}, the X-Ray
@@
 spectroscopy was secured using low-resolution spectrographs \citep[typically
-$R=\lambda/\Delta\lambda$$<$1000, e.g.,][]{Fynbo2009}. Spectroscopy is powerful
+$R=\lambda/\Delta\lambda$$<$1000, e.g.][]{Fynbo2009}. Spectroscopy is powerful
 as it allows us to secure information even for very faint targets
@@
 secured, and for these events much more information about conditions inside the
-host galaxies were extracted \citep[e.g.,][]{Fiore2005, Thone2007,
+host galaxies were extracted \citep[e.g.][]{Fiore2005, Thone2007,
 	Prochaska2007, Vreeswijk2007, DElia2009, Castro-Tirado2010}.
@@
 -- 2017 to use the X-shooter spectrograph to secure spectroscopic observations
-of afterglows and host galaxies of GRBs detected by {\it Swift}. We here make
+of afterglows and host galaxies of GRBs detected by {\it Swift}. Here we make
 all the data resulting from the survey publicly available in reduced form (see Sect
@@
 
-The paper is organized in the following way: In Sect.~\ref{sample} we describe
+The paper is organised in the following way: In Sect.~\ref{sample} we describe
 the sample including the sample selection criteria and the observational
@@
 in the data reduction process and auxiliary material. In Sect.~\ref{results} we
-describe the results of the survey, i.e. the efficiency of the follow-up effort
+describe the results of the survey, the efficiency of the follow-up effort,
 and the characteristics of the observed bursts. We also assess the
-completeness of the realized sample. Finally, we offer our conclusions
+completeness of the realised sample. Finally, we offer our conclusions
 in Sect.~\ref{conclusions}. We use the $\Lambda$CDM cosmology parameters
@@
 of the telescope or unconstrained burst localizations as reported by the alerting
-facility. To maximize the return of the follow-up campaign, we have chosen a
+facility. To maximise the return of the follow-up campaign, we have chosen a
 few selection criteria that facilitate an unbiased selection of bursts, while
@@
 brightness distribution. When investigating a specific distribution function, a
-high degree of \textit{completeness} is desired \citep[e.g.,][]{Perley2016b}.
-
-In defining the selection criteria, we simultaneously aim to minimize any biases
-against intrinsic astrophysical conditions while at the same time maximizing the
+high degree of completeness is desired \citep[e.g.][]{Perley2016b}.
+
+In defining the selection criteria, we simultaneously aim to minimise any biases
+against intrinsic astrophysical conditions while at the same time maximising the
 likelihood of successful observations, hence allowing us to obtain a higher
@@
 GRBs in a fair way. The selection criteria used here are based on previous,
-similar studies \citep{Jakobsson2006b, Fynbo2009, Hjorth2012}. We characterize
+similar studies \citep{Jakobsson2006b, Fynbo2009, Hjorth2012}. We characterise
 the sample completeness in Sect. \ref{completeness}.
 
-The selection criteria that define a GRB as part of our initial \textit{statistical}
+The selection criteria that define a GRB as part of our initial statistical
 	sample are:
@@
 
-	\item XRT started observing within 10 minutes after the GRB; an XRT position
+	\item XRT started observing within ten minutes after the GRB; an XRT position
 	must be distributed within 12 hr.
@@
 	\item The target must be visible from Cerro Paranal for at least 60 minutes, 30
-degrees above the horizon, with the Sun below $-12$ degrees\footnote{Note that
+degrees above the horizon, with the Sun below $-12$ degrees\footnote{We note that
 	in the P84 proposal the criteria have been stated differently, the visibility
@@
 \citep{Barthelmy2000} within 12 hours and to account for observing constraints
-on \textit{Swift} that XRT began observations within 10 minutes. The additional
+on \textit{Swift} that XRT began observations within ten minutes. The additional
 timing requirement of the XRT follow-up means that all bursts in our sample have
@@
 To ensure a minimum of observability, we require that the GRB is visible from
-the telescope site at Cerro Paranal, Chile, for a least 1 hour after the trigger
+the telescope site at Cerro Paranal, Chile, for a least one hour after the trigger
 with the sun below $-12$ degrees. This secures time for the spectroscopic
@@
 	old \citet{Schlegel1998} values for consistency with the first semesters of our
-	program.}. Choosing low-extinction sightlines also reduces the problem of field
+	programme.}. Choosing low-extinction sightlines also reduces the problem of field
 crowding and the contamination from Galactic outbursts posing as GRB impostors.
@@
 Bursts that fulfill these selection criteria are what we define as our initial
-\textit{statistical} sample. To reach a higher degree of sample completeness, we
+statistical sample. To reach a higher degree of sample completeness, we
 impose additional cuts that increase our ability to infer population properties.
@@
 Two people are permanently on alert to manually supervise each event, for
-example to recognize events of special interest beyond their inclusion in the
+example to recognise events of special interest beyond their inclusion in the
 sample. Members of our collaboration have access to numerous facilities spread
@@
 	\item The GRB triggered \textit{Swift}.
-	\item The X-ray position must be available less than 1 hr after the GRB.
+	\item The X-ray position must be available less than one hour after the GRB.
 	%    \item X-shooter is the currently operating instrument at the telescope.
@@
 pivotal example is the study of temporal variability of GRB afterglow absorption
-systems due to effect of the GRB itself on the surrounding medium \citep[e.g.,
+systems due to effect of the GRB itself on the surrounding medium \citep[e.g.
 see][]{Dessauges-Zavadsky2006, Vreeswijk2007, DElia2009, Vreeswijk2013}.
@@
 	\ref{results}. The background shows the dust maps presented in
-	\citet{Schlegel1998}. Please note that we removed the background where the
+	\citet{Schlegel1998}. We note that we removed the background where the
 	sample criterion is violated ($A_V > 0.5$ mag) and replaced it with a white
@@
 098.A-0136 (PI: Greiner), and 095.B-0811 (PI: Levan). The total collection of
-spectra represents \textit{all} GRB afterglows that have been followed up by
+spectra represents all GRB afterglows that have been followed up by
 X-shooter up to \termdate, which marks the end of the XS-GRB legacy follow-up
-program.
+programme.
 
@@
 slightly different observing strategy was employed. Starting as rapidly as
-possible, a simple stare mode sequence was started, with 5 spectroscopic
+possible, a simple stare mode sequence was started, with FIVE spectroscopic
 integrations with increasing exposure times.
@@
 Due to a mechanical failure, the atmospheric dispersion corrector (ADC) was
-disabled from 1st of August 2012 until the end of this program. Only GRB~100728B
+disabled from 1st of August 2012 until the end of this programme. Only GRB~100728B
 was affected by the failing ADC prior to disablement, resulting in a
@@
 across the spatial direction of the slit as a function of wavelength. This
-effect has been modeled in the extraction procedure, as described in Sect.
+effect has been modelled in the extraction procedure, as described in Sect.
 \ref{extract}.
@@
 while propagating the pixel uncertainties derived by the pipeline. This
-rectification  process correlates neighboring pixels and in order to minimize
+rectification  process correlates neighboring pixels and in order to minimise
 the degree of correlation, we need to choose a physical sampling that matches
@@
 wavelength part of neither of the arms have a sampling lower than the Nyquist
-sampling rate of 2 pixels per resolution FWHM.
+sampling rate of two pixels per resolution FWHM.
 
@@
 spectrum is then convolved with an increasing seeing PSF and the width that
-minimizes $\chi^2$ with the updated observed sky is chosen to be the effective
+minimises $\chi^2$ with the updated observed sky is chosen to be the effective
 sky-PSF. Using the synthetic sky with the matched resolution, a final wavelength
@@
 
-The resulting offsets are generally small, on the order of $\sim 5$~km~s$^{-1}$,
+The resulting offsets are generally small, of the order of $\sim 5$~km~s$^{-1}$,
 and thus do not influence significantly the redshift measurements, whose
@@
 Target-of-Opportunity mode. In most cases, there is little possibility to tweak
-slit widths to the seeing at the time of observations (i.e. to optimize spectral
+slit widths to the seeing at the time of observations (for instance to optimise spectral
 resolution and signal-to-noise), and almost all our data are therefore taken
@@
 \citep{Horne1986}, which weights the extraction aperture by the spectral profile,
-or alternatively sum all pixels within 1 FWHM of the modeled profile. Where
+or alternatively sum all pixels within 1 FWHM of the modelled profile. Where
 possible, we have used the optimal extraction. In cases where the trace is very
@@
 	\centerline{\includegraphics[width=\columnwidth]{figures/flux_accuracy}}
-\caption{The left panel shows the acquisition camera magnitudes as a function
+\caption{Left panel shows the acquisition camera magnitudes as a function
 	of synthetic photometry derived from the spectra. The blue, dashed line
@@
 	the labels. The $R$- and $I$- band magnitudes have been calibrated against the
-	USNO-B1 catalog, which has a poor photometric accuracy; this contributes
+	USNO-B1 catalogue, which has a poor photometric accuracy; this contributes
 	significantly to the observed scatter, and for this reason we indicate these
@@
 to secure photometry for the vast majority of events. As photometric flux
-standards are rarely acquired, we resort to sky surveys such as the USNO catalog
+standards are rarely acquired, we resort to sky surveys such as the USNO catalogue
 (for the $R$ and $I$ bands; \citealt{Monet2003}), SDSS, or Pan-STARRS ($griz$;
@@
 with significant scatter. The right panel in Fig.~\ref{fig:flux_acq} shows a
-histogram of the differences, which can be summarized with $\Delta \mathrm{mag}
+histogram of the differences, which can be summarised with $\Delta \mathrm{mag}
 = 0.07 ^{+0.51} _{-0.45} $, where the errors are the 16th and 84th percentiles of
 the distribution.
-The scatter is smaller if we exclude measurements calibrated against the USNO catalog, which is known to have a poor accuracy, also because, given the small field-of-view, we often need to rely on faint reference stars.
+The scatter is smaller if we exclude measurements calibrated against the USNO catalogue, which is known to have a poor accuracy, also because, given the small field-of-view, we often need to rely on faint reference stars.
 
@@
 received spectrum. These telluric features can be corrected for in a multitude
-of ways. We employ a prioritized list of methods here, depending on the
+of ways. We employ a prioritised list of methods here, depending on the
 availability of the chosen methods. Since the observations are often taken at
-odd times under varying conditions, this prioritized list ensures that we are
+odd times under varying conditions, this prioritised list ensures that we are
 always doing the best possible correction.
@@
 recover the continuum level even in regions of moderately low atmospheric
-transmission (see, e.g., the A- and B-bands in Fig.~\ref{fig:spectrum}).
+transmission (see, for example the A- and B-bands in Fig.~\ref{fig:spectrum}).
 Nonetheless, we recommend caution when fitting lines in regions significantly
@@
 	\centerline{\includegraphics[width=0.85\linewidth]{figures/GRB121024A.pdf}}
-	\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
+	\caption{Telluric corrected, normalised spectrum of GRB~121024A at $z = 2.300$
 		that illustrates the typical data quality. The continuum estimate is shown in
@@
 		Additionally, three intervening systems are seen in this sightline. This
-		spectrum has been analyzed in detail in \citet{Friis2015}.}
+		spectrum has been analysed in detail in \citet{Friis2015}.}
 	\label{fig:spectrum}
@@
 differed from the points by less than $\sim$ 5 per cent. This filtering is used
-because the intrinsic afterglow continuum of GRBs can be modeled by power laws
+because the intrinsic afterglow continuum of GRBs can be modelled by power laws
 \citep{Piran2005}, and removing points that differ significantly from a smooth
@@
 	material}. This release includes both prompt afterglow observations as well as
-late time observations of the associated hosts, and represents \textit{all}
+late time observations of the associated hosts, and represents all
 afterglow spectra of GRBs carried out by the X-shooter spectrograph since the
 commissioning of the instrument, \startdate, and until the end of the last
-period of the program 098.A-0055, \termdate~and thus constitutes eight years of GRB
+period of the programme 098.A-0055, \termdate~and thus constitutes eight years of GRB
 afterglow observations with X-shooter. An overview of all the spectra and their
@@
 obtained sample is representative for the full \textit{Swift} sample. An
-important note is that here we provide a release for \textit{all} GRBs after
+important note is that here we provide a release for all GRBs after
 \startdate, that have been observed with X-shooter, while only a subset of these
-constitutes our \textit{statistical} sample. The statistical sample is based on
+constitutes our statistical sample. The statistical sample is based on
 the selection criteria described in Sect. \ref{samplecrit}. Some bursts not
 fulfilling the sample criteria have been followed up due to interesting
-characteristics, e.g., curious properties of their light curves, their
+characteristics, due to curious properties of their light curves, their
 brightness, etc. These bursts are not discussed as part of the investigation of
@@
 spectacular burst outside the statistical sample is the bright \textit{INTEGRAL}
-burst GRB~161023A (de Ugarte Postigo et al., in prep), that contains at least 15
+burst GRB~161023A \citep{2018A&A...620A.119D}, that contains at least 15
 intervening absorption systems (See \ref{161023}).
@@
 criteria specified in Sect. \ref{samplecrit}, since the commissioning of
-VLT/X-shooter. This sample constitutes the \textit{statistical sample} from
+VLT/X-shooter. This sample constitutes the statistical sample from
 which we will derive statistical properties of the GRB population. The redshift
@@
 or afterglow properties. The reasons include unavailability of the telescope due
-to technical maintenance (e.g., mirror re-coating), a visiting observer
+to technical maintenance (because of mirror re-coating), a visiting observer
 rejecting the ToO trigger, or bad weather. Because this cut is unrelated to the
@@
 $F_{\gamma}$ is the observed BAT fluence in the 15-150 keV band and $d_L$ is the
-luminosity distance to the burst at the given redshift. Note that this measure
+luminosity distance to the burst at the given redshift. We note that this measure
 of luminosity does not include any $k$-correction. As an indication of the
@@
 Because the redshift completeness of our statistical sample is 88 per cent,
-making an inference of the \textit{true} redshift distribution of GRBs based on
+making an inference of the true redshift distribution of GRBs based on
 this sample is impossible. %The biases introduced in the measured redshift
@@
 	are marked by blue stars whereas black dots show all GRBs observed with
-	X-shooter. All \swift~GRBs with measured redshifts are shown in gray. For
+	X-shooter. All \swift~GRBs with measured redshifts are shown in grey. For
 	comparison, we overplot with red stars the GRB \citep{Tanvir2009b,
@@
 \citep{Groot1998, Djorgovski2001, Fynbo2001}. The degree of optical extinction
-relative to the X-ray brightness has been parametrized in terms of their optical
+relative to the X-ray brightness has been parametrised in terms of their optical
 darkness, using the measurement of, or limit on, the optical to X-ray spectral
@@
 
-For all bursts with follow-up within 100 hours we calculate the
-"darkness"-parameter, $\beta_{OX}$ \citep{Jakobsson2004}. This requires the
+For all bursts with follow-up within 100 hours we calculated the `darkness'-parameter, $\beta_{OX}$ \citep{Jakobsson2004}. This requires the
 simultaneous measurement of the X-ray flux density and the optical flux density
@@
 	presented in \citet{Fynbo2009}, we exclude all bursts for which we do not have
-	a redshift. Darker colors represent a higher density of points. In blue are the
+	a redshift. Darker colours represent a higher density of points. In blue are the
 	points for the bursts presented here along with the marginal histograms. Limits
@@
 values for $N_{\mathrm{H, X}}$ in the comparison for which the GRB has a
-measured redshift, this measure is likely biased toward optically brighter
+measured redshift, this measure is likely biased towards optically brighter
 bursts \citep[e.g.][]{Watson2012}.
@@
 \citep{Hogg1999, Bloom2002, Fruchter2006, Lyman2017}. Because a significant fraction
-of the hydrogen along the line of sight in these regions has not been ionized,
+of the hydrogen along the line of sight in these regions has not been ionised,
 the optical depth at the wavelength of \lya~is very high, saturating not only
@@
 that GRBs are systematically located behind the highest \nh, previously noted
-\citep[e.g.,][]{Prochaska2007, Fynbo2009}, is very clear in this figure. The
+\citep[e.g.][]{Prochaska2007, Fynbo2009}, is very clear in this figure. The
 reason for this is that quasar sample sight-lines through galaxies that are
@@
 \swift. This work was initiated by the consortium that built X-shooter and
-included this project as a key part of the GTO program, but over the years the
+included this project as a key part of the GTO programme, but over the years the
 project continued in open time.
@@
 
-Our sample serves the purpose to further characterize the environments of GRBs
+Our sample serves the purpose to further characterise the environments of GRBs
 that was also much advanced by the previous surveys based primarily on
@@
 towards about 10$^5$ quasars, a handful systems with \nh~$ > 22$ have been
-identified \citep[e.g., five in][]{Noterdaeme2012b}. Long GRB afterglow spectra,
-by contrast, reveal such systems routinely \citep[][and this
+identified \citep[for instance five in][]{Noterdaeme2012b}. Long GRB afterglow
+spectra, by contrast, reveal such systems routinely \citep[][and this
 work]{Jakobsson2006b, Fynbo2009, Cucchiara2015}. With afterglow spectroscopy
 (throughout the electromagnetic spectrum from X-rays to the sub-mm) we are able
-to characterize the properties of star-forming galaxies over cosmic history in
+to characterise the properties of star-forming galaxies over cosmic history in
 terms of redshifts, metallicities, molecular contents, ISM temperatures, UV-flux
@@
 published or submitted for publication focusing on many of these specific issues
-of our sample such as extinction curves (\citealt{Japelj2015}, Zafar et al.
-submitted, see also \citealt{Fynbo2014, Heintz2017a}), emission lines from the
-underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
-\mgii~ absorbers \citep{Christensen2017}, \citet{Arabsalmani2018} on the
-metallicity-scaling relations, and escape of ionizing radiation
-\citep{Tanvir2017} . A number of additional companion papers are also planned,
-investigating the detailed properties of the sample presented here, including
-equivalent width distributions (de Ugarte Postigo et al., in preparation),
-metallicities and kinematics (Th{\"o}ne et al., in preparation), high ionization
-lines (Heintz et al., in submitted), molecular lines \citep{Bolmer2018},
-fine-structure lines (Vreeswijk et al., in preparation), and composite GRB
-afterglow spectrum (Selsing et al., in preparation).
+of our sample such as extinction curves (\citealt{Japelj2015, 2018MNRAS.480..108Z,
+	2018MNRAS.479.1542Z}, see also \citealt{Fynbo2014,
+	Heintz2017a}), emission lines from the underlying host galaxies
+\citep{Kruhler2015}, the frequency of intervening \mgii~ absorbers
+\citep{Christensen2017}, \citet{Arabsalmani2018} on the metallicity-scaling
+relations, and escape of ionising radiation \citep{Tanvir2017}. A number of
+additional companion papers are also planned, investigating the detailed
+properties of the sample presented here, including equivalent width
+distributions (de Ugarte Postigo et al., in preparation), metallicities and
+kinematics (Th{\"o}ne et al., in preparation), high ionization lines \citep{2018MNRAS.479.3456H}, molecular lines \citep{Bolmer2018}, fine-structure lines
+(Vreeswijk et al., in preparation), and composite GRB afterglow spectrum
+(Selsing et al., in preparation).
 
@@
 sightlines as we have shown and consistent with earlier findings from samples
-based on low-resolution spectroscopy \citep[e.g.,][]{Fynbo2009} and from studies
+based on low-resolution spectroscopy \citep[e.g.][]{Fynbo2009} and from studies
 of complete samples of GRB host galaxies \citep{Hjorth2012, Covino2013,
@@
 \section{The complex error function and the spectral PSF} \label{voigt}
-When modeling the spectral PSF, we need to evaluate the Voigt-profile.
-The Voigt profile, which is the convolution of the Gaussian and Lorentzian profiles, can, centered at zero, be written as \citep{pagnini2010}
+When modelling the spectral PSF, we need to evaluate the Voigt-profile.
+The Voigt profile, which is the convolution of the Gaussian and Lorentzian profiles, can, centred at zero, be written as \citep{pagnini2010}
 \begin{equation} 
@@
 
-The data presented here have been examined by \citet{Japelj2015} and Zafar et al.
-(submitted), who both find a significant amount of extinction, $A_V\approx 1$
-mag, also supported by previous measurements \citep{Morgan2014a}. Three epochs
-of observations have been obtained, the first two immediately after the burst,
-and the last one long after the afterglow had faded. Starting 1.4 hours after
-the \textit{Swift} trigger, the first epoch contains bright afterglow continuum.
-Rich in absorption features belonging to a multitude of ions at $z = 1.728$ is
-estimated for the host with intervening systems at $z = 1.476$, $z = 1.214$, $z
-= 0.662$ and $z = 0.632$. The second epoch, obtained 4.5 hours after the burst
-contains the fading afterglow. A third epoch is obtained $>1$ year after the GRB
-in which emission lines from \hb~and \ha~are found at the redshift of the host,
-confirming the association of the absorption line system and the host. We also
-detect C\,\textsc{i} in absorption which indicates the presence of cold gas.
+The data presented here have been examined by \citet{Japelj2015} and
+\citet{2018MNRAS.479.1542Z}, who both find a significant amount of extinction,
+$A_V\approx 1$ mag, also supported by previous measurements \citep{Morgan2014a}.
+Three epochs of observations have been obtained, the first two immediately after
+the burst, and the last one long after the afterglow had faded. Starting 1.4
+hours after the \textit{Swift} trigger, the first epoch contains bright
+afterglow continuum. Rich in absorption features belonging to a multitude of
+ions at $z = 1.728$ is estimated for the host with intervening systems at $z =
+1.476$, $z = 1.214$, $z = 0.662$ and $z = 0.632$. The second epoch, obtained 4.5
+hours after the burst contains the fading afterglow. A third epoch is obtained
+$>1$ year after the GRB in which emission lines from \hb~and \ha~are found at
+the redshift of the host, confirming the association of the absorption line
+system and the host. We also detect C\,\textsc{i} in absorption which indicates
+the presence of cold gas.
 
@@
 belonging to \hb, \oiii~and \ha~are detected at $z = 0.350$. The line belonging
-to \ha~shows some extended emission toward the afterglow continuum. The angular
+to \ha~shows some extended emission towards the afterglow continuum. The angular
 distance between the two sources correspond to a projected distance in the host
@@
 The data presented here also formed the basis of GCN 13477 \citep{GCN13477} and
-are discussed in Klose et al. (submitted), but are not published anywhere.
+are discussed in \citet{2018arXiv180802710K}, but are not published anywhere.
 Observations of this burst started 7.8 hours after the GRB trigger, lasting 4.8
@@
 The gross content of these spectra have previously been reported in GCN 20104
-\citep{GCN20104}. This extremely bright \textit{INTEGRAL} burst is not a part of
-the statistical sample. Starting 3 hours after the trigger, the spectra are
-based on $2 \times 600$ s. The GRB afterglow continuum is detected with the
-highest signal-to-noise of all the spectra presented here. A myriad of
-absorption features are visible against the continuum, many of which would not be
-visible against lower signal-to-noise spectra. We identify at least 15
-intervening absorbers between $z = 1.243$ and $z = 2.710$. In the host system we
-clearly see absorption from \lyb, \lya, \sii, \SIii, \oi, \SIiv, \civ, \alii,
-\aliii, \feii, \mgii, and \mgi. Several fine-structure transitions are
-additionally seen, clearly associating the $z = 2.710$ system with the GRB. From
-the \lya-line we infer \nh$=20.96 \pm 0.05$. This spectrum will be further analysed in de Ugarte Postigo et al. (in prep).
+\citep{GCN20104} and is published in \citet{2018A&A...620A.119D}. This extremely
+bright \textit{INTEGRAL} burst is not a part of the statistical sample. Starting
+3 hours after the trigger, the spectra are based on $2 \times 600$ s. The GRB
+afterglow continuum is detected with the highest signal-to-noise of all the
+spectra presented here. A myriad of absorption features are visible against the
+continuum, many of which would not be visible against lower signal-to-noise
+spectra. We identify at least 15 intervening absorbers between $z = 1.243$ and
+$z = 2.710$. In the host system we clearly see absorption from \lyb, \lya, \sii,
+\SIii, \oi, \SIiv, \civ, \alii, \aliii, \feii, \mgii, and \mgi. Several
+fine-structure transitions are additionally seen, clearly associating the $z =
+2.710$ system with the GRB. From the \lya-line we infer \nh$=20.96 \pm 0.05$.
 

### 2019-02-07 document/final/authorlist.tex
Accepted version

--- 2018-10-16 document/authorlist.tex
@@
 \author{
-J.~Selsing \inst{1}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}\addtocounter{footnote}{2}
+J.~Selsing \inst{1, 2}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}\addtocounter{footnote}{2}
 \and D.~Malesani \inst{1, 2, 3}\fnmsep\thanks{On-call observer}
-\and P.~Goldoni \inst{4}\fnmsep$^\dagger$
+\and P.~Goldoni \inst{4, 7}\fnmsep$^\dagger$
 \and J.~P.~U. Fynbo \inst{1, 2}\fnmsep$^\dagger$
@@
 \institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
-\and The Cosmic Dawn Center, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+\and The Cosmic Dawn Center (DAWN), Niels Bohr Institute, University
+of Copenhagen, Juliane Maries Vej 30, DK-2100 Copenhagen \O;
+DTU-Space, Technical University of Denmark, Elektrovej 327,
+DK-2800 Kgs.\ Lyngby
 \and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, 2800 Lyngby, Denmark

### 2019-02-07 document/final/tables/redshift_comparison.tex
Accepted version

--- 2018-02-05 document/tables/redshift_comparison.tex
@@
 		{\smallskip}
-		BAT6 & 58 & 95 \% &  1.90 &  $1.71_{-1.04}^{+1.31}$ \\
+		BAT6 & 58 & 97 \% &  1.90 &  $1.71_{-1.04}^{+1.31}$ \\
 		{\smallskip}
-		TOUGH & 69 &  77 \% & 2.20 & $2.11_{-1.36}^{+1.42}$ \\
+		TOUGH & 69 &  87 \% & 2.20 & $2.11_{-1.36}^{+1.42}$ \\
 		{\smallskip}
@@
 	samples. The SHOALS redshift characteristics are taken from \citet{Perley2016a},
-	the BAT6 redshifts are from \citet{Salvaterra2012}, TOUGH is from
-	\cite{Hjorth2012}, and Fynbo09 are from \citet{Fynbo2009}. The errors shown on
+	the BAT6 redshifts are from \citet{Salvaterra2012} with the update from
+	\citet{Pescalli2016}, TOUGH is from \cite{Hjorth2012} with the update from
+	\citet{Schulze2015}, and Fynbo09 are from \citet{Fynbo2009}. The errors shown on
 	the median redshift contain 68 per cent of the probability mass.

### 2019-02-07 py/database/timediff.py

@@ -16,7 +16,7 @@ def timediff(t1, t2):
 
 if __name__ == '__main__':
 
-    t_trigger = "2018:04:04T00:45:35"
-    t_obs = "2018:04:04T01:01:33"
+    t_trigger = "2019:01:14T20:57:03"
+    t_obs = "2019:01;18T00:59:27.051"
 
     print(timediff(t_trigger, t_obs))
